### 2013-04-26 tex/thesis.tex
[tex] expanded outline : new sections

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\title{Typed Functional Genetic Programming}
\author{Tomáš Křen}
\date{April 2013}

\begin{document}
	\maketitle

	\section{Introduction}

	\section{Definitions}
		\subsection{Lambda term}
		\subsection{Type}
		\subsection{Context}
		\subsection{Inference rules}
		\subsection{Term generating grammars}
		\subsection{Inhabitation tree}
			\subsubsection{Inhabitation Machine}
		\subsection{Roadmap}
		\subsection{Conversion to SKI combinators}

	\section{Top level view}

	\section{Term generating}
	\section{Crossover}
	\section{Mutation}

	\section{Problems}
	In this section we will present usage of the system on solving specific problems.

		\subsection{Even Parity Problem}
		\subsection{Big Context}
		\subsection{Fly}
		\subsection{Simple Symbolic Regression}
		\subsection{Artificial Ant}
		\subsection{Boolean Alternate}

	\section{Conclusion}

	\section{Sandbox, YAY!}

	Blah!

	% Example formula

  	\begin{align}
    	E &= mc^2                              \\
    	m &= \frac{m_0}{\sqrt{1-\frac{v^2}{c^2}}}
  	\end{align}

\end{document}

### 2013-05-04 tex/thesis.tex
[tex] definitions

@@
-\documentclass[12pt]{article}
+\documentclass[12pt]{report}
+
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage[utf8]{inputenc}
+
 \title{Typed Functional Genetic Programming}
@@
 
-	\section{Introduction}
+\chapter{Introduction}
 
-	\section{Definitions}
-		\subsection{Lambda term}
-		\subsection{Type}
-		\subsection{Context}
-		\subsection{Inference rules}
-		\subsection{Term generating grammars}
-		\subsection{Inhabitation tree}
-			\subsubsection{Inhabitation Machine}
-		\subsection{Roadmap}
-		\subsection{Conversion to SKI combinators}
+\chapter{Definitions}
 
+	Let us first say some basic definitions.
+
+	\section{Lambda term}
+
+		Let $V$ be set of {\it variable names}.  \\* 
+		Let $C$ be set of {\it constant names}.	 \\*		
+		Then $\Lambda$ is set of {\it $\lambda$-terms} inductively defined as follows:
+
+		\begin{align*}
+			 x   \in V                 &\Rightarrow     x   \in \Lambda \\
+			 c   \in C                 &\Rightarrow     c   \in \Lambda \\
+			 M,N \in \Lambda           &\Rightarrow ( M N ) \in \Lambda \\
+			 x   \in V , M \in \Lambda &\Rightarrow ( \lambda x . M ) \in \Lambda 
+		\end{align*}
+
+	\section{Type}
+
+		Let $A$ be set of {\it atomic types}. \\*
+		Then $\mathbb{T}$ is set of {\it types} inductively defined as follows:
+
+		\begin{align*}
+		\alpha      \in A          &\Rightarrow     \alpha   \in \mathbb{T} \\
+		\sigma,\tau \in \mathbb{T} &\Rightarrow ( \sigma \rightarrow  \tau ) \in \mathbb{T} 
+		\end{align*}
+
+	\section{Statement with :}
+	Let $\Lambda$ be set of {\it $\lambda$-terms}. \\*
+	Let $\mathbb{T}$ be set of {\it types}.       \\*
+	A {\it statement} is a pair $(M,\sigma) \in \Lambda \times \mathbb{T}$. \\*
+	Notation $M : \sigma$. \\* 
+	The type $\sigma$ is the {\it predicate} and the term $M$ is the
+	{\it subject} of the statement.
+
+	\section{Context}
+	Let $\Gamma$ be a set of {\it statements}.	\\*
+	Then $\Gamma$ is {\it context} if it obeys following conditions:
+	\begin{align*}
+		 \forall (x,\sigma) \in \Gamma &: x \in V \cup C \\
+		 \forall s_1,s_2 \in \Gamma &: s_1 \neq s_2 \Rightarrow fst(s_1) \neq fst(s_2)
+    \end{align*}
+
+	In other words context is a set of statements with distinct variables as subjects.
+
+	\section{Statement with $\vdash$}
+	By writing $ \Gamma \vdash M : \sigma$ we say 
+	{\it statement $M : \sigma$ is derivable from context $\Gamma$ }.
+
+	\section{Inference rule}	
+	We construct valid {\it statements with $\vdash$} by using inference rules.
+
+	\section{Term generating grammar}
+
+	\section{Inhabitation tree}
+		\subsection{Inhabitation Machine}
+	\section{Roadmap}
+	\section{Conversion to SKI combinators}
+
+	\section{Genetic Programming}
+		\subsection{Term generating}
+		\subsection{Crossover}
+		\subsection{Mutation}
+
+\chapter{ Designed system }	
 	\section{Top level view}
+	\section{Term generating}
+		\subsection{A* algorithm}
+	\section{Crossover}
+		\subsection{Finding same types}
+		\subsection{Two basic options}
+		Resolve problems with free variables or avoid variables completely.
 
-	\section{Term generating}
-	\section{Crossover}
 	\section{Mutation}
+		\subsection{Using term generation}
 
-	\section{Problems}
-	In this section we will present usage of the system on solving specific problems.
+\chapter{Problems}
+	In this section will be presented usage of the system in order to solve specific problems.
 
-		\subsection{Even Parity Problem}
-		\subsection{Big Context}
-		\subsection{Fly}
-		\subsection{Simple Symbolic Regression}
-		\subsection{Artificial Ant}
-		\subsection{Boolean Alternate}
+		\section{Even Parity Problem}
+		\section{Big Context}
+		\section{Fly}
+		\section{Simple Symbolic Regression}
+		\section{Artificial Ant}
+		\section{Boolean Alternate}
 
-	\section{Conclusion}
+\chapter{Conclusion}
 
-	\section{Sandbox, YAY!}
+\chapter{Sandbox ..}
 
-	Blah!
+	...
 

### 2013-05-05 tex/thesis.tex
[tex] inference rules, term generating grammar and more...

@@
-\documentclass[12pt]{report}
+\documentclass[12pt,a4paper]{report}
+
+\setlength\textwidth{145mm}
+\setlength\textheight{247mm}
+\setlength\oddsidemargin{15mm}
+\setlength\evensidemargin{15mm}
+\setlength\topmargin{0mm}
+\setlength\headsep{0mm}
+\setlength\headheight{0mm}
 
@@
 
+\usepackage{graphicx}
+\usepackage{amsthm}
+
+%% Balíček hyperref, kterým jdou vyrábět klikací odkazy v PDF,
+%% ale hlavně ho používáme k uložení metadat do PDF (včetně obsahu).
+%% POZOR, nezapomeňte vyplnit jméno práce a autora.
+%\usepackage[ps2pdf,unicode]{hyperref}   % Musí být za všemi ostatními balíčky
+%\hypersetup{pdftitle=Typed Functional Genetic Programming}
+%\hypersetup{pdfauthor=Tomáš Křen}
+
+% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
+% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
+\makeatletter
+\def\@makechapterhead#1{
+  {\parindent \z@ \raggedright \normalfont
+   \Huge\bfseries \thechapter. #1
+   \par\nobreak
+   \vskip 20\p@
+}}
+\def\@makeschapterhead#1{
+  {\parindent \z@ \raggedright \normalfont
+   \Huge\bfseries #1
+   \par\nobreak
+   \vskip 20\p@
+}}
+\makeatother
+
+\newcommand\Vtextvisiblespace[1][.3em]{%
+  \mbox{\kern.06em\vrule height.3ex}%
+  \vbox{\hrule width#1}%
+  \hbox{\vrule height.3ex}}
+
 \title{Typed Functional Genetic Programming}
 \author{Tomáš Křen}
-\date{April 2013}
+\date{Prague 2013}
 
 \begin{document}
-	\maketitle
+
+% Trochu volnější nastavení dělení slov, než je default.
+\lefthyphenmin=2
+\righthyphenmin=2
+
+%%% Titulní strana práce
+
+\pagestyle{empty}
+\begin{center}
+
+\large
+
+Charles University in Prague
+
+\medskip
+
+Faculty of Mathematics and Physics
+
+\vfill
+
+{\bf\Large MASTER THESIS}
+
+\vfill
+
+%%% \centerline{\mbox{\includegraphics[width=60mm]{../img/logo.eps}}}
+
+%\begin{figure}[!ht]
+%  \centering
+%  \includegraphics{logo.eps}
+%  \caption{Default}\label{fig:default}
+%\end{figure}
+
+% Stačí todle odkomentovat a dát nahoře LaTeX , pak DVI->PDF  a pak View PDF
+%\includegraphics[scale=0.5]{logo.eps}
+
+\vfill
+\vspace{5mm}
+
+{\LARGE Tomáš Křen}
+
+\vspace{15mm}
+
+% Název práce přesně podle zadání
+{\LARGE\bfseries Typed Functional Genetic Programming}
+
+\vfill
+
+% Název katedry nebo ústavu, kde byla práce oficiálně zadána
+% (dle Organizační struktury MFF UK)
+%%%%Name of the department or institute
+%Department of Theoretical Computer Science and Mathematical Logic\\
+%{\small Department of Theoretical Computer Science and Mathematical Logic} \\
+{\fontsize{0.46cm}{1em}\selectfont 
+Department of Theoretical Computer Science and Mathematical Logic}
+
+\vfill
+
+\begin{tabular}{rl}
+
+Supervisor of the master thesis: & RNDr. Petr Pudlák, Ph.D. \\
+\noalign{\vspace{2mm}}
+Study programme: & Theoretical Computer Science \\ %Teoretická informatika \\
+\noalign{\vspace{2mm}}
+Specialization: & 
+%Neprocedurální programování a umělá inteligence \\
+{\fontsize{0.3cm}{1em}\selectfont 
+%Neprocedurální programování a umělá inteligence} \\
+Non-Procedural Programming and Artificial Intelligence} \\
+\end{tabular}
+
+\vfill
+
+% Zde doplňte rok
+Prague 2013
+
+\end{center}
+
+\newpage
+
+\tableofcontents
 
@@
 
-		Let $A$ be set of {\it atomic types}. \\*
+	Let $A$ be set of {\it atomic type names}. \\*
 	Then $\mathbb{T}$ is set of {\it types} inductively defined as follows:
@@
 
-	\section{Statement with :}
+\section{Statement of a form $M : \sigma$}
+
 	Let $\Lambda$ be set of {\it $\lambda$-terms}. \\*
 	Let $\mathbb{T}$ be set of {\it types}.       \\*
-	A {\it statement} is a pair $(M,\sigma) \in \Lambda \times \mathbb{T}$. \\*
-	Notation $M : \sigma$. \\* 
+	A {\it statement} $M : \sigma$ is a pair $(M,\sigma) \in \Lambda \times \mathbb{T}$. \\*
+	$M : \sigma$ is vocalized as {\it "$M$ has type $\sigma$"}.\footnote{ 
+	$M : \sigma$ can be also imagined as $M \in \sigma$ } \\*
 	The type $\sigma$ is the {\it predicate} and the term $M$ is the
@@
 \section{Context}
-	Let $\Gamma$ be a set of {\it statements}.	\\*
-	Then $\Gamma$ is {\it context} if it obeys following conditions:
+
+	Let $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$. 
+	($\Gamma$ is a set of {\it statements} of a form $M : \sigma$.)	\\*
+	Then $\Gamma$ is {\it context} if it obeys following 
+	conditions\footnote{
+	The $\pi_1$ corresponds to the projection of the first component of the Cartesian product.
+	}:
 	\begin{align*}
 		 \forall (x,\sigma) \in \Gamma &: x \in V \cup C \\
-		 \forall s_1,s_2 \in \Gamma &: s_1 \neq s_2 \Rightarrow fst(s_1) \neq fst(s_2)
+		 \forall s_1,s_2 \in \Gamma &: s_1 \neq s_2 \Rightarrow \pi_1(s_1) \neq \pi_1(s_2)
     \end{align*}
 
-	In other words context is a set of statements with distinct variables as subjects.
+	In other words context is a set of statements with distinct variables or constants as subjects.
 
-	\section{Statement with $\vdash$}
-	By writing $ \Gamma \vdash M : \sigma$ we say 
+\section{Statement of a form $\Gamma \vdash M : \sigma$}
+
+	By writing $\Gamma \vdash M : \sigma$ we say 
 	{\it statement $M : \sigma$ is derivable from context $\Gamma$ }.
 
+	We construct valid statements of form $\Gamma \vdash M : \sigma$ by using inference rules.
+
 \section{Inference rule}
-	We construct valid {\it statements with $\vdash$} by using inference rules.
+
+	Basically speaking, inference rules are used for deriving statements of a form 
+	$\Gamma \vdash M : \sigma$ from yet derived statements of such a form.
+	Those inference rules are written in the following form:
+
+	\begin{equation*}
+		\frac{\Gamma_1 \vdash M_1 : \sigma_1 \qquad
+			  \Gamma_2 \vdash M_2 : \sigma_2 \quad
+			  \dotsm \quad
+		      \Gamma_n \vdash M_n : \sigma_n}
+		     {\Gamma_{n+1} \vdash M_{n+1} : \sigma_{n+1}}
+	\end{equation*}
+
+	Suppose we have yet derived statements 
+	$\Gamma_1 \vdash M_1 : \sigma_1 ,
+	 \Gamma_2 \vdash M_2 : \sigma_2 ,
+	 \dots ,
+	 \Gamma_n \vdash M_n : \sigma_n$. 
+	It allows as to use the inference rule to derive statement
+	\mbox{ $\Gamma_{n+1} \vdash M_{n+1} : \sigma_{n+1}$ }.
+
+	For deriving statements including types of a form $(\sigma \rightarrow \tau)$ are essential those two 
+	inference rules:
+
+	\begin{equation*}
+		\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
+			  \Gamma \vdash N : \sigma }
+		     {\Gamma \vdash (M N) : \tau }
+	\end{equation*}
+
+	\begin{equation*}
+		\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
+		     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
+	\end{equation*}
+
+	This kind of inference rules allows us to derive new statements from yet derived statements, but 
+	what if we do not have any statement yet? 
+	For this purpose we have other kinds of inference rules such as {\it axiom} inference rule:
+
+	\begin{equation*}
+		\frac{( x , \sigma )  \in \Gamma}
+		     {\Gamma \vdash x : \sigma}
+	\end{equation*}
+
+	Let us consider an example statement of a form $\Gamma \vdash M : \sigma$ :
+
+	\[
+		\{\} \vdash (\lambda f . (\lambda x . (f x) )) : 
+		(\sigma \rightarrow \tau) \rightarrow ( \sigma \rightarrow \tau ) 
+	\]
+
+	This statement is derived as follows:
+
+	\begin{equation*}
+	\dfrac{
+		\dfrac{ (f,\sigma \rightarrow \tau) \in \{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \}  }
+		     { \{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \} \vdash f : \sigma \rightarrow \tau }
+		\qquad
+		\dfrac{ (x,\sigma) \in \{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \}  }
+		     { \{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \} \vdash x : \sigma }
+		 }
+		 {
+			\dfrac{		 	
+		 		\{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \} \vdash (f x) : \tau
+		 	}{
+				\dfrac{\{ (f,\sigma \rightarrow \tau) \} \vdash (\lambda x . (f x) ) : 
+				\sigma \rightarrow \tau}
+				{ \{ \} \vdash (\lambda f . (\lambda x . (f x) ) ) 
+				  : (\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau) }
+		 	}
+		 }
+	\end{equation*}
 
 	\section{Term generating grammar}
+	Inference rules are good for deriving statements of a form $\Gamma \vdash M : \sigma$, but our
+	goal is slightly different; we would like to generate many $\lambda$-terms M for a given type 
+	$\sigma$ and context $\Gamma$.
+
+	Our approach will be to take each inference rule and transform it to a rule of term generating
+	grammar. With this term generating grammar it will be much easier to reason about generating 
+	$\lambda$-terms.
+
+It won't be a grammar in classical sense because we will be operating with infinite sets of
+nonterminal symbols and rules. \footnote{TODO : mention terminal symbols - situation around 
+variables and their construction with ' symbol.}
+
+Let $Non = Type \times Context $ be our {\it nonterminal} set. 
+So for every $i \in Non$ is $i = (\sigma_i , \Gamma_i )$.
+
+Let's consider each relevant inference rule and its corresponding grammar rule.
+
+First inference rule is {\it implication elimination} also known as 
+{\it modus ponens}: 
+\[
+	\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
+		  \Gamma \vdash N : \sigma }
+	     {\Gamma \vdash (M N) : \tau }
+\]
+\\
+For every $\sigma, \tau \in \mathbb{T}$ and for every {\it context} 
+$\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$ there is a grammar rule of a form\footnote{ 
+Terminal symbols for parenthesis and normally {\it space} now \textvisiblespace \quad (for {\it function application} operator) are visually highlighted. }: 
+\[	
+	( \tau , \Gamma )  \longmapsto
+	\bigg( ( \sigma \rightarrow \tau , \Gamma ) 
+	  \mbox{ \Vtextvisiblespace[1em] } ( \sigma , \Gamma ) \bigg)
+\]
+\\
+
+Second inference rule is {\it implication introduction}: 
+\[
+	\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
+	     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
+\]
+\\
+$\forall \sigma, \tau \in \mathbb{T}$ 
+$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
+$\forall x \in V $ such that there is no $(x,\rho) \in \Gamma$ 
+there is a grammar rule:
+\[ 
+	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
+	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
+\]
+\\
+
+Third inference rule is {\it axiom}: 
+\[
+		\frac{( x , \sigma )  \in \Gamma}
+		     {\Gamma \vdash x : \sigma}
+\]
+\\
+$\forall \sigma \in \mathbb{T}$ 
+$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
+$\forall x \in V \cup C $ such that $(x,\sigma) \in \Gamma$ 
+there is a grammar rule:
+\[ 
+	( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}}
+\]
+\\
+
+We will demonstrate $\lambda$-term generation on example. 
+Again on $(\lambda f . (\lambda x . (f x) ))$. 
+We would like to generate $\lambda$-term of a type 
+$(\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau)$
+with $\Gamma = \{\}$.
+\begin{align*}
+	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
+	\longmapsto & \bigg( \mbox{ {\Large $\lambda$ f . }}
+	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
+	\quad \bigg)
+	\\
+	\longmapsto & 
+	\bigg( \mbox{ {\Large $\lambda$ f . }}
+		\bigg( \mbox{ {\Large $\lambda$ x . }}
+	  	 	( \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
+		\quad \bigg)  	 
+	\quad \bigg)
+	\\
+	\longmapsto & 
+	\bigg( \mbox{ {\Large $\lambda$ f . }}
+		\bigg( \mbox{ {\Large $\lambda$ x . }}	  	 	
+	  	 	\bigg( 
+	  	 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
+			  \mbox{ \Vtextvisiblespace[1em] } 
+			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \bigg) 
+		\quad \bigg)  	 
+	\quad \bigg)
+	\\
+	\longmapsto & 
+	\bigg( \mbox{ {\Large $\lambda$ f . }}
+		\bigg( \mbox{ {\Large $\lambda$ x . }}	  	 	
+	  	 	\bigg( 
+	  	 	  \mbox{ {\Large f}} 
+			  \mbox{ \Vtextvisiblespace[1em] } 
+			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \bigg) 
+		\quad \bigg)  	 
+	\quad \bigg)		
+	\\
+	\longmapsto & 
+	\bigg( \mbox{ {\Large $\lambda$ f . }}
+		\bigg( \mbox{ {\Large $\lambda$ x . }}	  	 	
+	  	 	\bigg( 
+	  	 	  \mbox{ {\Large f}} 
+			  \mbox{ \Vtextvisiblespace[1em] } 
+			  \mbox{{\Large x}} \bigg) 
+		\quad \bigg)  	 
+	\quad \bigg)
+\end{align*}
+
+\subsection{"Barendregt-like" inference and grammar rules}
+
+Inference rule 1: 
+\[
+	\frac{\Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} \vdash M : \alpha }
+	     {\Gamma \vdash (\lambda x_1 \dots x_n . M) : 
+	     \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha }
+\]
+\\
+... there is a grammar rule:
+\[ 
+	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \longmapsto
+	\bigg( \mbox{ {\Large 
+	$\lambda x_1 \dots x_n .$ 
+	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
+\]
+\\
+
+Inference rule 2: 
+\[
+	\frac{ (f , \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha ) \in \Gamma \qquad
+	       \Gamma \vdash M_1 : \tau_1 \quad
+	       \dotsm \quad
+	       \Gamma \vdash M_n : \tau_n        
+	      }
+	     {\Gamma \vdash (f M_1 \dots M_n) : \alpha}
+\]
+\\
+... there is a grammar rule:
+\[ 
+	( \alpha , \Gamma )  \longmapsto
+	\bigg( \mbox{ {\Large f }}
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  ( \tau_1 , \Gamma )
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  \dots
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  ( \tau_n , \Gamma )
+	  \quad \bigg)
+\]
+\\
 
@@
 
-\chapter{Sandbox ..}
-
-	...
-
-	% Example formula
-
-  	\begin{align}
-    	E &= mc^2                              \\
-    	m &= \frac{m_0}{\sqrt{1-\frac{v^2}{c^2}}}
-  	\end{align}
-
 \end{document}

### 2013-05-06 v3/Eva.hs
folder v3 created

@@ -1,0 +1,8 @@
+module Eva where
+
+import Data.List
+import Data.Map.Strict (Map)
+import qualified Data.Map.Strict as Map
+
+import System.Random
+import Control.Monad.State

### 2013-05-06 v3/Eva.hs
some preparations in v3

@@ -6,3 +6,19 @@ import qualified Data.Map.Strict as Map
 
 import System.Random
 import Control.Monad.State
+import Data.Typeable
+
+import Heval (hevalsWith)
+
+
+type Eva = StateT EvaState IO
+
+data EvaState = EvaState {
+     evaGen   :: StdGen ,
+     evaJobID :: Int 
+  }
+
+
+
+evalsWith :: (Typeable a) => String -> [String] -> a -> Eva [a]
+evalsWith file strs as = liftIO $ hevalsWith file strs as

### 2013-05-06 tex/thesis.tex
some preparations in v3
[tex] ...

@@
 
+\usepackage[utf8]{inputenc}
+
+\usepackage{qtree}
+
 \usepackage{amsmath}
 \usepackage{amsfonts}
-\usepackage[utf8]{inputenc}
+\usepackage{amsthm}
 
 \usepackage{graphicx}
-\usepackage{amsthm}
+
+\usepackage{hyperref}
 
@@
 
-% Stačí todle odkomentovat a dát nahoře LaTeX , pak DVI->PDF  a pak View PDF
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Stačí todle odkomentovat a dát nahoře LaTeX (F2) , pak DVI->PDF (F9)  a pak View PDF
 %\includegraphics[scale=0.5]{logo.eps}
+\includegraphics[scale=0.15]{logomff.png}
 
@@
 
+%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
+%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
+
+%%% Na tomto místě mohou být napsána případná poděkování (vedoucímu práce,
+%%% konzultantovi, tomu, kdo zapůjčil software, literaturu apod.)
+
+%% on tam měl %% \openright
+
+\noindent
+Dedication.
+
+\newpage
+
+%%% Strana s čestným prohlášením k diplomové práci
+
+\vglue 0pt plus 1fill
+
+\noindent
+I declare that I carried out this master thesis independently, and only with the cited
+sources, literature and other professional sources.
+
+\medskip\noindent
+I understand that my work relates to the rights and obligations under the Act No.
+121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
+University in Prague has the right to conclude a license agreement on the use of this
+work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
+
+\vspace{10mm}
+
+\hbox{\hbox to 0.5\hsize{%
+In ........ date ............
+\hss}\hbox to 0.5\hsize{%
+signature of the author
+\hss}}
+
+\vspace{20mm}
+\newpage
+
+%%% Povinná informační strana diplomové práce
+
+\vbox to 0.5\vsize{
+\setlength\parindent{0mm}
+\setlength\parskip{5mm}
+
+Název práce:
+Název práce
+% přesně dle zadání
+
+Autor:
+Jméno a příjmení autora
+
+Katedra:  % Případně Ústav:
+Název katedry či ústavu, kde byla práce oficiálně zadána
+% dle Organizační struktury MFF UK
+
+Vedoucí diplomové práce:
+Jméno a příjmení s tituly, pracoviště
+% dle Organizační struktury MFF UK, případně plný název pracoviště mimo MFF UK
+
+Abstrakt:
+% abstrakt v rozsahu 80-200 slov; nejedná se však o opis zadání diplomové práce
+
+Klíčová slova:
+% 3 až 5 klíčových slov
+
+\vss}\nobreak\vbox to 0.49\vsize{
+\setlength\parindent{0mm}
+\setlength\parskip{5mm}
+
+Title:
+% přesný překlad názvu práce v angličtině
+
+Author:
+Jméno a příjmení autora
+
+Department:
+Název katedry či ústavu, kde byla práce oficiálně zadána
+% dle Organizační struktury MFF UK v angličtině
+
+Supervisor:
+Jméno a příjmení s tituly, pracoviště
+% dle Organizační struktury MFF UK, případně plný název pracoviště
+% mimo MFF UK v angličtině
+
+Abstract:
+% abstrakt v rozsahu 80-200 slov v angličtině; nejedná se však o překlad
+% zadání diplomové práce
+
+Keywords:
+% 3 až 5 klíčových slov v angličtině
+
+\vss}
+
+\newpage
+
 \tableofcontents
 
-\chapter{Introduction}
+%\chapter{Introduction}
+\chapter*{Introduction}
+\addcontentsline{toc}{chapter}{Introduction}
 
@@
 
-	For deriving statements including types of a form $(\sigma \rightarrow \tau)$ are essential those two 
+	For deriving statements including types of a form 
+	$(\sigma \rightarrow \tau)$ are essential those two 
 	inference rules:
@@
 	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
-	\longmapsto & \bigg( \mbox{ {\Large $\lambda$ f . }}
+	\longmapsto & \Big( \mbox{ {\Large $\lambda$f.}}
 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
-	\quad \bigg)
+	~ \Big)
 	\\
 	\longmapsto & 
-	\bigg( \mbox{ {\Large $\lambda$ f . }}
-		\bigg( \mbox{ {\Large $\lambda$ x . }}
+	\Big( \mbox{ {\Large $\lambda$f. }}
+		\Big( \mbox{ {\Large $\lambda$x. }}
 	  	 	( \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
-		\quad \bigg)  	 
-	\quad \bigg)
+		~ \Big)  	 
+	~ \Big)
 	\\
 	\longmapsto & 
-	\bigg( \mbox{ {\Large $\lambda$ f . }}
-		\bigg( \mbox{ {\Large $\lambda$ x . }}	  	 	
-	  	 	\bigg( 
+	\Big( \mbox{ {\Large $\lambda$f. }}
+		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
+	  	 	\Big( 
 	  	 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
 			  \mbox{ \Vtextvisiblespace[1em] } 
-			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \bigg) 
-		\quad \bigg)  	 
-	\quad \bigg)
+			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} )  \Big) 
+		~ \Big)  	 
+	 ~ \Big)
 	\\
 	\longmapsto & 
-	\bigg( \mbox{ {\Large $\lambda$ f . }}
-		\bigg( \mbox{ {\Large $\lambda$ x . }}	  	 	
-	  	 	\bigg( 
+	\Big( \mbox{ {\Large $\lambda$f. }}
+		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
+	  	 	\Big( 
 	  	 	  \mbox{ {\Large f}} 
 			  \mbox{ \Vtextvisiblespace[1em] } 
-			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \bigg) 
-		\quad \bigg)  	 
-	\quad \bigg)		
+			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \Big) 
+		~ \Big)  	 
+	~ \Big)		
 	\\
 	\longmapsto & 
-	\bigg( \mbox{ {\Large $\lambda$ f . }}
-		\bigg( \mbox{ {\Large $\lambda$ x . }}	  	 	
-	  	 	\bigg( 
+	\Big( \mbox{ {\Large $\lambda$f. }}
+		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
+	  	 	\Big( 
 	  	 	  \mbox{ {\Large f}} 
 			  \mbox{ \Vtextvisiblespace[1em] } 
-			  \mbox{{\Large x}} \bigg) 
-		\quad \bigg)  	 
-	\quad \bigg)
+			  \mbox{{\Large x}} \Big) 
+		~ \Big)  	 
+	~ \Big)
 \end{align*}
@@
 \subsection{"Barendregt-like" inference and grammar rules}
+\label{barlike}
 
@@
 \section{Inhabitation tree}
+
+Now we will introduce \textit{Inhabitation tree}, structure slightly different from
+\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Henk Barendregt.
+We can think about Inhabitation tree as about unfolded Inhabitation machine.
+\\\\
+\textbf{TODO}: TALK ABOUT And-Or-graph \\
+\textbf{TODO}: TALK ABOUT "Barendregt-like" subsection \ref{barlike}
+\\\\
+Let's see some example:
+
+\begin{align*}
+\sigma =& ~ \mathbb{B} \rightarrow  \mathbb{B} \rightarrow  \mathbb{B} \\ 
+\Gamma =& \{ ~ true : \mathbb{B}  \\
+        &  , ~ nand :  \mathbb{B} \rightarrow \mathbb{B} \rightarrow \mathbb{B} ~ \}
+\end{align*}
+
+\Tree
+[.\text{ $\mathbb{B} \rightarrow \mathbb{B} \rightarrow \mathbb{B}$ }  
+	[.\textbf{$\lambda$x_1 x_2 } 
+		[.\text{ $\mathbb{B}$ } 
+			\textbf{true}  
+			[.\textbf{nand} 
+				\qroof{ ~~ $\dotsb$ ~~ }.\text{ $\mathbb{B}$ }
+				\qroof{ ~~ $\dotsb$ ~~ }.\text{ $\mathbb{B}$ } 
+			]
+			\textbf{x_1}
+			\textbf{x_2}
+		]
+	]
+]
+
 		\subsection{Inhabitation Machine}
@@
 	\section{Top level view}
+		\subsection{Comments about main source files}
+			\subsubsection{ Eva.hs }
 	\section{Term generating}
@@
 
-\chapter{Conclusion}
+%\chapter{Conclusion}
+\chapter*{Conclusion}
+\addcontentsline{toc}{chapter}{Conclusion}
+
+\begin{thebibliography}{9}
+
+\bibitem{barendregt10}
+  Henk Barendregt, Wil Dekkers, Richard Statman,
+  \emph{Lambda Calculus With Types}.
+  Cambridge University Press,
+  2010. \\
+  \url{http://www.cs.ru.nl/~henk/book.pdf}
+
+\end{thebibliography}
 

### 2013-05-06 tex/thesis.tex
[tex] Inhabitation tree and more ..

@@
 
+	\textbf{TODO} 
+	\begin{itemize} 
+		\item TALK ABOUT "parenthesis" conventions. 
+	\end{itemize}
+
 \section{Type}
+\label{deftype}
 
@@
 	\end{align*}
+
+\textbf{TODO} 
+\begin{itemize}
+	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
+	\item TALK ABOUT arrow/parenthesis conventions. 
+\end{itemize}
 
@@
 	In other words context is a set of statements with distinct variables or constants as subjects.
+	\\\\
+	\textbf{TODO:} TALK ABOUT Context represents library/building blocks.
 
@@
 \section{Term generating grammar}
+
 Inference rules are good for deriving statements of a form $\Gamma \vdash M : \sigma$, but our
@@
 \[
-	\frac{ (f , \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha ) \in \Gamma \qquad
-	       \Gamma \vdash M_1 : \tau_1 \quad
+	\frac{ (f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma \qquad
+	       \Gamma \vdash M_1 : \rho_1 \quad
 	       \dotsm \quad
-	       \Gamma \vdash M_n : \tau_n        
+	       \Gamma \vdash M_m : \rho_m        
 	      }
-	     {\Gamma \vdash (f M_1 \dots M_n) : \alpha}
+	     {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
 \]
@@
 	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \tau_1 , \Gamma )
+	  ( \rho_1 , \Gamma )
 	  \mbox{ \Vtextvisiblespace[1em] } 
@@
 	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \tau_n , \Gamma )
+	  ( \rho_m , \Gamma )
 	  \quad \bigg)
 \]
-\\
+\\\\
+\textbf{TODO} 
+\begin{itemize}
+	\item SHOW correctness of those inference rules by composing them of 
+		  $E^{\rightarrow}$, $I^{\rightarrow}$ and \textit{axiom}.
+	\item SHOW more examples of inference rules transformed into grammar rules.
+	\item DESCRIBE general algorithm for this transformation.
+	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
+	\item TALK ABOUT $\beta \eta$-normal form which is generated by this method.
+\end{itemize}
 
@@
 We can think about Inhabitation tree as about unfolded Inhabitation machine.
-\\\\
-\textbf{TODO}: TALK ABOUT And-Or-graph \\
-\textbf{TODO}: TALK ABOUT "Barendregt-like" subsection \ref{barlike}
-\\\\
-Let's see some example:
+The motivation for using Inhabitation trees is belief that it will help us 
+reason about generation of $\lambda$-terms 
+of a given type $\sigma$ and with a given context $\Gamma$.
+
+\subsection{Definition of Inhabitation tree}
+
+\textit{Inhabitation tree} is a \textit{rooted tree}, possibly infinite. 
+It has two types of nodes:
+
+\begin{samepage}
+\begin{itemize}
+  \item Type nodes   
+  			- containing type $\sigma \in \mathbb{T}$ 
+  			- aka "OR-node" , Nonterminal-node. 
+  \item Symbol nodes 
+  			%- containing $ x^{*} \in \mathfrak P \left(V\right) -blobost pač ordered $ or $c \in C$
+  			- containing "$\lambda$-head" (nonempty finite sequence of variable names) or constant
+  			  name. 
+  			- aka "AND-node" , Terminal-node.
+\end{itemize}
+\end{samepage}
+
+We construct Inhabitation tree for given type $\sigma$ and context $\Gamma$.\\
+We will define Inhabitation tree by describing its construction for a given $(\sigma,\Gamma)$.
+Notice that it will closely follow the rules from \ref{barlike}:
+
+\begin{itemize}
+\item The root of Inhabitation tree for $(\sigma,\Gamma)$ is 
+      \textit{type node} with $\sigma$ as type.
+\item All \textit{type nodes} have as child nodes only \textit{symbol nodes}. 
+\item And all \textit{symbol nodes} have as child nodes only \textit{type nodes}. 
+\end{itemize}
+
+Now we will resolve the child nodes of the root node.
+
+There are two cases of $\sigma$ (recall \ref{deftype}): 
+\begin{description}
+	\item[Atomic type] 
+		$\sigma = \alpha $ where $\alpha \in A$. 
+	\item[Function type] 
+		$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
+		where $n \geq 1, \alpha \in A$.
+\end{description}
+
+First case \textbf{Atomic type} --- i.e., $\sigma = \alpha$ where $\alpha \in A$:\\
+For every $(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha) \in \Gamma$
+where $\alpha \in A$ there is a child \textit{symbol node} of the root containing constant name $f$.
+This symbol node containing $f$ has $m$ child subtrees corresponding to Inhabitation trees for 
+$(\rho_1,\Gamma),\dots,(\rho_n,\Gamma)$.
+
+~
+
+Compare this case with corresponding grammar rule:
+\[ 
+	( \alpha , \Gamma )  \longmapsto
+	\bigg( \mbox{ {\Large f }}
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  ( \rho_1 , \Gamma )
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  \dots
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  ( \rho_m , \Gamma )
+	  \quad \bigg)
+\]
+
+Second case \textbf{Function type} --- i.e., 
+$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
+where $n~\geq~1, \alpha~\in~A$:\\
+For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.
+
+There is one and only one child \textit{symbol node} of the root containing "$\lambda$-head" 
+$\lambda x_1 \dots x_n$ which stands for sequence of variable names $(x_1,\dots,x_n)$.
+This symbol node containing $\lambda x_1 \dots x_n$
+has one and only one child subtree corresponding to Inhabitation trees for 
+$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$.
+
+~
+
+Compare this case with corresponding grammar rule:
+\[ 
+	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \longmapsto
+	\bigg( \mbox{ {\Large 
+	$\lambda x_1 \dots x_n .$ 
+	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
+\]
+~\\
+Let's consider following $(\sigma,\Gamma)$ as a simple example:
 
@@
 \end{align*}
+
+This particular $(\sigma,\Gamma)$ results in the following tree:\\
 
@@
 
+~
+
+Second example features our well known example:
+
+\begin{align*}
+\sigma =& ~ (\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau \\ 
+\Gamma =& \{ \}
+\end{align*}
+
+Which results in following tree:
+
+\Tree
+[.\text{ $(\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau $ } 
+	[.\textbf{$\lambda$ f x }	
+		[.\text{ $\tau$ }		
+			[.\textbf{f} 
+				[.\text{ $\delta$ }
+					\textbf{x}					
+				]
+			]
+		]
+	]
+]
+
+\subsection{And-or tree and searching in Inhabitation tree}
+Let us consider following definition of 
+\textit{And-or tree}\footnote{
+\textbf{TODO}: Mention that on WIKI there is more general definition, but 
+for our purposes is this one sufficient.}:
+
+\textit{And-or tree} is a rooted tree where each node is labeled as either \textit{and-node} 
+or\footnote{xor} \textit{or-node}.
+
+By \textit{solving} And-or tree $T$ we mean finding $T'$
+subtree of $T$ such that it follows these conditions: 
+\begin{itemize}
+	\item The root of $T'$ is the root of $T$.
+	\item Each \textit{and-node} in $T'$ has all the child nodes as in $T$.
+	\item Each \textit{or-node}  in $T'$ has precisely one child 
+	      node.\footnote{\textbf{TODO}: MENTION why precisely one 
+	      and not at least one ..or CHANGE the def. }   
+\end{itemize}
+~
+Let us now consider following labeling of Inhabitation tree:
+
+\begin{itemize}
+  \item \textbf{Type nodes}   are labeled as \textbf{or-nodes}.   
+  \item \textbf{Symbol nodes} are labeled as \textbf{and-nodes}.
+\end{itemize}
+
+This labeling has following justification:
+
+Selection of exactly one
+child node in \textit{type node} corresponds to selection of exactly one
+grammar rule in order to rewrite nonterminal symbol.
+
+Selection of all the child nodes in \textit{symbol node} corresponds to 
+rewriting all the nonterminal symbols in string that is being generated.
+
+~\\\\
+\textbf{TODO}: TALK ABOUT : By \textit{solving} And-or tree for Inhabitation tree
+               we obtain that typed tree notation for typed $\lambda$-term,
+               which must be firstly INTRODUCED.   \\
+\textbf{TODO}: TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}
+\\\\
+
+\subsection{Our approach to \textit{solving} Inhabitation machine}
+
 \subsection{Inhabitation Machine}
+
 \section{Roadmap}

### 2013-05-08 tex/thesis.tex
[tex] proofs of barendrecht inference rules ...
[tex] tree representations of terms and solving of Inhab tree

@@
 
+\section{Genetic Programming}
+	\subsection{Term generating}
+	\subsection{Crossover}
+	\subsection{Mutation}
+
 \section{Lambda term}
+\label{deflam}
 
@@
 	\begin{align*}
-		 x   \in V                 &\Rightarrow     x   \in \Lambda \\
-		 c   \in C                 &\Rightarrow     c   \in \Lambda \\
+		 x   \in V \cup C          &\Rightarrow     x   \in \Lambda \\
 		 M,N \in \Lambda           &\Rightarrow ( M N ) \in \Lambda \\
@@
 	\begin{itemize} 
-		\item TALK ABOUT "parenthesis" conventions. 
+		\item TALK ABOUT "parenthesis" conventions (and packing of lambda abstractions). 
 	\end{itemize}
@@
 \\
+Proof of correctness:
+\[
+	\dfrac{
+		\dfrac
+		 {\Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} \vdash M : \alpha}
+		 {\dfrac
+		   {\Gamma \cup \{ (x_1,\tau_1),\dots,(x_{n-1},\tau_{n-1})\} \cup 
+		                \{(x_n,\tau_n) \} \vdash M : \alpha}
+		   {\dfrac{\Gamma \cup \{ (x_1,\tau_1),\dots,(x_{n-1},\tau_{n-1})\}  
+		                \vdash (\lambda x_n . M) : \tau_n \rightarrow \alpha}
+				  { \vdots }		   
+		   }
+		 }		 
+	 }
+	     {\Gamma \vdash (\lambda x_1 \dots x_n . M) : 
+	     \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha }
+\]
+\\
 ... there is a grammar rule:
@@
 	      }
+	     {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
+\]
+\\
+Proof of correctness (\textbf{TODO REPAIR} Conceptually it is ok but there is sazba-bug somewhere): 
+\[
+   \dfrac
+     {\dfrac
+      {\dfrac
+       {\dfrac         
+         {\dfrac  
+          {\dfrac
+           {\boxed{(f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma}}
+           {\Gamma \vdash f : \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha}
+           \quad
+           \boxed{\Gamma \vdash M_1 : \rho_1} }
+          {\Gamma \vdash (f M_1) : \rho_2 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha }
+          }{\vdots} 
+         \quad 
+         \ddots }
+       {\Gamma \vdash (f M_1 \dots M_{m-2}) : \rho_{m-1} \rightarrow \rho_m \rightarrow \alpha}
+       \quad
+       \boxed{\Gamma \vdash M_{m-1} : \rho_{m-1}}  }
+      {\Gamma \vdash (f M_1 \dots M_{m-1}) : \rho_m \rightarrow \alpha}       
+      \quad 
+      \boxed{\Gamma \vdash M_m : \rho_m} }
 	 {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
@@
 
-~\\\\
-\textbf{TODO}: TALK ABOUT : By \textit{solving} And-or tree for Inhabitation tree
-               we obtain that typed tree notation for typed $\lambda$-term,
-               which must be firstly INTRODUCED.   \\
-\textbf{TODO}: TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}
+The motivation for defining \textit{solving} of a And-or tree the way we did is that
+a found tree $T'$ corresponds to generated $\lambda$-term. 
+In order to understand this correspondence let's now talk about various tree representations
+of $\lambda$-terms.
+
+\subsection{Tree representations of $\lambda$-terms}
+
+From the definition of $\lambda$-term (\ref{deflam}) we can straightforwardly derive 
+the classical tree representation for $\lambda$-terms. Term M is translated into tree $T[M]$ by following rules:
+
+\begin{itemize}
+	\item $x \in V \cup C$ translates into \textit{leaf} $x$.
+	\item $(M N)$ translates into tree\\
+		\Tree
+			[.@	
+		 		\text{$T[M]$}
+		 		\text{$T[N]$}		 			
+			] 
+	\item $\lambda x . M$ translates into tree\\
+	 	\Tree
+			[.\text{$\lambda x$}	
+		 	 	\text{$T[M]$}	
+			] 
+\end{itemize}
+
+We can enhance this representation by compressing consecutive lambda abstractions into one
+tree node like this:
+
+\begin{itemize}
+	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
+	 	\Tree
+			[.\text{$\lambda x_1 \dots x_n$ }	
+		 	 	\text{$T[M]$}	
+			] 
+\end{itemize}
+
+As this representation comes directly from definition it is evident 
+that it covers all possible $\lambda$-terms.
+
+For representing expressions as trees it is however more common use a little different
+representation. It will also be the representation suitable for showing 
+that \textit{solving} Inhabitation tree generates wanted $\lambda$-term.
+
+\begin{itemize}
+    \item $x \in V \cup C$ translates into \textit{leaf} $x$.
+	\item $(f M_1 M_2 \dots M_n)$ where $f \in V \cup C, n \geq 1$ translates into tree\\
+		\Tree
+			[.f	
+		 		\text{$T[M_1]$}
+		 		\text{$T[M_2]$}
+		 		\text{$\dots$}
+		 		\text{$T[M_n]$}		 				 			
+			] 
+	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
+	 	\Tree
+			[.\text{$\lambda x_1 \dots x_n$ }	
+		 	 	\text{$T[M]$}	
+			] 
+\end{itemize}
+
+Notice that this representation does not cover all $\lambda$-terms, 
+e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.  
 \\\\
+Let's now consider representation for \textit{typed $\lambda$-terms}.
+Straightforward approach would be to add to each node a type entry which 
+would be the type of the $\lambda$-term corresponding to subtree having this
+node as the root node.
+
+Approach more suitable for our purpose is to add a special type node above each node.
+More specifically:
+
+Let us consider tree $t$ corresponding to a $\lambda$-term of a type
+$\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
+Then corresponding tree $TT[t]$ for typed $\lambda$-term is 
+obtained from the tree $t$ as follows:
+
+\begin{equation*}
+\mbox{ 
+TT[
+\Tree
+	[.r 	
+	  	  \text{$s_1$}
+		  \text{$s_2$}
+		  \text{$\dots$}
+		  \text{$s_n$}
+	] 
+}]=
+\mbox{
+\Tree
+	[.\text{$\sigma$ }
+	    [.r 	
+	  	  \text{$TT[s_1]$}
+		  \text{$TT[s_2]$}
+		  \text{$\dots$}
+		  \text{$TT[s_n]$}
+		]	  	
+	] 
+}
+\end{equation*}
+
+Now we can finally put the pieces together. 
+Every solution to a Inhabitation tree has this just described tree form of a typed $\lambda$-term.
+
+~\\
+\textbf{TODO} 
+\begin{itemize}
+\item EXAMPLES of tree representations of $\lambda$-terms  
+\item TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
+		Things about $\beta\eta$ normal form, etc.    
+\end{itemize}
 
@@
 
+\subsubsection{A* algorithm}
+
 \subsection{Inhabitation Machine}
+\textbf{TODO} 
+\begin{itemize}
+\item DESCRIBE Inhabitation Machine...
+\end{itemize}
 
@@
 	\section{Conversion to SKI combinators}
-
-	\section{Genetic Programming}
-		\subsection{Term generating}
-		\subsection{Crossover}
-		\subsection{Mutation}
 

### 2013-05-08 tex/thesis.tex
[tex] some new macros...

@@
 \usepackage[utf8]{inputenc}
-
 \usepackage{qtree}
-
+\usepackage{xspace}
 \usepackage{amsmath}
@@
 \usepackage{amsthm}
+\usepackage{graphicx}
+\usepackage{hyperref}
 
-\usepackage{graphicx}
+\newcommand{\Lets}{Let us }
+\newcommand{\lterm}{$\lambda$-term\xspace}
+\newcommand{\lterms}{$\lambda$-terms\xspace}
 
-\usepackage{hyperref}
+\newcommand{\turst}[3]{$#1 \vdash #2 : #3$\xspace}
+\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
 
@@
 
-	Let us first say some basic definitions.
+	\Lets first say some basic definitions.
 
@@
 	Let $C$ be set of {\it constant names}.	 \\*		
-	Then $\Lambda$ is set of {\it $\lambda$-terms} inductively defined as follows:
+	Then $\Lambda$ is set of {\it \lterms} inductively defined as follows:
 
@@
 
-	Let $\Lambda$ be set of {\it $\lambda$-terms}. \\*
+	Let $\Lambda$ be set of {\it \lterms}. \\*
 	Let $\mathbb{T}$ be set of {\it types}.       \\*
@@
 
-\section{Statement of a form $\Gamma \vdash M : \sigma$}
+\section{Statement of a form \GMS}
 
-	By writing $\Gamma \vdash M : \sigma$ we say 
+	By writing \GMS we say 
 	{\it statement $M : \sigma$ is derivable from context $\Gamma$ }.
 
-	We construct valid statements of form $\Gamma \vdash M : \sigma$ by using inference rules.
+	We construct valid statements of form \GMS by using inference rules.
 
@@
 	Basically speaking, inference rules are used for deriving statements of a form 
-	$\Gamma \vdash M : \sigma$ from yet derived statements of such a form.
+	\GMS from yet derived statements of such a form.
 	Those inference rules are written in the following form:
@@
 
-	Let us consider an example statement of a form $\Gamma \vdash M : \sigma$ :
+	\Lets consider an example statement of a form \GMS :
 
@@
 
-Inference rules are good for deriving statements of a form $\Gamma \vdash M : \sigma$, but our
-goal is slightly different; we would like to generate many $\lambda$-terms M for a given type 
+Inference rules are good for deriving statements of a form \GMS, but our
+goal is slightly different; we would like to generate many \lterms M for a given type 
 $\sigma$ and context $\Gamma$.
@@
 grammar. With this term generating grammar it will be much easier to reason about generating 
-$\lambda$-terms.
+\lterms.
 
@@
 
-Let's consider each relevant inference rule and its corresponding grammar rule.
+\Lets consider each relevant inference rule and its corresponding grammar rule.
 
@@
 
-We will demonstrate $\lambda$-term generation on example. 
+We will demonstrate \lterm generation on example. 
 Again on $(\lambda f . (\lambda x . (f x) ))$. 
-We would like to generate $\lambda$-term of a type 
+We would like to generate \lterm of a type 
 $(\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau)$
@@
 The motivation for using Inhabitation trees is belief that it will help us 
-reason about generation of $\lambda$-terms 
+reason about generation of \lterms 
 of a given type $\sigma$ and with a given context $\Gamma$.
@@
 ~\\
-Let's consider following $(\sigma,\Gamma)$ as a simple example:
+\Lets consider following $(\sigma,\Gamma)$ as a simple example:
 
@@
 \subsection{And-or tree and searching in Inhabitation tree}
-Let us consider following definition of 
+\Lets consider following definition of 
 \textit{And-or tree}\footnote{
@@
 ~
-Let us now consider following labeling of Inhabitation tree:
+\Lets now consider following labeling of Inhabitation tree:
 
@@
 The motivation for defining \textit{solving} of a And-or tree the way we did is that
-a found tree $T'$ corresponds to generated $\lambda$-term. 
+a found tree $T'$ corresponds to generated \lterm. 
 In order to understand this correspondence let's now talk about various tree representations
-of $\lambda$-terms.
+of \lterms.
 
-\subsection{Tree representations of $\lambda$-terms}
+\subsection{Tree representations of \lterms}
 
-From the definition of $\lambda$-term (\ref{deflam}) we can straightforwardly derive 
-the classical tree representation for $\lambda$-terms. Term M is translated into tree $T[M]$ by following rules:
+From the definition of \lterm (\ref{deflam}) we can straightforwardly derive 
+the classical tree representation for \lterms. Term M is translated into tree $T[M]$ by following rules:
 
@@
 As this representation comes directly from definition it is evident 
-that it covers all possible $\lambda$-terms.
+that it covers all possible \lterms.
 
@@
 representation. It will also be the representation suitable for showing 
-that \textit{solving} Inhabitation tree generates wanted $\lambda$-term.
+that \textit{solving} Inhabitation tree generates wanted \lterm.
 
@@
 
-Notice that this representation does not cover all $\lambda$-terms, 
+Notice that this representation does not cover all \lterms, 
 e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.  
 \\\\
-Let's now consider representation for \textit{typed $\lambda$-terms}.
+\Lets now consider representation for \textit{typed \lterms}.
 Straightforward approach would be to add to each node a type entry which 
-would be the type of the $\lambda$-term corresponding to subtree having this
+would be the type of the \lterm corresponding to subtree having this
 node as the root node.
@@
 
-Let us consider tree $t$ corresponding to a $\lambda$-term of a type
+\Lets consider tree $t$ corresponding to a \lterm of a type
 $\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
-Then corresponding tree $TT[t]$ for typed $\lambda$-term is 
+Then corresponding tree $TT[t]$ for typed \lterm is 
 obtained from the tree $t$ as follows:
@@
 Now we can finally put the pieces together. 
-Every solution to a Inhabitation tree has this just described tree form of a typed $\lambda$-term.
+Every solution to a Inhabitation tree has this just described tree form of a typed \lterm.
 
@@
 \begin{itemize}
-\item EXAMPLES of tree representations of $\lambda$-terms  
+\item EXAMPLES of tree representations of \lterms  
 \item TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\

### 2013-05-09 v3/DrawIM.hs
v3 now running ok (but..)
V zásedě jediná podstatná zmena oproti v2 je předělání Eva monády (+
pročištění).
Z nějakýho důvodu to ale je asi míň efektivní než to předtim což se
projevuje
u BA pro pop 5000 to hodi stackoverflow co to předtim v pohodě zvladalo.

@@ -1,0 +1,184 @@
+module DrawIM (imGraphInJSON) where
+
+import Data.List  ((\\),nub)
+import Data.Maybe (catMaybes)
+
+import qualified Data.Map as Map
+import Data.Map (Map)
+
+
+import TTerm (Symbol,Typ(..),Context,typeArgs)
+import Utils (newSymbol')
+
+import Text.JSON
+import JSONUtils
+
+
+type IMGraph = Map Typ [ForkEdge]
+
+
+-- ( "Edge label" , "Vertices where the fork-edge (= edge with one begin and multiple ends) is pointing" )
+type ForkEdge = ( EdgeLabel , [Typ] )
+
+
+data EdgeLabel = LLams [(Symbol,Typ)] 
+               | LVar  Symbol Typ 
+
+-- Structure used for IM construction. 
+-- It is graph with fork-edges instead of edges. 
+-- Vertices also contain Context needed during IM construction.
+type PreGraph  = Map Typ PreVertex
+type PreVertex = ( [ForkEdge] , Context )
+
+data UpdateCmd = MkVertex Typ Context
+               | DeltaCtx Typ Context PreVertex
+
+
+--  var dummyGraph = [
+--    { node : 'a' , opts : {name:'foobar'}               } , 
+--    { node : 'b' , opts : {color:'red',name:'Int',sq:1} } ,  
+--    { node : 'c' , opts : {sq:1}                        } ,
+--    { node : 'd' , opts : {sq:1}                        } ,
+--    { node : 'e' , opts : {sq:1}                        } ,   
+--    
+--    { edge : ['a','b'] , opts : {dir:1,n:2}             } ,
+--    { edge : ['c','c'] , opts : {}                      } ,
+--    { edge : ['a','c'] , opts : {}                      } ,
+--    { edge : ['a','d'] , opts : {}                      } ,
+--    { edge : ['a','e'] , opts : {}                      } 
+--  ];
+
+
+imGraphInJSON :: Typ -> Context -> JSValue
+imGraphInJSON typ ctx = imGraph2JSON typ $ mkIMGraph typ ctx
+
+imGraph2JSON :: Typ -> IMGraph -> JSValue
+imGraph2JSON typ graph = jsArr $ imTypes2nodes typ graph
+
+
[…]

### 2013-05-09 v3/Eva.hs

@@ -1,3 +1,5 @@
+{-# LANGUAGE TypeSynonymInstances, FlexibleInstances #-}
+
 module Eva where
 
 import Data.List
@@ -7,18 +9,105 @@ import qualified Data.Map.Strict as Map
 import System.Random
 import Control.Monad.State
 import Data.Typeable
+import Data.Time.Clock
 
-import Heval (hevalsWith)
+import Text.JSON ( encode , JSValue )
+
+import ServerInterface (jsEmptyObj, stdoutCmd, writeNextOutput)
+
+
+
+import Heval (hevalsWith, hevals)
+import Utils (Randable(..),Logable(..))
 
 
 type Eva = StateT EvaState IO
 
 data EvaState = EvaState {
-     evaGen   :: StdGen ,
-     evaJobID :: Int 
+     evaGen    :: StdGen ,
+     evaJobID  :: Maybe Int ,
+     evaStdout :: String
   }
+
+runEva :: Eva a -> IO a
+runEva eva = do
+ gen <- getStdGen
+ runEvaWith gen eva 
+
+setJobID :: Int -> Eva ()
+setJobID jobID = do
+  evaState <- get
+  put $ evaState{ evaJobID = Just jobID }
+
+runEvaWith :: StdGen -> Eva a -> IO a
+runEvaWith gen eva = do
+ startTime <- getCurrentTime
+ putStrLn $ "\nstdGen: " ++ show gen
+ --((ret,gen'),stats) <- runStateT (runStateT eva gen) (Map.empty , Map.empty)  
+ ( ret , evaState' ) <- runStateT eva (initEvaState gen)
+ finishTime <- getCurrentTime
+ putStrLn $ "Total time: " ++ show (diffUTCTime finishTime startTime) ++ "\n"
+ return ret
+
+initEvaState :: StdGen -> EvaState
+initEvaState gen = EvaState { 
+  evaGen    = gen,
+  evaJobID  = Nothing,
+  evaStdout = "" 
+ }
 
 
 
 evalsWith :: (Typeable a) => String -> [String] -> a -> Eva [a]
 evalsWith file strs as = liftIO $ hevalsWith file strs as
+
+evals :: (Typeable a) => [String] -> a -> Eva [a]
+evals strs as = liftIO $ hevals strs as
+
+flushStdout :: Eva JSValue
+flushStdout = do
+  evaState <- get
+  case evaJobID evaState of
+    Nothing -> return jsEmptyObj
+    Just _  -> do
[…]

### 2013-05-09 v3/GP_Core.hs

@@ -1,0 +1,236 @@
+{-# LANGUAGE MultiParamTypeClasses, FlexibleInstances, TupleSections #-}
+
+module GP_Core where
+
+import Control.Monad ( liftM )
+import Data.Typeable ( Typeable )
+import Text.Printf   ( printf )
+import Data.Maybe    ( fromJust )
+
+
+import Eva   ( Eva, runEva, setJobID, evalsWith, evals, flushStdout, sendJSON )
+import Dist  ( Dist , mkDist, distTake_new, distGet, distMax, distMin, distAvg )
+import Utils ( logIt, boxIt , JShow, jshow, putList)
+import ServerInterface ( graphCmd, multiCmd ) 
+
+
+class Gene term opt where
+  generateIt :: Int -> opt -> Eva [term]
+
+class Muta term opt where
+ mutateIt :: opt -> term -> Eva term 
+
+class Cros term opt where
+ crossIt :: opt -> term -> term -> Eva (term,term) 
+
+class Evolvable term a gOpt mOpt cOpt where
+  evolveIt :: RunInfo -> Problem term a gOpt mOpt cOpt -> Eva (term,FitVal,Maybe Int)
+
+
+data Problem term a gOpt mOpt cOpt = Problem
+ { problemName :: String
+ , popSize     :: PopSize
+ , numGene     :: NumGene
+ , genOps      :: Dist (GenOp term)
+ , gOpt        :: gOpt
+ , mOpt        :: mOpt
+ , cOpt        :: cOpt
+ , fitFun      :: FitFun term a
+ }
+
+
+type PopSize = Int
+type NumGene = Int
+type FitVal  = Double
+type Prob    = Double
+type RunInfo = (Int,Int)
+
+data FitFun term a = 
+  FF1 (term -> Eva FitVal) a | 
+  FF2 (term->String) a (a->Eva FitVal) |
+  FF3 (term->String) a (a->Eva (FitVal,Bool) ) |
+  FF4 (term->String) String a | -- poslední a je fake jako u FF1, String je jmeno ff
+  FF5 String String a | -- jemeno FF , jmeno modulu , fake (showStr předpokladá show)
+  FF6 a (a->(FitVal,Bool)) String -- asType , ff , jmeno modulu 
+
+type GenOpProbs = (Prob,Prob,Prob)
[…]

### 2013-05-09 v3/GP_Data.hs

@@ -1,0 +1,519 @@
+{-# LANGUAGE MultiParamTypeClasses #-}
+{-# LANGUAGE FlexibleInstances     #-}
+{-# LANGUAGE TupleSections         #-}
+
+module GP_Data where
+
+import Eva (Eva,evaSplitStdGen)
+import GP_Core ( Gene, Muta, Cros, generateIt, mutateIt, crossIt , Prob )
+import Utils   ( getRandom, getRandomR, getRandomL, getNormal, randIf, randCase,logIt,boxIt,boxThem )
+
+import TTerm
+import TTree
+
+import IM ( SearchOptions , prove , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
+import System.Random (StdGen)
+
+--import InhabitationMachines ( proveN , randProveN, randProveUnique, proveOneWithLimit, Limit, kozaProveN )
+
+import KozaTree ( KTree(KNode), KPos, kSubtree, kChangeSubtree, kPoses, kPoses2, kDepth ) 
+import Dist     ( Dist, mkDist, distSize, distToList )
+
+
+import Control.Monad (replicateM, liftM, liftM2, forM)
+
+import Data.List     (nub,groupBy)
+
+type Len    = Int
+type Mean   = Double
+type StdDev = Double
+
+-- CTT -------------------------------------------------------------------------
+
+instance Gene CTT CTTGen where generateIt = cttGen
+instance Cros CTT CTTCro where crossIt    = cttCro2
+
+data CTTGen  
+-- = CTTG_Koza     Typ Context Limit  -- pomocí starýho IM 
+ = CTTG_Koza2    Typ Context        -- a zbytek pomocí novýho IM ..
+ | CTTG_AllEdges Typ Context 
+ | CTTG_Geom     Typ Context Prob 
+
+data CTTCro = CTTC_Koza
+
+cttGen :: Int -> CTTGen -> Eva [CTT]
+cttGen n opt = case opt of
+  CTTG_AllEdges typ ctx   -> proveIt  allEdgesSearchOptions n typ ctx 
+  CTTG_Koza2    typ ctx   -> proveIt  kozaSearchOptions     n typ ctx 
+  CTTG_Geom     typ ctx p -> proveIt (geomSearchOptions p)  n typ ctx
+--  CTTG_Koza typ ctx limit ->  map mkCTT `liftM` kozaProveN n limit typ ctx
+ where
[…]

### 2013-05-09 v3/IM.hs

@@ -1,0 +1,753 @@
+--  Nový pokus o Inhabitation Machines bez potřeby generovat graf dopředu a vůbec to udělat pořádně
+
+--  Asi by to chtělo celý předelat tak, že ZTree bude stavová monada, ale radši 
+-- to nejdřív už dodělat když je to takle rozdělaný a až pak dyštak překopat.
+
+
+
+
+module IM 
+ ( SearchOptions(..) 
+ , defaultSearchOptions 
+ , kozaSearchOptions
+ , allEdgesSearchOptions
+ , geomSearchOptions 
+ , prove 
+ ) where
+
+import TTerm (Symbol,Typ(..),TTerm(..),Context,typeArgs,ttermTyp,toSki,toSki',checkTyp)
+
+import TTree (CTT,mkCTT2)
+
+import Data.List
+import Data.Maybe
+import qualified Data.Set as Set
+import qualified Data.Map as Map
+import Data.Set (Set)
+import Data.Map (Map)
+
+import System.Random
+
+import qualified Data.PSQueue as Q
+
+
+--proveAll :: Int -> Typ -> Context -> [CTT]
+
+prove :: SearchOptions -> [CTT]
+prove so = 
+  let ( so' , problemHead ) = problemHeadPreproccess so
+      trees = proveWith2 so'
+   in map ( mkCTT2 problemHead . toSki' . tree2tterm ) trees   -- <============== toSki s ' je s typeCheckem ...........
+
+problemHeadPreproccess :: SearchOptions -> ( SearchOptions , Context )
+problemHeadPreproccess so = 
+  let ( typ' , ctx' , problemHead ) = problemHeadPreproccess' (so_typ so) (so_ctx so)
+   in ( so{ so_typ = typ' , so_ctx = ctx' } , problemHead )
+
+problemHeadPreproccess' :: Typ -> Context -> (Typ,Context,Context)
+problemHeadPreproccess' typ ctx = 
+  let (ts,alpha)  = typeArgs typ
+      ss          = map (\i->'x':show i) [0..]
+      problemHead = zip ss ts
+   in (Typ alpha , problemHead ++ ctx , problemHead )
+
+
+
+
+
+
+data ZTree = ZTree { 
+
+  current     :: Tree , 
+  dads        :: [ DTree ]  ,
+
+  searchOpts  :: SearchOptions,
+  rand        :: StdGen,
+
+  locals      :: Table ,
+  globals     :: Table ,
+
+  depth       :: Int ,
+  nextVar     :: Int ,
[…]

### 2013-05-09 v3/Job.hs

@@ -1,0 +1,151 @@
+module Job ( problemList_ , job_ ) where
+
+import GP_Core ( nRunsByServer )
+import Text.JSON
+import JSONUtils
+import Problems.Utils ( ProblemOpts , po2json, poCode, json2po , runProblemOpts )
+import Data.Typeable ( Typeable )
+import Data.Map (Map)
+import qualified Data.Map as Map
+
+
+
+import qualified Problems.SSR.Problem        as SSR
+import qualified Problems.BigCtx.Problem     as BigCtx
+import qualified Problems.Fly.Problem        as Fly
+import qualified Problems.Ant.Problem        as Ant
+import qualified Problems.BA.Problem         as BA
+import qualified Problems.EvenParity.Problem as EvenParity
+
+
+regs = 
+ [ reg EvenParity.reg
+ , reg BigCtx.reg
+ , reg Fly.reg 
+ , reg SSR.reg 
+ , reg Ant.reg
+ , reg (BA.reg :: POU) 
+ ]
+
+
+
+
+
+
+
+
+
+
+type JobID = String
+type Code  = String
+type POU   = ProblemOpts ()
+
+
+
+reg :: Typeable a => ProblemOpts a -> ((String, JSValue -> JobID -> IO ()), JSValue)
+reg po = ( (poCode po , \ jsv jobID -> runProblemOpts (json2po po jsv) jobID ) , po2json po )
+
+jobs :: ( [ (Code , JSValue -> JobID -> IO()) ] , [JSValue] )
+jobs = unzip regs
+
+problemList_ :: JSValue
+problemList_ = jsArr $ snd jobs
+
+problemTab :: Map Code ( JSValue -> JobID -> IO () )
+problemTab  = Map.fromList $ fst jobs 
+
+job_ :: JobID -> String -> IO ()
+job_ jobID cmd = case ( decode cmd :: Result JSValue ) of
+  Error str  -> putStrLn $ "ERROR in job_ : " ++ str
+  Ok jsvalue -> do
+    let code = fromJsStr $ jsProp jsvalue "code" 
+        Just runFun = Map.lookup code problemTab 
+    putStrLn code
+    runFun jsvalue jobID 
+
+
+
+
+
+
+
+-- ( problemMap , problemList__ ) = registerThem 
+--  [  SSR.reg 
+--  -- ,  Fly.reg 
+--  ,   BA.reg ]
+-- 
[…]

### 2013-05-09 v3/Problems/Ant/Problem.hs

@@ -1,0 +1,49 @@
+module Problems.Ant.Problem where
+
+import Text.JSON
+import JSONUtils
+import DrawIM( imGraphInJSON )
+
+import TTerm (Typ(..),Context)
+import Problems.Utils 
+import GP_Core (FitFun(FF5))
+import GP_Data (CTTGen(..))
+
+
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "ant"                                        ,
+  cttp_info        = "Artifical Ant"                              ,
+  cttp_data        = jsData                                       ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = ant                                          ,
+  cttp_ctx         = ctx                                          ,
+  
+  cttp_gOpt        = CTTG_Koza2 ant ctx                           , 
+
+  cttp_ff          = FF5 "ff" "Problems.Ant.Funs" ()
+  
+}
+
+mainProblem = cttProblem3 "ant" "Problems.Ant.Funs" ant ctx
+
+jsData = jsObj [ ( "im" , imGraphInJSON ant ctx ) ]
+
+ctx :: Context
+ctx = [ ( "l"   , ant  ) , 
+        ( "r"   , ant  ) , 
+        ( "m"   , ant  ) , 
+        ( "ifa" , ant2 ) , 
+        ( "p2"  , ant2 ) , 
+        ( "p3"  , ant3 ) ]
+
+
+ant, ant2, ant3 :: Typ
+ant  = Typ "Ant"
+ant2 = ant :-> ant :-> ant
+ant3 = ant :-> ant :-> ant :-> ant
+
+
+

### 2013-05-09 v3/Problems/BA/Problem.hs

@@ -1,0 +1,35 @@
+module Problems.BA.Problem where
+
+--import Problems.Utils ( boolListProblem2 )
+import Problems.Utils 
+import Text.JSON
+
+
+reg :: ProblemOpts a
+reg = PO_BLP_ PO_BLP {
+  blp_code        = "ba"                                         ,
+  blp_info        = "Bool Alternate"                             ,
+  blp_data        = JSNull                                       ,
+  blp_numRuns     = IntSlider "Runs"            1 10    4    1   ,
+  blp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  blp_popSize     = IntSlider "Population size" 0 1000  500 100  , -- dřív 5000 ale začlo to s předěláním Evy dělat stackoverflow :(
+  blp_length      = IntSlider "Genom length"    0 1000  200  10  ,
+  blp_ff          = ff
+}
+
+
+
+--mainProblem_ = boolListProblem2 "ba" (33,33,33) 100 ff
+
+
+
+
+
+ff :: [Bool] -> Double
+ff bits = 
+ let dLen = fromIntegral . length
+     tNum = dLen . filter (\(n,b)->if n `mod` 2 == 0 then b else not b) $ zip [0..] bits 
+     len  = dLen $ bits
+  in 100*(tNum * tNum)/(len*len)
+
+

### 2013-05-09 v3/Problems/BigCtx/Problem.hs

@@ -1,0 +1,193 @@
+module Problems.BigCtx.Problem where
+
+import TTerm (Typ(..),Context)
+import Problems.Utils 
+import JSONUtils
+import GP_Core (FitFun(FF6))
+import GP_Data (CTTGen(..))
+
+import DrawIM( imGraphInJSON )
+
+ 
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "bigctx"                                     ,
+  cttp_info        = "Problem with basic funs for types generating elementary funs." ,
+  cttp_data        = jsObj [ ( "im" , imGraphInJSON prog_typ prog_ctx ) ]      ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = prog_typ                                     ,
+  cttp_ctx         = prog_ctx                                     ,
+  
+  cttp_gOpt        = CTTG_Koza2 prog_typ prog_ctx                 , 
+
+  cttp_ff          = FF6 prog_type ff_fst3 "Problems.BigCtx.Funs" 
+  
+}
+
+
+prog_type = asType :: Fst3_Type  -- asType :: [Int] -> Maybe Int
+prog_typ  = fst3_typ             -- l_int :-> m_int  
+prog_ctx  = ctx_fst3             -- ctx_combo
+
+--problem1 = go pr_head
+
+go (ff,ctx,typ,as) = cttProblem5 "BigCtx" ff typ ctx as
+
+
+bool   = Typ "Bool"
+int    = Typ "Int"
+m_int  = Typ "Maybe Int"
+l_int  = Typ "[Int]"
+ml_int = Typ "Maybe [Int]"
+
+
+
+
+--(ff,ctx,typ,as) = pr_head
+
+
+pr_head   = go ( ff_head   , ctx_combo , l_int :-> m_int                 , asType :: [Int] -> Maybe Int           )
+pr_tail   = go ( ff_tail   , ctx_combo , l_int :-> ml_int                , asType :: [Int] -> Maybe [Int]         )
+pr_map    = go ( ff_map    , ctx_combo , (int:->int) :-> l_int :-> l_int , asType :: (Int->Int) -> [Int] -> [Int] )
[…]

### 2013-05-09 v3/Problems/EvenParity/Problem.hs

@@ -1,0 +1,167 @@
+module Problems.EvenParity.Problem where
+
+import TTerm (Typ(..),Context)
+import Problems.Utils 
+import JSONUtils
+import GP_Core (FitFun(FF6))
+import GP_Data (CTTGen(..))
+import DrawIM( imGraphInJSON )
+
+import Problems.EvenParity.Funs
+
+
+prog_type = asType :: [Bool]  -> Bool  
+prog_typ  =           l_bool :-> bool             
+prog_ctx  = ctx_yu             
+
+
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "ep"                                     ,
+  cttp_info        = "Even-parity problem." ,
+  cttp_data        = jsObj [ ( "im" , imGraphInJSON prog_typ prog_ctx ) ] ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 500   50   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = prog_typ                                     ,
+  cttp_ctx         = prog_ctx                                     ,
+  
+  cttp_gOpt        = CTTG_Geom     prog_typ prog_ctx 0.75 , 
+--cttp_gOpt        = CTTG_Koza2    prog_typ prog_ctx  , 
+--cttp_gOpt        = CTTG_AllEdges prog_typ prog_ctx  , 
+
+  cttp_ff          = FF6 prog_type ff "Problems.EvenParity.Funs" 
+  
+}
+
+
+
+bool   = Typ "Bool"
+l_bool = Typ "[Bool]"
+
+ctx = [
+ ( "foldr" , (bool:->bool:->bool) :-> bool :-> l_bool :-> bool ),
+ ( "(==)"  , bool :-> bool :-> bool                            ),
+ ( "(/=)"  , bool :-> bool :-> bool                            ),
+ ( "(&&)"  , bool :-> bool :-> bool                            ),
+ ( "(||)"  , bool :-> bool :-> bool                            ),
[…]

### 2013-05-09 v3/Problems/Fly/Problem.hs

@@ -1,0 +1,302 @@
+module Problems.Fly.Problem where
+
+
+import Data.List
+
+
+import TTerm          (Typ(..),Context)
+import Problems.Utils 
+ --( cttProblem5 , asType )
+
+import Text.JSON
+import JSONUtils
+
+import DrawIM( imGraphInJSON )
+
+import Problems.Fly.Funs ( Input_ , Output_ , Dir_  , Energy , Pos , avg , 
+                           output_ , myApplePoses_ ,nearestFlyPos_, myPos_ , posToDir_  ,
+                           dStay , dUp , dDown , dLeft , dRight )
+import Problems.Fly.Fly
+
+import GP_Core (FitFun(FF6))
+import GP_Data (CTTGen(..))
+
+
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "fly"                                        ,
+  cttp_info        = "Fly eating apples and stuff... "            ,
+  cttp_data        = jsData                                       ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = prog_typ                                     ,
+  cttp_ctx         = ctx                                          ,
+
+  cttp_gOpt        = CTTG_Koza2 prog_typ ctx                      , 
+  
+  cttp_ff          = FF6 prog_type ff "Problems.Fly.Funs" 
+  
+}
+
+problem1 = cttProblem5 "Fly" ff prog_typ ctx prog_type
+
+
+
+
+jsData = jsObj [ 
+ ( "world" , worldToJSON ff_solutionFlyPos ff_world_withEnvirFlies ),
+ ( "im"    , imGraphInJSON prog_typ ctx )]
+
+
+prog_type = asType :: Input_ -> Output_
+prog_typ  = input_typ :-> output_typ
+
+input_typ  = Typ "Input_"
+output_typ = Typ "Output_"
+dir_typ    = Typ "Dir_"
+
+m_pos   = Typ "Maybe Pos"
+l_pos   = Typ "[Pos]"
+pos_typ = Typ "Pos"
+
+dou     = Typ "Double"
[…]

### 2013-05-09 v3/Problems/SSR/Funs.hs

@@ -1,0 +1,20 @@
+module Problems.SSR.Funs where
+
+
+rdiv :: Double -> Double -> Double
+rdiv _ 0 = 1
+rdiv x y = x / y
+
+rlog :: Double -> Double
+rlog 0 = 0
+rlog x = log $ abs x
+
+ff :: (Double->Double) -> (Double,Bool)
+ff prog = ( (1/) . (1+) . sum $ diffs , 20 == (length . filter (\dif-> (abs dif) < 0.01 ) $ diffs)  )  
+ where 
+  diffs = map (\x-> let dx = (prog x) - ( x*x*x*x+x*x*x+x*x+x ) in abs dx ) (linInterval (-1,1) 20)
+  linInterval :: (Double,Double) -> Int -> [Double]
+  linInterval (a,b) n = 
+   let step = (b-a) / (fromIntegral $ n-1)
+    in [ a + (fromIntegral i)*step | i <- [0..(n-1)] ]
+ 

### 2013-05-09 v3/Problems/SSR/Problem.hs

@@ -1,0 +1,60 @@
+module Problems.SSR.Problem where
+
+import Text.JSON
+import JSONUtils
+import DrawIM( imGraphInJSON )
+
+import TTerm (Typ(..),Context)
+--import Problems.Utils ( cttProblem5 , cttProblem3 , cttProblem' , cttProblem2' , cttProblem4 , asType )
+import Problems.SSR.Funs (ff)
+
+import Problems.Utils 
+import GP_Core (FitFun(FF6))
+import GP_Data (CTTGen(..))
+
+
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "ssr"                                        ,
+  cttp_info        = "Simple Symbolic Regression - new IM !"      ,
+  cttp_data        = jsData                                       ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = dou1                                         ,
+  cttp_ctx         = ctx                                          ,
+
+  cttp_gOpt        = CTTG_Koza2 dou1 ctx                          , 
+  
+  cttp_ff          = FF6 (asType::Double->Double) ff "Problems.SSR.Funs" 
+  
+}
+
+
+mainProblem = cttProblem5 "SSR" ff dou1 ctx (asType::Double->Double)
+
+jsData = jsObj [ ( "im" , imGraphInJSON dou1 ctx ) ]
+
+
+dou, dou1, dou2 :: Typ
+dou  = Typ "Double"
+dou1 = dou :-> dou
+dou2 = dou :-> dou :-> dou
+
+ctx :: Context
+ctx = [( "(+)"  , dou2 ),
+       ( "(-)"  , dou2 ),
+       ( "(*)"  , dou2 ),
+       ( "rdiv" , dou2 ),
+       ( "sin"  , dou1 ),
+       ( "cos"  , dou1 ),
+       ( "exp"  , dou1 ),
+       ( "rlog" , dou1 )]
+
+
+
+
+-- problem1 = cttProblem'  "ssr"  ff (asType::Double->Double) dou1 ctx
+-- problem2 = cttProblem2' "ssr2" ff (asType::Double->Double) dou1 ctx
[…]

### 2013-05-09 v3/Problems/Utils.hs

@@ -1,0 +1,234 @@
+module Problems.Utils (
+  cttProblem5,
+  cttProblem4,
+  cttProblem3,
+  cttProblem2,
+  cttProblem2',
+  boolListProblem2,
+  asType,
+  casesFF,
+  ProblemOpts(..),PO_BLP(..),PO_CTTP(..),IntSlider(..),JobID,runProblemOpts,po2json,poCode,json2po
+) where
+
+import GP_Core (FitFun(..),Problem(..),NumGene,PopSize,mkGenOps)
+import GP_Core (GenOpProbs,FitVal,mkFF1)
+import GP_Core ( nRunsByServer )
+import GP_Data (CTTGen(..),CTTCro(..))
+import GP_Data (ListGen(..),ListCro(..),ListMut(..), BoolGen(..), BoolMut(..), Len)
+
+
+import TTree (CTT)
+import TTerm (Typ,Context)
+
+import Text.JSON
+import JSONUtils
+
+import Data.Typeable ( Typeable )
+
+-- ---------------------------------------------------------------
+
+
+
+data ProblemOpts a = PO_BLP_ PO_BLP | PO_CTTP_ (PO_CTTP a) 
+
+
+poCode :: ProblemOpts a -> String
+poCode po = case po of
+  PO_BLP_  x ->  blp_code x
+  PO_CTTP_ x -> cttp_code x
+
+po2json :: ProblemOpts a -> JSValue
+po2json po = case po of
+  PO_BLP_  x ->  blp2json x
+  PO_CTTP_ x -> cttp2json x
+
+json2po :: ProblemOpts a -> JSValue -> ProblemOpts a
+json2po po json = case po of
+  PO_BLP_  x -> PO_BLP_  $ json2blp  x json
+  PO_CTTP_ x -> PO_CTTP_ $ json2cttp x json
+
+
+
+data PO_BLP = PO_BLP {
+  
+  blp_numRuns     :: IntSlider        , 
+  blp_numGene     :: IntSlider        , 
+  blp_popSize     :: IntSlider        , 
+  blp_length      :: IntSlider        ,
+
+  blp_code        :: String           ,
+  blp_info        :: String           ,
+  blp_data        :: JSValue          ,
+  blp_ff          :: [Bool] -> Double }
+
+
+blp2json :: PO_BLP -> JSValue
+blp2json blp = jsObj [
+  ( "code"    , jsStr          $ blp_code    blp                                  ),
+  ( "info"    , jsStr          $ blp_info    blp                                  ),
[…]

### 2013-05-09 v3/Server.hs

@@ -1,0 +1,247 @@
+{-# LANGUAGE OverloadedStrings #-}
+ 
+import Network.Wai
+import Network.Wai.Handler.Warp
+import Network.HTTP.Types (status200)
+import Blaze.ByteString.Builder (copyByteString)
+import qualified Data.ByteString.UTF8 as BU
+import Data.Monoid
+import Data.Enumerator (run_, enumList, ($$))
+
+import Text.JSON ( encode )
+ 
+-- já
+import Control.Concurrent  
+import Control.Monad
+import Control.Monad.State
+import System.IO
+import System.Directory
+import System.Environment
+
+import Utils ( unescape )
+--import Job (job,problemList,problemList_,job_)
+import Job (problemList_,job_)
+
+import ServerInterface
+
+main = do
+    clearServer
+    (port:_) <- getArgs
+    putStrLn $ "Fishtron GUI is on http://localhost:" ++ port 
+    run (read port) app
+
+app :: Application
+app req = 
+ case pathInfo req of
+
+    ["out",i,j] -> do
+      let workerId = ( read . init . tail . show $ i ) :: Int
+      let outputId = ( read . init . tail . show $ j ) :: Int
+      outString <- liftIO $ serveOutput workerId outputId
+      return $ myTextPlain outString
+
+    [ ] -> return $ myIndex
+
+--    ["run",cmd] -> do 
+--      workerId <- liftIO newWorker
+--      let cmdStr = init . tail . show $ cmd 
+--      liftIO . forkIO $ runCmd workerId cmdStr
+--      return . myTextPlain . show $ workerId
+
+    ["run_",cmd] -> do 
+      workerId <- liftIO newWorker
+      let cmdStr = init . tail . show $ cmd 
+      liftIO . forkIO $ runCmd_ workerId (unescape cmdStr)
+      return . myTextPlain . show $ workerId
+
+--    ["problems"] -> return . myTextPlain . encode . problemListToJSON $ problemList
+
+    ["problems_"]-> return . myTextPlain . encode $ problemList_    
+
+    ["js",filename] -> do
+      return $ myJSFile (init . tail . show $ filename)
+
+    ["js","libs",filename] -> do
+      return $ myJSFile $ "libs/" ++ (init . tail . show $ filename)
+    
+    ["js","Problems",filename] -> do
[…]

### 2013-05-09 v3/ServerInterface.hs

@@ -1,0 +1,83 @@
+module ServerInterface where
+
+import System.IO
+import System.Directory
+
+import Text.JSON
+import JSONUtils
+
+writeNextOutput :: Int -> String -> IO ()
+writeNextOutput i output = do
+  outI <- incrementFile ("server/output/" ++ (show i) ++ "/_hotovo.txt")
+  myWriteFile ("server/output/"++ (show i) ++ "/" ++ (show outI )++ ".txt" ) output
+
+
+incrementFile :: FilePath -> IO Int
+incrementFile filename = do 
+  file <- openFile filename ReadWriteMode
+  str <- hGetLine file 
+  let num  = 1 + ( read str ) :: Int
+      str' = show num
+  hSeek file AbsoluteSeek 0
+  hPutStr file str'
+  hClose file 
+  return num
+
+
+myWriteFile :: FilePath -> String -> IO ()
+myWriteFile filename str = do 
+  file <- openFile filename WriteMode
+  hPutStr file str
+  hClose file
+
+myReadFile :: FilePath -> IO String
+myReadFile filename = do
+ file <- openFile filename ReadMode 
+ str <- hGetLine file 
+ hClose file
+ return str
+
+
+problemListToJSON :: [(String,String,JSValue)] -> JSValue
+problemListToJSON problemList =  jsArr . map toProblemObj $ problemList
+ where
+  toProblemObj :: (String,String,JSValue) -> JSValue
+  toProblemObj ( code , name , datas ) = jsObj [
+    ( "code" , jsStr code  ) ,
+    ( "name" , jsStr name  ) ,
+    ( "data" , datas       )
+   ]
+
+
+jsEmptyObj :: JSValue
+jsEmptyObj = jsObj []
+
+stdoutCmd :: String -> JSValue
+stdoutCmd str = jsObj [ 
+  ("type" , jsStr "stdout" ) ,
+  ("msg"  , jsStr str      ) ]
+
+graphCmd :: Int -> Int -> (Double,Double,Double) -> JSValue
+graphCmd runI genI (best,avg,worst) = jsObj [ 
+  ("type"   , jsStr "generationInfo" ) ,
+  ("i"      , jsNum genI ) ,
+  ("run"    , jsNum runI ) ,
+  ("ffvals" , jsObj [
+      ( "best"  , jsNum best  ) ,
+      ( "avg"   , jsNum avg   ) ,
[…]

### 2013-05-09 v3/TTerm.hs

@@ -1,0 +1,406 @@
+module TTerm
+( Typ (..)
+, TTerm (..)
+, TTermZipper (..)
+, TTermDrobek (..)
+, Context
+, Symbol
+, typeArgs
+, ttermTyp , isInCtx
+, mkTz, showTTZ, tzDown, tzLeft, tzRight, tzGoTop,ttzTyp
+, subs
+, makeVarsUnique 
+, checkTyp
+, toSki, toSki'
+) where
+
+import Data.List ( nub, (\\) )
+import Data.Typeable ( Typeable )
+import Text.ParserCombinators.Parsec ( Parser, ParseError, parse, (<|>), eof, char, many1, noneOf, skipMany1, space )
+
+import Utils ( newSymbol )
+
+--import Util  ( newSymbol )
+--import Heval ( eval )
+
+---------------------------------------------------
+
+type Symbol = String
+
+data Typ  = 
+ Typ Symbol | 
+ Typ :-> Typ
+ deriving (Eq,Ord)
+infixr 7 :->
+
+type Context = [(Symbol,Typ)]
+
+data TTerm = 
+ TVar Symbol        Typ | 
+ TVal Symbol        Typ | 
+ TLam Symbol TTerm  Typ |
+ TApp TTerm  TTerm  Typ
+ deriving (Ord,Eq)
+
+-- substitution ------------------------------------
+
+subs :: Symbol -> TTerm -> TTerm -> TTerm 
+subs x m n = fst $ sub (varNames m) m x n   
+
+sub :: [Symbol] -> TTerm -> Symbol -> TTerm -> (TTerm,[Symbol])
+sub vns x@(TVal _ _) _ _ = (x,vns)  
+sub vns x@(TVar v _) var vloz 
+    | v == var  = (vloz,vns)
+    | otherwise = (x,vns)  
+sub vns (TApp p q typ) var vloz = (TApp t1 t2 typ , vns'')
+    where 
+    (t1,vns')  = sub vns  p var vloz
+    (t2,vns'') = sub vns' q var vloz
+sub vns x@(TLam v p typ) var vloz
+    | v == var  = (x,vns)
+    | not (elem v (fv vloz)) = (TLam v t typ , vns'') 
+    | otherwise = sub vns' term' var vloz
+    where
+    (t,vns'') = (sub vns p var vloz)
+    (term',vns') = rename vns x
+
+varNames :: TTerm -> [Symbol]
+varNames = nub . varNames'
+    where
[…]

### 2013-05-09 v3/TTree.hs

@@ -1,0 +1,400 @@
+{-# LANGUAGE OverlappingInstances #-}
+
+module TTree 
+( CTT (..),
+  TTree (..),
+  TTPos,
+  mkCTT, -- <============== možná bug v něm!!!! 
+  mkCTT2,
+  ttreeSubtree,
+  ttreeSubtrees,
+  ttreeChangeSubtree,
+  ttreeDepth,
+  ttreePoses2,
+
+  ttreePoses2WithTyps,
+  ttreePoses2ByTyp
+) where
+
+import Data.List
+import Data.Either
+import Data.Maybe
+
+import Text.JSON (JSValue (..) , toJSObject , toJSString )
+import Utils ( JShow , jshow_js )
+
+import JSONUtils
+
+import TTerm
+
+data CTT = CTT Context TTree 
+
+data TTree = TTree Symbol Typ [TTree] 
+
+type TTPos = [Int]
+
+
+mkCTT2 :: Context -> TTerm -> CTT
+mkCTT2 ctx tt = CTT ctx (fromTTerm tt)
+
+
+instance Show CTT where
+ show (CTT ctx ttree) = case ctx of
+  [] -> show ttree
+  _  -> let vars = intercalate " " . map fst $ ctx
+         in "\\ "++ vars ++ " -> " ++ show ttree 
+
+instance JShow CTT where
+  jshow_js ctt = Just . jsStr . jsShow $ ctt
+--  jshow ctt = JSObject $ toJSObject [ 
+--   ("type"     , JSString . toJSString $ "jsonout"     ) ,
+--   ("haskell"  , JSString . toJSString $ show ctt      ) ,
+--   ("js"       , JSString . toJSString $ jsShow ctt    ) 
+--   --("kutil"    , JSString . toJSString $ ctt2kutil ctt ) 
+--   ]
+
+
+jsShow :: CTT -> String
+jsShow (CTT ctx ttree) =
+ let vars = intercalate "," . map fst $ ctx
+  in "function("++ vars ++ "){return " ++ jsShowBody ttree ++ ";}"
+
+jsShowBody :: TTree -> String
+jsShowBody (TTree symbol typ ttrees ) =
+ let symbol' = transformExceptions symbol 
+ in case ttrees of
+     [] -> symbol'
+     _  ->
+      if isBinop symbol' && length ttrees == 2
+       then let [l,r] = ttrees
+                op    = tail . init $ symbol'
[…]

### 2013-05-09 v3/Utils.hs

@@ -1,0 +1,240 @@
+{-# LANGUAGE FlexibleInstances, UndecidableInstances, OverlappingInstances #-}
+
+module Utils where
+
+import Data.Maybe
+import Data.List
+import Data.Map (Map)
+import qualified Data.Map as Map
+
+import System.Random
+import Control.Monad.State
+import Data.Random.Normal
+
+
+import Text.JSON (JSValue(..))
+
+import JSONUtils
+
+-- general ------------------------------------
+
+putList :: (Show a) => [a] -> IO ()
+putList []     = return ()
+putList (x:xs) = do
+ putStrLn . show $ x
+ putList xs 
+
+asType :: a
+asType = undefined
+
+unescape :: String -> String
+unescape []     = []
+unescape (x:[]) = [x]
+unescape (x:y:rest)
+    | x == '\\' = y : unescape rest
+    | otherwise = x : unescape (y : rest)
+
+-- Pro již použitá slova vrací první takové, které je od všech rozdílné
+newSymbol :: [String] -> String
+newSymbol = newSymbol' ['a'..'z'] 
+
+newSymbol' :: [Char] -> [String] -> String
+newSymbol' abeceda vns = snd . head $ dropWhile (\(x,y)->x==y) 
+   (zip ((sortBy (nameOrderingBy abeceda) vns') ++ repeat "") (map (nthWord abeceda) [1..]))
+ where
+  vns' = filter (all (`elem` abeceda )) vns 
+
+-- Uspořádání primárne podle délky, sekundárne podle abecedy.
+nameOrderingBy :: (Eq a) => [a] -> [a] -> [a] -> Ordering
+nameOrderingBy abc x y
+  | lo /= EQ  = lo
+  | otherwise = compareByAbc abc x y
+ where lo = lenOrdering x y
+
+compareByAbc :: (Eq a) => [a] -> [a] -> [a] -> Ordering
+compareByAbc _ [] [] = EQ
+compareByAbc abc (x:xs) (y:ys) = case cByAbc abc x y of
+  EQ -> compareByAbc abc xs ys
+  x  -> x
+ where 
+  cByAbc :: (Eq a) => [a] -> a -> a -> Ordering
+  cByAbc abc x y
+   | x == y    = EQ
+   | otherwise = case dropWhile (\ch-> (ch/=x) && (ch/=y) ) abc  of
+      []  -> error "cByAbc : x and y are not in the alphabet"
[…]

### 2013-05-09 v3/server/js/Problems/Ant.js

@@ -1,0 +1,324 @@
+
+$(document).ready(function() {
+
+console.log( "Ant.js ready!" );  
+
+var antState; 
+
+
+Global.Problems.ant = {
+
+
+  onLoadedProblemData : function( data ){ 
+    this.antState.loadResources();
+  },
+
+
+  onGo : function(){
+    
+    $('#special').html( 
+      '<b>Each number</b> (=fitness) <b>link represents one best-of-generation individual. Click it and see!</b>'+
+      '<div id="ant-links"></div><br>' +
+      '<canvas id="antCanvas" width="517" height="517"></canvas>' );
+
+    this.solutions     = [] ;
+    this.i             = 0  ;
+
+    this.antState.init();
+
+  },
+
+
+  fenotyp : function( solution , ffVal ){
+
+    this.solutions[this.i] = solution ;
+    $('#ant-links').append( '<a href="#" title="'+ solution +'" '+
+      'onclick="Global.Problems.ant.link('+this.i+')">' +ffVal + '</a> ' );
+
+    this.i ++ ;
+
+  },
+
+
+
+
+  link : function( i ){
+
+    this.antState.add( this.solutions[i]() );
+
+  },
+
+  antState : {
+    pos : undefined,
+    dir : undefined ,
+    world : [],
+    limit : 600,
+    steps : undefined,
+    eaten : undefined,
+    ctx : undefined,
+    antImg : {},
+    seedImg : undefined,
+    isRunning : false,
+    progQueue : [],
+    actAnt : 0,
+    initWorldStrs :   
+     [ ".FFF............................"
+     , "...F............................"
+     , "...F.....................FFF...." 
+     , "...F....................F....F.." 
+     , "...F....................F....F.." 
+     , "...FFFF.FFFFF........FF........." 
+     , "............F................F.." 
+     , "............F.......F..........." 
+     , "............F.......F..........." 
+     , "............F.......F........F.." 
+     , "....................F..........." 
+     , "............F..................." 
+     , "............F................F.." 
+     , "............F.......F..........." 
+     , "............F.......F.....FFF..." 
+     , ".................F.....F........" 
+     , "................................" 
+     , "............F..................."             
+     , "............F...F.......F......."
+     , "............F...F..........F...." 
[…]

### 2013-05-09 v3/server/js/Problems/Fly.js

@@ -1,0 +1,536 @@
+
+$(document).ready(function() {
+
+console.log( "Fly.js ready!" );  
+
+
+
+
+
+Global.Problems.fly = {
+
+  progLib : {
+    prog1 : function (x0){return output_(posToDir_(  myPos_(x0)   , head_(myApplePoses_(x0))  )   );},
+    prog2 : function (x0){return output_(dRight);},
+    prog3 : function (x0){return output_(posToDir_(  myPos_(x0)   , nearestFlyPos_(x0) )  );},
+  },
+
+//prog_1 input = output_ $ posToDir_ (myPos_ input) (head_ $ myApplePoses_ input)
+//prog_2 _     = output_ $ dRight 
+//prog_3 input = output_ $ posToDir_ (myPos_ input) (nearestFlyPos_ input)
+
+
+
+  onLoadedProblemData : function( data ){
+    //console.log( data );
+    this.loadResources();
+    this.dataJSON = JSON.stringify( data.world );
+    this.loadFromDataJSON( );
+  },
+
+  restart : function( solution , drawIt ){
+
+    if( solution === undefined ){
+      solution = this.defaultSolution ;
+      this.actualSolution = solution  ;
+    }
+    if( drawIt   === undefined ){drawIt   = true;}
+
+    this.loadFromDataJSON( );
+    this.mapaAt( this.solutionFlyPos ).prog = solution ;
+    if(drawIt){ this.drawMapa(); }
+  },
+
+  loadFromDataJSON : function( ){
+
+    var data = JSON.parse( this.dataJSON );
+
+    this.mapa = data.mapa ;
+    
+    this.mapa[data.solutionFlyPos[0]][data.solutionFlyPos[1]] = {  
+      type       : 'fly' ,
+      progName   : '_'   ,
+      energy     : 1  
+    };
+
+    this.solutionFlyPos = data.solutionFlyPos ;
+
+    this.fliesToDo      = data.fliesToDo;
+    this.doneFlies      = data.doneFlies;
+    this.applePoses     = data.applePoses;
+    
+    this.numSteps       = data.numSteps;
+
+    this.fliesToDo.unshift( data.solutionFlyPos );
+
+    for( var y in this.mapa ){
+      for( var x in this.mapa[y] ){
+        var obj = this.mapa[x][y] ;
+        if( obj.type === 'fly' ){
+          obj.prog = this.progLib[ obj.progName ] ;
+        }
+      }
+    }
[…]

### 2013-05-09 v3/server/js/Problems/SSR.js

@@ -1,0 +1,69 @@
+
+$(document).ready(function() {
+console.log( "SSR.js ready!" );  
+Global.Problems.ssr = {
+
+
+
+  onLoadedProblemData : function( data ){ },
+
+
+  onGo : function(){
+    
+    $('#special').html( 
+      '<b>Each number</b> (=fitness) <b>link represents one best-of-generation individual. Click it and see!</b>'+
+      '<div id="ssr-links"></div><br>' +
+      '<div id="graphssr-div" class="graph"></div>' );
+
+    this.solutions     = [] ;
+    this.i             = 0  ;
+
+  },
+
+
+  fenotyp : function( solution , ffVal ){
+
+    ffVal = Math.round(ffVal*100)/100 ;
+
+    this.plotIt(solution);
+
+    this.solutions[this.i] = solution ;
+    $('#ssr-links').append( '<a href="#" title="'+ solution +'" '+
+      'onclick="Global.Problems.ssr.ssrLink('+this.i+')">' +ffVal + '</a> ' );
+
+    this.i ++ ;
+
+  },
+
+  plotIt : function( solution ){
+
+    var trueSolution = function(x){return x*x*x*x+x*x*x+x*x+x ;}
+
+    var xs = [ ] ; 
+    for (var i = -1 ; i <= 1 ; i+= 0.01 ) {
+      xs.push(i);
+    }
+
+    ysSol     = _.map( xs , function(x){return [x,    solution(x)];} );
+    ysTrueSol = _.map( xs , function(x){return [x,trueSolution(x)];} );
+
+    
+    var ssrPlotData = {
+      trueSolution : ysTrueSol,
+      solution     : ysSol
+    };
+
+    drawPlot( 'ssr' , ssrPlotData );
+
+  },
+
+
+  ssrLink : function( i ){
+    
+    this.plotIt( this.solutions[i] );
+
+  },
+
+
+
+};});

### 2013-05-09 v3/server/js/utils.js

@@ -1,0 +1,231 @@
+var curr = function( fun ){
+  return curryfy( fun.length , fun , [] );
+};
+
+var curryfy = function( ii , fun , args ){
+
+  if( ii == 0 ){
+    return fun.apply( null , args );
+  }
+
+  return function(){
+
+    var args2 = args;
+    for (var j = 0; j < arguments.length; j++) {
+      args2 = args2.concat( [arguments[j]] );
+    }
+
+    return curryfy( ii-arguments.length , fun , args2 );
+  };
+
+};
+
+
+var curt = function( fun ){
+  var origNumArgs = fun.length ;
+  return function(){
+    console.log( "num args original: " + origNumArgs );
+    console.log( "num args applied : " + arguments.length );
+    console.log( "arguments:         " + arguments );
+
+    return [fun.apply( null , arguments ),arguments];
+  };
+}
+
+var argsToArr = function( args ){
+    ret = [];
+    for (var j = 0; j < arguments.length; j++) {
+      args2 = args2.concat( [arguments[j]] );
+    }  
+}
+
+var te1 = function( a , b , c ){
+  return a + b + c;
+};
+
+
+var cur2 = function( fun ){
+  return function( x , y ){
+    if( y === undefined ){
+      return function(y){ return fun(x,y) ; }
+    } else {
+      return fun( x , y );
+    }
+  }
+}; 
+
+var cur3 = function( fun ){
+  return function( x , y , z ){
+    if( y === undefined ){
+      return cur2( function(y,z){return fun(x,y,z);} );
+    } else if( z === undefined ){
+      return function(z){return fun(x,y,z);} ;
+    } else {
+      return fun(x,y,z);
+    }
+  }
+};
+
+var plus2 = cur2(function(x,y){
+  return x + y;
+});
+
+var plus3 = cur3(function(x,y,z){
+  return x + y + z;
+});
+
+
+
+//var k = _.partial(curr( function( x , y ){
+//  return x;
+//});
+
+var slice = Array.prototype.slice ;
+
+var ap = function(func) {
+    var args = slice.call(arguments, 1);
+    return function() {
+      return func.apply(this, args.concat(slice.call(arguments)));
+    };
+  };
+
+
+
+
+var not = function( x ){
[…]

### 2013-05-13 v3/IM.hs
found the bug in algorithm design of polyIM, started work on new design with MGU

@@ -144,8 +144,7 @@ data SearchOptions = SearchOptions{
   so_runLen             :: Maybe Int          , -- Nothing means unlimited runLen 
   so_stdGen             :: StdGen             ,
   so_edgeSelectionModel :: EdgeSelectionModel ,
-  so_randomRunState     :: RandomRunState     ,
-  so_isPolymorphic      :: Bool
+  so_randomRunState     :: RandomRunState     
  }
 
 defaultSearchOptions :: Int -> Typ -> Context -> SearchOptions
@@ -156,8 +155,7 @@ defaultSearchOptions n typ ctx = SearchOptions {
   so_runLen             = Nothing ,
   so_stdGen             = mkStdGen 42424242 ,
   so_edgeSelectionModel = AllEdges ,
-  so_randomRunState     = NoRandomRunState ,
-  so_isPolymorphic      = False
+  so_randomRunState     = NoRandomRunState 
  }
 
 kozaSearchOptions :: Int -> Typ -> Context -> StdGen -> SearchOptions
@@ -168,8 +166,7 @@ kozaSearchOptions n typ ctx stdGen = SearchOptions {
     so_stdGen             = stdGen ,
     so_runLen             = Just 1 , 
     so_randomRunState     = KozaRandomRunState Nothing Nothing ,
-    so_edgeSelectionModel = KozaESM ,
-    so_isPolymorphic      = False
+    so_edgeSelectionModel = KozaESM 
   } 
 
 allEdgesSearchOptions :: Int -> Typ -> Context -> StdGen -> SearchOptions
@@ -180,8 +177,7 @@ allEdgesSearchOptions n typ ctx stdGen = SearchOptions {
     so_stdGen             = stdGen ,
     so_runLen             = Nothing   ,
     so_randomRunState     = NoRandomRunState , 
-    so_edgeSelectionModel = AllEdges ,
-    so_isPolymorphic      = False  
+    so_edgeSelectionModel = AllEdges   
   } 
 
 geomSearchOptions :: Double -> Int -> Typ -> Context -> StdGen -> SearchOptions
@@ -192,8 +188,7 @@ geomSearchOptions p n typ ctx stdGen = SearchOptions {
     so_stdGen             = stdGen ,
     so_runLen             = Nothing   ,
     so_randomRunState     = NoRandomRunState , 
-    so_edgeSelectionModel = ContinueProbGeomWithDepth p  ,
-    so_isPolymorphic      = False 
[…]

### 2013-05-13 v3/IM2.hs

@@ -604,6 +604,11 @@ fillSpaces = intercalate " " . map show
 
 -- examples/tests : -----------
 
+t0 = Just$  mkZTree $ defaultSearchOptions 2 type_poly_head ctx_poly_head
+t1 = t0 >>= step >>= return . head
+t2 = t1 >>= step >>= return . head
+t3 = t2 >>= step >>= return
+
 h0' = mkZTree $ defaultSearchOptions 100 type_head ctx_head 
 h0  = Just h0'
 h1  = h0 >>= step >>= return . head 
@@ -635,29 +640,40 @@ t_2 = TreeLam ["x","y"] (Typ "A" :-> Typ "B" :-> Typ "C")
 --t1 = mkZTree ctx_head t_2
 
 
-int :: Typ
 int   = Typ "Int"
+
 m_int = Typ "MaybeInt"
 l_int = Typ "[Int]"
+
 int1 = int :-> int
 int2 = int :-> int :-> int
 
+a_ = TypVar "a"
+b_ = TypVar "b"
 
-m_a   = TypFun "Maybe" [TypVar "a"]
-l_a   = TypFun "List"  [TypVar "a"]
-m_Int = TypFun "Maybe" [Typ "Int"]
-l_Int = TypFun "List"  [Typ "Int"]
+m_a   = TypFun "Maybe" [a_]
+l_a   = TypFun "List"  [a_]
+m_Int = TypFun "Maybe" [int]
+l_Int = TypFun "List"  [int]
+
 
 
 
 type_poly_head :: Typ
 type_poly_head = l_Int :-> m_Int
 
+-- ctx_poly_head =    
+--   --[  ( "listCase" , l_a :-> b :-> (a:->l_a:->b ) :-> b  )
+--   [  ( "listCase" , l_a :-> m_a :-> (a:->l_a:->m_a ) :-> m_a  )
+--   ,  ( "Nothing"  , m_a                                                    )
+--   ,  ( "Just"     , a :-> m_a                                              )]
+
+
 ctx_poly_head :: Context
 ctx_poly_head = 
-  [ ( "listCase" , l_Int :-> m_Int :-> (int:->l_Int:->m_Int) :-> m_Int )
+  [ ( "listCase" , l_a :-> b_ :-> (a_:->l_a:->b_) :-> b_ )
   , ( "Nothing"  , m_a )--m_a )
-  , ( "Just"     , int :-> m_Int ) ]
+  , ( "Just"     , a_ :-> m_a ) ]
 
 
 
[…]

### 2013-05-13 v3/PolyUtils.hs

@@ -1,4 +1,4 @@
-module PolyUtils ( Substi , match , applySubsti ) where
+module PolyUtils ( Substi , mgu , applySubsti , composeSubsti , match ) where
 
 -- import Data.List 
 -- import Data.Maybe
@@ -14,13 +14,79 @@ import TTerm (Typ(..),Symbol)
 type Substi = Map Symbol Typ
 
 applySubsti :: Substi -> Typ -> Typ
-applySubsti s typ = case typ of
- Typ    _    -> typ
- TypVar x    -> case Map.lookup x s of
-                 Nothing -> typ
-                 Just t  -> t
- TypFun f ts -> TypFun f (map (applySubsti s) ts)
- a :-> b     -> (applySubsti s a) :-> (applySubsti s b)
+applySubsti s typ 
+ | Map.null s = typ
+ | otherwise  = case typ of
+  Typ    _    -> typ
+  TypVar x    -> case Map.lookup x s of
+                  Nothing -> typ
+                  Just t  -> t
+  TypFun f ts -> TypFun f (map (applySubsti s) ts)
+  a :-> b     -> (applySubsti s a) :-> (applySubsti s b)
+
+composeSubsti :: Substi -> Substi -> Substi
+composeSubsti new old = 
+ let old' = fmap (applySubsti new) old
+  in Map.filterWithKey (\ x t -> t /= TypVar x ) $ Map.union old' new 
+
+
+s1 = Map.fromList [ ("x",a_:->b_),("k",m_) ]
+
+s2 = Map.fromList [ ("a",TypFun "F" [i_,j_,k_] ) , ("m",k_) , ("k",i_) ]
+
+
+a_ = TypVar "a"
+b_ = TypVar "b"
+j_ = TypVar "j"
+i_ = TypVar "i"
+e_ = TypVar "e"
+g_ = TypVar "g"
+k_ = TypVar "k"
+m_ = TypVar "m"
+
+testik = mgu ((a_:->a_):->a_) ((((j_:->k_):->i_):->((e_:->e_):->(g_:->k_))):->m_)
+
+mgu :: Typ -> Typ -> Maybe Substi
+mgu t1 t2 = mgu' (Map.empty) [(t1,t2)]
+ where
+
+  mgu' :: Substi -> [(Typ,Typ)] -> Maybe Substi
+  mgu' s todo = case todo of
+    [] -> Just s
[…]

### 2013-05-31 v3/TTree.hs
[bugfix] bug was in Class/Instance for JShow ..
Bug se objevil po tom co sem na notebooku použil novou haskell
platformu, která se trošičku jinak chová k overlaping instances fíčuře.
Byl to divnej bug pač všecko běželo akorat se neobjevovali linky ve
fenotypu .... poučení: bacha na overlaping instances a dat je asi pryč i
z toho druhyho mista kde někde sou

@@ -1,4 +1,4 @@
-{-# LANGUAGE OverlappingInstances #-}
+-- {-# LANGUAGE OverlappingInstances #-}
 
 module TTree 
 ( CTT (..),

### 2013-05-31 v3/Utils.hs

@@ -1,4 +1,5 @@
-{-# LANGUAGE FlexibleInstances, UndecidableInstances, OverlappingInstances #-}
+{-# LANGUAGE FlexibleInstances, UndecidableInstances #-}
+-- , OverlappingInstances 
 
 module Utils where
 
@@ -218,6 +219,7 @@ align width str =
 
 class (Show a) => JShow a where
   jshow_js :: a -> Maybe JSValue
+  jshow_js _ = Nothing
   jshow    :: a -> JSValue
   jshow x = jsObj $ l_js ++ [ 
     ("type"     , jsStr "jsonout" ) ,
@@ -227,9 +229,6 @@ class (Show a) => JShow a where
      Nothing    -> []
      Just jsval -> [("js",jsval)] 
 
-
-instance Show a => JShow a where 
-  jshow_js _ = Nothing
 
 instance JShow Bool where
   jshow_js x = Just $ JSBool x 

### 2013-06-21 tex/thesis.tex
v3 now running ok (but..)
V zásedě jediná podstatná zmena oproti v2 je předělání Eva monády (+
pročištění).
Z nějakýho důvodu to ale je asi míň efektivní než to předtim což se
projevuje
u BA pro pop 5000 to hodi stackoverflow co to předtim v pohodě zvladalo.
found the bug in algorithm design of polyIM, started work on new design with MGU
..
[tex] outline of GP..

@@
 \section{Genetic Programming}
-	\subsection{Term generating}
-	\subsection{Crossover}
-	\subsection{Mutation}
+
+	\textit{Genetic programming} (GP) is a technique inspired by biological evolution
+	that for a given problem tries to find computer programs able to solve that problem.
+
+	A problem to be solved is given to GP in a form of \textit{fitness function}. 
+	Fitness function is a function which takes computer program as its input and 
+	returns numerical value called \textit{fitness} as output. 
+	The bigger fitness of a computer program is, the better solution of a problem it is.
+
+	GP maintains a collection of computer programs called \textit{population}. 
+	A member of population is called \textit{individual}. 
+	By running GP algorithm evolution of those individuals is performed.
+
+	\Lets describe GP algorithm briefly.
+
+	At the beginning initial population is generated randomly.
+
+	A step of GP algorithm is stochastic transformation of the current population into 	
+	the next population.
+
+	This step consists of two sub steps\footnote{TODO : Technically it is done in a little  
+	bit different fashion which is equivalent.}:
+	\begin{itemize} 
+		\item Selection of \textit{parents} for individuals of the next population based on the   		          fitness.
+		      The bigger fitness of an individual of the current population is, 
+		      the better chance of success being selected as parent it has.  
+		\item Application of genetic operators (such as \textit{crossover}, 
+		      \textit{reproduction} and \textit{mutation}) 
+			  on parent individuals producing new individuals of the next population.  
+	\end{itemize}	  
+	This transformation is repeatedly applied for a predefined number of steps (which is called 
+	number of \textit{generations}) or until some predefined criterion is met.
+
+	...
+	\\\\
+	\textbf{TODO}
+	\begin{itemize} 
+		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
+		      specifying the differences or something like that...    
+		\item History, citations, etc ....
+	\end{itemize}
+
+	\subsection{Input}
+	\subsection{Output}
+	\subsection{GP algorithm}
+		\subsubsection{Generating initial population}
+		\subsubsection{Selection}
+		\subsubsection{Crossover}
+		\subsubsection{Reproduction}
+		\subsubsection{Mutation}
 
@@
 		\item TALK ABOUT "parenthesis" conventions (and packing of lambda abstractions).
+		\item BETTER SPECIFICATION $V$ is infinite spočetná (?countable)
 	\end{itemize}
@@
 			\subsubsection{ Eva.hs }
+			\subsubsection{ GP\_{}Core.hs }
 	\section{Term generating}

### 2013-06-22 tex/thesis.tex
[tex] GP pseudocode...

@@
 \usepackage{qtree}
+
+\usepackage[vlined]{algorithm2e}
+%% \usepackage{algpseudocode}
+\usepackage{framed}
+
 \usepackage{xspace}
@@
 \newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
+
+\newcommand{\Pseudokod}[2]{
+	\begin{framed}
+	\begin{algorithm}[H]
+		\DontPrintSemicolon
+		\SetKwProg{Fn}{Algorithm}{}{}
+		\Fn{#1}{#2}
+	\end{algorithm}
+	\end{framed}
+}
 
@@
 that for a given problem tries to find computer programs able to solve that problem. 
+GP was developed by John Koza \cite{koza92} in 1992.
 
@@
 returns numerical value called \textit{fitness} as output. 
-	The bigger fitness of a computer program is, the better solution of a problem it is.
+The bigger fitness of a computer program is, the better solution of a problem.
 
@@
 
-	\Lets describe GP algorithm briefly.
-
-	At the beginning initial population is generated randomly.
+Individuals are computer program \textit{expressions} kept as \textit{syntactic trees}. 
+Basically those trees are rooted trees with a function symbol in each internal node 
+and with constant symbol or variable symbol in each leaf node. 
+Number of child nodes for each internal node corresponds to the number of arguments of a function whose
+symbol is in that node.
+
+Another crucial input besides fitness function is a collection of \textit{building blocks}.
+It is collection of symbols (accompanied with an information about number of arguments).
+Those symbols are used to construct trees representing individuals.  
+\\\\
+\Lets describe GP algorithm briefly:
+
+At the beginning initial population is generated from building blocks randomly.
 
@@
 number of \textit{generations}) or until some predefined criterion is met.	
-
-	...
 \\\\
-	\textbf{TODO}
-	\begin{itemize} 
-		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
-		      specifying the differences or something like that...    
-		\item History, citations, etc ....
-	\end{itemize}
+\Lets now look on GP at more detail. Here is pseudocode for GP algorithm:
+
+\Pseudokod{GP(fitnessFun,buildingBlocks,numGens,popSize,probabs)}{
+	$gen \leftarrow$ 0 \;
+	$pop \leftarrow generateInitialPopuletion( buildingBlocks ) $ \;
+	($popWithFitnesses,terminate) \leftarrow$ evaluate($fitnessFun$, $pop$) \;
+
+	\While{ $gen < numGens$ $\wedge$ $\neg terminate$  }{
+
+	$newPop \leftarrow$ empty population \;
+	$i \leftarrow 0$ \;
+
+	\While{ $i < popSize$ }{
+		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;		
+		\Switch{op}{
+			\Case{Crossover}{
+				$indiv1 \leftarrow$ selection( $popWithFitnesses$ ) \;
+				$indiv2 \leftarrow$ selection( $popWithFitnesses$ ) \;
+				($child1$,$child2$) = crossover( $indiv1$ , $indiv2$ ) \;
+				$newPop$.insert( $child1$ ) \;
+				$newPop$.insert( $child2$ ) \;
+				$i \leftarrow i + 2$ \;
+			}
+			\Case{Reproduction}{
+				$indiv \leftarrow$ selection( $popWithFitnesses$ ) \;
+				$newPop$.insert( $indiv$ ) \;
+				$i \leftarrow i + 1$ \;
+			}
+			\Case{Mutation}{
+				$indiv \leftarrow$ selection( $popWithFitnesses$ ) \;
+				$mutant \leftarrow$ mutate( $indiv$ ) \;
+				$newPop$.insert( $mutant$ ) \;
+				$i \leftarrow i + 1$ \;			
+			}		
+		}
+	}
+
+	$pop \leftarrow newPop$  \;
+	$gen \leftarrow gen + 1$ \;
+	\;
+	($popWithFitnesses,terminate) \leftarrow$ evaluate($fitnessFun$, $pop$) \; 
+ }
+ \Return pop \;
+}
+
+In order to clarify the code let us describe in greater detail inputs and
+contained procedures
+generateInitialPopuletion(),
+evaluate(),
+probabilisticallySelectOperation(),
+selection(),
+crossover() 
+and mutation().
 
@@
 		\subsubsection{Mutation}
+
+    ...
+	\\\\
+	\textbf{TODO}
+	\begin{itemize} 
+		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
+		      specifying the differences or something like that...    
+		\item History, citations, etc ....
+		\item ten algoritmus v kozovi dělá "Designate Result" já tam vracim poslední populaci
+	\end{itemize}
 
@@
 
+\bibitem{koza92}
+
+  John R. Koza,
+  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
+  MIT Press, Cambridge, MA,
+  1992.
+
 \bibitem{barendregt10}
+
   Henk Barendregt, Wil Dekkers, Richard Statman,

### 2013-06-22 tex/vystrizek.tex
[tex] GP pseudocode...

--- 2013-06-21 tex/thesis.tex
@@
 
-\setlength\textwidth{145mm}
-\setlength\textheight{247mm}
-\setlength\oddsidemargin{15mm}
-\setlength\evensidemargin{15mm}
-\setlength\topmargin{0mm}
-\setlength\headsep{0mm}
-\setlength\headheight{0mm}
-
-\usepackage[utf8]{inputenc}
-\usepackage{qtree}
-\usepackage{xspace}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amsthm}
-\usepackage{graphicx}
-\usepackage{hyperref}
-
 \newcommand{\Lets}{Let us }
-\newcommand{\lterm}{$\lambda$-term\xspace}
-\newcommand{\lterms}{$\lambda$-terms\xspace}
-
-\newcommand{\turst}[3]{$#1 \vdash #2 : #3$\xspace}
-\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
-
-%% Balíček hyperref, kterým jdou vyrábět klikací odkazy v PDF,
-%% ale hlavně ho používáme k uložení metadat do PDF (včetně obsahu).
-%% POZOR, nezapomeňte vyplnit jméno práce a autora.
-%\usepackage[ps2pdf,unicode]{hyperref}   % Musí být za všemi ostatními balíčky
-%\hypersetup{pdftitle=Typed Functional Genetic Programming}
-%\hypersetup{pdfauthor=Tomáš Křen}
-
-% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
-% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
-\makeatletter
-\def\@makechapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries \thechapter. #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\def\@makeschapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\makeatother
-
-\newcommand\Vtextvisiblespace[1][.3em]{%
-  \mbox{\kern.06em\vrule height.3ex}%
-  \vbox{\hrule width#1}%
-  \hbox{\vrule height.3ex}}
-
-\title{Typed Functional Genetic Programming}
-\author{Tomáš Křen}
-\date{Prague 2013}
 
 \begin{document}
-
-% Trochu volnější nastavení dělení slov, než je default.
-\lefthyphenmin=2
-\righthyphenmin=2
-
-%%% Titulní strana práce
-
-\pagestyle{empty}
-\begin{center}
-
-\large
-
-Charles University in Prague
-
-\medskip
-
-Faculty of Mathematics and Physics
-
-\vfill
-
-{\bf\Large MASTER THESIS}
-
-\vfill
-
-%%% \centerline{\mbox{\includegraphics[width=60mm]{../img/logo.eps}}}
-
-%\begin{figure}[!ht]
-%  \centering
-%  \includegraphics{logo.eps}
-%  \caption{Default}\label{fig:default}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Stačí todle odkomentovat a dát nahoře LaTeX (F2) , pak DVI->PDF (F9)  a pak View PDF
-%\includegraphics[scale=0.5]{logo.eps}
-\includegraphics[scale=0.15]{logomff.png}
-
-\vfill
-\vspace{5mm}
-
-{\LARGE Tomáš Křen}
-
-\vspace{15mm}
-
-% Název práce přesně podle zadání
-{\LARGE\bfseries Typed Functional Genetic Programming}
-
-\vfill
-
-% Název katedry nebo ústavu, kde byla práce oficiálně zadána
-% (dle Organizační struktury MFF UK)
-%%%%Name of the department or institute
-%Department of Theoretical Computer Science and Mathematical Logic\\
-%{\small Department of Theoretical Computer Science and Mathematical Logic} \\
-{\fontsize{0.46cm}{1em}\selectfont 
-Department of Theoretical Computer Science and Mathematical Logic}
-
-\vfill
-
-\begin{tabular}{rl}
-
-Supervisor of the master thesis: & RNDr. Petr Pudlák, Ph.D. \\
-\noalign{\vspace{2mm}}
-Study programme: & Theoretical Computer Science \\ %Teoretická informatika \\
-\noalign{\vspace{2mm}}
-Specialization: & 
-%Neprocedurální programování a umělá inteligence \\
-{\fontsize{0.3cm}{1em}\selectfont 
-%Neprocedurální programování a umělá inteligence} \\
-Non-Procedural Programming and Artificial Intelligence} \\
-\end{tabular}
-
-\vfill
-
-% Zde doplňte rok
-Prague 2013
-
-\end{center}
-
-\newpage
-
-%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
-%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
-
-%%% Na tomto místě mohou být napsána případná poděkování (vedoucímu práce,
-%%% konzultantovi, tomu, kdo zapůjčil software, literaturu apod.)
-
-%% on tam měl %% \openright
-
-\noindent
-Dedication.
-
-\newpage
-
-%%% Strana s čestným prohlášením k diplomové práci
-
-\vglue 0pt plus 1fill
-
-\noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
-
-\medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
-
-\vspace{10mm}
-
-\hbox{\hbox to 0.5\hsize{%
-In ........ date ............
-\hss}\hbox to 0.5\hsize{%
-signature of the author
-\hss}}
-
-\vspace{20mm}
-\newpage
-
-%%% Povinná informační strana diplomové práce
-
-\vbox to 0.5\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Název práce:
-Název práce
-% přesně dle zadání
-
-Autor:
-Jméno a příjmení autora
-
-Katedra:  % Případně Ústav:
-Název katedry či ústavu, kde byla práce oficiálně zadána
-% dle Organizační struktury MFF UK
-
-Vedoucí diplomové práce:
-Jméno a příjmení s tituly, pracoviště
-% dle Organizační struktury MFF UK, případně plný název pracoviště mimo MFF UK
-
-Abstrakt:
-% abstrakt v rozsahu 80-200 slov; nejedná se však o opis zadání diplomové práce
-
-Klíčová slova:
-% 3 až 5 klíčových slov
-
-\vss}\nobreak\vbox to 0.49\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Title:
-% přesný překlad názvu práce v angličtině
-
-Author:
-Jméno a příjmení autora
-
-Department:
-Název katedry či ústavu, kde byla práce oficiálně zadána
-% dle Organizační struktury MFF UK v angličtině
-
-Supervisor:
-Jméno a příjmení s tituly, pracoviště
-% dle Organizační struktury MFF UK, případně plný název pracoviště
-% mimo MFF UK v angličtině
-
-Abstract:
-% abstrakt v rozsahu 80-200 slov v angličtině; nejedná se však o překlad
-% zadání diplomové práce
-
-Keywords:
-% 3 až 5 klíčových slov v angličtině
-
-\vss}
-
-\newpage
-
-\tableofcontents
-
-%\chapter{Introduction}
-\chapter*{Introduction}
-\addcontentsline{toc}{chapter}{Introduction}
-
-\chapter{Definitions}
-
-	\Lets first say some basic definitions.
-
-\section{Genetic Programming}
 
@@
 
-	\Lets describe GP algorithm briefly.
+Individuals are computer program \textit{expressions} kept as \textit{syntactic trees}. 
+Basically those trees are rooted trees with a function symbol in each internal node 
+and with constant symbol or variable symbol in each leaf node. 
+Number of child nodes for each internal node corresponds to the number of arguments of a function whose
+symbol is in that node.
 
-	At the beginning initial population is generated randomly.
+Another crucial input besides fitness function is a collection of \textit{building blocks}.
+It is collection of symbols (accompanied with an information about number of arguments).
+Those symbols are used to construct trees representing individuals.  
+\\\\
+\Lets describe GP algorithm briefly:
+
+At the beginning initial population is generated from building blocks randomly.
 
@@
 
-	This step consists of two sub steps\footnote{TODO : Technically it is done in a little  
-	bit different fashion which is equivalent.}:
+This step consists of two sub steps:
 \begin{itemize} 
@@
 number of \textit{generations}) or until some predefined criterion is met.	
-
-	...
 \\\\
-	\textbf{TODO}
-	\begin{itemize} 
-		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
-		      specifying the differences or something like that...    
-		\item History, citations, etc ....
-	\end{itemize}
-
-	\subsection{Input}
-	\subsection{Output}
-	\subsection{GP algorithm}
-		\subsubsection{Generating initial population}
-		\subsubsection{Selection}
-		\subsubsection{Crossover}
-		\subsubsection{Reproduction}
-		\subsubsection{Mutation}
-
-\section{Lambda term}
-\label{deflam}
-
-	Let $V$ be set of {\it variable names}.  \\* 
-	Let $C$ be set of {\it constant names}.	 \\*		
-	Then $\Lambda$ is set of {\it \lterms} inductively defined as follows:
-
-	\begin{align*}
-		 x   \in V \cup C          &\Rightarrow     x   \in \Lambda \\
-		 M,N \in \Lambda           &\Rightarrow ( M N ) \in \Lambda \\
-		 x   \in V , M \in \Lambda &\Rightarrow ( \lambda x . M ) \in \Lambda 
-	\end{align*}
-
-	\textbf{TODO} 
-	\begin{itemize} 
-		\item TALK ABOUT "parenthesis" conventions (and packing of lambda abstractions).
-		\item BETTER SPECIFICATION $V$ is infinite spočetná (?countable)
-	\end{itemize}
-
-\section{Type}
-\label{deftype}
-
-	Let $A$ be set of {\it atomic type names}. \\*
-	Then $\mathbb{T}$ is set of {\it types} inductively defined as follows:
-
-	\begin{align*}
-	\alpha      \in A          &\Rightarrow     \alpha   \in \mathbb{T} \\
-	\sigma,\tau \in \mathbb{T} &\Rightarrow ( \sigma \rightarrow  \tau ) \in \mathbb{T} 
-	\end{align*}
-
-\textbf{TODO} 
-\begin{itemize}
-	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
-	\item TALK ABOUT arrow/parenthesis conventions. 
-\end{itemize}
-
-\section{Statement of a form $M : \sigma$}
-
-	Let $\Lambda$ be set of {\it \lterms}. \\*
-	Let $\mathbb{T}$ be set of {\it types}.       \\*
-	A {\it statement} $M : \sigma$ is a pair $(M,\sigma) \in \Lambda \times \mathbb{T}$. \\*
-	$M : \sigma$ is vocalized as {\it "$M$ has type $\sigma$"}.\footnote{ 
-	$M : \sigma$ can be also imagined as $M \in \sigma$ } \\*
-	The type $\sigma$ is the {\it predicate} and the term $M$ is the
-	{\it subject} of the statement.
-
-\section{Context}
-
-	Let $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$. 
-	($\Gamma$ is a set of {\it statements} of a form $M : \sigma$.)	\\*
-	Then $\Gamma$ is {\it context} if it obeys following 
-	conditions\footnote{
-	The $\pi_1$ corresponds to the projection of the first component of the Cartesian product.
-	}:
-	\begin{align*}
-		 \forall (x,\sigma) \in \Gamma &: x \in V \cup C \\
-		 \forall s_1,s_2 \in \Gamma &: s_1 \neq s_2 \Rightarrow \pi_1(s_1) \neq \pi_1(s_2)
-    \end{align*}
-
-	In other words context is a set of statements with distinct variables or constants as subjects.
-	\\\\
-	\textbf{TODO:} TALK ABOUT Context represents library/building blocks.
-
-\section{Statement of a form \GMS}
-
-	By writing \GMS we say 
-	{\it statement $M : \sigma$ is derivable from context $\Gamma$ }.
-
-	We construct valid statements of form \GMS by using inference rules.
-
-\section{Inference rule}
-
-	Basically speaking, inference rules are used for deriving statements of a form 
-	\GMS from yet derived statements of such a form.
-	Those inference rules are written in the following form:
-
-	\begin{equation*}
-		\frac{\Gamma_1 \vdash M_1 : \sigma_1 \qquad
-			  \Gamma_2 \vdash M_2 : \sigma_2 \quad
-			  \dotsm \quad
-		      \Gamma_n \vdash M_n : \sigma_n}
-		     {\Gamma_{n+1} \vdash M_{n+1} : \sigma_{n+1}}
-	\end{equation*}
-
-	Suppose we have yet derived statements 
-	$\Gamma_1 \vdash M_1 : \sigma_1 ,
-	 \Gamma_2 \vdash M_2 : \sigma_2 ,
-	 \dots ,
-	 \Gamma_n \vdash M_n : \sigma_n$. 
-	It allows as to use the inference rule to derive statement
-	\mbox{ $\Gamma_{n+1} \vdash M_{n+1} : \sigma_{n+1}$ }.
-
-	For deriving statements including types of a form 
-	$(\sigma \rightarrow \tau)$ are essential those two 
-	inference rules:
-
-	\begin{equation*}
-		\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
-			  \Gamma \vdash N : \sigma }
-		     {\Gamma \vdash (M N) : \tau }
-	\end{equation*}
-
-	\begin{equation*}
-		\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
-		     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
-	\end{equation*}
-
-	This kind of inference rules allows us to derive new statements from yet derived statements, but 
-	what if we do not have any statement yet? 
-	For this purpose we have other kinds of inference rules such as {\it axiom} inference rule:
-
-	\begin{equation*}
-		\frac{( x , \sigma )  \in \Gamma}
-		     {\Gamma \vdash x : \sigma}
-	\end{equation*}
-
-	\Lets consider an example statement of a form \GMS :
-
-	\[
-		\{\} \vdash (\lambda f . (\lambda x . (f x) )) : 
-		(\sigma \rightarrow \tau) \rightarrow ( \sigma \rightarrow \tau ) 
-	\]
-
-	This statement is derived as follows:
-
-	\begin{equation*}
-	\dfrac{
-		\dfrac{ (f,\sigma \rightarrow \tau) \in \{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \}  }
-		     { \{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \} \vdash f : \sigma \rightarrow \tau }
-		\qquad
-		\dfrac{ (x,\sigma) \in \{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \}  }
-		     { \{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \} \vdash x : \sigma }
-		 }
-		 {
-			\dfrac{		 	
-		 		\{ (f,\sigma \rightarrow \tau) , (x,\sigma)  \} \vdash (f x) : \tau
-		 	}{
-				\dfrac{\{ (f,\sigma \rightarrow \tau) \} \vdash (\lambda x . (f x) ) : 
-				\sigma \rightarrow \tau}
-				{ \{ \} \vdash (\lambda f . (\lambda x . (f x) ) ) 
-				  : (\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau) }
-		 	}
-		 }
-	\end{equation*}
-
-\section{Term generating grammar}
-
-Inference rules are good for deriving statements of a form \GMS, but our
-goal is slightly different; we would like to generate many \lterms M for a given type 
-$\sigma$ and context $\Gamma$.
-
-Our approach will be to take each inference rule and transform it to a rule of term generating
-grammar. With this term generating grammar it will be much easier to reason about generating 
-\lterms.
-
-It won't be a grammar in classical sense because we will be operating with infinite sets of
-nonterminal symbols and rules. \footnote{TODO : mention terminal symbols - situation around 
-variables and their construction with ' symbol.}
-
-Let $Non = Type \times Context $ be our {\it nonterminal} set. 
-So for every $i \in Non$ is $i = (\sigma_i , \Gamma_i )$.
-
-\Lets consider each relevant inference rule and its corresponding grammar rule.
-
-First inference rule is {\it implication elimination} also known as 
-{\it modus ponens}: 
-\[
-	\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
-		  \Gamma \vdash N : \sigma }
-	     {\Gamma \vdash (M N) : \tau }
-\]
-\\
-For every $\sigma, \tau \in \mathbb{T}$ and for every {\it context} 
-$\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$ there is a grammar rule of a form\footnote{ 
-Terminal symbols for parenthesis and normally {\it space} now \textvisiblespace \quad (for {\it function application} operator) are visually highlighted. }: 
-\[	
-	( \tau , \Gamma )  \longmapsto
-	\bigg( ( \sigma \rightarrow \tau , \Gamma ) 
-	  \mbox{ \Vtextvisiblespace[1em] } ( \sigma , \Gamma ) \bigg)
-\]
-\\
-
-Second inference rule is {\it implication introduction}: 
-\[
-	\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
-	     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
-\]
-\\
-$\forall \sigma, \tau \in \mathbb{T}$ 
-$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
-$\forall x \in V $ such that there is no $(x,\rho) \in \Gamma$ 
-there is a grammar rule:
-\[ 
-	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
-\]
-\\
-
-Third inference rule is {\it axiom}: 
-\[
-		\frac{( x , \sigma )  \in \Gamma}
-		     {\Gamma \vdash x : \sigma}
-\]
-\\
-$\forall \sigma \in \mathbb{T}$ 
-$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
-$\forall x \in V \cup C $ such that $(x,\sigma) \in \Gamma$ 
-there is a grammar rule:
-\[ 
-	( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}}
-\]
-\\
-
-We will demonstrate \lterm generation on example. 
-Again on $(\lambda f . (\lambda x . (f x) ))$. 
-We would like to generate \lterm of a type 
-$(\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau)$
-with $\Gamma = \{\}$.
-\begin{align*}
-	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
-	\longmapsto & \Big( \mbox{ {\Large $\lambda$f.}}
-	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
-	~ \Big)
-	\\
-	\longmapsto & 
-	\Big( \mbox{ {\Large $\lambda$f. }}
-		\Big( \mbox{ {\Large $\lambda$x. }}
-	  	 	( \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
-		~ \Big)  	 
-	~ \Big)
-	\\
-	\longmapsto & 
-	\Big( \mbox{ {\Large $\lambda$f. }}
-		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
-	  	 	\Big( 
-	  	 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
-			  \mbox{ \Vtextvisiblespace[1em] } 
-			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} )  \Big) 
-		~ \Big)  	 
-	 ~ \Big)
-	\\
-	\longmapsto & 
-	\Big( \mbox{ {\Large $\lambda$f. }}
-		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
-	  	 	\Big( 
-	  	 	  \mbox{ {\Large f}} 
-			  \mbox{ \Vtextvisiblespace[1em] } 
-			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \Big) 
-		~ \Big)  	 
-	~ \Big)		
-	\\
-	\longmapsto & 
-	\Big( \mbox{ {\Large $\lambda$f. }}
-		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
-	  	 	\Big( 
-	  	 	  \mbox{ {\Large f}} 
-			  \mbox{ \Vtextvisiblespace[1em] } 
-			  \mbox{{\Large x}} \Big) 
-		~ \Big)  	 
-	~ \Big)
-\end{align*}
-
-\subsection{"Barendregt-like" inference and grammar rules}
-\label{barlike}
-
-Inference rule 1: 
-\[
-	\frac{\Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} \vdash M : \alpha }
-	     {\Gamma \vdash (\lambda x_1 \dots x_n . M) : 
-	     \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha }
-\]
-\\
-Proof of correctness:
-\[
-	\dfrac{
-		\dfrac
-		 {\Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} \vdash M : \alpha}
-		 {\dfrac
-		   {\Gamma \cup \{ (x_1,\tau_1),\dots,(x_{n-1},\tau_{n-1})\} \cup 
-		                \{(x_n,\tau_n) \} \vdash M : \alpha}
-		   {\dfrac{\Gamma \cup \{ (x_1,\tau_1),\dots,(x_{n-1},\tau_{n-1})\}  
-		                \vdash (\lambda x_n . M) : \tau_n \rightarrow \alpha}
-				  { \vdots }		   
-		   }
-		 }		 
-	 }
-	     {\Gamma \vdash (\lambda x_1 \dots x_n . M) : 
-	     \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha }
-\]
-\\
-... there is a grammar rule:
-\[ 
-	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large 
-	$\lambda x_1 \dots x_n .$ 
-	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
-\]
-\\
-
-Inference rule 2: 
-\[
-	\frac{ (f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma \qquad
-	       \Gamma \vdash M_1 : \rho_1 \quad
-	       \dotsm \quad
-	       \Gamma \vdash M_m : \rho_m        
-	      }
-	     {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
-\]
-\\
-Proof of correctness (\textbf{TODO REPAIR} Conceptually it is ok but there is sazba-bug somewhere): 
-\[
-   \dfrac
-     {\dfrac
-      {\dfrac
-       {\dfrac         
-         {\dfrac  
-          {\dfrac
-           {\boxed{(f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma}}
-           {\Gamma \vdash f : \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha}
-           \quad
-           \boxed{\Gamma \vdash M_1 : \rho_1} }
-          {\Gamma \vdash (f M_1) : \rho_2 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha }
-          }{\vdots} 
-         \quad 
-         \ddots }
-       {\Gamma \vdash (f M_1 \dots M_{m-2}) : \rho_{m-1} \rightarrow \rho_m \rightarrow \alpha}
-       \quad
-       \boxed{\Gamma \vdash M_{m-1} : \rho_{m-1}}  }
-      {\Gamma \vdash (f M_1 \dots M_{m-1}) : \rho_m \rightarrow \alpha}       
-      \quad 
-      \boxed{\Gamma \vdash M_m : \rho_m} }
-	 {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
-\]
-\\
-... there is a grammar rule:
-\[ 
-	( \alpha , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large f }}
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \rho_1 , \Gamma )
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  \dots
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \rho_m , \Gamma )
-	  \quad \bigg)
-\]
-\\\\
-\textbf{TODO} 
-\begin{itemize}
-	\item SHOW correctness of those inference rules by composing them of 
-		  $E^{\rightarrow}$, $I^{\rightarrow}$ and \textit{axiom}.
-	\item SHOW more examples of inference rules transformed into grammar rules.
-	\item DESCRIBE general algorithm for this transformation.
-	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
-	\item TALK ABOUT $\beta \eta$-normal form which is generated by this method.
-\end{itemize}
-
-\section{Inhabitation tree}
-
-Now we will introduce \textit{Inhabitation tree}, structure slightly different from
-\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Henk Barendregt.
-We can think about Inhabitation tree as about unfolded Inhabitation machine.
-The motivation for using Inhabitation trees is belief that it will help us 
-reason about generation of \lterms 
-of a given type $\sigma$ and with a given context $\Gamma$.
-
-\subsection{Definition of Inhabitation tree}
-
-\textit{Inhabitation tree} is a \textit{rooted tree}, possibly infinite. 
-It has two types of nodes:
-
-\begin{samepage}
-\begin{itemize}
-  \item Type nodes   
-  			- containing type $\sigma \in \mathbb{T}$ 
-  			- aka "OR-node" , Nonterminal-node. 
-  \item Symbol nodes 
-  			%- containing $ x^{*} \in \mathfrak P \left(V\right) -blobost pač ordered $ or $c \in C$
-  			- containing "$\lambda$-head" (nonempty finite sequence of variable names) or constant
-  			  name. 
-  			- aka "AND-node" , Terminal-node.
-\end{itemize}
-\end{samepage}
-
-We construct Inhabitation tree for given type $\sigma$ and context $\Gamma$.\\
-We will define Inhabitation tree by describing its construction for a given $(\sigma,\Gamma)$.
-Notice that it will closely follow the rules from \ref{barlike}:
-
-\begin{itemize}
-\item The root of Inhabitation tree for $(\sigma,\Gamma)$ is 
-      \textit{type node} with $\sigma$ as type.
-\item All \textit{type nodes} have as child nodes only \textit{symbol nodes}. 
-\item And all \textit{symbol nodes} have as child nodes only \textit{type nodes}. 
-\end{itemize}
-
-Now we will resolve the child nodes of the root node.
-
-There are two cases of $\sigma$ (recall \ref{deftype}): 
-\begin{description}
-	\item[Atomic type] 
-		$\sigma = \alpha $ where $\alpha \in A$. 
-	\item[Function type] 
-		$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
-		where $n \geq 1, \alpha \in A$.
-\end{description}
-
-First case \textbf{Atomic type} --- i.e., $\sigma = \alpha$ where $\alpha \in A$:\\
-For every $(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha) \in \Gamma$
-where $\alpha \in A$ there is a child \textit{symbol node} of the root containing constant name $f$.
-This symbol node containing $f$ has $m$ child subtrees corresponding to Inhabitation trees for 
-$(\rho_1,\Gamma),\dots,(\rho_n,\Gamma)$.
-
-~
-
-Compare this case with corresponding grammar rule:
-\[ 
-	( \alpha , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large f }}
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \rho_1 , \Gamma )
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  \dots
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \rho_m , \Gamma )
-	  \quad \bigg)
-\]
-
-Second case \textbf{Function type} --- i.e., 
-$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
-where $n~\geq~1, \alpha~\in~A$:\\
-For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.
-
-There is one and only one child \textit{symbol node} of the root containing "$\lambda$-head" 
-$\lambda x_1 \dots x_n$ which stands for sequence of variable names $(x_1,\dots,x_n)$.
-This symbol node containing $\lambda x_1 \dots x_n$
-has one and only one child subtree corresponding to Inhabitation trees for 
-$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$.
-
-~
-
-Compare this case with corresponding grammar rule:
-\[ 
-	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large 
-	$\lambda x_1 \dots x_n .$ 
-	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
-\]
-~\\
-\Lets consider following $(\sigma,\Gamma)$ as a simple example:
-
-\begin{align*}
-\sigma =& ~ \mathbb{B} \rightarrow  \mathbb{B} \rightarrow  \mathbb{B} \\ 
-\Gamma =& \{ ~ true : \mathbb{B}  \\
-        &  , ~ nand :  \mathbb{B} \rightarrow \mathbb{B} \rightarrow \mathbb{B} ~ \}
-\end{align*}
-
-This particular $(\sigma,\Gamma)$ results in the following tree:\\
-
-\Tree
-[.\text{ $\mathbb{B} \rightarrow \mathbb{B} \rightarrow \mathbb{B}$ }  
-	[.\textbf{$\lambda$x_1 x_2 } 
-		[.\text{ $\mathbb{B}$ } 
-			\textbf{true}  
-			[.\textbf{nand} 
-				\qroof{ ~~ $\dotsb$ ~~ }.\text{ $\mathbb{B}$ }
-				\qroof{ ~~ $\dotsb$ ~~ }.\text{ $\mathbb{B}$ } 
-			]
-			\textbf{x_1}
-			\textbf{x_2}
-		]
-	]
-]
-
-~
-
-Second example features our well known example:
-
-\begin{align*}
-\sigma =& ~ (\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau \\ 
-\Gamma =& \{ \}
-\end{align*}
-
-Which results in following tree:
-
-\Tree
-[.\text{ $(\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau $ } 
-	[.\textbf{$\lambda$ f x }	
-		[.\text{ $\tau$ }		
-			[.\textbf{f} 
-				[.\text{ $\delta$ }
-					\textbf{x}					
-				]
-			]
-		]
-	]
-]
-
-\subsection{And-or tree and searching in Inhabitation tree}
-\Lets consider following definition of 
-\textit{And-or tree}\footnote{
-\textbf{TODO}: Mention that on WIKI there is more general definition, but 
-for our purposes is this one sufficient.}:
-
-\textit{And-or tree} is a rooted tree where each node is labeled as either \textit{and-node} 
-or\footnote{xor} \textit{or-node}.
-
-By \textit{solving} And-or tree $T$ we mean finding $T'$
-subtree of $T$ such that it follows these conditions: 
-\begin{itemize}
-	\item The root of $T'$ is the root of $T$.
-	\item Each \textit{and-node} in $T'$ has all the child nodes as in $T$.
-	\item Each \textit{or-node}  in $T'$ has precisely one child 
-	      node.\footnote{\textbf{TODO}: MENTION why precisely one 
-	      and not at least one ..or CHANGE the def. }   
-\end{itemize}
-~
-\Lets now consider following labeling of Inhabitation tree:
-
-\begin{itemize}
-  \item \textbf{Type nodes}   are labeled as \textbf{or-nodes}.   
-  \item \textbf{Symbol nodes} are labeled as \textbf{and-nodes}.
-\end{itemize}
-
-This labeling has following justification:
-
-Selection of exactly one
-child node in \textit{type node} corresponds to selection of exactly one
-grammar rule in order to rewrite nonterminal symbol.
-
-Selection of all the child nodes in \textit{symbol node} corresponds to 
-rewriting all the nonterminal symbols in string that is being generated.
-
-The motivation for defining \textit{solving} of a And-or tree the way we did is that
-a found tree $T'$ corresponds to generated \lterm. 
-In order to understand this correspondence let's now talk about various tree representations
-of \lterms.
-
-\subsection{Tree representations of \lterms}
-
-From the definition of \lterm (\ref{deflam}) we can straightforwardly derive 
-the classical tree representation for \lterms. Term M is translated into tree $T[M]$ by following rules:
-
-\begin{itemize}
-	\item $x \in V \cup C$ translates into \textit{leaf} $x$.
-	\item $(M N)$ translates into tree\\
-		\Tree
-			[.@	
-		 		\text{$T[M]$}
-		 		\text{$T[N]$}		 			
-			] 
-	\item $\lambda x . M$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x$}	
-		 	 	\text{$T[M]$}	
-			] 
-\end{itemize}
-
-We can enhance this representation by compressing consecutive lambda abstractions into one
-tree node like this:
-
-\begin{itemize}
-	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x_1 \dots x_n$ }	
-		 	 	\text{$T[M]$}	
-			] 
-\end{itemize}
-
-As this representation comes directly from definition it is evident 
-that it covers all possible \lterms.
-
-For representing expressions as trees it is however more common use a little different
-representation. It will also be the representation suitable for showing 
-that \textit{solving} Inhabitation tree generates wanted \lterm.
-
-\begin{itemize}
-    \item $x \in V \cup C$ translates into \textit{leaf} $x$.
-	\item $(f M_1 M_2 \dots M_n)$ where $f \in V \cup C, n \geq 1$ translates into tree\\
-		\Tree
-			[.f	
-		 		\text{$T[M_1]$}
-		 		\text{$T[M_2]$}
-		 		\text{$\dots$}
-		 		\text{$T[M_n]$}		 				 			
-			] 
-	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x_1 \dots x_n$ }	
-		 	 	\text{$T[M]$}	
-			] 
-\end{itemize}
-
-Notice that this representation does not cover all \lterms, 
-e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.  
-\\\\
-\Lets now consider representation for \textit{typed \lterms}.
-Straightforward approach would be to add to each node a type entry which 
-would be the type of the \lterm corresponding to subtree having this
-node as the root node.
-
-Approach more suitable for our purpose is to add a special type node above each node.
-More specifically:
-
-\Lets consider tree $t$ corresponding to a \lterm of a type
-$\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
-Then corresponding tree $TT[t]$ for typed \lterm is 
-obtained from the tree $t$ as follows:
-
-\begin{equation*}
-\mbox{ 
-TT[
-\Tree
-	[.r 	
-	  	  \text{$s_1$}
-		  \text{$s_2$}
-		  \text{$\dots$}
-		  \text{$s_n$}
-	] 
-}]=
-\mbox{
-\Tree
-	[.\text{$\sigma$ }
-	    [.r 	
-	  	  \text{$TT[s_1]$}
-		  \text{$TT[s_2]$}
-		  \text{$\dots$}
-		  \text{$TT[s_n]$}
-		]	  	
-	] 
-}
-\end{equation*}
-
-Now we can finally put the pieces together. 
-Every solution to a Inhabitation tree has this just described tree form of a typed \lterm.
-
-~\\
-\textbf{TODO} 
-\begin{itemize}
-\item EXAMPLES of tree representations of \lterms  
-\item TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
-		Things about $\beta\eta$ normal form, etc.    
-\end{itemize}
-
-\subsection{Our approach to \textit{solving} Inhabitation machine}
-
-\subsubsection{A* algorithm}
-
-\subsection{Inhabitation Machine}
-\textbf{TODO} 
-\begin{itemize}
-\item DESCRIBE Inhabitation Machine...
-\end{itemize}
-
-\section{Roadmap}
-	\section{Conversion to SKI combinators}
-
-\chapter{ Designed system }	
-	\section{Top level view}
-		\subsection{Comments about main source files}
-			\subsubsection{ Eva.hs }
-			\subsubsection{ GP\_{}Core.hs }
-	\section{Term generating}
-		\subsection{A* algorithm}
-	\section{Crossover}
-		\subsection{Finding same types}
-		\subsection{Two basic options}
-		Resolve problems with free variables or avoid variables completely.
-
-	\section{Mutation}
-		\subsection{Using term generation}
-
-\chapter{Problems}
-	In this section will be presented usage of the system in order to solve specific problems.
-
-		\section{Even Parity Problem}
-		\section{Big Context}
-		\section{Fly}
-		\section{Simple Symbolic Regression}
-		\section{Artificial Ant}
-		\section{Boolean Alternate}
-
-%\chapter{Conclusion}
-\chapter*{Conclusion}
-\addcontentsline{toc}{chapter}{Conclusion}
-
-\begin{thebibliography}{9}
-
-\bibitem{barendregt10}
-  Henk Barendregt, Wil Dekkers, Richard Statman,
-  \emph{Lambda Calculus With Types}.
-  Cambridge University Press,
-  2010. \\
-  \url{http://www.cs.ru.nl/~henk/book.pdf}
-
-\end{thebibliography}
+\Lets now look on GP at more detail.
 

### 2013-06-23 tex/thesis.tex
[tex] ...

@@
 \\\\
-\Lets now look on GP at more detail. Here is pseudocode for GP algorithm:
+\Lets now look on GP at more detail.
+
+\subsection{Program trees}
+
+In GP programs are represented as expressions. \Lets define expression inductively:
+
+\begin{itemize}
+	\item Constant or variable symbol $s$ is expression.
+	\item Let there be a function with symbol $f$ which has $n$ arguments. 
+	      And let there be expressions $e_{1}, ..., e_{n}$. 
+	      Then ( $f$ $e_{1}$ ... $e_{n}$ ) is expression.   
+\end{itemize}
+
+There is straightforward tree representation corresponding to these two cases:
+
+\begin{itemize}
+	\item One node tree $s$.
+    \item \Tree
+			[.$f$	
+		 		\text{$t_{1}$}
+		 		\text{...}
+		 		\text{$t_{n}$} ]\\\\
+		 Where $t_{1}, ..., t_{n}$ are trees corresponding to expressions $e_{1}, ..., e_{n}$.	   
+\end{itemize}
+
+Computer program with inputs $x_{1}, ..., x_{n}$ is realized as expression in which 
+may occur variables $x_{1}, ..., x_{n}$.
+
+\subsection{Building blocks}
+
+Set of building blocks consists of two sets.
+
+\begin{itemize}
+	\item Terminal set $T$ : Set of symbols used as leaf nodes of program trees.
+	\item Function set $F$ : Set of symbols used as internal nodes of program trees.
+\end{itemize}
+
+Therefore $building$ $blocks = T \cup F$.
+
+\subsection{GP algorithm in pseudocode}
+
+Here is pseudocode for GP algorithm:
 

### 2013-06-23 tex/thesis.tex
[tex]

@@
 and with constant symbol or variable symbol in each leaf node. 
-Number of child nodes for each internal node corresponds to the number of arguments of a function whose
-symbol is in that node.
+Number of child nodes for each internal node corresponds to the number of arguments of a function whose symbol is in that node.
 
@@
 \begin{itemize}
-	\item Constant or variable symbol $s$ is expression.
+	\item Constant\footnote{By constants we also mean procedures with zero  	
+		  arguments.} or variable symbol $s$ is expression.
 	\item Let there be a function with symbol $f$ which has $n$ arguments. 
 	      And let there be expressions $e_{1}, ..., e_{n}$. 
-	      Then ( $f$ $e_{1}$ ... $e_{n}$ ) is expression.   
+	      Then ( $f$ $e_{1}$ ... $e_{n}$ ) is expression
+	      \footnote{This notation comes from Lisp programming language, 
+	      classical notation would be $f(e_{1}, ... ,e_{n})$. }.   
 \end{itemize}
@@
 \begin{itemize}
-	\item Terminal set $T$ : Set of symbols used as leaf nodes of program trees.
-	\item Function set $F$ : Set of symbols used as internal nodes of program trees.
-\end{itemize}
-
-Therefore $building$ $blocks = T \cup F$.
+	\item Terminal set $T$ : Set of symbols used as leaf nodes of program trees 		
+		  standing for constants and variables.
+	\item Function set $F$ : Set of symbols used as internal nodes of program trees
+	      standing for functions.
+\end{itemize}
+
+Therefore $building$ $blocks = T \cup F$.\\
+
+%For brevity we will als
+
+There is one important constrain on function implementations for functions from F:
+There should be type $A$ that for every function symbol $f \in F$ the corresponding function implementation standing behind this symbol should be of a type 
+$A \times ... \times A \rightarrow A$ and should be total (defined on all
+combinations of inputs). And analogically for $t \in T$ being of a type A.
+
+Satisfying this constrain ensures that every program tree build 
+from $T \cup F$ will be total.
+
+We can say that the motivation behind this work is to construct system where we
+eliminate this constraint.
+
+\subsection{Generating initial population}
+We will describe tree generating method used by Koza in \cite{koza92} called
+\textit{ramped half-and-half}.
 

### 2013-06-24 tex/thesis.tex
[tex]

@@
 \newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
+
+\newcommand{\EA}{EA } % TODO nahradit správnym "evoluťionar algorithms"
+
+\newcommand{\setDots}[2]{ 
+	\lbrace #1 \dots #2 \rbrace
+}
 
@@
 \begin{itemize}
-	\item Terminal set $T$ : Set of symbols used as leaf nodes of program trees 		
-		  standing for constants and variables.
-	\item Function set $F$ : Set of symbols used as internal nodes of program trees
-	      standing for functions.
+	\item Terminal set $T$ : Set of symbols used as leaf nodes of 	               
+	      program trees standing for constants and variables.
+	\item Function set $F$ : Set of symbols used as internal nodes 
+	      of program trees standing for functions.
 \end{itemize}
@@
 
-%For brevity we will als
+Beside symbols in $T \cup F$ there must also be 
+an implementation for each symbol which is not variable. 
+And for every function symbol from $F$ there must be specified 
+number of arguments.\\
 
@@
 Satisfying this constrain ensures that every program tree build 
-from $T \cup F$ will be total.
+from $T \cup F$ will be total.\\
 
@@
 \subsection{Generating initial population}
+
 We will describe tree generating method used by Koza in \cite{koza92} called
@@
 
+\textbf{TODO!}
+
+\subsection{Selection}
+
+In \EA there is plenty of options for selection mechanisms 
+for us to choose from. Again we will describe mechanism Koza
+used in his first book on GP. It is \textit{Roulette selection}.
+It uses fitness value of each individual in straightforward way to determine probability of selecting this individual.\\
+
+Let there be $popSize$ individuals in the population.\\
+And let $f_{i}$ be fitness value for individual i 
+where $i \in \setDots{1}{popSize}$.
+
+Then $p_{i}$ probability of selection of individual $i$ is computed
+as follows:\\
+
+$p_{i} = \dfrac{ f_{i}  }{ \Sigma_{j=1}^{popSize} f_{j} } $
+
+\subsection{Crossover}
+
+For John Koza the most important genetic operator in GP is 
+crossover. It is operator inspired by sexual reproduction
+occurring in the nature. Generally speaking crossover takes
+two (parent) individuals and combines theirs genomes to produce 
+two possibly new (child) individuals.
+
+In GP the most common mutation is \textit{Subtree swapping mutation}.
+This crossover randomly selects one node in each parent tree.
+Two new child individuals are constructed by swapping subtrees 
+which have roots in those selected nodes.\\
+
+Example should clarify this process. Here are two parent trees with 
+selected nodes in bold:
+
+\Tree [.$ifneq$ $1$
+		 	   [.\textbf{iflt} $0$ $x$ [.$-$ $0$ $x$ ] $1$ ]
+		 	   [.$+$ \text{$x$} \text{$2$} ]
+		 	   $1$ ]
+\Tree [.$\%$ \text{$x$}
+         	 [.\textbf{ifeq} \text{$1$} \text{$x$} \text{$x$} \text{$0$} ] ]\\
+
+And here are two child trees with swapped subtrees:
+
+\Tree [.$ifneq$ $1$
+		 	   [.\textbf{ifeq} \textbf{1} \textbf{x} \textbf{x} 
+		 	     \textbf{0} ]
+		 	   [.$+$ \text{$x$} \text{$2$} ]
+		 	   $1$ ]
+\Tree [.$\%$ \text{$x$}
+         	 [.\textbf{iflt} \textbf{0} \textbf{x} 
+         	   [.\textbf{-} \textbf{0} \textbf{x} ] \textbf{1} ] ]\\
+
+\subsection{Reproduction}
+\subsection{Mutation}
+
+\newpage
 \subsection{GP algorithm in pseudocode}
@@
 
-\Pseudokod{GP(fitnessFun,buildingBlocks,numGens,popSize,probabs)}{
+\Pseudokod{GP(fitness,buildingBlocks,numGens,popSize,probabs)}{
 	$gen \leftarrow$ 0 \;
 	$pop \leftarrow generateInitialPopuletion( buildingBlocks ) $ \;
-	($popWithFitnesses,terminate) \leftarrow$ evaluate($fitnessFun$, $pop$) \;
+	($popWithF,terminate,best) \leftarrow$ evaluate($fitness$, $pop$)\;
 
@@
 	$newPop \leftarrow$ empty population \;
-	$i \leftarrow 0$ \;
+	$newPop$.insert( $best$ )\;
+	$i \leftarrow 1$ \;
 
@@
 			\Case{Crossover}{
-				$indiv1 \leftarrow$ selection( $popWithFitnesses$ ) \;
-				$indiv2 \leftarrow$ selection( $popWithFitnesses$ ) \;
-				($child1$,$child2$) = crossover( $indiv1$ , $indiv2$ ) \;
+				$parent1 \leftarrow$ selection( $popWithF$ ) \;
+				$parent2 \leftarrow$ selection( $popWithF$ ) \;
+				($child1$,$child2$) = crossover( $parent1$ , $parent2$ ) \;
 				$newPop$.insert( $child1$ ) \;
@@
 			\Case{Reproduction}{
-				$indiv \leftarrow$ selection( $popWithFitnesses$ ) \;
+				$indiv \leftarrow$ selection( $popWithF$ ) \;
 				$newPop$.insert( $indiv$ ) \;
@@
 			\Case{Mutation}{
-				$indiv \leftarrow$ selection( $popWithFitnesses$ ) \;
+				$indiv \leftarrow$ selection( $popWithF$ ) \;
 				$mutant \leftarrow$ mutate( $indiv$ ) \;
@@
 	$pop \leftarrow newPop$  \;
+	($popWithF,terminate,best) \leftarrow$ evaluate($fitness$, $pop$)\;
 	$gen \leftarrow gen + 1$ \; 
-	\;
-	($popWithFitnesses,terminate) \leftarrow$ evaluate($fitnessFun$, $pop$) \; 
  }
@@
 	\subsection{Output}
-	\subsection{GP algorithm}
-		\subsubsection{Generating initial population}
-		\subsubsection{Selection}
-		\subsubsection{Crossover}
-		\subsubsection{Reproduction}
-		\subsubsection{Mutation}
 
@@
 	\\\\
-	\textbf{TODO}
+\subsection{TODO}
 	\begin{itemize} 

### 2013-06-24 tex/thesis.tex
[tex] GP outline

@@
 
-\subsection{Generating initial population}
+\subsection{Generating individual}
 
@@
 
-\textbf{TODO!}
+\textbf{TODO!}\\
+\textbf{Talk about Generating initial population}
 
@@
 
-For John Koza the most important genetic operator in GP is 
+For Koza the most important genetic operator in GP is 
 crossover. It is operator inspired by sexual reproduction
@@
 \subsection{Reproduction}
+
+Reproduction is simple mechanism providing preservation of solutions
+from the current population to the next one. It simply copies 
+one individual to the next population.
+
 \subsection{Mutation}
+
+Mutation is genetic operator modifying one individual.
+There are many options for mutation mechanisms 
+for us to choose from. Here will be described 
+\textit{Subtree generating mutation} 
+witch uses mechanism for generating individuals.
+
+This mutation randomly selects one node in the individual tree.
+New mutant individual is constructed by replacement of 
+subtree with root in the selected node by new generated
+tree. This new tree is generated by mechanism for generating 
+individuals.
+
+\subsection{Construction of a next population}
+
+Let $pop_{t}$ be the current population which we want to 
+transform into the next population $pop_{t+1}$. 
+We start by initializing $pop_{t+1}$ by empty population.
+
+Then we iteratively fill $pop_{t+1}$ by individuals 
+returned by genetic operators.
+
+In each iteration one genetic operator is randomly selected.
+Then required amount of individuals for the operation is selected
+by individual selection mechanism described above (one or two
+individuals for our genetic operators).
+And those selected individuals are used as input for selected
+genetic operator which produces some new individuals.
+Those new individuals are inserted into $pop_{t+1}$.
+
+This process continues until $pop_{t+1}$ is filled with
+$popSize$ individuals.
+
+Selection of genetic operation in each operation is 
+controlled by probabilities for each genetic operator.
+
+Koza in \cite{koza92} uses as default probabilities those values:
+
+\begin{itemize}
+	\item \textit{Crossover}    : $70\%$
+	\item \textit{Reproduction} : $30\%$
+	\item \textit{Mutation}     :  $0\%$
+\end{itemize}
 
@@
 
-Here is pseudocode for GP algorithm:
-
-\Pseudokod{GP(fitness,buildingBlocks,numGens,popSize,probabs)}{
+Bellow is described GP algorithm in pseudocode.
+Inputs are:
+
+\begin{itemize}
+	\item \textit{fitness} - Fitness function.
+	\item \textit{TuF} - Building blocks accompanied with an  	
+	      information about implementations and argument numbers.
+	\item \textit{popSize} - Population size.
+	\item \textit{numGens} - Number of generations.
+	\item \textit{probabs} - Genetic operators probabilities.
+\end{itemize}
+
+\Pseudokod{GP(fitness,TuF,popSize,numGens,probabs)}{
 	$gen \leftarrow$ 0 \;
-	$pop \leftarrow generateInitialPopuletion( buildingBlocks ) $ \;
-	($popWithF,terminate,best) \leftarrow$ evaluate($fitness$, $pop$)\;
+	$pop \leftarrow generateInitialPopuletion( TuF ) $ \;
+	($popWithF,terminate,best) \leftarrow$ 
+	evaluate($fitness$, $TuF$, $pop$)\;
 
@@
 				$indiv \leftarrow$ selection( $popWithF$ ) \;
-				$mutant \leftarrow$ mutate( $indiv$ ) \;
+				$mutant \leftarrow$ mutate( $indiv$ , $TuF$ ) \;
 				$newPop$.insert( $mutant$ ) \;
@@
 	$pop \leftarrow newPop$  \;
-	($popWithF,terminate,best) \leftarrow$ evaluate($fitness$, $pop$)\;
+	($popWithF,terminate,best) \leftarrow$ 
+	evaluate($fitness$, $TuF$, $pop$)\;
 	$gen \leftarrow gen + 1$ \; 
@@
 
+\textbf{TODO mention best preservation and same popSize checking. }
+
 	\subsection{Input}
@@
 
+\newpage		
 \section{Lambda term}

### 2013-06-24 tex/thesis.tex
[minor][tex]

@@
 				$parent2 \leftarrow$ selection( $popWithF$ ) \;
-				($child1$,$child2$) = crossover( $parent1$ , $parent2$ ) \;
+				($child1$,$child2$) = crossover( $parent1$ , $parent2$ )\;
 				$newPop$.insert( $child1$ ) \;
@@
 
-	\subsection{Input}
-	\subsection{Output}
-
-    ...
-	\\\\
+...\\\
 \subsection{TODO}

### 2013-06-25 tex/thesis.tex
[tex] Our approach to solving Inhabitation tree and stuff...

@@
 \newcommand{\lterms}{$\lambda$-terms\xspace}
+\newcommand{\lhead}{$\lambda$-head\xspace}
+\newcommand{\lheads}{$\lambda$-heads\xspace}
 
@@
 
-\newcommand{\EA}{EA } % TODO nahradit správnym "evoluťionar algorithms"
+\newcommand{\EA}{EA\xspace} % TODO nahradit správnym "evoluťionar algorithms"
 
@@
 for us to choose from. Again we will describe mechanism Koza
-used in his first book on GP. It is \textit{Roulette selection}.
+used in his first book on GP. It is the \textit{Roulette selection}.
 It uses fitness value of each individual in straightforward way to determine probability of selecting this individual.\\
@@
 Then $p_{i}$ probability of selection of individual $i$ is computed
-as follows:\\
-
-$p_{i} = \dfrac{ f_{i}  }{ \Sigma_{j=1}^{popSize} f_{j} } $
+as follows:
+
+$$ p_{i} = \dfrac{ f_{i}  }{ \sum\limits_{j=1}^{popSize}{f_{j} }  } $$
 
@@
 
+\begin{samepage}
 Now we will resolve the child nodes of the root node.
-
 There are two cases of $\sigma$ (recall \ref{deftype}): 
@@
 \end{description}
+\end{samepage}
 
@@
 		Things about $\beta\eta$ normal form, etc.   
-\end{itemize}
-
-\subsection{Our approach to \textit{solving} Inhabitation machine}
+\item DESCRIBE Inhabitation Machine... 
+\end{itemize}
+
+\subsection{Our approach to \textit{solving} Inhabitation tree}
+
+The main design choice behind our approach to solving inhabitation trees
+is ability to be variable enough to enable choice between 
+\textit{systematic} way of generating terms 
+and \textit{classical-GP-like} way of generating terms 
+by choice of simple search strategy.\\
+
+\Lets explain what is meant by \textit{systematic} and \textit{classical-GP-like}. \\
+
+\textit{Systematic} way of solving inhabitation tree corresponds to
+generating terms in order from smallest to largest in number of symbols and \lheads.
+That means minimizing number of and-nodes in and-or-tree view of point on inhabitation tree.
+But since inhabitation tree is a special kind of and-or-tree with alternating layers of 
+and-nodes (symbols and \lheads) and or-nodes (types) it is also minimizing number of 
+size of whole inhabitation/and-or tree. \\
+
+\textit{Classical-GP-like} ... \textbf{TODO}
 
@@
 
-\subsection{Inhabitation Machine}
-\textbf{TODO} 
-\begin{itemize}
-\item DESCRIBE Inhabitation Machine...
-\end{itemize}
-
-\section{Roadmap}
+A* is a general informed search algorithm used for finding shortest path in a state space.
+According to \cite{AIAMA} A* is the most widely known form of best-first search.
+
 \section{Conversion to SKI combinators}
@@
 	\section{Term generating}
-		\subsection{A* algorithm}
 	\section{Crossover}
@@
 
+\chapter{Ideas yet not implemented}
+
+\section{Roadmap}
+\section{Tree ants}
+\section{[Propojení těch dvou barandrechtskejch odvozovacích pravidel do jedinýho]}
+\section{[Chytřejší heuristika pro A* která si to předpočítá na $\overline{\Gamma}$ ]}
+
 %\chapter{Conclusion}
@@
 
+\bibitem{AIAMA}
+
+	Stuart J. Russell, Peter Norvig,
+	\emph{Artificial Intelligence: A Modern Approach}.
+	Pearson Education,
+	2003.
+
 \end{thebibliography}

### 2013-06-25 tex/thesis.tex
[tex] little bit more on A*

@@
 
-A* is a general informed search algorithm used for finding shortest path in a state space.
+A* is a general informed search algorithm used for finding least-cost 
+path from a given initial state to some goal state in a state space.
 According to \cite{AIAMA} A* is the most widely known form of best-first search.
+
+As A* traverses the state space, it follows a path of the lowest expected total cost.
+It uses a priority queue of alternative unfinished paths.
+
+...\\\\\\\\
+\textbf{TODO : čeknout jak je to v kódu jestli ten A* co tam je čekuje ty
+ nódy co už navštívil při přidávání
+}
 

### 2013-06-29 tex/thesis.tex
[tex] A* and our A* search space ..

@@
 \documentclass[12pt,a4paper]{report}
+
+% Nejistý fráze který tam všude používám
+% ... of a form ...
+% Let us ...
 
@@
 
-\newcommand{\Lets}{Let us }
+\newcommand{\Lets}{Let us\xspace}
+\newcommand{\lets}{let us\xspace}
 \newcommand{\lterm}{$\lambda$-term\xspace}
@@
 
+\newcommand{\la}{\leftarrow\xspace}
+\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
+
 \newcommand{\turst}[3]{$#1 \vdash #2 : #3$\xspace}
@@
 		\DontPrintSemicolon
-		\SetKwProg{Fn}{Algorithm}{}{}
+		\SetKwProg{Fn}{function}{}{}
 		\Fn{#1}{#2}
@@
 	\end{itemize}
+
+\section{Conversion to SKI combinators}
+\textbf{TODO}
 
@@
 
+\newpage
 \section{Inhabitation tree}
@@
   			- containing type $\sigma \in \mathbb{T}$ 
-  			- aka "OR-node" , Nonterminal-node. 
+  			- aka "OR-node" %, Nonterminal-node. 
   \item Symbol nodes 
@@
   			  name. 
-  			- aka "AND-node" , Terminal-node.
+  			- aka "AND-node" %, Terminal-node. - ale nemusí, proto jsem to zakomentoval
 \end{itemize}
@@
 This symbol node containing $f$ has $m$ child subtrees corresponding to Inhabitation trees for 
-$(\rho_1,\Gamma),\dots,(\rho_n,\Gamma)$.
+$(\rho_1,\Gamma),\dots,(\rho_m,\Gamma)$.
 
@@
 
-By \textit{solving} And-or tree $T$ we mean finding $T'$
-subtree of $T$ such that it follows these conditions: 
+By \textit{solving} And-or tree $T$ we mean finding finite tree $T'$
+subgraph of $T$ such that it follows these conditions: 
 \begin{itemize}
@@
 \subsection{Tree representations of \lterms}
+\label{tree-reps}
 
@@
 	\item $x \in V \cup C$ translates into \textit{leaf} $x$.
-	\item $(M N)$ translates into tree\\
+	\item $(P$ $Q)$ translates into tree\\
 		\Tree
 			[.@	
-		 		\text{$T[M]$}
-		 		\text{$T[N]$}		 			
-			] 
-	\item $\lambda x . M$ translates into tree\\
+		 		\text{$T[P]$}
+		 		\text{$T[Q]$}		 			
+			] 
+	\item $\lambda x . P$ translates into tree\\
 	 	\Tree
 			[.\text{$\lambda x$}	
-		 	 	\text{$T[M]$}	
+		 	 	\text{$T[P]$}	
 			] 
@@
 \begin{itemize}
-	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
+	\item $\lambda x_1 \dots x_n . P$ translates into tree\\
 	 	\Tree
 			[.\text{$\lambda x_1 \dots x_n$ }	
-		 	 	\text{$T[M]$}	
+		 	 	\text{$T[P]$}	
 			] 
@@
 
-For representing expressions as trees it is however more common use a little different
+For representing expressions as trees it is however more common to use a little different
 representation. It will also be the representation suitable for showing 
@@
 \item TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
-		Things about $\beta\eta$ normal form, etc.   
+		Things about $\beta\eta^{-1}$ normal form, etc.   
 \item DESCRIBE Inhabitation Machine... 
@@
 
-\subsection{Our approach to \textit{solving} Inhabitation tree}
+\chapter{ A mathematical view of the designed system }
+
+\section{Our approach to \textit{solving} Inhabitation tree}
 
@@
 
-\subsubsection{A* algorithm}
+\subsection{A* algorithm}
 
@@
 As A* traverses the state space, it follows a path of the lowest expected total cost.
-It uses a priority queue of alternative unfinished paths.
+It uses a priority queue of alternative unfinished paths.\\
+
+\textbf{TODO : víc to tu rozepsat hlavně zmínit heuristiku 
+a že stačí aby byla optimistická atd.. }
+
+\Pseudokod{A*( start , isGoal , heuristic )}{\;
+
+	$open \la$ empty priority queue \;
+	$closed \la$ empty set \;\;
+
+	$start$.G = 0 \;
+	$start$.F = 0 + $heuristic$( $start$ ) \;\;
+
+	$open$.insert( $start$ )\;\;
+
+	\While{$\neg$ $open$.isEmpty() }{\;
+
+		$state \la$ $open$.popWithLowestF()\;\;
+
+		\If{ isGoal( $state$ ) }{
+			\Return $state$ \;
+		}\;
+
+		closed.insert( $state$ ) \;\;
+
+		\For{$next \in$ $state$.nexts() }{\;
+
+		    $newG \la$ $state$.G + dist($state$,$next$)\;\;
+
+			\If{ $( next \not\in open \wedge next \not\in closed ) \vee newG < next.$G  }{\;
+
+				$next$.G $\la newG$                          \;
+				$next$.F $\la newG + heuristic$( $next$ )  \;\;
+
+				\If{ $next \not\in open$ }{
+					$open$.insert( $next$ )\;
+				}
+
+			}						
+		}
+	}\;
+
+	\Return no-path-exists 
+}
 
@@
 \textbf{TODO : čeknout jak je to v kódu jestli ten A* co tam je čekuje ty
- nódy co už navštívil při přidávání
-}
-
-\section{Conversion to SKI combinators}
-
-\chapter{ Designed system }	
+ nódy co už navštívil při přidávání - VŽDYŤ díky tomu žeje to tree-search 
+ (vs. graph search) to neni potřeba - pořádně se podívat do AIAMA jakj
+ je terminologie a napsat to sem.
+}
+
+\subsection{Our state space}
+
+% - Stavový prostor 
+%   Nedokončený term
+%   Navíc unfin typ gama
+%   my jen podmnozinu urcenou funkci nexts() a vsemi triv koreny unfin typ gama
+%   Vzdy vezmeme nejlevejsi unfin a udelame z nej konkrétnější term pomoci 
+%   vyberu v or uzlu tim nam vzniknou 
+%   nove unfin uzly odpovidajici pozadavkum and uzlu.
+%   Protoze vyplnovani jednotlivych vetvi je na sobe nezavisle 
+%   tak je jedno v jakem poradi je budem vyplnovat 
+%   a tak je nase "z leva doprava" ok.
+% - isFinal(state)
+%   Term který neobsahuje unfin uzel
+% - Heuristika 
+%   simple -pocet unfin uzlů. 
+%   Lepsi - predpocitat si to chytre s gama' a udelat si mapu v kery kdyz to neni tak dat 1.
+% - Vic nez 1 reseni
+%   staci pokracovat pri nalezeni reseni.
+
+In order to describe our state space to search in \lets introduce
+extended version of our tree representation for typed \lterms from \ref{tree-reps}.
+This extension consists in adding a new type of leaf node $(\sigma,\Gamma)$ 
+standing for unfinished tree of a type $\sigma$ with context $\Gamma$ to use
+as set of building blocks
+\footnote{\textbf{TODO}: Tady to napsat nějak líp s tim gamma }.\\
+
+Initial start states are all of a form "one node tree $(\sigma,\Gamma)$" since we
+are trying to generate a typed \lterm of a type $\sigma$ from building blocks $\Gamma$.\\
+
+A tree with no unfinished leaf is considered as final state.  \\
+
+We will define our state space inductively by specifying algorithm for obtaining 
+successors of a state.
+
+A non-final state must have one or more unfinished leafs. 
+One of those unfinished leafs is selected. 
+It is the first one found by depth-first search (from the root of a state tree)
+i.e., the leftmost unfinished leaf (if we consider that tree written as expression).
+
+Let $(\sigma,\Gamma)$ be the selected leaf. Successor state is constructed 
+by replacing this leaf by new subtree. We must distinguish two cases of $\sigma$ :
+\textit{Atomic type} $\sigma$ and \textit{function type} $\sigma$.\\
+
+\textbf{Atomic type} $\sigma = \alpha $, $\alpha \in A$.
+
+For atomic $\sigma$ there are as many successors as there are members of $\Gamma$ of a form 
+$(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \boldsymbol{\alpha} )$,
+where $m \geq 0 $. In other words all the members of $\Gamma$ of a type
+$\alpha$ or of a type for function which returns $\alpha$.
+
+For each 
+$(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \boldsymbol{\alpha} ) \in \Gamma$,
+where $m > 0$ the new subtree which will replace $(\sigma,\Gamma)$ has following form: \\
+
+\Tree
+   [.$\alpha$
+	[.f	
+ 		\text{$(\rho_1,\Gamma)$}
+ 		\text{$(\rho_2,\Gamma)$}
+ 		\text{$\dots$} 		
+ 		\text{$(\rho_m,\Gamma)$}		 				 			
+	]   
+   ]\\
+
+And for m = 0:
+
+\Tree [.$\alpha$ f ] \\
+
+\textbf{Function type} 
+$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
+where $n \geq 1, \alpha \in A$.
+
+For function type $\sigma$ there is exactly one successor.
+
+For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.
+
+The new subtree which will replace $(\sigma,\Gamma)$ has following form: \\
+
+\Tree
+   [.\text{$\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$}
+	[.\text{$\lambda x_1 \dots x_n$}	
+ 		\text{$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$}		 				
+	]   
+   ]\\
+
+\textbf{TODO : Discuss that it is OK to expand only the DFS-found node. 
+The order of unfinished nodes is irrelevant since generation of each branch is 
+independent on generating other branches. Můžem na to koukat jako na takovou tu redukci 
+simetrií, najít to v AIAMA a pořádně popsat.  }
+
+\subsection{Our heuristic function}
+
+An heuristic function takes as input a state and returns as output an estimation of number
+of steps needed to get to a final state by the shortest possible path.
+
+Our heuristic function is a simple one: The number of unfinished leafs in the state tree.
+
+\textbf{TODO : rozebrat to, ukázat že je optimistická atd...}
+
+\chapter{ An implementation view of the designed system }
+
 \section{Top level view}

### 2013-07-03 tex/thesis.tex
[tex]

@@
 
+\usepackage[ampersand]{easylist}
+
 \usepackage[vlined]{algorithm2e}
@@
 \newcommand{\setDots}[2]{ 
-	\lbrace #1 \dots #2 \rbrace
+	\lbrace #1 , \dots , #2 \rbrace
 }
@@
 
+Témata co zmínit v úvodu :\\
+
+\begin{easylist}[itemize]
+& Začít asi tim že evoluce je cool a že
+  tohle bude o tom jak do GP dát typy.
+& Říct jak jde text:
+  && Začnem popsanim klasickýho GP podle Kozy
+  && Vyslovíme definice okolo lambda termu a typu
+  && Vyslovíme definici inhabitačního stromu
+     kterej je stromovou verzí IM od Barendrechta.
+  && Matematicky popíšem jak na to dem my.
+  && Popíšem rozšíření zakladního systému na 
+     polymorfní.
+  && Popíšem to z programátorskýho hlediska
+  && Popíšem příklady na kterejch to demonstrujem
+  && Srovnáme to s nějakejma přístupama co už 
+     existujou 
+& Možná se zmínit o tý představě dvou pólu
+  induktivním a deduktivním -
+  klasickym GP a Dokazovačem, čim složitější
+  typovej sytem tim víc se jedná o deduktivní   
+  systém a blíží se to dokazovači.
+& Nějak naťuknout už tady v úvodu jak se to má při
+  srovnání s tim jak na to dou jiný (Yu atd)
+& Výkřiky co zmínit:
+  && Na co zaměřený:
+     &&& Proof of a concept
+     &&& Pořádně to vymáchat v typový teorii
+     &&& Spíš než bejt další z řady nástroju
+         se snaží zobecnovat ty už existující
+     &&& Spíš jednoduchost
+         &&&& Ve většině aspektú to sleduje
+              Kozu - tzn i v málo řídících
+              parametru
+     &&& Duraz na generování termu (!)              
+     &&& Po implementační stránce
+         &&&& Sranda s JS překladem
+         &&&& Servrová architektura
+         &&&& oboje spíš srandy
+  && Na co nezaměřený (ale snad taky ok v tom):
+     &&& Performance   
+     &&& Mature knihovna co by šla použít robustně
+         v cizim kódu - spíš to počítá že to 
+         bude dál rozvíjeno a tohle je mezizastávka
+\end{easylist}
+
 \chapter{Definitions}
@@
 
-	Let $V$ be set of {\it variable names}.  \\* 
+\newcommand{\then}{\Rightarrow\xspace}
+
+Let $V$ be infinite countable set of {\it 
+variable names}.  \\* 
 Let $C$ be set of {\it constant names}.	 \\*		
@@
 
+\newcommand{\lam}[2]{( \lambda #1 . #2 )}
+
 \begin{align*}
-		 x   \in V \cup C          &\Rightarrow     x   \in \Lambda \\
-		 M,N \in \Lambda           &\Rightarrow ( M N ) \in \Lambda \\
-		 x   \in V , M \in \Lambda &\Rightarrow ( \lambda x . M ) \in \Lambda 
+x   \in V \cup C          &\then x          \in \Lambda \\
+M,N \in \Lambda           &\then (M~N)      \in \Lambda \\
+x   \in V , M \in \Lambda &\then \lam{x}{M} \in \Lambda 
 \end{align*}
+\\
+It is usual to use $V = \Sigma^+$ set of all non-empty finite strings of symbols 
+from $\Sigma$ where $\Sigma$ is some alphabet set, e.g.  
+$
+\Sigma =
+\setDots{a}{z} \cup 
+\setDots{_0}{_9} $.
+
+And $C$ is set supplied by "the user" to enrich 
+the language with constant names standing
+for some predefined behavior.
+
+A constant may stand just for another \lterm
+or it may stand for some predefined constant 
+such as 0,1,2,... and primitive operations on
+them, e.g. addition. 
+But the specific implementations 
+of these constants will not be ours big concern 
+while reasoning about methods for generating 
+\lterms .\\
+
+We use upper case symbols such as 
+$M,N,M_1,M_2,...,N_1,N_2,...$
+to denote arbitrary lambda terms in contrast to
+those in lower case such as
+$a,b,c,...,x,y,z,$
+$f_1,f_2,...,g_1,g_2,...$
+denoting variable names, elements of $V$.\\
+
+Examples of \lterms:\\
+
+\begin{easylist}[itemize]
+& $x$
+& $\lam{x}{x}$
+& $(\lam{x}{(x~x)}~\lam{x}{(x~x)})$
+& $\lam{x}{\lam{y}{x}}$
+& $\lam{f_1}{\lam{f_2}{\lam{x}{((f_1~x)~(f_2~x))}}}$
+\end{easylist}~
+
+We use following notation
+abbreviations for better readability:
+
+\begin{easylist}[itemize]
+& $M_1~M_2~M_3~\dots~M_n$ for 
+  $(\dots((M_1~M_2)~M_3)~\dots~M_n)$ 
+& $\lam{x_1 x_2 \dots x_n }{M}$ for
+  $\lam{x_1}{\lam{x_2}{\dots\lam{x_n}{M}\dots}}$
+\end{easylist}~
+
+\subsection{$\beta$-reduction}
+
+\subsection{$\eta$-reduction}
+
+\subsection{Reductions for $c \in C$}
+TODO možná vynechat...
+
+\subsection{Normal form}
+
+\subsection{Long normal form}
 
@@
 \begin{itemize} 
-		\item TALK ABOUT "parenthesis" conventions (and packing of lambda abstractions).
-		\item BETTER SPECIFICATION $V$ is infinite spočetná (?countable)
-	\end{itemize}
-
-\section{Conversion to SKI combinators}
-\textbf{TODO}
+	\item TALK ABOUT "parenthesis" conventions 
+	      (and packing of lambda abstractions).
+	\item Popsat ve stručnosti beta redukci,
+	      etha redukci, atd.
+	      Ale hlavně promluvit krátce o intuici
+	      a nějak to napojit na to jak to používám já
+	      pak v tom GP.
+	\item promluvit zde o long normal form což propojuje 
+		  tuhle sekci se sekcí o barenregt-like 	
+		  pravidlech \ref{barlike}.
+\end{itemize}
 
@@
 
-\subsection{"Barendregt-like" inference and grammar rules}
+%"Barendregt-like"
+\subsection{Inference and grammar rules producing 
+terms in Long normal form}
 \label{barlike}
+
+As is said in \cite{barendregt10} there are two
+inference/grammar rules which together generate
+simply typed lambda terms in their long normal forms
+(from section \ref{deflam}).
 
@@
 Now we will introduce \textit{Inhabitation tree}, structure slightly different from
-\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Henk Barendregt.
+\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Barendregt.
 We can think about Inhabitation tree as about unfolded Inhabitation machine.
@@
 
+\section{Our approach to crossover}
+
+\subsection{Conversion to SKI combinators}
+\textbf{TODO}
+
 \chapter{ An implementation view of the designed system }
+
+Předběžný náhodně uspořádaný body osnovy ..\\
+
+\begin{easylist}[itemize]
+& že je to v haskellu
+&& Jak moc hrotit vysvětlení haskellu?
+& že je to servrově řešený
+& architektura evolučního algoritmu 
+&& ty tři typový třídy a ta čtvrtá typová třída
+& evaluace
+&& vychytavka s vyhodnocenim celýho seznamu aby s eto nemuselo furt inicializovat - inicializace vyhodnocovače jednou za generaci 
+& hlavní soubory
+&& Eva.hs     - Eva monada
+&& GP-Core.hs - Evoluční algoritmus
+&& GP-Data.hs - Instance tříd z GP-Core
+&& IMx.hs - Term generating
+&& TTerm.hs   - typed lambda term
+&& TTree.hs   - CTT
+&& Server.hs  - main serveru
+&& Job.hs     - Registrace problemu
+&& Heval.hs   - Evaluator 
+&& složka Problems
+&&& pro každej problém složka 
+&&& ta typicky strukturu
+&&&& Problem.hs
+&&&& Funs.hs
+&& Slozka server
+&&& index.html
+&&& slozka js/Problems
+&&& .js pro kazdej problem
+\end{easylist}
 

### 2013-07-05 tex/thesis.tex
[tex]

@@
 
-\newcommand{\lam}[2]{( \lambda #1 . #2 )}
+\newcommand{\lamb}[2]{( \lambda #1 . #2 )}
+\newcommand{\lam}[2]{\lambda #1 . #2}
 
@@
 x   \in V \cup C          &\then x           \in \Lambda \\
-M,N \in \Lambda           &\then (M~N)      \in \Lambda \\
-x   \in V , M \in \Lambda &\then \lam{x}{M} \in \Lambda 
+M,N \in \Lambda           &\then (M~N)       \in \Lambda 
+\textit{~~~~~~(Function application)} \\
+x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
+\textit{~~~~~($\lambda$-abstraction)} 
 \end{align*}
@@
 $f_1,f_2,...,g_1,g_2,...$
-denoting variable names, elements of $V$.\\
-
-Examples of \lterms:\\
+denoting variable names, elements of $V$.
+Sometimes we will use lower case symbols
+to denote arbitrary variable names (such as
+$x$ in $\lamb{x}{M}$),
+and at other times, we will use them
+as specific variable names in specific terms 
+(such as $x$ in $\lamb{x}{x}$).
+The distinction between the two should be
+clear from the context.\\
+
+Examples of \lterms:
 
@@
 & $x$
-& $\lam{x}{x}$
-& $(\lam{x}{(x~x)}~\lam{x}{(x~x)})$
-& $\lam{x}{\lam{y}{x}}$
-& $\lam{f_1}{\lam{f_2}{\lam{x}{((f_1~x)~(f_2~x))}}}$
+& $\lamb{x}{x}$
+& $(\lamb{x}{(x~x)}~\lamb{x}{(x~x)})$
+& $\lamb{x}{\lamb{y}{x}}$
+& $\lamb{f_1}{\lamb{f_2}{\lamb{x}{((f_1~x)~(f_2~x))}}}$
 \end{easylist}~
@@
 & $\lam{x_1 x_2 \dots x_n }{M}$ for
-  $\lam{x_1}{\lam{x_2}{\dots\lam{x_n}{M}\dots}}$
+  $\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$
 \end{easylist}~
 
+For example we can write
+$\lamb{f_1}{\lamb{f_2}{\lamb{x}{((f_1~x)~(f_2~x))}}}$
+as $\lam{f_1 f_2 x}{f_1~x~(f_2~x)}$.
+
+\subsection{ Subterms, free variables and substitution }
+
+Let us quickly define $ST(M)$ set of subterms of a term $M$:\\
+
+\begin{easylist}[itemize]
+& If $M = x$ then $ST(M) = {x}$.
+& If $M = (P~Q)$ then $ST(M) = \{(P~Q)\} \cup ST(P) \cup ST(Q)$
+& If $M = \lam{x}{M}$ then $ST(M) = \{\lam{x}{M}\} \cup ST(M)$
+\end{easylist}~
+
+\textbf{TODO FV a Substituci...}
+
 \subsection{$\beta$-reduction}
+
+In order to perform computation there must be some
+mechanism for term evaluation. In $\lambda$-calculus there
+is $\beta$-reduction for this reason.\\
+
+Any term of a form $\lamb{x}{M}N$ can be $\beta$-reduced to 
+term $M[x:=N]$. This fact is written as relation of those
+two terms in the following way:
+
+$$\lamb{x}{M}N \rightarrow_\beta M[x:=N]$$
 

### 2013-07-07 tex/thesis.tex
[tex]

@@
 % Let us ...
+
+%takle asi spíš psát ty anglický věty !!!!!!!!!!!!!!!!!!!!!! 
+% takle nak to píše barendrecht :
+%$\FV(M)$ the set of free variables of term M is defined inductively as follows.
 
@@
 \usepackage[ampersand]{easylist}
+
+\usepackage{amssymb}
 
@@
 
+\newcommand{\lamb}[2]{( \lambda #1 . #2 )}
+\newcommand{\lam}[2]{\lambda #1 . #2}
+
 Let $V$ be infinite countable set of {\it 
@@
 Let $C$ be set of {\it constant names}.	 \\*		
-Then $\Lambda$ is set of {\it \lterms} inductively defined as follows:
-
-\newcommand{\lamb}[2]{( \lambda #1 . #2 )}
-\newcommand{\lam}[2]{\lambda #1 . #2}
+Then $\Lambda$ is set of {\it \lterms} defined inductively as follows
+\footnote{Function application and $\lambda$-abstraction are concepts
+well known from common programming languages. 
+For example in JavaScript 
+$(M~N)$ translates to expression \texttt{$M$($N$)} and
+$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
+In other words, the $\lambda$-abstraction is equivalent to 
+\textit{anonymous function}. 
+}.
 
@@
 $M,N,M_1,M_2,...,N_1,N_2,...$
-to denote arbitrary lambda terms in contrast to
+to denote arbitrary lambda terms in contrast with
 those in lower case such as
@@
 
-Let us quickly define $ST(M)$ set of subterms of a term $M$:\\
-
-\begin{easylist}[itemize]
-& If $M = x$ then $ST(M) = {x}$.
-& If $M = (P~Q)$ then $ST(M) = \{(P~Q)\} \cup ST(P) \cup ST(Q)$
-& If $M = \lam{x}{M}$ then $ST(M) = \{\lam{x}{M}\} \cup ST(M)$
-\end{easylist}~
-
-\textbf{TODO FV a Substituci...}
+\newcommand{\ST}{\mathop{\mathrm{ST}}}
+\newcommand{\FV}{\mathop{\mathrm{FV}}}
+
+In order to demonstrate what is a subterm of a term
+\lets define inductively $\ST(M)$ set of all subterms of a term $M$.
+\begin{align*}
+\ST(x)          &= \{x\} \\
+\ST((P~Q))      &= \{(P~Q)\} \cup \ST(P) \cup \ST(Q) \\
+\ST(\lam{x}{P}) &= \{\lam{x}{P}\} \cup \ST(P) 
+\end{align*}
+
+%takle asi spíš psát ty anglický věty !!!!!!!!!!!!!!!!!!!!!! 
+% takle nak to píše barendrecht :
+
+$\FV(M)$ the set of free variables of a term M is defined inductively as follows.
+\begin{align*}
+\FV(x)          &= \{x\} \\
+\FV((P~Q))      &= \FV(P) \cup \FV(Q) \\
+\FV(\lam{x}{P}) &= \FV(P) \smallsetminus \{x\}
+\end{align*}
+
+$M[x:=N]$ the substitution of a term $N$ for the free occurrences of 
+a variable $x$ in a term $M$ is defined inductively as follows.
+\begin{align*}
+x[x:=N]           &= N \\
+y[x:=N]           &= y \text{ ~~~~where } x \not= y  \\
+(P~Q)[x:=N]       &= ( P[x:=N]  ~ Q[x:=N] ) \\
+\lamb{x}{P}[x:=N] &= \lam{x}{P}\\
+\lamb{y}{P}[x:=N] &= \lam{y}{( P[x:=N] )} \text{ ~~~~where } x \not= y
+\end{align*}
+
+\textbf{TODO možná zmínit že y nesmí být v N volně v připadě\\
+$\lamb{y}{P}[x:=N] = \lam{y}{( P[x:=N] )}$ }, ale řekl bych že zbytečnej detail
 
@@
 
-Any term of a form $\lamb{x}{M}N$ can be $\beta$-reduced to 
-term $M[x:=N]$. This fact is written as relation of those
-two terms in the following way:
-
-$$\lamb{x}{M}N \rightarrow_\beta M[x:=N]$$
+\newcommand{\bRedex}{$\beta$-redex\xspace}
+\newcommand{\bRedexes}{$\beta$-redexes\xspace}
+\newcommand{\bArrow}{\rightarrow_\beta\xspace}
+\newcommand{\eArrow}{\rightarrow_\eta\xspace}
+\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}
+
+A term of a form $\lamb{x}{M}N$ is called \textit{\bRedex}.
+A \bRedex can be $\beta$-reduced to term $M[x:=N]$. 
+This fact is written as \textit{relation} $\bArrow$ 
+of those two terms:
+\begin{equation} \label{eq:bRed}
+\lamb{x}{M}N \bArrow M[x:=N]
+\end{equation}
+It is also possible to reduce \textit{subterm \bRedexes} 
+which can be formally stated as:
+\begin{align*}
+P \bArrow Q &\then (R~P)      \bArrow (R~Q) \\
+P \bArrow Q &\then (P~R)      \bArrow (Q~R) \\
+P \bArrow Q &\then \lam{x}{P} \bArrow \lam{x}{Q}  
+\end{align*}
 
@@
 
+Similarly as for $\beta$-reduction we can define $\eta$-reduction 
+except that instead of \ref{eq:bRed} we use:
+
+$$\lamb{x}{(M~x)} \eArrow M \text{ ~~~~where } x \not\in FV(M) $$
+
+\subsection{$\eta^{-1}$-reduction}
+
+$\eta^{-1}$-reduction (also called $\eta$-expansion) is 
+the reduction converse to $\eta$-reduction.
+Again it may be obtained by replacing \ref{eq:bRed}, now with:
+
+$$M \eeArrow \lamb{x}{(M~x)} \text{ ~~~~where } x \not\in FV(M) $$
+
 \subsection{Reductions for $c \in C$}
-TODO možná vynechat...
-
-\subsection{Normal form}
-
-\subsection{Long normal form}
-
+\textbf{TODO možná vynechat...}
+
+\subsection{$\beta$-normal form}
+
+A \lterm is a \textit{$\beta$-normal form} if it does not have a $\beta$-redex as
+subterm.
+\\\\
+A normal form may be thought of as a result of a term evaluation.
+
+~\\\\
 \textbf{TODO} 
 \begin{itemize} 
-	\item TALK ABOUT "parenthesis" conventions 
-	      (and packing of lambda abstractions).
-	\item Popsat ve stručnosti beta redukci,
-	      etha redukci, atd.
-	      Ale hlavně promluvit krátce o intuici
+	\item promluvit krátce o intuici
 	      a nějak to napojit na to jak to používám já
 	      pak v tom GP.
-	\item promluvit zde o long normal form což propojuje 
-		  tuhle sekci se sekcí o barenregt-like 	
-		  pravidlech \ref{barlike}.
+	\item Citovat nějakýho bárendrechta jako zdroj podrobnějších informací
 \end{itemize}
@@
 \end{itemize}
+
+\subsection{Long normal form}
+\label{lnf}
+
+\textbf{TODO : promluvit zde o long normal form což propojuje 
+		  tuhle sekci se sekcí o barenregt-like 	
+		  pravidlech \ref{barlike}.
+}
 
@@
 simply typed lambda terms in their long normal forms
-(from section \ref{deflam}).
+(defined in \ref{lnf}).
 
@@
 	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
-	\item TALK ABOUT $\beta \eta$-normal form which is generated by this method.
+	\item TALK ABOUT $\beta \eta^{-1}$-normal form which is generated by this method.
 \end{itemize}
@@
 simetrií, najít to v AIAMA a pořádně popsat.  }
+
+\textbf{TODO : Odiskutovat nějak tohle:
+přirozeně chceme při produkci velkýho množství termů obhospodařovávat nějaký
+velký množství rozdělanejch termů v nějaký struktuře a postuipně je dodělávat.
+Proto přichází jako tahle struktura na mysl nějaká prioritní fronta,
+otázka je kde vzít ty priority a A* je jednoduchá ale přitom optimum hledající 
+odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.  }
+
+\textbf{TODO: určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
+a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
+udělali jeden důkaz typicky složitýho tvrzení, my sme v situaci kdy
+chceme co nejvíc (a co možná co nejvíc rozmanitejch) důkazů jednoduchýho problému.
+ }
 

### 2013-07-07 v3/Server.hs
Rebuilding server architecture to solve bug caused by System.Directory
.. using STM for concurrent comunication
TODO : in order to make it working jobID must be replaced everywhere by
TVar or something...

@@ -15,7 +15,7 @@ import Control.Concurrent
 import Control.Monad
 import Control.Monad.State
 import System.IO
-import System.Directory
+import System.Directory (doesFileExist, removeFile, createDirectory, doesDirectoryExist,removeDirectoryRecursive,getDirectoryContents)
 import System.Environment
 
 import Utils ( unescape )

### 2013-07-07 v3/Server2.hs

@@ -1,0 +1,392 @@
+{-# LANGUAGE OverloadedStrings #-}
+ 
+--import Network.Wai
+--import Network.Wai.Handler.Warp
+--import Network.HTTP.Types (status200)
+--import Blaze.ByteString.Builder (copyByteString)
+--import qualified Data.ByteString.UTF8 as BU
+--import Data.Monoid
+--import Data.Enumerator (run_, enumList, ($$))
+--
+--import Text.JSON ( encode )
+-- 
+---- já
+--import Control.Concurrent  
+--import Control.Monad
+--import Control.Monad.State
+--import System.IO
+--import System.Directory (doesFileExist, removeFile, createDirectory, doesDirectoryExist,removeDirectoryRecursive,getDirectoryContents)
+--import System.Environment
+--
+--import Utils ( unescape )
+----import Job (job,problemList,problemList_,job_)
+--import Job (problemList_,job_)
+--
+--import ServerInterface
+
+
+
+import Network.Wai (Application,Request(..),Response(..),responseLBS)
+import Network.HTTP.Types (status200)
+import Network.Wai.Handler.Warp (run)
+
+import Control.Monad.State (liftIO)
+import Control.Concurrent (threadDelay,forkIO)
+
+import Control.Monad.STM (atomically)
+import Control.Concurrent.STM.TVar (TVar,newTVar,modifyTVar,swapTVar,readTVarIO,readTVar,writeTVar)
+
+import Data.ByteString.Lazy.UTF8 (fromString)
+import Data.ByteString.Lazy as B (concat, ByteString, append)
+
+import Text.JSON ( encode )
+
+
+
+import Job (problemList_,job_)
+
+
+
+data OutRecord = OutStr String | OutEnd
+type OutputBuffer = TVar [OutRecord]
+type InputCmd = String
+
+data ServerState = ServerState{
+  ssBuff :: OutputBuffer  
+ }
+
+newServerState :: IO ServerState
+newServerState = do
+  buff <- atomically $ newTVar []
+  return ServerState{
+    ssBuff = buff   
+   }
+
+main = do
+    --(port:_) <- getArgs
+    let port = "3000"
+    putStrLn $ "FAKE GUI is on http://localhost:" ++ port 
+    ss     <- newServerState
+    tvarSS <- atomically $ newTVar ss
+    run (read port) (app tvarSS)
+
+app :: TVar ServerState -> Application
[…]

### 2013-07-07 v3/ServerInterface.hs

@@ -1,7 +1,7 @@
 module ServerInterface where
 
 import System.IO
-import System.Directory
+--import System.Directory
 
 import Text.JSON
 import JSONUtils

### 2013-07-08 tex/thesis.tex
[tex]

@@
 
-\subsection{Reductions for $c \in C$}
-\textbf{TODO možná vynechat...}
+%\subsection{Reductions for $c \in C$}
+%\textbf{TODO možná vynechat...}
 
@@
 
+\newcommand{\ar}{\rightarrow\xspace}
+\newcommand{\T}{\mathbb{T}\xspace}
+
 Let $A$ be set of {\it atomic type names}. \\*
@@
 \begin{align*}
-	\alpha      \in A          &\Rightarrow     \alpha   \in \mathbb{T} \\
-	\sigma,\tau \in \mathbb{T} &\Rightarrow ( \sigma \rightarrow  \tau ) \in \mathbb{T} 
+\alpha      \in A  &\then   \alpha \in \T \\
+\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
 \end{align*}
 
+Type $\sigma \ar \tau$ is type for functions taking as input
+something of a type $\sigma$ and returning 
+as output something of a type $\tau$. \\
+
+We use following notation
+abbreviation for better readability:\\
+
+$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ means 
+$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.\\
+
+In other words, the operator $\ar$ is right-associative.\\
+
+It is easy to see that every $\sigma \in \T$ may be expressed as 
+$\tau_1 \ar \dots \ar \tau_n \ar \alpha$ 
+where $\alpha \in A$ and $n \geq 0$.\footnote{ 
+We can prove it by induction on size of $\sigma$. 
+For $\sigma = \alpha \in \T$ is $n = 0$ and for $\sigma = \tau_0 \ar \rho$
+we have from induction hypothesis that $\rho = \tau_1 \ar \dots \ar \tau_n \ar \alpha$,
+therefore
+$\sigma = \tau_0 \ar \tau_1 \ar \dots \ar \tau_n \ar \alpha$.}
+
+Technically speaking, we have types only for functions with one argument, but
+this property suggests how we can "simulate" function which takes $n$ arguments.
+
+% We will utilize this property ... asi netřeba řikat
+
+~\\
 \textbf{TODO} 
 \begin{itemize}
-	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
-	\item TALK ABOUT arrow/parenthesis conventions. 
-\end{itemize}
-
-\subsection{Long normal form}
-\label{lnf}
-
-\textbf{TODO : promluvit zde o long normal form což propojuje 
-		  tuhle sekci se sekcí o barenregt-like 	
-		  pravidlech \ref{barlike}.
-}
+	\item určitě promluvit nějak v úvodu k tomu že se jedná o 
+	Simply typed lambda kalkulus, zmínit že klasicky se těm alfa z A 
+	řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
+	\item možná se hlobějc zmínit o curryingu  
+\end{itemize}
 
@@
 
+\section{Long normal form}
+\label{lnf}
+
+Let \GMS where $\sigma = \tau_0 \ar \tau_1 \ar \dots \ar \tau_n \ar \alpha$.\\
+Then $M$ is in \textit{long normal form} if following 
+conditions are satisfied:
+
+\begin{itemize}
+ \item $M$ is term of a form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$,
+ \item each $M_i$ is in \textit{long normal form}.
+\end{itemize}
+
 %"Barendregt-like"
-\subsection{Inference and grammar rules producing 
+\section{Inference and grammar rules producing 
 terms in Long normal form}
@@
 inference/grammar rules which together generate
-simply typed lambda terms in their long normal forms
-(defined in \ref{lnf}).
+simply typed lambda terms in their long normal forms.
 
@@
 
-\chapter{ A mathematical view of the designed system }
+\chapter{ Our approach }
 

### 2013-07-08 tex/thesis.tex
[tex]

@@
 }
+
+\newenvironment{enum}
+{\begin{easylist}[itemize]}
+{\end{easylist}}
 
@@
 \end{itemize}
+
+As is shown in \cite{barendregt10} long normal form has those 
+nice properties:
+
+\begin{easylist}[itemize]
+& Every $M$ in $\beta$-normal form has a long normal form 
+$N$ such that $N \twoheadrightarrow_{\eta} M$.
+& \textbf{TODO jen zkopčeno přepsat!!} .. if M has a beta-nf, which according to Theorem 2B.4 is
+always the case, then it also has a unique lnf and this will be its unique $\beta\eta^{-1}$-nf.
+& \textbf{TODO : potřebuju eštěnějaký z tyhle knihy?}
+\end{easylist}
+
+\textbf{TODO : popsat tu jak ji předelat na beta nf
+pokud by se nám chtělo... 
+V barendrech84 (ta nová kniha s žlutejma deskama) je ukázano že je 
+eta CR (čurč roserovská) což snad stačí společně s 
+tim tvrzeničkem z book.pdf že vezmu lnf
+a dělám eta redukce dokud to de. až to nejde tak bych to měl mít v 
+beta-nf }
+
+~\\\textbf{TODO : přepsat to nějak hezčejc, ten seznam heskejch vlastností}
 
@@
 
+~\\\textbf{TODO : Napsat tu (nebo asi někam jinam
+ale hlavně někam), že de generovat přímo i termy v beta normalní 
+formě, ale že se mi to zdálo o dost pracnější díky tomu, že
+musíme čekovat jestli neni něco toho typu i v ne atomickejch
+vrcholech/typech... }
+
 \newpage
@@
 
-\section{Our approach to \textit{solving} Inhabitation tree}
+\section{Generating individuals}
+
+\textbf{TODO\\
+Dřív se to menovalo 
+"Our approach to \textit{solving} Inhabitation tree"
+tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
+termy a že je  budeme generovat pomocí 
+řešení inhabitačních stromů.\\\\
+}
 
@@
 
-\section{Our approach to crossover}
+\section{Crossover}
+
+\textbf{TODO}
+\begin{enum}
+& Prostě budem křížit jen podstromy stejnýho typu
+& ALE! Hlavní problém jsou proměnný
+& ve stručnosti kouknout na dvě hlavní metody
+  && křížit prostě lambda stromy a vzniklý problémy řešit online 
+  && zbavit se proměnejch 
+& říct že nakonec sme přilnuli k zbavit se proměnejch
+\end{enum}
+
+\subsection{Finding same types}
+\subsection{Two basic options}
+Resolve problems with free variables or avoid variables 
+completely.
 
@@
 
-\chapter{ An implementation view of the designed system }
+\section{Mutation}
+
+\textbf{TODO: 
+říct že na tu sme se moc nesoustředili ze dvou důvodů
+ktery to ospravedlňujou \\
+1) koza taky\\
+2) ta s vygereovanim podstromu je v zásadě použití
+   algoritmu na generovaní jedince a tak nám přišlo
+   lepší pořádně se soustředit na generování jedince a
+   začít do toho mutaci tahat až bude generování pořádně
+   prozkoumáno.}
+
+\subsection{Using term generation}
+
+\chapter{ Implementation }
 
@@
 \section{Top level view}
-	\subsection{Comments about main source files}
+\section{Comments about main source files}
  \subsubsection{ Eva.hs }
@@
 \section{Crossover}
-	\subsection{Finding same types}
-	\subsection{Two basic options}
-	Resolve problems with free variables or avoid variables completely.
-
 \section{Mutation}
-	\subsection{Using term generation}
 
@@
 \chapter*{Conclusion}
+
+\textbf{Poznámky :\\
+bylo by asi mimojiný dobrý 
+ospravedlnit/okomentovat tak fancy (=javascript frontend) 
+implementaci při 
+dost nedodělanejch jinejch věcech (heuristika by 
+mohla být lepší, mutace by mohli bejt, atd)
+ale věřil sem, že je to proces iterativní
+a že je po celou dobu dobrý mít na čem to testovat.
+Javascript byl vybranej pač HTML je nejpromáklejší
+UI jaký vubec je na zemi teď (state of dzí art)
+a tak to snad nebyla stráta času.
+navíc se naskitla díky tomu další pěkná příležitost
+k ukázání toho, že to neni omezený na haskell
+ale že se kombinatorový konstrukty daj v zásadě přeložit do libovolnýho jazyka celkem bez problémů
+nebo minimalně do JS. Ale todle možná nepsat tolik do závěru
+spíš nějaky podrobnosti přehodit do implementačních detailu
+a tady to jen tak líznout.}
+
 \addcontentsline{toc}{chapter}{Conclusion}

### 2013-07-09 tex/thesis.tex
[tex]

@@
 %$\FV(M)$ the set of free variables of term M is defined inductively as follows.
+
+% !!! místo we can atd psát one can ...
 
@@
 {\end{easylist}}
+
+\newenvironment{todo}
+{ ~\\[1em]
+  \textbf{TODO}
+  \begin{easylist}[itemize]}
+{\end{easylist}~\\[1em]}
 
@@
 
-Therefore $building$ $blocks = T \cup F$.\\
-
-Beside symbols in $T \cup F$ there must also be 
+\newcommand{\TuF}{$T \cup F$\xspace}
+
+Therefore $building$ $blocks = $ \TuF.\\
+
+Beside symbols in \TuF there must also be 
 an implementation for each symbol which is not variable. 
@@
 Satisfying this constrain ensures that every program tree build 
-from $T \cup F$ will be total.\\
+from \TuF will be total.\\
 
@@
 
-We will describe tree generating method used by Koza in \cite{koza92} called
-\textit{ramped half-and-half}.
-
-\textbf{TODO!}\\
-\textbf{Talk about Generating initial population}
+We will describe three tree generating method described by Koza 
+in \cite{koza92} called \textit{ramped half-and-half}. \\
+
+An individual tree is constructed by random (uniform) selection of symbol from 
+subset of \TuF for the root node and after that by generation of its 
+subtrees recursively. 
+Number of subtrees of a node corresponds 
+to the number of arguments for the selected symbol.
+
+There are two generating sub-methods called \textit{full} and 
+\textit{grow} each restricting differently the subset of \TuF. 
+This restriction depends on depth of the
+node for which the symbol is being selected.
+
+In order to perform \textit{ramped half-and-half} method
+one of those two sub-methods is randomly (uniformly) 
+selected and the selected one is performed.\\
+
+In the root (\textit{depth} = 0) both \textit{full} and \textit{grow}
+select the symbol from the set $F$.
+
+A maximum depth $D_{initial}$ is defined, i.e. 6.
+
+In nodes with \textit{depth} $\leq D_{initial}$ the \textit{full}
+method selects from the set $F$, whereas the \textit{grow} method 
+selects from whole \TuF.
+
+In nodes with \textit{depth} $= D_{initial}$ both methods select from the
+set $T$.
 
@@
 
+\begin{todo}
+ & zmínit maximální hloubku vzniklýho dítěte (17)
+\end{todo}
+
 \subsection{Reproduction}
@@
 Selection of genetic operation in each operation is 
-controlled by probabilities for each genetic operator.
+controlled by probabilities for each genetic operator.\\
 
@@
 \begin{itemize}
-	\item \textit{Crossover}    : $70\%$
-	\item \textit{Reproduction} : $30\%$
+	\item \textit{Crossover}    : $90\%$
+	\item \textit{Reproduction} : $10\%$
 	\item \textit{Mutation}     :  $0\%$
@@
 Bellow is described GP algorithm in pseudocode.
-Inputs are:
-
-\begin{itemize}
-	\item \textit{fitness} - Fitness function.
-	\item \textit{TuF} - Building blocks accompanied with an  	
-	      information about implementations and argument numbers.
-	\item \textit{popSize} - Population size.
-	\item \textit{numGens} - Number of generations.
-	\item \textit{probabs} - Genetic operators probabilities.
-\end{itemize}
-
-\Pseudokod{GP(fitness,TuF,popSize,numGens,probabs)}{
-	$gen \leftarrow$ 0 \;
-	$pop \leftarrow generateInitialPopuletion( TuF ) $ \;
+
+\Pseudokod{GP( fitness, \TuF, popSize, numGens, probabs )}{
+  $gen \leftarrow$ 0 \;\;
+
+  $pop \leftarrow generateInitialPopuletion( $ \TuF$ ) $ \;
   ($popWithF,terminate,best) \leftarrow$ 
-	evaluate($fitness$, $TuF$, $pop$)\;
+  evaluate($fitness$, \TuF, $pop$)\;\;
 
   \While{ $gen < numGens$ $\wedge$ $\neg terminate$  }{  	
-
 	$newPop \leftarrow$ empty population \;
-	$newPop$.insert( $best$ )\;
+	$newPop$.insert( $best$ )\;\;
+
 	$i \leftarrow 1$ \;
-
 	\While{ $i < popSize$ }{	
-		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;		
+		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;\;
+
 		\Switch{op}{
+
 			\Case{Crossover}{
 				$parent1 \leftarrow$ selection( $popWithF$ ) \;
-				$parent2 \leftarrow$ selection( $popWithF$ ) \;
-				($child1$,$child2$) = crossover( $parent1$ , $parent2$ )\;
+				$parent2 \leftarrow$ selection( $popWithF$ ) \;\;
+
+				($child1$,$child2$) = crossover( $parent1$ , $parent2$ )\;\;
+
 				$newPop$.insert( $child1$ ) \;
@@
 				$indiv \leftarrow$ selection( $popWithF$ ) \;
-				$mutant \leftarrow$ mutate( $indiv$ , $TuF$ ) \;
+				$mutant \leftarrow$ mutate( $indiv$ , \TuF ) \;
 				$newPop$.insert( $mutant$ ) \;
@@
 		}
-	}
+	}\;
 
@@
 	$gen \leftarrow gen + 1$ \; 
- }
+ }\;
+
  \Return pop \;
@@
 
-In order to clarify the code let us describe in greater detail inputs and
-contained procedures
+\begin{samepage}
+\Lets clarify the input arguments:
+\begin{itemize}
+	\item \textit{fitness} - Fitness function.
+	\item \textit{\TuF} - Building blocks accompanied with an  	
+	      information about implementations and argument numbers.
+	\item \textit{popSize} - Population size.
+	\item \textit{numGens} - Number of generations.
+	\item \textit{probabs} - Genetic operators probabilities.
+\end{itemize} 
+\end{samepage}
+
+%In order to clarify the code let us describe in greater detail inputs and
+
+Behaviors of contained procedures
 generateInitialPopuletion(),
@@
 crossover() 
-and mutation().
-
-\textbf{TODO mention best preservation and same popSize checking. }
-
-...\\\
-\subsection{TODO}
-	\begin{itemize} 
-		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
-		      specifying the differences or something like that...    
-		\item History, citations, etc ....
-		\item ten algoritmus v kozovi dělá "Designate Result" já tam vracim poslední populaci
-	\end{itemize}
+and mutation() are hopefully clear from verbal description.
+
+\begin{todo}
+ & mention best preservation and same popSize checking.
+ & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
+   defining EA and then specifying the differences or something 
+   like that...    
+ & History, citations, etc ....
+ & ten algoritmus v kozovi dělá "Designate Result" já tam vracim 
+   poslední populaci
+\end{todo}
 
@@
 
-\newcommand{\lamb}[2]{( \lambda #1 . #2 )}
-\newcommand{\lam}[2]{\lambda #1 . #2}
+\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
+\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
+
+\newcommand{\ST}{\mathop{\mathrm{ST}}}
+\newcommand{\FV}{\mathop{\mathrm{FV}}}
+
+\newcommand{\Scomb }{\mathbf{S}}
+\newcommand{\Kcomb }{\mathbf{K}}
+\newcommand{\Icomb }{\mathbf{I}}
 
@@
 Let $C$ be set of {\it constant names}.	 \\*		
-Then $\Lambda$ is set of {\it \lterms} defined inductively as follows
-\footnote{Function application and $\lambda$-abstraction are concepts
-well known from common programming languages. 
-For example in JavaScript 
-$(M~N)$ translates to expression \texttt{$M$($N$)} and
-$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
-In other words, the $\lambda$-abstraction is equivalent to 
-\textit{anonymous function}. 
-}.
-
+Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
 \begin{align*}
@@
 \end{align*}
-\\
+
+Function application and $\lambda$-abstraction are concepts
+well known from common programming languages. 
+For example in JavaScript 
+$(M~N)$ translates to expression \texttt{$M$($N$)} and
+$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
+In other words, the function application 
+corresponds to the act of supplying a function 
+with an argument and
+the $\lambda$-abstraction is equivalent to 
+\textit{anonymous function}. \\
+
 It is usual to use $V = \Sigma^+$ set of all non-empty finite strings of symbols 
@@
 \Sigma =
-\setDots{a}{z} \cup 
-\setDots{_0}{_9} $.
+\setDots{a}{z}$.
 
@@
 denoting variable names, elements of $V$.
+
 Sometimes we will use lower case symbols
@@
 The distinction between the two should be
-clear from the context.\\
+clear from the context.
+
+We use bold upper case symbols such as 
+$\Scomb, \Kcomb, \Icomb$ as abbreviation
+for specific terms, e.i. 
+$\Icomb = \lamb{x}{x}$.  
+\\
 
@@
 & $\lamb{x}{x}$
+& $\lamb{x}{f}$
 & $(\lamb{x}{(x~x)}~\lamb{x}{(x~x)})$
 & $\lamb{x}{\lamb{y}{x}}$
-& $\lamb{f_1}{\lamb{f_2}{\lamb{x}{((f_1~x)~(f_2~x))}}}$
+& $\lamb{f}{\lamb{g}{\lamb{x}{((f~x)~(g~x))}}}$
 \end{easylist}~
@@
 For example we can write
-$\lamb{f_1}{\lamb{f_2}{\lamb{x}{((f_1~x)~(f_2~x))}}}$
-as $\lam{f_1 f_2 x}{f_1~x~(f_2~x)}$.
+$\lamb{f}{\lamb{g}{\lamb{x}{((f~x)\,(g~x))}}}$
+as $\lam{f\,g\,x}{f~x\,(g~x)}$.
 
 \subsection{ Subterms, free variables and substitution }
-
-\newcommand{\ST}{\mathop{\mathrm{ST}}}
-\newcommand{\FV}{\mathop{\mathrm{FV}}}
 
@@
 \end{align*}
+
+A variable in \lterm $M$ is called \textit{bound} if it is not free.
+
+\lterm M is called \textit{combinator} if $\FV(M)=\emptyset$.\\
+
+Notable examples of combinators are combinators
+$\Scomb$, $\Kcomb$ and $\Icomb$.
+\begin{align*}
+\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
+\Kcomb &= \lam{x\,y}{x} \\
+\Icomb &= \lam{x}{x} 
+\end{align*}
+
+As will be shown in \ref{toSKI}
+these combinators will help us in performing crossover of \lterms.\\
 
@@
 
-\textbf{TODO možná zmínit že y nesmí být v N volně v připadě\\
-$\lamb{y}{P}[x:=N] = \lam{y}{( P[x:=N] )}$ }, ale řekl bych že zbytečnej detail
+\begin{todo}
+& možná zmínit že y nesmí být v N volně v připadě\\
+  $\lamb{y}{P}[x:=N] = \lam{y}{( P[x:=N] )}$ , 
+  ale řekl bych že zbytečnej detail
+\end{todo}
 
@@
 
+In other words, $\beta$-reduction is the process 
+of insertion of arguments supplied to a function into 
+its body.
+
 \subsection{$\eta$-reduction}
@@
 
-~\\\\
-\textbf{TODO} 
-\begin{itemize} 
-	\item promluvit krátce o intuici
+\begin{todo}
+	& říct co to vubec je ty intutivně (v par větách jen) 
 	  a nějak to napojit na to jak to používám já
 	  pak v tom GP.
-	\item Citovat nějakýho bárendrechta jako zdroj podrobnějších informací
-\end{itemize}
+	& Citovat nějakýho bárendrechta jako zdroj podrobnějších informací
+\end{todo}
 
@@
 
-~\\
-\textbf{TODO} 
-\begin{itemize}
-	\item určitě promluvit nějak v úvodu k tomu že se jedná o 
+\begin{todo}
+   & určitě promluvit nějak v úvodu k tomu že se jedná o 
 	 Simply typed lambda kalkulus, zmínit že klasicky se těm alfa z A 
 	 řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
-	\item možná se hlobějc zmínit o curryingu  
-\end{itemize}
+   & možná se hlobějc zmínit o curryingu  
+\end{todo}
 
@@
 	In other words context is a set of statements with distinct variables or constants as subjects.
-	\\\\
-	\textbf{TODO:} TALK ABOUT Context represents library/building blocks.
+
+\begin{todo}
+ & TALK ABOUT Context represents library/building blocks.
+\end{todo}
 
@@
 
-\textbf{TODO : popsat tu jak ji předelat na beta nf
+\begin{todo}
+ & popsat tu jak ji předelat na beta nf
 	pokud by se nám chtělo... 
@@
 	a dělám eta redukce dokud to de. až to nejde tak bych to měl mít v 
-beta-nf }
-
-~\\\textbf{TODO : přepsat to nějak hezčejc, ten seznam heskejch vlastností}
+	beta-nf
+ & přepsat to nějak hezčejc, ten seznam heskejch vlastností
+\end{todo}
 
@@
 \]
-\\\\
-\textbf{TODO} 
-\begin{itemize}
-	\item SHOW correctness of those inference rules by composing them of 
+
+\begin{todo}
+& SHOW correctness of those inference rules by composing them of 
 	  $E^{\rightarrow}$, $I^{\rightarrow}$ and \textit{axiom}.
-	\item SHOW more examples of inference rules transformed into grammar rules.
-	\item DESCRIBE general algorithm for this transformation.
-	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
-	\item TALK ABOUT $\beta \eta^{-1}$-normal form which is generated by this method.
-\end{itemize}
-
-~\\\textbf{TODO : Napsat tu (nebo asi někam jinam
+& SHOW more examples of inference rules transformed into grammar rules.
+& DESCRIBE general algorithm for this transformation.
+& TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
+& TALK ABOUT $\beta \eta^{-1}$-normal form which is generated by this method.
+& Napsat tu (nebo asi někam jinam
 	ale hlavně někam), že de generovat přímo i termy v beta normalní 
@@
 	musíme čekovat jestli neni něco toho typu i v ne atomickejch
-vrcholech/typech... }
+	vrcholech/typech...
+\end{todo}
 
@@
 
-~\\
-\textbf{TODO} 
-\begin{itemize}
-\item EXAMPLES of tree representations of \lterms  
-\item TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
+\begin{todo}
+& EXAMPLES of tree representations of \lterms  
+& TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
 		Things about $\beta\eta^{-1}$ normal form, etc.   
-\item DESCRIBE Inhabitation Machine... 
-\end{itemize}
+& DESCRIBE Inhabitation Machine... 
+\end{todo}
 
@@
 
-\textbf{TODO\\
-Dřív se to menovalo 
+\begin{todo}
+& Dřív se to menovalo 
 "Our approach to \textit{solving} Inhabitation tree"
@@
 termy a že je  budeme generovat pomocí 
-řešení inhabitačních stromů.\\\\
-}
+řešení inhabitačních stromu.
+\end{todo}
 
@@
 
-\textit{Classical-GP-like} ... \textbf{TODO}
+\textit{Classical-GP-like}
+
+\begin{todo}
+!!!
+\end{todo}
 
@@
 
-\textbf{TODO : víc to tu rozepsat hlavně zmínit heuristiku 
-a že stačí aby byla optimistická atd.. }
+\begin{todo}
+& víc to tu rozepsat hlavně zmínit heuristiku 
+a že stačí aby byla optimistická atd..
+\end{todo}
 
@@
 
-...\\\\\\\\
-\textbf{TODO : čeknout jak je to v kódu jestli ten A* co tam je čekuje ty
+\begin{todo}
+& čeknout jak je to v kódu jestli ten A* co tam je čekuje ty
  nódy co už navštívil při přidávání - VŽDYŤ díky tomu žeje to tree-search 
@@
  je terminologie a napsat to sem.
-}
+\end{todo}
 
@@
 
-\textbf{TODO : Discuss that it is OK to expand only the DFS-found node. 
+\begin{todo}
+ & Discuss that it is OK to expand only the DFS-found node. 
 The order of unfinished nodes is irrelevant since generation of each branch is 
-independent on generating other branches. Můžem na to koukat jako na takovou tu redukci 
-simetrií, najít to v AIAMA a pořádně popsat.  }
-
-\textbf{TODO : Odiskutovat nějak tohle:
-přirozeně chceme při produkci velkýho množství termů obhospodařovávat nějaký
-velký množství rozdělanejch termů v nějaký struktuře a postuipně je dodělávat.
+independent on generating other branches. Mužem na to koukat jako na takovou tu redukci simetrií, najít to v AIAMA a pořádně popsat.
+
+& Odiskutovat nějak tohle:
+přirozeně chceme při produkci velkýho množství termu obhospodařovávat nějaký
+velký množství rozdělanejch termu v nějaký struktuře a postuipně je dodělávat.
 Proto přichází jako tahle struktura na mysl nějaká prioritní fronta,
 otázka je kde vzít ty priority a A* je jednoduchá ale přitom optimum hledající 
-odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.  }
-
-\textbf{TODO: určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
+odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.
+
+& určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
 a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
-udělali jeden důkaz typicky složitýho tvrzení, my sme v situaci kdy
-chceme co nejvíc (a co možná co nejvíc rozmanitejch) důkazů jednoduchýho problému.
- }
+udělali jeden dukaz typicky složitýho tvrzení, my sme v situaci kdy
+chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
+\end{todo}
 
@@
 
-\textbf{TODO : rozebrat to, ukázat že je optimistická atd...}
+\begin{todo}
+ & rozebrat to, ukázat že je optimistická atd...
+\end{todo}
 
@@
 
-\textbf{TODO}
-\begin{enum}
+\begin{todo}
 & Prostě budem křížit jen podstromy stejnýho typu
@@
 & říct že nakonec sme přilnuli k zbavit se proměnejch
-\end{enum}
+\end{todo}
 
@@
 
-\subsection{Conversion to SKI combinators}
-\textbf{TODO}
+\subsection{Abstraction elimination}
+\label{toSKI}
+%... Conversion to SKI combinators
+
+\textit{Abstraction elimination} is a process of transforming 
+an arbitrary \lterm into \lterm witch contains no lambda abstractions
+and no bound variables.
+The newly produced \lterm may contain function applications, 
+free symbols from former \lterm and some new symbols standing for 
+combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
+
+Those combinators are defined as:
+\begin{align*}
+\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
+\Kcomb &= \lam{x\,y}{x} \\
+\Icomb &= \lam{x}{x} 
+\end{align*}
+
+\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
+
+\Lets describe transformation $\Ae$ performing this process.
+\begin{align*}
+\Ae[x]           &= x &\\[0.4em]
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
+\Ae[\lam{x}{x}]  &= \Icomb &\\
+\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
+\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
+&\textbf{if } x \in \FV(M)\\
+\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
+&\textbf{if } x \in \FV(M) \vee x \in \FV(N)\\
+\end{align*}
+
+\begin{todo}
+& přidat (ale i do kódu) B a C
+& říct že my musíme řešit i typy 
+& říct o složitosti , a že je to docela
+disadventič vzhledem k bloatu
+& citovat Simon Peyton Jones implementaition of FP languages
+jako source of further reading on this subject.
+\end{todo}
 
@@
 
-\textbf{TODO: 
-říct že na tu sme se moc nesoustředili ze dvou důvodů
-ktery to ospravedlňujou \\
-1) koza taky\\
-2) ta s vygereovanim podstromu je v zásadě použití
+\begin{todo}
+& říct že na tu sme se moc nesoustředili ze dvou duvodu
+  ktery to ospravedlňujou 
+&& koza taky
+&& ta s vygereovanim podstromu je v zásadě použití
    algoritmu na generovaní jedince a tak nám přišlo
@@
    začít do toho mutaci tahat až bude generování pořádně
-   prozkoumáno.}
+   prozkoumáno.
+\end{todo}
 
@@
 
+\chapter{Comparison with others}
+
 \chapter{Ideas yet not implemented}

### 2013-07-10 tex/thesis.tex
[tex]

@@
 \newenvironment{todo}
-{ ~\\[1em]
+{ ~\\[0.5em]
   \textbf{TODO}
   \begin{easylist}[itemize]}
-{\end{easylist}~\\[1em]}
+{ \end{easylist}
+  ~}
 
@@
 \section{Genetic Programming}
+\label{GP}
 
@@
 
+\begin{todo}
+ & Asi to přejmenvat na S-expression.  
+\end{todo}
+
 \begin{itemize}
@@
 \subsection{Building blocks}
+\label{building-blocks}
 
@@
 Satisfying this constrain ensures that every program tree build 
-from \TuF will be total.\\
+from \TuF will be total.
+
+We will refer to this constrain as \textit{over-one-type} constrain
+in the further text.\\
 
@@
 \subsection{Generating individual}
+\label{GPgene}
 
@@
 
-\subsection{Tree representations of \lterms}
+\section{Tree representations of \lterms}
 \label{tree-reps}
+
+\newcommand{\sexprTree}{sexpr-tree\xspace}
+\newcommand{\atTree}{@-tree\xspace}
 
@@
 that it covers all possible \lterms.
+
+We will refer to this representation as to \textit{\atTree}.\\
 
@@
 e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.
-\\\\
+
+We will refer to this representation as to \textit{\sexprTree}.\\
+
 \Lets now consider representation for \textit{typed \lterms}.
@@
 
+\begin{todo}
+ & říct že to je přesně ta notace co používá koza
+ & EXAMPLES of tree representations of \lterms  
+ & napsat transformace mezi \atTree a \sexprTree
+ & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
+   přímo \sexprTree a říct proč se tomu tak řiká 
+   (na tom místě co je v textu dřív)
+
+\end{todo}
+
+\section{ Generating terms in Long normal form by solving Inhabitation tree }
+
 Now we can finally put the pieces together. 
-Every solution to a Inhabitation tree has this just described tree form of a typed \lterm.
-
-\begin{todo}
-& EXAMPLES of tree representations of \lterms  
+Every solution to an Inhabitation tree has this just described tree form of a typed \lterm.
+
+\begin{todo}
+& !!!
 & TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
 		Things about $\beta\eta^{-1}$ normal form, etc.   
+\end{todo}
+
+\section{ Inhabitation machine }
+
+\begin{todo}
   & DESCRIBE Inhabitation Machine...
+   && něco jako na závěrještě ve stručnosti popišme 
+      IM, strukturu ze který je odvozenej IT. 
+  & říct, že je to hezkej konstrukt vzhledem k tomu že je konečnej
+  & ale je popsanej velmi zrychla (v tý knize) a implementační
+     detaily jsou celkem vynechaný
+  & na stromový verzi jsou implementační detaily hezky vidět
+  & a navíc pro polymorfní verzi jsou už i IM nekonečný takže
+     se strácí i ta konečnostní výhoda
+  & dát tam obrázek 
 \end{todo}
@@
 
+In this chapter we will describe our system trying to 
+dive into implementation details as little as possible.
+Some of those details will be revealed in the next chapter
+devoted to the implementation.
+
+This chapter will be
+presented in the mix between the mathematical terminology 
+described in the previous chapter and the algorithmic terminology.\\
+
+We can summarize our approach to design of 
+the system for Genetic programming with types
+by its main design goals:\\
+
+\begin{enum}
+ & It should eliminate the \textit{over-one-type} constrain of classical GP
+   (see \ref{building-blocks}). 
+ & It should be generalization of the classical GP described in \ref{GP},
+   by which is meant that for \textit{building blocks} satisfying 
+   constrains of classical GP there should be simple setup of control parameters
+   of the system which will make the system behave in precisely the same way
+   as the classical GP behaves. 
+ & It should utilize the theory around typed $\lambda$-calculus.  
+\end{enum}
+
+\section{Individuals and Building blocks}
+
+\begin{todo}
+ & několik fází, několik reprezentací
+ & během generování se reprezentace mění 
+   && nejdří obsahuje lambdy i volný proměnný pak už ne
+      &&& vygeneruje se v \sexprTree
+      &&& ta se převedena na \atTree aby jsme provedli abstraction elimination
+ & dvě možnosti jak uchovávat jedince po vygenerování
+   && \sexprTree
+      &&& currently implemented
+   && \atTree
+      &&& currently not implemented, ale neni problém dodělat
+      &&& DOĎELAT! :)
+   && ale de mezi nima volně převádět
+      &&& díky tomu že nemaj lambda abstrakce, jinak by museli bejt ve 
+          WHNF/HNF nebo něco takovýho
+   && dusledky pro křížení
+      &&& \atTree umožňuje víc křížení než \sexprTree
+          &&&& každý co de v \sexprTree umí i \atTree ale obráceně to neplatí
+      &&& \atTree je zase ekvivalentní tomu kozovskýmu  
+ & říct že bilding blocks je prostě kontext
+\end{todo}
+
 \section{Generating individuals}
 
-\begin{todo}
-& Dřív se to menovalo 
-"Our approach to \textit{solving} Inhabitation tree"
-tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
-termy a že je  budeme generovat pomocí 
-řešení inhabitačních stromu.
-\end{todo}
-
+%\begin{todo}
+%& Dřív se to menovalo 
+%"Our approach to \textit{solving} Inhabitation tree"
+%tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
+%termy a že je  budeme generovat pomocí 
+%řešení inhabitačních stromu.
+%\end{todo}
+
+Input for term generating algorithm is pair $(\sigma,\Gamma)$ where:\\
+
+\begin{enum}
+ & $\sigma$ is desired type of each generated individual.
+ & $\Gamma$ is \textit{building blocks} context. 
+\end{enum}~
+
+Each generated \lterms $M$ should  satisfy \GMS. \\
+
+We generate terms in long normal form.
+This is achieved by \textit{solving} inhabitation tree for $(\sigma,\Gamma)$.  
 The main design choice behind our approach to solving inhabitation trees
 is ability to be variable enough to enable choice between 
-\textit{systematic} way of generating terms 
-and \textit{classical-GP-like} way of generating terms 
+\textit{systematic}  
+and \textit{ramped half-and-half} (described in \ref{GPgene}) 
+methods of generating terms 
 by choice of simple search strategy.\\
 
-\Lets explain what is meant by \textit{systematic} and \textit{classical-GP-like}. \\
-
-\textit{Systematic} way of solving inhabitation tree corresponds to
-generating terms in order from smallest to largest in number of symbols and \lheads.
+\Lets explain what is meant by \textit{systematic} 
+method of generating terms.
+
+Systematic way of solving inhabitation tree corresponds to
+generating terms in their long normal forms
+in order from smallest to largest in number of symbols and \lheads.
 That means minimizing number of and-nodes in and-or-tree view of point on inhabitation tree.
@@
 and-nodes (symbols and \lheads) and or-nodes (types) it is also minimizing number of 
-size of whole inhabitation/and-or tree. \\
-
-\textit{Classical-GP-like}
-
-\begin{todo}
-!!!
+size of whole inhabitation/and-or tree.
+
+\begin{todo}
+ & Ten poslední odstavec je napsanej divně, přepsat.
+   && říct v něm že na to budem teda používat A*. 
 \end{todo}
@@
 
+\subsection{Other tried approach to generating individuals}
+
+\begin{todo}
+ & popsat tu ten puvodní přístup, kde se bralo přímo 
+   zavedení a eliminace pro každej typovej konstrukt 
+   a podle toho se pak udělali příslušný term generating grammar pravidla
+   kerý odpovídaj manipulacímse strommem
+ & ale má nevýhody
+   && hlavně to neni v normální formě, takže generujeme spoustu 
+      ruzně dopočítanejch beta-rovnáse-ekvivalentních termu
+   && u pravidla zavedení implikace neni uplně zřejmý
+      co tam použít za typ
+   && takže sice to na to ze začátku běhalo ale nakonec to bylo vyměněno
+      za IM metodu since ta je jednoduší, líp se přemejšlí o ní,
+      konstrukty jako and a or nemusíme zavádět na urovni pravidel
+      ale stačí tam dat vhodný polymorfní konstruktory/destruktory 
+ & nějak tu metodu pomenovat a přemenovat podle toho nadpis
+\end{todo}
+
 \section{Crossover}
@@
 	   prozkoumáno.
+& To ale neznamená že jí nemáme radi, naopak se těšíme
+  že její vhodný zavedení celou metodu ještě zlepší.
 \end{todo}
@@
 \subsection{Using term generation}
+
+\section{ Complexity discussion }
+
+\begin{todo}
+& !!!
+& zmínit že mi moc neřešíme formální/asymptotický složitosti
+  && de nám hlavně o realný časy
+& vyjít z komplexnosti A* asi
+  && složitost výrazně ovlivněná kvalitou heuristický funkce
+\end{todo}
 
@@
 \section{[Chytřejší heuristika pro A* která si to předpočítá na $\overline{\Gamma}$ ]}
+\section{[šlechtění těch strategií na generování termů]}
+\section{[šlechtění fitness funkcí a nápad s "turnajem olympioniků"]}
 

### 2013-07-11 tex/thesis.tex
[tex] + IM3 test s Nothing v kozovi...

@@
 		\Fn{#1}{#2}
+	\end{algorithm}
+	\end{framed}
+}
+
+\newcommand{\pseudo}[1]{
+	\begin{framed}
+	\begin{algorithm}[H]
+		\DontPrintSemicolon
+		#1
 	\end{algorithm}
@@
 \begin{align*}
-\FV(x)          &= \{x\} \\
-\FV((P~Q))      &= \FV(P) \cup \FV(Q) \\
-\FV(\lam{x}{P}) &= \FV(P) \smallsetminus \{x\}
+\FV(x)          &= \{\}                        &\textbf{if } x \in C  \\
+\FV(x)          &= \{x\}                       &\textbf{if } x \in V  \\
+\FV((P~Q))      &= \FV(P) \cup \FV(Q)          &          \\
+\FV(\lam{x}{P}) &= \FV(P) \smallsetminus \{x\} &
 \end{align*}
@@
 \begin{todo}
+& Mužem nás zajímat buď cesta nebo jen cílovej stav, nás bude zajímat
+  cílovej stav - proto nebudem zaznamenávat cestu.
 & víc to tu rozepsat hlavně zmínit heuristiku 
@@
 
-			\If{ $( next \not\in open \wedge next \not\in closed ) \vee newG < next.$G  }{\;
+			\If{ $( next \not\in open \wedge next \not\in closed ) \vee newG < next.$G}{\;
 
@@
 
-	\Return no-path-exists 
+	\Return no-reachable-goal 
+}
+
+\begin{todo}
+& náš stavovej prostor je strom takže 
+  && vime že $next \not\in open \wedge next \not\in closed$
+  && a tedy ani nepotřebujeme udržovat množinu closed
+& dist(state,next) = 1 
+\end{todo}
+
+... thus ...
+
+\Pseudokod{Our-A*( start , isGoal , heuristic )}{
+
+	$open \la$ empty priority queue \;\;
+
+	$start$.G = 0 \;
+	$start$.F = $heuristic$( $start$ ) \;\;
+
+	$open$.insert( $start$ )\;\;
+
+	\While{$\neg$ $open$.isEmpty() }{
+
+		$state \la$ $open$.popWithLowestF()\;\;
+
+		\If{ isGoal( $state$ ) }{
+			\Return $state$ \;
+		}\;
+
+		closed.insert( $state$ ) \;\;
+
+		\For{$next \in$ $state$.nexts() }{
+			$next$.G $\la $state$.G + 1$ \;
+			$next$.F $\la $next$.G + heuristic$( $next$ )  \;
+			$open$.insert( $next$ )\;			
+		}
+	}\;
+
+	\Return no-reachable-goal 
 }

### 2013-07-11 v3/IM3.hs

@@ -804,6 +804,17 @@ testSO2 soFun = do
   mapM_ (putStrLn . show) ctts
   putStrLn $ "num terms : " ++ (show $ length ctts)
 
+testSO3 :: (StdGen->SearchOptions) -> IO ()
+testSO3 soFun = do
+  let stdGen = mkStdGen 424242
+  let ctts = prove (soFun stdGen)
+  mapM_ (putStrLn . show) ctts
+  putStrLn $ "num terms : " ++ (show $ length ctts)
+
+
+
+
+
 test_ep = testSO2 $ \ stdGen -> SearchOptions {
     so_n                  = 500            ,
     so_typ                = l_bool :-> bool ,
@@ -913,6 +924,28 @@ test_ssr_koza = testSO2 $ \ stdGen -> SearchOptions {
     so_ctx                = ctx_ttSSR  ,
     so_stdGen             = stdGen    ,
     so_runLen             = Just 1    ,  --Nothing   ,
+    so_isPolymorphic      = False,
+    so_randomRunState     = KozaRandomRunState Nothing Nothing ,
+    so_edgeSelectionModel = KozaESM
+  } 
+
+test_ssr_koza_1 = testSO3 $ \ stdGen -> SearchOptions {
+    so_n                  = 500       ,
+    so_typ                = dou1 ,
+    so_ctx                = ctx_ttSSR  ,
+    so_stdGen             = stdGen    ,
+    so_runLen             = Just 1    ,  --Nothing   ,
+    so_isPolymorphic      = False,
+    so_randomRunState     = KozaRandomRunState Nothing Nothing ,
+    so_edgeSelectionModel = KozaESM
+  } 
+
+test_ssr_koza_2 = testSO3 $ \ stdGen -> SearchOptions {
+    so_n                  = 500       ,
+    so_typ                = dou1 ,
+    so_ctx                = ctx_ttSSR  ,
+    so_stdGen             = stdGen    ,
+    so_runLen             = Nothing   ,
     so_isPolymorphic      = False,
     so_randomRunState     = KozaRandomRunState Nothing Nothing ,
     so_edgeSelectionModel = KozaESM

### 2013-07-12 tex/thesis.tex
[tex]

@@
 \usepackage{qtree}
+
+\usepackage{color}
 
@@
 { ~\\[0.5em]
-  \textbf{TODO}
+  {\color{red}\textbf{TODO}}
   \begin{easylist}[itemize]}
@@
 in order from smallest to largest in number of symbols and \lheads.
-That means minimizing number of and-nodes in and-or-tree view of point on inhabitation tree.
-But since inhabitation tree is a special kind of and-or-tree with alternating layers of 
-and-nodes (symbols and \lheads) and or-nodes (types) it is also minimizing number of 
-size of whole inhabitation/and-or tree.
-
-\begin{todo}
- & Ten poslední odstavec je napsanej divně, přepsat.
-   && říct v něm že na to budem teda používat A*. 
-\end{todo}
+
+We will achieve this by using \textit{A* algorithm}.
+
+%That means minimizing number of and-nodes in 
+%and-or-tree view of point on inhabitation tree.
+%But since inhabitation tree is a special kind of 
+%and-or-tree with alternating layers of 
+%and-nodes (symbols and \lheads) and or-nodes 
+%(types) it is also minimizing number of 
+%size of whole inhabitation/and-or tree.
 
@@
 A* is a general informed search algorithm used for finding least-cost 
-path from a given initial state to some goal state in a state space.
+path from a given start state to some goal state in a state space.
 According to \cite{AIAMA} A* is the most widely known form of best-first search.
 
-As A* traverses the state space, it follows a path of the lowest expected total cost.
-It uses a priority queue of alternative unfinished paths.\\
-
-\begin{todo}
-& Mužem nás zajímat buď cesta nebo jen cílovej stav, nás bude zajímat
-  cílovej stav - proto nebudem zaznamenávat cestu.
-& víc to tu rozepsat hlavně zmínit heuristiku 
-a že stačí aby byla optimistická atd..
-\end{todo}
-
-\Pseudokod{A*( start , isGoal , heuristic )}{\;
-
+By state space is meant oriented weighted graph with states as vertices and edges 
+labeled with numbers corresponding to distance or cost. 
+Edge from $s_i$ to $s_j$ means that state 
+$s_j$ is reachable from state $s_i$ in one step with cost $dist(s_i,s_j)$.
+In pseudocode we will refer to set of states reachable from 
+state $s$ in one step as to $s.nexts()$.
+
+Typically we use A* algorithm to find path from one specific start state to another 
+specific goal state. But sometimes  
+we can be interested only in finding a goal state which 
+is specified by some $isGoal$ predicate give to the algorithm as
+input (and the state space may contain many such goal states). 
+This later variant corresponds to our situation, 
+so we will continue by describing this variant.
+
+As A* traverses the state space, it follows a path of the lowest 
+expected total cost. The expectation is based on \textit{heuristic}
+function which for state $s$ predicts distance to the nearest goal state. 
+It uses a priority queue of states as its crucial data structure.
+In this queue the priority of state $s$ is 
+the total expected cost of the path from start state 
+going through $s$ and continuing to the nearest goal state. \\
+
+Here follows A* algorithm written in pseudocode:
+
+\Pseudokod{A*( start , isGoal , heuristic )}{
 	$open \la$ empty priority queue \;
@@
 
-	\While{$\neg$ $open$.isEmpty() }{\;
-
-		$state \la$ $open$.popWithLowestF()\;\;
+	\While{$\neg$ $open$.isEmpty() }{
+		$state \la$ $open$.popStateWithLowestF()\;\;
 
@@
 
-		\For{$next \in$ $state$.nexts() }{\;
-
+		\For{$next \in$ $state$.nexts() }{		
 		    $newG \la$ $state$.G + dist($state$,$next$)\;\;
@@
 
-\begin{todo}
-& náš stavovej prostor je strom takže 
-  && vime že $next \not\in open \wedge next \not\in closed$
-  && a tedy ani nepotřebujeme udržovat množinu closed
-& dist(state,next) = 1 
-\end{todo}
-
-... thus ...
-
-\Pseudokod{Our-A*( start , isGoal , heuristic )}{
-
-	$open \la$ empty priority queue \;\;
+Now we will modify this general A* algorithm so
+it will be more suited for our needs. We want following 
+properties:\\
+
+\begin{enum}
+& The state space is a tree.
+& $dist(s_i,s_j) = 1$
+& We do not want to find one, but $n$ goal states. 
+\end{enum}~
+
+For state space which is a tree the condition
+$(next \not\in open \wedge next \not\in closed)$ 
+is always true, since it is possible
+to come to a state only from its parent state. 
+This fact results in simplification of algorithm's code and behavior;
+we may completely omit the $closed$ set and the checks in the for loop.\\
+
+Those changes result in the following code:
+
+\Pseudokod{Our-A*( start , n , isGoal , heuristic )}{
+
+	$open    \la$ empty priority queue \;
+	$results \la$ empty set\;\;
 
@@
 
-		$state \la$ $open$.popWithLowestF()\;\;
+		$state \la$ $open$.popStateWithLowestF()\;\;
 
 		\If{ isGoal( $state$ ) }{
-			\Return $state$ \;
+
+			$results$.insert( $state$ )\;
+
+			\If{ $\vert results \vert = n$ }{
+				\Return $results$ \;
+			}
 		}\;
-
-		closed.insert( $state$ ) \;\;
 
@@
 
-	\Return no-reachable-goal 
+	\Return results
 }
 
-\begin{todo}
-& čeknout jak je to v kódu jestli ten A* co tam je čekuje ty
- nódy co už navštívil při přidávání - VŽDYŤ díky tomu žeje to tree-search 
- (vs. graph search) to neni potřeba - pořádně se podívat do AIAMA jakj
- je terminologie a napsat to sem.
-\end{todo}
+The most important part of A* algorithm is the heuristic function.
+A heuristic function must be \textit{admissible} in order to make
+A* algorithm find the shortest possible path.
+
+A heuristic function $h$ is said to be \textit{admissible} if it satisfies 
+following condition for every state $s$:
+$$ h(s) \leq h^{*}(s) $$
+
+where $h^{*}(s)$ is the true length of the shortest path from $s$ to 
+the nearest goal state.
+
+In other words, a heuristic function is \textit{admissible} if it is
+\textit{optimistic}.
 
@@
 extended version of our tree representation for typed \lterms from \ref{tree-reps}.
-This extension consists in adding a new type of leaf node $(\sigma,\Gamma)$ 
+This extension consists in adding a new kind of leaf node $(\sigma,\Gamma)$ 
 standing for unfinished tree of a type $\sigma$ with context $\Gamma$ to use
@@
 
-For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.
+For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet used anywhere in the whole tree.
+%included in context $\Gamma$ as variable or constant name.
 
@@
 	]   
-   ]\\
+   ]\\[3em]
+
+We should discuss that it is correct to have as successors of a state those
+created by expansion of the leftmost unfinished leaf (in contrast with
+more broad successor set where any unfinished leaf may be expanded).
+What we need to show is that no final state will be omitted by the method 
+and that there is no shorter path omitted.
+
+One can see that it is correct by considering that 
+the order of expanding unfinished leafs is irrelevant 
+since expansion of each unfinished leaf is 
+independent from expansion of any other unfinished leaf
+(up to renaming of variables which is irrelevant).
+
+Every unfinished leaf of a state must be replaced by 
+tree containing no unfinished leafs in order to be final state. 
+But since expansion of an unfinished leaf has no impact 
+(besides new variable names) on expansion of other unfinished
+leafs we may choose the order of expansions arbitrarily. \\
+
+We prefer this variant because:
+ \begin{enum}
+ 	& It has smaller or equal number successors for each state, 
+ 	  i.e. smaller or equal branching factor.
+ 	& It makes the state space a tree, in contrast with
+ 	  the naive state space which is not a tree. 
+ \end{enum}
 
 \begin{todo}
- & Discuss that it is OK to expand only the DFS-found node. 
-The order of unfinished nodes is irrelevant since generation of each branch is 
-independent on generating other branches. Mužem na to koukat jako na takovou tu redukci simetrií, najít to v AIAMA a pořádně popsat.
+Mužem na to koukat jako na takovou tu redukci simetrií, 
+najít to v AIAMA a pořádně popsat.
+\end{todo}
+
+\subsection{Our heuristic function}
+
+An heuristic function takes as input a state and returns as output an estimation of number
+of steps needed to get to a final state by the shortest possible path.
+
+Our heuristic function is a simple one: The number of unfinished leafs in the state tree.
+
+\begin{todo}
+& rozebrat to, ukázat že je optimistická atd...
+\end{todo}
+
+\subsection{Search strategy}
+
+\begin{todo}
+ & Internal state
+ & Edge selection model
+
+\end{todo}
+
+\begin{todo}
+
+& sjednotit terminologii nexts successors atd
 
@@
 chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
-\end{todo}
-
-\subsection{Our heuristic function}
-
-An heuristic function takes as input a state and returns as output an estimation of number
-of steps needed to get to a final state by the shortest possible path.
-
-Our heuristic function is a simple one: The number of unfinished leafs in the state tree.
-
-\begin{todo}
- & rozebrat to, ukázat že je optimistická atd...
+
 \end{todo}

### 2013-07-13 tex/thesis.tex
[tex]

@@
 
-\chapter{Definitions}
-
-	\Lets first say some basic definitions.
-
-\section{Genetic Programming}
+\chapter{Genetic Programming}
 \label{GP}
@@
 
-\subsection{Program trees}
+\section{Program trees}
 
@@
 
-\subsection{Building blocks}
+\section{Building blocks}
 \label{building-blocks}
@@
 
-\subsection{Generating individual}
+\section{Generating individual}
 \label{GPgene}
@@
 
-\subsection{Selection}
+\section{Selection}
 
@@
 
-\subsection{Crossover}
+\section{Crossover}
 
@@
 
-\subsection{Reproduction}
+\section{Reproduction}
 
@@
 
-\subsection{Mutation}
+\section{Mutation}
 
@@
 
-\subsection{Construction of a next population}
+\section{Construction of a next population}
 
@@
 \newpage
-\subsection{GP algorithm in pseudocode}
+\section{GP algorithm in pseudocode}
 
@@
 
-\newpage		
+\chapter{Mathematical background}
+
+\begin{todo}
+ & Shrnout tu nějak ve skratce tuhle kapitolu.. 
+\end{todo}
+
 \section{Lambda term}
@@
 
-\chapter{ Our approach }
+\chapter{ Design of our system }
 
@@
 
-\begin{todo}
-Mužem na to koukat jako na takovou tu redukci simetrií, 
-najít to v AIAMA a pořádně popsat.
-\end{todo}
+%\begin{todo}
+%Mužem na to koukat jako na takovou tu redukci simetrií, 
+%najít to v AIAMA a pořádně popsat.
+%\end{todo}
 
@@
 
+Now we will describe further modification of the A* algorithm 
+which is generalizing it enough to enable use of \textit{ramped half-and-half}
+term generating method and other term generating methods.
+
+\begin{enum}
+	& Co se přidalo 
+	  && strategy
+	     &&& selectedNexts
+	  && outer while loop 	
+\end{enum}
+
+\Pseudokod{A*-with-strategy( start , n , isGoal , heuristic , strategy )}{
+
+	$results \la$ empty set\;\;
+
+	\While{ $\vert results \vert < n$ }{
+
+		$open    \la$ empty priority queue \;
+		$strategy.init()$\;\;
+
+		$start$.G = 0 \;
+		$start$.F = $heuristic$( $start$ ) \;\;
+
+		$open$.insert( $start$ )\;\;
+
+		\While{$\neg$ $open$.isEmpty() }{
+
+			$state \la$ $open$.popStateWithLowestF()\;\;
+
+			\If{ isGoal( $state$ ) }{
+
+				$results$.insert( $state$ )\;
+
+				\If{ $\vert results \vert = n$ }{
+					\Return $results$ \;
+				}
+			}\;
+
+			$nexts \la strategy.filter$( $state$.nexts() , $state$.depth() ) \;\;
+
+			\For{$next \in nexts$ }{
+				$next$.G $\la $state$.G + 1$ \;
+				$next$.F $\la $next$.G + heuristic$( $next$ )  \;
+				$open$.insert( $next$ )\;			
+			}
+		}\;
+
+	}
+
+	\Return results
+}
+
 \begin{todo}
@@
  & Edge selection model
-
 \end{todo}

### 2013-07-14 tex/thesis.tex
[tex]

@@
 % !!! místo we can atd psát one can ...
+% ujednotit i.e. , ty čaky kolem toho atd
+
+% myslim že používam málo čárek, přečíst to a vygooglit kde mi příde že by mohla bejt
+
+% ujednotit kurzívu a normální písmo v pseudocodu
 
@@
 
+\begin{todo}
+& koza tomu řiká closure, přečíst si tu pasáž (str 81) a 
+upravit podle toho tady tu terminologii
+\end{todo}
+
 Satisfying this constrain ensures that every program tree build 
@@
 
-\section{Generating individual}
+\section{Generating individuals}
 \label{GPgene}
@@
 one of those two sub-methods is randomly (uniformly) 
-selected and the selected one is performed.\\
+selected and the selected one is performed.
+This selection is done for each generated tree, so there
+is 50\% chance to be generated by \textit{full} 
+and 50\% chance to be generated by \textit{grow}.\\
+
+A maximum depth $D_{initial}$ is defined (e.g. 6).
+
+For each tree is selected its 
+depth $d$ from $\setDots{2}{D_{initial}}$ randomly (uniformly).\\
 
@@
 
-A maximum depth $D_{initial}$ is defined, i.e. 6.
-
-In nodes with \textit{depth} $\leq D_{initial}$ the \textit{full}
+In nodes with \textit{depth} $\leq d$ the \textit{full}
 method selects from the set $F$, whereas the \textit{grow} method 
@@
 
-In nodes with \textit{depth} $= D_{initial}$ both methods select from the
+In nodes with \textit{depth} $= d$ both methods select from the
 set $T$.
@@
 	 řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
+   & říct že se vždycky zastavěj (jak se to správně jmenuje?) ale
+     asi to říct na jinym místě než tady, tady se teprv definujou typy, 
+     říct to někde kde už se ví jak sou napojený na termy    
    & možná se hlobějc zmínit o curryingu  
@@
 
-An heuristic function takes as input a state and returns as output an estimation of number
+A heuristic function takes as input a state and returns as output an estimation of number
 of steps needed to get to a final state by the shortest possible path.
@@
 which is generalizing it enough to enable use of \textit{ramped half-and-half}
-term generating method and other term generating methods.
-
-\begin{enum}
-	& Co se přidalo 
-	  && strategy
-	     &&& selectedNexts
-	  && outer while loop 	
-\end{enum}
+term generating method and other term generating methods.\\
+
+It is done by adding one more input argument for the algorithm.
+This additional argument is a search strategy whose purpose is
+to determine which successor states of all possible successor states 
+will be added to the priority queue, i.e., a search strategy 
+is there to filter out some successor states from $state.nexts()$.
+This filtering may be non-deterministic. It is based on
+the depth of expanded unfinished leaf and on properties of 
+the newly added subtree, e.g., whether the newly added subtree
+contains one or more unfinished leafs (so it corresponds to non-terminal) 
+or not (so it corresponds to terminal).
+
+By \textit{depth} we really mean $\lfloor depth/2 \rfloor$, because our 
+formal state trees contain those type nodes above each symbol node. 
+\\
+
+In previous situation where there is no filtering of states involved 
+we know that if queue becomes empty, then we have exhaustively searched
+all possible states.
+
+But since filtering is involved this is no longer true.
+Therefore, if the queue becomes empty before the required number of
+terms is generated, then the generating process is restarted and 
+it continues to generate terms.
+
+We also enable strategy to have some internal state which is 
+initialized (possibly non-deterministically) each time the 
+generating process restarts. \\
+
+Hopefully those changes will be clarified by the following pseudocode:
 
@@
 
-		$open    \la$ empty priority queue \;
 		$strategy.init()$\;\;
+
+		$open    \la$ empty priority queue \;\;
 
@@
 			}
+		}
 	}\;
 
-	}
-
 	\Return results
@@
 
+\Lets show examples of search strategies. 
+They are described by describing 
+their $filter(nexts,depth)$ 
+and $init()$ methods. Again, in order to avoid confusion, 
+\lets stress that $depth$ is really $\lfloor depth/2 \rfloor$ since
+our formal states contain above each symbol node a type node.
+
+\subsubsection{Systematic strategy}
+
+The previous systematic behavior may be obtained by trivial 
+$filter$ function which returns all $nexts$
+and $init$ is doing nothing.
+
+\subsubsection{Ramped half-and-half}
+
+In the $init$ are randomly (uniformly) initialized two variables of the
+strategy internal state:
+
+\begin{enum}
+ & \textit{isFull} - A boolean value, determining whether \textit{full}
+                     or \textit{grow} method will be performed.
+ & \textit{d} - A integer value from $\setDots{2}{D_{init}}$, where 
+                $D_{init}$ is predefined maximal depth (e.g. 6).                    
+\end{enum}~
+
+The method $filter(nexts,depth)$ returns precisely one randomly
+(uniformly) selected  element from a subset of $nexts$
+(or zero elements if $nexts$ is empty). 
+This means that the queue always contains only one (or zero) state.
+
+The subset to select from is determined by $depth, d$ and $isFull$.\\
+
+Those elements of $nexts$ whose newly added subtree contains one ore more 
+unfinished leafs are regarded as \textit{non-terminals}, whereas 
+those whose newly added subtree contains no unfinished leaf are regarded as 
+\textit{terminals}.\\
+
+If $depth = 0$, then the subset to select from is  
+set of all \textit{non-terminals} of $nexts$.
+
+If $depth = d$, then the subset to select from is
+set of all \textit{terminals} of $nexts$.\\
+
+In other cases of $depth$ it depends on value of $isFull$.
+
+If $isFull = true$, then the subset to select from is 
+set of all \textit{non-terminals} of $nexts$.
+
+If $isFull = false$, then the subset to select from is 
+whole $nexts$ set.
+
 \begin{todo}
- & Internal state
- & Edge selection model
+& eště se musí nějak prohrotit to, že tam je navíc ta lambda hlava
+kerá ještě o něco kazí tu $depth$
+& ukázat že je to fakt to samý co kozovský klasický
 \end{todo}
@@
 chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
+
+& nekde zmínit že s těma nexts nezacházíme takle dementně, že samozřejmě vybíráme 
+jen z těch přidávanejch podstromu ale že tim nechcem komplikovat popis
 

### 2013-07-15 tex/thesis.tex
[tex]

@@
 % ujednotit kurzívu a normální písmo v pseudocodu
+
+% zčekovat: za any musí bejt plural, nejde any term ale musí bejt any terms atd
+%           pro singular je a term.
 
@@
   systém a blíží se to dokazovači.
+   && Ve světle toho, je pak hezky vidět, že
+      je chytrý exhaustivně systematicky prohledávat od nejmenšího
+      po největší- jak budeme zesložiťovat typovej systém, tak
+      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
+      jich generovat kvanta
 & Nějak naťuknout už tady v úvodu jak se to má při
@@
 The previous systematic behavior may be obtained by trivial 
-$filter$ function which returns all $nexts$
-and $init$ is doing nothing.
-
-\subsubsection{Ramped half-and-half}
+$filter$ function which returns all $nexts$.
+Systematic strategy needs no internal state, 
+therefore $init$ is doing nothing.
+
+\subsubsection{Ramped half-and-half strategy}
 
@@
 & ukázat že je to fakt to samý co kozovský klasický
+& zavýst tam pojem běhu pro ten jeden vnější while cyklus
 \end{todo}
+
+\subsubsection{Geometric strategy}
+
+We can see those two previous strategies as two extremes on the spectrum of 
+possible strategies.
+
+\textit{Systematic strategy} filters no successor state thus performing
+exhaustive search resulting in discovery of $n$ smallest terms in one run.
+
+On the other hand, \textit{ramped half-and-half strategy} filters 
+all but one successor states resulting in degradation of 
+priority queue to "fancy variable". 
+And each term is generated in its own run.\\
+
+\textit{Geometric strategy} is simple yet fairly effective term generating 
+strategy somewhere in the middle of this spectrum.
+
+It is parameterized by parametr $q \in (0,1)$.
+
+For each element of $nexts$ it is probabilistically decided whether
+it will be returned or omitted. A probability $p$ of returning is
+same for all elements, but depends on $depth$. 
+It is computed as follows:
+
+$$ p = q^{depth} $$
+
+This formula is motivated by idea that it is important to
+explore all possible root symbols, but as the $depth$ 
+increases it becomes less "dangerous" to omit 
+an exploration branch.
+
+We can see this by considering that this strategy results in
+somehow undisciplined A* search.
+With each omission we make the search space smaller. But with
+increasing depth these omissions have smaller impact on the search space,
+i.e., they cut out lesser portion of the search space.
+
+Another slightly esoteric argument supporting this formula is that "root 
+parts" of a program usually stand for more crucial parts
+with radical impact on global behavior of a program, 
+whereas "leaf parts" of a program usually
+stand for less important local parts (e.g. constants).
+
+It also plays nicely with the idea that "too big trees should be killed".\\
+
+Systematic strategy does not have an internal state, thus $init$ does nothing.
 
@@
   že její vhodný zavedení celou metodu ještě zlepší.
+
+& Nápad na to jak to dělat:
+  && nutno ošetřit - co všechno do gammy alemyslim že v phodě když je uplne 
+     stejná jako ta generujicí, pač ve vy SKI-ovanym termu stejne nejsou zadny 
+     proměný 
+  && mít nějakej předem danej počet pohybu na frontě, abych mohl efektivně
+     kontrolovat že to přidání mutací moc nespomalý
+  && Po tom počtu pohybu končim generování
+  && mam nějakej počet vygenerovanejch termu
+  && pro chytrou mutaci mužu chtít třeba 10 termu
+     &&& pokud se mi jich vygenerovalo míň, tak beru míň
+     &&& pokud se mi jich vygenerovalo víc, tak vemu 10 náhodně
+     &&& těch 10 vyhodnotim (jakože dosazenejch na jejich místo)
+     &&& nejlepší si vezmu
+     &&& dobrý udělat to zase najednou pro všechny v tý generaci mutovaný
+         pánčto je nákladný spouštět Hint
+& Další mutace - beta eta redukce (?? asi moc naroční viz dale) 
+  && zmenšující mutace
+  && udělaná že převedem SKIBC do lambda termu a to redukujem? to asi ne pač
+     pak převod to spíš prodlouží, spíš asi mít redukční pravidla pro SKIBC
+     a těma redukovat, ale je otázka jak moc tam budou vznikat SKIBC-redexi
+     - ty mohou vznikat jen při @-tree reprezentaci
+  && to otevírá otázku toho, že by elementy gammy mohla mít k sobě redukční pravidla
+& Další mutace - jednoduše náhrada podstromu/uzlu za nějakej z gammy
+  && nezvětšující mutace
+& Další mutace - mutace číselnejch-like konstant
+& To mě přivádí k myšlence sice pro generování ale myslim podstatný:
+  jakmile jsou v problému číselný-like konstanty tak mužu udělat pro 
+  jeden tvar hodně jedincu - což šetří jak čas na generování, tak to prozkomává 
+  víc danej tvar
+
 \end{todo}
@@
 \section{[šlechtění fitness funkcí a nápad s "turnajem olympioniků"]}
+\section{[šlechtění search strategií pro prohledávání]}
 

### 2013-07-16 tex/thesis.tex
[tex]

@@
 % !!! místo we can atd psát one can ...
+%      - ale ted to vypada že to neni tak žhavý pravidlo
+
 % ujednotit i.e. , ty čaky kolem toho atd
@@
 \section{Crossover}
+\label{GPxover}
 
@@
 
-In GP the most common mutation is \textit{Subtree swapping mutation}.
+In GP the most common crossover is \textit{Subtree swapping crossover}.
 This crossover randomly selects one node in each parent tree.
@@
 \begin{todo}
+
+& ujistit se že je v části o tree reprezentacích 
+  je napsaný co to znamená že podstrom má nějakej typ,
+  že podstromy odpovídaj podvýrazum atd
+
  & říct že to je přesně ta notace co používá koza
@@
 
+\newpage
 \section{Crossover}
 
-\begin{todo}
-& Prostě budem křížit jen podstromy stejnýho typu
-& ALE! Hlavní problém jsou proměnný
-& ve stručnosti kouknout na dvě hlavní metody
-  && křížit prostě lambda stromy a vzniklý problémy řešit online 
-  && zbavit se proměnejch 
-& říct že nakonec sme přilnuli k zbavit se proměnejch
-\end{todo}
-
-\subsection{Finding same types}
-\subsection{Two basic options}
-Resolve problems with free variables or avoid variables 
-completely.
+The crossover operation in classical GP is performed 
+by swapping randomly selected subtrees in each parent 
+\sexprTree (see \ref{GPxover}).
+
+For typed lambda terms two difficulties arise: Types and variables.
+
+We will show how to crossover typed \lterm trees in both 
+\atTree and \sexprTree notation.\\
+
+As in classical GP our crossover will be performed by swapping
+two subtrees. But now with constraint that both subtrees have
+the same type. This is discussed in grater detail in \ref{typed-swapping}.\\
+
+Variables bring more difficulties then types do.
+This problem arises from variables that are free in subterms corresponding 
+to swapped subtrees.
+
+Following example illustrates the problem. \Lets have these two
+parent trees with selected nodes in bold.\\
+
+\newcommand{\lh}[1]{\lambda #1}
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]
+
+~\\The swap of subtrees results in following trees:\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]
+
+~\\The problem is that variable $x_2$ in second tree
+is not bound by any $\lambda$-head and since
+it is not element of $\Gamma$, the second tree is not well-typed \lterm.
+
+% - This is only one example of a problem caused by swapping subtrees with variables,
+% - Resolve problems with free variables or avoid variables completely.
+
+In \ref{repairing-method} is presented more detailed discussion of problems with
+variables and method which repairs such "broken" terms.
+
+But we can avoid dealing with this problem by avoiding use of variables.
+This can be achieved by process called abstraction elimination.
 
@@
 
+\subsection{Typed subtree swapping}
+\label{typed-swapping}
+
+\subsection{ Repairing method }
+\label{repairing-method}
+
+\begin{todo}
+& !!!
+\end{todo}
+
+\subsection{ Comparison of the two methods }
+\label{comarison-ski-repairing}
+
+\begin{todo}
+& !!!
+& říct že nakonec sme přilnuli k zbavit se proměnejch
+\end{todo}
+
+\newpage
 \section{Mutation}
@@
 
+\chapter{Polymorphic variant}
+
+\begin{todo}
+ & poznamky o čem se mimojiné zmínit:
+   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
+      jako alternativa k ADF ! Myslim že hodně zajimavý
+\end{todo}
+
 \chapter{ Implementation }

### 2013-07-16 tex/thesis.tex
[tex] vyřešen problem s neočíslovanejma stránkama
[tex]

@@
 %           pro singular je a term.
+
+% The general rule you can use is “that" is for restricting the meaning of the subject, whereas "which" is for decorating the subject without adding to the meaning. 
+%http://www.wikihow.com/Use-%22That%22-and-%22Which%22-Correctly
 
@@
 
-%% on tam měl %% \openright
+%% on tam měl %% \openright %!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
@@
 
+%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
+%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
+%%% na začátku práce, místo na jejím konci.
+
+%% on tam měl %% \openright %!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+\pagestyle{plain}
+\setcounter{page}{1}
+
 \tableofcontents
@@
 
-Témata co zmínit v úvodu :\\
+Genetic programming (GP) is an AI technique, which falls into 
+broader category of evolutionary algorithms --- 
+mataheuristic search algorithms inspired by 
+biological evolution by natural selection.
+I dare say many people perceive its beauty in the fact that GP is
+a computer program constructing other computer programs 
+with desired properties by breeding them. 
+It is also pretty successful technique in a number of areas. 
+And perhaps that is why GP has become very popular.
+
+~\\[3em]Témata co zmínit v úvodu :\\
 
 \begin{easylist}[itemize]
+
+& Co je to GP, 
+ && že je to část EVA
+ && že v clasickym GP je closure
+ && a ruzný prace ukazujou jak se ho zbavit
+
+& Co je to typ
+  && typ jedobře zaběhlá programátorská konstrukce
+  && matematicky uchopuje typy a programy v teorii typovaného lambda kalkulu
+
+& cílem této práce je 
+  && ukázat jak spolu hezky souvysej
+     &&& např skrzestavební bloky a context
+  && a jak využít konstrukty typový teorie k prospěchu v GP
+     &&& such as AE, beta,etha reduction,...
+  && na základě takto zmapované podobnosti navrhnout systém a implementovat
+     rather then build yet another GP system, generalize ten klasickej
+
 & Začít asi tim že evoluce je cool a že
@@
 
-\textit{Genetic programming} (GP) is a technique inspired by biological evolution
+\textit{Genetic programming} is a technique inspired by biological evolution
 that for a given problem tries to find computer programs able to solve that problem. 
@@
 
+\section{Brief history of the project}
+\begin{todo}
+ & zmínit třeba tyhle fáze:
+  && že sem si delal evaluator sám co sem pak s performenc duvodu opustil
+  && že sem hledal intuicionistickej dokazovač ale že nejsem nenarazil na vhodnej
+     pač nejsou delany na to plivat hodně dukazu  
+  && že sem nejdriv pouzival ten vypsychlej dokazovač co tam měl nějaký ty 
+     spešl rozhodovací stromy
+  && pak sem to přepsal tak aby to nebylo jen na GP šitý ale obecně na jakoukoliv 
+     evoluci
+  && ...
+\end{todo}
+
 \chapter{Problems}

### 2013-07-17 tex/thesis.tex
[tex]

@@
 %http://www.wikihow.com/Use-%22That%22-and-%22Which%22-Correctly
+
+% možná místo classical GP řikat standard GP jako montana95
 
@@
 
+%	Seznamte se s typovanými lambda kalkuly [1] 
+%	a s jejich praktickou implementací [5]. 
+%	Prostudujte standardní techniky genetického 
+%	programování [2-4]. Navrhněte a implementujte 
+%	prototyp systému řešící úlohu genetického 
+%	programování nad typovaným funkcionálním 
+%	jazykem. Porovnejte dosažené výsledky 
+%	v závislosti na zvolených vstupních parametrech.
+
 Genetic programming (GP) is an AI technique, which falls into 
@@
 a computer program constructing other computer programs 
+%(property sometimes called \textit{automatic programming}) 
 with desired properties by breeding them. 
@@
 And perhaps that is why GP has become very popular.
+
+GP constructs the population of solution programs from 
+two sets of building blocks; set F of functions 
+and set T of constants and variables. 
+In classical GP \textit{type consistency} of those 
+building blocks is required, property also called \textit{closure}.
+This means that all functions must accept
 

### 2013-07-17 tex/thesis.tex
[tex]

@@
 \noindent
-Dedication.
+$<$dedication$>$
 
@@
 
-Genetic programming (GP) is an AI technique, which falls into 
-broader category of evolutionary algorithms --- 
-mataheuristic search algorithms inspired by 
+Genetic programming (GP) is an AI technique, which falls into broader category 
+of evolutionary algorithms  ---  mataheuristic  search algorithms inspired  by 
 biological evolution by natural selection.
-I dare say many people perceive its beauty in the fact that GP is
-a computer program constructing other computer programs 
-%(property sometimes called \textit{automatic programming}) 
+I dare say many people perceive its beauty in the  fact that GP is a computer program  constructing other computer programs %(property sometimes called \textit{automatic programming})
 with desired properties by breeding them. 
-It is also pretty successful technique in a number of areas. 
+It is also pretty successful technique in a number of areas \cite{koza05}. 
 And perhaps that is why GP has become very popular.
@@
 GP constructs the population of solution programs from 
-two sets of building blocks; set F of functions 
-and set T of constants and variables. 
-In classical GP \textit{type consistency} of those 
-building blocks is required, property also called \textit{closure}.
-This means that all functions must accept
+two sets of building blocks; set $F$ of functions 
+and set $T$ of constants and variables. 
+In classical GP \textit{closure} of those 
+building blocks is required.
+This means that all functions in $F$ must accept all 
+constants and variables from $T$ as its arguments and produce 
+such values that are also acceptable by all functions in $F$ as arguments.
+Main motivation behind this thesis is to design and implement a system that
+eliminates this constraint. 
+This can be achieved by utilization of \textit{types}.\\
+
+Utilization of types in GP is not a new idea; 
+it was dealt with in various previous works 
+\cite{yu01,montana95,haynes96,olsson94}.
+The contribution of this thesis can be seen 
+in that it deals with this task using concepts from
+theorie of typed $\lambda$-calculus and combinatory logic
+and shows how they can be used for benefit in GP.\\
+
+Notable highlights are: 
+\begin{itemize}
+ \item \textit{Inhabitation tree} (a slight modification 
+   of \textit{inhabitation machine} presented in \cite{barendregt10}
+   suited better for implementation purposes) 
+   is presented and used for \textit{generating of individuals}.
+ \item \textit{Abstraction elimination} is utilized
+   to enable use of simple \textit{mutation operator}.
+ \item \textit{Polymorphic} generalization of inhabitation tree,
+   enabling use of polymorphic functions as building blocks, 
+   is presented.
+
+\end{itemize}
+
+Design of a GP system utilizing these methods is presented.
+This design aims to be generalization of classical GP 
+developed by Koza \cite{koza92}, rather than being 
+specialized variant of GP using more recent techniques.
+This decision is based on belief that modern specialized 
+techniques can be incorporated later and on belief
+that simplicity of design of classical GP plays nicely
+with testing those $\lambda$-calculus concepts in 
+the evolutionary context.
+
+The implementation aims
 
@@
      rather then build yet another GP system, generalize ten klasickej
+
+& zmínit že existuje i jinej pohled na to proč tam zavádět typy:
+  && mužem koukat na to tak že konstrejnt je naše TuF tim, že 
+     v ní nemužou bejt libovolný funkce
+  && ale tuhle constrain mužeme omezit tim že konstrjnujeme 
+     právě použití tech bilding bloku, kde constrains jsou daný těma typama
+     - jakoby výměna global constrain za localní konstrejny na tech bilding blockách
+  && druhá možná ješte silnější motivace je že pomocí typu pak mohou vznikat 
+     smysluplnější programy, tím že dostavají vstupy které přesně čekají
+     a dávají výstupy které mají nějaké očekávané vlastnosti
 
@@
 
+% TODO bacha neni tam jednotnej formát!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+
 \bibitem{koza92}
@@
 
+\bibitem{koza05}
+
+  Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
+  \emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
+  Springer, 2005. ISBN 978-0-387-26417-2
+
+\bibitem{yu01}
+
+  T. Yu. 
+  \emph{Hierachical processing for evolving recursive and modular 
+        programs using higher order functions and lambda abstractions}. 
+  Genetic Programming and Evolvable Machines,
+  2(4):345–380, December 2001. ISSN 1389-2576.
+
+\bibitem{montana95}
+
+D. J. Montana. 
+\emph{Strongly typed genetic programming.} 
+Evolutionary Computation, 3(2): 199–230, 1995.
+%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha
+
+\bibitem{haynes96}
+
+T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
+\emph{Type inheritance in strongly typed genetic programming.} 
+In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
+in Genetic Programming 2, chapter 18, pages 359–376.
+MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1.\\ 
+URL 
+\url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.
+
+\bibitem{olsson94}
+
+J. R. Olsson. 
+\emph{Inductive functional programming using incremental program 
+transformation and Execution of logic programs by 
+iterative-deepening A* SLD-tree search.} 
+Dr scient thesis, University of Oslo, Norway, 1994.
+
 \bibitem{barendregt10}
@@
   2010. \\
-  \url{http://www.cs.ru.nl/~henk/book.pdf}
+  URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
 

### 2013-07-18 tex/thesis.tex
[tex]

@@
 \documentclass[12pt,a4paper]{report}
+
+% dát pryč will, používat přítomnej čas
 
@@
 This means that all functions in $F$ must accept all 
-constants and variables from $T$ as its arguments and produce 
+constants and variables from $T$ as their arguments and produce 
 such values that are also acceptable by all functions in $F$ as arguments.
@@
 in that it deals with this task using concepts from
-theorie of typed $\lambda$-calculus and combinatory logic
-and shows how they can be used for benefit in GP.\\
+theory of typed $\lambda$-calculus and combinatory logic
+and shows how to use them in GP.
+Hopefully, it is also shown that those two fields are
+nicely linked together, e.g., the concept of \textit{building blocks}
+corresponds to concept of \textit{context}. \\
 
@@
    of \textit{inhabitation machine} presented in \cite{barendregt10}
-   suited better for implementation purposes) 
-   is presented and used for \textit{generating of individuals}.
+   suited better for reasoning about implementation details) 
+   is presented and used for \textit{generation of individuals}.
+   This generation process is parameterized by simple 
+   \textit{search strategy}. Several search strategies are presented.   
  \item \textit{Abstraction elimination} is utilized
@@
 specialized variant of GP using more recent techniques.
-This decision is based on belief that modern specialized 
-techniques can be incorporated later and on belief
+This decision is based on belief
 that simplicity of design of classical GP plays nicely
 with testing those $\lambda$-calculus concepts in 
-the evolutionary context.
-
-The implementation aims
-
-~\\[3em]Témata co zmínit v úvodu :\\
-
-\begin{easylist}[itemize]
-
-& Co je to GP, 
- && že je to část EVA
- && že v clasickym GP je closure
- && a ruzný prace ukazujou jak se ho zbavit
-
-& Co je to typ
-  && typ jedobře zaběhlá programátorská konstrukce
-  && matematicky uchopuje typy a programy v teorii typovaného lambda kalkulu
-
-& cílem této práce je 
-  && ukázat jak spolu hezky souvysej
-     &&& např skrzestavební bloky a context
-  && a jak využít konstrukty typový teorie k prospěchu v GP
-     &&& such as AE, beta,etha reduction,...
-  && na základě takto zmapované podobnosti navrhnout systém a implementovat
-     rather then build yet another GP system, generalize ten klasickej
-
-& zmínit že existuje i jinej pohled na to proč tam zavádět typy:
-  && mužem koukat na to tak že konstrejnt je naše TuF tim, že 
-     v ní nemužou bejt libovolný funkce
-  && ale tuhle constrain mužeme omezit tim že konstrjnujeme 
-     právě použití tech bilding bloku, kde constrains jsou daný těma typama
-     - jakoby výměna global constrain za localní konstrejny na tech bilding blockách
-  && druhá možná ješte silnější motivace je že pomocí typu pak mohou vznikat 
-     smysluplnější programy, tím že dostavají vstupy které přesně čekají
-     a dávají výstupy které mají nějaké očekávané vlastnosti
-
-& Začít asi tim že evoluce je cool a že
-  tohle bude o tom jak do GP dát typy.
-& Říct jak jde text:
-  && Začnem popsanim klasickýho GP podle Kozy
-  && Vyslovíme definice okolo lambda termu a typu
-  && Vyslovíme definici inhabitačního stromu
-     kterej je stromovou verzí IM od Barendrechta.
-  && Matematicky popíšem jak na to dem my.
-  && Popíšem rozšíření zakladního systému na 
-     polymorfní.
-  && Popíšem to z programátorskýho hlediska
-  && Popíšem příklady na kterejch to demonstrujem
-  && Srovnáme to s nějakejma přístupama co už 
-     existujou 
-& Možná se zmínit o tý představě dvou pólu
-  induktivním a deduktivním -
-  klasickym GP a Dokazovačem, čim složitější
-  typovej sytem tim víc se jedná o deduktivní   
-  systém a blíží se to dokazovači.
-   && Ve světle toho, je pak hezky vidět, že
-      je chytrý exhaustivně systematicky prohledávat od nejmenšího
-      po největší- jak budeme zesložiťovat typovej systém, tak
-      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
-      jich generovat kvanta
-& Nějak naťuknout už tady v úvodu jak se to má při
-  srovnání s tim jak na to dou jiný (Yu atd)
-& Výkřiky co zmínit:
-  && Na co zaměřený:
-     &&& Proof of a concept
-     &&& Pořádně to vymáchat v typový teorii
-     &&& Spíš než bejt další z řady nástroju
-         se snaží zobecnovat ty už existující
-     &&& Spíš jednoduchost
-         &&&& Ve většině aspektú to sleduje
-              Kozu - tzn i v málo řídících
-              parametru
-     &&& Duraz na generování termu (!)              
-     &&& Po implementační stránce
-         &&&& Sranda s JS překladem
-         &&&& Servrová architektura
-         &&&& oboje spíš srandy
-  && Na co nezaměřený (ale snad taky ok v tom):
-     &&& Performance   
-     &&& Mature knihovna co by šla použít robustně
-         v cizim kódu - spíš to počítá že to 
-         bude dál rozvíjeno a tohle je mezizastávka
-\end{easylist}
+the evolutionary context and on belief that modern specialized 
+techniques can be incorporated later.
+
+%The implementation of our system aims 
+Described system is implemented in purely functional language 
+\textit{Haskell}. Evaluation of individual \lterms is 
+performed by translating them to
+Haskell programs and evaluating them in Haskell interpreter. 
+The core part of the system runs as server, which is 
+managed through user interface accessible via
+web browser. Solution individuals are also translated
+to \textit{JavaScript} expressions, which enables  
+an interactive solution presentation.
+This approach helps us with demonstrating 
+that \lterms can be easily translated to a 
+widely used programming language.
+
+Several example problems for the designed system are presented 
+and examined. Those examples are not intended to 
+bring breakthrough results; they are intended to 
+serve as demonstration of that the system is working properly
+and as demonstration of several directions in which the system
+can be used.
+
+\begin{todo}
+ & možná přidat ještě něco z poznámek co jsou teď v Conclusion
+\end{todo}
 
@@
 that for a given problem tries to find computer programs able to solve that problem. 
-GP was developed by John Koza \cite{koza92} in 1992.
+GP was developed by John Koza in 1992 \cite{koza92}.
 
@@
 It is collection of symbols (accompanied with an information about number of arguments).
-Those symbols are used to construct trees representing individuals.  
-\\\\
+Those symbols are used to construct trees representing individuals.\\
+
 \Lets describe GP algorithm briefly:
@@
 
-This step consists of two sub steps\footnote{TODO : Technically it is done in a little  
-bit different fashion which is equivalent.}:
+This step consists of two sub steps:
 \begin{itemize} 
@@
 
-In GP programs are represented as expressions. \Lets define expression inductively:
-
-\begin{todo}
- & Asi to přejmenvat na S-expression.  
-\end{todo}
-
+In GP programs are represented as \textit{S-expressions}. 
+\Lets define S-expression inductively:
+
+\begin{samepage}
 \begin{itemize}
-	\item Constant\footnote{By constants we also mean procedures with zero  	
-		  arguments.} or variable symbol $s$ is expression.
+	\item Constant or variable symbol $s$ is S-expression.\footnote{
+	      By constants we also mean procedures with zero arguments.}
 	\item Let there be a function with symbol $f$ which has $n$ arguments.
-	      And let there be expressions $e_{1}, ..., e_{n}$. 
-	      Then ( $f$ $e_{1}$ ... $e_{n}$ ) is expression
-	      \footnote{This notation comes from Lisp programming language, 
-	      classical notation would be $f(e_{1}, ... ,e_{n})$. }.   
+
+	      And let there be S-expressions $e_{1}, ..., e_{n}$.
+
+	      Then ( $f$ $e_{1}$ ... $e_{n}$ ) is expression.\footnote{
+	      This notation comes from Lisp programming language, 
+	      a more standard notation would be $f(e_{1}, ... ,e_{n})$.}   
 \end{itemize}
+\end{samepage}
 
@@
 		 		\text{$t_{n}$} ]\\\\
-		 Where $t_{1}, ..., t_{n}$ are trees corresponding to expressions $e_{1}, ..., e_{n}$.	   
+		 Where $t_{1}, ..., t_{n}$ are trees corresponding 
+		 to S-expressions $e_{1}, ..., e_{n}$.	   
 \end{itemize}
@@
 \begin{itemize}
-	\item Terminal set $T$ : Set of symbols used as leaf nodes of 	               
+	\item \textit{Terminal set $T$}: Set of symbols used as leaf nodes of 	               
 	      program trees standing for constants and variables.
-	\item Function set $F$ : Set of symbols used as internal nodes 
+	\item \textit{Function set $F$}: Set of symbols used as internal nodes 
 	      of program trees standing for functions.
@@
 
-Beside symbols in \TuF there must also be 
-an implementation for each symbol which is not variable. 
+For each symbol in \TuF that is not variable 
+there must also be an implementation.
 And for every function symbol from $F$ there must be specified 
@@
 
-\begin{todo}
-& koza tomu řiká closure, přečíst si tu pasáž (str 81) a 
-upravit podle toho tady tu terminologii
-\end{todo}
-
 Satisfying this constrain ensures that every program tree build 
-from \TuF will be total.
-
-We will refer to this constrain as \textit{over-one-type} constrain
-in the further text.\\
-
-We can say that the motivation behind this work is to construct system where we
-eliminate this constraint.
+from \TuF will be total. We will refer to this constrain as to \textit{closure}.\\
+
+We can say that the motivation behind this thesis is to construct system,
+where we eliminate this constraint.
 
@@
 
-We will describe three tree generating method described by Koza 
-in \cite{koza92} called \textit{ramped half-and-half}. \\
+We will describe tree generating method described by Koza 
+called \textit{ramped half-and-half}  \cite{koza92}. \\
 
@@
 There are two generating sub-methods called \textit{full} and 
-\textit{grow} each restricting differently the subset of \TuF. 
+\textit{grow}; each of them restricting differently the subset of \TuF. 
 This restriction depends on depth of the
@@
 This selection is done for each generated tree, so there
-is 50\% chance to be generated by \textit{full} 
-and 50\% chance to be generated by \textit{grow}.\\
-
-A maximum depth $D_{initial}$ is defined (e.g. 6).
+is 50\% chance for tree to be generated by \textit{full} 
+and 50\% chance for tree to be generated by \textit{grow}.\\
+
+A maximum depth $D_{initial}$ is defined (e.g. $D_{initial} = 6$).
 
@@
 
-In \EA there is plenty of options for selection mechanisms 
+In the field of evolutionary algorithms there is plenty of 
+options for selection mechanisms 
 for us to choose from. Again we will describe mechanism Koza
@@
 two (parent) individuals and combines theirs genomes to produce 
-two possibly new (child) individuals.
+two possibly new (offspring) individuals.
 
@@
 Two new child individuals are constructed by swapping subtrees 
-which have roots in those selected nodes.\\
+that have roots in those selected nodes.\\
 
@@
 
-\begin{todo}
- & zmínit maximální hloubku vzniklýho dítěte (17)
-\end{todo}
+A maximum permissible depth $D_{created}$ 
+for offspring individuals is defined (e.g. $D_{created} = 17$).
+If one of the offspring has greater depth than this limit, then 
+this offspring is replaced by the first parent in the result of 
+the crossover operator. If both offspring exceeds this limit, than 
+both are replaced by both parents.
 
@@
 
+Similarly as for crossover, a maximum permissible depth for mutant individual 
+is defined.
+
 \section{Construction of a next population}
 
-Let $pop_{t}$ be the current population which we want to 
+Let $pop_{t}$ be the current population that we want to 
 transform into the next population $pop_{t+1}$. 
@@
 \end{itemize}
+
+\section{Evaluation}
+
+\begin{todo}
+ & !!!
+ & zmínit tu terminate bool
+\end{todo}
 
@@
  & mention best preservation and same popSize checking.
- & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
-   defining EA and then specifying the differences or something 
-   like that...    
- & History, citations, etc ....
+ & popsat co je to popWithF
+ & popsat co je to terminate 
  & ten algoritmus v kozovi dělá "Designate Result" já tam vracim 
@@
 
-\begin{todo}
- & Shrnout tu nějak ve skratce tuhle kapitolu.. 
-\end{todo}
-
-\section{Lambda term}
+In this chapter we briefly describe key concepts
+from simply typed $\lambda$-calculus.
+We try to favor readability over formality,
+for more detailed and formal treatment of this subject see 
+\cite{barendregt92}.
+
+After that, \textit{inhabitation tree} is introduced.
+It is structure used for reasoning about generation of individuals.
+
+\theoremstyle{plain} 
+\newtheorem{theorem}{Theorem} 
+\newtheorem{proposition}{Proposition} 
+\newtheorem{lemma}{Lemma} 
+\newtheorem*{corollary}{Corollary}
+
+\theoremstyle{definition} 
+\newtheorem*{definition}{Definition} 
+\newtheorem{conjecture}{Conjecture}
+ \newtheorem*{example}{Example}
+
+\theoremstyle{remark} 
+\newtheorem*{remark}{Remark} 
+\newtheorem*{note}{Note} 
+\newtheorem{case}{Case}
+
+\section{\lterms}
 \label{deflam}
+
+Now we will formally describe programming language 
+in which we will generate individual programs --- so called \lterms.
 
@@
 
+\begin{definition}
+
 Let $V$ be infinite countable set of {\it 
-variable names}.  \\* 
-Let $C$ be set of {\it constant names}.	 \\*		
+variable names}. Let $C$ be set of {\it constant names}.	 	
 Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
@@
 \textit{~~~~~($\lambda$-abstraction)} 
-\end{align*}
-
-Function application and $\lambda$-abstraction are concepts
+\end{align*}~
+\end{definition}
+
+\textit{Function application} and 
+\textit{$\lambda$-abstraction} are concepts
 well known from common programming languages. 
@@
 $\Icomb = \lamb{x}{x}$.  
-\\
+\\[1em]
 
@@
 
-\begin{todo}
-	& říct co to vubec je ty intutivně (v par větách jen) 
-	  a nějak to napojit na to jak to používám já
-	  pak v tom GP.
-	& Citovat nějakýho bárendrechta jako zdroj podrobnějších informací
-\end{todo}
-
-\section{Type}
+\section{Types}
 \label{deftype}
@@
 
-Let $A$ be set of {\it atomic type names}. \\*
-Then $\mathbb{T}$ is set of {\it types} inductively defined as follows:
-
+A \lterm as described above
+corresponds to a program expression with no type information
+included. Now we will describe \textit{types} (or \textit{type terms}).
+After putting those two pieces 
+(\textit{\lterms} and \textit{types}) together 
+we will get system called \textit{simply typed $\lambda$-calculus}.
+
+\begin{definition}
+Let $A$ be set of {\it atomic type names}. \\
+Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
 \begin{align*}
@@
 \sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
-\end{align*}
+\end{align*}~
+
+\end{definition}
 
@@
 something of a type $\sigma$ and returning 
-as output something of a type $\tau$. \\
+as output something of a type $\tau$.
+
+Usually A is called \textit{set of type variables}, 
+but since it might be source of confusion with polymorphic type variables
+we prefer this name, which also better suits our usage. 
+\\
 
 We use following notation
-abbreviation for better readability:\\
+abbreviation for better readability:
 
@@
 \begin{todo}  
-   & určitě promluvit nějak v úvodu k tomu že se jedná o 
-	 Simply typed lambda kalkulus, zmínit že klasicky se těm alfa z A 
-	 řiká proměný, ale že je to matoucí vzhledem k polymorfizmu
-   & říct že se vždycky zastavěj (jak se to správně jmenuje?) ale
-     asi to říct na jinym místě než tady, tady se teprv definujou typy, 
-     říct to někde kde už se ví jak sou napojený na termy    
    & možná se hlobějc zmínit o curryingu  
@@
 	\end{equation*}
+
+\section{Simply typed $\lambda$-calculus}
+
+\begin{todo}
+   & říct že se vždycky zastavěj (strong normalization) ale
+     asi to říct na jinym místě než tady, tady se teprv definujou typy, 
+     říct to někde kde už se ví jak sou napojený na termy    
+\end{todo}
 
@@
 \begin{enum}
- & It should eliminate the \textit{over-one-type} constrain of classical GP
+ & It should eliminate the \textit{closure} constrain of classical GP
    (see \ref{building-blocks}). 
@@
 
-\subsection{Using term generation}
+%\subsection{Using term generation}
 
@@
 
-\textbf{Poznámky :\\
-bylo by asi mimojiný dobrý 
+\begin{todo}
+ & !!!!!!!!!!!!!!!!!
+\end{todo}
+
+Další témata co zmínit v úvodu nebo v závěru :\\
+
+\begin{easylist}[itemize]
+
+& zmínit že existuje i jinej pohled na to proč tam zavádět typy:
+  && mužem koukat na to tak že konstrejnt je naše TuF tim, že 
+     v ní nemužou bejt libovolný funkce
+  && ale tuhle constrain mužeme omezit tim že konstrjnujeme 
+     právě použití tech bilding bloku, kde constrains jsou daný těma typama
+     - jakoby výměna global constrain za localní konstrejny na tech bilding  
+       blockách
+  && druhá možná ješte silnější motivace je že pomocí typu pak mohou vznikat 
+     smysluplnější programy, tím že dostavají vstupy které přesně čekají
+     a dávají výstupy které mají nějaké očekávané vlastnosti
+
+& Možná se zmínit o tý představě dvou pólu
+  induktivním a deduktivním -
+  klasickym GP a Dokazovačem, čim složitější
+  typovej sytem tim víc se jedná o deduktivní   
+  systém a blíží se to dokazovači.
+   && Ve světle toho, je pak hezky vidět, že
+      je chytrý exhaustivně systematicky prohledávat od nejmenšího
+      po největší- jak budeme zesložiťovat typovej systém, tak
+      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
+      jich generovat kvanta
+
+& že imlementace je delana těma vtipnejamtajpklasama umožnujicima libovolný
+  evoluční struktury na evolvovani
+
+& říct že vnekterejch aspektech je videt, že to je work in progress
+  ale že si myaslime že i tak to ukazuje pěkný výsledky 
+  a že doufame že ty nedodělky budou další prací odstraněny
+
+& Říct jak jde text:
+  && Začnem popsanim klasickýho GP podle Kozy
+  && Vyslovíme definice okolo lambda termu a typu
+  && Vyslovíme definici inhabitačního stromu
+     kterej je stromovou verzí IM od Barendrechta.
+  && Matematicky popíšem jak na to dem my.
+  && Popíšem rozšíření zakladního systému na 
+     polymorfní.
+  && Popíšem to z programátorskýho hlediska
+  && Popíšem příklady na kterejch to demonstrujem
+  && Srovnáme to s nějakejma přístupama co už 
+     existujou
+
+& Nějak naťuknout už tady v úvodu jak se to má při
+  srovnání s tim jak na to dou jiný (Yu atd)
+
+& Výkřiky co zmínit:
+  && Na co zaměřený:
+     &&& Proof of a concept
+     &&& Spíš jednoduchost
+         &&&& Ve většině aspektú to sleduje
+              Kozu - tzn i v málo řídících
+              parametru
+     &&& Duraz na generování termu (!)
+
+  && Na co nezaměřený (ale snad taky ok v tom):
+     &&& Performance   
+     &&& Mature knihovna co by šla použít robustně
+         v cizim kódu - spíš to počítá že to 
+         bude dál rozvíjeno a tohle je mezizastávka
+
+ & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
+   defining EA and then specifying the differences or something 
+   like that...
+
+& bylo by asi mimojiný dobrý 
 ospravedlnit/okomentovat tak fancy (=javascript frontend) 
@@
 k ukázání toho, že to neni omezený na haskell
-ale že se kombinatorový konstrukty daj v zásadě přeložit do libovolnýho jazyka celkem bez problémů
+ale že se kombinatorový konstrukty daj 
+v zásadě přeložit do libovolnýho jazyka celkem bez problému
 nebo minimalně do JS. Ale todle možná nepsat tolik do závěru
 spíš nějaky podrobnosti přehodit do implementačních detailu
-a tady to jen tak líznout.}
+a tady to jen tak líznout.
+
+\end{easylist}
 
@@
 
+\bibitem{barendregt92}
+H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum , H.  P.
+\emph{Lambda Calculi with Types.} 
+Handbook of Logic in Computer Science, 1992.
+
 \bibitem{barendregt10}
@@
 
+\bibitem{jones87}
+Simon Peyton Jones. 
+\emph{The Implementation of Functional Programming Languages}. 
+Prentice Hall, 1987.
+
 \bibitem{AIAMA}

### 2013-07-19 tex/thesis.tex
[tex]

@@
 \newcommand{\Lp}  {\Lambda^{\prime}\xspace}
+
+\newcommand{\tur}[3]{#1 \vdash #2 : #3}
 
@@
 Let $V$ be infinite countable set of {\it 
-variable names}. Let $C$ be set of {\it constant names}.	 	
+variable names}. Let $C$ be set of {\it constant names}, 
+$V \cap C = \emptyset$.	 	
 Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
@@
 
-It is usual to use $V = \Sigma^+$ set of all non-empty finite strings of symbols 
+It is usual to not distinguish between $V$ and $U$ and just use
+one set of variable names. But because this distinction is quite convenient
+for us in the implementation, we use it even here. 
+Hopefully, it will not confuse the reader.
+
+We can use $V = \Sigma^+$, set of all non-empty finite strings of symbols 
 from $\Sigma$ where $\Sigma$ is some alphabet set, e.g.  
@@
 \begin{definition}
-Let $A$ be set of {\it atomic type names}. \\
+Let $A$ be set of {\it atomic type names}. 
 Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
@@
 we prefer this name, which also better suits our usage.
-\\
+
+Specific definition of $A$ is not so important, but for practical purposes 
+we can think of it as the set of all non-function types relevant in our
+situation, e.g., $A = \{ Int , IntList , Bool, BoolList , Char, String \}$.
+Or we can think of $A$ more generally as of set of all non-empty strings
+of symbols of some alphabet. \\
 
@@
 
-% We will utilize this property ... asi netřeba řikat
-
-\begin{todo}  
-   & možná se hlobějc zmínit o curryingu  
-\end{todo}
-
-\section{Statement of a form $M : \sigma$}
-
-	Let $\Lambda$ be set of {\it \lterms}. \\*
-	Let $\mathbb{T}$ be set of {\it types}.       \\*
-	A {\it statement} $M : \sigma$ is a pair $(M,\sigma) \in \Lambda \times \mathbb{T}$. \\*
-	$M : \sigma$ is vocalized as {\it "$M$ has type $\sigma$"}.\footnote{ 
-	$M : \sigma$ can be also imagined as $M \in \sigma$ } \\*
-	The type $\sigma$ is the {\it predicate} and the term $M$ is the
-	{\it subject} of the statement.
-
-\section{Context}
-
-	Let $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$. 
-	($\Gamma$ is a set of {\it statements} of a form $M : \sigma$.)	\\*
-	Then $\Gamma$ is {\it context} if it obeys following 
-	conditions\footnote{
-	The $\pi_1$ corresponds to the projection of the first component of the Cartesian product.
-	}:
+\subsection{Statements of the form $M : \sigma$}
+
+\begin{definition}
+	Let $\Lambda$ be set of {\it \lterms}. 
+	Let $\mathbb{T}$ be set of {\it types}.       
+	A {\it statement} $M : \sigma$ is a pair 
+	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
+	\\ 
+\end{definition}
+
+Statement $M : \sigma$ is vocalized as {\it "$M$ has type $\sigma$"}.
+The type $\sigma$ is called the {\it predicate} and 
+the term $M$ is called the {\it subject} of the statement $M : \sigma$. \\
+
+We can look on this relation in tree different ways.
+
+\begin{description}
+  \item[Programming point of view.] 
+   Statement $M : \sigma$ means that value $M$ is 
+   an instance of data type
+   $\sigma$. So statement $M : \sigma$ is
+   somehow analogous to \textit{Java}
+   expression \texttt{$M$ instanceof $\sigma$}.
+
+   Type \textit{constrains} the value 
+   and gives \textit{information} about it, 
+   e.g., function $f : \sigma \ar \tau$ 
+   is constrained to only accept values of a type $\sigma$
+   as input and must return value of a type $\tau$ as output.
+
+  \item[Set point of view.] 
+     Statement $M : \sigma$ can also be viewed as relation
+     $M \in \sigma$. This makes us look on types as on
+     collections of their instances.
+
+  \item[Logic point of view.] According to 
+  the \textit{Curry–Howard correspondence}
+  we can read statement $M : \sigma$ as 
+  \textit{"$M$ is a proof of formula $\sigma$"}.
+  We will not treat this interesting subject in much detail,
+  \lets just say that $\ar$ corresponds to the
+  implication $\then$ and \textit{atomic types} correspond to
+  \textit{propositional variables}. The whole system of
+  \textit{simply typed $\lambda$-calculus} then corresponds
+  to the \textit{implicational fragment of propositional 
+  intuitionistic logic}. We are mentioning this beautiful
+  correspondence in hope that it will help clarify 
+  use of \textit{inference rules} which are key concept
+  in the process of deriving valid statements about
+  terms and types.
+
+\end{description}
+
+\subsection{Context}
+
+%\begin{definition}
+%Let $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$. 
+%($\Gamma$ is a set of statements of a form $M : \sigma$.)	
+%Then $\Gamma$ is {\it context} if it obeys following 
+%conditions.
+%%\footnote{The $\pi_1$ corresponds to the projection of the 
+%%first component of the Cartesian product.}
+%\begin{align*}
+%\forall (x,\sigma) \in \Gamma &: x \in V \cup C  
+%&\textit{(x is variable or constant)}\\
+%\forall (x_1,\sigma_1),(x_2,\sigma_2) \in \Gamma &: 
+%x_1,x_2 \in V , x_1 \neq x_2 \then \sigma_1 \neq \sigma_2
+%&\textit{(variables are distinct)}
+%\end{align*}~
+%Ten druhej je navíc blbě, má tam bejt celej stejtment se nerovná pak...
+%\end{definition}
+
+\begin{definition}
+~
+\begin{enumerate}
+ \item A \textit{declaration} is a statement 
+ $x : \sigma$ where $x \in V \cup C$.
+
+ \item A \textit{context} (or \textit{basis}) 
+ is set of declarations with distinct variables as subjects.
+\end{enumerate}~
+\end{definition}
+
+Again our definition slightly differs from the standard one
+in that it permits more than one declaration with the same 
+\textit{constant}, which enables us to simulate very primitive
+kind of ad-hoc polymorphism by \textit{"constant overloading"}.  
+And again it does not affect the theory; it just seems less
+confusing to me to state these definitions as they are later 
+used in the implementation.   \\
+
+In our system, the set of \textit{building blocks} 
+(the equivalent of the set $T \cup F$ in classical GP)
+will be a context with all declarations having \textit{constants}
+as subjects, i.e., a set of constants together with their types.
+
+In the process of term generation, 
+this initial context is being further extend with variables
+ --- by "entering inner scopes of bodies of lambda functions".
+
+%\subsection{Statements of the form \GMS}
+
+%\subsection{Inference rules}
+\subsection{Simply typed $\lambda$-calculus}
+
+\begin{definition}
+A statement $M : \sigma$ is \textit{derivable from}
+a context $\Gamma$ if it can be produced by the following rules.
 \begin{align*}
-		 \forall (x,\sigma) \in \Gamma &: x \in V \cup C \\
-		 \forall s_1,s_2 \in \Gamma &: s_1 \neq s_2 \Rightarrow \pi_1(s_1) \neq \pi_1(s_2)
-    \end{align*}
-
-	In other words context is a set of statements with distinct variables or constants as subjects.
-
-\begin{todo}
- & TALK ABOUT Context represents library/building blocks.
-\end{todo}
-
-\section{Statement of a form \GMS}
+(x,\sigma) \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
+\tur{\Gamma}{M}{(\sigma \ar \tau)}~,~\tur{\Gamma}{N}{\sigma} 
+&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
+\tur{\Gamma \cup (x,\sigma)}{M}{\tau}
+&~\then~ \tur{\Gamma}{\lamb{x}{M}}{(\sigma \ar \tau)} 
+\end{align*}~
+\end{definition}
 
@@
 
-	We construct valid statements of form \GMS by using inference rules.
-
-\section{Inference rule}
+We construct valid statements of the form \GMS 
+by using inference rules.\\
 
@@
 
-\section{Simply typed $\lambda$-calculus}
+\subsection{Simply typed $\lambda$-calculus}
 
 \begin{todo}
+   & tady to nějak okomentovat že to je teda vono
    & říct že se vždycky zastavěj (strong normalization) ale

### 2013-07-19 tex/thesis.tex
[tex]

@@
 
-\newcommand{\tur}[3]{#1 \vdash #2 : #3}
-
-\newcommand{\turst}[3]{$#1 \vdash #2 : #3$\xspace}
+\newcommand{\tur}[3]{#1\vdash{}#2:#3}
+
+\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
 \newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
@@
    This generation process is parameterized by simple 
-   \textit{search strategy}. Several search strategies are presented.   
+   \textit{search strategy}. Several search strategies are presented,
+   such as \textit{systematic} strategy and strategy performing
+   classical GP individual generation.     
  \item \textit{Abstraction elimination} is utilized
@@
 
+\begin{todo}
+& zmínit $=_\beta$ a $\twoheadrightarrow_\beta$ 
+jestli se k tomu někde dál v textu odkazuju
+\end{todo}
+
 \subsection{$\eta$-reduction}
@@
 \begin{definition}
-A statement $M : \sigma$ is \textit{derivable from}
-a context $\Gamma$ if it can be produced by the following rules.
+A statement $M\colon\sigma$ is \textit{derivable from}
+a context $\Gamma$ (notation $\Gamma\vdash{}M\colon\sigma$) 
+if it can be produced by the following rules.
 \begin{align*}
 (x,\sigma) \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
-\tur{\Gamma}{M}{(\sigma \ar \tau)}~,~\tur{\Gamma}{N}{\sigma} 
+\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
 &~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
-\tur{\Gamma \cup (x,\sigma)}{M}{\tau}
-&~\then~ \tur{\Gamma}{\lamb{x}{M}}{(\sigma \ar \tau)} 
+\tur{\Gamma \cup \{(x,\sigma)\}}{M}{\tau}
+&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
 \end{align*}~
@@
 
-By writing \GMS we say 
-{\it statement $M : \sigma$ is derivable from context $\Gamma$ }.
-
-We construct valid statements of the form \GMS 
-by using inference rules.\\
-
-	Basically speaking, inference rules are used for deriving statements of a form 
-	\GMS from yet derived statements of such a form.
-	Those inference rules are written in the following form:
-
-	\begin{equation*}
-		\frac{\Gamma_1 \vdash M_1 : \sigma_1 \qquad
-			  \Gamma_2 \vdash M_2 : \sigma_2 \quad
-			  \dotsm \quad
-		      \Gamma_n \vdash M_n : \sigma_n}
-		     {\Gamma_{n+1} \vdash M_{n+1} : \sigma_{n+1}}
-	\end{equation*}
-
-	Suppose we have yet derived statements 
-	$\Gamma_1 \vdash M_1 : \sigma_1 ,
-	 \Gamma_2 \vdash M_2 : \sigma_2 ,
-	 \dots ,
-	 \Gamma_n \vdash M_n : \sigma_n$. 
-	It allows as to use the inference rule to derive statement
-	\mbox{ $\Gamma_{n+1} \vdash M_{n+1} : \sigma_{n+1}$ }.
-
-	For deriving statements including types of a form 
-	$(\sigma \rightarrow \tau)$ are essential those two 
-	inference rules:
-
-	\begin{equation*}
-		\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
-			  \Gamma \vdash N : \sigma }
-		     {\Gamma \vdash (M N) : \tau }
-	\end{equation*}
-
-	\begin{equation*}
-		\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
-		     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
-	\end{equation*}
-
-	This kind of inference rules allows us to derive new statements from yet derived statements, but 
-	what if we do not have any statement yet? 
-	For this purpose we have other kinds of inference rules such as {\it axiom} inference rule:
-
-	\begin{equation*}
-		\frac{( x , \sigma )  \in \Gamma}
-		     {\Gamma \vdash x : \sigma}
-	\end{equation*}
-
-	\Lets consider an example statement of a form \GMS :
-
-	\[
+Those rules are usually stated in the form of \textit{inference rules}.
+
+\newcommand{\ruleI}[2]{\dfrac{#1}{#2}}
+\newcommand{\ruleII}[3]{\dfrac{#1 \qquad #2}{#3}}
+\newcommand{\ruleIII}[4]{\dfrac{#1 \qquad #2 \qquad #3}{#4}}
+\newcommand{\ruleIV}[5]{\dfrac{#1 \qquad #2 \qquad #3 \qquad #4}{#5}}
+
+\begin{align*}
+\ruleI { (x:\sigma) \in \Gamma }
+       { \tur{\Gamma}{x}{\sigma} }
+&& \textit{(axiom)} \\[1em]
+\ruleII{ \tur{\Gamma}{M}{\sigma \ar \tau} }
+       { \tur{\Gamma}{N}{\sigma} }
+       { \tur{\Gamma}{(M~N)}{\tau} }
+&& \textit{($\ar$-elimination)} \\[1em]
+\ruleI { \tur{\Gamma,x:\sigma}{M}{\tau} }
+       { \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} }
+&& \textit{($\ar$-introduction)} \\[1em]
+\end{align*}
+
+\Lets consider following example statement of a form \GMS.
+
+$$
 	\{\} \vdash (\lambda f . (\lambda x . (f x) )) : 
 	(\sigma \rightarrow \tau) \rightarrow ( \sigma \rightarrow \tau ) 
-	\]
-
-	This statement is derived as follows:
+$$~
+
+This statement is derived as follows.
 
@@
 
-\subsection{Simply typed $\lambda$-calculus}
+If we turn the derivation upside down, we can see that it is a tree with 
+desired statement in the root and \textit{axioms} in the leafs. \\
+
+New valid inference rules may be produced by finite composition of 
+other valid inference rules.
+
+\Lets consider following rule.
+
+$$
+\ruleII
+ { \tur{\Gamma,x:\sigma}{M}{\tau} }
+ { \tur{\Gamma}{N}{\sigma} }
+ { \tur{\Gamma}{(\lamb{x}{M}~N)}{\tau} }
+$$
+
+Its validity is shown by the following rule composition.
+
+$$
+\ruleII
+	{ \ruleI 
+	     { \tur{\Gamma,x:\sigma}{M}{\tau} }
+	     { \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} }
+	}
+	{ \tur{\Gamma}{N}{\sigma} }
+	{ \tur{\Gamma}{(\lamb{x}{M}~N)}{\tau} }
+$$~
+
+\Lets consider following strange rule.
+
+$$
+\ruleII
+ { \tur{\Gamma,x:\rho \ar \pi}{M}{\tau} }
+ { \tur{\Gamma,y:\rho}{N}{\pi} }
+ { \tur{\Gamma}{(\lamb{x}{M}~\lamb{y}{N})}{\tau} }
+$$
+
+Its validity is shown by the following rule composition.
+
+$$
+\ruleII
+	{ \ruleI 
+	     { \tur{\Gamma,x:\rho \ar \pi}{M}{\tau} }
+	     { \tur{\Gamma}{\lamb{x}{M}}{(\rho \ar \pi) \ar \tau} }
+	}
+	{ \ruleI 
+	     { \tur{\Gamma,y:\rho}{N}{\pi} }
+	     { \tur{\Gamma}{\lamb{y}{N}}{\rho \ar \pi} }
+	}	 { \tur{\Gamma}{(\lamb{x}{M}~\lamb{y}{N})}{\tau} }
+$$~
+
+Generally speaking, inference rules are used for deriving statements of a form 
+\GMS from yet derived statements of such a form (with exception of 
+\textit{axiom} rule, which serves as a "starting point").
+Those inference rules are written in the following form.
+
+$$ 
+\ruleIV
+  { \tur{\Gamma_1}{M_1}{\sigma_1} }
+  { \tur{\Gamma_2}{M_2}{\sigma_2} }
+  { \dotsm }
+  { \tur{\Gamma_n}{M_n}{\sigma_n} }
+  { \tur{\Gamma_{n+1}}{ \mathbf{T}(M_1,M_2,\dots,M_n)}{\sigma_{n+1}} } 
+$$
+
+Where $\mathbf{T}(M_1,M_2,\dots,M_n)$ stands for some 
+specific \lterm that may contain \textit{"meta-variables"}
+$M_1,M_2,\dots,M_n$.
+
+In the example above $\mathbf{T}(M,N) = (\lamb{x}{M}~\lamb{y}{N})$
+
+Suppose we have derived statements 
+$\Gamma_1 \vdash M_1 : \sigma_1 ,
+ \Gamma_2 \vdash M_2 : \sigma_2 ,
+ \dots ,
+ \Gamma_n \vdash M_n : \sigma_n$. 
+Then we are allowed to use the inference rule to derive statement
+\mbox{ $\Gamma_{n+1} \vdash \mathbf{T}(M_1,M_2,\dots,M_n) : \sigma_{n+1}$ }.\\
+
+%For deriving statements including types of a form 
+%$(\sigma \rightarrow \tau)$ are essential those two 
+%inference rules:
+%
+%\begin{equation*}
+%	\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
+%		  \Gamma \vdash N : \sigma }
+%	     {\Gamma \vdash (M N) : \tau }
+%\end{equation*}	
+%
+%\begin{equation*}
+%	\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
+%	     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
+%\end{equation*}		 
+% 
+%This kind of inference rules allows us to derive new statements from yet derived statements, but 
+%what if we do not have any statement yet? 
+%For this purpose we have other kinds of inference rules such as {\it axiom} inference rule:   
+%	
+%\begin{equation*}
+%	\frac{( x , \sigma )  \in \Gamma}
+%	     {\Gamma \vdash x : \sigma}
+%\end{equation*}
 
 \begin{todo}
-   & tady to nějak okomentovat že to je teda vono
-   & říct že se vždycky zastavěj (strong normalization) ale
-     asi to říct na jinym místě než tady, tady se teprv definujou typy, 
-     říct to někde kde už se ví jak sou napojený na termy    
+   & říct že se vždycky zastavěj (strong normalization),
+     && potřebný definice pro to uvést v \lterms sekci
+   & možná zmínit že se jedná o \textit{gentzen style proof}
+   & říct že jsou to modus ponens a jeden směr věty o dedukci  
 \end{todo}
@@
 \section{Term generating grammar}
-
-Inference rules are good for deriving statements of a form \GMS, but our
+Inference rules are good for deriving statements of the form \GMS, but our
 goal is slightly different; we would like to generate many \lterms M for a given type 
-$\sigma$ and context $\Gamma$.
-
-Our approach will be to take each inference rule and transform it to a rule of term generating
-grammar. With this term generating grammar it will be much easier to reason about generating 
-\lterms.
-
-It won't be a grammar in classical sense because we will be operating with infinite sets of
-nonterminal symbols and rules. \footnote{TODO : mention terminal symbols - situation around 
-variables and their construction with ' symbol.}
-
-Let $Non = Type \times Context $ be our {\it nonterminal} set. 
-So for every $i \in Non$ is $i = (\sigma_i , \Gamma_i )$.
-
-\Lets consider each relevant inference rule and its corresponding grammar rule.
-
-First inference rule is {\it implication elimination} also known as 
-{\it modus ponens}: 
-\[
-	\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
-		  \Gamma \vdash N : \sigma }
-	     {\Gamma \vdash (M N) : \tau }
-\]
-\\
-For every $\sigma, \tau \in \mathbb{T}$ and for every {\it context} 
-$\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$ there is a grammar rule of a form\footnote{ 
-Terminal symbols for parenthesis and normally {\it space} now \textvisiblespace \quad (for {\it function application} operator) are visually highlighted. }: 
-\[	
+$\sigma$ and context $\Gamma$.\\
+
+\Lets start our reasoning about generation of \lterms with
+notion of term generating grammar. We will show how to
+transform an inference rule into a rule of term generating grammar.
+With the appropriate term generating grammar 
+in hands it is just a simple step to \textit{inhabitation trees}.\\
+
+It is not a grammar in classical sense, because we will be operating with infinite sets of nonterminal symbols and rules. Technically it it
+2-level grammer \cite{todo}.\\
+
+Our terminals are symbols used for writing \lterms 
+.%,i.e., parentheses,  .
+
+Our non-terminals are pairs $(\tau,\Gamma)$ where 
+$\tau \in \T$ and $\Gamma$ is a \textit{context}.
+
+Our initial non-terminal is $S = (\sigma_S,\Gamma_S)$
+where $\sigma_S$ is desired type of generated \lterms and
+$\Gamma_S$ is \textit{building blocks} context. \\
+
+\Lets begin with \textit{axiom} inference rule.
+
+$$
+\ruleI { (x:\sigma) \in \Gamma }
+       { \tur{\Gamma}{x}{\sigma} }
+$$~
+
+%It corresponds to following grammar rule schema.
+%$$ ( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}} 
+%~~~~~~~~~~~~~ \textbf{if } (x,\sigma) \in \Gamma$$
+
+For every non-terminal $(\sigma,\Gamma)$ 
+and for every $x$ such that $(x:\sigma) \in \Gamma$,
+there is following grammar rule.
+$$ ( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}} $$ \\
+
+Next is \textit{$\ar$-elimination}.
+
+$$
+\ruleII{ \tur{\Gamma}{M}{\sigma \ar \tau} }
+       { \tur{\Gamma}{N}{\sigma} }
+       { \tur{\Gamma}{(M~N)}{\tau} }
+$$~
+
+For every non-terminal $(\tau,\Gamma)$ 
+and for every type $\sigma \in \T$ 
+there is following grammar rule.\footnote{ 
+Terminal symbols for parenthesis and normally {\it space} 
+now \textvisiblespace \quad (for {\it function application} operator) 
+are visually highlighted. } 
+$$
 	( \tau , \Gamma )  \longmapsto
@@
 	\mbox{ \Vtextvisiblespace[1em] } ( \sigma , \Gamma ) \bigg)
-\]
-\\
-
-Second inference rule is {\it implication introduction}: 
-\[
-	\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
-	     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
-\]
-\\
-$\forall \sigma, \tau \in \mathbb{T}$ 
-$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
-$\forall x \in V $ such that there is no $(x,\rho) \in \Gamma$ 
-there is a grammar rule:
-\[ 
+$$\\
+
+And finally {\it $\ar$-introduction}. 
+$$
+\ruleI { \tur{\Gamma,x:\sigma}{M}{\tau} }
+       { \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} }
+$$
+
+For every non-terminal $(\sigma \ar \tau,\Gamma)$
+and for every variable $x$
+there is following grammar rule.\footnote{
+\textbf{TODO !!!!} nějak okomentovat že by měla být freš,
+pokudmožno v celym termu ale to že byse dělalo blě aby to
+zůstala bezkontextová gramatika, ale je to celkem nepodtatnej detail
+protože tyhle term generating gramatiky sou tu spíš
+pro zajímavost a pro pevnější kramfleky, než pro praktický 
+generování}
+
+%and let $x$ be variable name that is
+%not present in any declaration in $\Gamma$.
+%(since we want to generate "fresh" variables).
+
+$$ 
 	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
 	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
-\]
-\\
-
-Third inference rule is {\it axiom}: 
-\[
-		\frac{( x , \sigma )  \in \Gamma}
-		     {\Gamma \vdash x : \sigma}
-\]
-\\
-$\forall \sigma \in \mathbb{T}$ 
-$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
-$\forall x \in V \cup C $ such that $(x,\sigma) \in \Gamma$ 
-there is a grammar rule:
-\[ 
-	( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}}
-\]
-\\
+$$\\
 
@@
 	~ \Big)
-\end{align*}
+\end{align*}~\\
+
+\Lets consider a general inference rule again.
+$$ 
+\ruleIV
+  { \tur{\Gamma_1}{M_1}{\sigma_1} }
+  { \tur{\Gamma_2}{M_2}{\sigma_2} }
+  { \dotsm }
+  { \tur{\Gamma_n}{M_n}{\sigma_n} }
+  { \tur{\Gamma_{n+1}}{ \mathbf{T}(M_1,M_2,\dots,M_n)}{\sigma_{n+1}} } 
+$$
+
+The corresponding grammar rule can be summarized as follows.
+
+$$
+ (\sigma_{n+1},\Gamma_{n+1})
+ \longmapsto
+ \mbox{ {\Large \textbf{T}}} 
+ \bigg( 
+ (\sigma_1,\Gamma_1)
+ \mbox{ {\Large ,}}
+ (\sigma_2,\Gamma_2)
+ \mbox{ {\Large ,}}
+ ~\dots
+ \mbox{ {\Large ,}}
+ (\sigma_n,\Gamma_n)
+ \bigg)  
+$$
+
+\begin{todo}
+   & asi ocitovat toho van Wijngaarden [1981] s těma 2-level grammar
+     jak je to v Barendregtovi \cite{barendregt10} 
+   & vyřešit jak je to stou freš proměnnou x v introdukci...
+     && stejně tak v tom obecnym případě to dělá neplechu
+\end{todo}
+
+%\section{Term generating grammar}
+%
+%Inference rules are good for deriving statements of the form \GMS, but our
+%goal is slightly different; we would like to generate many \lterms M for a given type 
+%$\sigma$ and context $\Gamma$.
+%
+%Our approach will be to take each inference rule and transform it to a rule of term generating
+%grammar. With this term generating grammar it will be much easier to reason about generating 
+%\lterms.
+%	
+%It won't be a grammar in classical sense because we will be operating with infinite sets of
+%nonterminal symbols and rules. \footnote{TODO : mention terminal symbols - situation around 
+%variables and their construction with ' symbol.}
+%
+%Let $Non = Type \times Context $ be our {\it nonterminal} set. 
+%So for every $i \in Non$ is $i = (\sigma_i , \Gamma_i )$.
+%
+%\Lets consider each relevant inference rule and its corresponding grammar rule.
+%
+%First inference rule is {\it implication elimination} also known as 
+%{\it modus ponens}: 
+%\[
+%	\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
+%		  \Gamma \vdash N : \sigma }
+%	     {\Gamma \vdash (M N) : \tau }
+%\]
+%\\
+%For every $\sigma, \tau \in \mathbb{T}$ and for every {\it context} 
+%$\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$ there is a grammar rule of a form\footnote{ 
+%Terminal symbols for parenthesis and normally {\it space} now \textvisiblespace \quad (for {\it function application} operator) are visually highlighted. }: 
+%\[	
+%	( \tau , \Gamma )  \longmapsto
+%	\bigg( ( \sigma \rightarrow \tau , \Gamma ) 
+%	  \mbox{ \Vtextvisiblespace[1em] } ( \sigma , \Gamma ) \bigg)
+%\]
+%\\
+%
+%Second inference rule is {\it implication introduction}: 
+%\[
+%	\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
+%	     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
+%\]
+%\\
+%$\forall \sigma, \tau \in \mathbb{T}$ 
+%$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
+%$\forall x \in V $ such that there is no $(x,\rho) \in \Gamma$ 
+%there is a grammar rule:
+%\[ 
+%	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
+%	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
+%\]
+%\\	
+%
+%Third inference rule is {\it axiom}: 
+%\[
+%		\frac{( x , \sigma )  \in \Gamma}
+%		     {\Gamma \vdash x : \sigma}
+%\]
+%\\
+%$\forall \sigma \in \mathbb{T}$ 
+%$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
+%$\forall x \in V \cup C $ such that $(x,\sigma) \in \Gamma$ 
+%there is a grammar rule:
+%\[ 
+%	( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}}
+%\]
+%\\
+%
+%We will demonstrate \lterm generation on example. 
+%Again on $(\lambda f . (\lambda x . (f x) ))$. 
+%We would like to generate \lterm of a type 
+%$(\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau)$
+%with $\Gamma = \{\}$.
+%\begin{align*}
+%	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
+%	\longmapsto & \Big( \mbox{ {\Large $\lambda$f.}}
+%	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
+%	~ \Big)
+%	\\
+%	\longmapsto & 
+%	\Big( \mbox{ {\Large $\lambda$f. }}
+%		\Big( \mbox{ {\Large $\lambda$x. }}
+%	  	 	( \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
+%		~ \Big)  	 
+%	~ \Big)
+%	\\
+%	\longmapsto & 
+%	\Big( \mbox{ {\Large $\lambda$f. }}
+%		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
+%	  	 	\Big( 
+%	  	 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
+%			  \mbox{ \Vtextvisiblespace[1em] } 
+%			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} )  \Big) 
+%		~ \Big)  	 
+%	 ~ \Big)
+%	\\
+%	\longmapsto & 
+%	\Big( \mbox{ {\Large $\lambda$f. }}
+%		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
+%	  	 	\Big( 
+%	  	 	  \mbox{ {\Large f}} 
+%			  \mbox{ \Vtextvisiblespace[1em] } 
+%			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \Big) 
+%		~ \Big)  	 
+%	~ \Big)		
+%	\\
+%	\longmapsto & 
+%	\Big( \mbox{ {\Large $\lambda$f. }}
+%		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
+%	  	 	\Big( 
+%	  	 	  \mbox{ {\Large f}} 
+%			  \mbox{ \Vtextvisiblespace[1em] } 
+%			  \mbox{{\Large x}} \Big) 
+%		~ \Big)  	 
+%	~ \Big)
+%\end{align*}
+%
 
@@
 
-Let \GMS where $\sigma = \tau_0 \ar \tau_1 \ar \dots \ar \tau_n \ar \alpha$.\\
+Let \GMS where 
+$\sigma = \tau_0 \ar \tau_1 \ar \dots \ar \tau_n \ar \alpha$,
+$n \geq 0$.\\
 Then $M$ is in \textit{long normal form} if following 
@@
 %"Barendregt-like"
-\section{Inference and grammar rules producing 
-terms in Long normal form}
+\section{Inference and grammar rules producing terms in Long normal form}
 \label{barlike}

### 2013-07-20 tex/thesis.tex
[tex]

@@
 \newtheorem{theorem}{Theorem} 
-\newtheorem{proposition}{Proposition} 
+\newtheorem*{proposition}{Proposition} 
 \newtheorem{lemma}{Lemma} 
@@
 \begin{todo}
-& zmínit $=_\beta$ a $\twoheadrightarrow_\beta$ 
-jestli se k tomu někde dál v textu odkazuju
+& !!!!!!! zmínit $=_\beta$ a $\twoheadrightarrow_\beta$  a i pro ostatní redukce!!!!!
 \end{todo}
@@
 
+\begin{todo}
+ & zapsat sem notaci s $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
+\end{todo}
+
 %\subsection{Inference rules}
@@
 \newcommand{\ruleII}[3]{\dfrac{#1 \qquad #2}{#3}}
+\newcommand{\ruleIIs}[3]{\dfrac{#1 ~ #2}{#3}}
 \newcommand{\ruleIII}[4]{\dfrac{#1 \qquad #2 \qquad #3}{#4}}
@@
 \section{Term generating grammar}
+
+\newcommand{\gar}{\longmapsto}
+
 Inference rules are good for deriving statements of the form \GMS, but our
@@
 %It corresponds to following grammar rule schema.
-%$$ ( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}} 
+%$$ ( \sigma , \Gamma )  \gar \mbox{ {\Large x}} 
 %~~~~~~~~~~~~~ \textbf{if } (x,\sigma) \in \Gamma$$
@@
 $$
-	( \tau , \Gamma )  \longmapsto
+	( \tau , \Gamma )  \gar
 	\bigg( ( \sigma \rightarrow \tau , \Gamma ) 
@@
 $$ 
-	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
+	( \sigma \rightarrow \tau , \Gamma )  \gar
 	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
@@
 	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
-	\longmapsto & \Big( \mbox{ {\Large $\lambda$f.}}
+	\gar & \Big( \mbox{ {\Large $\lambda$f.}}
 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
@@
 	\\
-	\longmapsto & 
+	\gar & 
 	\Big( \mbox{ {\Large $\lambda$f. }}
@@
 
-The corresponding grammar rule can be summarized as follows.
+The corresponding grammar rule can be summarized as following schema.
 
@@
  (\sigma_{n+1},\Gamma_{n+1})
- \longmapsto
+ \gar
  \mbox{ {\Large \textbf{T}}} 
@@
 \begin{todo}
-   & asi ocitovat toho van Wijngaarden [1981] s těma 2-level grammar
+   & buď dokázat ty převody na grammar rules explicitně
+     nebo říct že se to dokáže podobnym stylem jako v dukazu 
+     v sekci \ref{barlike}
+   & možná ocitovat toho van Wijngaarden [1981] s těma 2-level grammar
      jak je to v Barendregtovi \cite{barendregt10} 
@@
 \end{todo}
+
+\section{Long normal form}
+\label{lnf}
+
+\newcommand{\lnf}{\textit{lnf}\xspace}
+
+\begin{definition}
+Let \GMS where 
+$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha$,
+$n \geq 0$.
+Then $M$ is in \textit{long normal form} (\textit{lnf}) if following 
+conditions are satisfied.
+\begin{enumerate}
+ \item $M$ is term of the form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_m}$\\
+  (specially for $n = 0$, $M$ is term of the form $f$),
+ \item each $M_i$ is in \textit{long normal form}.
+\end{enumerate}
+\end{definition}~
+
+As is shown in \cite{barendregt10} long normal form has those 
+nice properties:
+
+\begin{easylist}[itemize]
+& Every $M$ in $\beta$-normal form has a long normal form 
+$N$ such that $N \twoheadrightarrow_{\eta} M$.
+& \textbf{TODO jen zkopčeno přepsat!!} .. if M has a beta-nf, which according to Theorem 2B.4 is
+always the case, then it also has a unique lnf and this will be its unique $\beta\eta^{-1}$-nf.
+& \textbf{TODO : potřebuju eštěnějaký z tyhle knihy?}
+\end{easylist}
+
+\begin{todo}
+ & v barendrechtovi je to v definici lnf $\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$ místo 
+   $M_m$ ale to musí bejt určitě překlep...
+ & popsat tu jak ji předelat na beta nf
+	pokud by se nám chtělo... 
+	V barendrech84 (ta nová kniha s žlutejma deskama) je ukázano že je 
+	eta CR (čurč roserovská) což snad stačí společně s 
+	tim tvrzeničkem z book.pdf že vezmu lnf
+	a dělám eta redukce dokud to de. až to nejde tak bych to měl mít v 
+	beta-nf
+ & přepsat to nějak hezčejc, ten seznam heskejch vlastností
+\end{todo}
+
+\section{Grammar producing \lterms in \lnf}
+\label{barlike}
+
+In \cite{barendregt10} is shown term generating grammar with 
+following rules (our notation is used, but we will not 
+highlighted terminals anymore).\footnote{
+I was using term generating grammars for term generation
+on my own before I encountered with \cite{barendregt10},
+where i happily discovered that Barendregt is using
+almost identical notation.
+}
+\begin{align*}
+( \alpha , \Gamma )  
+&\gar
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
+& \textbf{if } \alpha \in A,
+(f : \rho_1 \ar \dots \ar \rho_m \ar \alpha) \in \Gamma
+\\ 
+( \sigma \rightarrow \tau , \Gamma )  
+&\gar
+(~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)
+&   
+\end{align*}
+
+The second rule can be replaced by more effective one.
+\[ 
+( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
+\gar
+(~\lambda~x_1~\dots~x_n~.~
+( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
+\]
+
+This rule packs consecutive uses of the second rule into one use.
+This is valid since the use of the second rule is deterministic;
+it is used if and only if the non-terminal's type is not atomic.\\
+
+Following proposition about this grammar is stated (but not proven)
+in~\cite{barendregt10}. Since it is key proposition for our 
+term generating method, we also provide a proof.
+
+\newcommand{\garr}{\leadsto}%{\Rrightarrow}%
+\newcommand{\Gp}{\Gamma^\prime}
+
+\begin{proposition}
+Let context $\Gamma$, \lterm $M$ and type $\sigma$ be given.\\
+Let $\garr$ be transitive reflexive closure of $\gar$. 
+Then
+$$ 
+(\sigma,\Gamma) \garr M 
+\Leftrightarrow
+\tur{\Gamma}{M}{\sigma} \text{ and $M$ is in lnf}.   
+$$
+\end{proposition}
+
+\begin{proof}~
+
+\Lets start with proving
+
+$ (\sigma,\Gamma) \garr M 
+\then
+\tur{\Gamma}{M}{\sigma} \text{ and $M$ is in \textit{lnf}}$
+
+by induction on the length of rewriting chain of $(\sigma,\Gamma) \garr M$.\\
+
+Let $\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha (n \geq 0)$,
+
+$\Gp = \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n$.\\
+
+$(\tau_1 \ar \dots \ar \tau_n \ar \alpha,\Gamma), n>0$ must rewrite to
+$(~\lambda~x_1~\dots~x_n~.~( \alpha , \Gp )~)$,
+
+thus (even for $n = 0$) the chain must go through
+$(~\lambda~x_1~\dots~x_n~.~( \alpha , \Gp )~)$.\\
+
+$(\alpha,\Gp)$ must rewrite to 
+$(~f~( \rho_1 , \Gp )~\dots~( \rho_m , \Gp )~)$
+
+for some $(f,\rho_1\ar\dots\ar\rho_m\ar\alpha) \in \Gp$,
+
+thus the chain must go through 
+$(~\lambda~x_1~\dots~x_n~.~(~f~( \rho_1 , \Gp )~\dots~( \rho_m , \Gp )~)~)$.\\
+
+After that each $( \rho_i , \Gp ) \garr M_i$
+with shorter chain than $(\sigma,\Gamma) \garr M $.
+
+Thus we have from the induction hypothesis for each $M_i$ that
+
+$\tur{\Gp}{M_i}{\rho_i}$ and $M_i$ is in \textit{lnf}.\\
+
+$M = (~\lambda~x_1~\dots~x_n~.~(~f~M_1~\dots~M_m~)~)$ 
+and each $M_i$ is in \lnf,
+
+thus M is in \lnf.\\
+
+Following derivation shows how to derive \GMS.\\
+
+$
+\ruleI{
+\ruleI{ 
+\ruleI{
+\ruleI{
+\ruleIIs{
+\ruleIIs{
+\ruleIIs{
+\ruleIIs{ 
+\ruleI
+{\boxed{(f,\rho_1\ar\dots\ar\rho_m\ar\alpha) \in \Gp} }
+{\tur{\Gp}{f}{\rho_1\ar\dots\ar\rho_m\ar\alpha}}}
+{\boxed{
+ \tur{\Gp }
+     {M_1 }
+     {\rho_1 }}}
+{
+\tur{\Gp}{(~f~M_1~)}{\rho_2\ar\dots\ar\rho_m\ar\alpha} 
+}}
+{\boxed{
+ \tur{\Gp }
+     {M_2 }
+     {\rho_2 }}}
+{\vdots}}
+{\ddots}
+{\tur{\Gp }
+     {(~f~M_1~\dots~M_{m-1}~) }
+     {\rho_m \ar \alpha}}}
+{\boxed{
+ \tur{\Gp }
+     {M_m }
+     {\rho_m}} }
+{\tur{\Gp }
+     {(~f~M_1~\dots~M_m~) }
+     {\alpha}}}
+{\tur{\Gamma,x_1:\tau_1,\dots,x_{n-1}:\tau_{n-1} }
+     {\lam{x_n}{(~f~M_1~\dots~M_m~)} }
+     {\tau_n \ar \alpha} }}
+{\vdots}}
+{\tur{\Gamma,x_1:\tau_1 }
+       {\lam{x_2~\dots~x_n}{(~f~M_1~\dots~M_m~)} }
+       {\tau_2 \ar \dots \ar \tau_n \ar \alpha} } }
+{\boxed{
+ \tur{\Gamma }
+     {\lam{x_1~x_2~\dots~x_n}{(~f~M_1~\dots~M_m~)} }
+     {\tau_1 \ar \tau_2 \ar \dots \ar \tau_n \ar \alpha} }}
+$
+
+~\\~\\
+
+Now \lets prove the opposite direction
+
+$\tur{\Gamma}{M}{\sigma} \text{ and $M$ is in \textit{lnf}}
+\then
+(\sigma,\Gamma) \garr M$
+
+by induction on the size of M.\\
+
+Let $M = \lam{x_1 \dots x_n}{f~M_1~\dots~M_m}$,
+
+$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha$,
+
+$\Gp = \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n$.\\
+
+Then
+
+$
+( \sigma , \Gamma ) =
+( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
+\gar
+(~\lambda~x_1~\dots~x_n~.~
+( \alpha , \Gp )~)
+$
+
+by using the updated second grammar rule.\\
+
+From \GMS we have that $(f,\rho_1\ar\dots\ar\rho_m\ar\alpha) \in \Gp$
+and $\tur{\Gp}{M_i}{\rho_i}$.
+
+(This can be easily observed from definition of \GMS.)\\
+
+Therefore
+
+$
+(~\lambda~x_1~\dots~x_n~.~( \alpha , \Gp )~)
+\gar
+(~\lambda~x_1~\dots~x_n~.~
+	(~f~( \rho_1 , \Gp )~\dots~( \rho_m , \Gp )~)
+~)
+$
+
+by using the first rule.\\
+
+Each $M_i$ is in \lnf, 
+is smaller in size then $M$
+and $\tur{\Gp}{M_i}{\rho_i}$;
+
+therefore we have from the induction hypothesis that 
+$(\rho_i,\Gp) \garr M_i$.\\
+
+Therefore 
+$
+( \sigma , \Gamma )
+\garr
+(~\lambda~x_1~\dots~x_n~.~
+	(~f~M_1~\dots~M_m~)
+~)
+$.
+
+\end{proof}~
+
+This means that we can use this grammar to generate all \lterms $M$ in \lnf
+with desired type $\sigma$ from \textit{building blocks} context $\Gamma$.\\
+
+\subsection{Benefits of generating \lterms in \lnf}
+
+By generating \lterms in \textit{lnf} we avoid generating 
+\lterms $M$,$N$ such that $M \not= N$, but $M =_{\beta\eta} N$.
+In other words, we avoid generating two programs with different 
+source codes, but performing the same computation.
+
+Every \lterm $M$ such that \GMS has its unique \lnf $L$, 
+for which $L =_{\beta\eta} M$.
+Therefore the computation performed by \lterm $M$ 
+is not omitted, because it is the same computation
+as the computation performed by \lterm $L$. \\
+
+\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
+\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
+
+Generating \lterms in \lnf is even better than generating 
+\lterms in \bnf. Since \lnf is the same thing as \beenf,
+every \lterm in \lnf is also in \bnf.
+
+This comes straight from the definition of \beenf, 
+but one can also see it from observing method for generating
+terms in \bnf. As is shown in \cite{barendregt10}, 
+this method is obtained simply by replacement of 
+the first grammar rule by slightly modified rule,
+resulting in following grammar.
+\begin{align*}
+( \pi , \Gamma )  
+&\gar
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
+& \textbf{if } (f : \rho_1 \ar \dots \ar \rho_m \ar \pi) \in \Gamma
+\\ 
+( \sigma \rightarrow \tau , \Gamma )  
+&\gar
+(~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)
+&   
+\end{align*}
+
+The difference lies in that $f$'s type is no longer obligated to be fully expanded
+($\pi \in \T$ instead of $\alpha \in A$). This makes the grammar less deterministic,
+resulting in bigger search space. The new rule is generalization of the old one,
+thus all terms in \lnf will be generated, along with many new terms in \bnf that 
+are not in \lnf.
+
+By generating \lterms in \lnf we avoid generating 
+\lterms $M$,$N$ such that $M \not= N$ and $M =_{\eta} N$; 
+but by generating in \bnf we do not avoid this.
+
+\newpage
+\section{Inhabitation tree}
+
+Now we will introduce \textit{Inhabitation tree}, structure slightly different from
+\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Barendregt.
+We can think about Inhabitation tree as about unfolded Inhabitation machine.
+The motivation for using Inhabitation trees is belief that it will help us 
+reason about generation of \lterms 
+of a given type $\sigma$ and with a given context $\Gamma$.
+
+\subsection{Definition of Inhabitation tree}
+
+\textit{Inhabitation tree} is a \textit{rooted tree}, possibly infinite. 
+It has two types of nodes:
+
+\begin{samepage}
+\begin{itemize}
+  \item Type nodes   
+  			- containing type $\sigma \in \mathbb{T}$ 
+  			- aka "OR-node" %, Nonterminal-node. 
+  \item Symbol nodes 
+  			%- containing $ x^{*} \in \mathfrak P \left(V\right) -blobost pač ordered $ or $c \in C$
+  			- containing "$\lambda$-head" (nonempty finite sequence of variable names) or constant
+  			  name. 
+  			- aka "AND-node" %, Terminal-node. - ale nemusí, proto jsem to zakomentoval
+\end{itemize}
+\end{samepage}
+
+We construct Inhabitation tree for given type $\sigma$ and context $\Gamma$.\\
+We will define Inhabitation tree by describing its construction for a given $(\sigma,\Gamma)$.
+Notice that it will closely follow the rules from \ref{barlike}:
+
+\begin{itemize}
+\item The root of Inhabitation tree for $(\sigma,\Gamma)$ is 
+      \textit{type node} with $\sigma$ as type.
+\item All \textit{type nodes} have as child nodes only \textit{symbol nodes}. 
+\item And all \textit{symbol nodes} have as child nodes only \textit{type nodes}. 
+\end{itemize}
+
+\begin{samepage}
+Now we will resolve the child nodes of the root node.
+There are two cases of $\sigma$ (recall \ref{deftype}): 
+\begin{description}
+	\item[Atomic type] 
+		$\sigma = \alpha $ where $\alpha \in A$. 
+	\item[Function type] 
+		$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
+		where $n \geq 1, \alpha \in A$.
+\end{description}
+\end{samepage}
+
+First case \textbf{Atomic type} --- i.e., $\sigma = \alpha$ where $\alpha \in A$:\\
+For every $(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha) \in \Gamma$
+where $\alpha \in A$ there is a child \textit{symbol node} of the root containing constant name $f$.
+This symbol node containing $f$ has $m$ child subtrees corresponding to Inhabitation trees for 
+$(\rho_1,\Gamma),\dots,(\rho_m,\Gamma)$.
+
+~
+
+Compare this case with corresponding grammar rule:
+\[ 
+	( \alpha , \Gamma )  \gar
+	\bigg( \mbox{ {\Large f }}
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  ( \rho_1 , \Gamma )
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  \dots
+	  \mbox{ \Vtextvisiblespace[1em] } 
+	  ( \rho_m , \Gamma )
+	  \quad \bigg)
+\]
+
+Second case \textbf{Function type} --- i.e., 
+$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
+where $n~\geq~1, \alpha~\in~A$:\\
+For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.
+
+There is one and only one child \textit{symbol node} of the root containing "$\lambda$-head" 
+$\lambda x_1 \dots x_n$ which stands for sequence of variable names $(x_1,\dots,x_n)$.
+This symbol node containing $\lambda x_1 \dots x_n$
+has one and only one child subtree corresponding to Inhabitation trees for 
+$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$.
+
+~
+
+Compare this case with corresponding grammar rule:
+\[ 
+	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \gar
+	\bigg( \mbox{ {\Large 
+	$\lambda x_1 \dots x_n .$ 
+	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
+\]
+~\\
+\Lets consider following $(\sigma,\Gamma)$ as a simple example:
+
+\begin{align*}
+\sigma =& ~ \mathbb{B} \rightarrow  \mathbb{B} \rightarrow  \mathbb{B} \\ 
+\Gamma =& \{ ~ true : \mathbb{B}  \\
+        &  , ~ nand :  \mathbb{B} \rightarrow \mathbb{B} \rightarrow \mathbb{B} ~ \}
+\end{align*}
+
+This particular $(\sigma,\Gamma)$ results in the following tree:\\
+
+\Tree
+[.\text{ $\mathbb{B} \rightarrow \mathbb{B} \rightarrow \mathbb{B}$ }  
+	[.\textbf{$\lambda$x_1 x_2 } 
+		[.\text{ $\mathbb{B}$ } 
+			\textbf{true}  
+			[.\textbf{nand} 
+				\qroof{ ~~ $\dotsb$ ~~ }.\text{ $\mathbb{B}$ }
+				\qroof{ ~~ $\dotsb$ ~~ }.\text{ $\mathbb{B}$ } 
+			]
+			\textbf{x_1}
+			\textbf{x_2}
+		]
+	]
+]
+
+~
+
+Second example features our well known example:
+
+\begin{align*}
+\sigma =& ~ (\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau \\ 
+\Gamma =& \{ \}
+\end{align*}
+
+Which results in following tree:
+
+\Tree
+[.\text{ $(\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau $ } 
+	[.\textbf{$\lambda$ f x }	
+		[.\text{ $\tau$ }		
+			[.\textbf{f} 
+				[.\text{ $\delta$ }
+					\textbf{x}					
+				]
+			]
+		]
+	]
+]
+
+\subsection{And-or tree and searching in Inhabitation tree}
+\Lets consider following definition of 
+\textit{And-or tree}\footnote{
+\textbf{TODO}: Mention that on WIKI there is more general definition, but 
+for our purposes is this one sufficient.}:
+
+\textit{And-or tree} is a rooted tree where each node is labeled as either \textit{and-node} 
+or\footnote{xor} \textit{or-node}.
+
+By \textit{solving} And-or tree $T$ we mean finding finite tree $T'$
+subgraph of $T$ such that it follows these conditions: 
+\begin{itemize}
+	\item The root of $T'$ is the root of $T$.
+	\item Each \textit{and-node} in $T'$ has all the child nodes as in $T$.
+	\item Each \textit{or-node}  in $T'$ has precisely one child 
+	      node.\footnote{\textbf{TODO}: MENTION why precisely one 
+	      and not at least one ..or CHANGE the def. }   
+\end{itemize}
+~
+\Lets now consider following labeling of Inhabitation tree:
+
+\begin{itemize}
+  \item \textbf{Type nodes}   are labeled as \textbf{or-nodes}.   
+  \item \textbf{Symbol nodes} are labeled as \textbf{and-nodes}.
+\end{itemize}
+
+This labeling has following justification:
+
+Selection of exactly one
+child node in \textit{type node} corresponds to selection of exactly one
+grammar rule in order to rewrite nonterminal symbol.
+
+Selection of all the child nodes in \textit{symbol node} corresponds to 
+rewriting all the nonterminal symbols in string that is being generated.
+
+The motivation for defining \textit{solving} of a And-or tree the way we did is that
+a found tree $T'$ corresponds to generated \lterm. 
+In order to understand this correspondence let's now talk about various tree representations
+of \lterms.
+
+\section{Tree representations of \lterms}
+\label{tree-reps}
+
+\newcommand{\sexprTree}{sexpr-tree\xspace}
+\newcommand{\atTree}{@-tree\xspace}
+
+From the definition of \lterm (\ref{deflam}) we can straightforwardly derive 
+the classical tree representation for \lterms. Term M is translated into tree $T[M]$ by following rules:
+
+\begin{itemize}
+	\item $x \in V \cup C$ translates into \textit{leaf} $x$.
+	\item $(P$ $Q)$ translates into tree\\
+		\Tree
+			[.@	
+		 		\text{$T[P]$}
+		 		\text{$T[Q]$}		 			
+			] 
+	\item $\lambda x . P$ translates into tree\\
+	 	\Tree
+			[.\text{$\lambda x$}	
+		 	 	\text{$T[P]$}	
+			] 
+\end{itemize}
+
+We can enhance this representation by compressing consecutive lambda abstractions into one
+tree node like this:
+
+\begin{itemize}
+	\item $\lambda x_1 \dots x_n . P$ translates into tree\\
+	 	\Tree
+			[.\text{$\lambda x_1 \dots x_n$ }	
+		 	 	\text{$T[P]$}	
+			] 
+\end{itemize}
+
+As this representation comes directly from definition it is evident 
+that it covers all possible \lterms.
+
+We will refer to this representation as to \textit{\atTree}.\\
+
+For representing expressions as trees it is however more common to use a little different
+representation. It will also be the representation suitable for showing 
+that \textit{solving} Inhabitation tree generates wanted \lterm.
+
+\begin{itemize}
+    \item $x \in V \cup C$ translates into \textit{leaf} $x$.
+	\item $(f M_1 M_2 \dots M_n)$ where $f \in V \cup C, n \geq 1$ translates into tree\\
+		\Tree
+			[.f	
+		 		\text{$T[M_1]$}
+		 		\text{$T[M_2]$}
+		 		\text{$\dots$}
+		 		\text{$T[M_n]$}		 				 			
+			] 
+	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
+	 	\Tree
+			[.\text{$\lambda x_1 \dots x_n$ }	
+		 	 	\text{$T[M]$}	
+			] 
+\end{itemize}
+
+Notice that this representation does not cover all \lterms, 
+e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.
+
+We will refer to this representation as to \textit{\sexprTree}.\\
+
+\Lets now consider representation for \textit{typed \lterms}.
+Straightforward approach would be to add to each node a type entry which 
+would be the type of the \lterm corresponding to subtree having this
+node as the root node.
+
+Approach more suitable for our purpose is to add a special type node above each node.
+More specifically:
+
+\Lets consider tree $t$ corresponding to a \lterm of a type
+$\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
+Then corresponding tree $TT[t]$ for typed \lterm is 
+obtained from the tree $t$ as follows:
+
+\begin{equation*}
+\mbox{ 
+TT[
+\Tree
+	[.r 	
+	  	  \text{$s_1$}
+		  \text{$s_2$}
+		  \text{$\dots$}
+		  \text{$s_n$}
+	] 
+}]=
+\mbox{
+\Tree
+	[.\text{$\sigma$ }
+	    [.r 	
+	  	  \text{$TT[s_1]$}
+		  \text{$TT[s_2]$}
+		  \text{$\dots$}
+		  \text{$TT[s_n]$}
+		]	  	
+	] 
+}
+\end{equation*}
+
+\begin{todo}
+
+& ujistit se že je v části o tree reprezentacích 
+  je napsaný co to znamená že podstrom má nějakej typ,
+  že podstromy odpovídaj podvýrazum atd
+
+ & říct že to je přesně ta notace co používá koza
+ & EXAMPLES of tree representations of \lterms  
+ & napsat transformace mezi \atTree a \sexprTree
+ & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
+   přímo \sexprTree a říct proč se tomu tak řiká 
+   (na tom místě co je v textu dřív)
+
+\end{todo}
+
+\section{ Generating terms in Long normal form by solving Inhabitation tree }
+
+Now we can finally put the pieces together. 
+Every solution to an Inhabitation tree has this just described tree form of a typed \lterm.
+
+\begin{todo}
+& !!!
+& TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
+		Things about $\beta\eta^{-1}$ normal form, etc.   
+\end{todo}
+
+\section{ Inhabitation machine }
+
+\begin{todo}
+  & DESCRIBE Inhabitation Machine...
+   && něco jako na závěrještě ve stručnosti popišme 
+      IM, strukturu ze který je odvozenej IT. 
+  & říct, že je to hezkej konstrukt vzhledem k tomu že je konečnej
+  & ale je popsanej velmi zrychla (v tý knize) a implementační
+     detaily jsou celkem vynechaný
+  & na stromový verzi jsou implementační detaily hezky vidět
+  & a navíc pro polymorfní verzi jsou už i IM nekonečný takže
+     se strácí i ta konečnostní výhoda
+  & dát tam obrázek 
+\end{todo}
+
+\chapter{ Design of our system }
+
+In this chapter we will describe our system trying to 
+dive into implementation details as little as possible.
+Some of those details will be revealed in the next chapter
+devoted to the implementation.
+
+This chapter will be
+presented in the mix between the mathematical terminology 
+described in the previous chapter and the algorithmic terminology.\\
+
+We can summarize our approach to design of 
+the system for Genetic programming with types
+by its main design goals:\\
+
+\begin{enum}
+ & It should eliminate the \textit{closure} constrain of classical GP
+   (see \ref{building-blocks}). 
+ & It should be generalization of the classical GP described in \ref{GP},
+   by which is meant that for \textit{building blocks} satisfying 
+   constrains of classical GP there should be simple setup of control parameters
+   of the system which will make the system behave in precisely the same way
+   as the classical GP behaves. 
+ & It should utilize the theory around typed $\lambda$-calculus.  
+\end{enum}
+
+\section{Individuals and Building blocks}
+
+\begin{todo}
+ & několik fází, několik reprezentací
+ & během generování se reprezentace mění 
+   && nejdří obsahuje lambdy i volný proměnný pak už ne
+      &&& vygeneruje se v \sexprTree
+      &&& ta se převedena na \atTree aby jsme provedli abstraction elimination
+ & dvě možnosti jak uchovávat jedince po vygenerování
+   && \sexprTree
+      &&& currently implemented
+   && \atTree
+      &&& currently not implemented, ale neni problém dodělat
+      &&& DOĎELAT! :)
+   && ale de mezi nima volně převádět
+      &&& díky tomu že nemaj lambda abstrakce, jinak by museli bejt ve 
+          WHNF/HNF nebo něco takovýho
+   && dusledky pro křížení
+      &&& \atTree umožňuje víc křížení než \sexprTree
+          &&&& každý co de v \sexprTree umí i \atTree ale obráceně to neplatí
+      &&& \atTree je zase ekvivalentní tomu kozovskýmu  
+ & říct že bilding blocks je prostě kontext
+\end{todo}
+
+\section{Generating individuals}
+
+%\begin{todo}
+%& Dřív se to menovalo 
+%"Our approach to \textit{solving} Inhabitation tree"
+%tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
+%termy a že je  budeme generovat pomocí 
+%řešení inhabitačních stromu.
+%\end{todo}
+
+Input for term generating algorithm is pair $(\sigma,\Gamma)$ where:\\
+
+\begin{enum}
+ & $\sigma$ is desired type of each generated individual.
+ & $\Gamma$ is \textit{building blocks} context. 
+\end{enum}~
+
+Each generated \lterms $M$ should  satisfy \GMS. \\
+
+We generate terms in long normal form.
+This is achieved by \textit{solving} inhabitation tree for $(\sigma,\Gamma)$.  
+The main design choice behind our approach to solving inhabitation trees
+is ability to be variable enough to enable choice between 
+\textit{systematic}  
+and \textit{ramped half-and-half} (described in \ref{GPgene}) 
+methods of generating terms 
+by choice of simple search strategy.\\
+
+\Lets explain what is meant by \textit{systematic} 
+method of generating terms.
+
+Systematic way of solving inhabitation tree corresponds to
+generating terms in their long normal forms
+in order from smallest to largest in number of symbols and \lheads.
+
+We will achieve this by using \textit{A* algorithm}.
+
+%That means minimizing number of and-nodes in 
+%and-or-tree view of point on inhabitation tree.
+%But since inhabitation tree is a special kind of 
+%and-or-tree with alternating layers of 
+%and-nodes (symbols and \lheads) and or-nodes 
+%(types) it is also minimizing number of 
+%size of whole inhabitation/and-or tree.
+
+\subsection{A* algorithm}
+
+A* is a general informed search algorithm used for finding least-cost 
+path from a given start state to some goal state in a state space.
+According to \cite{AIAMA} A* is the most widely known form of best-first search.
+
+By state space is meant oriented weighted graph with states as vertices and edges 
+labeled with numbers corresponding to distance or cost. 
+Edge from $s_i$ to $s_j$ means that state 
+$s_j$ is reachable from state $s_i$ in one step with cost $dist(s_i,s_j)$.
+In pseudocode we will refer to set of states reachable from 
+state $s$ in one step as to $s.nexts()$.
+
+Typically we use A* algorithm to find path from one specific start state to another 
+specific goal state. But sometimes  
+we can be interested only in finding a goal state which 
+is specified by some $isGoal$ predicate give to the algorithm as
+input (and the state space may contain many such goal states). 
+This later variant corresponds to our situation, 
+so we will continue by describing this variant.
+
+As A* traverses the state space, it follows a path of the lowest 
+expected total cost. The expectation is based on \textit{heuristic}
+function which for state $s$ predicts distance to the nearest goal state. 
+It uses a priority queue of states as its crucial data structure.
+In this queue the priority of state $s$ is 
+the total expected cost of the path from start state 
+going through $s$ and continuing to the nearest goal state. \\
+
+Here follows A* algorithm written in pseudocode:
+
+\Pseudokod{A*( start , isGoal , heuristic )}{
+	$open \la$ empty priority queue \;
+	$closed \la$ empty set \;\;
+
+	$start$.G = 0 \;
+	$start$.F = 0 + $heuristic$( $start$ ) \;\;
+
+	$open$.insert( $start$ )\;\;
+
+	\While{$\neg$ $open$.isEmpty() }{
+		$state \la$ $open$.popStateWithLowestF()\;\;
+
+		\If{ isGoal( $state$ ) }{
+			\Return $state$ \;
+		}\;
+
+		closed.insert( $state$ ) \;\;
+
+		\For{$next \in$ $state$.nexts() }{		
+		    $newG \la$ $state$.G + dist($state$,$next$)\;\;
+
+			\If{ $( next \not\in open \wedge next \not\in closed ) \vee newG < next.$G}{\;
+
+				$next$.G $\la newG$                          \;
+				$next$.F $\la newG + heuristic$( $next$ )  \;\;
+
+				\If{ $next \not\in open$ }{
+					$open$.insert( $next$ )\;
+				}
+
+			}						
+		}
+	}\;
+
+	\Return no-reachable-goal 
+}
+
+Now we will modify this general A* algorithm so
+it will be more suited for our needs. We want following 
+properties:\\
+
+\begin{enum}
+& The state space is a tree.
+& $dist(s_i,s_j) = 1$
+& We do not want to find one, but $n$ goal states. 
+\end{enum}~
+
+For state space which is a tree the condition
+$(next \not\in open \wedge next \not\in closed)$ 
+is always true, since it is possible
+to come to a state only from its parent state. 
+This fact results in simplification of algorithm's code and behavior;
+we may completely omit the $closed$ set and the checks in the for loop.\\
+
+Those changes result in the following code:
+
+\Pseudokod{Our-A*( start , n , isGoal , heuristic )}{
+
+	$open    \la$ empty priority queue \;
+	$results \la$ empty set\;\;
+
+	$start$.G = 0 \;
+	$start$.F = $heuristic$( $start$ ) \;\;
+
+	$open$.insert( $start$ )\;\;
+
+	\While{$\neg$ $open$.isEmpty() }{
+
+		$state \la$ $open$.popStateWithLowestF()\;\;
+
+		\If{ isGoal( $state$ ) }{
+
+			$results$.insert( $state$ )\;
+
+			\If{ $\vert results \vert = n$ }{
+				\Return $results$ \;
+			}
+		}\;
+
+		\For{$next \in$ $state$.nexts() }{
+			$next$.G $\la $state$.G + 1$ \;
+			$next$.F $\la $next$.G + heuristic$( $next$ )  \;
+			$open$.insert( $next$ )\;			
+		}
+	}\;
+
+	\Return results
+}
+
+The most important part of A* algorithm is the heuristic function.
+A heuristic function must be \textit{admissible} in order to make
+A* algorithm find the shortest possible path.
+
+A heuristic function $h$ is said to be \textit{admissible} if it satisfies 
+following condition for every state $s$:
+$$ h(s) \leq h^{*}(s) $$
+
+where $h^{*}(s)$ is the true length of the shortest path from $s$ to 
+the nearest goal state.
+
+In other words, a heuristic function is \textit{admissible} if it is
+\textit{optimistic}.
+
+\subsection{Our state space}
+
+% - Stavový prostor 
+%   Nedokončený term
+%   Navíc unfin typ gama
+%   my jen podmnozinu urcenou funkci nexts() a vsemi triv koreny unfin typ gama
+%   Vzdy vezmeme nejlevejsi unfin a udelame z nej konkrétnější term pomoci 
+%   vyberu v or uzlu tim nam vzniknou 
+%   nove unfin uzly odpovidajici pozadavkum and uzlu.
+%   Protoze vyplnovani jednotlivych vetvi je na sobe nezavisle 
+%   tak je jedno v jakem poradi je budem vyplnovat 
+%   a tak je nase "z leva doprava" ok.
+% - isFinal(state)
+%   Term který neobsahuje unfin uzel
+% - Heuristika 
+%   simple -pocet unfin uzlů. 
+%   Lepsi - predpocitat si to chytre s gama' a udelat si mapu v kery kdyz to neni tak dat 1.
+% - Vic nez 1 reseni
+%   staci pokracovat pri nalezeni reseni.
+
+In order to describe our state space to search in \lets introduce
+extended version of our tree representation for typed \lterms from \ref{tree-reps}.
+This extension consists in adding a new kind of leaf node $(\sigma,\Gamma)$ 
+standing for unfinished tree of a type $\sigma$ with context $\Gamma$ to use
+as set of building blocks
+\footnote{\textbf{TODO}: Tady to napsat nějak líp s tim gamma }.\\
+
+Initial start states are all of a form "one node tree $(\sigma,\Gamma)$" since we
+are trying to generate a typed \lterm of a type $\sigma$ from building blocks $\Gamma$.\\
+
+A tree with no unfinished leaf is considered as final state.  \\
+
+We will define our state space inductively by specifying algorithm for obtaining 
+successors of a state.
+
+A non-final state must have one or more unfinished leafs. 
+One of those unfinished leafs is selected. 
+It is the first one found by depth-first search (from the root of a state tree)
+i.e., the leftmost unfinished leaf (if we consider that tree written as expression).
+
+Let $(\sigma,\Gamma)$ be the selected leaf. Successor state is constructed 
+by replacing this leaf by new subtree. We must distinguish two cases of $\sigma$ :
+\textit{Atomic type} $\sigma$ and \textit{function type} $\sigma$.\\
+
+\textbf{Atomic type} $\sigma = \alpha $, $\alpha \in A$.
+
+For atomic $\sigma$ there are as many successors as there are members of $\Gamma$ of a form 
+$(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \boldsymbol{\alpha} )$,
+where $m \geq 0 $. In other words all the members of $\Gamma$ of a type
+$\alpha$ or of a type for function which returns $\alpha$.
+
+For each 
+$(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \boldsymbol{\alpha} ) \in \Gamma$,
+where $m > 0$ the new subtree which will replace $(\sigma,\Gamma)$ has following form: \\
+
+\Tree
+   [.$\alpha$
+	[.f	
+ 		\text{$(\rho_1,\Gamma)$}
+ 		\text{$(\rho_2,\Gamma)$}
+ 		\text{$\dots$} 		
+ 		\text{$(\rho_m,\Gamma)$}		 				 			
+	]   
+   ]\\
+
+And for m = 0:
+
+\Tree [.$\alpha$ f ] \\
+
+\textbf{Function type} 
+$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
+where $n \geq 1, \alpha \in A$.
+
+For function type $\sigma$ there is exactly one successor.
+
+For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet used anywhere in the whole tree.
+%included in context $\Gamma$ as variable or constant name.
+
+The new subtree which will replace $(\sigma,\Gamma)$ has following form: \\
+
+\Tree
+   [.\text{$\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$}
+	[.\text{$\lambda x_1 \dots x_n$}	
+ 		\text{$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$}		 				
+	]   
+   ]\\[3em]
+
+We should discuss that it is correct to have as successors of a state those
+created by expansion of the leftmost unfinished leaf (in contrast with
+more broad successor set where any unfinished leaf may be expanded).
+What we need to show is that no final state will be omitted by the method 
+and that there is no shorter path omitted.
+
+One can see that it is correct by considering that 
+the order of expanding unfinished leafs is irrelevant 
+since expansion of each unfinished leaf is 
+independent from expansion of any other unfinished leaf
+(up to renaming of variables which is irrelevant).
+
+Every unfinished leaf of a state must be replaced by 
+tree containing no unfinished leafs in order to be final state. 
+But since expansion of an unfinished leaf has no impact 
+(besides new variable names) on expansion of other unfinished
+leafs we may choose the order of expansions arbitrarily. \\
+
+We prefer this variant because:
+ \begin{enum}
+ 	& It has smaller or equal number successors for each state, 
+ 	  i.e. smaller or equal branching factor.
+ 	& It makes the state space a tree, in contrast with
+ 	  the naive state space which is not a tree. 
+ \end{enum}
+
+%\begin{todo}
+%Mužem na to koukat jako na takovou tu redukci simetrií, 
+%najít to v AIAMA a pořádně popsat.
+%\end{todo}
+
+\subsection{Our heuristic function}
+
+A heuristic function takes as input a state and returns as output an estimation of number
+of steps needed to get to a final state by the shortest possible path.
+
+Our heuristic function is a simple one: The number of unfinished leafs in the state tree.
+
+\begin{todo}
+& rozebrat to, ukázat že je optimistická atd...
+\end{todo}
+
+\subsection{Search strategy}
+
+Now we will describe further modification of the A* algorithm 
+which is generalizing it enough to enable use of \textit{ramped half-and-half}
+term generating method and other term generating methods.\\
+
+It is done by adding one more input argument for the algorithm.
+This additional argument is a search strategy whose purpose is
+to determine which successor states of all possible successor states 
+will be added to the priority queue, i.e., a search strategy 
+is there to filter out some successor states from $state.nexts()$.
+This filtering may be non-deterministic. It is based on
+the depth of expanded unfinished leaf and on properties of 
+the newly added subtree, e.g., whether the newly added subtree
+contains one or more unfinished leafs (so it corresponds to non-terminal) 
+or not (so it corresponds to terminal).
+
+By \textit{depth} we really mean $\lfloor depth/2 \rfloor$, because our 
+formal state trees contain those type nodes above each symbol node. 
+\\
+
+In previous situation where there is no filtering of states involved 
+we know that if queue becomes empty, then we have exhaustively searched
+all possible states.
+
+But since filtering is involved this is no longer true.
+Therefore, if the queue becomes empty before the required number of
+terms is generated, then the generating process is restarted and 
+it continues to generate terms.
+
+We also enable strategy to have some internal state which is 
+initialized (possibly non-deterministically) each time the 
+generating process restarts. \\
+
+Hopefully those changes will be clarified by the following pseudocode:
+
+\Pseudokod{A*-with-strategy( start , n , isGoal , heuristic , strategy )}{
+
+	$results \la$ empty set\;\;
+
+	\While{ $\vert results \vert < n$ }{
+
+		$strategy.init()$\;\;
+
+		$open    \la$ empty priority queue \;\;
+
+		$start$.G = 0 \;
+		$start$.F = $heuristic$( $start$ ) \;\;
+
+		$open$.insert( $start$ )\;\;
+
+		\While{$\neg$ $open$.isEmpty() }{
+
+			$state \la$ $open$.popStateWithLowestF()\;\;
+
+			\If{ isGoal( $state$ ) }{
+
+				$results$.insert( $state$ )\;
+
+				\If{ $\vert results \vert = n$ }{
+					\Return $results$ \;
+				}
+			}\;
+
+			$nexts \la strategy.filter$( $state$.nexts() , $state$.depth() ) \;\;
+
+			\For{$next \in nexts$ }{
+				$next$.G $\la $state$.G + 1$ \;
+				$next$.F $\la $next$.G + heuristic$( $next$ )  \;
+				$open$.insert( $next$ )\;			
+			}
+		}
+	}\;
+
+	\Return results
+}
+
+\Lets show examples of search strategies. 
+They are described by describing 
+their $filter(nexts,depth)$ 
+and $init()$ methods. Again, in order to avoid confusion, 
+\lets stress that $depth$ is really $\lfloor depth/2 \rfloor$ since
+our formal states contain above each symbol node a type node.
+
+\subsubsection{Systematic strategy}
+
+The previous systematic behavior may be obtained by trivial 
+$filter$ function which returns all $nexts$.
+Systematic strategy needs no internal state, 
+therefore $init$ is doing nothing.
+
+\subsubsection{Ramped half-and-half strategy}
+
+In the $init$ are randomly (uniformly) initialized two variables of the
+strategy internal state:
+
+\begin{enum}
+ & \textit{isFull} - A boolean value, determining whether \textit{full}
+                     or \textit{grow} method will be performed.
+ & \textit{d} - A integer value from $\setDots{2}{D_{init}}$, where 
+                $D_{init}$ is predefined maximal depth (e.g. 6).                    
+\end{enum}~
+
+The method $filter(nexts,depth)$ returns precisely one randomly
+(uniformly) selected  element from a subset of $nexts$
+(or zero elements if $nexts$ is empty). 
+This means that the queue always contains only one (or zero) state.
+
+The subset to select from is determined by $depth, d$ and $isFull$.\\
+
+Those elements of $nexts$ whose newly added subtree contains one ore more 
+unfinished leafs are regarded as \textit{non-terminals}, whereas 
+those whose newly added subtree contains no unfinished leaf are regarded as 
+\textit{terminals}.\\
+
+If $depth = 0$, then the subset to select from is  
+set of all \textit{non-terminals} of $nexts$.
+
+If $depth = d$, then the subset to select from is
+set of all \textit{terminals} of $nexts$.\\
+
+In other cases of $depth$ it depends on value of $isFull$.
+
+If $isFull = true$, then the subset to select from is 
+set of all \textit{non-terminals} of $nexts$.
+
+If $isFull = false$, then the subset to select from is 
+whole $nexts$ set.
+
+\begin{todo}
+& eště se musí nějak prohrotit to, že tam je navíc ta lambda hlava
+kerá ještě o něco kazí tu $depth$
+& ukázat že je to fakt to samý co kozovský klasický
+& zavýst tam pojem běhu pro ten jeden vnější while cyklus
+\end{todo}
+
+\subsubsection{Geometric strategy}
+
+We can see those two previous strategies as two extremes on the spectrum of 
+possible strategies.
+
+\textit{Systematic strategy} filters no successor state thus performing
+exhaustive search resulting in discovery of $n$ smallest terms in one run.
+
+On the other hand, \textit{ramped half-and-half strategy} filters 
+all but one successor states resulting in degradation of 
+priority queue to "fancy variable". 
+And each term is generated in its own run.\\
+
+\textit{Geometric strategy} is simple yet fairly effective term generating 
+strategy somewhere in the middle of this spectrum.
+
+It is parameterized by parametr $q \in (0,1)$.
+
+For each element of $nexts$ it is probabilistically decided whether
+it will be returned or omitted. A probability $p$ of returning is
+same for all elements, but depends on $depth$. 
+It is computed as follows:
+
+$$ p = q^{depth} $$
+
+This formula is motivated by idea that it is important to
+explore all possible root symbols, but as the $depth$ 
+increases it becomes less "dangerous" to omit 
+an exploration branch.
+
+We can see this by considering that this strategy results in
+somehow undisciplined A* search.
+With each omission we make the search space smaller. But with
+increasing depth these omissions have smaller impact on the search space,
+i.e., they cut out lesser portion of the search space.
+
+Another slightly esoteric argument supporting this formula is that "root 
+parts" of a program usually stand for more crucial parts
+with radical impact on global behavior of a program, 
+whereas "leaf parts" of a program usually
+stand for less important local parts (e.g. constants).
+
+It also plays nicely with the idea that "too big trees should be killed".\\
+
+Systematic strategy does not have an internal state, thus $init$ does nothing.
+
+\begin{todo}
+
+& sjednotit terminologii nexts successors atd
+
+& Odiskutovat nějak tohle:
+přirozeně chceme při produkci velkýho množství termu obhospodařovávat nějaký
+velký množství rozdělanejch termu v nějaký struktuře a postuipně je dodělávat.
+Proto přichází jako tahle struktura na mysl nějaká prioritní fronta,
+otázka je kde vzít ty priority a A* je jednoduchá ale přitom optimum hledající 
+odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.
+
+& určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
+a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
+udělali jeden dukaz typicky složitýho tvrzení, my sme v situaci kdy
+chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
+
+& nekde zmínit že s těma nexts nezacházíme takle dementně, že samozřejmě vybíráme 
+jen z těch přidávanejch podstromu ale že tim nechcem komplikovat popis
+
+\end{todo}
+
+\subsection{Other tried approach to generating individuals}
+
+\begin{todo}
+ & popsat tu ten puvodní přístup, kde se bralo přímo 
+   zavedení a eliminace pro každej typovej konstrukt 
+   a podle toho se pak udělali příslušný term generating grammar pravidla
+   kerý odpovídaj manipulacímse strommem
+ & ale má nevýhody
+   && hlavně to neni v normální formě, takže generujeme spoustu 
+      ruzně dopočítanejch beta-rovnáse-ekvivalentních termu
+   && u pravidla zavedení implikace neni uplně zřejmý
+      co tam použít za typ
+   && takže sice to na to ze začátku běhalo ale nakonec to bylo vyměněno
+      za IM metodu since ta je jednoduší, líp se přemejšlí o ní,
+      konstrukty jako and a or nemusíme zavádět na urovni pravidel
+      ale stačí tam dat vhodný polymorfní konstruktory/destruktory 
+ & nějak tu metodu pomenovat a přemenovat podle toho nadpis
+\end{todo}
+
+\newpage
+\section{Crossover}
+
+The crossover operation in classical GP is performed 
+by swapping randomly selected subtrees in each parent 
+\sexprTree (see \ref{GPxover}).
+
+For typed lambda terms two difficulties arise: Types and variables.
+
+We will show how to crossover typed \lterm trees in both 
+\atTree and \sexprTree notation.\\
+
+As in classical GP our crossover will be performed by swapping
+two subtrees. But now with constraint that both subtrees have
+the same type. This is discussed in grater detail in \ref{typed-swapping}.\\
+
+Variables bring more difficulties then types do.
+This problem arises from variables that are free in subterms corresponding 
+to swapped subtrees.
+
+Following example illustrates the problem. \Lets have these two
+parent trees with selected nodes in bold.\\
+
+\newcommand{\lh}[1]{\lambda #1}
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]
+
+~\\The swap of subtrees results in following trees:\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]
+
+~\\The problem is that variable $x_2$ in second tree
+is not bound by any $\lambda$-head and since
+it is not element of $\Gamma$, the second tree is not well-typed \lterm.
+
+% - This is only one example of a problem caused by swapping subtrees with variables,
+% - Resolve problems with free variables or avoid variables completely.
+
+In \ref{repairing-method} is presented more detailed discussion of problems with
+variables and method which repairs such "broken" terms.
+
+But we can avoid dealing with this problem by avoiding use of variables.
+This can be achieved by process called abstraction elimination.
+
+\subsection{Abstraction elimination}
+\label{toSKI}
+%... Conversion to SKI combinators
+
+\textit{Abstraction elimination} is a process of transforming 
+an arbitrary \lterm into \lterm witch contains no lambda abstractions
+and no bound variables.
+The newly produced \lterm may contain function applications, 
+free symbols from former \lterm and some new symbols standing for 
+combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
+
+Those combinators are defined as:
+\begin{align*}
+\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
+\Kcomb &= \lam{x\,y}{x} \\
+\Icomb &= \lam{x}{x} 
+\end{align*}
+
+\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
+
+\Lets describe transformation $\Ae$ performing this process.
+\begin{align*}
+\Ae[x]           &= x &\\[0.4em]
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
+\Ae[\lam{x}{x}]  &= \Icomb &\\
+\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
+\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
+&\textbf{if } x \in \FV(M)\\
+\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
+&\textbf{if } x \in \FV(M) \vee x \in \FV(N)\\
+\end{align*}
+
+\begin{todo}
+& přidat (ale i do kódu) B a C
+& říct že my musíme řešit i typy 
+& říct o složitosti , a že je to docela
+disadventič vzhledem k bloatu
+& citovat Simon Peyton Jones implementaition of FP languages
+jako source of further reading on this subject.
+\end{todo}
+
+\subsection{Typed subtree swapping}
+\label{typed-swapping}
+
+\subsection{ Repairing method }
+\label{repairing-method}
+
+\begin{todo}
+& !!!
+\end{todo}
+
+\subsection{ Comparison of the two methods }
+\label{comarison-ski-repairing}
+
+\begin{todo}
+& !!!
+& říct že nakonec sme přilnuli k zbavit se proměnejch
+\end{todo}
+
+\newpage
+\section{Mutation}
+
+\begin{todo}
+& říct že na tu sme se moc nesoustředili ze dvou duvodu
+  ktery to ospravedlňujou 
+	&& koza taky
+	&& ta s vygereovanim podstromu je v zásadě použití
+	   algoritmu na generovaní jedince a tak nám přišlo
+	   lepší pořádně se soustředit na generování jedince a
+	   začít do toho mutaci tahat až bude generování pořádně
+	   prozkoumáno.
+& To ale neznamená že jí nemáme radi, naopak se těšíme
+  že její vhodný zavedení celou metodu ještě zlepší.
+
+& Nápad na to jak to dělat:
+  && nutno ošetřit - co všechno do gammy alemyslim že v phodě když je uplne 
+     stejná jako ta generujicí, pač ve vy SKI-ovanym termu stejne nejsou zadny 
+     proměný 
+  && mít nějakej předem danej počet pohybu na frontě, abych mohl efektivně
+     kontrolovat že to přidání mutací moc nespomalý
+  && Po tom počtu pohybu končim generování
+  && mam nějakej počet vygenerovanejch termu
+  && pro chytrou mutaci mužu chtít třeba 10 termu
+     &&& pokud se mi jich vygenerovalo míň, tak beru míň
+     &&& pokud se mi jich vygenerovalo víc, tak vemu 10 náhodně
+     &&& těch 10 vyhodnotim (jakože dosazenejch na jejich místo)
+     &&& nejlepší si vezmu
+     &&& dobrý udělat to zase najednou pro všechny v tý generaci mutovaný
+         pánčto je nákladný spouštět Hint
+& Další mutace - beta eta redukce (?? asi moc naroční viz dale) 
+  && zmenšující mutace
+  && udělaná že převedem SKIBC do lambda termu a to redukujem? to asi ne pač
+     pak převod to spíš prodlouží, spíš asi mít redukční pravidla pro SKIBC
+     a těma redukovat, ale je otázka jak moc tam budou vznikat SKIBC-redexi
+     - ty mohou vznikat jen při @-tree reprezentaci
+  && to otevírá otázku toho, že by elementy gammy mohla mít k sobě redukční pravidla
+& Další mutace - jednoduše náhrada podstromu/uzlu za nějakej z gammy
+  && nezvětšující mutace
+& Další mutace - mutace číselnejch-like konstant
+& To mě přivádí k myšlence sice pro generování ale myslim podstatný:
+  jakmile jsou v problému číselný-like konstanty tak mužu udělat pro 
+  jeden tvar hodně jedincu - což šetří jak čas na generování, tak to prozkomává 
+  víc danej tvar
+
+\end{todo}
+
+%\subsection{Using term generation}
+
+\section{ Complexity discussion }
+
+\begin{todo}
+& !!!
+& zmínit že mi moc neřešíme formální/asymptotický složitosti
+  && de nám hlavně o realný časy
+& vyjít z komplexnosti A* asi
+  && složitost výrazně ovlivněná kvalitou heuristický funkce
+\end{todo}
+
+\chapter{Polymorphic variant}
+
+\begin{todo}
+ & poznamky o čem se mimojiné zmínit:
+   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
+      jako alternativa k ADF ! Myslim že hodně zajimavý
+\end{todo}
+
+\chapter{ Implementation }
+
+Předběžný náhodně uspořádaný body osnovy ..\\
+
+\begin{easylist}[itemize]
+& že je to v haskellu
+&& Jak moc hrotit vysvětlení haskellu?
+& že je to servrově řešený
+& architektura evolučního algoritmu 
+&& ty tři typový třídy a ta čtvrtá typová třída
+& evaluace
+&& vychytavka s vyhodnocenim celýho seznamu aby s eto nemuselo furt inicializovat - inicializace vyhodnocovače jednou za generaci 
+& hlavní soubory
+&& Eva.hs     - Eva monada
+&& GP-Core.hs - Evoluční algoritmus
+&& GP-Data.hs - Instance tříd z GP-Core
+&& IMx.hs - Term generating
+&& TTerm.hs   - typed lambda term
+&& TTree.hs   - CTT
+&& Server.hs  - main serveru
+&& Job.hs     - Registrace problemu
+&& Heval.hs   - Evaluator 
+&& složka Problems
+&&& pro každej problém složka 
+&&& ta typicky strukturu
+&&&& Problem.hs
+&&&& Funs.hs
+&& Slozka server
+&&& index.html
+&&& slozka js/Problems
+&&& .js pro kazdej problem
+\end{easylist}
+
+\section{Top level view}
+\section{Comments about main source files}
+ \subsubsection{ Eva.hs }
+ \subsubsection{ GP\_{}Core.hs }
+\section{Term generating}
+\section{Crossover}
+\section{Mutation}
+
+\section{Brief history of the project}
+\begin{todo}
+ & zmínit třeba tyhle fáze:
+  && že sem si delal evaluator sám co sem pak s performenc duvodu opustil
+  && že sem hledal intuicionistickej dokazovač ale že nejsem nenarazil na vhodnej
+     pač nejsou delany na to plivat hodně dukazu  
+  && že sem nejdriv pouzival ten vypsychlej dokazovač co tam měl nějaký ty 
+     spešl rozhodovací stromy
+  && pak sem to přepsal tak aby to nebylo jen na GP šitý ale obecně na jakoukoliv 
+     evoluci
+  && ...
+\end{todo}
+
+\chapter{Problems}
+	In this section will be presented usage of the system in order to solve specific problems.
+
+		\section{Even Parity Problem}
+		\section{Big Context}
+		\section{Fly}
+		\section{Simple Symbolic Regression}
+		\section{Artificial Ant}
+		\section{Boolean Alternate}
+
+\chapter{Comparison with others}
+
+%\chapter{Ideas yet not implemented}
+\chapter{Future work}
+
+\section{Roadmap}
+\section{Tree ants}
+\section{[Propojení těch dvou barandrechtskejch odvozovacích pravidel do jedinýho]}
+\section{[Chytřejší heuristika pro A* která si to předpočítá na $\overline{\Gamma}$ ]}
+\section{[šlechtění těch strategií na generování termů]}
+\section{[šlechtění fitness funkcí a nápad s "turnajem olympioniků"]}
+\section{[šlechtění search strategií pro prohledávání]}
+
+%\chapter{Conclusion}
+\chapter*{Conclusion}
+
+\begin{todo}
+ & !!!!!!!!!!!!!!!!!
+\end{todo}
+
+Další témata co zmínit v úvodu nebo v závěru :\\
+
+\begin{easylist}[itemize]
+
+& zmínit že existuje i jinej pohled na to proč tam zavádět typy:
+  && mužem koukat na to tak že konstrejnt je naše TuF tim, že 
+     v ní nemužou bejt libovolný funkce
+  && ale tuhle constrain mužeme omezit tim že konstrjnujeme 
+     právě použití tech bilding bloku, kde constrains jsou daný těma typama
+     - jakoby výměna global constrain za localní konstrejny na tech bilding  
+       blockách
+  && druhá možná ješte silnější motivace je že pomocí typu pak mohou vznikat 
+     smysluplnější programy, tím že dostavají vstupy které přesně čekají
+     a dávají výstupy které mají nějaké očekávané vlastnosti
+
+& Možná se zmínit o tý představě dvou pólu
+  induktivním a deduktivním -
+  klasickym GP a Dokazovačem, čim složitější
+  typovej sytem tim víc se jedná o deduktivní   
+  systém a blíží se to dokazovači.
+   && Ve světle toho, je pak hezky vidět, že
+      je chytrý exhaustivně systematicky prohledávat od nejmenšího
+      po největší- jak budeme zesložiťovat typovej systém, tak
+      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
+      jich generovat kvanta
+
+& že imlementace je delana těma vtipnejamtajpklasama umožnujicima libovolný
+  evoluční struktury na evolvovani
+
+& říct že vnekterejch aspektech je videt, že to je work in progress
+  ale že si myaslime že i tak to ukazuje pěkný výsledky 
+  a že doufame že ty nedodělky budou další prací odstraněny
+
+& Říct jak jde text:
+  && Začnem popsanim klasickýho GP podle Kozy
+  && Vyslovíme definice okolo lambda termu a typu
+  && Vyslovíme definici inhabitačního stromu
+     kterej je stromovou verzí IM od Barendrechta.
+  && Matematicky popíšem jak na to dem my.
+  && Popíšem rozšíření zakladního systému na 
+     polymorfní.
+  && Popíšem to z programátorskýho hlediska
+  && Popíšem příklady na kterejch to demonstrujem
+  && Srovnáme to s nějakejma přístupama co už 
+     existujou
+
+& Nějak naťuknout už tady v úvodu jak se to má při
+  srovnání s tim jak na to dou jiný (Yu atd)
+
+& Výkřiky co zmínit:
+  && Na co zaměřený:
+     &&& Proof of a concept
+     &&& Spíš jednoduchost
+         &&&& Ve většině aspektú to sleduje
+              Kozu - tzn i v málo řídících
+              parametru
+     &&& Duraz na generování termu (!)
+
+  && Na co nezaměřený (ale snad taky ok v tom):
+     &&& Performance   
+     &&& Mature knihovna co by šla použít robustně
+         v cizim kódu - spíš to počítá že to 
+         bude dál rozvíjeno a tohle je mezizastávka
+
+ & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
+   defining EA and then specifying the differences or something 
+   like that...
+
+& bylo by asi mimojiný dobrý 
+ospravedlnit/okomentovat tak fancy (=javascript frontend) 
+implementaci při 
+dost nedodělanejch jinejch věcech (heuristika by 
+mohla být lepší, mutace by mohli bejt, atd)
+ale věřil sem, že je to proces iterativní
+a že je po celou dobu dobrý mít na čem to testovat.
+Javascript byl vybranej pač HTML je nejpromáklejší
+UI jaký vubec je na zemi teď (state of dzí art)
+a tak to snad nebyla stráta času.
+navíc se naskitla díky tomu další pěkná příležitost
+k ukázání toho, že to neni omezený na haskell
+ale že se kombinatorový konstrukty daj 
+v zásadě přeložit do libovolnýho jazyka celkem bez problému
+nebo minimalně do JS. Ale todle možná nepsat tolik do závěru
+spíš nějaky podrobnosti přehodit do implementačních detailu
+a tady to jen tak líznout.
+
+\end{easylist}
+
+\addcontentsline{toc}{chapter}{Conclusion}
+
+\begin{thebibliography}{9}
+
+% TODO bacha neni tam jednotnej formát!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+
+\bibitem{koza92}
+
+  John R. Koza,
+  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
+  MIT Press, Cambridge, MA,
+  1992.
+
+\bibitem{koza05}
+
+  Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
+  \emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
+  Springer, 2005. ISBN 978-0-387-26417-2
+
+\bibitem{yu01}
+
+  T. Yu. 
+  \emph{Hierachical processing for evolving recursive and modular 
+        programs using higher order functions and lambda abstractions}. 
+  Genetic Programming and Evolvable Machines,
+  2(4):345–380, December 2001. ISSN 1389-2576.
+
+\bibitem{montana95}
+
+D. J. Montana. 
+\emph{Strongly typed genetic programming.} 
+Evolutionary Computation, 3(2): 199–230, 1995.
+%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha
+
+\bibitem{haynes96}
+
+T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
+\emph{Type inheritance in strongly typed genetic programming.} 
+In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
+in Genetic Programming 2, chapter 18, pages 359–376.
+MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1.\\ 
+URL 
+\url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.
+
+\bibitem{olsson94}
+
+J. R. Olsson. 
+\emph{Inductive functional programming using incremental program 
+transformation and Execution of logic programs by 
+iterative-deepening A* SLD-tree search.} 
+Dr scient thesis, University of Oslo, Norway, 1994.
+
+\bibitem{barendregt92}
+H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum , H.  P.
+\emph{Lambda Calculi with Types.} 
+Handbook of Logic in Computer Science, 1992.
+
+\bibitem{barendregt10}
+
+  Henk Barendregt, Wil Dekkers, Richard Statman,
+  \emph{Lambda Calculus With Types}.
+  Cambridge University Press,
+  2010. \\
+  URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
+
+\bibitem{jones87}
+Simon Peyton Jones. 
+\emph{The Implementation of Functional Programming Languages}. 
+Prentice Hall, 1987.
+
+\bibitem{AIAMA}
+
+	Stuart J. Russell, Peter Norvig,
+	\emph{Artificial Intelligence: A Modern Approach}.
+	Pearson Education,
+	2003.
+
+\end{thebibliography}
+
+\end{document}
+
+%-------------------------------------------------------------------------------
+%% STARY pasaže co vnich je možna eště něco co by se dalo použit
+%-------------------------------------------------------------------------------
 
@@
 %\[	
-%	( \tau , \Gamma )  \longmapsto
+%	( \tau , \Gamma )  \gar
 %	\bigg( ( \sigma \rightarrow \tau , \Gamma ) 
@@
 %\[ 
-%	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
+%	( \sigma \rightarrow \tau , \Gamma )  \gar
 %	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
@@
 %\[ 
-%	( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}}
+%	( \sigma , \Gamma )  \gar \mbox{ {\Large x}}
 %\]
@@
 %	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
-%	\longmapsto & \Big( \mbox{ {\Large $\lambda$f.}}
+%	\gar & \Big( \mbox{ {\Large $\lambda$f.}}
 %	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
@@
 %	\\
-%	\longmapsto & 
+%	\gar & 
 %	\Big( \mbox{ {\Large $\lambda$f. }}
@@
 
-\section{Long normal form}
-\label{lnf}
-
-Let \GMS where 
-$\sigma = \tau_0 \ar \tau_1 \ar \dots \ar \tau_n \ar \alpha$,
-$n \geq 0$.\\
-Then $M$ is in \textit{long normal form} if following 
-conditions are satisfied:
-
-\begin{itemize}
- \item $M$ is term of a form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$,
- \item each $M_i$ is in \textit{long normal form}.
-\end{itemize}
-
-As is shown in \cite{barendregt10} long normal form has those 
-nice properties:
-
-\begin{easylist}[itemize]
-& Every $M$ in $\beta$-normal form has a long normal form 
-$N$ such that $N \twoheadrightarrow_{\eta} M$.
-& \textbf{TODO jen zkopčeno přepsat!!} .. if M has a beta-nf, which according to Theorem 2B.4 is
-always the case, then it also has a unique lnf and this will be its unique $\beta\eta^{-1}$-nf.
-& \textbf{TODO : potřebuju eštěnějaký z tyhle knihy?}
-\end{easylist}
-
-\begin{todo}
- & popsat tu jak ji předelat na beta nf
-	pokud by se nám chtělo... 
-	V barendrech84 (ta nová kniha s žlutejma deskama) je ukázano že je 
-	eta CR (čurč roserovská) což snad stačí společně s 
-	tim tvrzeničkem z book.pdf že vezmu lnf
-	a dělám eta redukce dokud to de. až to nejde tak bych to měl mít v 
-	beta-nf
- & přepsat to nějak hezčejc, ten seznam heskejch vlastností
-\end{todo}
-
-%"Barendregt-like"
-\section{Inference and grammar rules producing terms in Long normal form}
-\label{barlike}
-
-As is said in \cite{barendregt10} there are two
-inference/grammar rules which together generate
-simply typed lambda terms in their long normal forms.
-
-Inference rule 1: 
-\[
-	\frac{\Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} \vdash M : \alpha }
-	     {\Gamma \vdash (\lambda x_1 \dots x_n . M) : 
-	     \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha }
-\]
-\\
-Proof of correctness:
-\[
-	\dfrac{
-		\dfrac
-		 {\Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} \vdash M : \alpha}
-		 {\dfrac
-		   {\Gamma \cup \{ (x_1,\tau_1),\dots,(x_{n-1},\tau_{n-1})\} \cup 
-		                \{(x_n,\tau_n) \} \vdash M : \alpha}
-		   {\dfrac{\Gamma \cup \{ (x_1,\tau_1),\dots,(x_{n-1},\tau_{n-1})\}  
-		                \vdash (\lambda x_n . M) : \tau_n \rightarrow \alpha}
-				  { \vdots }		   
-		   }
-		 }		 
-	 }
-	     {\Gamma \vdash (\lambda x_1 \dots x_n . M) : 
-	     \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha }
-\]
-\\
-... there is a grammar rule:
-\[ 
-	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large 
-	$\lambda x_1 \dots x_n .$ 
-	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
-\]
-\\
-
-Inference rule 2: 
-\[
-	\frac{ (f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma \qquad
-	       \Gamma \vdash M_1 : \rho_1 \quad
-	       \dotsm \quad
-	       \Gamma \vdash M_m : \rho_m        
-	      }
-	     {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
-\]
-\\
-Proof of correctness (\textbf{TODO REPAIR} Conceptually it is ok but there is sazba-bug somewhere): 
-\[
-   \dfrac
-     {\dfrac
-      {\dfrac
-       {\dfrac         
-         {\dfrac  
-          {\dfrac
-           {\boxed{(f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma}}
-           {\Gamma \vdash f : \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha}
-           \quad
-           \boxed{\Gamma \vdash M_1 : \rho_1} }
-          {\Gamma \vdash (f M_1) : \rho_2 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha }
-          }{\vdots} 
-         \quad 
-         \ddots }
-       {\Gamma \vdash (f M_1 \dots M_{m-2}) : \rho_{m-1} \rightarrow \rho_m \rightarrow \alpha}
-       \quad
-       \boxed{\Gamma \vdash M_{m-1} : \rho_{m-1}}  }
-      {\Gamma \vdash (f M_1 \dots M_{m-1}) : \rho_m \rightarrow \alpha}       
-      \quad 
-      \boxed{\Gamma \vdash M_m : \rho_m} }
-	 {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
-\]
-\\
-... there is a grammar rule:
-\[ 
-	( \alpha , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large f }}
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \rho_1 , \Gamma )
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  \dots
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \rho_m , \Gamma )
-	  \quad \bigg)
-\]
-
-\begin{todo}
-& SHOW correctness of those inference rules by composing them of 
-	  $E^{\rightarrow}$, $I^{\rightarrow}$ and \textit{axiom}.
-& SHOW more examples of inference rules transformed into grammar rules.
-& DESCRIBE general algorithm for this transformation.
-& TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
-& TALK ABOUT $\beta \eta^{-1}$-normal form which is generated by this method.
-& Napsat tu (nebo asi někam jinam
-	ale hlavně někam), že de generovat přímo i termy v beta normalní 
-	formě, ale že se mi to zdálo o dost pracnější díky tomu, že
-	musíme čekovat jestli neni něco toho typu i v ne atomickejch
-	vrcholech/typech...
-\end{todo}
-
-\newpage
-\section{Inhabitation tree}
-
-Now we will introduce \textit{Inhabitation tree}, structure slightly different from
-\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Barendregt.
-We can think about Inhabitation tree as about unfolded Inhabitation machine.
-The motivation for using Inhabitation trees is belief that it will help us 
-reason about generation of \lterms 
-of a given type $\sigma$ and with a given context $\Gamma$.
-
-\subsection{Definition of Inhabitation tree}
-
-\textit{Inhabitation tree} is a \textit{rooted tree}, possibly infinite. 
-It has two types of nodes:
-
-\begin{samepage}
-\begin{itemize}
-  \item Type nodes   
-  			- containing type $\sigma \in \mathbb{T}$ 
-  			- aka "OR-node" %, Nonterminal-node. 
-  \item Symbol nodes 
-  			%- containing $ x^{*} \in \mathfrak P \left(V\right) -blobost pač ordered $ or $c \in C$
-  			- containing "$\lambda$-head" (nonempty finite sequence of variable names) or constant
-  			  name. 
-  			- aka "AND-node" %, Terminal-node. - ale nemusí, proto jsem to zakomentoval
-\end{itemize}
-\end{samepage}
-
-We construct Inhabitation tree for given type $\sigma$ and context $\Gamma$.\\
-We will define Inhabitation tree by describing its construction for a given $(\sigma,\Gamma)$.
-Notice that it will closely follow the rules from \ref{barlike}:
-
-\begin{itemize}
-\item The root of Inhabitation tree for $(\sigma,\Gamma)$ is 
-      \textit{type node} with $\sigma$ as type.
-\item All \textit{type nodes} have as child nodes only \textit{symbol nodes}. 
-\item And all \textit{symbol nodes} have as child nodes only \textit{type nodes}. 
-\end{itemize}
-
-\begin{samepage}
-Now we will resolve the child nodes of the root node.
-There are two cases of $\sigma$ (recall \ref{deftype}): 
-\begin{description}
-	\item[Atomic type] 
-		$\sigma = \alpha $ where $\alpha \in A$. 
-	\item[Function type] 
-		$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
-		where $n \geq 1, \alpha \in A$.
-\end{description}
-\end{samepage}
-
-First case \textbf{Atomic type} --- i.e., $\sigma = \alpha$ where $\alpha \in A$:\\
-For every $(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha) \in \Gamma$
-where $\alpha \in A$ there is a child \textit{symbol node} of the root containing constant name $f$.
-This symbol node containing $f$ has $m$ child subtrees corresponding to Inhabitation trees for 
-$(\rho_1,\Gamma),\dots,(\rho_m,\Gamma)$.
-
-~
-
-Compare this case with corresponding grammar rule:
-\[ 
-	( \alpha , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large f }}
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \rho_1 , \Gamma )
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  \dots
-	  \mbox{ \Vtextvisiblespace[1em] } 
-	  ( \rho_m , \Gamma )
-	  \quad \bigg)
-\]
-
-Second case \textbf{Function type} --- i.e., 
-$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
-where $n~\geq~1, \alpha~\in~A$:\\
-For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.
-
-There is one and only one child \textit{symbol node} of the root containing "$\lambda$-head" 
-$\lambda x_1 \dots x_n$ which stands for sequence of variable names $(x_1,\dots,x_n)$.
-This symbol node containing $\lambda x_1 \dots x_n$
-has one and only one child subtree corresponding to Inhabitation trees for 
-$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$.
-
-~
-
-Compare this case with corresponding grammar rule:
-\[ 
-	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \longmapsto
-	\bigg( \mbox{ {\Large 
-	$\lambda x_1 \dots x_n .$ 
-	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
-\]
-~\\
-\Lets consider following $(\sigma,\Gamma)$ as a simple example:
-
-\begin{align*}
-\sigma =& ~ \mathbb{B} \rightarrow  \mathbb{B} \rightarrow  \mathbb{B} \\ 
-\Gamma =& \{ ~ true : \mathbb{B}  \\
-        &  , ~ nand :  \mathbb{B} \rightarrow \mathbb{B} \rightarrow \mathbb{B} ~ \}
-\end{align*}
-
-This particular $(\sigma,\Gamma)$ results in the following tree:\\
-
-\Tree
-[.\text{ $\mathbb{B} \rightarrow \mathbb{B} \rightarrow \mathbb{B}$ }  
-	[.\textbf{$\lambda$x_1 x_2 } 
-		[.\text{ $\mathbb{B}$ } 
-			\textbf{true}  
-			[.\textbf{nand} 
-				\qroof{ ~~ $\dotsb$ ~~ }.\text{ $\mathbb{B}$ }
-				\qroof{ ~~ $\dotsb$ ~~ }.\text{ $\mathbb{B}$ } 
-			]
-			\textbf{x_1}
-			\textbf{x_2}
-		]
-	]
-]
-
-~
-
-Second example features our well known example:
-
-\begin{align*}
-\sigma =& ~ (\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau \\ 
-\Gamma =& \{ \}
-\end{align*}
-
-Which results in following tree:
-
-\Tree
-[.\text{ $(\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau $ } 
-	[.\textbf{$\lambda$ f x }	
-		[.\text{ $\tau$ }		
-			[.\textbf{f} 
-				[.\text{ $\delta$ }
-					\textbf{x}					
-				]
-			]
-		]
-	]
-]
-
-\subsection{And-or tree and searching in Inhabitation tree}
-\Lets consider following definition of 
-\textit{And-or tree}\footnote{
-\textbf{TODO}: Mention that on WIKI there is more general definition, but 
-for our purposes is this one sufficient.}:
-
-\textit{And-or tree} is a rooted tree where each node is labeled as either \textit{and-node} 
-or\footnote{xor} \textit{or-node}.
-
-By \textit{solving} And-or tree $T$ we mean finding finite tree $T'$
-subgraph of $T$ such that it follows these conditions: 
-\begin{itemize}
-	\item The root of $T'$ is the root of $T$.
-	\item Each \textit{and-node} in $T'$ has all the child nodes as in $T$.
-	\item Each \textit{or-node}  in $T'$ has precisely one child 
-	      node.\footnote{\textbf{TODO}: MENTION why precisely one 
-	      and not at least one ..or CHANGE the def. }   
-\end{itemize}
-~
-\Lets now consider following labeling of Inhabitation tree:
-
-\begin{itemize}
-  \item \textbf{Type nodes}   are labeled as \textbf{or-nodes}.   
-  \item \textbf{Symbol nodes} are labeled as \textbf{and-nodes}.
-\end{itemize}
-
-This labeling has following justification:
-
-Selection of exactly one
-child node in \textit{type node} corresponds to selection of exactly one
-grammar rule in order to rewrite nonterminal symbol.
-
-Selection of all the child nodes in \textit{symbol node} corresponds to 
-rewriting all the nonterminal symbols in string that is being generated.
-
-The motivation for defining \textit{solving} of a And-or tree the way we did is that
-a found tree $T'$ corresponds to generated \lterm. 
-In order to understand this correspondence let's now talk about various tree representations
-of \lterms.
-
-\section{Tree representations of \lterms}
-\label{tree-reps}
-
-\newcommand{\sexprTree}{sexpr-tree\xspace}
-\newcommand{\atTree}{@-tree\xspace}
-
-From the definition of \lterm (\ref{deflam}) we can straightforwardly derive 
-the classical tree representation for \lterms. Term M is translated into tree $T[M]$ by following rules:
-
-\begin{itemize}
-	\item $x \in V \cup C$ translates into \textit{leaf} $x$.
-	\item $(P$ $Q)$ translates into tree\\
-		\Tree
-			[.@	
-		 		\text{$T[P]$}
-		 		\text{$T[Q]$}		 			
-			] 
-	\item $\lambda x . P$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x$}	
-		 	 	\text{$T[P]$}	
-			] 
-\end{itemize}
-
-We can enhance this representation by compressing consecutive lambda abstractions into one
-tree node like this:
-
-\begin{itemize}
-	\item $\lambda x_1 \dots x_n . P$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x_1 \dots x_n$ }	
-		 	 	\text{$T[P]$}	
-			] 
-\end{itemize}
-
-As this representation comes directly from definition it is evident 
-that it covers all possible \lterms.
-
-We will refer to this representation as to \textit{\atTree}.\\
-
-For representing expressions as trees it is however more common to use a little different
-representation. It will also be the representation suitable for showing 
-that \textit{solving} Inhabitation tree generates wanted \lterm.
-
-\begin{itemize}
-    \item $x \in V \cup C$ translates into \textit{leaf} $x$.
-	\item $(f M_1 M_2 \dots M_n)$ where $f \in V \cup C, n \geq 1$ translates into tree\\
-		\Tree
-			[.f	
-		 		\text{$T[M_1]$}
-		 		\text{$T[M_2]$}
-		 		\text{$\dots$}
-		 		\text{$T[M_n]$}		 				 			
-			] 
-	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x_1 \dots x_n$ }	
-		 	 	\text{$T[M]$}	
-			] 
-\end{itemize}
-
-Notice that this representation does not cover all \lterms, 
-e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.
-
-We will refer to this representation as to \textit{\sexprTree}.\\
-
-\Lets now consider representation for \textit{typed \lterms}.
-Straightforward approach would be to add to each node a type entry which 
-would be the type of the \lterm corresponding to subtree having this
-node as the root node.
-
-Approach more suitable for our purpose is to add a special type node above each node.
-More specifically:
-
-\Lets consider tree $t$ corresponding to a \lterm of a type
-$\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
-Then corresponding tree $TT[t]$ for typed \lterm is 
-obtained from the tree $t$ as follows:
-
-\begin{equation*}
-\mbox{ 
-TT[
-\Tree
-	[.r 	
-	  	  \text{$s_1$}
-		  \text{$s_2$}
-		  \text{$\dots$}
-		  \text{$s_n$}
-	] 
-}]=
-\mbox{
-\Tree
-	[.\text{$\sigma$ }
-	    [.r 	
-	  	  \text{$TT[s_1]$}
-		  \text{$TT[s_2]$}
-		  \text{$\dots$}
-		  \text{$TT[s_n]$}
-		]	  	
-	] 
-}
-\end{equation*}
-
-\begin{todo}
-
-& ujistit se že je v části o tree reprezentacích 
-  je napsaný co to znamená že podstrom má nějakej typ,
-  že podstromy odpovídaj podvýrazum atd
-
- & říct že to je přesně ta notace co používá koza
- & EXAMPLES of tree representations of \lterms  
- & napsat transformace mezi \atTree a \sexprTree
- & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
-   přímo \sexprTree a říct proč se tomu tak řiká 
-   (na tom místě co je v textu dřív)
-
-\end{todo}
-
-\section{ Generating terms in Long normal form by solving Inhabitation tree }
-
-Now we can finally put the pieces together. 
-Every solution to an Inhabitation tree has this just described tree form of a typed \lterm.
-
-\begin{todo}
-& !!!
-& TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
-		Things about $\beta\eta^{-1}$ normal form, etc.   
-\end{todo}
-
-\section{ Inhabitation machine }
-
-\begin{todo}
-  & DESCRIBE Inhabitation Machine...
-   && něco jako na závěrještě ve stručnosti popišme 
-      IM, strukturu ze který je odvozenej IT. 
-  & říct, že je to hezkej konstrukt vzhledem k tomu že je konečnej
-  & ale je popsanej velmi zrychla (v tý knize) a implementační
-     detaily jsou celkem vynechaný
-  & na stromový verzi jsou implementační detaily hezky vidět
-  & a navíc pro polymorfní verzi jsou už i IM nekonečný takže
-     se strácí i ta konečnostní výhoda
-  & dát tam obrázek 
-\end{todo}
-
-\chapter{ Design of our system }
-
-In this chapter we will describe our system trying to 
-dive into implementation details as little as possible.
-Some of those details will be revealed in the next chapter
-devoted to the implementation.
-
-This chapter will be
-presented in the mix between the mathematical terminology 
-described in the previous chapter and the algorithmic terminology.\\
-
-We can summarize our approach to design of 
-the system for Genetic programming with types
-by its main design goals:\\
-
-\begin{enum}
- & It should eliminate the \textit{closure} constrain of classical GP
-   (see \ref{building-blocks}). 
- & It should be generalization of the classical GP described in \ref{GP},
-   by which is meant that for \textit{building blocks} satisfying 
-   constrains of classical GP there should be simple setup of control parameters
-   of the system which will make the system behave in precisely the same way
-   as the classical GP behaves. 
- & It should utilize the theory around typed $\lambda$-calculus.  
-\end{enum}
-
-\section{Individuals and Building blocks}
-
-\begin{todo}
- & několik fází, několik reprezentací
- & během generování se reprezentace mění 
-   && nejdří obsahuje lambdy i volný proměnný pak už ne
-      &&& vygeneruje se v \sexprTree
-      &&& ta se převedena na \atTree aby jsme provedli abstraction elimination
- & dvě možnosti jak uchovávat jedince po vygenerování
-   && \sexprTree
-      &&& currently implemented
-   && \atTree
-      &&& currently not implemented, ale neni problém dodělat
-      &&& DOĎELAT! :)
-   && ale de mezi nima volně převádět
-      &&& díky tomu že nemaj lambda abstrakce, jinak by museli bejt ve 
-          WHNF/HNF nebo něco takovýho
-   && dusledky pro křížení
-      &&& \atTree umožňuje víc křížení než \sexprTree
-          &&&& každý co de v \sexprTree umí i \atTree ale obráceně to neplatí
-      &&& \atTree je zase ekvivalentní tomu kozovskýmu  
- & říct že bilding blocks je prostě kontext
-\end{todo}
-
-\section{Generating individuals}
+%\section{-OLD-Inference and grammar rules producing terms in Long normal form}
+%
+%As is said in \cite{barendregt10} there are two
+%inference/grammar rules which together generate
+%simply typed lambda terms in their long normal forms.
+%
+%
+%Inference rule 1: 
+%\[
+%	\frac{\Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} \vdash M : \alpha }
+%	     {\Gamma \vdash (\lambda x_1 \dots x_n . M) : 
+%	     \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha }
+%\]
+%\\
+%Proof of correctness:
+%\[
+%	\dfrac{
+%		\dfrac
+%		 {\Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} \vdash M : \alpha}
+%		 {\dfrac
+%		   {\Gamma \cup \{ (x_1,\tau_1),\dots,(x_{n-1},\tau_{n-1})\} \cup 
+%		                \{(x_n,\tau_n) \} \vdash M : \alpha}
+%		   {\dfrac{\Gamma \cup \{ (x_1,\tau_1),\dots,(x_{n-1},\tau_{n-1})\}  
+%		                \vdash (\lambda x_n . M) : \tau_n \rightarrow \alpha}
+%				  { \vdots }		   
+%		   }
+%		 }		 
+%	 }
+%	     {\Gamma \vdash (\lambda x_1 \dots x_n . M) : 
+%	     \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha }
+%\]
+%\\
+%... there is a grammar rule:
+%\[ 
+%	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \gar
+%	\bigg( \mbox{ {\Large 
+%	$\lambda x_1 \dots x_n .$ 
+%	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
+%\]
+%\\
+%
+%Inference rule 2: 
+%\[
+%	\frac{ (f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma \qquad
+%	       \Gamma \vdash M_1 : \rho_1 \quad
+%	       \dotsm \quad
+%	       \Gamma \vdash M_m : \rho_m        
+%	      }
+%	     {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
+%\]
+%\\
+%Proof of correctness (\textbf{TODO REPAIR} Conceptually it is ok but there is sazba-bug somewhere): 
+%\[
+%   \dfrac
+%     {\dfrac
+%      {\dfrac
+%       {\dfrac         
+%         {\dfrac  
+%          {\dfrac
+%           {\boxed{(f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma}}
+%           {\Gamma \vdash f : \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha}
+%           \quad
+%           \boxed{\Gamma \vdash M_1 : \rho_1} }
+%          {\Gamma \vdash (f M_1) : \rho_2 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha }
+%          }{\vdots} 
+%         \quad 
+%         \ddots }
+%       {\Gamma \vdash (f M_1 \dots M_{m-2}) : \rho_{m-1} \rightarrow \rho_m \rightarrow \alpha}
+%       \quad
+%       \boxed{\Gamma \vdash M_{m-1} : \rho_{m-1}}  }
+%      {\Gamma \vdash (f M_1 \dots M_{m-1}) : \rho_m \rightarrow \alpha}       
+%      \quad 
+%      \boxed{\Gamma \vdash M_m : \rho_m} }
+%	 {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
+%\]
+%\\
+%... there is a grammar rule:
+%\[ 
+%	( \alpha , \Gamma )  \gar
+%	\bigg( \mbox{ {\Large f }}
+%	  \mbox{ \Vtextvisiblespace[1em] } 
+%	  ( \rho_1 , \Gamma )
+%	  \mbox{ \Vtextvisiblespace[1em] } 
+%	  \dots
+%	  \mbox{ \Vtextvisiblespace[1em] } 
+%	  ( \rho_m , \Gamma )
+%	  \quad \bigg)
+%\]
 
 %\begin{todo}
-%& Dřív se to menovalo 
-%"Our approach to \textit{solving} Inhabitation tree"
-%tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
-%termy a že je  budeme generovat pomocí 
-%řešení inhabitačních stromu.
+%& SHOW correctness of those inference rules by composing them of 
+%	  $E^{\rightarrow}$, $I^{\rightarrow}$ and \textit{axiom}.
+%& SHOW more examples of inference rules transformed into grammar rules.
+%& DESCRIBE general algorithm for this transformation.
+%& TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
+%& TALK ABOUT $\beta \eta^{-1}$-normal form which is generated by this method.
+%& Napsat tu (nebo asi někam jinam
+%	ale hlavně někam), že de generovat přímo i termy v beta normalní 
+%	formě, ale že se mi to zdálo o dost pracnější díky tomu, že
+%	musíme čekovat jestli neni něco toho typu i v ne atomickejch
+%	vrcholech/typech...
 %\end{todo}
-
-Input for term generating algorithm is pair $(\sigma,\Gamma)$ where:\\
-
-\begin{enum}
- & $\sigma$ is desired type of each generated individual.
- & $\Gamma$ is \textit{building blocks} context. 
-\end{enum}~
-
-Each generated \lterms $M$ should  satisfy \GMS. \\
-
-We generate terms in long normal form.
-This is achieved by \textit{solving} inhabitation tree for $(\sigma,\Gamma)$.  
-The main design choice behind our approach to solving inhabitation trees
-is ability to be variable enough to enable choice between 
-\textit{systematic}  
-and \textit{ramped half-and-half} (described in \ref{GPgene}) 
-methods of generating terms 
-by choice of simple search strategy.\\
-
-\Lets explain what is meant by \textit{systematic} 
-method of generating terms.
-
-Systematic way of solving inhabitation tree corresponds to
-generating terms in their long normal forms
-in order from smallest to largest in number of symbols and \lheads.
-
-We will achieve this by using \textit{A* algorithm}.
-
-%That means minimizing number of and-nodes in 
-%and-or-tree view of point on inhabitation tree.
-%But since inhabitation tree is a special kind of 
-%and-or-tree with alternating layers of 
-%and-nodes (symbols and \lheads) and or-nodes 
-%(types) it is also minimizing number of 
-%size of whole inhabitation/and-or tree.
-
-\subsection{A* algorithm}
-
-A* is a general informed search algorithm used for finding least-cost 
-path from a given start state to some goal state in a state space.
-According to \cite{AIAMA} A* is the most widely known form of best-first search.
-
-By state space is meant oriented weighted graph with states as vertices and edges 
-labeled with numbers corresponding to distance or cost. 
-Edge from $s_i$ to $s_j$ means that state 
-$s_j$ is reachable from state $s_i$ in one step with cost $dist(s_i,s_j)$.
-In pseudocode we will refer to set of states reachable from 
-state $s$ in one step as to $s.nexts()$.
-
-Typically we use A* algorithm to find path from one specific start state to another 
-specific goal state. But sometimes  
-we can be interested only in finding a goal state which 
-is specified by some $isGoal$ predicate give to the algorithm as
-input (and the state space may contain many such goal states). 
-This later variant corresponds to our situation, 
-so we will continue by describing this variant.
-
-As A* traverses the state space, it follows a path of the lowest 
-expected total cost. The expectation is based on \textit{heuristic}
-function which for state $s$ predicts distance to the nearest goal state. 
-It uses a priority queue of states as its crucial data structure.
-In this queue the priority of state $s$ is 
-the total expected cost of the path from start state 
-going through $s$ and continuing to the nearest goal state. \\
-
-Here follows A* algorithm written in pseudocode:
-
-\Pseudokod{A*( start , isGoal , heuristic )}{
-	$open \la$ empty priority queue \;
-	$closed \la$ empty set \;\;
-
-	$start$.G = 0 \;
-	$start$.F = 0 + $heuristic$( $start$ ) \;\;
-
-	$open$.insert( $start$ )\;\;
-
-	\While{$\neg$ $open$.isEmpty() }{
-		$state \la$ $open$.popStateWithLowestF()\;\;
-
-		\If{ isGoal( $state$ ) }{
-			\Return $state$ \;
-		}\;
-
-		closed.insert( $state$ ) \;\;
-
-		\For{$next \in$ $state$.nexts() }{		
-		    $newG \la$ $state$.G + dist($state$,$next$)\;\;
-
-			\If{ $( next \not\in open \wedge next \not\in closed ) \vee newG < next.$G}{\;
-
-				$next$.G $\la newG$                          \;
-				$next$.F $\la newG + heuristic$( $next$ )  \;\;
-
-				\If{ $next \not\in open$ }{
-					$open$.insert( $next$ )\;
-				}
-
-			}						
-		}
-	}\;
-
-	\Return no-reachable-goal 
-}
-
-Now we will modify this general A* algorithm so
-it will be more suited for our needs. We want following 
-properties:\\
-
-\begin{enum}
-& The state space is a tree.
-& $dist(s_i,s_j) = 1$
-& We do not want to find one, but $n$ goal states. 
-\end{enum}~
-
-For state space which is a tree the condition
-$(next \not\in open \wedge next \not\in closed)$ 
-is always true, since it is possible
-to come to a state only from its parent state. 
-This fact results in simplification of algorithm's code and behavior;
-we may completely omit the $closed$ set and the checks in the for loop.\\
-
-Those changes result in the following code:
-
-\Pseudokod{Our-A*( start , n , isGoal , heuristic )}{
-
-	$open    \la$ empty priority queue \;
-	$results \la$ empty set\;\;
-
-	$start$.G = 0 \;
-	$start$.F = $heuristic$( $start$ ) \;\;
-
-	$open$.insert( $start$ )\;\;
-
-	\While{$\neg$ $open$.isEmpty() }{
-
-		$state \la$ $open$.popStateWithLowestF()\;\;
-
-		\If{ isGoal( $state$ ) }{
-
-			$results$.insert( $state$ )\;
-
-			\If{ $\vert results \vert = n$ }{
-				\Return $results$ \;
-			}
-		}\;
-
-		\For{$next \in$ $state$.nexts() }{
-			$next$.G $\la $state$.G + 1$ \;
-			$next$.F $\la $next$.G + heuristic$( $next$ )  \;
-			$open$.insert( $next$ )\;			
-		}
-	}\;
-
-	\Return results
-}
-
-The most important part of A* algorithm is the heuristic function.
-A heuristic function must be \textit{admissible} in order to make
-A* algorithm find the shortest possible path.
-
-A heuristic function $h$ is said to be \textit{admissible} if it satisfies 
-following condition for every state $s$:
-$$ h(s) \leq h^{*}(s) $$
-
-where $h^{*}(s)$ is the true length of the shortest path from $s$ to 
-the nearest goal state.
-
-In other words, a heuristic function is \textit{admissible} if it is
-\textit{optimistic}.
-
-\subsection{Our state space}
-
-% - Stavový prostor 
-%   Nedokončený term
-%   Navíc unfin typ gama
-%   my jen podmnozinu urcenou funkci nexts() a vsemi triv koreny unfin typ gama
-%   Vzdy vezmeme nejlevejsi unfin a udelame z nej konkrétnější term pomoci 
-%   vyberu v or uzlu tim nam vzniknou 
-%   nove unfin uzly odpovidajici pozadavkum and uzlu.
-%   Protoze vyplnovani jednotlivych vetvi je na sobe nezavisle 
-%   tak je jedno v jakem poradi je budem vyplnovat 
-%   a tak je nase "z leva doprava" ok.
-% - isFinal(state)
-%   Term který neobsahuje unfin uzel
-% - Heuristika 
-%   simple -pocet unfin uzlů. 
-%   Lepsi - predpocitat si to chytre s gama' a udelat si mapu v kery kdyz to neni tak dat 1.
-% - Vic nez 1 reseni
-%   staci pokracovat pri nalezeni reseni.
-
-In order to describe our state space to search in \lets introduce
-extended version of our tree representation for typed \lterms from \ref{tree-reps}.
-This extension consists in adding a new kind of leaf node $(\sigma,\Gamma)$ 
-standing for unfinished tree of a type $\sigma$ with context $\Gamma$ to use
-as set of building blocks
-\footnote{\textbf{TODO}: Tady to napsat nějak líp s tim gamma }.\\
-
-Initial start states are all of a form "one node tree $(\sigma,\Gamma)$" since we
-are trying to generate a typed \lterm of a type $\sigma$ from building blocks $\Gamma$.\\
-
-A tree with no unfinished leaf is considered as final state.  \\
-
-We will define our state space inductively by specifying algorithm for obtaining 
-successors of a state.
-
-A non-final state must have one or more unfinished leafs. 
-One of those unfinished leafs is selected. 
-It is the first one found by depth-first search (from the root of a state tree)
-i.e., the leftmost unfinished leaf (if we consider that tree written as expression).
-
-Let $(\sigma,\Gamma)$ be the selected leaf. Successor state is constructed 
-by replacing this leaf by new subtree. We must distinguish two cases of $\sigma$ :
-\textit{Atomic type} $\sigma$ and \textit{function type} $\sigma$.\\
-
-\textbf{Atomic type} $\sigma = \alpha $, $\alpha \in A$.
-
-For atomic $\sigma$ there are as many successors as there are members of $\Gamma$ of a form 
-$(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \boldsymbol{\alpha} )$,
-where $m \geq 0 $. In other words all the members of $\Gamma$ of a type
-$\alpha$ or of a type for function which returns $\alpha$.
-
-For each 
-$(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \boldsymbol{\alpha} ) \in \Gamma$,
-where $m > 0$ the new subtree which will replace $(\sigma,\Gamma)$ has following form: \\
-
-\Tree
-   [.$\alpha$
-	[.f	
- 		\text{$(\rho_1,\Gamma)$}
- 		\text{$(\rho_2,\Gamma)$}
- 		\text{$\dots$} 		
- 		\text{$(\rho_m,\Gamma)$}		 				 			
-	]   
-   ]\\
-
-And for m = 0:
-
-\Tree [.$\alpha$ f ] \\
-
-\textbf{Function type} 
-$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
-where $n \geq 1, \alpha \in A$.
-
-For function type $\sigma$ there is exactly one successor.
-
-For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet used anywhere in the whole tree.
-%included in context $\Gamma$ as variable or constant name.
-
-The new subtree which will replace $(\sigma,\Gamma)$ has following form: \\
-
-\Tree
-   [.\text{$\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$}
-	[.\text{$\lambda x_1 \dots x_n$}	
- 		\text{$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$}		 				
-	]   
-   ]\\[3em]
-
-We should discuss that it is correct to have as successors of a state those
-created by expansion of the leftmost unfinished leaf (in contrast with
-more broad successor set where any unfinished leaf may be expanded).
-What we need to show is that no final state will be omitted by the method 
-and that there is no shorter path omitted.
-
-One can see that it is correct by considering that 
-the order of expanding unfinished leafs is irrelevant 
-since expansion of each unfinished leaf is 
-independent from expansion of any other unfinished leaf
-(up to renaming of variables which is irrelevant).
-
-Every unfinished leaf of a state must be replaced by 
-tree containing no unfinished leafs in order to be final state. 
-But since expansion of an unfinished leaf has no impact 
-(besides new variable names) on expansion of other unfinished
-leafs we may choose the order of expansions arbitrarily. \\
-
-We prefer this variant because:
- \begin{enum}
- 	& It has smaller or equal number successors for each state, 
- 	  i.e. smaller or equal branching factor.
- 	& It makes the state space a tree, in contrast with
- 	  the naive state space which is not a tree. 
- \end{enum}
-
-%\begin{todo}
-%Mužem na to koukat jako na takovou tu redukci simetrií, 
-%najít to v AIAMA a pořádně popsat.
-%\end{todo}
-
-\subsection{Our heuristic function}
-
-A heuristic function takes as input a state and returns as output an estimation of number
-of steps needed to get to a final state by the shortest possible path.
-
-Our heuristic function is a simple one: The number of unfinished leafs in the state tree.
-
-\begin{todo}
-& rozebrat to, ukázat že je optimistická atd...
-\end{todo}
-
-\subsection{Search strategy}
-
-Now we will describe further modification of the A* algorithm 
-which is generalizing it enough to enable use of \textit{ramped half-and-half}
-term generating method and other term generating methods.\\
-
-It is done by adding one more input argument for the algorithm.
-This additional argument is a search strategy whose purpose is
-to determine which successor states of all possible successor states 
-will be added to the priority queue, i.e., a search strategy 
-is there to filter out some successor states from $state.nexts()$.
-This filtering may be non-deterministic. It is based on
-the depth of expanded unfinished leaf and on properties of 
-the newly added subtree, e.g., whether the newly added subtree
-contains one or more unfinished leafs (so it corresponds to non-terminal) 
-or not (so it corresponds to terminal).
-
-By \textit{depth} we really mean $\lfloor depth/2 \rfloor$, because our 
-formal state trees contain those type nodes above each symbol node. 
-\\
-
-In previous situation where there is no filtering of states involved 
-we know that if queue becomes empty, then we have exhaustively searched
-all possible states.
-
-But since filtering is involved this is no longer true.
-Therefore, if the queue becomes empty before the required number of
-terms is generated, then the generating process is restarted and 
-it continues to generate terms.
-
-We also enable strategy to have some internal state which is 
-initialized (possibly non-deterministically) each time the 
-generating process restarts. \\
-
-Hopefully those changes will be clarified by the following pseudocode:
-
-\Pseudokod{A*-with-strategy( start , n , isGoal , heuristic , strategy )}{
-
-	$results \la$ empty set\;\;
-
-	\While{ $\vert results \vert < n$ }{
-
-		$strategy.init()$\;\;
-
-		$open    \la$ empty priority queue \;\;
-
-		$start$.G = 0 \;
-		$start$.F = $heuristic$( $start$ ) \;\;
-
-		$open$.insert( $start$ )\;\;
-
-		\While{$\neg$ $open$.isEmpty() }{
-
-			$state \la$ $open$.popStateWithLowestF()\;\;
-
-			\If{ isGoal( $state$ ) }{
-
-				$results$.insert( $state$ )\;
-
-				\If{ $\vert results \vert = n$ }{
-					\Return $results$ \;
-				}
-			}\;
-
-			$nexts \la strategy.filter$( $state$.nexts() , $state$.depth() ) \;\;
-
-			\For{$next \in nexts$ }{
-				$next$.G $\la $state$.G + 1$ \;
-				$next$.F $\la $next$.G + heuristic$( $next$ )  \;
-				$open$.insert( $next$ )\;			
-			}
-		}
-	}\;
-
-	\Return results
-}
-
-\Lets show examples of search strategies. 
-They are described by describing 
-their $filter(nexts,depth)$ 
-and $init()$ methods. Again, in order to avoid confusion, 
-\lets stress that $depth$ is really $\lfloor depth/2 \rfloor$ since
-our formal states contain above each symbol node a type node.
-
-\subsubsection{Systematic strategy}
-
-The previous systematic behavior may be obtained by trivial 
-$filter$ function which returns all $nexts$.
-Systematic strategy needs no internal state, 
-therefore $init$ is doing nothing.
-
-\subsubsection{Ramped half-and-half strategy}
-
-In the $init$ are randomly (uniformly) initialized two variables of the
-strategy internal state:
-
-\begin{enum}
- & \textit{isFull} - A boolean value, determining whether \textit{full}
-                     or \textit{grow} method will be performed.
- & \textit{d} - A integer value from $\setDots{2}{D_{init}}$, where 
-                $D_{init}$ is predefined maximal depth (e.g. 6).                    
-\end{enum}~
-
-The method $filter(nexts,depth)$ returns precisely one randomly
-(uniformly) selected  element from a subset of $nexts$
-(or zero elements if $nexts$ is empty). 
-This means that the queue always contains only one (or zero) state.
-
-The subset to select from is determined by $depth, d$ and $isFull$.\\
-
-Those elements of $nexts$ whose newly added subtree contains one ore more 
-unfinished leafs are regarded as \textit{non-terminals}, whereas 
-those whose newly added subtree contains no unfinished leaf are regarded as 
-\textit{terminals}.\\
-
-If $depth = 0$, then the subset to select from is  
-set of all \textit{non-terminals} of $nexts$.
-
-If $depth = d$, then the subset to select from is
-set of all \textit{terminals} of $nexts$.\\
-
-In other cases of $depth$ it depends on value of $isFull$.
-
-If $isFull = true$, then the subset to select from is 
-set of all \textit{non-terminals} of $nexts$.
-
-If $isFull = false$, then the subset to select from is 
-whole $nexts$ set.
-
-\begin{todo}
-& eště se musí nějak prohrotit to, že tam je navíc ta lambda hlava
-kerá ještě o něco kazí tu $depth$
-& ukázat že je to fakt to samý co kozovský klasický
-& zavýst tam pojem běhu pro ten jeden vnější while cyklus
-\end{todo}
-
-\subsubsection{Geometric strategy}
-
-We can see those two previous strategies as two extremes on the spectrum of 
-possible strategies.
-
-\textit{Systematic strategy} filters no successor state thus performing
-exhaustive search resulting in discovery of $n$ smallest terms in one run.
-
-On the other hand, \textit{ramped half-and-half strategy} filters 
-all but one successor states resulting in degradation of 
-priority queue to "fancy variable". 
-And each term is generated in its own run.\\
-
-\textit{Geometric strategy} is simple yet fairly effective term generating 
-strategy somewhere in the middle of this spectrum.
-
-It is parameterized by parametr $q \in (0,1)$.
-
-For each element of $nexts$ it is probabilistically decided whether
-it will be returned or omitted. A probability $p$ of returning is
-same for all elements, but depends on $depth$. 
-It is computed as follows:
-
-$$ p = q^{depth} $$
-
-This formula is motivated by idea that it is important to
-explore all possible root symbols, but as the $depth$ 
-increases it becomes less "dangerous" to omit 
-an exploration branch.
-
-We can see this by considering that this strategy results in
-somehow undisciplined A* search.
-With each omission we make the search space smaller. But with
-increasing depth these omissions have smaller impact on the search space,
-i.e., they cut out lesser portion of the search space.
-
-Another slightly esoteric argument supporting this formula is that "root 
-parts" of a program usually stand for more crucial parts
-with radical impact on global behavior of a program, 
-whereas "leaf parts" of a program usually
-stand for less important local parts (e.g. constants).
-
-It also plays nicely with the idea that "too big trees should be killed".\\
-
-Systematic strategy does not have an internal state, thus $init$ does nothing.
-
-\begin{todo}
-
-& sjednotit terminologii nexts successors atd
-
-& Odiskutovat nějak tohle:
-přirozeně chceme při produkci velkýho množství termu obhospodařovávat nějaký
-velký množství rozdělanejch termu v nějaký struktuře a postuipně je dodělávat.
-Proto přichází jako tahle struktura na mysl nějaká prioritní fronta,
-otázka je kde vzít ty priority a A* je jednoduchá ale přitom optimum hledající 
-odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.
-
-& určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
-a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
-udělali jeden dukaz typicky složitýho tvrzení, my sme v situaci kdy
-chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
-
-& nekde zmínit že s těma nexts nezacházíme takle dementně, že samozřejmě vybíráme 
-jen z těch přidávanejch podstromu ale že tim nechcem komplikovat popis
-
-\end{todo}
-
-\subsection{Other tried approach to generating individuals}
-
-\begin{todo}
- & popsat tu ten puvodní přístup, kde se bralo přímo 
-   zavedení a eliminace pro každej typovej konstrukt 
-   a podle toho se pak udělali příslušný term generating grammar pravidla
-   kerý odpovídaj manipulacímse strommem
- & ale má nevýhody
-   && hlavně to neni v normální formě, takže generujeme spoustu 
-      ruzně dopočítanejch beta-rovnáse-ekvivalentních termu
-   && u pravidla zavedení implikace neni uplně zřejmý
-      co tam použít za typ
-   && takže sice to na to ze začátku běhalo ale nakonec to bylo vyměněno
-      za IM metodu since ta je jednoduší, líp se přemejšlí o ní,
-      konstrukty jako and a or nemusíme zavádět na urovni pravidel
-      ale stačí tam dat vhodný polymorfní konstruktory/destruktory 
- & nějak tu metodu pomenovat a přemenovat podle toho nadpis
-\end{todo}
-
-\newpage
-\section{Crossover}
-
-The crossover operation in classical GP is performed 
-by swapping randomly selected subtrees in each parent 
-\sexprTree (see \ref{GPxover}).
-
-For typed lambda terms two difficulties arise: Types and variables.
-
-We will show how to crossover typed \lterm trees in both 
-\atTree and \sexprTree notation.\\
-
-As in classical GP our crossover will be performed by swapping
-two subtrees. But now with constraint that both subtrees have
-the same type. This is discussed in grater detail in \ref{typed-swapping}.\\
-
-Variables bring more difficulties then types do.
-This problem arises from variables that are free in subterms corresponding 
-to swapped subtrees.
-
-Following example illustrates the problem. \Lets have these two
-parent trees with selected nodes in bold.\\
-
-\newcommand{\lh}[1]{\lambda #1}
-
-\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
-\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]
-
-~\\The swap of subtrees results in following trees:\\
-
-\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
-\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]
-
-~\\The problem is that variable $x_2$ in second tree
-is not bound by any $\lambda$-head and since
-it is not element of $\Gamma$, the second tree is not well-typed \lterm.
-
-% - This is only one example of a problem caused by swapping subtrees with variables,
-% - Resolve problems with free variables or avoid variables completely.
-
-In \ref{repairing-method} is presented more detailed discussion of problems with
-variables and method which repairs such "broken" terms.
-
-But we can avoid dealing with this problem by avoiding use of variables.
-This can be achieved by process called abstraction elimination.
-
-\subsection{Abstraction elimination}
-\label{toSKI}
-%... Conversion to SKI combinators
-
-\textit{Abstraction elimination} is a process of transforming 
-an arbitrary \lterm into \lterm witch contains no lambda abstractions
-and no bound variables.
-The newly produced \lterm may contain function applications, 
-free symbols from former \lterm and some new symbols standing for 
-combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
-
-Those combinators are defined as:
-\begin{align*}
-\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
-\Kcomb &= \lam{x\,y}{x} \\
-\Icomb &= \lam{x}{x} 
-\end{align*}
-
-\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
-
-\Lets describe transformation $\Ae$ performing this process.
-\begin{align*}
-\Ae[x]           &= x &\\[0.4em]
-\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
-\Ae[\lam{x}{x}]  &= \Icomb &\\
-\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
-\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
-&\textbf{if } x \in \FV(M)\\
-\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
-&\textbf{if } x \in \FV(M) \vee x \in \FV(N)\\
-\end{align*}
-
-\begin{todo}
-& přidat (ale i do kódu) B a C
-& říct že my musíme řešit i typy 
-& říct o složitosti , a že je to docela
-disadventič vzhledem k bloatu
-& citovat Simon Peyton Jones implementaition of FP languages
-jako source of further reading on this subject.
-\end{todo}
-
-\subsection{Typed subtree swapping}
-\label{typed-swapping}
-
-\subsection{ Repairing method }
-\label{repairing-method}
-
-\begin{todo}
-& !!!
-\end{todo}
-
-\subsection{ Comparison of the two methods }
-\label{comarison-ski-repairing}
-
-\begin{todo}
-& !!!
-& říct že nakonec sme přilnuli k zbavit se proměnejch
-\end{todo}
-
-\newpage
-\section{Mutation}
-
-\begin{todo}
-& říct že na tu sme se moc nesoustředili ze dvou duvodu
-  ktery to ospravedlňujou 
-	&& koza taky
-	&& ta s vygereovanim podstromu je v zásadě použití
-	   algoritmu na generovaní jedince a tak nám přišlo
-	   lepší pořádně se soustředit na generování jedince a
-	   začít do toho mutaci tahat až bude generování pořádně
-	   prozkoumáno.
-& To ale neznamená že jí nemáme radi, naopak se těšíme
-  že její vhodný zavedení celou metodu ještě zlepší.
-
-& Nápad na to jak to dělat:
-  && nutno ošetřit - co všechno do gammy alemyslim že v phodě když je uplne 
-     stejná jako ta generujicí, pač ve vy SKI-ovanym termu stejne nejsou zadny 
-     proměný 
-  && mít nějakej předem danej počet pohybu na frontě, abych mohl efektivně
-     kontrolovat že to přidání mutací moc nespomalý
-  && Po tom počtu pohybu končim generování
-  && mam nějakej počet vygenerovanejch termu
-  && pro chytrou mutaci mužu chtít třeba 10 termu
-     &&& pokud se mi jich vygenerovalo míň, tak beru míň
-     &&& pokud se mi jich vygenerovalo víc, tak vemu 10 náhodně
-     &&& těch 10 vyhodnotim (jakože dosazenejch na jejich místo)
-     &&& nejlepší si vezmu
-     &&& dobrý udělat to zase najednou pro všechny v tý generaci mutovaný
-         pánčto je nákladný spouštět Hint
-& Další mutace - beta eta redukce (?? asi moc naroční viz dale) 
-  && zmenšující mutace
-  && udělaná že převedem SKIBC do lambda termu a to redukujem? to asi ne pač
-     pak převod to spíš prodlouží, spíš asi mít redukční pravidla pro SKIBC
-     a těma redukovat, ale je otázka jak moc tam budou vznikat SKIBC-redexi
-     - ty mohou vznikat jen při @-tree reprezentaci
-  && to otevírá otázku toho, že by elementy gammy mohla mít k sobě redukční pravidla
-& Další mutace - jednoduše náhrada podstromu/uzlu za nějakej z gammy
-  && nezvětšující mutace
-& Další mutace - mutace číselnejch-like konstant
-& To mě přivádí k myšlence sice pro generování ale myslim podstatný:
-  jakmile jsou v problému číselný-like konstanty tak mužu udělat pro 
-  jeden tvar hodně jedincu - což šetří jak čas na generování, tak to prozkomává 
-  víc danej tvar
-
-\end{todo}
-
-%\subsection{Using term generation}
-
-\section{ Complexity discussion }
-
-\begin{todo}
-& !!!
-& zmínit že mi moc neřešíme formální/asymptotický složitosti
-  && de nám hlavně o realný časy
-& vyjít z komplexnosti A* asi
-  && složitost výrazně ovlivněná kvalitou heuristický funkce
-\end{todo}
-
-\chapter{Polymorphic variant}
-
-\begin{todo}
- & poznamky o čem se mimojiné zmínit:
-   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
-      jako alternativa k ADF ! Myslim že hodně zajimavý
-\end{todo}
-
-\chapter{ Implementation }
-
-Předběžný náhodně uspořádaný body osnovy ..\\
-
-\begin{easylist}[itemize]
-& že je to v haskellu
-&& Jak moc hrotit vysvětlení haskellu?
-& že je to servrově řešený
-& architektura evolučního algoritmu 
-&& ty tři typový třídy a ta čtvrtá typová třída
-& evaluace
-&& vychytavka s vyhodnocenim celýho seznamu aby s eto nemuselo furt inicializovat - inicializace vyhodnocovače jednou za generaci 
-& hlavní soubory
-&& Eva.hs     - Eva monada
-&& GP-Core.hs - Evoluční algoritmus
-&& GP-Data.hs - Instance tříd z GP-Core
-&& IMx.hs - Term generating
-&& TTerm.hs   - typed lambda term
-&& TTree.hs   - CTT
-&& Server.hs  - main serveru
-&& Job.hs     - Registrace problemu
-&& Heval.hs   - Evaluator 
-&& složka Problems
-&&& pro každej problém složka 
-&&& ta typicky strukturu
-&&&& Problem.hs
-&&&& Funs.hs
-&& Slozka server
-&&& index.html
-&&& slozka js/Problems
-&&& .js pro kazdej problem
-\end{easylist}
-
-\section{Top level view}
-\section{Comments about main source files}
- \subsubsection{ Eva.hs }
- \subsubsection{ GP\_{}Core.hs }
-\section{Term generating}
-\section{Crossover}
-\section{Mutation}
-
-\section{Brief history of the project}
-\begin{todo}
- & zmínit třeba tyhle fáze:
-  && že sem si delal evaluator sám co sem pak s performenc duvodu opustil
-  && že sem hledal intuicionistickej dokazovač ale že nejsem nenarazil na vhodnej
-     pač nejsou delany na to plivat hodně dukazu  
-  && že sem nejdriv pouzival ten vypsychlej dokazovač co tam měl nějaký ty 
-     spešl rozhodovací stromy
-  && pak sem to přepsal tak aby to nebylo jen na GP šitý ale obecně na jakoukoliv 
-     evoluci
-  && ...
-\end{todo}
-
-\chapter{Problems}
-	In this section will be presented usage of the system in order to solve specific problems.
-
-		\section{Even Parity Problem}
-		\section{Big Context}
-		\section{Fly}
-		\section{Simple Symbolic Regression}
-		\section{Artificial Ant}
-		\section{Boolean Alternate}
-
-\chapter{Comparison with others}
-
-\chapter{Ideas yet not implemented}
-
-\section{Roadmap}
-\section{Tree ants}
-\section{[Propojení těch dvou barandrechtskejch odvozovacích pravidel do jedinýho]}
-\section{[Chytřejší heuristika pro A* která si to předpočítá na $\overline{\Gamma}$ ]}
-\section{[šlechtění těch strategií na generování termů]}
-\section{[šlechtění fitness funkcí a nápad s "turnajem olympioniků"]}
-\section{[šlechtění search strategií pro prohledávání]}
-
-%\chapter{Conclusion}
-\chapter*{Conclusion}
-
-\begin{todo}
- & !!!!!!!!!!!!!!!!!
-\end{todo}
-
-Další témata co zmínit v úvodu nebo v závěru :\\
-
-\begin{easylist}[itemize]
-
-& zmínit že existuje i jinej pohled na to proč tam zavádět typy:
-  && mužem koukat na to tak že konstrejnt je naše TuF tim, že 
-     v ní nemužou bejt libovolný funkce
-  && ale tuhle constrain mužeme omezit tim že konstrjnujeme 
-     právě použití tech bilding bloku, kde constrains jsou daný těma typama
-     - jakoby výměna global constrain za localní konstrejny na tech bilding  
-       blockách
-  && druhá možná ješte silnější motivace je že pomocí typu pak mohou vznikat 
-     smysluplnější programy, tím že dostavají vstupy které přesně čekají
-     a dávají výstupy které mají nějaké očekávané vlastnosti
-
-& Možná se zmínit o tý představě dvou pólu
-  induktivním a deduktivním -
-  klasickym GP a Dokazovačem, čim složitější
-  typovej sytem tim víc se jedná o deduktivní   
-  systém a blíží se to dokazovači.
-   && Ve světle toho, je pak hezky vidět, že
-      je chytrý exhaustivně systematicky prohledávat od nejmenšího
-      po největší- jak budeme zesložiťovat typovej systém, tak
-      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
-      jich generovat kvanta
-
-& že imlementace je delana těma vtipnejamtajpklasama umožnujicima libovolný
-  evoluční struktury na evolvovani
-
-& říct že vnekterejch aspektech je videt, že to je work in progress
-  ale že si myaslime že i tak to ukazuje pěkný výsledky 
-  a že doufame že ty nedodělky budou další prací odstraněny
-
-& Říct jak jde text:
-  && Začnem popsanim klasickýho GP podle Kozy
-  && Vyslovíme definice okolo lambda termu a typu
-  && Vyslovíme definici inhabitačního stromu
-     kterej je stromovou verzí IM od Barendrechta.
-  && Matematicky popíšem jak na to dem my.
-  && Popíšem rozšíření zakladního systému na 
-     polymorfní.
-  && Popíšem to z programátorskýho hlediska
-  && Popíšem příklady na kterejch to demonstrujem
-  && Srovnáme to s nějakejma přístupama co už 
-     existujou
-
-& Nějak naťuknout už tady v úvodu jak se to má při
-  srovnání s tim jak na to dou jiný (Yu atd)
-
-& Výkřiky co zmínit:
-  && Na co zaměřený:
-     &&& Proof of a concept
-     &&& Spíš jednoduchost
-         &&&& Ve většině aspektú to sleduje
-              Kozu - tzn i v málo řídících
-              parametru
-     &&& Duraz na generování termu (!)
-
-  && Na co nezaměřený (ale snad taky ok v tom):
-     &&& Performance   
-     &&& Mature knihovna co by šla použít robustně
-         v cizim kódu - spíš to počítá že to 
-         bude dál rozvíjeno a tohle je mezizastávka
-
- & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
-   defining EA and then specifying the differences or something 
-   like that...
-
-& bylo by asi mimojiný dobrý 
-ospravedlnit/okomentovat tak fancy (=javascript frontend) 
-implementaci při 
-dost nedodělanejch jinejch věcech (heuristika by 
-mohla být lepší, mutace by mohli bejt, atd)
-ale věřil sem, že je to proces iterativní
-a že je po celou dobu dobrý mít na čem to testovat.
-Javascript byl vybranej pač HTML je nejpromáklejší
-UI jaký vubec je na zemi teď (state of dzí art)
-a tak to snad nebyla stráta času.
-navíc se naskitla díky tomu další pěkná příležitost
-k ukázání toho, že to neni omezený na haskell
-ale že se kombinatorový konstrukty daj 
-v zásadě přeložit do libovolnýho jazyka celkem bez problému
-nebo minimalně do JS. Ale todle možná nepsat tolik do závěru
-spíš nějaky podrobnosti přehodit do implementačních detailu
-a tady to jen tak líznout.
-
-\end{easylist}
-
-\addcontentsline{toc}{chapter}{Conclusion}
-
-\begin{thebibliography}{9}
-
-% TODO bacha neni tam jednotnej formát!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-
-\bibitem{koza92}
-
-  John R. Koza,
-  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
-  MIT Press, Cambridge, MA,
-  1992.
-
-\bibitem{koza05}
-
-  Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
-  \emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
-  Springer, 2005. ISBN 978-0-387-26417-2
-
-\bibitem{yu01}
-
-  T. Yu. 
-  \emph{Hierachical processing for evolving recursive and modular 
-        programs using higher order functions and lambda abstractions}. 
-  Genetic Programming and Evolvable Machines,
-  2(4):345–380, December 2001. ISSN 1389-2576.
-
-\bibitem{montana95}
-
-D. J. Montana. 
-\emph{Strongly typed genetic programming.} 
-Evolutionary Computation, 3(2): 199–230, 1995.
-%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha
-
-\bibitem{haynes96}
-
-T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
-\emph{Type inheritance in strongly typed genetic programming.} 
-In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
-in Genetic Programming 2, chapter 18, pages 359–376.
-MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1.\\ 
-URL 
-\url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.
-
-\bibitem{olsson94}
-
-J. R. Olsson. 
-\emph{Inductive functional programming using incremental program 
-transformation and Execution of logic programs by 
-iterative-deepening A* SLD-tree search.} 
-Dr scient thesis, University of Oslo, Norway, 1994.
-
-\bibitem{barendregt92}
-H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum , H.  P.
-\emph{Lambda Calculi with Types.} 
-Handbook of Logic in Computer Science, 1992.
-
-\bibitem{barendregt10}
-
-  Henk Barendregt, Wil Dekkers, Richard Statman,
-  \emph{Lambda Calculus With Types}.
-  Cambridge University Press,
-  2010. \\
-  URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
-
-\bibitem{jones87}
-Simon Peyton Jones. 
-\emph{The Implementation of Functional Programming Languages}. 
-Prentice Hall, 1987.
-
-\bibitem{AIAMA}
-
-	Stuart J. Russell, Peter Norvig,
-	\emph{Artificial Intelligence: A Modern Approach}.
-	Pearson Education,
-	2003.
-
-\end{thebibliography}
-
-\end{document}

### 2013-07-21 tex/thesis.tex
[tex]

@@
      && potřebný definice pro to uvést v \lterms sekci
+     && musí se říc t co, tzn simpli typed l. kalkulus termy
+        a betaeta-redukce
+     && řekl bych že eta redukce je SN vždycky (tzn i untyped lk)
+        pač plati (CR+terminující pak SN) a je CR vzhledem k barendrecht-stará-žlutá
+        a terminující je rpotože zmenšuje term s nemuže zmenšovat donekonečna    
    & možná zmínit že se jedná o \textit{gentzen style proof}
@@
 terms in \bnf. As is shown in \cite{barendregt10}, 
-this method is obtained simply by replacement of 
+this method is obtained simply by replacing  
 the first grammar rule by slightly modified rule,
-resulting in following grammar.
+resulting in the following grammar.
 \begin{align*}
@@
 
-The difference lies in that $f$'s type is no longer obligated to be fully expanded
+The difference lies in that $f$'s type is no longer needed to be fully expanded
 ($\pi \in \T$ instead of $\alpha \in A$). This makes the grammar less deterministic,
-resulting in bigger search space. The new rule is generalization of the old one,
+resulting in a bigger search space. The new rule is generalization of the old one,
 thus all terms in \lnf will be generated, along with many new terms in \bnf that 
@@
 \lterms $M$,$N$ such that $M \not= N$ and $M =_{\eta} N$; 
-but by generating in \bnf we do not avoid this.
+but by generating in \bnf we do not avoid it.\\
+
+The disadvantage of the \lnf, as the name suggests, is that it is long.
+Terms in \lnf are said to be \textit{fully $\eta$-expanded}. 
+Relevant property of $\eta$-reduction is that it always shortens the term
+that is being reduced be it. And conversely, $\eta$-expansion prolongs.
+$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$
+
+Chceme ukazat že pomocí libovolné posloupnosti eta-redukcí se dostaneme do
+unikátní eta-nf. To plyne z toho, že
+
+ (a) každej lambda term má eta-NF (eta zmenšuje a nemuže se zmenšovat donekonečna)
+
+ (b) eta-redukce je CR (z žlutýho barendrechta) a z toho plyne že má unikátní NF (lemátko)
+
+pak chcem ukazat že eta redukcema nepřibude b-redex, když to dk
+tak víme že sme se tema eta redukce dostali do betaeta-NF.
+
+to se ukaže tak, že víme že eta redukce ničí @, tn ten novej redex by musel
+mít svůj stare @. Zajímá nas levej "argument" @ , aby vznikl b-redex
+musel by se zmenit něco co nebylo abstrakce na abstrakci ale to se nemohlo stát
+protože se mohla tak akorat posunout blíž ta abstrakce M k tomu @
+ale na zavinači určitě nevysel ten redukovanej podterm protože to by tam předtim měl ten redex.
 

### 2013-07-21 tex/thesis.tex
[tex]

@@
 }
+
+\newcommand{\lh}[1]{\lambda #1}
 
@@
 & !!!!!!! zmínit $=_\beta$ a $\twoheadrightarrow_\beta$  a i pro ostatní redukce!!!!!
+& definovat \textbf{CR} a říct o tom lemátku, že má CR redukce má unikátní nf.
 \end{todo}
@@
   to the \textit{implicational fragment of propositional 
-  intuitionistic logic}. We are mentioning this beautiful
+  intuitionistic logic}.
+
+  Intuition is that the proof of $A \then B$ 
+  can be imagined as function expecting proof of $A$
+  and utilizing it in production of the proof of $B$
+  that is returned as output of the function.
+
+  We are mentioning this beautiful
   correspondence in hope that it will help clarify 
-  use of \textit{inference rules} which are key concept
+  use of \textit{inference rules}, which are key concept
   in the process of deriving valid statements about
@@
 
+\newcommand{\etar}{\twoheadrightarrow_\eta}
+\newcommand{\ered}{$\eta$-reduction\xspace}
 \newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
+\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
+\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
 \newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
+\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
+\newcommand{\bredex}{$\beta$-redex\xspace}
 
@@
 The disadvantage of the \lnf, as the name suggests, is that it is long.
-Terms in \lnf are said to be \textit{fully $\eta$-expanded}. 
+Terms in \lnf are said to be \textit{fully $\eta$-expanded} \cite{barendregt10}. 
 Relevant property of $\eta$-reduction is that it always shortens the term
@@
 
-Chceme ukazat že pomocí libovolné posloupnosti eta-redukcí se dostaneme do
-unikátní eta-nf. To plyne z toho, že
-
- (a) každej lambda term má eta-NF (eta zmenšuje a nemuže se zmenšovat donekonečna)
-
- (b) eta-redukce je CR (z žlutýho barendrechta) a z toho plyne že má unikátní NF (lemátko)
-
-pak chcem ukazat že eta redukcema nepřibude b-redex, když to dk
-tak víme že sme se tema eta redukce dostali do betaeta-NF.
-
-to se ukaže tak, že víme že eta redukce ničí @, tn ten novej redex by musel
-mít svůj stare @. Zajímá nas levej "argument" @ , aby vznikl b-redex
-musel by se zmenit něco co nebylo abstrakce na abstrakci ale to se nemohlo stát
-protože se mohla tak akorat posunout blíž ta abstrakce M k tomu @
-ale na zavinači určitě nevysel ten redukovanej podterm protože to by tam předtim měl ten redex.
+Now we show that for every \lterm $M$ 
+every sequence of $\eta$-reductions is finite and
+leads to unique \enf $N$.
+
+\begin{enumerate}
+ \item Every application of \ered shortens the term.
+       Since every term has finite size, this process must 
+       end at some point. Thus every \lterm has \enf.
+ \item Since \ered is \textbf{CR}, \enf is unique (see \ref{todo}). 
+\end{enumerate}
+
+So we can take every generated \lterm $M$ in 
+\lnf and transform it to shorter term in \enf. 
+The question is whether it remains in \bnf, thus being in \benf.
+The answer is yes; it can be proven by showing that no 
+new \bredex is created by \ered.
+
+\begin{proposition}
+Let $P$ be in \bnf and $P \eArrow Q$. Then $Q$ is in \bnf.    
+\end{proposition}
+\begin{proof}
+
+For better clarity \lets show the \ered and \bredex using trees.\\
+\ered: \\
+
+\Tree [.$\lh{x}$ [.@ $M$ $x$ ] ] 
+~~~~$\eArrow$~~~~
+$M$
+
+And \bredex:
+
+\Tree [.@ [.$\lh{x}$ $M$ ] $N$ ] \\
+
+\Lets assume that $P \eArrow Q$ creates a new \bredex $B$ in $Q$.
+
+Since \ered only destroys and never creates \textit{function applications} (i.e. @),
+the root @ of $B$ must be present in $P$.  
+But since $P$ contains no \bredex, the left subterm $L$ of this root @
+is not $\lambda$-abstraction.
+Only possible way for $L$ to be changed by $\eArrow$ into 
+a $\lambda$-abstraction is that $L$ is the reduced subterm (so that
+$L$ is changed for its subterm).
+But that is in contradiction with $P$ not containing any \bredex,
+because it would cause $L$ be a $\lambda$-abstraction.
+
+\end{proof}
+
+Notable property of \lnf and \benf is that there is \textit{bijection} 
+(i.e. one-to-one correspondence) of 
+the set of simply typed \lterms in \lnf and 
+the set of simply typed \lterms in \benf.
+
+\begin{proposition}
+
+Reduction to \enf is bijection between  
+the set of simply typed \lterms in \lnf and 
+the set of simply typed \lterms in \benf.
+\end{proposition}
+\begin{proof}
+
+Since reduction to \enf always leads to an unique term, it is a function.
+In previous proposition is shown that \ered of \lnf
+leads to a term in \benf.\\
+
+In order to show that a function is bijection it is sufficient to show that it is
+both \textit{injection} and \textit{surjection}.\\
+
+Suppose it is not injection.
+
+So there must be $M_1,M_2$ in \lnf such that $M_1 \not= M_2$
+and $N$ in \benf such that $M_1 \etar N$, $M_2 \etar N$.
+Therefore $M_1 =_\eta M_2$, 
+so $M_1 =_{\beta\eta^{-1}} M_2$.
+This contradicts with $M_1,M_2$ being distinct \lnf{}s.\\
+
+Every $M$ in \bnf has a \lnf $N$ such that 
+$N \twoheadrightarrow_{\eta} M$ (proposition from \ref{lnf}).
+Term $M$ in \benf is in \bnf, thus it has desired \lnf $N$
+which reduces to it.
+
+Therefore it is surjection. 
+\end{proof}
+
+Suppose we have systematic (i.e. gradually generating all terms, 
+but no term twice) method for generating terms in \lnf,
+we may transform it to systematic method for generating terms in \benf
+by simply reducing each generated term to its \enf.
+
+\begin{todo}
+& možná eště zmínit něco jako : \\
+Jedna drobná nevýhoda je, že pokud chceme generovat termy od nejmenšího do největšího
+a pokud lnf generujeme od nejmenšího po největší tak pak generovaní \benf
+timhle trikem nemusí bejt od nejmenšího po nějvětší. V naší aplikaci (tzn GP)
+ale nijak nebazírtujeme na týhle vlastnosti a nakonec budeme spíš stejně používat
+nějaký nedeterminizmy aby to fungovalo líp, takže to vubec nevadí, že budou trochu přeházený. 
+\end{todo}
 
@@
 
-Now we will introduce \textit{Inhabitation tree}, structure slightly different from
-\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Barendregt.
-We can think about Inhabitation tree as about unfolded Inhabitation machine.
-The motivation for using Inhabitation trees is belief that it will help us 
-reason about generation of \lterms 
-of a given type $\sigma$ and with a given context $\Gamma$.
-
-\subsection{Definition of Inhabitation tree}
-
-\textit{Inhabitation tree} is a \textit{rooted tree}, possibly infinite. 
-It has two types of nodes:
+Now we introduce \textit{inhabitation tree}; structure slightly different from
+\textit{inhabitation machine}, which was introduced by Barendregt\cite{barendregt10}.
+We can think about inhabitation tree as about unfolded inhabitation machine.
+Inhabitation tree approach is further refinement 
+of the term generating grammar approach, which is too much textual.
+Whereas inhabitation tree approach aims at tree structure of \lterms.
+
+~\\
+
+\newcommand{\IT}{\mathop{\mathrm{IT}}}
+
+\begin{definition}
+
+Let context $\Gamma$ and type $\sigma$ be given.
+\textit{Inhabitation tree} for $(\sigma,\Gamma)$, denoted as $\IT(\sigma,\Gamma)$,  
+is a rooted possibly infinite tree that is recursively defined as follows.\\
+
+If $\sigma = \alpha$ such that $\alpha \in A$, then
+
+$\IT(\sigma,\Gamma) = $ \\
+
+\Tree 
+[.$\alpha$ 
+   [.$f_1$ $\IT(\rho^1_1,\Gamma)$ $\dots$ $\IT(\rho^1_{m_1},\Gamma)$ ]
+   [.$\dots$ ]
+   [.$f_k$ $\IT(\rho^k_1,\Gamma)$ $\dots$ $\IT(\rho^k_{m_k},\Gamma)$ ]           
+]\\[1em]
+
+~~~~~~where
+
+$~~~~~~~~~~\{ \,
+f_1 : \rho^1_1 \ar \dots \ar \rho^1_{m_1} \ar \alpha ~,~\dots~,~
+f_k : \rho^k_1 \ar \dots \ar \rho^k_{m_k} \ar \alpha \} \\ ~\,~~~~~~~~~~= 
+\{\, f \: : \rho_1 \ar \dots \ar \rho_m \; \ar \alpha ~|~ 
+  (f,\rho_1 \ar \dots \ar \rho_m \ar \alpha ) \in \Gamma \,\}$.\\
+
+Otherwise, for $\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha$ 
+such that $n> 0, \alpha \in A$
+
+$\IT(\sigma,\Gamma) =$\\
+
+\Tree [.$\tau_1\ar\dots\ar\tau_n\ar\alpha$ 
+ [.$\lh{x_1~\dots~x_n}$ [.$\IT(\alpha;\Gamma,x_1:\tau_1,\dots,x_n:\tau_n)$ ] ] ]
+
+\end{definition}~\\[1em]
+
+\newpage
+
+\Lets discuss this definition in grater detail. 
+An inhabitation tree has two kinds of nodes.
 
 \begin{samepage}
-\begin{itemize}
-  \item Type nodes   
-  			- containing type $\sigma \in \mathbb{T}$ 
-  			- aka "OR-node" %, Nonterminal-node. 
-  \item Symbol nodes 
-  			%- containing $ x^{*} \in \mathfrak P \left(V\right) -blobost pač ordered $ or $c \in C$
-  			- containing "$\lambda$-head" (nonempty finite sequence of variable names) or constant
-  			  name. 
-  			- aka "AND-node" %, Terminal-node. - ale nemusí, proto jsem to zakomentoval
-\end{itemize}
-\end{samepage}
-
-We construct Inhabitation tree for given type $\sigma$ and context $\Gamma$.\\
-We will define Inhabitation tree by describing its construction for a given $(\sigma,\Gamma)$.
-Notice that it will closely follow the rules from \ref{barlike}:
-
-\begin{itemize}
+\begin{enumerate}
+  \item \textit{Type node} (or \textit{OR-node}).
+
+  	    This kind of node contains type. %$\sigma \in \mathbb{T}$  
+  \item \textit{Symbol node} (or \textit{AND-node}).
+
+  		This kind of node contains 
+  		constant name, variable name 
+  		\\or \textit{$\lambda$-head} 
+  	    --- non-empty finite sequence of variable names. 
+\end{enumerate}
+\end{samepage}~
+
+Following observation clarifies the alternation of node kinds.
+\begin{enumerate}
 \item The root of Inhabitation tree for $(\sigma,\Gamma)$ is 
@@
 \item All \textit{type nodes} have as child nodes only \textit{symbol nodes}. 
-\item And all \textit{symbol nodes} have as child nodes only \textit{type nodes}. 
-\end{itemize}
-
-\begin{samepage}
-Now we will resolve the child nodes of the root node.
-There are two cases of $\sigma$ (recall \ref{deftype}): 
-\begin{description}
-	\item[Atomic type] 
-		$\sigma = \alpha $ where $\alpha \in A$. 
-	\item[Function type] 
-		$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
-		where $n \geq 1, \alpha \in A$.
-\end{description}
-\end{samepage}
-
-First case \textbf{Atomic type} --- i.e., $\sigma = \alpha$ where $\alpha \in A$:\\
+\item All \textit{symbol nodes} have as child nodes only \textit{type nodes}. 
+\end{enumerate}~
+
+%~\\~\\~\\
+
+There are two cases of $\sigma$ --- 
+\textit{atomic type} and \textit{function type}. \\
+
+First case is \textbf{atomic type}, i.e., $\sigma = \alpha$ where $\alpha \in A$.
+
 For every $(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha) \in \Gamma$
-where $\alpha \in A$ there is a child \textit{symbol node} of the root containing constant name $f$.
-This symbol node containing $f$ has $m$ child subtrees corresponding to Inhabitation trees for 
-$(\rho_1,\Gamma),\dots,(\rho_m,\Gamma)$.
+there is a child \textit{symbol node} containing constant or variable name $f$.
+This symbol node containing $f$ has $m$ child subtrees corresponding to  
+$\IT(\rho_1,\Gamma),\dots,\IT(\rho_m,\Gamma)$.
+
+Therefore specially for $m = 0$ node containing $f$ is 
+a leaf of the inhabitation tree.
 
@@
 
-Compare this case with corresponding grammar rule:
+Compare this case with corresponding grammar rule.
 \[ 
@@
 
-Second case \textbf{Function type} --- i.e., 
+Second case is \textbf{function type}, i.e., 
 $\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
-where $n~\geq~1, \alpha~\in~A$:\\
-For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.
-
-There is one and only one child \textit{symbol node} of the root containing "$\lambda$-head" 
-$\lambda x_1 \dots x_n$ which stands for sequence of variable names $(x_1,\dots,x_n)$.
+where $n~\geq~1, \alpha~\in~A$.\\
+For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ 
+that is not yet included in context $\Gamma$ as variable or constant name.
+
+There is one and only one child \textit{symbol node} 
+of function type node containing $\lambda$-head
+$\lambda x_1 \dots x_n$, which stands for 
+sequence of variable names $(x_1,\dots,x_n)$.
 This symbol node containing $\lambda x_1 \dots x_n$
-has one and only one child subtree corresponding to Inhabitation trees for 
-$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$.
+has one and only one child subtree corresponding to  
+$\IT(\alpha;\Gamma,x_1:\tau_1 , \dots , x_n:\tau_n )$.
 
@@
 	$\lambda x_1 \dots x_n .$ 
-	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
+	}}( \alpha ; \Gamma , x_1:\tau_1,\dots,x_n:\tau_n ) \quad \bigg)
 \]
@@
 
+\newpage
 \subsection{And-or tree and searching in Inhabitation tree}
+
 \Lets consider following definition of 
-\textit{And-or tree}\footnote{
-\textbf{TODO}: Mention that on WIKI there is more general definition, but 
-for our purposes is this one sufficient.}:
-
-\textit{And-or tree} is a rooted tree where each node is labeled as either \textit{and-node} 
-or\footnote{xor} \textit{or-node}.
-
-By \textit{solving} And-or tree $T$ we mean finding finite tree $T'$
-subgraph of $T$ such that it follows these conditions: 
-\begin{itemize}
+\textit{and-or tree}\footnote{
+\textbf{TODO}: Mention that on [WIKI] there is more general definition, but 
+for our purposes is this one sufficient.}.
+
+\begin{definition}
+~
+	\begin{enumerate}
+	\item \textit{And-or tree} is a rooted possibly infinite 
+	      tree with every node labeled as either \textit{and-node} 
+	      or
+	      %\footnote{xor} 
+	      \textit{or-node}.
+
+	\item By \textit{solving and-or tree} $T$ is meant finding finite tree $T'$
+	      subgraph of $T$ such that following conditions are met. 
+	    \begin{enumerate}
 		  \item The root of $T'$ is the root of $T$.
@@
 		      and not at least one ..or CHANGE the def. }   
-\end{itemize}
-~
-\Lets now consider following labeling of Inhabitation tree:
+	    \end{enumerate}
+
+	\end{enumerate}
+
+\end{definition}
+
+~\\[1em]
+\Lets now consider following labeling of Inhabitation tree.
 
@@
   \item \textbf{Symbol nodes} are labeled as \textbf{and-nodes}.
-\end{itemize}
-
-This labeling has following justification:
+\end{itemize}~
+
+This labeling has following justification.\\
 
@@
 child node in \textit{type node} corresponds to selection of exactly one
-grammar rule in order to rewrite nonterminal symbol.
+grammar rule in order to rewrite nonterminal symbol.  \\
 
@@
 rewriting all the nonterminal symbols in string that is being generated.
+Or put in different words, it corresponds to filling all the slots
+for arguments with values.\\
 
@@
 of \lterms.
+
+\begin{todo}
+ & Přepsat poslední odstavec --- počítal ještě s tím že 
+   pasaže o stromové reprezentaci jsou až po této kapitole, což bylo změněno.
+\end{todo}
 
@@
 \begin{todo}
-  & DESCRIBE Inhabitation Machine...
-   && něco jako na závěrještě ve stručnosti popišme 
+  & Když na to zbude čas, tak popsat Inhabitation Machine...
+   && něco jako na závěr ještě ve stručnosti popišme 
       IM, strukturu ze který je odvozenej IT. 
@@
 parent trees with selected nodes in bold.\\
-
-\newcommand{\lh}[1]{\lambda #1}
 

### 2013-07-21 tex/thesis.tex
[tex]

@@
 \section{Program trees}
+\label{GP-prog-trees}
 
@@
 may occur variables $x_{1}, ..., x_{n}$.
+
+\begin{todo}
+ & nějak to tu přepsat aby to bylo ve stejnym stylu jako v \ref{tree-reps}
+\end{todo}
 
@@
 
+\newcommand{\then}{\Rightarrow\xspace}
+
+\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
+\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
+
+\newcommand{\ST}{\mathop{\mathrm{ST}}}
+\newcommand{\FV}{\mathop{\mathrm{FV}}}
+
+\newcommand{\Scomb }{\mathbf{S}}
+\newcommand{\Kcomb }{\mathbf{K}}
+\newcommand{\Icomb }{\mathbf{I}}
+
+\newcommand{\etar}{\twoheadrightarrow_\eta}
+\newcommand{\ered}{$\eta$-reduction\xspace}
+\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
+\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
+\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
+\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
+\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
+\newcommand{\bredex}{$\beta$-redex\xspace}
+
+\newcommand{\lnf}{\textit{lnf}\xspace}
+
 In this chapter we briefly describe key concepts
-from simply typed $\lambda$-calculus.
+of simply typed $\lambda$-calculus.
 We try to favor readability over formality,
 for more detailed and formal treatment of this subject see 
-\cite{barendregt92}.
-
-After that, \textit{inhabitation tree} is introduced.
-It is structure used for reasoning about generation of individuals.
+\cite{barendregt92,barendregt10}.
+
+After introductory description, an analysis of term generation follows; 
+a path from inference rules to term generating grammar rules is shown,
+\textit{long normal form (lnf)} is described,
+followed by term generating grammar for \textit{lnf}
+and analysis of its relevant properties (such as its relation to \benf),
+then \textit{inhabitation tree} is presented --- 
+an implementation friendly refinement of 
+\textit{lnf} term generating grammar.
+
+\begin{todo}
+ & Až to bude dostatečně sepsáno, tak do této kapitoly taky přidat
+   část o abstraction elimination a o polymorfním zobecnění inhabitation tree.
+   Nezapomenout to pak zmínit i v tomhle úvodu.
+\end{todo}
 
@@
 
-Now we will formally describe programming language 
-in which we will generate individual programs --- so called \lterms.
-
-\newcommand{\then}{\Rightarrow\xspace}
-
-\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
-\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
-
-\newcommand{\ST}{\mathop{\mathrm{ST}}}
-\newcommand{\FV}{\mathop{\mathrm{FV}}}
-
-\newcommand{\Scomb }{\mathbf{S}}
-\newcommand{\Kcomb }{\mathbf{K}}
-\newcommand{\Icomb }{\mathbf{I}}
+Here we  %formally 
+describe programming language, 
+in which we generate individual programs --- so called \lterms.
 
@@
 x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
-\textit{~~~~~($\lambda$-abstraction)} 
+\textit{~~~~($\lambda$-abstraction)} 
 \end{align*}~
@@
 
-Examples of \lterms:
-
-\begin{easylist}[itemize]
-& $x$
-& $\lamb{x}{x}$
-& $\lamb{x}{f}$
-& $(\lamb{x}{(x~x)}~\lamb{x}{(x~x)})$
-& $\lamb{x}{\lamb{y}{x}}$
-& $\lamb{f}{\lamb{g}{\lamb{x}{((f~x)~(g~x))}}}$
-\end{easylist}~
-
 We use following notation
-abbreviations for better readability:
-
-\begin{easylist}[itemize]
+abbreviations for better readability. \vspace{3mm}
+
+\begin{easylist}[enumerate]
 & $M_1~M_2~M_3~\dots~M_n$ for 
@@
 as $\lam{f\,g\,x}{f~x\,(g~x)}$.
+
+\newcommand{\ods}{\hspace{3mm}}
+
+~
+
+Here follow some examples of \lterms. \vspace{1mm}
+
+\ods $x$
+
+\ods $\lamb{x}{x}$
+
+\ods $\lamb{x}{f}$
+
+\ods $(\lamb{x}{(x~x)}~\lamb{x}{(x~x)})$
+
+\ods $\lamb{x}{\lamb{y}{x}}$
+
+\ods $\lamb{f}{\lamb{g}{\lamb{x}{((f~x)~(g~x))}}}$
 
@@
 \begin{todo}
+& asi to přepsat nějak obecně pro redukci R a pak zadat beta,eta,eta-1 
+  jako specialni připady
 & !!!!!!! zmínit $=_\beta$ a $\twoheadrightarrow_\beta$  a i pro ostatní redukce!!!!!
 & definovat \textbf{CR} a říct o tom lemátku, že má CR redukce má unikátní nf.
+&& V \cite{barendregt84} (ta nová kniha s žlutejma deskama) je ukázano že je 
+	 eta CR (Church Rosserovská)
+& zmínit Theorem 2B.4 z \cite{barendregt10}
+& co to znamená mít nf
 \end{todo}
@@
 
+\subsection{Tree representations of \lterms}
+\label{tree-reps}
+
+\newcommand{\sexprTree}{sexpr-tree\xspace}
+\newcommand{\atTree}{@-tree\xspace}
+
+From the definition of \lterm we can straightforwardly derive 
+the standard tree representation for \lterms, which we call 
+\textit{\atTree}. Term M is 
+translated into \atTree $T_@[M]$ by following rules.
+
+\begin{enumerate}
+	\item $T_@[x] := x$ (i.e. $x \in V \cup C$ becomes one node tree $x$)
+	\item \mbox{$T_@[(P$ $Q)] := $ \Tree
+			[.@	
+		 		\text{$T_@[P]$}
+		 		\text{$T_@[Q]$}		 			
+			] }
+	\item \mbox{$T_@[\lambda x . P] := $ \Tree
+			[.\text{$\lambda x$}	
+		 	 	\text{$T_@[P]$}	
+			] }
+\end{enumerate}
+
+We can enhance this representation by replacing third rule
+by following one. It compressing consecutive 
+$\lambda$-abstractions into one \atTree node. \\
+
+\mbox{$T_@[\lambda x_1 \dots x_n . P] := $ \Tree
+	[.\text{$\lambda x_1 \dots x_n$ }	
+ 	 	\text{$T_@[P]$}	
+	] }
+
+	~\\
+
+Since this representation comes directly from definition, 
+it is evident that it covers all possible \lterms.\\
+
+Following tree representation is extension of 
+\textit{S-expression trees} described in~\ref{GP-prog-trees},
+thus we call it \textit{\sexprTree}.
+Unlike \atTree, it is not able to cover all \lterms, but only those
+in \bnf. 
+Term M is translated into \sexprTree $T_S[M]$ by following rules.
+
+%It will also be the representation suitable for showing 
+%that \textit{solving} Inhabitation tree generates wanted \lterm.
+
+\begin{enumerate}
+    \item $T_S[x] := x$
+	\item \mbox{ $T_S[(f M_1 M_2 \dots M_m)] := $ 
+	%where $f \in V \cup C, n \geq 1$ translates into tree\\
+		\Tree
+			[.f	
+		 		\text{$T_S[M_1]$}
+		 		\text{$T_S[M_2]$}
+		 		\text{$\dots$}
+		 		\text{$T_S[M_m]$}		 				 			
+			] }
+	\item \mbox{$T_S[\lambda x_1 \dots x_n . M] := $
+	 	\Tree
+			[.\text{$\lambda x_1 \dots x_n$ }	
+		 	 	\text{$T_S[M]$}	
+			] }
+\end{enumerate}~
+
+\begin{todo}
+
+& ujistit se že je v části o tree reprezentacích 
+  je napsaný co to znamená že podstrom má nějakej typ,
+  že podstromy odpovídaj podvýrazum atd
+
+ & říct že to je přesně ta notace co používá koza
+ & EXAMPLES of tree representations of \lterms  
+ & napsat transformace mezi \atTree a \sexprTree
+ & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
+   přímo \sexprTree a říct proč se tomu tak řiká 
+   (na tom místě co je v textu dřív)
+
+ & zkontrolovat že reference tree-reps se odkazuje na dobrý místo,
+   mohlo by se to rozbít tim že tree reprezentations - typovany/netypovany 
+   jsou od sebe oddelený
+\end{todo}
+
+\newpage
 \section{Types}
@@
    & říct že jsou to modus ponens a jeden směr věty o dedukci  
+\end{todo}
+
+\subsection{Tree representations of typed \lterms}
+
+\newcommand{\lsCurry}{$\lambda^{\ar}$\nobreakdash-Curry\xspace}
+\newcommand{\lsChurch}{$\lambda^{\ar}$\nobreakdash-Church\xspace}
+
+Simply typed $\lambda$-calculus as described above is formally
+called the \textit{system \lsCurry}. 
+Other slightly different formalization is called 
+\textit{system \lsChurch}. 
+They differ in that \lsCurry uses as terms above described
+untyped terms, whereas \lsChurch uses slightly modified
+terms with information about types of variables written in
+lambda abstraction along with variable name.
+Since those systems are proven to be equivalent,
+we are not giving this subject much attention here.
+For more detailed information on this subject see \cite{barendregt92}.\\
+
+In our implementation we also use modified representation of \lterms
+storing information about types. But our representation is more 
+radical than that of \lsChurch; it stores information about types
+of all subterms.
+
+Regarding tree representation of such a typed
+\lterm, most of the time it is more convenient to 
+understand this information as invisibly sitting
+in the tree nodes alongside with visible
+term symbols.
+
+But as we will see below in the section about \textit{inhabitation trees},
+there are situations when the explicit type information in the tree representation
+comes in handy. Thus we are presenting following tree representation.\\
+
+Straightforward approach would be to add to each node a visible type entry 
+that would be the type of the subterm corresponding to subtree having this
+node as the root node.
+
+Approach more suitable for our further purpose 
+is to add a special type node \textit{above} each node.
+
+More specifically,
+\lets consider \textit{\sexprTree} $t$ corresponding to a \lterm of the type
+$\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
+Then corresponding \textit{typed \sexprTree} $TT_S[t]$ for typed \lterm is 
+obtained from the tree $t$ as follows.
+
+\begin{equation*}
+\mbox{ 
+TT[
+\Tree
+	[.r 	
+	  	  \text{$s_1$}
+		  \text{$s_2$}
+		  \text{$\dots$}
+		  \text{$s_n$}
+	] 
+}]=
+\mbox{
+\Tree
+	[.\text{$\sigma$ }
+	    [.r 	
+	  	  \text{$TT[s_1]$}
+		  \text{$TT[s_2]$}
+		  \text{$\dots$}
+		  \text{$TT[s_n]$}
+		]	  	
+	] 
+}
+\end{equation*}
+
+We silently suppose here that type information about every subterm is available.
+
+\begin{todo}
+ & možná konec týhle subsekce trochu přeformulovat, je to takový kostrbatý...
 \end{todo}
@@
      jak je to v Barendregtovi \cite{barendregt10} 
-   & vyřešit jak je to stou freš proměnnou x v introdukci...
+   & vyřešit jak to tu popsat s tou freš proměnnou x v introdukci...
      && stejně tak v tom obecnym případě to dělá neplechu
@@
 
-\newcommand{\lnf}{\textit{lnf}\xspace}
+%v barendrechtovi je to v definici lnf
+%$\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$ místo $M_m$ 
+%ale to musí bejt určitě překlep...
 
@@
 Let \GMS where 
-$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha$,
-$n \geq 0$.
-Then $M$ is in \textit{long normal form} (\textit{lnf}) if following 
+$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha, n \geq 0$.
+	\begin{enumerate}
+	  \item	
+		Then $M$ is in \textit{long normal form} (\lnf) if following 
 		conditions are satisfied.
@@
 		 \item $M$ is term of the form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_m}$\\
-  (specially for $n = 0$, $M$ is term of the form $f$),
- \item each $M_i$ is in \textit{long normal form}.
+		  (specially for $n = 0$, $M$ is term of the form $f$).
+		 \item Each $M_i$ is in \lnf.
+		\end{enumerate}	
+	  \item 
+	    $M$ has a \lnf if $M =_{\beta\eta} N$ and $N$ is in \lnf.
 	\end{enumerate}
@@
 
-As is shown in \cite{barendregt10} long normal form has those 
-nice properties:
-
-\begin{easylist}[itemize]
-& Every $M$ in $\beta$-normal form has a long normal form 
-$N$ such that $N \twoheadrightarrow_{\eta} M$.
-& \textbf{TODO jen zkopčeno přepsat!!} .. if M has a beta-nf, which according to Theorem 2B.4 is
-always the case, then it also has a unique lnf and this will be its unique $\beta\eta^{-1}$-nf.
-& \textbf{TODO : potřebuju eštěnějaký z tyhle knihy?}
-\end{easylist}
+As is shown in \cite{barendregt10}, \lnf has following nice properties.
+
+\begin{proposition}
+If $M$ has a \bnf, 
+%which according to Theorem 2B.4 is always the case, 
+then it also has a unique \lnf, 
+which is also its unique \beenf.
+\end{proposition}
+
+\begin{proposition}
+Every $B$ in \bnf has a \lnf 
+$L$ such that $L \twoheadrightarrow_{\eta} B$.
+\end{proposition}
 
 \begin{todo}
- & v barendrechtovi je to v definici lnf $\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$ místo 
-   $M_m$ ale to musí bejt určitě překlep...
- & popsat tu jak ji předelat na beta nf
-	pokud by se nám chtělo... 
-	V barendrech84 (ta nová kniha s žlutejma deskama) je ukázano že je 
-	eta CR (čurč roserovská) což snad stačí společně s 
-	tim tvrzeničkem z book.pdf že vezmu lnf
-	a dělám eta redukce dokud to de. až to nejde tak bych to měl mít v 
-	beta-nf
- & přepsat to nějak hezčejc, ten seznam heskejch vlastností
+ & co to znamená mít lnf
+ & potřebuju eště nějaký \textbf{propositions} z tyhle knihy?
+ & ta definice je tam daná pro $\tur{}{M}{\sigma}$,
+   ukazat ze ta naše je ok, že tam má navíc to $\Gamma$.
+   Stačí udělat tríček že beru funkci co má lambda hlavu odpovídající 
+   tomu kontextu.
 \end{todo}
@@
 almost identical notation. 
+But his \lnf grammar was more clever then my grammar
+that was using grammar rules corresponding to the three basic
+inference rules. 
 }
@@
 
+\begin{todo}
+ & zmínit že podobně jako mužeme transformovat inference rules
+   na grammar rules, mužeme i obráceně grammar rules
+   převést na inference rules - to bylo ostatně uděláno v tom dukazu ale
+   je to podle mě hezký říct ty pravidla explicitně.
+\end{todo}
+
 \subsection{Benefits of generating \lterms in \lnf}
@@
 as the computation performed by \lterm $L$. \\
-
-\newcommand{\etar}{\twoheadrightarrow_\eta}
-\newcommand{\ered}{$\eta$-reduction\xspace}
-\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
-\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
-\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
-\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
-\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
-\newcommand{\bredex}{$\beta$-redex\xspace}
 
@@
 
-$\IT(\sigma,\Gamma) = $ \\
+$\IT(\sigma,\Gamma) := $ \\
 
@@
 
-$\IT(\sigma,\Gamma) =$\\
+$\IT(\sigma,\Gamma) := $\\
 
@@
    pasaže o stromové reprezentaci jsou až po této kapitole, což bylo změněno.
-\end{todo}
-
-\section{Tree representations of \lterms}
-\label{tree-reps}
-
-\newcommand{\sexprTree}{sexpr-tree\xspace}
-\newcommand{\atTree}{@-tree\xspace}
-
-From the definition of \lterm (\ref{deflam}) we can straightforwardly derive 
-the classical tree representation for \lterms. Term M is translated into tree $T[M]$ by following rules:
-
-\begin{itemize}
-	\item $x \in V \cup C$ translates into \textit{leaf} $x$.
-	\item $(P$ $Q)$ translates into tree\\
-		\Tree
-			[.@	
-		 		\text{$T[P]$}
-		 		\text{$T[Q]$}		 			
-			] 
-	\item $\lambda x . P$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x$}	
-		 	 	\text{$T[P]$}	
-			] 
-\end{itemize}
-
-We can enhance this representation by compressing consecutive lambda abstractions into one
-tree node like this:
-
-\begin{itemize}
-	\item $\lambda x_1 \dots x_n . P$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x_1 \dots x_n$ }	
-		 	 	\text{$T[P]$}	
-			] 
-\end{itemize}
-
-As this representation comes directly from definition it is evident 
-that it covers all possible \lterms.
-
-We will refer to this representation as to \textit{\atTree}.\\
-
-For representing expressions as trees it is however more common to use a little different
-representation. It will also be the representation suitable for showing 
-that \textit{solving} Inhabitation tree generates wanted \lterm.
-
-\begin{itemize}
-    \item $x \in V \cup C$ translates into \textit{leaf} $x$.
-	\item $(f M_1 M_2 \dots M_n)$ where $f \in V \cup C, n \geq 1$ translates into tree\\
-		\Tree
-			[.f	
-		 		\text{$T[M_1]$}
-		 		\text{$T[M_2]$}
-		 		\text{$\dots$}
-		 		\text{$T[M_n]$}		 				 			
-			] 
-	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
-	 	\Tree
-			[.\text{$\lambda x_1 \dots x_n$ }	
-		 	 	\text{$T[M]$}	
-			] 
-\end{itemize}
-
-Notice that this representation does not cover all \lterms, 
-e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.
-
-We will refer to this representation as to \textit{\sexprTree}.\\
-
-\Lets now consider representation for \textit{typed \lterms}.
-Straightforward approach would be to add to each node a type entry which 
-would be the type of the \lterm corresponding to subtree having this
-node as the root node.
-
-Approach more suitable for our purpose is to add a special type node above each node.
-More specifically:
-
-\Lets consider tree $t$ corresponding to a \lterm of a type
-$\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
-Then corresponding tree $TT[t]$ for typed \lterm is 
-obtained from the tree $t$ as follows:
-
-\begin{equation*}
-\mbox{ 
-TT[
-\Tree
-	[.r 	
-	  	  \text{$s_1$}
-		  \text{$s_2$}
-		  \text{$\dots$}
-		  \text{$s_n$}
-	] 
-}]=
-\mbox{
-\Tree
-	[.\text{$\sigma$ }
-	    [.r 	
-	  	  \text{$TT[s_1]$}
-		  \text{$TT[s_2]$}
-		  \text{$\dots$}
-		  \text{$TT[s_n]$}
-		]	  	
-	] 
-}
-\end{equation*}
-
-\begin{todo}
-
-& ujistit se že je v části o tree reprezentacích 
-  je napsaný co to znamená že podstrom má nějakej typ,
-  že podstromy odpovídaj podvýrazum atd
-
- & říct že to je přesně ta notace co používá koza
- & EXAMPLES of tree representations of \lterms  
- & napsat transformace mezi \atTree a \sexprTree
- & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
-   přímo \sexprTree a říct proč se tomu tak řiká 
-   (na tom místě co je v textu dřív)
-
 \end{todo}
@@
 \bibitem{koza92}
-
   John R. Koza,
@@
 \bibitem{koza05}
-
   Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
@@
 \bibitem{yu01}
-
   T. Yu. 
@@
 \bibitem{montana95}
-
 D. J. Montana. 
@@
 \bibitem{haynes96}
-
 T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
@@
 \bibitem{olsson94}
-
 J. R. Olsson. 
@@
 
+\bibitem{barendregt84}
+H. P. Barendregt,
+\emph{The Lambda Calculus: its Syntax and Semantics}, 
+revised ed., North-Holland, 1984.
+
 \bibitem{barendregt92}
-H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum , H.  P.
+H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
 \emph{Lambda Calculi with Types.} 
@@
 \bibitem{AIAMA}
-
 	Stuart J. Russell, Peter Norvig,
@@
 %\end{todo}
+
+%\section{---OLD---Tree representations of \lterms}
+
+%From the definition of \lterm (\ref{deflam}) we can straightforwardly derive 
+%the classical tree representation for \lterms. Term M is translated into tree $T[M]$ by following rules:
+%
+%\begin{itemize}
+%	\item $x \in V \cup C$ translates into \textit{leaf} $x$.
+%	\item $(P$ $Q)$ translates into tree\\
+%		\Tree
+%			[.@	
+%		 		\text{$T[P]$}
+%		 		\text{$T[Q]$}		 			
+%			] 
+%	\item $\lambda x . P$ translates into tree\\
+%	 	\Tree
+%			[.\text{$\lambda x$}	
+%		 	 	\text{$T[P]$}	
+%			] 
+%\end{itemize}
+%
+%We can enhance this representation by compressing consecutive lambda abstractions into one
+%tree node like this: 
+%
+%\begin{itemize}
+%	\item $\lambda x_1 \dots x_n . P$ translates into tree\\
+%	 	\Tree
+%			[.\text{$\lambda x_1 \dots x_n$ }	
+%		 	 	\text{$T[P]$}	
+%			] 
+%\end{itemize}
+%
+%As this representation comes directly from definition it is evident 
+%that it covers all possible \lterms.
+%
+%We will refer to this representation as to \textit{\atTree}.\\
+% 
+%
+%For representing expressions as trees it is however more common to use a little different
+%representation. It will also be the representation suitable for showing 
+%that \textit{solving} Inhabitation tree generates wanted \lterm.
+%
+%\begin{itemize}
+%    \item $x \in V \cup C$ translates into \textit{leaf} $x$.
+%	\item $(f M_1 M_2 \dots M_n)$ where $f \in V \cup C, n \geq 1$ translates into tree\\
+%		\Tree
+%			[.f	
+%		 		\text{$T[M_1]$}
+%		 		\text{$T[M_2]$}
+%		 		\text{$\dots$}
+%		 		\text{$T[M_n]$}		 				 			
+%			] 
+%	\item $\lambda x_1 \dots x_n . M$ translates into tree\\
+%	 	\Tree
+%			[.\text{$\lambda x_1 \dots x_n$ }	
+%		 	 	\text{$T[M]$}	
+%			] 
+%\end{itemize}
+%
+%Notice that this representation does not cover all \lterms, 
+%e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us. 
+%
+%We will refer to this representation as to \textit{\sexprTree}.\\ 
+%
+%\Lets now consider representation for \textit{typed \lterms}.
+%Straightforward approach would be to add to each node a type entry which 
+%would be the type of the \lterm corresponding to subtree having this
+%node as the root node. 
+%
+%Approach more suitable for our purpose is to add a special type node above each node.
+%More specifically:
+%
+%\Lets consider tree $t$ corresponding to a \lterm of a type
+%$\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
+%Then corresponding tree $TT[t]$ for typed \lterm is 
+%obtained from the tree $t$ as follows:  
+%
+%\begin{equation*}
+%\mbox{ 
+%TT[
+%\Tree
+%	[.r 	
+%	  	  \text{$s_1$}
+%		  \text{$s_2$}
+%		  \text{$\dots$}
+%		  \text{$s_n$}
+%	] 
+%}]=
+%\mbox{
+%\Tree
+%	[.\text{$\sigma$ }
+%	    [.r 	
+%	  	  \text{$TT[s_1]$}
+%		  \text{$TT[s_2]$}
+%		  \text{$\dots$}
+%		  \text{$TT[s_n]$}
+%		]	  	
+%	] 
+%}
+%\end{equation*}

### 2013-07-22 tex/thesis.tex
[tex]

@@
 set $T$.
+
+\begin{todo}
+ & Dozjistit a dopsat jak to Koza má s vyžadovánim unikátnosti při generování.
+\end{todo}
 
@@
 \newtheorem{theorem}{Theorem} 
-\newtheorem*{proposition}{Proposition} 
+\newtheorem{proposition}{Proposition} 
 \newtheorem{lemma}{Lemma} 
@@
 \begin{todo}
-
-& ujistit se že je v části o tree reprezentacích 
-  je napsaný co to znamená že podstrom má nějakej typ,
-  že podstromy odpovídaj podvýrazum atd
-
- & říct že to je přesně ta notace co používá koza
- & EXAMPLES of tree representations of \lterms  
  & napsat transformace mezi \atTree a \sexprTree
- & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
-   přímo \sexprTree a říct proč se tomu tak řiká 
-   (na tom místě co je v textu dřív)
-
- & zkontrolovat že reference tree-reps se odkazuje na dobrý místo,
-   mohlo by se to rozbít tim že tree reprezentations - typovany/netypovany 
-   jsou od sebe oddelený
+%& ujistit se že je v části o tree reprezentacích 
+%  je napsaný co to znamená že podstrom má nějakej typ,
+%  že podstromy odpovídaj podvýrazum atd
+% & EXAMPLES of tree representations of \lterms  
+% & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
+%   přímo \sexprTree a říct proč se tomu tak řiká 
+%   (na tom místě co je v textu dřív) 
+% & zkontrolovat že reference tree-reps se odkazuje na dobrý místo,
+%   mohlo by se to rozbít tim že tree reprezentations - typovany/netypovany 
+%   jsou od sebe oddelený
 \end{todo}
@@
 \begin{todo}
-   & říct že se vždycky zastavěj (strong normalization),
+   & říct že se vždycky zastavěj (strong normalization)
+     && a jaký to má výhody pro nás v GP (duležitý),
      && potřebný definice pro to uvést v \lterms sekci
-     && musí se říc t co, tzn simpli typed l. kalkulus termy
-        a betaeta-redukce
-     && řekl bych že eta redukce je SN vždycky (tzn i untyped lk)
-        pač plati (CR+terminující pak SN) a je CR vzhledem k barendrecht-stará-žlutá
-        a terminující je rpotože zmenšuje term s nemuže zmenšovat donekonečna    
+     && musí se říct co je strongly normalizing, 
+        tzn simpli typed $lambda$-kalkulus termy a $\beta$-redukce /
+        $\beta\eta$-redukce
+%     && řekl bych že eta redukce je SN vždycky (tzn i untyped lk)
+%        pač plati (CR+terminující pak SN) a je CR vzhledem k barendrecht-stará-žlutá
+%        a terminující je rpotože zmenšuje term s nemuže zmenšovat donekonečna    
    & možná zmínit že se jedná o \textit{gentzen style proof}
@@
 \subsection{Tree representations of typed \lterms}
+\label{typed-tree-reps}
 
@@
 $\sigma$ with root $r$ and subtrees $s_1 , \dots , s_n$. 
-Then corresponding \textit{typed \sexprTree} $TT_S[t]$ for typed \lterm is 
+Then corresponding \textit{typed \sexprTree} $TT[t]$ for typed \lterm is 
 obtained from the tree $t$ as follows.
@@
 	] 
-}]=
+]} =
 \mbox{
@@
 We silently suppose here that type information about every subterm is available.
+
+%And analogically we can define transformation $TTtoTerm[]$ as follows.
 
@@
 \begin{todo}
- & co to znamená mít lnf
- & potřebuju eště nějaký \textbf{propositions} z tyhle knihy?
+ & potřebuju eště nějaký \textbf{propositions} z \cite{barendregt10} tu uvíst ?
  & ta definice je tam daná pro $\tur{}{M}{\sigma}$,
@@
 ( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
+~~~~ \textbf{if } n > 0
 \]
@@
 \begin{proposition}
+\label{gram-lnf-prop}
 Let context $\Gamma$, \lterm $M$ and type $\sigma$ be given.\\
@@
 \begin{todo}
- & zmínit že podobně jako mužeme transformovat inference rules
+ & možná zmínit že podobně jako mužeme transformovat inference rules
    na grammar rules, mužeme i obráceně grammar rules
@@
 
-	\item By \textit{solving and-or tree} $T$ is meant finding finite tree $T'$
-	      subgraph of $T$ such that following conditions are met. 
+%	\item By \textit{solving and-or tree} $T$ is meant finding finite tree $T'$
+%	      subgraph of $T$ such that following conditions are met. 
+    \item $T'$ \textit{solves} and-or tree $T$ if $T'$ satisfy following conditions.
 	    \begin{enumerate}
+	      \item $T'$ is a finite tree, subgraph of $T$. 
 		  \item The root of $T'$ is the root of $T$.
@@
 
-The motivation for defining \textit{solving} of a And-or tree the way we did is that
-a found tree $T'$ corresponds to generated \lterm. 
-In order to understand this correspondence let's now talk about various tree representations
-of \lterms.
+\newpage
+\section{ Generating terms in \lnf by solving inhabitation tree }
+
+The motivation for defining \textit{solving} and-or tree the way we did is that
+solution trees of $\IT(\sigma,\Gamma)$ are  
+\textit{typed \sexprTree{}s} representing all \lterms M in \lnf such that \GMS.
+
+\begin{proposition}
+\label{gram-it-prop}
+(Correctness)\\
+If \sexprTree $M'$ solves $\IT(\sigma,\Gamma)$,\\ 
+then $M'$ represents \lterm $M$ in \lnf such that \GMS. 
+\end{proposition}
+\begin{proof}
+
+We will prove following statement from which follows 
+the rest by proposition \ref{gram-lnf-prop}:
+If \sexprTree $M'$ solves $\IT(\sigma,\Gamma)$,
+then $M'$ represents \lterm $M$ such that $(\sigma,\Gamma) \garr M$.\\
+
+By induction on size of $M'$.\\
+
+\textbf{If} $\sigma = \alpha$ such that $\alpha \in A$, then
+
+\mbox{
+$M' = $ 
+\Tree 
+[.$\alpha$ 
+   [.$f$ $M'_1$ $\dots$ $M'_m$ ]
+]}\\
+
+For some $f : \rho_1 \ar \dots \ar \rho_m \ar \alpha \in \Gamma$.\\
+
+One can easily see that $M'_i$ solves $\IT(\rho_i,\Gamma)$.
+
+Thus by induction hypothesis we get that
+
+$M'_i$ represents \lterm $M_i$ such that $(\rho_i,\Gamma) \garr M_i$.\\
+
+$(\sigma,\Gamma) \gar (~f~(\rho_1,\Gamma)~\dots~(\rho_m,\Gamma)~)$,
+thus $(\sigma,\Gamma) \garr (~f~M_i~\dots~M_i~)$.
+
+Since $M'$ represents $(~f~M_i~\dots~M_i~)$, the proof for the first case is complete.\\
+
+\textbf{Otherwise}, for $\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha$ 
+such that $n> 0, \alpha \in A$
+
+\mbox{$M' = $
+\Tree [.$\tau_1\ar\dots\ar\tau_n\ar\alpha$ 
+ [.$\lh{x_1~\dots~x_n}$ [.$N'$ ] ] ]    }
+
+Again, one can easily see that $N'$ solves 
+$\IT(\alpha;\Gamma,x_1:\tau_1,\dots,x_n:\tau_n)$.
+
+Thus by induction hypothesis we get that
+
+$N'$ represents \lterm $N$ such that 
+$( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  ) \garr N$.\\
+
+$
+( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
+\gar
+(~\lambda~x_1~\dots~x_n~.~
+( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
+$,
+
+thus $(\sigma,\Gamma) \garr (~\lambda~x_1~\dots~x_n~.~N~)$.
+
+Since $M'$ represents $(~\lambda~x_1~\dots~x_n~.~N~)$, the proof is complete.
+
+\end{proof}
+
+\newpage
+\begin{proposition}
+(Completeness)\\
+If $M$ is \lterm in \lnf such that \GMS,\\
+then there exists exactly one \sexprTree $M'$ such that 
+$M'$ represents $M$ and
+$M'$ solves $\IT(\sigma,\Gamma)$.
+\end{proposition}
+\begin{proof}
+We will prove following statement from which follows 
+the rest by proposition \ref{gram-lnf-prop}:
+If $M$ is \lterm such that $(\sigma,\Gamma) \garr M$,
+then there exists exactly one \sexprTree $M'$ such that 
+$M'$ represents $M$ and
+$M'$ solves $\IT(\sigma,\Gamma)$.\\
+
+By induction on the length of rewriting chain of $(\sigma,\Gamma) \garr M$.\\
+
+\textbf{If} $\sigma = \alpha$ such that $\alpha \in A$, then:
+
+$(\sigma,\Gamma)$ must rewrite to
+$(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)$.
+
+And there is only one possibility for $f$ since $M$ is given.
+
+$M = (~f~M_1~\dots~M_m~)$ where $( \rho_i , \Gamma ) \garr M_i$.\\
+
+From induction hypothesis we have that 
+there exists exactly one \sexprTree $M_i'$ such that 
+$M_i'$ represents $M_i$ and
+$M_i'$ solves $\IT(\rho_i,\Gamma)$.\\
+
+In order to construct $M'$ solving $\IT(\sigma,\Gamma)$ 
+we have only one possibility since we know the right $f$ to take:
+
+\mbox{
+$M' = $ 
+\Tree 
+[.$\alpha$ 
+   [.$f$ $M'_1$ $\dots$ $M'_m$ ]
+]}\\
+
+\textbf{Otherwise}, for $\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha,$ $(n > 0)$:\\
+
+$(\sigma,\Gamma)$ must rewrite to
+$(~\lambda~x_1~\dots~x_n~.~( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n )~)$.
+
+$M = (~\lambda~x_1~\dots~x_n~.~N~)$ where 
+$( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n ) \garr N$.\\
+
+From induction hypothesis we have that 
+there exists exactly one \sexprTree $N'$ such that 
+$N'$ represents $N$ and
+$N'$ solves $\IT(\alpha,\Gp)$.\\
+
+In order to construct $M'$ solving $\IT(\sigma,\Gamma)$ in non-atomic $\sigma$
+we have only one possibility:
+
+\mbox{$M' = $
+\Tree [.$\tau_1\ar\dots\ar\tau_n\ar\alpha$ 
+ [.$\lh{x_1~\dots~x_n}$ [.$N'$ ] ] ]    }
+
+\end{proof}
 
 \begin{todo}
- & Přepsat poslední odstavec --- počítal ještě s tím že 
-   pasaže o stromové reprezentaci jsou až po této kapitole, což bylo změněno.
-\end{todo}
-
-\section{ Generating terms in Long normal form by solving Inhabitation tree }
-
-Now we can finally put the pieces together. 
-Every solution to an Inhabitation tree has this just described tree form of a typed \lterm.
-
-\begin{todo}
-& !!!
-& TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
-		Things about $\beta\eta^{-1}$ normal form, etc.   
+& skontrolovat ty dukazy pořádně sepsal sem 
+  je dost rychle a mužou bejt divný. 
+& A možná to ještě to nějak okomentovat
 \end{todo}
@@
 
-\begin{todo}
- & několik fází, několik reprezentací
- & během generování se reprezentace mění 
-   && nejdří obsahuje lambdy i volný proměnný pak už ne
-      &&& vygeneruje se v \sexprTree
-      &&& ta se převedena na \atTree aby jsme provedli abstraction elimination
- & dvě možnosti jak uchovávat jedince po vygenerování
-   && \sexprTree
-      &&& currently implemented
-   && \atTree
-      &&& currently not implemented, ale neni problém dodělat
-      &&& DOĎELAT! :)
-   && ale de mezi nima volně převádět
-      &&& díky tomu že nemaj lambda abstrakce, jinak by museli bejt ve 
-          WHNF/HNF nebo něco takovýho
-   && dusledky pro křížení
-      &&& \atTree umožňuje víc křížení než \sexprTree
-          &&&& každý co de v \sexprTree umí i \atTree ale obráceně to neplatí
-      &&& \atTree je zase ekvivalentní tomu kozovskýmu  
- & říct že bilding blocks je prostě kontext
-\end{todo}
+We have two competing variants for representing individuals ---
+\textit{\atTree} and \mbox{\textit{\sexprTree}} (see \ref{tree-reps}).
+
+Individuals generated by solving inhabitation tree 
+are \sexprTree{}s, which can be easily translated to \atTree{}s. \\
+
+Before we start generating we do one important trick.
+Suppose we want individuals of the type 
+$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha$
+from building blocks context $\Gamma$.
+
+All our individuals will have form $\lam{x_1\,\dots\,x_n}{M}$.
+%Later in this chapter we will discuss problems occurring with variables
+%when performing subtree swap in order to perform crossover.
+But since those variables are same for all individuals, 
+we may do following trick.
+
+Instead of generating individuals for
+$(\tau_1 \ar \dots \ar \tau_n \ar \alpha,\Gamma)$
+we generate individuals for $(\alpha;\Gamma,x_1:\tau_1,\dots,x_n:\tau_n)$
+and remember to add the $\lambda$-head $\lh{x_1\,\dots\,x_n}$ to
+each individual when evaluating it.\\
+
+This trick has two reasons:
+\begin{enumerate}
+ \item Later in this chapter we will discuss problems occurring with variables
+       when performing subtree swap in order to perform crossover.
+       We solve those problems by performing \textit{abstraction elimination},
+       which eliminates all occurences of variables and $lambda$-abstractions
+       in a \lterm. But those problems are caused by local variables. 
+       Since $x_1\,\dots\,x_n$ are common for all \lterms, it is valid
+       to exclude them from this elimination. Since abstraction elimination
+       prolongs size of \lterms it is beneficial to lessen 
+       the effect of abstraction elimination in this way.
+
+ \item It also makes our method compatible with classical GP, where
+       variables are treated as ordinary element of the set $T$
+       and no $\lambda$-abstraction is used.
+
+       This ensures that S-expressions will be generated for
+       $\Gamma$ satisfying constrains of classical GP and that
+       abstraction elimination will have no effect on such individuals.
+\end{enumerate}
+
+%\begin{todo}
+% & několik fází, několik reprezentací
+% & během generování se reprezentace mění 
+%   && nejdří obsahuje lambdy i volný proměnný pak už ne
+%      &&& vygeneruje se v \sexprTree
+%      &&& ta se převedena na \atTree aby jsme provedli abstraction elimination
+%\end{todo}
 
@@
 In order to describe our state space to search in \lets introduce
-extended version of our tree representation for typed \lterms from \ref{tree-reps}.
+extended version of our tree representation for typed \lterms from \ref{typed-tree-reps}.
 This extension consists in adding a new kind of leaf node $(\sigma,\Gamma)$ 
@@
 & eště se musí nějak prohrotit to, že tam je navíc ta lambda hlava
-kerá ještě o něco kazí tu $depth$
-& ukázat že je to fakt to samý co kozovský klasický
-& zavýst tam pojem běhu pro ten jeden vnější while cyklus
+kerá ještě o něco kazí tu $depth$  - to už uplně neplatí tim že je v
+uvodu popsaný to že se ta hlava ukousne ale stejně to ještě pořádně zkontrolovat
+& ukázat že je to fakt to samý co kozovský klasický generování
+& zavést tam pojem run pro ten jeden vnější while cyklus 
+ (i když to je asi jasný z kontextu)
 \end{todo}
@@
 \label{typed-swapping}
+
+\begin{todo}
+   & rozdíli při křížení ruznejch reprezentací 
+      && \atTree umožňuje víc křížení než \sexprTree
+          &&& každý co de v \sexprTree umí i \atTree ale obráceně to neplatí
+      && \atTree je zase ekvivalentní tomu kozovskýmu  
+\end{todo}
 

### 2013-07-23 v4/Eva.hs
v4

@@ -1,0 +1,110 @@
+{-# LANGUAGE TypeSynonymInstances, FlexibleInstances #-}
+
+module Eva where
+
+import Data.List
+import Data.Map.Strict (Map)
+import qualified Data.Map.Strict as Map
+
+import System.Random       ( StdGen, getStdGen, split )
+import Control.Monad.State ( StateT, runStateT, put, get, liftIO )
+import Data.Typeable       ( Typeable )
+import Data.Time.Clock     ( getCurrentTime, diffUTCTime )
+
+import Text.JSON ( encode , JSValue )
+
+import ServerInterface (jsEmptyObj, stdoutCmd, writeNextOutput)
+import Heval (hevalsWith, hevals)
+import Utils (Randable(..),Logable(..))
+
+
+type Eva = StateT EvaState IO
+
+type JobID = Int
+
+data EvaState = EvaState {
+     evaGen    :: StdGen ,
+     evaJobID  :: Maybe JobID ,
+     evaStdout :: String
+  }
+
+runEva :: Eva a -> IO a
+runEva eva = do
+ gen <- getStdGen
+ runEvaWith gen eva 
+
+runEvaWith :: StdGen -> Eva a -> IO a
+runEvaWith gen eva = do
+ startTime <- getCurrentTime
+ putStrLn $ "\nstdGen: " ++ show gen
+ ( ret , evaState' ) <- runStateT eva (initEvaState gen)
+ finishTime <- getCurrentTime
+ putStrLn $ "Total time: " ++ show (diffUTCTime finishTime startTime) ++ "\n"
+ return ret
+
+initEvaState :: StdGen -> EvaState
+initEvaState gen = EvaState { 
+  evaGen    = gen,
+  evaJobID  = Nothing,
+  evaStdout = "" 
+ }
+
+
+setJobID :: Int -> Eva ()
+setJobID jobID = do
+  evaState <- get
+  put $ evaState{ evaJobID = Just jobID }
+
+evalsWith :: (Typeable a) => String -> [String] -> a -> Eva [a]
+evalsWith file strs as = liftIO $ hevalsWith file strs as
+
+evals :: (Typeable a) => [String] -> a -> Eva [a]
+evals strs as = liftIO $ hevals strs as
+
+flushStdout :: Eva JSValue
+flushStdout = do
+  evaState <- get
+  case evaJobID evaState of
+    Nothing -> return jsEmptyObj
+    Just _  -> do
+      let stdout = evaStdout evaState
[…]

### 2013-07-23 v4/GP_Data.hs

@@ -1,0 +1,520 @@
+{-# LANGUAGE MultiParamTypeClasses #-}
+{-# LANGUAGE FlexibleInstances     #-}
+{-# LANGUAGE TupleSections         #-}
+
+module GP_Data where
+
+import Eva (Eva,evaSplitStdGen)
+import GP_Core ( Gene, Muta, Cros, generateIt, mutateIt, crossIt , Prob )
+import Utils   ( getRandom, getRandomR, getRandomL, getNormal, randIf, randCase,logIt,boxIt,boxThem )
+
+import TTerm
+import TTree
+
+import IM3 ( SearchOptions , prove , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
+--import IM2 ( SearchOptions , prove , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
+import System.Random (StdGen)
+
+--import InhabitationMachines ( proveN , randProveN, randProveUnique, proveOneWithLimit, Limit, kozaProveN )
+
+import KozaTree ( KTree(KNode), KPos, kSubtree, kChangeSubtree, kPoses, kPoses2, kDepth ) 
+import Dist     ( Dist, mkDist, distSize, distToList )
+
+
+import Control.Monad (replicateM, liftM, liftM2, forM)
+
+import Data.List     (nub,groupBy)
+
+type Len    = Int
+type Mean   = Double
+type StdDev = Double
+
+-- CTT -------------------------------------------------------------------------
+
+instance Gene CTT CTTGen where generateIt = cttGen
+instance Cros CTT CTTCro where crossIt    = cttCro2
+
+data CTTGen  
+-- = CTTG_Koza     Typ Context Limit  -- pomocí starýho IM 
+ = CTTG_Koza2    Typ Context        -- a zbytek pomocí novýho IM ..
+ | CTTG_AllEdges Typ Context 
+ | CTTG_Geom     Typ Context Prob 
+
+data CTTCro = CTTC_Koza
+
+cttGen :: Int -> CTTGen -> Eva [CTT]
+cttGen n opt = case opt of
+  CTTG_AllEdges typ ctx   -> proveIt  allEdgesSearchOptions n typ ctx 
+  CTTG_Koza2    typ ctx   -> proveIt  kozaSearchOptions     n typ ctx 
+  CTTG_Geom     typ ctx p -> proveIt (geomSearchOptions p)  n typ ctx
[…]

### 2013-07-23 v4/IM2.hs

@@ -1,0 +1,867 @@
+--  Nový pokus o Inhabitation Machines bez potřeby generovat graf dopředu a vůbec to udělat pořádně
+
+--  Asi by to chtělo celý předelat tak, že ZTree bude stavová monada, ale radši 
+-- to nejdřív už dodělat když je to takle rozdělaný a až pak dyštak překopat.
+
+
+-- IM2 = přidání polymorfizmu
+
+module IM2 
+ ( SearchOptions(..) 
+ , defaultSearchOptions 
+ , kozaSearchOptions
+ , allEdgesSearchOptions
+ , geomSearchOptions 
+ , prove 
+ ) where
+
+import Debug.Trace
+
+import TTerm (Symbol,Typ(..),TTerm(..),Context,ttermTyp,toSki,toSki',checkTyp)
+
+import TTree (CTT,mkCTT2)
+
+import Data.List
+import Data.Maybe
+import qualified Data.Set as Set
+import qualified Data.Map as Map
+import Data.Set (Set)
+import Data.Map (Map)
+
+import System.Random
+
+import qualified Data.PSQueue as Q
+
+import PolyUtils ( Substi , match , applySubsti , composeSubsti, emptySubsti , mgu )
+
+
+typeArgz :: Typ -> ([Typ],Typ)
+typeArgz typ = case typ of
+ (a :-> b) -> let (  as, alpha) = typeArgz b
+               in (a:as, alpha)
+ typHead   -> ([],typHead)
+ 
+
+
+--proveAll :: Int -> Typ -> Context -> [CTT]
+
+prove :: SearchOptions -> [CTT]
+prove so = 
+  let ( so' , problemHead ) = problemHeadPreproccess so
+      trees = proveWith2 so'
+   in map ( mkCTT2 problemHead . toSki . tree2tterm ) trees   -- <============== toSki s ' je s typeCheckem ...........
+
+problemHeadPreproccess :: SearchOptions -> ( SearchOptions , Context )
+problemHeadPreproccess so = 
+  let ( typ' , ctx' , problemHead ) = problemHeadPreproccess' (so_typ so) (so_ctx so)
+   in ( so{ so_typ = typ' , so_ctx = ctx' } , problemHead )
+
+problemHeadPreproccess' :: Typ -> Context -> (Typ,Context,Context)
+problemHeadPreproccess' typ ctx = 
+  let (ts,typHead) = typeArgz typ
+      ss           = map (\i->'x':show i) [0..]
+      problemHead  = zip ss ts
[…]

### 2013-07-23 v4/ProblemUtils.hs

@@ -1,0 +1,234 @@
+module ProblemUtils (
+  cttProblem5,
+  cttProblem4,
+  cttProblem3,
+  cttProblem2,
+  cttProblem2',
+  boolListProblem2,
+  asType,
+  casesFF,
+  ProblemOpts(..),PO_BLP(..),PO_CTTP(..),IntSlider(..),JobID,runProblemOpts,po2json,poCode,json2po
+) where
+
+import GP_Core (FitFun(..),Problem(..),NumGene,PopSize,mkGenOps)
+import GP_Core (GenOpProbs,FitVal,mkFF1)
+import GP_Core ( nRunsByServer )
+import GP_Data (CTTGen(..),CTTCro(..))
+import GP_Data (ListGen(..),ListCro(..),ListMut(..), BoolGen(..), BoolMut(..), Len)
+
+
+import TTree (CTT)
+import TTerm (Typ,Context)
+
+import Text.JSON
+import JSONUtils
+
+import Data.Typeable ( Typeable )
+
+-- ---------------------------------------------------------------
+
+
+
+data ProblemOpts a = PO_BLP_ PO_BLP | PO_CTTP_ (PO_CTTP a) 
+
+
+poCode :: ProblemOpts a -> String
+poCode po = case po of
+  PO_BLP_  x ->  blp_code x
+  PO_CTTP_ x -> cttp_code x
+
+po2json :: ProblemOpts a -> JSValue
+po2json po = case po of
+  PO_BLP_  x ->  blp2json x
+  PO_CTTP_ x -> cttp2json x
+
+json2po :: ProblemOpts a -> JSValue -> ProblemOpts a
+json2po po json = case po of
+  PO_BLP_  x -> PO_BLP_  $ json2blp  x json
+  PO_CTTP_ x -> PO_CTTP_ $ json2cttp x json
+
+
+
+data PO_BLP = PO_BLP {
+  
+  blp_numRuns     :: IntSlider        , 
+  blp_numGene     :: IntSlider        , 
+  blp_popSize     :: IntSlider        , 
+  blp_length      :: IntSlider        ,
+
+  blp_code        :: String           ,
+  blp_info        :: String           ,
+  blp_data        :: JSValue          ,
+  blp_ff          :: [Bool] -> Double }
+
+
+blp2json :: PO_BLP -> JSValue
+blp2json blp = jsObj [
+  ( "code"    , jsStr          $ blp_code    blp                                  ),
+  ( "info"    , jsStr          $ blp_info    blp                                  ),
[…]

### 2013-07-23 v4/Problems/Ant/Ant.hs

@@ -1,0 +1,239 @@
+--{-# LANGUAGE FlexibleInstances #-}
+--{-# LANGUAGE TypeSynonymInstances #-}
+
+-- Problém s mravencem a jídlem z Kozy
+
+module Problems.Ant.Ant (
+ AAnt, toAAnt , toAAnt2 , toAAnt3, unAAnt , 
+ Ant , World , antWorld , evalAnt , runAnt,
+ doLeft, doRight , doMove , ifFoodAhead , progn2 , progn3
+) where 
+
+import qualified Data.Set as Set
+import Control.Monad.State
+
+type Size  = (Int,Int)
+type Limit = Int
+type Pos   = (Int,Int)
+type AntPos= Pos
+data Dir   = DUp | DDown | DLeft | DRight
+data Action= ALeft | ARight | AMove
+data World = World Size Limit (AntPos,Dir) Stats ( Set.Set Pos ) ( Set.Set Pos ) --deriving (Show)
+
+type Steps = Int
+type Eaten = Int
+type Stats = (Eaten,Steps)
+
+type Ant = State World ()
+
+newtype AAnt = AAnt { unAAnt :: Ant } 
+
+toAAnt :: Ant -> AAnt
+toAAnt ant = AAnt ant
+
+toAAnt2 :: (Ant -> Ant -> Ant) -> (AAnt -> AAnt -> AAnt)
+toAAnt2 f aa1 aa2 = AAnt $ f (unAAnt aa1) (unAAnt aa2) 
+
+toAAnt3 :: (Ant -> Ant -> Ant -> Ant) -> (AAnt -> AAnt -> AAnt -> AAnt)
+toAAnt3 f aa1 aa2 aa3 = AAnt $ f (unAAnt aa1) (unAAnt aa2) (unAAnt aa3) 
+
+
+---------data ANT = ANT (State World ()) deriving Typeable
+
+--ant :: Typ
+--ant  = Typ "Ant"
+--
+--
+--instance Typeable Ant where
+--  typeOf _ = mkTyConApp (mkTyCon "Ant") []
+--
+-- 
+--antEnv :: Env
+--antEnv =  [
+--          ( "l"   , ant                         ,  toDyn' doLeft )
+--          ,("r"   , ant                         ,  toDyn' doRight )
+--          ,("m"   , ant                         ,  toDyn' doMove )
+--          ,("ifa" , ant :-> ant :-> ant         ,  toDyn' ifFoodAhead )
[…]

### 2013-07-23 v4/Problems/BA/Problem.hs

@@ -1,0 +1,35 @@
+module Problems.BA.Problem where
+
+--import Problems.Utils ( boolListProblem2 )
+import ProblemUtils 
+import Text.JSON
+
+
+reg :: ProblemOpts a
+reg = PO_BLP_ PO_BLP {
+  blp_code        = "ba"                                         ,
+  blp_info        = "Bool Alternate"                             ,
+  blp_data        = JSNull                                       ,
+  blp_numRuns     = IntSlider "Runs"            1 10    4    1   ,
+  blp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  blp_popSize     = IntSlider "Population size" 0 1000  500 100  , -- dřív 5000 ale začlo to s předěláním Evy dělat stackoverflow :(
+  blp_length      = IntSlider "Genom length"    0 1000  200  10  ,
+  blp_ff          = ff
+}
+
+
+
+--mainProblem_ = boolListProblem2 "ba" (33,33,33) 100 ff
+
+
+
+
+
+ff :: [Bool] -> Double
+ff bits = 
+ let dLen = fromIntegral . length
+     tNum = dLen . filter (\(n,b)->if n `mod` 2 == 0 then b else not b) $ zip [0..] bits 
+     len  = dLen $ bits
+  in 100*(tNum * tNum)/(len*len)
+
+

### 2013-07-23 v4/Problems/BigCtx/Problem.hs

@@ -1,0 +1,229 @@
+module Problems.BigCtx.Problem where
+
+import TTerm (Typ(..),Context)
+import ProblemUtils 
+import JSONUtils
+import GP_Core (FitFun(FF6))
+import GP_Data (CTTGen(..))
+
+import DrawIM( imGraphInJSON )
+
+ 
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "bigctx"                                     ,
+  cttp_info        = "Problem with basic funs for types generating elementary funs." ,
+  cttp_data        = jsObj [ ( "im" , imGraphInJSON prog_typ prog_ctx ) ]      ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = prog_typ                                     ,
+  cttp_ctx         = prog_ctx                                     ,
+  
+  cttp_gOpt        = CTTG_Koza2 prog_typ prog_ctx                 , 
+
+  cttp_ff          = FF6 prog_type ff_fst3 "Problems.BigCtx.Funs" 
+  
+}
+
+prog_type = asType :: Fst3_Type  -- asType :: [Int] -> Maybe Int
+prog_typ  = fst3_typ             -- l_int :-> m_int  
+prog_ctx  = ctx_fst3             -- ctx_combo
+
+
+reg_head =
+ let head_type = asType :: [Int] -> Maybe Int
+     head_typ  = l_Int :-> m_Int
+     head_ctx  = ctx_poly_head
+  in PO_CTTP_ PO_CTTP {
+      cttp_code        = "head"                                     ,
+      cttp_info        = "Evolving safe polymorphic head function ", --from big context." ,
+      cttp_data        = jsObj [] ,  --jsObj [ ( "im" , imGraphInJSON head_typ head_ctx ) ]      ,
+      cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+      cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+      cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+      
+      cttp_typ         = head_typ                                     ,
[…]

### 2013-07-23 v4/Problems/EvenParity/Problem.hs

@@ -1,0 +1,173 @@
+module Problems.EvenParity.Problem where
+
+import TTerm (Typ(..),Context)
+import ProblemUtils 
+import JSONUtils
+import GP_Core (FitFun(FF6))
+import GP_Data (CTTGen(..))
+import DrawIM( imGraphInJSON )
+
+import Problems.EvenParity.Funs
+
+
+prog_type = asType :: [Bool]  -> Bool  
+prog_typ  =           l_bool :-> bool             
+prog_ctx  = ctx_yu             
+
+
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "ep"                                     ,
+  cttp_info        = "Even-parity problem." ,
+  cttp_data        = jsObj [ ( "im" , imGraphInJSON prog_typ prog_ctx ) ] ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 500   50   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = prog_typ                                     ,
+  cttp_ctx         = prog_ctx                                     ,
+  
+  cttp_gOpt        = CTTG_Geom     prog_typ prog_ctx 0.75 , 
+--cttp_gOpt        = CTTG_Koza2    prog_typ prog_ctx  , 
+--cttp_gOpt        = CTTG_AllEdges prog_typ prog_ctx  , 
+
+  cttp_ff          = FF6 prog_type ff "Problems.EvenParity.Funs" 
+  
+}
+
+
+
+bool   = Typ "Bool"
+l_bool = Typ "[Bool]"
+
+ctx = [
+ ( "foldr" , (bool:->bool:->bool) :-> bool :-> l_bool :-> bool ),
+ ( "(==)"  , bool :-> bool :-> bool                            ),
+ ( "(/=)"  , bool :-> bool :-> bool                            ),
+ ( "(&&)"  , bool :-> bool :-> bool                            ),
+ ( "(||)"  , bool :-> bool :-> bool                            ),
[…]

### 2013-07-23 v4/Problems/Fly/Problem.hs

@@ -1,0 +1,302 @@
+module Problems.Fly.Problem where
+
+
+import Data.List
+
+
+import TTerm          (Typ(..),Context)
+import ProblemUtils 
+ --( cttProblem5 , asType )
+
+import Text.JSON
+import JSONUtils
+
+import DrawIM( imGraphInJSON )
+
+import Problems.Fly.Funs ( Input_ , Output_ , Dir_  , Energy , Pos , avg , 
+                           output_ , myApplePoses_ ,nearestFlyPos_, myPos_ , posToDir_  ,
+                           dStay , dUp , dDown , dLeft , dRight )
+import Problems.Fly.Fly
+
+import GP_Core (FitFun(FF6))
+import GP_Data (CTTGen(..))
+
+
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "fly"                                        ,
+  cttp_info        = "Fly eating apples and stuff... "            ,
+  cttp_data        = jsData                                       ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = prog_typ                                     ,
+  cttp_ctx         = ctx                                          ,
+
+  cttp_gOpt        = CTTG_Koza2 prog_typ ctx                      , 
+  
+  cttp_ff          = FF6 prog_type ff "Problems.Fly.Funs" 
+  
+}
+
+problem1 = cttProblem5 "Fly" ff prog_typ ctx prog_type
+
+
+
+
+jsData = jsObj [ 
+ ( "world" , worldToJSON ff_solutionFlyPos ff_world_withEnvirFlies ),
+ ( "im"    , imGraphInJSON prog_typ ctx )]
+
+
+prog_type = asType :: Input_ -> Output_
+prog_typ  = input_typ :-> output_typ
+
+input_typ  = Typ "Input_"
+output_typ = Typ "Output_"
+dir_typ    = Typ "Dir_"
+
+m_pos   = Typ "Maybe Pos"
+l_pos   = Typ "[Pos]"
+pos_typ = Typ "Pos"
+
+dou     = Typ "Double"
[…]

### 2013-07-23 v4/Problems/Fly02/Problem.hs

@@ -1,0 +1,1167 @@
+module Problems.Fly02.Problem where
+
+import Data.List
+
+import Problems.Fly02.Funs  --(  )
+import Problems.Fly02.Fly2  --(  )
+
+import TTerm  (Typ(..),Context)
+import GP_Core (FitFun(FF6))
+import GP_Data (CTTGen(..))
+
+import DrawIM( imGraphInJSON )
+
+import ProblemUtils 
+
+import JSONUtils
+import Text.JSON
+
+
+--import Heval
+
+reg = PO_CTTP_ PO_CTTP {
+  cttp_code        = "fly02"                                      ,
+  cttp_info        = "Fly 0.2 eating apples and stuff... "        ,
+  cttp_data        = jsData                                       ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttp_typ         = prog_typ                                     ,
+  cttp_ctx         = ctx                                          ,
+
+  cttp_gOpt        = CTTG_Geom     prog_typ ctx 0.75 ,  --CTTG_Koza2 prog_typ ctx    , 
+  
+  cttp_ff          = FF6 prog_type ff "Problems.Fly02.Funs" 
+  
+}
+
+prog_type  = asType :: Prog_
+
+prog_typ   = input_typ :-> output_typ
+input_typ  = Typ "Input_"
+output_typ = Typ "Output_"
+move_typ   = Typ "Move_"
+regs_typ   = Typ "Registers_"
+dir_typ    = Typ "Dir_"
+bool_typ   = Typ "Bool" 
+int_typ    = Typ "Int" 
+dist_typ   = Typ "Double"
+
+
+--function (x0){return output_(if_(lte(xGet_(x0),zGet_(x0)),easySplit(x0),travel_(dLeft)),xInc_(myRegs_(x0)));} 
+
+ctx :: Context
+ctx = [
+
+
+  ( "output_"          , move_typ :-> regs_typ :-> output_typ         ) ,
+
[…]

### 2013-07-23 v4/Server.hs

@@ -1,0 +1,247 @@
+{-# LANGUAGE OverloadedStrings #-}
+ 
+import Network.Wai
+import Network.Wai.Handler.Warp
+import Network.HTTP.Types (status200)
+import Blaze.ByteString.Builder (copyByteString)
+import qualified Data.ByteString.UTF8 as BU
+import Data.Monoid
+import Data.Enumerator (run_, enumList, ($$))
+
+import Text.JSON ( encode )
+ 
+-- já
+import Control.Concurrent  
+import Control.Monad
+import Control.Monad.State
+import System.IO
+import System.Directory (doesFileExist, removeFile, createDirectory, doesDirectoryExist,removeDirectoryRecursive,getDirectoryContents)
+import System.Environment
+
+import Utils ( unescape )
+--import Job (job,problemList,problemList_,job_)
+import Job (problemList_,job_)
+
+import ServerInterface
+
+main = do
+    clearServer
+    (port:_) <- getArgs
+    putStrLn $ "Fishtron GUI is on http://localhost:" ++ port 
+    run (read port) app
+
+app :: Application
+app req = 
+ case pathInfo req of
+
+    ["out",i,j] -> do
+      let workerId = ( read . init . tail . show $ i ) :: Int
+      let outputId = ( read . init . tail . show $ j ) :: Int
+      outString <- liftIO $ serveOutput workerId outputId
+      return $ myTextPlain outString
+
+    [ ] -> return $ myIndex
+
+--    ["run",cmd] -> do 
+--      workerId <- liftIO newWorker
+--      let cmdStr = init . tail . show $ cmd 
+--      liftIO . forkIO $ runCmd workerId cmdStr
+--      return . myTextPlain . show $ workerId
+
+    ["run_",cmd] -> do 
+      workerId <- liftIO newWorker
+      let cmdStr = init . tail . show $ cmd 
+      liftIO . forkIO $ runCmd_ workerId (unescape cmdStr)
+      return . myTextPlain . show $ workerId
+
+--    ["problems"] -> return . myTextPlain . encode . problemListToJSON $ problemList
+
+    ["problems_"]-> return . myTextPlain . encode $ problemList_    
+
+    ["js",filename] -> do
+      return $ myJSFile (init . tail . show $ filename)
+
+    ["js","libs",filename] -> do
[…]

### 2013-07-23 v4/ServerInterface.hs

@@ -1,0 +1,83 @@
+module ServerInterface where
+
+import System.IO
+--import System.Directory
+
+import Text.JSON
+import JSONUtils
+
+writeNextOutput :: Int -> String -> IO ()
+writeNextOutput i output = do
+  outI <- incrementFile ("server/output/" ++ (show i) ++ "/_hotovo.txt")
+  myWriteFile ("server/output/"++ (show i) ++ "/" ++ (show outI )++ ".txt" ) output
+
+
+incrementFile :: FilePath -> IO Int
+incrementFile filename = do 
+  file <- openFile filename ReadWriteMode
+  str <- hGetLine file 
+  let num  = 1 + ( read str ) :: Int
+      str' = show num
+  hSeek file AbsoluteSeek 0
+  hPutStr file str'
+  hClose file 
+  return num
+
+
+myWriteFile :: FilePath -> String -> IO ()
+myWriteFile filename str = do 
+  file <- openFile filename WriteMode
+  hPutStr file str
+  hClose file
+
+myReadFile :: FilePath -> IO String
+myReadFile filename = do
+ file <- openFile filename ReadMode 
+ str <- hGetLine file 
+ hClose file
+ return str
+
+
+problemListToJSON :: [(String,String,JSValue)] -> JSValue
+problemListToJSON problemList =  jsArr . map toProblemObj $ problemList
+ where
+  toProblemObj :: (String,String,JSValue) -> JSValue
+  toProblemObj ( code , name , datas ) = jsObj [
+    ( "code" , jsStr code  ) ,
+    ( "name" , jsStr name  ) ,
+    ( "data" , datas       )
+   ]
+
+
+jsEmptyObj :: JSValue
+jsEmptyObj = jsObj []
+
+stdoutCmd :: String -> JSValue
+stdoutCmd str = jsObj [ 
+  ("type" , jsStr "stdout" ) ,
+  ("msg"  , jsStr str      ) ]
+
+graphCmd :: Int -> Int -> (Double,Double,Double) -> JSValue
+graphCmd runI genI (best,avg,worst) = jsObj [ 
+  ("type"   , jsStr "generationInfo" ) ,
+  ("i"      , jsNum genI ) ,
+  ("run"    , jsNum runI ) ,
+  ("ffvals" , jsObj [
+      ( "best"  , jsNum best  ) ,
[…]

### 2013-07-23 v4/TTerm.hs

@@ -1,0 +1,410 @@
+module TTerm
+( Typ (..)
+, TTerm (..)
+, TTermZipper (..)
+, TTermDrobek (..)
+, Context
+, Symbol
+, typeArgs
+, ttermTyp , isInCtx
+, mkTz, showTTZ, tzDown, tzLeft, tzRight, tzGoTop,ttzTyp
+, subs
+, makeVarsUnique 
+, checkTyp
+, toSki, toSki'
+) where
+
+import Data.List ( nub, (\\) , intercalate )
+import Data.Typeable ( Typeable )
+import Text.ParserCombinators.Parsec ( Parser, ParseError, parse, (<|>), eof, char, many1, noneOf, skipMany1, space )
+
+import Utils ( newSymbol )
+
+--import Util  ( newSymbol )
+--import Heval ( eval )
+
+---------------------------------------------------
+
+type Symbol = String
+
+data Typ  = 
+ Typ Symbol    |
+ TypVar Symbol | 
+ TypFun Symbol [Typ] |
+ Typ :-> Typ 
+ deriving (Eq,Ord)
+infixr 7 :->
+
+type Context = [(Symbol,Typ)]
+
+data TTerm = 
+ TVar Symbol        Typ | 
+ TVal Symbol        Typ | 
+ TLam Symbol TTerm  Typ |
+ TApp TTerm  TTerm  Typ
+ deriving (Ord,Eq)
+
+-- substitution ------------------------------------
+
+subs :: Symbol -> TTerm -> TTerm -> TTerm 
+subs x m n = fst $ sub (varNames m) m x n   
+
+sub :: [Symbol] -> TTerm -> Symbol -> TTerm -> (TTerm,[Symbol])
+sub vns x@(TVal _ _) _ _ = (x,vns)  
+sub vns x@(TVar v _) var vloz 
+    | v == var  = (vloz,vns)
+    | otherwise = (x,vns)  
+sub vns (TApp p q typ) var vloz = (TApp t1 t2 typ , vns'')
+    where 
+    (t1,vns')  = sub vns  p var vloz
+    (t2,vns'') = sub vns' q var vloz
+sub vns x@(TLam v p typ) var vloz
+    | v == var  = (x,vns)
+    | not (elem v (fv vloz)) = (TLam v t typ , vns'') 
+    | otherwise = sub vns' term' var vloz
+    where
+    (t,vns'') = (sub vns p var vloz)
+    (term',vns') = rename vns x
+
+varNames :: TTerm -> [Symbol]
[…]

### 2013-07-23 v4/Utils.hs

@@ -1,0 +1,239 @@
+{-# LANGUAGE FlexibleInstances, UndecidableInstances #-}
+-- , OverlappingInstances 
+
+module Utils where
+
+import Data.Maybe
+import Data.List
+import Data.Map (Map)
+import qualified Data.Map as Map
+
+import System.Random
+import Control.Monad.State
+import Data.Random.Normal
+
+
+import Text.JSON (JSValue(..))
+
+import JSONUtils
+
+-- general ------------------------------------
+
+putList :: (Show a) => [a] -> IO ()
+putList []     = return ()
+putList (x:xs) = do
+ putStrLn . show $ x
+ putList xs 
+
+asType :: a
+asType = undefined
+
+unescape :: String -> String
+unescape []     = []
+unescape (x:[]) = [x]
+unescape (x:y:rest)
+    | x == '\\' = y : unescape rest
+    | otherwise = x : unescape (y : rest)
+
+-- Pro již použitá slova vrací první takové, které je od všech rozdílné
+newSymbol :: [String] -> String
+newSymbol = newSymbol' ['a'..'z'] 
+
+newSymbol' :: [Char] -> [String] -> String
+newSymbol' abeceda vns = snd . head $ dropWhile (\(x,y)->x==y) 
+   (zip ((sortBy (nameOrderingBy abeceda) vns') ++ repeat "") (map (nthWord abeceda) [1..]))
+ where
+  vns' = filter (all (`elem` abeceda )) vns 
+
+-- Uspořádání primárne podle délky, sekundárne podle abecedy.
+nameOrderingBy :: (Eq a) => [a] -> [a] -> [a] -> Ordering
+nameOrderingBy abc x y
+  | lo /= EQ  = lo
+  | otherwise = compareByAbc abc x y
+ where lo = lenOrdering x y
+
+compareByAbc :: (Eq a) => [a] -> [a] -> [a] -> Ordering
+compareByAbc _ [] [] = EQ
+compareByAbc abc (x:xs) (y:ys) = case cByAbc abc x y of
+  EQ -> compareByAbc abc xs ys
+  x  -> x
+ where 
+  cByAbc :: (Eq a) => [a] -> a -> a -> Ordering
+  cByAbc abc x y
+   | x == y    = EQ
+   | otherwise = case dropWhile (\ch-> (ch/=x) && (ch/=y) ) abc  of
+      []  -> error "cByAbc : x and y are not in the alphabet"
[…]

### 2013-07-23 v4/fly-js/FlySim.js

@@ -1,0 +1,1022 @@
+
+var FlySim = function(){
+
+  var log = function(x){ console.log(x) };
+
+  var cur2 = function( fun ){
+    return function( x , y ){
+      if( y === undefined ){
+        return function(y){ return fun(x,y) ; }
+      } else {
+        return fun( x , y );
+      }
+    }
+  }; 
+  
+  var cur3 = function( fun ){
+    return function( x , y , z ){
+      if( y === undefined ){
+        return cur2( function(y,z){return fun(x,y,z);} );
+      } else if( z === undefined ){
+        return function(z){return fun(x,y,z);} ;
+      } else {
+        return fun(x,y,z);
+      }
+    }
+  };
+  
+
+  var minus = function( pos1 , pos2 ){ 
+    return [ pos1[0]-pos2[0] , pos1[1]-pos2[1] ]; 
+  };
+  
+  var dist = function( pos1 , pos2 ) {
+      var d = minus( pos1 , pos2 );
+      var dx = d[0] , dy = d[1];
+      return Math.sqrt(  dx*dx + dy*dy );
+  };
+
+
+  var getNearestPos = function( pos , poses ){
+ 
+    var bestPos  = undefined ;
+    var bestDist = 999999999 ;
+ 
+    for( var i in poses ){
+      var actPos = poses[i];
+      var dista = dist( pos , actPos );
+      if( dista < bestDist  ){
+        bestDist = dista;
+        bestPos  = actPos;
+      }
+    }
+  
+    //log(pos +' - ' + bestPos+' - '+bestDist);
+
+    if( bestPos === undefined ){ return null ;}
+    return [bestPos,bestDist] ;
+  };
+
+  var posToDir = function( posMy , posHer ){ 
+         
+    var d = minus( posHer , posMy );
+    var dx = d[0] , dy = d[1];
+         
+    if(   dx  > dy && (-dx) > dy ) { return dUp   ; }
+    if(   dx  > dy               ) { return dRight; }
+    if( (-dx) > dy               ) { return dLeft ; }
+    if( true                     ) { return dDown ; }
+        
+  };
+
+  var posPlusDir = function( pos , dir ){
+    var x = pos[0];
+    var y = pos[1];
+
+    switch( dir ){
+      case dUp    : return [x  ,y-1] ; 
+      case dDown  : return [x  ,y+1] ; 
+      case dLeft  : return [x-1,y  ] ; 
[…]

### 2013-07-23 v4/fly-js/fly-origo.js

@@ -1,0 +1,554 @@
+
+$(document).ready(function() {
+
+console.log( "Fly02.js ready!" );  
+
+
+var dStay   = 0; 
+var dUp     = 1;
+var dDown   = 2;
+var dLeft   = 3; 
+var dRight  = 4; 
+
+var lt = function(x,y){
+  return (x<=y) ;
+};
+
+var head_ = function( xs ){
+  if( xs.length == 0 ) return null;
+  return xs[0];
+};
+
+var avg = function( poses ){
+
+  if( poses.length == 0 ) return null;
+
+  var sum0 = 0;
+  var sum1 = 0;
+
+  for( var i in poses ){
+    var pos = poses[i];
+    sum0 += pos[0];
+    sum1 += pos[1];
+  }
+
+  return [ Math.round( sum0 / poses.length ) , Math.round( sum1 / poses.length ) ];
+};
+
+var output_ = function( dir_ ){
+  return dir_ ;
+};
+
+var myApplePoses_ = function( input_ ){
+  return input_[0];
+};
+
+var nearestFlyPos_ = function( input_ ){
+  return input_[1];
+};
+
+var myPos_ = function( input_ ){
+  return input_[2];
+};
+
+var inputEnergy_ = function( input_ ){
+  return input_[3];
+};
+
+
+var posToDir_ = function( posMy , posHer ){ 
+      
+  if( posHer === null ){
+    return dStay ;
+  }
+       
+  var d = minus( posHer , posMy );
+  var dx = d[0] , dy = d[1];
+       
+  if(   dx  > dy && (-dx) > dy ) { return dUp   ; }
+  if(   dx  > dy               ) { return dRight; }
+  if( (-dx) > dy               ) { return dLeft ; }
+  if( true                     ) { return dDown ; }
+      
+};
+
+var minus = function( pos1 , pos2 ){ 
+  return [ pos1[0]-pos2[0] , pos1[1]-pos2[1] ]; 
+};
+
+var dist = function( pos1 , pos2 ) {
+    var d = minus( pos1 , pos2 );
+    var dx = d[0] , dy = d[1];
+    return Math.sqrt(  dx*dx + dy*dy );
+};
+
+
+Global.Problems.fly02 = {
+
+  Funs : {
+    dStay           : dStay          ,   
+    dUp             : dUp            ,   
+    dDown           : dDown          ,   
+    dLeft           : dLeft          ,             
+    dRight          : dRight         ,   
+    lt              : lt             ,        
+    head_           : head_          ,
[…]

### 2013-07-24 tex/thesis.tex
v4 ...

@@
  & možná přidat ještě něco z poznámek co jsou teď v Conclusion
+ & vlastně i koza hrotí evolution of constrained syntactic structures !!!
+ & my tu řikáme že to závadíme kuli typum, ale my děláme v víc! -dáváme tam
+   i abstrakce atd - ale můžeme říct, že tim že se to pak převede zase na 
+   to bez abstrakcí tak setrváváme dal v krasnym bez-lambda abstrakčí ale
+   s tou výhodou že naše metoda gerneruje všechny struktury pro ten danej typ ...
+   nebo tak něco - asi to dát spíš do design of our system nebo někam
+   ale někde to asi zmínit
 \end{todo}
@@
 both are replaced by both parents.
+
+\begin{todo}
+ & POZOR POZOR !!!!! zapoměl jsem tam že je 90\% procent šance 
+   že se vybere internal node a jen 10\% šance že terminal node
+\end{todo}
 

### 2013-07-24 v4/Eva.hs

@@ -1,10 +1,16 @@
 {-# LANGUAGE TypeSynonymInstances, FlexibleInstances #-}
 
-module Eva where
+module Eva 
+( Eva
+, runEva
+, evalsWith
+, evals
+, setOutputBuffer
+, sendJSON
+, flushStdout
+, evaSplitStdGen
+) where
 
-import Data.List
-import Data.Map.Strict (Map)
-import qualified Data.Map.Strict as Map
 
 import System.Random       ( StdGen, getStdGen, split )
 import Control.Monad.State ( StateT, runStateT, put, get, liftIO )
@@ -13,7 +19,7 @@ import Data.Time.Clock     ( getCurrentTime, diffUTCTime )
 
 import Text.JSON ( encode , JSValue )
 
-import ServerInterface (jsEmptyObj, stdoutCmd, writeNextOutput, writeNextOutput_new ,OutputBuffer)
+import ServerInterface (jsEmptyObj, stdoutCmd, writeNextOutput ,OutputBuffer) 
 import Heval (hevalsWith, hevals)
 import Utils (Randable(..),Logable(..))
 
@@ -24,7 +30,6 @@ type JobID = Int
 
 data EvaState = EvaState {
      evaGen          :: StdGen ,
-     evaJobID        :: Maybe JobID ,
      evaOutputBuffer :: Maybe OutputBuffer,
      evaStdout       :: String
   }
@@ -46,16 +51,10 @@ runEvaWith gen eva = do
 initEvaState :: StdGen -> EvaState
 initEvaState gen = EvaState { 
   evaGen          = gen,
-  evaJobID        = Nothing,
   evaOutputBuffer = Nothing,
   evaStdout       = "" 
  }
 
-
-setJobID :: Int -> Eva ()
-setJobID jobID = do
-  evaState <- get
-  put $ evaState{ evaJobID = Just jobID }
 
 setOutputBuffer :: OutputBuffer -> Eva ()
 setOutputBuffer buff = do
@@ -69,19 +68,10 @@ evalsWith file strs as = liftIO $ hevalsWith file strs as
 evals :: (Typeable a) => [String] -> a -> Eva [a]
 evals strs as = liftIO $ hevals strs as
 
+
 flushStdout :: Eva JSValue
 flushStdout = do
   evaState <- get
-  case evaJobID evaState of
-    Nothing -> return jsEmptyObj
-    Just _  -> do
-      let stdout = evaStdout evaState
[…]

### 2013-07-24 v4/GP_Core.hs

@@ -1,14 +1,28 @@
 {-# LANGUAGE MultiParamTypeClasses, FlexibleInstances, TupleSections #-}
 
-module GP_Core where
+module GP_Core 
+( Gene(..)
+, Muta(..)
+, Cros(..)
+, Prob
+, FitFun(..)
+, Problem(..)
+, NumGene
+, PopSize
+, GenOpProbs
+, mkGenOps
+, FitVal
+, mkFF1
+, nRunsByServer
+) where
 
 import Control.Monad ( liftM )
 import Data.Typeable ( Typeable )
 import Text.Printf   ( printf )
-import Data.Maybe    ( fromJust )
+import Data.Maybe    ( fromJust,isJust )
 
 
-import Eva   ( Eva, runEva, setJobID, evalsWith, evals, flushStdout, sendJSON_ , setOutputBuffer, sendJSON_new, flushStdout_new )
+import Eva   ( Eva, runEva, evalsWith, evals, setOutputBuffer, sendJSON, flushStdout )
 import Dist  ( Dist , mkDist, distTake_new, distGet, distMax, distMin, distAvg )
 import Utils ( logIt, boxIt , JShow, jshow, putList)
 import ServerInterface ( graphCmd, multiCmd, OutputBuffer ) 
@@ -81,8 +95,8 @@ evolveBegin :: ( Gene t go, Typeable a,JShow t) => RunInfo -> Problem t a go mo 
 evolveBegin runInfo p = do
  let n = popSize p 
  terms <- generateIt n (gOpt p)
- ret@(pop0,_) <- evalFF (fitFun p) terms
- logGeneration runInfo 0 pop0
+ ret@(pop0,mWin) <- evalFF (fitFun p) terms
+ logGeneration runInfo 0 (isJust mWin) pop0
  return ret
 
 evolveStep ::(Muta t mo,Cros t co,Typeable a,JShow t)=> RunInfo->Problem t a go mo co -> Int -> Dist t -> Eva(Dist t,Maybe(t,FitVal))
@@ -90,8 +104,8 @@ evolveStep runInfo p i pop = do
  let best     =  getBest pop 
  terms        <- distTake_new (popSize p - 1) pop   
  terms'       <- performOps (genOps p) terms 
- ret@(pop',_) <- evalFF (fitFun p) ( best : terms' )
- logGeneration runInfo i pop'
+ ret@(pop',mWin) <- evalFF (fitFun p) ( best : terms' )
[…]

### 2013-07-24 v4/GP_Data.hs

@@ -8,11 +8,15 @@ import Eva (Eva,evaSplitStdGen)
 import GP_Core ( Gene, Muta, Cros, generateIt, mutateIt, crossIt , Prob )
 import Utils   ( getRandom, getRandomR, getRandomL, getNormal, randIf, randCase,logIt,boxIt,boxThem )
 
-import TTerm
-import TTree
+import TTerm (Typ,Context)
+import TTree (CTT(..),ttreeDepth,ttreeChangeSubtree,ttreeSubtree,TTree,ttreePoses2, 
+              TTPos,ttreePoses2WithTyps,ttreePoses2ByTyp, ttreePoses2WithTyps_onlyCompatible )
 
 import IM3 ( SearchOptions , prove , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
 --import IM2 ( SearchOptions , prove , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
+
+--import IM ( SearchOptions , prove , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
+
 import System.Random (StdGen)
 
 --import InhabitationMachines ( proveN , randProveN, randProveUnique, proveOneWithLimit, Limit, kozaProveN )
@@ -32,7 +36,7 @@ type StdDev = Double
 -- CTT -------------------------------------------------------------------------
 
 instance Gene CTT CTTGen where generateIt = cttGen
-instance Cros CTT CTTCro where crossIt    = cttCro2
+instance Cros CTT CTTCro where crossIt    = cttCro3
 
 data CTTGen  
 -- = CTTG_Koza     Typ Context Limit  -- pomocí starýho IM 
@@ -81,7 +85,9 @@ cttCro2 CTTC_Koza ctt1@(CTT v1 tree1) ctt2@(CTT v2 tree2) = do
   (cPos1,typ) <- crossPos1 tree1
   maybe_cPos2 <- crossPos2 tree2 typ
   case maybe_cPos2 of
-   Nothing -> return ( ctt1 , ctt2 )
+   Nothing -> do 
+    logIt "BUM!"
+    return ( ctt1 , ctt2 )
    Just cPos2 -> do
     let sub1          = ttreeSubtree tree1 cPos1
         (tree2',sub2) = ttreeChangeSubtree tree2 cPos2 sub1
[…]

### 2013-07-24 v4/Job.hs

@@ -1,12 +1,11 @@
-module Job ( problemList_ , job_ , job_new, problemList_new ) where
+module Job ( job, problemList ) where
 
-import GP_Core ( nRunsByServer )
-import Text.JSON
-import JSONUtils
-import ProblemUtils ( ProblemOpts , po2json, poCode, json2po , runProblemOpts, runProblemOpts_new )
+import Text.JSON (decode, Result(..), JSValue )
+import JSONUtils (fromJsStr,jsProp,jsArr)
+import ProblemUtils ( ProblemOpts , po2json, poCode, json2po , runProblemOpts )
 import Data.Typeable ( Typeable )
 import Data.Map (Map)
-import qualified Data.Map as Map
+import qualified Data.Map as Map (lookup, fromList)
 
 import ServerInterface ( OutputBuffer ) 
 
@@ -21,48 +20,6 @@ import qualified Problems.BA.Problem         as BA
 import qualified Problems.EvenParity.Problem as EvenParity
 
 
-regs_new = 
- [ reg_new Fly02.reg
- , reg_new Fly.reg 
- , reg_new BigCtx.reg_head
- , reg_new EvenParity.reg
- , reg_new BigCtx.reg
- , reg_new SSR.reg 
- , reg_new Ant.reg
- , reg_new (BA.reg :: POU) 
- ]
-
-
-
-job_new :: OutputBuffer -> String -> IO ()
-job_new buff cmd = case ( decode cmd :: Result JSValue ) of
-  Error str  -> do
-    putStrLn $ "ERROR in job_new : " ++ str
-    putStrLn $ cmd
-  Ok jsvalue -> do
-    let code = fromJsStr $ jsProp jsvalue "code" 
-        Just runFun = Map.lookup code problemTab_new 
-    putStrLn code
-    runFun jsvalue buff
-
-problemList_new :: JSValue
-problemList_new = jsArr $ snd jobs_new
-
-jobs_new :: ( [ (Code , JSValue -> OutputBuffer -> IO()) ] , [JSValue] )
-jobs_new = unzip regs_new
-
-reg_new :: Typeable a => ProblemOpts a -> ((String, JSValue -> OutputBuffer -> IO ()), JSValue)
-reg_new po = ( (poCode po , \ jsv buff -> runProblemOpts_new (json2po po jsv) buff ) , po2json po )
-
-
[…]

### 2013-07-24 v4/ProblemUtils.hs

@@ -7,13 +7,13 @@ module ProblemUtils (
   boolListProblem2,
   asType,
   casesFF,
-  ProblemOpts(..),PO_BLP(..),PO_CTTP(..),IntSlider(..),JobID,runProblemOpts,po2json,poCode,json2po,
-  runProblemOpts_new
+  ProblemOpts(..),PO_BLP(..),PO_CTTP(..),IntSlider(..),JobID,po2json,poCode,json2po,
+  runProblemOpts
 ) where
 
 import GP_Core (FitFun(..),Problem(..),NumGene,PopSize,mkGenOps)
 import GP_Core (GenOpProbs,FitVal,mkFF1)
-import GP_Core ( nRunsByServer, nRunsByServer_new )
+import GP_Core ( nRunsByServer )
 import GP_Data (CTTGen(..),CTTCro(..))
 import GP_Data (ListGen(..),ListCro(..),ListMut(..), BoolGen(..), BoolMut(..), Len)
 
@@ -21,8 +21,8 @@ import GP_Data (ListGen(..),ListCro(..),ListMut(..), BoolGen(..), BoolMut(..), L
 import TTree (CTT)
 import TTerm (Typ,Context)
 
-import Text.JSON
-import JSONUtils
+import Text.JSON (JSValue)
+import JSONUtils (jsObj,jsStr,jsProp,jsArr,jsNum,fromJsStr,fromJsInt)
 
 import Data.Typeable ( Typeable )
 
@@ -149,8 +149,10 @@ json2intSlider json = IntSlider {
 
 type JobID = String
 
-runProblemOpts :: (Typeable a) => ProblemOpts a -> JobID -> IO ()
-runProblemOpts pOpts jobID = case pOpts of
+
+
+runProblemOpts :: (Typeable a) => ProblemOpts a -> OutputBuffer -> IO ()
+runProblemOpts pOpts buff = case pOpts of
   PO_BLP_ blp ->
    let numRuns     = ( slider_value . blp_numRuns $ blp )
        numGene     = ( slider_value . blp_numGene $ blp )
@@ -162,7 +164,7 @@ runProblemOpts pOpts jobID = case pOpts of
        genOps      = mkGenOps (mOpt,cOpt) (33,33,33)
        fitFun      = mkFF1 $ return . (blp_ff blp)
        problemName = blp_code blp
-    in nRunsByServer jobID numRuns $ Problem problemName popSize numGene genOps gOpt mOpt cOpt fitFun
[…]

### 2013-07-24 v4/Server.hs

@@ -7,20 +7,17 @@ import Network.HTTP.Types (status200)
 import Network.Wai.Handler.Warp (run)
 
 import Control.Monad.State (liftIO)
-import Control.Concurrent (threadDelay,forkIO)
+import Control.Concurrent (forkIO)
 
 import Control.Monad.STM (atomically)
 import Control.Concurrent.STM.TVar (TVar,newTVar,modifyTVar,swapTVar,readTVarIO,readTVar,writeTVar)
 
 import Data.ByteString.Lazy.UTF8 (fromString)
-import Data.ByteString.Lazy as B (concat, ByteString, append)
 
 import Text.JSON ( encode )
 
 import Utils ( unescape )
-
-import Job (problemList_new,job_new)
-
+import Job (problemList,job)
 import ServerInterface( OutputBuffer , OutRecord(..) )
 
 
@@ -60,13 +57,13 @@ app tvarSS req = case pathInfo req of
  ["js","libs",filename]     -> return $ myJSFile $ "libs/"     ++ (myUnpack filename)
  ["js","Problems",filename] -> return $ myJSFile $ "Problems/" ++ (myUnpack filename)  
 
- ["problems"] -> return . myTextPlain . fromString . encode $ problemList_new
+ ["problems"] -> return . myTextPlain . fromString . encode $ problemList
 
  ["run",cmd] -> do 
    let cmdStr = myUnpack $ cmd 
    ss <- liftIO . readTVarIO $ tvarSS
    liftIO . forkIO $ do
-     job_new (ssBuff ss) (unescape cmdStr) 
+     job (ssBuff ss) (unescape cmdStr) 
      atomically $ modifyTVar (ssBuff ss) (\b->OutEnd:b)
    return $ myTextPlain "OK"
 
@@ -78,20 +75,6 @@ app tvarSS req = case pathInfo req of
  x -> do 
   liftIO $ putStrLn $ "404: " ++ show x
   return $ myTextPlain "404"
-
-myTextPlain x = responseLBS  status200 [ ("Content-Type", "text/plain") ] $ x
-
-serveFile      filename = ResponseFile status200 [ ("Content-Type", "text/html") ]       ("data/"++filename)               Nothing
-myFile         filename = ResponseFile status200 [  ]                                    ("data/files/" ++ filename )      Nothing
-myJSFile       filename = ResponseFile status200 [ ("Content-Type", "text/javascript") ] ("data/js/" ++ filename )         Nothing
[…]

### 2013-07-24 v4/ServerInterface.hs

@@ -1,10 +1,15 @@
-module ServerInterface where
+module ServerInterface 
+( OutputBuffer
+, OutRecord(..) 
+, writeNextOutput
+, jsEmptyObj
+, stdoutCmd 
+, graphCmd 
+, multiCmd ) where
 
-import System.IO
---import System.Directory
 
-import Text.JSON
-import JSONUtils
+import Text.JSON (JSValue)
+import JSONUtils (jsArr,jsObj,jsStr,jsNum)
 
 import Control.Monad.STM (atomically)
 import Control.Concurrent.STM.TVar (TVar,modifyTVar)
@@ -13,52 +18,9 @@ type OutputBuffer = TVar [OutRecord]
 data OutRecord = OutStr String | OutEnd
 
 
-writeNextOutput_new :: OutputBuffer -> String -> IO ()
-writeNextOutput_new buff output 
+writeNextOutput :: OutputBuffer -> String -> IO ()
+writeNextOutput buff output 
   = atomically $ modifyTVar buff (\b->(OutStr output):b)
-
-
-writeNextOutput :: Int -> String -> IO ()
-writeNextOutput i output = do
-  outI <- incrementFile ("server/output/" ++ (show i) ++ "/_hotovo.txt")
-  myWriteFile ("server/output/"++ (show i) ++ "/" ++ (show outI )++ ".txt" ) output
-
-
-incrementFile :: FilePath -> IO Int
-incrementFile filename = do 
-  file <- openFile filename ReadWriteMode
-  str <- hGetLine file 
-  let num  = 1 + ( read str ) :: Int
-      str' = show num
-  hSeek file AbsoluteSeek 0
-  hPutStr file str'
-  hClose file 
-  return num
-
-
-myWriteFile :: FilePath -> String -> IO ()
-myWriteFile filename str = do 
-  file <- openFile filename WriteMode
-  hPutStr file str
-  hClose file
-
-myReadFile :: FilePath -> IO String
-myReadFile filename = do
- file <- openFile filename ReadMode 
- str <- hGetLine file 
- hClose file
- return str
-
-
-problemListToJSON :: [(String,String,JSValue)] -> JSValue
-problemListToJSON problemList =  jsArr . map toProblemObj $ problemList
- where
-  toProblemObj :: (String,String,JSValue) -> JSValue
[…]

### 2013-07-24 v4/TTree.hs

@@ -4,7 +4,7 @@ module TTree
 ( CTT (..),
   TTree (..),
   TTPos,
-  mkCTT, -- <============== možná bug v něm!!!! 
+  --mkCTT, -- <============== možná bug v něm!!!! 
   mkCTT2,
   ttreeSubtree,
   ttreeSubtrees,
@@ -13,7 +13,8 @@ module TTree
   ttreePoses2,
 
   ttreePoses2WithTyps,
-  ttreePoses2ByTyp
+  ttreePoses2ByTyp,
+  ttreePoses2WithTyps_onlyCompatible
 ) where
 
 import Data.List
@@ -27,11 +28,20 @@ import JSONUtils
 
 import TTerm
 
+
+import qualified Data.Set as Set
+--import qualified Data.Map as Map
+import Data.Set (Set)
+--import Data.Map (Map)
+
+
 data CTT = CTT Context TTree 
 
 data TTree = TTree Symbol Typ [TTree] 
 
 type TTPos = [Int]
+
+
 
 
 mkCTT2 :: Context -> TTerm -> CTT
@@ -142,6 +152,23 @@ ttreePoses2WithTyps t =
   poses2xx pos (TTree _ typ []) = [Left (pos,typ)]
   poses2xx pos (TTree _ typ ts) = 
    (Right (pos,typ)) : (concatMap (\(i,t)-> poses2xx (i:pos) t ) (zip [1..] ts) )
+
+ 
+
+ttreePoses2WithTyps_onlyCompatible :: TTree -> TTree -> ([(TTPos,Typ)],[(TTPos,Typ)])
+ttreePoses2WithTyps_onlyCompatible prvni druhej =
+ let (ters,nonters) = ttreePoses2WithTyps prvni
+     setTypuVDruhym = ttreeTypsSet druhej
+     vyhodCoNemajTypVDruhym poziceSTypy 
+       = filter ( \(_,typ) -> Set.member typ setTypuVDruhym ) poziceSTypy
+  in ( vyhodCoNemajTypVDruhym ters , vyhodCoNemajTypVDruhym nonters  )
+
+
+ttreeTypsSet :: TTree -> Set Typ
+ttreeTypsSet (TTree _ typ ts) = Set.unions $ (Set.singleton typ) : (map ttreeTypsSet ts)    
+
+
+
 
 ttreePoses2ByTyp :: Typ -> TTree -> ([TTPos],[TTPos])
 ttreePoses2ByTyp typ t = 

### 2013-07-25 tex/thesis.tex
saveBest settable, etaRedukce, pokusy

@@
  & my tu řikáme že to závadíme kuli typum, ale my děláme v víc! -dáváme tam
-   i abstrakce atd - ale můžeme říct, že tim že se to pak převede zase na 
+   i abstrakce atd - ale mužeme říct, že tim že se to pak převede zase na 
    to bez abstrakcí tak setrváváme dal v krasnym bez-lambda abstrakčí ale
@@
 \section{[šlechtění search strategií pro prohledávání]}
+\section{[šlechtění alternative family trees]}
+
+\section{vypisování jedinců tak že nahradíme kombinátory \lterm{}ama a redukujem }
 

### 2013-07-25 v4/GP_Core.hs

@@ -50,6 +50,7 @@ data Problem term a gOpt mOpt cOpt = Problem
  , mOpt        :: mOpt
  , cOpt        :: cOpt
  , fitFun      :: FitFun term a
+ , saveBest    :: Bool
  }
 
 
@@ -102,9 +103,9 @@ evolveBegin runInfo p = do
 evolveStep ::(Muta t mo,Cros t co,Typeable a,JShow t)=> RunInfo->Problem t a go mo co -> Int -> Dist t -> Eva(Dist t,Maybe(t,FitVal))
 evolveStep runInfo p i pop = do
  let best     =  getBest pop 
- terms        <- distTake_new (popSize p - 1) pop   
+ terms        <- distTake_new (popSize p - (if saveBest p then 1 else 0) ) pop   
  terms'       <- performOps (genOps p) terms 
- ret@(pop',mWin) <- evalFF (fitFun p) ( best : terms' )
+ ret@(pop',mWin) <- evalFF (fitFun p) (if saveBest p then ( best : terms' ) else terms' )
  logGeneration runInfo i (isJust mWin) pop'
  return ret
 

### 2013-07-25 v4/IM3.hs

@@ -17,7 +17,7 @@ module IM3
 
 import Debug.Trace
 
-import TTerm (Symbol,Typ(..),TTerm(..),Context,ttermTyp,toSki,toSki',checkTyp)
+import TTerm (Symbol,Typ(..),TTerm(..),Context,ttermTyp,toSki,toSki',checkTyp,fullEtaReduce)
 
 import TTree (CTT,mkCTT2)
 
@@ -50,7 +50,7 @@ prove :: SearchOptions -> [CTT]
 prove so = 
   let ( so' , problemHead ) = problemHeadPreproccess so
       trees = proveWith2 so'
-   in map ( mkCTT2 problemHead . toSki . tree2tterm ) trees   -- <============== toSki s ' je s typeCheckem ...........
+   in map ( mkCTT2 problemHead . toSki . fullEtaReduce . tree2tterm ) trees   -- <============== toSki s ' je s typeCheckem ...........
 
 problemHeadPreproccess :: SearchOptions -> ( SearchOptions , Context )
 problemHeadPreproccess so = 

### 2013-07-25 v4/Job.hs

@@ -21,11 +21,11 @@ import qualified Problems.EvenParity.Problem as EvenParity
 
 
 regs = 
- [ reg Fly02.reg
+ [ reg BigCtx.reg
+ , reg Fly02.reg
  , reg Fly.reg 
  , reg BigCtx.reg_head
  , reg EvenParity.reg
- , reg BigCtx.reg
  , reg SSR.reg 
  , reg Ant.reg
  , reg (BA.reg :: POU) 

### 2013-07-25 v4/ProblemUtils.hs

@@ -62,7 +62,8 @@ data PO_BLP = PO_BLP {
   blp_code        :: String           ,
   blp_info        :: String           ,
   blp_data        :: JSValue          ,
-  blp_ff          :: [Bool] -> Double }
+  blp_ff          :: [Bool] -> Double ,
+  blp_saveBest    :: Bool             }
 
 
 blp2json :: PO_BLP -> JSValue
@@ -97,7 +98,8 @@ data PO_CTTP a = PO_CTTP {
   cttp_ff          :: FitFun CTT a       ,
   cttp_typ         :: Typ                ,
   cttp_ctx         :: Context            ,
-  cttp_gOpt        :: CTTGen         
+  cttp_gOpt        :: CTTGen             ,
+  cttp_saveBest    :: Bool 
  }
 
 cttp2json :: PO_CTTP a -> JSValue
@@ -164,7 +166,8 @@ runProblemOpts pOpts buff = case pOpts of
        genOps      = mkGenOps (mOpt,cOpt) (33,33,33)
        fitFun      = mkFF1 $ return . (blp_ff blp)
        problemName = blp_code blp
-    in nRunsByServer buff numRuns $ Problem problemName popSize numGene genOps gOpt mOpt cOpt fitFun
+       saveBest    = blp_saveBest blp
+    in nRunsByServer buff numRuns $ Problem problemName popSize numGene genOps gOpt mOpt cOpt fitFun saveBest
   PO_CTTP_ cttp ->
    let numRuns     = ( slider_value . cttp_numRuns $ cttp )
        numGene     = ( slider_value . cttp_numGene $ cttp )
@@ -180,7 +183,9 @@ runProblemOpts pOpts buff = case pOpts of
        genOpProbs  = (10,0,90)
        genOps      = mkGenOps (mOpt,cOpt) genOpProbs
        problemName = cttp_code cttp
-    in nRunsByServer buff numRuns $ Problem problemName popSize numGene genOps gOpt mOpt cOpt fitFun
+       saveBest    = cttp_saveBest cttp
+
+    in nRunsByServer buff numRuns $ Problem problemName popSize numGene genOps gOpt mOpt cOpt fitFun saveBest
 
 
 -- ---------------------------------------------------------------
@@ -215,7 +220,7 @@ cttProblem2 problemName ff typ ctx numGene popSize =
      gOpt       = CTTG_Koza2 typ ctx 
[…]

### 2013-07-25 v4/Problems/Ant/Problem.hs

@@ -23,7 +23,9 @@ reg = PO_CTTP_ PO_CTTP {
   
   cttp_gOpt        = CTTG_Koza2 ant ctx                           , 
 
-  cttp_ff          = FF5 "ff" "Problems.Ant.Funs" ()
+  cttp_ff          = FF5 "ff" "Problems.Ant.Funs" (), 
+  
+  cttp_saveBest    = True
   
 }
 

### 2013-07-25 v4/Problems/BA/Problem.hs

@@ -14,7 +14,8 @@ reg = PO_BLP_ PO_BLP {
   blp_numGene     = IntSlider "Generations"     0 100   10   10  ,
   blp_popSize     = IntSlider "Population size" 0 1000  500 100  , -- dřív 5000 ale začlo to s předěláním Evy dělat stackoverflow :(
   blp_length      = IntSlider "Genom length"    0 1000  200  10  ,
-  blp_ff          = ff
+  blp_ff          = ff                                           ,
+  blp_saveBest    = True
 }
 
 

### 2013-07-25 v4/Problems/BigCtx/Funs.hs

@@ -8,6 +8,9 @@ mkAll h t m f e = (h,t,m,f,e)
 mkFst3 :: h -> t -> m -> (h,t,m)
 mkFst3 h t m = (h,t,m)
 
+mkFst2 :: h -> t -> (h,t)
+mkFst2 h t = (h,t)
+
 
 s :: (a->b->c) -> (a->b) -> a -> c
 s x y z = x z (y z)

### 2013-07-25 v4/Problems/BigCtx/Problem.hs

@@ -13,8 +13,8 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_code        = "bigctx"                                     ,
   cttp_info        = "Problem with basic funs for types generating elementary funs." ,
   cttp_data        = jsObj [ ( "im" , imGraphInJSON prog_typ prog_ctx ) ]      ,
-  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
-  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    10   1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   50   10  ,
   cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
   
   cttp_typ         = prog_typ                                     ,
@@ -22,13 +22,15 @@ reg = PO_CTTP_ PO_CTTP {
   
   cttp_gOpt        = CTTG_Koza2 prog_typ prog_ctx                 , 
 
-  cttp_ff          = FF6 prog_type ff_fst3 "Problems.BigCtx.Funs" 
+  cttp_ff          = FF6 prog_type ff_fst2 "Problems.BigCtx.Funs" , 
+  
+  cttp_saveBest    = True
   
 }
 
-prog_type = asType :: Fst3_Type  -- asType :: [Int] -> Maybe Int
-prog_typ  = fst3_typ             -- l_int :-> m_int  
-prog_ctx  = ctx_fst3             -- ctx_combo
+prog_type = asType :: Fst2_Type   -- asType :: [Int] -> Maybe [Int]  -- asType :: Fst3_Type  -- asType :: [Int] -> Maybe Int
+prog_typ  = fst2_typ              -- l_int :-> ml_int                -- fst3_typ             -- l_int :-> m_int  
+prog_ctx  = ctx_fst2              -- ctx_combo                       -- ctx_fst3             -- ctx_combo
 
 
 reg_head =
@@ -48,7 +50,9 @@ reg_head =
       
       cttp_gOpt        = CTTG_Koza2 head_typ head_ctx                 , 
     
-      cttp_ff          = FF6 head_type ff_head "Problems.BigCtx.Funs"  
[…]

### 2013-07-25 v4/Problems/EvenParity/Problem.hs

@@ -19,7 +19,7 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_code        = "ep"                                     ,
   cttp_info        = "Even-parity problem." ,
   cttp_data        = jsObj [ ( "im" , imGraphInJSON prog_typ prog_ctx ) ] ,
-  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
+  cttp_numRuns     = IntSlider "Runs"            1 100   50   1   ,
   cttp_numGene     = IntSlider "Generations"     0 500   50   10  ,
   cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
   
@@ -30,9 +30,11 @@ reg = PO_CTTP_ PO_CTTP {
 --cttp_gOpt        = CTTG_Koza2    prog_typ prog_ctx  , 
 --cttp_gOpt        = CTTG_AllEdges prog_typ prog_ctx  , 
 
-  cttp_ff          = FF6 prog_type ff "Problems.EvenParity.Funs" 
+  cttp_ff          = FF6 prog_type ff "Problems.EvenParity.Funs", 
   
+  cttp_saveBest    = True
 }
+
 
 
 

### 2013-07-25 v4/Problems/Fly/Problem.hs

@@ -35,7 +35,9 @@ reg = PO_CTTP_ PO_CTTP {
 
   cttp_gOpt        = CTTG_Koza2 prog_typ ctx                      , 
   
-  cttp_ff          = FF6 prog_type ff "Problems.Fly.Funs" 
+  cttp_ff          = FF6 prog_type ff "Problems.Fly.Funs" , 
+  
+  cttp_saveBest    = True
   
 }
 

### 2013-07-25 v4/Problems/Fly02/Problem.hs

@@ -32,7 +32,9 @@ reg = PO_CTTP_ PO_CTTP {
 
   cttp_gOpt        = CTTG_Geom     prog_typ ctx 0.75 ,  --CTTG_Koza2 prog_typ ctx    , 
   
-  cttp_ff          = FF6 prog_type ff "Problems.Fly02.Funs" 
+  cttp_ff          = FF6 prog_type ff "Problems.Fly02.Funs" , 
+  
+  cttp_saveBest    = True
   
 }
 

### 2013-07-25 v4/Problems/SSR/Problem.hs

@@ -26,7 +26,9 @@ reg = PO_CTTP_ PO_CTTP {
 
   cttp_gOpt        = CTTG_Koza2 dou1 ctx                          , 
   
-  cttp_ff          = FF6 (asType::Double->Double) ff "Problems.SSR.Funs" 
+  cttp_ff          = FF6 (asType::Double->Double) ff "Problems.SSR.Funs", 
+  
+  cttp_saveBest    = False -- True 
   
 }
 

### 2013-07-25 v4/TTerm.hs

@@ -12,6 +12,8 @@ module TTerm
 , makeVarsUnique 
 , checkTyp
 , toSki, toSki'
+
+,fullEtaReduce
 ) where
 
 import Data.List ( nub, (\\) , intercalate )
@@ -43,6 +45,43 @@ data TTerm =
  TLam Symbol TTerm  Typ |
  TApp TTerm  TTerm  Typ
  deriving (Ord,Eq)
+
+-----------------------------------------------------------
+-- eta-reduction ------------------------------------------
+
+fullEtaReduce :: TTerm -> TTerm
+fullEtaReduce t = case oneEtaReduce t of
+  Nothing -> t
+  Just t' -> fullEtaReduce t'
+
+oneEtaReduce :: TTerm -> Maybe TTerm
+oneEtaReduce t = case t of
+ TLam x ( TApp m (TVar x' xPrimeTyp ) typApp ) typLam 
+  | x == x' && (not (elem x (etaFV m))) -> Just m  
+  | otherwise -> do
+      m' <- oneEtaReduce m
+      return $ TLam x ( TApp m' (TVar x' xPrimeTyp ) typApp ) typLam
+ 
+ TVar _ _     -> Nothing 
+ TVal _ _     -> Nothing 
+ TLam x m typ -> do
+  m' <- oneEtaReduce m
+  return $ TLam x m' typ
+ TApp m n typ -> case oneEtaReduce m of
+  Nothing -> do
+    n' <- oneEtaReduce n
+    return $ TApp m n' typ
+  Just m' -> Just $ TApp m' n typ
+ 
+
+
+etaFV :: TTerm -> [Symbol]
+etaFV (TVar v   _) = [v]
+etaFV (TVal v   _) = [] 
+etaFV (TApp p q _) = nub $ (fv p) ++ (fv q) -- neefektivni
+etaFV (TLam v p _) = (fv p) \\ [v]
+
+
 
 -- substitution ------------------------------------
 

### 2013-07-28 tex/thesis.tex
tex

@@
 \newcommand{\sexprTree}{sexpr-tree\xspace}
+\newcommand{\SexprTree}{Sexpr-tree\xspace}
 \newcommand{\atTree}{@-tree\xspace}
@@
 \subsection{Benefits of generating \lterms in \lnf}
+\label{benefits}
 
@@
 
-It is parameterized by parametr $q \in (0,1)$.
+It is parameterized by parametr $q \in (0,1)$, its default well-performing 
+value is $q = 0.75$. \\
 
@@
 \chapter{Problems}
-	In this section will be presented usage of the system in order to solve specific problems.
-
-		\section{Even Parity Problem}
+	In this section will be presented usage of the system in order to solve specific problems. Several problems will be described and some of them compared with results
+of others.
+
+\section{Even-parity problem}
+The even-parity function is a boolean function taking as inputs $N$
+boolean values and returning \textit{True} if an even number of inputs 
+are \textit{True}. For odd number it returns \textit{False}.
+
+The reason why examine this problem is that it has been used by many researchers
+as a benchmark for GP \cite{todo,yu01}.
+
+We compare our results with that obtained by Yu in \cite{yu01},
+where is presented approach to evolve recursive and modular programs
+by use of higher-order functions and $\lambda$-abstractions.
+Since the results are highly sensitive to set of \textit{building blocks}, 
+we use very similar set of \textit{building blocks} in order to make 
+the results comparable.
+
+\begin{align*}
+\sigma = [Boo&l] \ar Bool\\
+\Gamma = \{
+  and   &: Bool \ar Bool \ar Bool                              ,\\
+  or    &: Bool \ar Bool \ar Bool                              ,\\
+  nand  &: Bool \ar Bool \ar Bool                              ,\\
+  nor   &: Bool \ar Bool \ar Bool                              ,\\
+  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
+  head' &: [Bool] \ar Bool                                   ,\\
+  tail' &: [Bool] \ar [Bool]                              \}\\
+\end{align*}
+
+The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
+this problem is considered atomic.
+
+Unlike in \cite{yu01}, we use specific instance of polymorphic 
+function \texttt{foldr}. 
+And modifications of functions \texttt{head} 
+(returning the first element of the list) 
+and \texttt{tail} (returning the list without the first element) are used; 
+making them total by returning default value \textit{False}
+and \texttt{[]}, respectively.
+
+The \texttt{fold} function is higher order function widely used in \textit{Haskell}
+programming language, recursively defined as follows.
+
+\texttt{~\\
+foldr :: (a -> b -> b) -> b -> [a] -> b \\
+foldr f z []     = z \\
+foldr f z (x:xs) = f x (foldr f z xs)  
+}\\
+
+\Lets demonstrate \texttt{foldr}'s behavior on the following simple example.
+
+\texttt{~\\
+foldr xor False [True,False,True] \\
+= xor True (foldr False [False,True]) \\
+= xor True ( xor False ( foldr False [True] ) )\\
+= xor True ( xor False ( xor True (foldr False [] ) )\\
+= xor True ( xor False ( xor True False) )\\
+= False
+}\\
+
+The reasons why is this difficult problem 
+for GP to solve are following \cite{todo}:
+
+\begin{itemize}
+ \item Since a single change of one input generates a different output,
+       the output is highly sensitive to all inputs.
+ \item The building blocks set does not contain functions \textit{xor}
+       or \textit{eq}, which are keys to solving the problem.
+       But those two functions may be discover by GP.  
+\end{itemize}
+
+We use the same fitness function as in \cite{yu01}. 
+The fitness function examines the individual by giving
+it all possible boolean lists of length 2 and 3.
+Thus resulting in $2^2 + 2^3 = 12$ test cases.
+For each correct response it receives $1/12$, 
+therefore correct solution receives fitness value $1$.\\
+
+\begin{todo}
+ & zmínit tu, že to při týhle building block stačí na general solution...
+\end{todo}
+
+\subsection{Experiments}
+
+Series of five experiments have been performed to test properties of our system.
+Each experiment consisted of 50 independent runs of GP algorithm.
+Each run had maximally 50 generations
+(50 + 1 for generation 0) and 500 individuals as population size.\\
+
+In those experiments we examined influences of the following factors:
+
+\begin{itemize}
+ \item Option of preserving the best individual into the next population.
+ \item Option of performing $\eta$-normalization of generated individuals 
+      (conversion from \lnf to \benf, see \ref{benefits}).
+ \item Representation of individuals: \sexprTree vs \atTree.
+ \item Option of performing optimized abstraction elimination
+       (see \ref{toSKI}) instead of the simple one.
+\end{itemize}
+
+We analyze the ability of the system to produce a correct solution.
+We are interested in percentage of runs that led to discovery of
+correct solution. And we also use the popular measurement  
+method within GP field --- the \textit{performance curves}
+described in \cite{koza92}.
+
+First performance curve is the cumulative \textit{probability of success}
+$P(M,i)$, where $i$ is generation number ($i = 0$ for initial population) 
+and $M$ is the size of the population. 
+$P(M,i)$ is probability that a correct solution will
+be find in one run of GP with maximal number of generations $i$ and 
+population size $M$. $P(M,i)$ is computed as percentage of 
+runs yielding a correct solution in generation $i$ or earlier.
+
+From values of $P(M,i)$ are computed values of the second performance curve
+$I(M,i,z)$
+--- the total number of individuals that must be processed to yield a correct
+solution with probability $z$ by multiple runs of GP with maximal 
+number of generations $i$ and population size $M$ (usually $z = 0.99$).
+
+As is described in \cite{koza92},
+The number $R(M,i,z)$ of independent runs required to yield a correct solution 
+is computed as
+
+$$ R(M,i,z) = \lceil\frac{\log(1-z)}{\log(1-P(M,i))}\rceil ,$$
+
+thus $I(M,i,z)$ is computed as
+
+$$I(M,i,z) = (i+1) \cdot M \cdot R(M,i,z) .$$\\
+
+The setup of the first experiment was following:\\
+
+\begin{enum}
+ & \textit{Geometric} search strategy for generating individuals with default $q=0.75$.
+   This option is unchanged for all five experiments.
+ & No best individual preservation.
+ & $\eta$-normalization was not performed.
+ & Individuals represented as \sexprTree{}s. 
+ & Simple abstraction elimination.
+\end{enum}~
+
+Thus the simplest one setup; all extra features are off. 
+In each following experiment one extra feature will be turned on,
+in hope that it will enhance the performance.
+
+The correct individual was yield 
+in 17 out of 50 runs, i.e., 34\% success rate.
+The performance curve for $P(M,i)$ is following.
+
+\includegraphics[scale=0.65]{reports/ep/report1/probabs.png}
+
+Interestingly, in 2 runs out of 50, a correct solution was present in the initial population.
+
+The performance curve for $I(M,i,z)$ is following.
+
+\includegraphics[scale=0.65]{reports/ep/report1/indivs.png}
+
+The smallest value in this graph is used to indicate the minimum effort
+(i.e. number of individuals to be processed) for GP to yield a correct solution.
+
+Thus for this experiment this analysis suggest
+to perform only term generating and no evolution at all, 
+because the smallest value 56500 is in generation 0.
+
+This experiment took 59 minutes (on average desktop computer).\\
+
+Second experiment was different from the first one in that it 
+preserved the best individual to the next generation.
+
+It scored 22/50, i.e., 40\% success rate. A correct solution was
+created once in the initial generation and again it was
+sufficient to make the value in generation 0
+be the smallest value in the $I(M,i,z)$ curve; now 114000
+ individuals to be processed.
+This experiment also took 59 minutes.\\
+
+Third experiment was different from the second one in that 
+$\eta$-normalization of generated individuals was performed.
+
+It scored 27/50, i.e., 54\% success rate. 
+Two runs contained a correct solution in the initial population.
+But now the $I(M,i,z)$ was best for generation 1 with 56000
+ individuals to be processed.
+This experiment took 53 minutes.\\
+
+Fourth experiment was different from the third one in that
+\atTree representation was used instead of \sexprTree.
+
+It scored 34/50, i.e., 68\% success rate. Again two runs
+contained a correct solution in the initial population.
+The $I(M,i,z)$ was best for generation 0 with 56500 
+individuals to be processed.
+This experiment took 64 minutes.
+
+As we will see, this setup was the most successful one,
+thus we again show the performance curves.
+
+\includegraphics[scale=0.65]{reports/ep/report4/probabs.png}
+
+\includegraphics[scale=0.65]{reports/ep/report4/indivs.png}
+
+Fifth, last experiment was different from the fourth one in that
+optimized abstraction elimination was used instead of the simple one.
+
+It scored 33/50, i.e., 66\% success rate. One run
+contained a correct solution in the initial population.
+The $I(M,i,z)$ was best for generation 6 with 59500 
+individuals to be processed.
+This experiment took only 30 minutes. \\
+
+\subsection{Comparison with other results}
+
+Results presented by Yu \cite{yu01} are better then our results;
+they scored 40/50, i.e., 80\% success rate.  
+Her $I(M,i,z)$ was best for generation 4 with 17500 
+individuals to be processed.
+Yu also uses crossover as the only genetic operator, thus we
+can narrow the discussion to discussion of generating method and 
+crossover method.
+
+When we observe results of our experiments, it seems
+that our system is pretty good in term generation, but not so good 
+in crossover. Since the main focus point of this thesis is
+the term generation, it is positive rather then negative observation.
+
+Design of our crossover comes from that we wanted it to be
+very simple generalization of standard tree-swapping crossover.
+Whereas the crossover used by Yu is more structure oriented
+--- as is stated in \cite{yu01} the crossover \textit{can only operate 
+between nodes in two main programs or between nodes in the same
+kind of $\lambda$ abstraction (i.e., $\lambda$ abstractions that 
+represent the same function arguments to the same higher order function).}
+%This seems to be great advantage for optimizing the
+
+Since our method has better success rate for generation 0, it would 
+be interesting to try combination of our generating 
+method with crossover used by Yu.\\
+
+Positive observation is that our result at least outperforms 
+other results mentioned in \cite{yu01} with which her method is compared.
+
+\textit{Generic genetic programming} scored 17/60, i.e. 28\%; 
+minimum $I(M,i,z) = 220,000$.
+
+\textit{GP with ADFs} scored 10/29, i.e. 34\%; 
+minimum $I(M,i,z) = 1,440,000$.
+
+\begin{todo}
+& napsat tam jak vypadaj nějaký nalezeneý řešení
+\end{todo}
+
+\section{Flies and Apples}
+
+\textit{Flies and Apples} is problem consisting in breeding 
+control programs for a fly agent in a simple world.
+
+Fly world is a square grid. 
+Each cell is ether empty or contains precisely one object.
+There are three kinds of objects:
+
+~\begin{enum}
+ & Wall
+ & Apple
+ & Fly
+\end{enum}~
+
+Apples and flies have \textit{energy}.
+
+Moreover, each fly has its control program and inner state, 
+so called \textit{registers}.
+
+There is a queue of all flies in the fly world determining which 
+fly is next to make an action.
+After the fly in the front of the queue performs its action, 
+it is put to the back. In other words, flies are cyclically
+taking turns similarly like in common card games.\\
+
+Fly can perform two kinds of action:
+
+~\begin{enum}
+ & \textit{Trave.} Moves the fly to one of four adjacent cells determined by 
+   direction specified in the action.  
+   Result of the travel action depends on the content of the target cell.    
+   && If it is empty, then the fly simply moves to that cell.
+   && If it contains wall, then no movement is performed.
+   && If it contains an apple, then the fly moves to that cell, the apple 
+      is eaten and its energy is added to energy of the fly.  
+   && If it contains another fly, then the fly moves to that cell, the fly with bigger 
+      energy stays alive and the weaker one is eaten
+      and its energy is added to energy of the stronger fly.  
+ & \textit{Split.} Puts a daughter fly to one of four adjacent cells.
+    Together with the direction, amount of energy and new registers 
+    for the fly are specified by this action. The mother gives this energy to 
+    its child. In order to be successful, the target cell must be empty
+    and the mother fly must have energy greater then 1. 
+\end{enum}~
+
+\section{Simple Symbolic Regression}
+
+\section{Artificial Ant}
+
 \section{Big Context}
-		\section{Fly}
-		\section{Simple Symbolic Regression}
-		\section{Artificial Ant}
+
 \section{Boolean Alternate}

### 2013-07-29 tex/thesis.tex
pondělí

@@
    to enable use of simple \textit{mutation operator}.
- \item \textit{Polymorphic} generalization of inhabitation tree,
-   enabling use of polymorphic functions as building blocks, 
-   is presented.
+% \item \textit{Polymorphic} generalization of inhabitation tree,
+%   enabling use of polymorphic functions as building blocks, 
+%   is presented.
 
@@
 
-\chapter{Polymorphic variant}
-
-\begin{todo}
- & poznamky o čem se mimojiné zmínit:
-   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
-      jako alternativa k ADF ! Myslim že hodně zajimavý
-\end{todo}
-
 \chapter{ Implementation }
@@
 
-Results presented by Yu \cite{yu01} are better then our results;
+Our results are less successful then those 
+presented by Yu \cite{yu01};
 they scored 40/50, i.e., 80\% success rate.  
@@
 individuals to be processed.
+(We scored 34/50 (68\%); min $I(M,i,z) = 56,500$.)
+
 Yu also uses crossover as the only genetic operator, thus we
@@
 be interesting to try combination of our generating 
-method with crossover used by Yu.\\
+method with crossover used by Yu.
+
+From an optimist's point of view 34/50 versus 40/50 
+is not such a huge difference when we take to account that
+our general tree-swapping crossover is competing with specialized one.
 
 Positive observation is that our result at least outperforms 
-other results mentioned in \cite{yu01} with which her method is compared.
-
-\textit{Generic genetic programming} scored 17/60, i.e. 28\%; 
-minimum $I(M,i,z) = 220,000$.
-
-\textit{GP with ADFs} scored 10/29, i.e. 34\%; 
-minimum $I(M,i,z) = 1,440,000$.
+other results mentioned in \cite{yu01} with which Yu's method is compared.
+
+\textit{Generic genetic programming} scored 17/60 (28\%); 
+min $I(M,i,z) = 220,000$.
+
+\textit{GP with ADFs} scored 10/29 (34\%); 
+min $I(M,i,z) = 1,440,000$.
 
@@
 \textit{Flies and Apples} is problem consisting in breeding 
-control programs for a fly agent in a simple world.
+control programs for a fly agent in a simple world.\\
+
+Whereas previous problem is a kind of problem where we know precisely
+what problem we are solving and where we are more interested in 
+performance statistics than in the solutions since we know the correct solution in 
+advance, problem we are going to present now is of more playful nature.
+In this problem we are not interested in numerical results, instead 
+we are hoping for interesting \textit{behavior} to emerge.
+
+We start with defining a simple simulation and goal for agents.
+Then we prepare first testing simulation world. 
+After that we write some agent programs by hand. After that we 
+identify some useful recurring 
+themes occurring in those programs and pack them into functions
+which will become building blocks. During this process are also identified
+useful sensory data to be prepared for the agent program.
+After that we run the evolution and
+iteratively continue in this process  --- we may add more testing worlds,
+add more functions or change them, etc.
+
+Since we do not need to satisfy the \textit{closure} constraint, 
+it is much easier to perform such a "human-computer jam" than it
+would be in standard GP. And it is further simplified by use of
+functional language (such as \textit{Haskell}) where almost
+every construct can be expressed as function (or value).
+
+Another advantage of typed GP over standard GP is that 
+it is more capable of handling huge set of building blocks,
+since the search space can be significantly smaller due to 
+fact that types must match.
+\\
 
@@
 There is a queue of all flies in the fly world determining which 
-fly is next to make an action.
-After the fly in the front of the queue performs its action, 
+fly is next to make a move.
+After the fly in the front of the queue performs its move, 
 it is put to the back. In other words, flies are cyclically
@@
 
-Fly can perform two kinds of action:
+A control program of a fly has as input a collection of various sensory information
+and returns as output agents move and agents registers for the next turn.
+
+Fly can perform two kinds of moves:
 
@@
  & \textit{Trave.} Moves the fly to one of four adjacent cells determined by 
-   direction specified in the action.  
-   Result of the travel action depends on the content of the target cell.    
+   direction specified in the move.  
+   Result of the travel move depends on the content of the target cell.    
    && If it is empty, then the fly simply moves to that cell.
@@
     Together with the direction, amount of energy and new registers 
-    for the fly are specified by this action. The mother gives this energy to 
+    for the fly are specified by this move. The mother gives this energy to 
     its child. In order to be successful, the target cell must be empty
@@
 
+Agent's fitness is computed as sum of fitness cases; each 
+fitness case consists of one fly world and fitness in a
+fly world is computed as sum of agent's energy and of energy of all his descendants
+(daughters, granddaughter, ...) after predefined number of rounds.
+
+\Lets look in more detail on input for fly. It is collection of 
+various useful information. Its components are:
+
+~\begin{enum}
+ & Current value of fly's Registers. 
+ & Energy of the fly.
+ & Direction of the last successful travel move.
+ & Boolean value indicating whether the last move was successful.
+ & Direction of the nearest apple.
+ & Distance to the nearest apple.
+ & Energy of the nearest apple.
+ & Direction of the nearest fly.
+ & Distance to the nearest fly.
+ & Energy of the nearest fly.
+ & Direction of the "center of gravity" of all apples.
+ & Distance to the "center of gravity" of all apples.
+\end{enum} ~
+
+Set of building blocks is following.
+
+\begin{align*}
+\sigma = Input &\ar Output\\
+\Gamma = \{
+dUp              &: Direction                                       ,\\              
+dDown            &: Direction                                       ,\\              
+dLeft            &: Direction                                       ,\\              
+dRight           &: Direction                                       ,\\            
+output           &: Move \ar Registers \ar Output                   ,\\            
+travel           &: Direction \ar Move                              ,\\          
+split            &: Direction \ar Int \ar Registers \ar Move        ,\\       
+easySplit        &: Input \ar Move                                  ,\\     
+myEnergy         &: Input \ar Int                                   ,\\     
+myLastTravel     &: Input \ar Direction                             ,\\                
+myWasSuccess     &: Input \ar Bool                                  ,\\      
+nAppleDir        &: Input \ar Direction                             ,\\          
+nAppleDir        &: Input \ar Direction                             ,\\                    
+nAppleDist       &: Input \ar Distance                              ,\\                   
+nAppleEnergy     &: Input \ar Int                                   ,\\              
+nFlyDir          &: Input \ar Direction                             ,\\                      
+nFlyDist         &: Input \ar Distance                              ,\\                    
+nFlyEnergy       &: Input \ar Int                                   ,\\              
+cAppleDir        &: Input \ar Direction                             ,\\          
+cAppleDist       &: Input \ar Distance                              ,\\        
+myRegs           &: Input \ar Registers                             ,\\         
+\end{align*}
+\begin{align*}
+xGet             &: Input \ar Int                                   ,\\     
+yGet             &: Input \ar Int                                   ,\\     
+zGet             &: Input \ar Int                                   ,\\     
+dGet             &: Input \ar Direction                             ,\\          
+xSet             &: Int \ar Registers \ar Registers                 ,\\          
+ySet             &: Int \ar Registers \ar Registers                 ,\\          
+zSet             &: Int \ar Registers \ar Registers                 ,\\          
+dSet             &: Direction \ar Registers \ar Registers           ,\\                
+xInc             &: Registers \ar Registers                         ,\\              
+yInc             &: Registers \ar Registers                         ,\\              
+zInc             &: Registers \ar Registers                         ,\\              
+rotCW            &: Direction \ar Direction                         ,\\                                                             
+(==)             &: Int \ar Int \ar Bool                            ,\\                                      
+(<=)             &: Int \ar Int \ar Bool                            ,\\               
+if'              &: Bool \ar Output \ar Output \ar Output           ,\\                                  
+if'              &: Bool \ar Move \ar Move \ar Move                 ,\\                                  
+if'              &: Bool \ar Direction \ar Direction \ar Direction  ,\\              
+if'              &: Bool \ar Int \ar Int \ar Int                    ,\\     
+if'              &: Bool \ar Distance \ar Distance \ar Distance     ,\\        
+if'              &: Bool \ar Registers \ar Registers \ar Registers  ,\\                           
+0                &: Int                                             ,\\                                     
+1                &: Int                                             ,\\                                    
+2                &: Int                                             \}\\
+\end{align*}
+
+We will not discuss it in great detail, only briefly: First four 
+elements are four direction values. Next four functions are
+constructors for $Output$ and $Move$ type. Next thirteen functions
+are for accessing components of $Input$. Next eleven functions are 
+related to $Registers$. The $rotCW$ performs clock-wise rotation of a direction.
+Next two functions are comparison functions. Next six are various \textit{if}s. And last 
+three are integer constants.
+
+Tedy celkem 44 prvků množiny stavebních bloků, tedy poměrně obrovská množina
+na GP poměry.
+
+Takže množina atomických typů je následující:
+Move Bool Registers Input Direction Int Output Distance
+
+Používáme tři levely. První level vypadá takle
+
+[obrázek]
+
+Další dva levely jsou reakce na typické chování mouchy 
+která používala pro předchozí level triviální strategii
+jdi furt za nejbližšim jablkemn. roto jsou následující dva levely 
+léčka na tuto strategii.
+
+V druhém levelu se zmíněná moucha nachytá a vleze do 
+chodby kde je mnohem mín jablek než v v tom mega jablku z jakblek.
+
+[obrázek]
+
+Třetí dovádí předchozí myšlenku do extrému. 
+Moucha se zasekne a nesebere žádné jablko.
+
+V počátečních pokusech byl problém s tím, že mochy s počáteční energii 1
+neměli moc tendenci se množit. Proto byla simulace předělána tak, aby
+že nyní je počáteční energie 100. To mělo na Mouchy blahodárný vliv.
+Ale i pro mouchy sw počátečni energii 1 vznikli některé zajímavé strategie
+
+[povídání o zajímavých strategiích pro poč en 1]
+
+Po navíšení poč energie vznikli tyto zajímavá chování
+
+[povídání o zajímavých strat por poč en 100]\\
+
+Aplikace: například v oboru počítačových her, kde nám často jde o
+zajímavé nelineární chování by se tento přístup k využití typovaného funkcionálního
+programování dal využít.
+
+Předvedený postup by taky ohlo být možné použít na
+
 \section{Simple Symbolic Regression}
 
+\textit{Simple Symbolic Regression} is a problem described
+in \cite{koza92}. Objective of this problem is to 
+find a function $f(x)$ that fits a sample
+of twenty given points. The target function is 
+function $f_{t}(x) = x^4 + x^3 + x^2 + x$.
+
+Desired type of generated programs $\sigma$ and 
+building blocks context $\Gamma$ are following.
+\newcommand{\Real}{\mathbb{R}}
+\begin{align*}
+\sigma = \Real \ar &\Real\\
+\Gamma = \{
+  (+)  &: \Real \ar \Real \ar \Real    ,\\
+  (-)  &: \Real \ar \Real \ar \Real    ,\\
+  (*)  &: \Real \ar \Real \ar \Real    ,\\
+  rdiv &: \Real \ar \Real \ar \Real    ,\\
+  sin  &: \Real \ar \Real              ,\\
+  cos  &: \Real \ar \Real              ,\\
+  exp  &: \Real \ar \Real              ,\\ 
+  rlog &: \Real \ar \Real              \}\\
+\end{align*}
+
+Fitness function is computed as follows
+
+$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
+
+where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.\\
+
+An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
+considered as a correct individual.\\
+
+On this problem we demonstrate that if control parameters are set to be
+compatible with standard GP, then the results obtained fit the results
+presented in \cite{koza92} relatively well.
+
+Those control parameters are following two:
+
+~\begin{enum}
+ & \textit{Ramped half-and-half strategy} is used as search strategy.
+ & Option of preserving the best individual into the next population is set \textbf{off}.
+\end{enum}~
+
+The experiment consisted of 50 independent runs of GP algorithm.
+Each run had maximally 50 generations and 500 individuals as population size.\\
+
+It scored 17/50 (34\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 23 
+with 192,000 individuals to be processed.
+The experiment took 46 minutes.
+
+\Lets compare it with results from \cite{koza92} (based on 113 runs):
+
+Success rate 35\%; minimal $I(M,i,z)$ in generation 24 
+with 162,500 individuals to be processed.\\
+
+Performance curves are following.
+
+\includegraphics[scale=0.65]{reports/SSR/1/probabs.png}
+
+\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}
+
+Second experiment was performed to compare 
+\textit{ramped half-and-half strategy} with our
+\textit{geometric strategy} (with default parameter $q=0.75$). 
+Thus all control parameters stay same except for the 
+search strategy for generating algorithm.
+
+It scored 21/50 (42\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 29 
+with 150,000 individuals to be processed.
+The experiment took 26 minutes.\\
+
+[todo OBRAZKY GRAFů]
+
+So our geometric strategy is slightly more successful then 
+ramped half-and-half strategy. But more interesting is that
+it made the whole experiment almost twice as fast.
+
 \section{Artificial Ant}
 
-\section{Big Context}
-
-\section{Boolean Alternate}
-
-\chapter{Comparison with others}
+\textit{Artificial Ant} is another problem described
+in \cite{koza92}. Objective of this problem is to 
+find a control program for an artificial ant so
+that it can find all food located on "Santa Fe" trail.
+
+The Santa Fe trail has following layout.
+
+~\\~[ TODO OBRAZEK Santa Fe]\\
+
+The ant is able to move forward, turn left, and sense if a food 
+piece is ahead of him.\\
+
+Desired type of generated programs $\sigma$ and 
+building blocks context $\Gamma$ are following.
+\begin{align*}
+\sigma = An&tAction\\
+\Gamma = \{~~
+  l    &: AntAction                              ,\\
+  r    &: AntAction                              ,\\
+  m    &: AntAction                              ,\\
+  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
+  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
+  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
+\end{align*}
+
+Action $l$ turns the ant left.
+
+Action $r$ turns the ant right.
+
+Action $m$ moves the ant forward.
+
+Action $ifa x y$ performs action $x$ if a food piece is ahead of the ant,
+otherwise it performs action $y$.
+
+Action $p2 x y$ first performs action $x$ and after it action $y$.
+
+Action $p3 x y z$ first performs action $x$, after that action $y$ and finally $z$.\\
+
+Actions $l, r$ and $m$ each take one time step to execute.
+
+Tento problém je také převzatý z \cite{koza92}.
+
+Jde v něm o to [podivat ze do kozy a převypravet to]
+
+sigma a gama jsou
+
+fitnes je následující
+
+koza uvádí že jedinec [ten jedinec] je jedinec který mu vznikl
+při vyšlechtění. Na našem systému ale trvá mu to nakejch tech 500něco
+krků přičemž by to mělo trvat 400 max.
+
+Nevíme zda je chyba na mé straně (nepřesná implementace) nebo
+se jedná o nepřesnost v koza92. Díky tomuto rozporu však nepoužijeme
+tuto ulohu pro dokladování toho že se to chová jako 
+standard GP. My jsme tedy oproti knize použili max počet kroků 600
+tak aby popsaný jedinec v koza92 byl korektní i vnašem systému.
+
+Uvádíme pokus o 50 behach s (geom strategie default q, best preservation on)
+...
+
+\section{Results summary}
+
+\chapter{Previous related work}
+
+\section{Evolution of Constrained Syntactic Structures }
+by Koza
+
+\section{STGP}
+by Montana
+
+\section{Higher-Order Functions and Lambda Abstractions} 
+by Yu
+Požívá dvou levelovou hierarchii (najít tu pasaž)
+
+oproti nám umí i polymorfní typy
+
+K  lambda abstrakcím přistupuje jako k jedinému bodu který se dělí až 
+vevnitř.
+
+Používá na to speciálně přichystaný křížení
+
+Generování počáteční generace je v zásadě to samí co koza, když příde na vygenerování 
+abstrakce tak se to zase generuje klasicky s tim že proměný jsou \#1 \#2 atd
+
+klade velkej důraz na foldr jakožto na implicitní rekurzi
+
+v even parity scoruje o něco lépe než my.
+
+\section{Genetic Programming with Polymorphic Types and Higher-Order Functions}
+by Franck Binard and Amy Felty
 
@@
 \chapter{Future work}
+
+\section{Polymorphic variant}
+
+\begin{todo}
+ & poznamky o čem se mimojiné zmínit:
+   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
+      jako alternativa k ADF ! Myslim že hodně zajimavý
+\end{todo}
+
+\section{Big Context}
+
+TODO asi přenda do future work
+
+Tento problém je napůl rozpracovaný, hlavně protože by se na něj o dost více
+hodilo použít polymorfní typy, které jsou také ne uplně dodělané a véto thesis
+se o nich zminujeme jen velmi okrajově. Přestose myslím jedná o zajímavou myšlanku
+a proto ho popíšeme.
 

### 2013-07-29 v4/IM3.hs

@@ -59,7 +59,8 @@ prove :: SearchOptions -> [CTT]
 prove so = 
   let ( so' , problemHead ) = problemHeadPreproccess so
       trees = proveWith2 so'
-   in map ( mkCTT2 problemHead . optSki . fullEtaReduce . tree2tterm ) trees   -- <============== toSki s ' je s typeCheckem ...........
+   in map ( mkCTT2 problemHead . toSki . fullEtaReduce . tree2tterm ) trees   -- <============== toSki s ' je s typeCheckem .. nebo tam de optSki
+
 
 problemHeadPreproccess :: SearchOptions -> ( SearchOptions , Context )
 problemHeadPreproccess so = 

### 2013-07-29 v4/Problems/Ant/Problem.hs

@@ -21,11 +21,11 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_typ         = ant                                          ,
   cttp_ctx         = ctx                                          ,
   
-  cttp_gOpt        = CTTG_Koza2 ant ctx                           , 
+  cttp_gOpt        = CTTG_Koza2 ant ctx ,  --CTTG_Geom     ant ctx 0.75 ,  --CTTG_Koza2 ant ctx ,    
 
   cttp_ff          = FF5 "ff" "Problems.Ant.Funs" (), 
   
-  cttp_saveBest    = False
+  cttp_saveBest    = True --False
   
 }
 

### 2013-07-29 v4/Problems/SSR/Problem.hs

@@ -17,14 +17,14 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_code        = "ssr"                                        ,
   cttp_info        = "Simple Symbolic Regression - new IM !"      ,
   cttp_data        = jsData                                       ,
-  cttp_numRuns     = IntSlider "Runs"            1 10    1    1   ,
-  cttp_numGene     = IntSlider "Generations"     0 100   10   10  ,
+  cttp_numRuns     = IntSlider "Runs"            1 10    50   1   ,
+  cttp_numGene     = IntSlider "Generations"     0 100   50   10  ,
   cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
   
   cttp_typ         = dou1                                         ,
   cttp_ctx         = ctx                                          ,
 
-  cttp_gOpt        = CTTG_Koza2 dou1 ctx                          , 
+  cttp_gOpt        = CTTG_Geom  dou1 ctx 0.75  , --CTTG_Koza2 dou1 ctx , 
   
   cttp_ff          = FF6 (asType::Double->Double) ff "Problems.SSR.Funs", 
   

### 2013-07-29 tex/thesis.tex
uterý rano

@@
 with 192,000 individuals to be processed.
-The experiment took 46 minutes.
+The experiment took 46 minutes.\\
+
+Performance curves are following.
+
+\includegraphics[scale=0.65]{reports/SSR/1/probabs.png}
+
+\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}\\
 
@@
 with 162,500 individuals to be processed.\\
-
-Performance curves are following.
-
-\includegraphics[scale=0.65]{reports/SSR/1/probabs.png}
-
-\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}
 
@@
 
-[todo OBRAZKY GRAFů]
+\includegraphics[scale=0.65]{reports/SSR/2/probabs.png}
+
+\includegraphics[scale=0.65]{reports/SSR/2/indivs.png}\\
 
@@
 it made the whole experiment almost twice as fast.
+
+~\\TODO dát sem ty grafy průměrný term size
 
@@
 
-The Santa Fe trail has following layout.
+The Santa Fe trail lies on toroidal square grid and has following layout.
 
@@
 
-Action $ifa x y$ performs action $x$ if a food piece is ahead of the ant,
+Action $ifa x y$ (if-food-ahead) performs action $x$ if a food piece is ahead of the ant,
 otherwise it performs action $y$.
@@
 
-Tento problém je také převzatý z \cite{koza92}.
-
-Jde v něm o to [podivat ze do kozy a převypravet to]
-
-sigma a gama jsou
-
-fitnes je následující
-
-koza uvádí že jedinec [ten jedinec] je jedinec který mu vznikl
-při vyšlechtění. Na našem systému ale trvá mu to nakejch tech 500něco
-krků přičemž by to mělo trvat 400 max.
-
-Nevíme zda je chyba na mé straně (nepřesná implementace) nebo
-se jedná o nepřesnost v koza92. Díky tomuto rozporu však nepoužijeme
-tuto ulohu pro dokladování toho že se to chová jako 
-standard GP. My jsme tedy oproti knize použili max počet kroků 600
-tak aby popsaný jedinec v koza92 byl korektní i vnašem systému.
-
-Uvádíme pokus o 50 behach s (geom strategie default q, best preservation on)
-...
+Ants action is performed over and over again until it reaches predefined
+maximal number of steps. In \cite{koza92} this limit is set to be 400 time steps.
+
+Fitness value is equal to number of eaten food pieces.
+
+An individual such that eats all 89 pieces of food is 
+considered as a correct solution.\\
+
+There is also mentioned following ant program which is described as correct solution.
+
+\texttt{~\\
+(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))\\
+}
+
+But here arises a problem: In our implementation this program needs 545
+time steps; if it is given only 400 time steps, then it eats only 79 pieces
+of food. It is uncertain whether it is caused by my mistake in implementation
+or by inaccuracy in \cite{koza92}. Since everything else seems working just 
+fine, I have decided to set the limit to 600 in my experiments.
+
+Due to those uncertainties, we will not perform comparing experiments 
+as we did in the problem of simple symbolic regression.\\
+
+In the following experiments, the option of preserving the best individual 
+into the next generation is active (since we are not trying to compare 
+with results in \cite{koza92}).
+
+As usually, first experiment consisted of 50 independent runs of GP algorithm.
+Each run had maximally 50 generations and 500 individuals as population size.
+\textit{Ramped half-and-half strategy} is used as search strategy for the first 
+experiment.\\
+
+It scored 7/50 (14\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 28 
+with 449,500 individuals to be processed.
+The experiment took 265 minutes.
+
+Performance curves are following.
+
+Similarly as in previous problem, second experiment was performed to compare 
+\textit{ramped half-and-half strategy} with our
+\textit{geometric strategy} (with default parameter $q=0.75$). 
+Thus all control parameters stay same except for the 
+search strategy for generating algorithm.
+
+It scored 19/50 (38\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 10 
+with 115,500 individuals to be processed.
+The experiment took 107 minutes.\\
+
+\includegraphics[scale=0.65]{reports/Ant/1/probabs.png}
+
+\includegraphics[scale=0.65]{reports/Ant/1/indivs.png}\\
+
+This is big improvement in all three watched factors:
+
+\begin{enumerate}
+ \item 38\% vs 14\% success rate.
+ \item 115,500 vs 449,500 individuals needed to be processed.
+ \item 107 vs 265 minutes.  
+\end{enumerate}
+
+Third factor, the time difference might give us clue about what is going on.
+
+Those two experiments differs only in used term generating method.
+During one run the vast majority of the time is spend by evolution,
+the generation process only slightly prolongs the time spend in generation 0.
+The run time of the evolution is influenced by two important factors:
+Number of generations and time that crossover takes.
+
+The number of generations of more successful method is surly smaller,
+but it seems improbable that it would have such a big impact 
+(this is idea is supported by results of experiments with
+even-parity problem, where times are almost the same but success 
+rates differs significantly).
+
+Time of crossover is proportionate to number of nodes that 
+an individual has. Thus it seems that observed slowdown is
+due to terms being too big. \\
+
+In order to test this hypothesis, we run additional experiments
+to get statistics about dynamics of individuals' sizes.
+
+~\\TODO dát sem ty grafy průměrný term size
 

### 2013-07-29 v4/GP_Core.hs

@@ -24,7 +24,8 @@ import Data.Maybe    ( fromJust,isJust )
 
 import Eva   ( Eva, runEva, evalsWith, evals, setOutputBuffer, sendJSON, flushStdout )
 import Dist  ( Dist , mkDist, distTake_new, distGet, distMax, distMin, distAvg )
-import Utils ( logIt, boxIt , JShow, jshow, putList)
+import Dist  ( distToList )
+import Utils ( logIt, boxIt , JShow, jshow, putList, jss_size)
 import ServerInterface ( graphCmd, multiCmd, OutputBuffer ) 
 
 
@@ -207,9 +208,16 @@ logGeneration (actRun,allRuns) i isWinner pop = do
  logIt  $ " │ Worst   " ++ p2 w ++ " │"
  logIt  $ " └────────────────────────┘" 
  if isWinner then logIt "WINNER!!!!!!!" else return ()
+ let m_averageTermSize = if True 
+                          then let popList = map fst $ distToList 32 pop
+                                   sizeOf  = fromIntegral . fromJust . jss_size
+                                   len     = fromIntegral $ length popList
+                                   isOK    = isJust . jss_size . head $ popList
+                                in if isOK then Just (  ( sum $ map sizeOf popList  ) / len ) else Nothing 
+                          else Nothing 
  boxIt  $ show best 
  stdout <- flushStdout 
- sendJSON $ multiCmd [ graphCmd actRun i (b,a,w) isWinner , stdout , jshow best ]
+ sendJSON $ multiCmd [ graphCmd actRun i (b,a,w) isWinner m_averageTermSize , stdout , jshow best ]
 
 
 

### 2013-07-29 v4/Job.hs

@@ -22,16 +22,16 @@ import qualified Problems.EvenParity.Problem as EvenParity
 
 
 regs = 
- [ reg Ant2.reg
- , reg BigCtx.reg
- , reg Fly02.reg
- , reg Fly.reg 
- , reg BigCtx.reg_head
+ [ reg SSR.reg 
+ , reg Ant.reg
  , reg EvenParity.reg
  , reg EvenParity.rege
- , reg SSR.reg 
- , reg Ant.reg
+ , reg Fly02.reg
+ , reg BigCtx.reg
  , reg (BA.reg :: POU) 
+ --, reg BigCtx.reg_head
+ -- , reg Fly.reg 
+ -- ,  reg Ant2.reg
  ]
 
 type Code  = String

### 2013-07-29 v4/Problems/Ant/Ant.hs

@@ -237,3 +237,37 @@ antWorld = mkWorld 600
   , "..FFFF.........................." 
   , "................................" ]
 
+
+antWorld2 = mkWorld 545 
+  [ ".FFF............................"
+  , "...F............................"
+  , "...F.....................FFF...." 
+  , "...F....................F....F.." 
+  , "...F....................F....F.." 
+  , "...FFFF.FFFFF........FF........." 
+  , "............F................F.." 
+  , "............F.......F..........." 
+  , "............F.......F..........." 
+  , "............F.......F........F.." 
+  , "....................F..........." 
+  , "............F..................." 
+  , "............F................F.." 
+  , "............F.......F..........." 
+  , "............F.......F.....FFF..." 
+  , ".................F.....F........" 
+  , "................................" 
+  , "............F..................."             
+  , "............F...F.......F......."
+  , "............F...F..........F...." 
+  , "............F...F..............."             
+  , "............F...F..............."
+  , "............F.............F....." 
+  , "............F..........F........" 
+  , "...FF..FFFFF....F..............." 
+  , ".F..............F..............." 
+  , ".F..............F..............." 
+  , ".F......FFFFFFF................." 
+  , ".F.....F........................" 
+  , ".......F........................" 
+  , "..FFFF.........................." 
+  , "................................" ]

### 2013-07-29 v4/Problems/Ant/Problem.hs

@@ -14,14 +14,15 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_code        = "ant"                                        ,
   cttp_info        = "Artifical Ant"                              ,
   cttp_data        = jsData                                       ,
-  cttp_numRuns     = IntSlider "Runs"            1 10    50   1   ,
+  cttp_numRuns     = IntSlider "Runs"            1 100   50   1   ,
   cttp_numGene     = IntSlider "Generations"     0 100   50   10  ,
   cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
   
   cttp_typ         = ant                                          ,
   cttp_ctx         = ctx                                          ,
   
-  cttp_gOpt        = CTTG_Koza2 ant ctx ,  --CTTG_Geom     ant ctx 0.75 ,  --CTTG_Koza2 ant ctx ,    
+--cttp_gOpt        = CTTG_Geom     ant ctx 0.75 ,  
+  cttp_gOpt        = CTTG_Koza2 ant ctx ,    
 
   cttp_ff          = FF5 "ff" "Problems.Ant.Funs" (), 
   

### 2013-07-29 v4/Problems/SSR/Problem.hs

@@ -24,7 +24,7 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_typ         = dou1                                         ,
   cttp_ctx         = ctx                                          ,
 
-  cttp_gOpt        = CTTG_Geom  dou1 ctx 0.75  , --CTTG_Koza2 dou1 ctx , 
+  cttp_gOpt        = CTTG_Geom  dou1 ctx 0.75 , --CTTG_Geom  dou1 ctx 0.75  , --CTTG_Koza2 dou1 ctx , 
   
   cttp_ff          = FF6 (asType::Double->Double) ff "Problems.SSR.Funs", 
   

### 2013-07-29 v4/ServerInterface.hs

@@ -31,8 +31,13 @@ stdoutCmd str = jsObj [
   ("type" , jsStr "stdout" ) ,
   ("msg"  , jsStr str      ) ]
 
-graphCmd :: Int -> Int -> (Double,Double,Double) -> Bool -> JSValue
-graphCmd runI genI (best,avg,worst) isWinner = jsObj [ 
+graphCmd :: Int -> Int -> (Double,Double,Double) -> Bool -> Maybe Double -> JSValue
+graphCmd runI genI (best,avg,worst) isWinner m_averageTermSize = jsObj $ 
+ ( case m_averageTermSize of
+     Nothing -> []
+     Just x  -> [("averageTermSize" , jsNum x )]
+ ) ++
+ [ 
   ("type"   , jsStr "generationInfo" ) ,
   ("i"      , jsNum genI ) ,
   ("run"    , jsNum runI ) ,

### 2013-07-29 v4/TTree.hs

@@ -22,7 +22,7 @@ import Data.Either
 import Data.Maybe
 
 import Text.JSON (JSValue (..) , toJSObject , toJSString )
-import Utils ( JShow , jshow_js )
+import Utils ( JShow , jshow_js, jss_size )
 
 import JSONUtils
 
@@ -55,6 +55,7 @@ instance Show CTT where
          in "\\ "++ vars ++ " -> " ++ show ttree 
 
 instance JShow CTT where
+  jss_size (CTT _ ttree) = Just $ ttreeSize ttree 
   jshow_js ctt = Just . jsStr . jsShow $ ctt
 --  jshow ctt = JSObject $ toJSObject [ 
 --   ("type"     , JSString . toJSString $ "jsonout"     ) ,
@@ -115,6 +116,8 @@ ttreeDepth (TTree _ _ ts) = case ts of
  [] -> 0
  _  -> (1+) . maximum $ map ttreeDepth ts
 
+ttreeSize :: TTree -> Int
+ttreeSize (TTree _ _ ts) = 1 + ( sum $ map ttreeSize ts )
 
 ttreeSubtree :: TTree -> TTPos -> TTree
 ttreeSubtree t [] = t

### 2013-07-29 v4/Utils.hs

@@ -218,6 +218,8 @@ align width str =
 -- JSON-show stuff ----
 
 class (Show a) => JShow a where
+  jss_size :: a -> Maybe Int
+  jss_size _ = Nothing
   jshow_js :: a -> Maybe JSValue
   jshow_js _ = Nothing
   jshow    :: a -> JSValue

### 2013-07-30 v4/Problems/SSR/Problem.hs
uterý 3:30

@@ -15,10 +15,10 @@ import GP_Data (CTTGen(..))
 
 reg = PO_CTTP_ PO_CTTP {
   cttp_code        = "ssr"                                        ,
-  cttp_info        = "Simple Symbolic Regression - new IM !"      ,
+  cttp_info        = "Simple Symbolic Regression."                ,
   cttp_data        = jsData                                       ,
-  cttp_numRuns     = IntSlider "Runs"            1 10    50   1   ,
-  cttp_numGene     = IntSlider "Generations"     0 100   50   10  ,
+  cttp_numRuns     = IntSlider "Runs"            1 100   1    1   ,
+  cttp_numGene     = IntSlider "Generations"     0 200   50   10  ,
   cttp_popSize     = IntSlider "Population size" 0 5000  500  100 ,
   
   cttp_typ         = dou1                                         ,

### 2013-07-30 v4/data/js/Problems/Fly02.js

@@ -18,7 +18,7 @@ var log = function(x){ return console.log(x); };
 //---------------------------------------------------------------------
 
 var onLoadedProblemData = function( data ){ 
-  log(data);
+  //log(data);
   problemData = data; 
 
 };

### 2013-07-30 v4/data/js/Problems/SSR.js

@@ -27,14 +27,18 @@ Global.Problems.ssr = {
   },
 
 
-  fenotyp : function( solution , ffVal ){
+  fenotyp : function( solution , ffVal , actGen ){
 
     ffVal = Math.round(ffVal*100)/100 ;
 
     this.plotIt(solution);
 
+    if( actGen == 0 ){
+      $('#ssr-links').append('<hr>');  
+    }
+
     this.solutions[this.i] = solution ;
-    $('#ssr-links').append( '<a href="#" title="'+ solution +'" '+
+    $('#ssr-links').append( '<a style="cursor:pointer" title="'+ solution +'" '+
       'onclick="Global.Problems.ssr.ssrLink('+this.i+')">' +ffVal + '</a> ' );
 
     this.i ++ ;

### 2013-07-30 tex/thesis.tex
uterý 4:07

@@
 \end{align*}
+where
+\begin{align*}
+rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
+p/q & \mbox{otherwise } \end{cases}  \\
+rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
+log(\vert x\vert) & \mbox{otherwise}. \end{cases}
+\end{align*}
 
@@
 
+\newpage
 \section{Artificial Ant}
@@
 
-The Santa Fe trail lies on toroidal square grid and has following layout.
-
-~\\~[ TODO OBRAZEK Santa Fe]\\
+The Santa Fe trail lies on toroidal square grid.
+The ant is in the upper left corner, facing right.
+It has following layout.
+
+\includegraphics[scale=1]{santafe.png}
 
@@
 \section{Big Context}
-
-TODO asi přenda do future work
 

### 2013-07-31 tex/thesis.tex
čtvrtek 0:23

@@
 %http://www.wikihow.com/Use-%22That%22-and-%22Which%22-Correctly
-
-% možná místo classical GP řikat standard GP jako montana95
 
@@
 
-\newcommand{\EA}{EA\xspace} % TODO nahradit správnym "evoluťionar algorithms"
+\newcommand{\atTree}{@-tree\xspace}
+
+%\newcommand{\EA}{EA\xspace} % TODO nahradit správnym "evoluťionar algorithms"
 
@@
 \noalign{\vspace{2mm}}
-Study programme: & Theoretical Computer Science \\ %Teoretická informatika \\
+Study programme: & Computer Science \\ %Teoretická informatika \\
 \noalign{\vspace{2mm}}
-Specialization: & 
+Specialization: & Theoretical Computer Science
 %Neprocedurální programování a umělá inteligence \\
-{\fontsize{0.3cm}{1em}\selectfont 
+%{\fontsize{0.36cm}{1em}\selectfont 
 %Neprocedurální programování a umělá inteligence} \\
-Non-Procedural Programming and Artificial Intelligence} \\
+%Non-procedural Programming and Artificial Intelligence} \\
+%}\\
 \end{tabular}
@@
 \noindent
-$<$dedication$>$
+%$<$dedication$>$
+I would like to express my gratitude to my supervisor RNDr. Petr Pudlák, Ph.D.
+for the useful comments, remarks and help with this master thesis.
+I would also like to thank my parents, who have supported me throughout
+my studies.
 
@@
 Název práce:
-Název práce
+Typed Functional Genetic Programming
 % přesně dle zadání
@@
 Autor:
-Jméno a příjmení autora
+Tomáš Křen
 
 Katedra:  % Případně Ústav:
-Název katedry či ústavu, kde byla práce oficiálně zadána
+Katedra teoretické informatiky a matematické logiky
 % dle Organizační struktury MFF UK
@@
 Vedoucí diplomové práce:
-Jméno a příjmení s tituly, pracoviště
+RNDr. Petr Pudlák, Ph.D., 
+Katedra teoretické informatiky a matematické logiky
+%Jméno a příjmení s tituly, pracoviště
 % dle Organizační struktury MFF UK, případně plný název pracoviště mimo MFF UK
@@
 Klíčová slova:
+geneticke programovani, funkcionalni programovani
 % 3 až 5 klíčových slov
@@
 Title:
+Typed Functional Genetic Programming
 % přesný překlad názvu práce v angličtině
@@
 Author:
-Jméno a příjmení autora
+Tomáš Křen
 
 Department:
-Název katedry či ústavu, kde byla práce oficiálně zadána
+Department of Theoretical Computer Science and Mathematical Logic
 % dle Organizační struktury MFF UK v angličtině
@@
 Supervisor:
-Jméno a příjmení s tituly, pracoviště
+RNDr. Petr Pudlák, Ph.D., 
+Department of Theoretical Computer Science and Mathematical Logic
+%Jméno a příjmení s tituly, pracoviště
 % dle Organizační struktury MFF UK, případně plný název pracoviště
@@
 Keywords:
+genetic programming, functional programming
 % 3 až 5 klíčových slov v angličtině
@@
 and set $T$ of constants and variables. 
-In classical GP \textit{closure} of those 
+In standard GP \textit{closure} of those 
 building blocks is required.
@@
 eliminates this constraint. 
-This can be achieved by utilization of \textit{types}.\\
-
+This can be achieved by utilization of \textit{types}.
 Utilization of types in GP is not a new idea; 
 it was dealt with in various previous works 
-\cite{yu01,montana95,haynes96,olsson94}.
+\cite{yu01,montana95,haynes96,olsson94}.\\
+
 The contribution of this thesis can be seen 
@@
 Hopefully, it is also shown that those two fields are
-nicely linked together, e.g., the concept of \textit{building blocks}
-corresponds to concept of \textit{context}. \\
-
-Notable highlights are: 
+nicely linked together.
+%, e.g., the concept of \textit{building blocks}
+%corresponds to concept of \textit{context}.
+Several enhancements are derived from this theoretical background
+and shown to have positive effect on performance of our system.  
+Our main focus lies in the process of generating the initial population.
+
+Design of a GP system utilizing these methods is presented.
+This design aims to be generalization of standard GP 
+developed by Koza\cite{koza92}, rather than being 
+specialized variant of GP using more recent techniques.
+This decision is based on belief
+that simplicity of design of standard GP plays nicely
+with testing those $\lambda$-calculus concepts in 
+the evolutionary context and on belief that modern specialized 
+techniques can be incorporated later.\\
+
+Notable highlights of the thesis are: 
 \begin{itemize}
  \item \textit{Inhabitation tree} (a slight modification 
-   of \textit{inhabitation machine} presented in \cite{barendregt10}
+   of \textit{inhabitation machine}\cite{barendregt10}
    suited better for reasoning about implementation details) 
@@
    \textit{search strategy}. Several search strategies are presented,
-   such as \textit{systematic} strategy and strategy performing
-   classical GP individual generation.     
+   such as \textit{systematic} strategy performing exhaustive enumeration
+   of individuals and strategy performing same individual generation 
+   as standard GP method \textit{ramped half-and-half}.  
+ \item One of those search strategies, the \textit{geometric} strategy, is
+   examined in a couple of experiments and shown to have 
+   several desirable properties.
+ \item The \textit{$\eta$-normalization} of generated individuals
+       and \textit{\atTree} representation of individuals, 
+       two enhancements of the system inspired from the theoretical
+       background of the system are examined and shown to have 
+       improving effect on performance of the system.     
  \item \textit{Abstraction elimination} is utilized
@@
 \end{itemize}
-
-Design of a GP system utilizing these methods is presented.
-This design aims to be generalization of classical GP 
-developed by Koza \cite{koza92}, rather than being 
-specialized variant of GP using more recent techniques.
-This decision is based on belief
-that simplicity of design of classical GP plays nicely
-with testing those $\lambda$-calculus concepts in 
-the evolutionary context and on belief that modern specialized 
-techniques can be incorporated later.
 
@@
 that \lterms can be easily translated to a 
-widely used programming language.
+widely used programming language.\\
 
@@
 bring breakthrough results; they are intended to 
-serve as demonstration of that the system is working properly
-and as demonstration of several directions in which the system
+serve as demonstration of that the system is working properly,
+as instrument for examining presented techniques and 
+as demonstration of several directions in which the system
 can be used.
-
-\begin{todo}
- & možná přidat ještě něco z poznámek co jsou teď v Conclusion
- & vlastně i koza hrotí evolution of constrained syntactic structures !!!
- & my tu řikáme že to závadíme kuli typum, ale my děláme v víc! -dáváme tam
-   i abstrakce atd - ale mužeme říct, že tim že se to pak převede zase na 
-   to bez abstrakcí tak setrváváme dal v krasnym bez-lambda abstrakčí ale
-   s tou výhodou že naše metoda gerneruje všechny struktury pro ten danej typ ...
-   nebo tak něco - asi to dát spíš do design of our system nebo někam
-   ale někde to asi zmínit
-\end{todo}
 
@@
 
-Another crucial input besides fitness function is a collection of \textit{building blocks}.
-It is collection of symbols (accompanied with an information about number of arguments).
+Another crucial input besides fitness function is a collection of 
+\textit{building blocks}\footnote{Throughout this thesis, 
+we use the term \textit{building blocks} 
+differently than it is sometimes used in GP context. 
+Hopefully it will not confuse the reader.}.
+It is a collection of symbols (accompanied with an information about number of arguments).
 Those symbols are used to construct trees representing individuals.\\
 
-\Lets describe GP algorithm briefly:
-
-At the beginning initial population is generated from building blocks randomly.
-
+\Lets describe GP algorithm briefly.
+At the beginning, initial population is generated from building blocks randomly.
 A step of GP algorithm is stochastic transformation of the current population into 	
@@
 \\\\
-\Lets now look on GP at more detail.
-
+\Lets now look on GP in greater detail.
+
+\newpage
 \section{Program trees}
@@
 \begin{samepage}
-\begin{itemize}
+\begin{enumerate}
 	\item Constant or variable symbol $s$ is S-expression.\footnote{
@@
 	      a more standard notation would be $f(e_{1}, ... ,e_{n})$.}   
-\end{itemize}
-\end{samepage}
-
+\end{enumerate}
+\end{samepage}~\\
 There is straightforward tree representation corresponding to these two cases:
 
-\begin{itemize}
+\begin{enumerate}
 	\item One node tree $s$.
-    \item \Tree
+    \item \mbox{\Tree
 			[.$f$	
@@
 		 		\text{...}
-		 		\text{$t_{n}$} ]\\\\
-		 Where $t_{1}, ..., t_{n}$ are trees corresponding 
+		 		\text{$t_{n}$} ]}\\\\
+		 where $t_{1}, ..., t_{n}$ are trees corresponding 
 		 to S-expressions $e_{1}, ..., e_{n}$.	   
-\end{itemize}
+\end{enumerate}
 
@@
 may occur variables $x_{1}, ..., x_{n}$.
-
-\begin{todo}
- & nějak to tu přepsat aby to bylo ve stejnym stylu jako v \ref{tree-reps}
-\end{todo}
 
@@
 We can say that the motivation behind this thesis is to construct system,
-where we eliminate this constraint.
+where building blocks may be of any type.
 
@@
 
-We will describe tree generating method described by Koza 
-called \textit{ramped half-and-half}  \cite{koza92}. \\
+We will describe tree generating method described by Koza \cite{koza92}
+called \textit{ramped half-and-half}. \\
 
@@
 
-\begin{todo}
- & Dozjistit a dopsat jak to Koza má s vyžadovánim unikátnosti při generování.
-\end{todo}
+Generated individual is added into the initial population only if 
+the initial population does not yet contain this individual.
+This way it is guaranteed that every individual is unique 
+in the initial population.
 
@@
 It uses fitness value of each individual in straightforward way to determine probability of selecting this individual.\\
-
+\newpage
 Let there be $popSize$ individuals in the population.\\
@@
 This crossover randomly selects one node in each parent tree.
+
+With probability $p_{ip}$ (e.g. $p_{ip} = 90\%$) the crossover node
+is randomly (uniformly) selected from internal nodes (i.e. function nodes) and
+with probability $1 - p_{ip}$ (e.g 10\%) from leaf nodes.
+This is done for both parents.
+
 Two new child individuals are constructed by swapping subtrees 
@@
 
-And here are two child trees with swapped subtrees:
+And here are two child trees with swapped subtrees:\\
 
@@
 
-\begin{todo}
- & POZOR POZOR !!!!! zapoměl jsem tam že je 90\% procent šance 
-   že se vybere internal node a jen 10\% šance že terminal node
-\end{todo}
-
 \section{Reproduction}
@@
 
-\section{Construction of a next population}
+\section{Construction of next population}
 
@@
 
-\begin{todo}
- & !!!
- & zmínit tu terminate bool
-\end{todo}
+Evaluation is process of assigning fitness value to all 
+individuals in the population. It is also usual to check
+individuals for being correct solution.
+If such a correct solution is find, the algorithm terminates. 
+In our implementation fitness function returns
+pair of real number (fitness value) and boolean value (indicating
+whether tested individual is correct solution).
 
@@
 Bellow is described GP algorithm in pseudocode. 
+It is slight modification of the standard GP algorithm, 
+because the best individual of the 
+generation is preserved unchanged into the next generation.
 
 \Pseudokod{GP( fitness, \TuF, popSize, numGens, probabs )}{
-  $gen \leftarrow$ 0 \;\;
-
+  $gen \leftarrow$ 0 \;
   $pop \leftarrow generateInitialPopuletion( $ \TuF$ ) $ \;
-  ($popWithF,terminate,best) \leftarrow$ 
-  evaluate($fitness$, \TuF, $pop$)\;\;
+  ($popWithFitness,terminate,best) \leftarrow$ 
+  evaluate($fitness$, $pop$)\;\;
 
@@
 			\Case{Crossover}{
-				$parent1 \leftarrow$ selection( $popWithF$ ) \;
-				$parent2 \leftarrow$ selection( $popWithF$ ) \;\;
+				$parent1 \leftarrow$ selection( $popWithFitness$ ) \;
+				$parent2 \leftarrow$ selection( $popWithFitness$ ) \;\;
 
@@
 			\Case{Reproduction}{
-				$indiv \leftarrow$ selection( $popWithF$ ) \;
+				$indiv \leftarrow$ selection( $popWithFitness$ ) \;
 				$newPop$.insert( $indiv$ ) \;
@@
 			\Case{Mutation}{
-				$indiv \leftarrow$ selection( $popWithF$ ) \;
+				$indiv \leftarrow$ selection( $popWithFitness$ ) \;
 				$mutant \leftarrow$ mutate( $indiv$ , \TuF ) \;
@@
 	$pop \leftarrow newPop$  \;
-	($popWithF,terminate,best) \leftarrow$ 
-	evaluate($fitness$, $TuF$, $pop$)\;
+	($popWithFitness,terminate,best) \leftarrow$ 
+	evaluate($fitness$, $pop$)\;
 	$gen \leftarrow gen + 1$ \; 
@@
 
- \Return pop \;
+ \Return $popWithFitness$ \;
 }
@@
 
-Behaviors of contained procedures
-generateInitialPopuletion(),
-evaluate(),
-probabilisticallySelectOperation(),
-selection(),
-crossover() 
-and mutation() are hopefully clear from verbal description.
-
-\begin{todo}
- & mention best preservation and same popSize checking.
- & popsat co je to popWithF
- & popsat co je to terminate 
- & ten algoritmus v kozovi dělá "Designate Result" já tam vracim 
-   poslední populaci
-\end{todo}
+Behaviors of contained procedures are hopefully clear from verbal description.
+\\[2em]
+
+For more detailed information on subject of GP and its many variants
+see \cite{fg} containing many references to other works.
 
@@
 \newcommand{\Icomb }{\mathbf{I}}
+
+\newcommand{\bbarr}{\twoheadrightarrow_\beta}
+\newcommand{\barr}{\rightarrow_\beta}
+\newcommand{\beq}{=_\beta}
+
+\newcommand{\eearr}{\twoheadrightarrow_\eta}
+\newcommand{\earr}{\rightarrow_\eta}
+\newcommand{\eeq}{=_\eta}
+
+\newcommand{\bearr}{\rightarrow_{\beta\eta}}
+\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
+\newcommand{\beeq}{=_{\beta\eta}}
 
@@
 \begin{todo}
+ & všude nahradit ty množinový za $\Gamma , x:\sigma$ a taky $(x,\Sigma)$ kde se to nehodí.
  & Až to bude dostatečně sepsáno, tak do této kapitoly taky přidat
@@
 $\Scomb, \Kcomb, \Icomb$ as abbreviation
-for specific terms, e.i. 
-$\Icomb = \lamb{x}{x}$.  
+for specific fixed terms, e.g. 
+$\Icomb$ always means $\lamb{x}{x}$.  
 \\[1em]
@@
 x[x:=N]           &= N \\
-y[x:=N]           &= y \text{ ~~~~where } x \not= y  \\
+y[x:=N]           &= y &\text{ where } x \not= y  \\
 (P~Q)[x:=N]       &= ( P[x:=N]  ~ Q[x:=N] ) \\
 \lamb{x}{P}[x:=N] &= \lam{x}{P}\\
-\lamb{y}{P}[x:=N] &= \lam{y}{( P[x:=N] )} \text{ ~~~~where } x \not= y
+\lamb{y}{P}[x:=N] &= \lam{y}{( P[x:=N] )} &\text{ where } x \not= y, y\not\in \FV(N)\\
 \end{align*}
 
-\begin{todo}
-& možná zmínit že y nesmí být v N volně v připadě\\
-  $\lamb{y}{P}[x:=N] = \lam{y}{( P[x:=N] )}$ , 
-  ale řekl bych že zbytečnej detail
-\end{todo}
+In order to satisfy the $y\not\in \FV(N)$ condition, names of bound variables
+are chosen such that they differ from the free ones in the term.
 
@@
 of insertion of arguments supplied to a function into 
-its body.
+its body. \\
+
+Another useful relations are $\bbarr$ and $\beq$ defined as follows.
+
+\begin{enumerate}
+ \item \begin{enumerate}
+ 	\item $M \bbarr M$
+ 	\item $M \barr N \then M \bbarr N$
+ 	\item $M \bbarr N , N \bbarr L \then M \bbarr L$ 	
+ \end{enumerate}
+ \item \begin{enumerate}
+ 	\item $M \bbarr N \then M \beq N$
+ 	\item $M \beq N \then N \beq M$
+ 	\item $M \beq N , N \beq L \then M \beq L$
+ \end{enumerate}
+
+\end{enumerate}
+
+We read those relations as follows.
+\begin{enumerate}
+ 	\item $M \bbarr N$ --- "$M$ $\beta$-reduces to $N$."  
+ 	\item $M \barr N$  --- "$M$ $\beta$-reduces to $N$
+ 	      in one step."
+ 	\item $M \beq N$ --- "$M$ is $\beta$-convertible to $N$."	
+ \end{enumerate}
 
 \begin{todo}
-& asi to přepsat nějak obecně pro redukci R a pak zadat beta,eta,eta-1 
-  jako specialni připady
-& !!!!!!! zmínit $=_\beta$ a $\twoheadrightarrow_\beta$  a i pro ostatní redukce!!!!!
 & definovat \textbf{CR} a říct o tom lemátku, že má CR redukce má unikátní nf.
-&& V \cite{barendregt84} (ta nová kniha s žlutejma deskama) je ukázano že je 
-	 eta CR (Church Rosserovská)
+&& V \cite{barendregt84} (ta nová kniha s žlutejma deskama) je ukázano že je eta CR (Church Rosserovská)
 & zmínit Theorem 2B.4 z \cite{barendregt10}
-& co to znamená mít nf
 \end{todo}
@@
 
-$$\lamb{x}{(M~x)} \eArrow M \text{ ~~~~where } x \not\in FV(M) $$
+$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$~
+
+Analogically, a term of a form $\lamb{x}{(M~x)}$ is called 
+\textit{$\eta$-redex}.
+
+Relation $\bearr\;=\;\barr \cup \earr$. 
+(Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.)
+
+Similarly as for $\bbarr$ and $\beq$ we can define relations 
+$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.
 
@@
 
-$$M \eeArrow \lamb{x}{(M~x)} \text{ ~~~~where } x \not\in FV(M) $$
-
-%\subsection{Reductions for $c \in C$}
-%\textbf{TODO možná vynechat...}
+$$M \eeArrow \lamb{x}{(M~x)} \textbf{ ~~~~if } x \not\in FV(M) $$
 
@@
 
-A \lterm is a \textit{$\beta$-normal form} if it does not have a $\beta$-redex as
-subterm.
-\\\\
+~\begin{easylist}[enumerate]
+& A \lterm is a \textit{$\beta$-normal form} (\bnf) 
+if it does not have a $\beta$-redex as subterm.
+& A \lterm M \textit{has} a \bnf if $M \beq N$
+and $N$ is a \bnf.\\
+\end{easylist}
 A normal form may be thought of as a result of a term evaluation.
+
+Similarly we can define \enf and \benf.
 
@@
 \newcommand{\SexprTree}{Sexpr-tree\xspace}
-\newcommand{\atTree}{@-tree\xspace}
 
@@
 
-\begin{todo}
- & napsat transformace mezi \atTree a \sexprTree
-%& ujistit se že je v části o tree reprezentacích 
-%  je napsaný co to znamená že podstrom má nějakej typ,
-%  že podstromy odpovídaj podvýrazum atd
-% & EXAMPLES of tree representations of \lterms  
-% & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
-%   přímo \sexprTree a říct proč se tomu tak řiká 
-%   (na tom místě co je v textu dřív) 
-% & zkontrolovat že reference tree-reps se odkazuje na dobrý místo,
-%   mohlo by se to rozbít tim že tree reprezentations - typovany/netypovany 
-%   jsou od sebe oddelený
-\end{todo}
-
 \newpage
@@
 
-%\begin{definition}
-%Let $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$. 
-%($\Gamma$ is a set of statements of a form $M : \sigma$.)	
-%Then $\Gamma$ is {\it context} if it obeys following 
-%conditions.
-%%\footnote{The $\pi_1$ corresponds to the projection of the 
-%%first component of the Cartesian product.}
-%\begin{align*}
-%\forall (x,\sigma) \in \Gamma &: x \in V \cup C  
-%&\textit{(x is variable or constant)}\\
-%\forall (x_1,\sigma_1),(x_2,\sigma_2) \in \Gamma &: 
-%x_1,x_2 \in V , x_1 \neq x_2 \then \sigma_1 \neq \sigma_2
-%&\textit{(variables are distinct)}
-%\end{align*}~
-%Ten druhej je navíc blbě, má tam bejt celej stejtment se nerovná pak...
-%\end{definition}
-
 \begin{definition}
@@
 In our system, the set of \textit{building blocks} 
-(the equivalent of the set $T \cup F$ in classical GP)
+(the equivalent of the set $T \cup F$ in standard GP)
 will be a context with all declarations having \textit{constants}
@@
 this initial context is being further extend with variables
- --- by "entering inner scopes of bodies of lambda functions".
-
-%\subsection{Statements of the form \GMS}
-
-\begin{todo}
- & zapsat sem notaci s $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
-\end{todo}
-
-%\subsection{Inference rules}
+ --- by "entering inner scopes of bodies of lambda functions".\\
+
+Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
+such that $\Gamma$ does not contain any declaration with $x$ as subject.
+
 \subsection{Simply typed $\lambda$-calculus}
@@
 %        a terminující je rpotože zmenšuje term s nemuže zmenšovat donekonečna    
-   & možná zmínit že se jedná o \textit{gentzen style proof}
-   & říct že jsou to modus ponens a jeden směr věty o dedukci  
 \end{todo}
@@
 
-\begin{todo}
- & možná konec týhle subsekce trochu přeformulovat, je to takový kostrbatý...
-\end{todo}
-
 \section{Term generating grammar}
+\label{Term-generating-grammar}
 
@@
 For every non-terminal $(\sigma \ar \tau,\Gamma)$
-and for every variable $x$
-there is following grammar rule.\footnote{
-\textbf{TODO !!!!} nějak okomentovat že by měla být freš,
-pokudmožno v celym termu ale to že byse dělalo blě aby to
-zůstala bezkontextová gramatika, ale je to celkem nepodtatnej detail
-protože tyhle term generating gramatiky sou tu spíš
-pro zajímavost a pro pevnější kramfleky, než pro praktický 
-generování}
+and for every variable $x$ such that $\Gamma$ does not contain any 
+declaration with $x$ as subject
+there is following grammar rule.
 
@@
 	( \sigma \rightarrow \tau , \Gamma )  \gar
-	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
+	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau ; \Gamma ,x:\sigma ) \quad \bigg)
 $$\\
@@
 
-\begin{todo}
-   & buď dokázat ty převody na grammar rules explicitně
-     nebo říct že se to dokáže podobnym stylem jako v dukazu 
-     v sekci \ref{barlike}
-   & možná ocitovat toho van Wijngaarden [1981] s těma 2-level grammar
-     jak je to v Barendregtovi \cite{barendregt10} 
-   & vyřešit jak to tu popsat s tou freš proměnnou x v introdukci...
-     && stejně tak v tom obecnym případě to dělá neplechu
-\end{todo}
-
 \section{Long normal form}
@@
 \end{proposition}
-
-\begin{todo}
- & potřebuju eště nějaký \textbf{propositions} z \cite{barendregt10} tu uvíst ?
- & ta definice je tam daná pro $\tur{}{M}{\sigma}$,
-   ukazat ze ta naše je ok, že tam má navíc to $\Gamma$.
-   Stačí udělat tríček že beru funkci co má lambda hlavu odpovídající 
-   tomu kontextu.
-\end{todo}
 
@@
 
-\begin{todo}
- & možná zmínit že podobně jako mužeme transformovat inference rules
-   na grammar rules, mužeme i obráceně grammar rules
-   převést na inference rules - to bylo ostatně uděláno v tom dukazu ale
-   je to podle mě hezký říct ty pravidla explicitně.
-\end{todo}
-
 \subsection{Benefits of generating \lterms in \lnf}
@@
 
-\begin{todo}
-& možná eště zmínit něco jako : \\
-Jedna drobná nevýhoda je, že pokud chceme generovat termy od nejmenšího do největšího
-a pokud lnf generujeme od nejmenšího po největší tak pak generovaní \benf
-timhle trikem nemusí bejt od nejmenšího po nějvětší. V naší aplikaci (tzn GP)
-ale nijak nebazírtujeme na týhle vlastnosti a nakonec budeme spíš stejně používat
-nějaký nedeterminizmy aby to fungovalo líp, takže to vubec nevadí, že budou trochu přeházený. 
-\end{todo}
-
 \newpage
@@
 \Lets consider following definition of 
-\textit{and-or tree}\footnote{
-\textbf{TODO}: Mention that on [WIKI] there is more general definition, but 
-for our purposes is this one sufficient.}.
+\textit{and-or tree}.
+%\footnote{
+%\textbf{TODO}: Mention that on [WIKI] there is more general definition, but 
+%for our purposes is this one sufficient.}
 
@@
 		  \item Each \textit{or-node}  in $T'$ has precisely one child 
-		      node.\footnote{\textbf{TODO}: MENTION why precisely one 
-		      and not at least one ..or CHANGE the def. }   
+		      node.
+		      %\footnote{\textbf{TODO}: MENTION why precisely one 
+		      %and not at least one ..or CHANGE the def. }   
 	    \end{enumerate}
@@
 \newpage
-\section{ Generating terms in \lnf by solving inhabitation tree }
+\subsection{ Generating terms in \lnf by solving inhabitation tree }
 
@@
   je dost rychle a mužou bejt divný. 
-& A možná to ještě to nějak okomentovat
 \end{todo}
+
+\newpage
+\subsection{In defense of inhabitation trees}
+
+It may seem that we are overthinking the problem by introduction
+of this new structure, whereas it would be sufficient to
+use the term generating grammars instead.\\
+
+I see two reasons why inhabitation tree is useful concept.
+
+\begin{enumerate}
+\item The first reason is rather subjective.
+      In my opinion, it simplifies the problem to be
+      better graspable. My arguments are following: 
+      \begin{enumerate}
+      \item
+      It reduces the unpleasant fact
+      that term generating grammars are not grammars in
+      the standard sense. 
+      \item
+      It brings the problem to 
+      its more natural domain, 
+      from (textual) grammars 
+      to (syntactical) trees.
+      \item
+      We may think about the whole structure of generating 
+      at once since the whole tree, although infinite, is
+      fairly simple structure of alternating layers of \textit{and} 
+      and \textit{or} nodes. 
+      \item
+      It enables us to ask such questions
+      as: What would be usage of inhabitation trees with 
+      number labeled edges coming from \textit{or}-nodes? Et cetera.
+      \end{enumerate}   
+\item We can think about the inhabitation tree as about 
+      some kind of search space. But another view is also 
+      possible: We can think about it as about collection
+      of typed lambda terms that are generable by that inhabitation
+      tree.
+
+      This is even more visible when we introduce
+      concept of an \textit{incomplete inhabitation tree}
+
+\end{enumerate}
+
+In the next chapter we will use A*-algorithm together with
+inhabitation tree ideas to perform systematic generation and other
+kinds of generation of lambda terms;
+but maybe even more interesting application of inhabitation trees
+involving some above mentioned ideas will be revealed in the 
+chapter about future work, since those ideas are not yet implemented.
 
@@
 \begin{enum}
- & It should eliminate the \textit{closure} constrain of classical GP
+ & It should eliminate the \textit{closure} constrain of standard GP
    (see \ref{building-blocks}). 
- & It should be generalization of the classical GP described in \ref{GP},
+ & It should be generalization of the standard GP described in \ref{GP},
    by which is meant that for \textit{building blocks} satisfying 
-   constrains of classical GP there should be simple setup of control parameters
+   constrains of standard GP there should be simple setup of control parameters
    of the system which will make the system behave in precisely the same way
-   as the classical GP behaves. 
+   as the standard GP behaves. 
  & It should utilize the theory around typed $\lambda$-calculus.  
@@
 
- \item It also makes our method compatible with classical GP, where
+ \item It also makes our method compatible with standard GP, where
        variables are treated as ordinary element of the set $T$
@@
        This ensures that S-expressions will be generated for
-       $\Gamma$ satisfying constrains of classical GP and that
+       $\Gamma$ satisfying constrains of standard GP and that
        abstraction elimination will have no effect on such individuals.
@@
 
+This heuristic function is admissible because every unfinished
+leaf must be at least once expanded to become final.
+
+\subsection{Future improvements}
+
 \begin{todo}
-& rozebrat to, ukázat že je optimistická atd...
+& !!!
 \end{todo}
@@
 our formal states contain above each symbol node a type node.
+
+\begin{todo}
+ & Přidat že pak mužeme čekovat unikátnost při generování, protože už neni zajištěna 
+   jako u systematickýho. Nebo říct že musíme ale pak asi změnit implementaci tý geometrický což znevalidní experimenty
+\end{todo}
 
@@
 
-The crossover operation in classical GP is performed 
+Design goal behind our approach to crossover operation is
+to try to generalize standard tree swapping mutation.
+This is not because of dislike of more specialized crossovers;
+it comes from systematic approach for exploring the domain of
+typed functional GP --- thus I would like to start with the most common one.\\
+
+%In the future, after finding sufficiently nice generalization of the standard tree %swapping crossover I plan to move my attention to more 
+%specialized crossovers, like the one presented in \cite{yu01}.  \\
+
+The crossover operation in standard GP is performed 
 by swapping randomly selected subtrees in each parent 
-\sexprTree (see \ref{GPxover}).
-
-For typed lambda terms two difficulties arise: Types and variables.
-
-We will show how to crossover typed \lterm trees in both 
+S-expression (see \ref{GPxover}).
+
+For typed lambda terms two difficulties arise: Types and variables.\\
+
+We will show how to crossover \textit{typed \lterm trees} in both 
 \atTree and \sexprTree notation.\\
 
-As in classical GP our crossover will be performed by swapping
+As in standard GP our crossover will be performed by swapping
 two subtrees. But now with constraint that both subtrees have
@@
 
-In \ref{repairing-method} is presented more detailed discussion of problems with
-variables and method which repairs such "broken" terms.
+%In \ref{repairing-method} is presented more detailed discussion of problems with
+%variables and method which repairs such "broken" terms.
+
+First approach tried at this problem was to repair the broken term,
+but since this method is not currently implemented we will describe it
+in just a few words. If the moved free variable ends in the scope 
+where is defined variable with the same name and type, then it
+is OK. If it is not this case, we may rename this variable to some 
+other variable "visible" in its new scope if such a variable has the same 
+type. If there is no such a suitable variable we may generate new value 
+to replace the variable by term generating method.
+
+As you can see, this is not much elegant solution.
 
@@
 \textit{Abstraction elimination} is a process of transforming 
-an arbitrary \lterm into \lterm witch contains no lambda abstractions
+an arbitrary \lterm into \lterm that contains no lambda abstractions
 and no bound variables.
@@
 
-\Lets describe transformation $\Ae$ performing this process.
+\Lets describe transformation $\Ae$ performing this 
+process.\footnote{In our implementation we must also deal with types, because 
+our individuals are annotated with types. But since this process
+is fairly straightforward but cumbersome to describe, we will 
+skip explaining it here. Reader interested in seeing it can see 
+the source code of our implementation. }
 \begin{align*}
@@
 
-\begin{todo}
-& přidat (ale i do kódu) B a C
-& říct že my musíme řešit i typy 
-& říct o složitosti , a že je to docela
-disadventič vzhledem k bloatu
-& citovat Simon Peyton Jones implementaition of FP languages
-jako source of further reading on this subject.
-\end{todo}
+\newcommand{\Bcomb }{\mathbf{B}}   
+\newcommand{\BBcomb }{\mathbf{B*}}
+\newcommand{\Ccomb }{\mathbf{C}}   
+\newcommand{\CCcomb }{\mathbf{C'}}
+\newcommand{\SScomb }{\mathbf{S'}}
+
+This is simple version of this process. More optimized version,
+in the means of the size of resulting term and its performance
+is following one, presented in \cite{jones87}.\\
+
+This version operates with more combinators:\\
+
+\begin{align*}
+\Bcomb  &= \lam{f\,g\,x}{f\,(g\,x)} \\
+\Ccomb  &= \lam{f\,g\,x}{f\,x\,g} \\
+\SScomb &= \lam{c\,f\,g\,x}{c\,(f\,x)\,(g\,x)} \\
+\BBcomb &= \lam{c\,f\,g\,x}{c\,(f\,(g\,x))} \\
+\CCcomb &= \lam{c\,f\,g\,x}{c\,(f\,x)\,g} \\
+\end{align*}
+
+\newpage
+And the transformation can be written as follows.
+
+\begin{align*}
+\Ae[x]           &= x &\\
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\
+\Ae[\lam{x}{M}]  &= A[ x ; \Ae[M] ] &
+\\[1em]
+A[x;x]           &= \Icomb &\\
+A[x;y]           &= (\Kcomb~y)\\
+A[x;\,(M\,N)\,]  &= Opt[~\Scomb~(A[x;M])~(A[x;N])~]
+\\[1em]
+Opt[~\Scomb~(\Kcomb~M)~(\Kcomb~N)~]   &= \Kcomb~(M~N)\\
+Opt[~\Scomb~(\Kcomb~M)~\Icomb~]       &= M\\
+Opt[~\Scomb~(\Kcomb~M)~(\Bcomb~N~L)~] &= \BBcomb~M~N~L\\
+Opt[~\Scomb~(\Kcomb~M)~N~]            &= \Bcomb~M~N\\
+Opt[~\Scomb~(\Bcomb~M~N)~(\Kcomb~L)~] &= \CCcomb~M~N~L\\
+Opt[~\Scomb~M~(\Kcomb~N)~]            &= \Ccomb~M~N\\
+Opt[~\Scomb~(\Bcomb~M~N)~L~]          &= \SScomb~M~N~L
+\end{align*}
+\\
+
+As is stated in \cite{jones87},
+the biggest disadvantage of this technique is that the translated
+term is often much larger than in its lambda form --- the size of
+the translated term can be proportional to the
+square of the size of the original term.
+
+But the advantage is also tempting --- no need to deal with variables
+and lambda heads.
 
@@
 
+First thing to do in standard subtree swapping is to select random node
+in the first parent.
+
+We modify this procedure so that we allow
+selection only of those nodes with such a type that there exists  
+a node in the second parent with the same type.
+
+Standard subtree swapping crossover as a first thing selects 
+whether the selected node will be inner node (usually with probability 
+$p_{ip} = 90\%$) or leaf node (with probability 10\%).
+
+We are in a more complicated situation, because one of those 
+sets may be empty, because of allowing only nodes with possible "partner"
+in the second parent. Thus we do this step only if both sets are
+nonempty.
+
+After selecting a node in the first parent we select node in the
+second parent such that type of that node must by the same as the type 
+of the first node. Again, this may eliminate the "90-10" step of
+first deciding whether the selected node will be internal node 
+or leaf node.
+
+When both nodes are selected we may swap the trees. \\
+
+If the abstraction elimination was performed, then 
+since the trees are of the same type and there are no variables to be 
+moved from their scope, the offspring trees are well typed.\\
+
+Both \sexprTree and \atTree are able to be crossed by this 
+mechanism. But \atTree has more possibilities then \atTree.
+This comes from the fact that every subtree of the \sexprTree
+corresponds to a subtree of \atTree, but there are subtrees
+of \atTree that do not correspond to a subtree of a \sexprTree.\\
+
+Following example should clarify this.
+
+\Tree[.@	
+   [.@ \textbf{f} x ]
+   [.y ]  		 			
+]
+\Tree[.\textbf{f} x y ]~\\
+
+In \atTree, \textbf{f} is leaf thus subtree, 
+whereas in \sexprTree it is internal node thus not a subtree.\\
+
+Another nice property of \sexprTree{}s with no lambdas 
+is that they are the same representation as S-expressions
+used by standard GP.
+
+%\subsection{ Repairing method }
+%\label{repairing-method}
+%
+% 
+%
+%\begin{todo}
+%& !!!
+%\end{todo}
+%
+%
+%\subsection{ Comparison of the two methods }
+%\label{comarison-ski-repairing}
+%
+%\begin{todo}
+%& !!!
+%& říct že nakonec sme přilnuli k zbavit se proměnejch
+%\end{todo}
+
+\chapter{ Implementation }
+
+In this chapter we will briefly discus some parts of the implementation
+of our system. Some previous knowledge of Haskell is assumed, but one can
+freely skip this chapter since it is not crucial for the 
+overall message of this thesis. \\
+
+Our system is implemented in functional language \textit{Haskell}.
+Program implementing it is called \textit{Fishtron}.
+It runs as a simple server application and its GUI is
+accessible through web browser.\\
+
+Evaluation of individuals is performed by translating them
+from \lterms into Haskell programs which are evaluated
+in the Haskell interpreter \textit{Hint} accessible 
+in the code as \texttt{Interpreter} monad.  \\
+
+Each problem definition typically consists of two Haskell 
+source code files. First file (usually called \texttt{Problem.hs}) 
+contains problem definition involving such things as desired 
+type of individual, context of building blocks, 
+fitness function implementation and various control parameters 
+like term generating method, crossover,
+default number of runs, generations, population size etc.
+
+Those functions or constants of building block context
+that are not part of the standard Haskell \textit{Prelude}\footnote{
+A standard module imported by default into all Haskell modules.}
+need to be defined in the second file (usually called \texttt{Funs.hs}).\\
+
+After start of the program one can access the interface
+by pointing a browser to \texttt{http://localhost:3000}. Here is simple 
+form where registered problems are listed. Beside problem selection
+one can also change some control parameters. After hitting the "start"
+button evolution begins. From that point until the end of the running
+experiment, the front-end JavaScript program periodically asks
+server for new run information. Server makes this information available 
+at the end of each generation. This information contains various
+data: Haskell and JavaScript source codes of the best individual,
+some statistical data about the just evaluated generation,  
+standard output of the server program, etc.
+
+These data are immediately available in the form of
+continuously updated graphs and if 
+the \textit{fenotype} JavaScript file is present 
+for the problem, then is also available immediate 
+presentation of the best individual from each generation.\\
+
+Now \lets move our attention to the architecture of the evolution engine.
+Various dirty parts of the outside world (such as textual output,
+communication with front-end, access to random number generator) are  
+packed into \texttt{Eva} state monad defined in \texttt{Eva.hs}.\\
+
+Core of the evolution engine (defined in \texttt{GP\_{}Core.hs})
+is done so it enables various structures
+to be evolved - such as \sexprTree{}s, \atTree{}s or even boolean lists
+of fixed length as to support beside GP also Genetic Algorithms.\\[1em]
+
+This is possible due to following type classes.
+
+~\\\texttt{
+class Gene term opt where\\
+$~\;\;$generateIt :: Int -> opt -> Eva [term]\\[1em]
+class Muta term opt where\\
+$~\;\;$mutateIt :: opt -> term -> Eva term \\[1em]
+class Cros term opt where\\
+$~\;\;$crossIt :: opt -> term -> term -> Eva (term,term) \\[1em]
+class Evolvable term a gOpt mOpt cOpt where\\
+$~\;\;$evolveIt :: RunInfo -> Problem term a gOpt mOpt cOpt  \\
+$~~~~~~~~~~~~~~~~~~~~~~$
+-> Eva (term,FitVal,Maybe Int)
+} ~\\
+
+Put simply, if we define instances for 
+\texttt{Gene}, \texttt{Muta} and \texttt{Cros}, then we get
+instance of \texttt{Evolvable}.
+
+\texttt{Gene} correspond to individual generating method for
+structure \texttt{term} parametrized by option \texttt{opt}.
+Analogically for mutation and crossover.
+
+Instances of those three type classes are located in the file
+\texttt{GP\_{}Data.hs}.
+
+We will skip deeper explanation of this topic now.\\
+
+One technical detail that I think is worth mentioning here
+is a trick involved in evaluation of Haskell programs 
+corresponding to individuals. Straightforward evaluation 
+technique would be to simply call evaluation function on every
+string corresponding to individual expression.
+
+But since most time consuming activity
+of evaluating using \textit{Hint} is its initialization,
+it is much more effective to collect all the individuals  
+to be evaluated in the current generation and create huge
+string representing list of expressions
+and evaluate this string of all individual. \\
+
+Another notable comment about implementation comes
+from fact that \textit{Hint} needs to know type of the expression
+being evaluated. This particular thing had enormous enlarging effect
+on pain produced by need to fight the type system in order to
+make it reasonably working piece of software. This may be caused
+by the fact that I am still pretty inexperienced Haskell programmer,
+since this is my first bigger program done in this language.
+
+Moral of this remark is the following one: 
+If you plan to use the \textit{Hint}
+in your Haskell program on place where it will be used to eval 
+expressions of various types, think through the integration of the
+evaluator into the whole system in advance.\\
+
+\newpage
+\Lets finish this brief chapter about implementation of our system
+by listing other not yet mentioned source files with a little information 
+about them.   \\[1em]
+
+\begin{easylist}[itemize]
+& \texttt{Server.hs}    --- File containing main of the server.
+& \texttt{Register.hs}  --- Place for registering the problem to make it appear 
+                            in the list of problems in GUI.
+& \texttt{InhabTree.hs} --- Core of the term generating method.
+& \texttt{TTerm.hs}     --- Code related to typed \lterms 
+                            represented as \atTree{}s. 
+& \texttt{TTree.hs}     --- Code related to typed \lterms 
+                            represented as \sexprTree{}s.
+& \texttt{Heval.hs}     --- Evaluator wrapper code.
+& \texttt{Dist.hs}      --- Implementation of "probability distribution" data 
+                            structure used as container for population.
+\end{easylist}
+
+\chapter{Results}
+	In this section will be presented usage of the system in order to solve specific problems. Several problems will be described and some of them compared with results
+of others.
+
+\section{Simple Symbolic Regression}
+
+\textit{Simple Symbolic Regression} is a problem described
+in \cite{koza92}. Objective of this problem is to 
+find a function $f(x)$ that fits a sample
+of twenty given points. The target function is 
+function $f_{t}(x) = x^4 + x^3 + x^2 + x$.
+
+Desired type of generated programs $\sigma$ and 
+building blocks context $\Gamma$ are following.
+\newcommand{\Real}{\mathbb{R}}
+\begin{align*}
+\sigma = \Real \ar &\Real\\
+\Gamma = \{
+  (+)  &: \Real \ar \Real \ar \Real    ,\\
+  (-)  &: \Real \ar \Real \ar \Real    ,\\
+  (*)  &: \Real \ar \Real \ar \Real    ,\\
+  rdiv &: \Real \ar \Real \ar \Real    ,\\
+  sin  &: \Real \ar \Real              ,\\
+  cos  &: \Real \ar \Real              ,\\
+  exp  &: \Real \ar \Real              ,\\ 
+  rlog &: \Real \ar \Real              \}\\
+\end{align*}
+where
+\begin{align*}
+rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
+p/q & \mbox{otherwise } \end{cases}  \\
+rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
+log(\vert x\vert) & \mbox{otherwise}. \end{cases}
+\end{align*}
+
+Fitness function is computed as follows
+
+$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
+
+where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.\\
+
+An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
+considered as a correct individual.\\
+
+\newpage
+\subsection{Experiments}
+
+On this problem we demonstrate that if control parameters are set to be
+compatible with standard GP, then the results obtained fit the results
+presented in \cite{koza92} relatively well. \\
+
+Those control parameters are following two:
+
+~\begin{enum}
+ & \textit{Ramped half-and-half strategy} is used as search strategy.
+ & Option of preserving the best individual into the next population is set \textbf{off}.
+\end{enum}~
+
+The experiment consisted of 50 independent runs of GP algorithm.
+Each run had maximally 50 generations (50 + 1 for generation 0)
+and 500 individuals as population size.\\
+
+We analyze the ability of the system to produce a correct solution.
+We are interested in percentage of runs that led to discovery of
+correct solution. And we also use the popular measurement  
+method within GP field --- the \textit{performance curves}
+described in \cite{koza92}.
+
+First performance curve is the cumulative \textit{probability of success}
+$P(M,i)$, where $i$ is generation number ($i = 0$ for initial population) 
+and $M$ is the size of the population. 
+$P(M,i)$ is probability that a correct solution will
+be find in one run of GP with maximal number of generations $i$ and 
+population size $M$. $P(M,i)$ is computed as percentage of 
+runs yielding a correct solution in generation $i$ or earlier.
+
+From values of $P(M,i)$ are computed values of the second performance curve
+$I(M,i,z)$
+--- the total number of individuals that must be processed to yield a correct
+solution with probability $z$ by multiple runs of GP with maximal 
+number of generations $i$ and population size $M$ (usually $z = 0.99$).
+
+As is described in \cite{koza92},
+The number $R(M,i,z)$ of independent runs required to yield a correct solution 
+is computed as
+
+$$ R(M,i,z) = \lceil\frac{\log(1-z)}{\log(1-P(M,i))}\rceil ,$$
+
+thus $I(M,i,z)$ is computed as
+
+$$I(M,i,z) = (i+1) \cdot M \cdot R(M,i,z) .$$\\
+
+It scored 17/50 (34\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 23 
+with 192,000 individuals to be processed.
+The experiment took 46 minutes.\\
+
+Performance curves are following.
+
+\includegraphics[scale=0.65]{reports/SSR/1/probabs.png}
+
+\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}\\
+
+\Lets compare it with results from \cite{koza92} (based on 113 runs):
+
+Success rate 35\%; minimal $I(M,i,z)$ in generation 24 
+with 162,500 individuals to be processed.
+
+We have 34\%; minimal $I(M,i,z)$ in generation 23 
+with 192,000. This seems like a pretty good match.
+\\
+
+Second experiment was performed to compare 
+\textit{ramped half-and-half strategy} with our
+\textit{geometric strategy} (with default parameter $q=0.75$). 
+Thus all control parameters stay same except for the 
+search strategy for generating algorithm.
+
+It scored 21/50 (42\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 29 
+with 150,000 individuals to be processed.
+The experiment took 26 minutes.\\
+
+\includegraphics[scale=0.65]{reports/SSR/2/probabs.png}
+
+\includegraphics[scale=0.65]{reports/SSR/2/indivs.png}\\
+
+So our geometric strategy is slightly more successful then 
+ramped half-and-half strategy. But more interesting is that
+it made the whole experiment almost twice as fast; 
+reason of this will be discussed in more detail in the next problem.
+
+\newpage
+\section{Artificial Ant}
+
+\textit{Artificial Ant} is another problem described
+in \cite{koza92}. Objective of this problem is to 
+find a control program for an artificial ant so
+that it can find all food located on "Santa Fe" trail.
+
+The Santa Fe trail lies on toroidal square grid.
+The ant is in the upper left corner, facing right.
+It has following layout.
+
+\includegraphics[scale=1]{santafe.png}
+
+The ant is able to move forward, turn left, and sense if a food 
+piece is ahead of him.\\
+
+Desired type of generated programs $\sigma$ and 
+building blocks context $\Gamma$ are following.
+\begin{align*}
+\sigma = An&tAction\\
+\Gamma = \{~~
+  l    &: AntAction                              ,\\
+  r    &: AntAction                              ,\\
+  m    &: AntAction                              ,\\
+  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
+  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
+  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
+\end{align*}
+
+Action $l$ turns the ant left.
+
+Action $r$ turns the ant right.
+
+Action $m$ moves the ant forward.
+
+Action $ifa x y$ (if-food-ahead) performs action $x$ if a food piece is ahead of the ant,
+otherwise it performs action $y$.
+
+Action $p2 x y$ first performs action $x$ and after it action $y$.
+
+Action $p3 x y z$ first performs action $x$, after that action $y$ and finally $z$.\\
+
+Actions $l, r$ and $m$ each take one time step to execute.
+
+Ants action is performed over and over again until it reaches predefined
+maximal number of steps. In \cite{koza92} this limit is set to be 400 time steps.
+
+Fitness value is equal to number of eaten food pieces.
+
+An individual such that eats all 89 pieces of food is 
+considered as a correct solution.\\
+
+There is also mentioned following ant program which is described as correct solution.
+
+\texttt{~\\
+(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))\\
+}
+
+But here arises a problem: In our implementation this program needs 545
+time steps; if it is given only 400 time steps, then it eats only 79 pieces
+of food. It is uncertain whether it is caused by my mistake in implementation
+or by inaccuracy in \cite{koza92}. Since everything else seems working just 
+fine, I have decided to set the limit to 600 in my experiments.
+
+\subsection{Experiments}
+
+Due to those uncertainties, we will not perform comparing experiments 
+as we did in the problem of simple symbolic regression.\\
+
+In the following experiments, the option of preserving the best individual 
+into the next generation is active (since we are not trying to compare 
+with results in \cite{koza92}).
+
+As usually, first experiment consisted of 50 independent runs of GP algorithm.
+Each run had maximally 50 generations and 500 individuals as population size.
+\textit{Ramped half-and-half strategy} is used as search strategy for the first 
+experiment.\\
+
+It scored 7/50 (14\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 28 
+with 449,500 individuals to be processed.
+The experiment took 265 minutes.
+
+Performance curves are following.
+
+\includegraphics[scale=0.65]{reports/Ant/2/probabs.png}
+
+\includegraphics[scale=0.65]{reports/Ant/2/indivs.png}\\
+
+Similarly as in previous problem, second experiment was performed to compare 
+\textit{ramped half-and-half strategy} with our
+\textit{geometric strategy} (with default parameter $q=0.75$). 
+Thus all control parameters stay same except for the 
+search strategy for generating algorithm.
+
+It scored 19/50 (38\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 10 
+with 115,500 individuals to be processed.
+The experiment took 107 minutes.\\
+
+\includegraphics[scale=0.65]{reports/Ant/1/probabs.png}
+
+\includegraphics[scale=0.65]{reports/Ant/1/indivs.png}\\
+
+This is big improvement in all three watched factors:
+
+\begin{enumerate}
+ \item 38\% vs 14\% success rate.
+ \item 115,500 vs 449,500 individuals needed to be processed.
+ \item 107 vs 265 minutes.  
+\end{enumerate}
+
+Third factor, the time difference might give us clue about what is going on.
+
+Those two experiments differs only in used term generating method.
+During one run the vast majority of the time is spend by evolution,
+the generation process only slightly prolongs the time spend in generation 0.
+The run time of the evolution is influenced by two important factors:
+Number of generations and time that crossover takes.
+
+The number of generations of more successful method is surly smaller,
+but it seems improbable that it would have such a big impact 
+(this is idea is supported by results of experiments with
+even-parity problem, where times are almost the same but success 
+rates differs significantly).
+
+Time of crossover is proportionate to number of nodes that 
+an individual has. Thus it seems that observed slowdown is
+due to terms being too big. \\
+
+In order to test this hypothesis, we run additional experiments
+to get statistics about dynamics of individuals' sizes.
+
+Following graphs are base upon 10 run experiments.\\
+
+Ramped half-and-half strategy:
+
+\includegraphics[scale=0.65]{reports/Ant/term-sizes-ANT-10behu-ramped.png}
+
+Our geometric strategy:
+
+\includegraphics[scale=0.65]{reports/Ant/term-sizes-ANT-10behu-geom.png}
+
+Those graphs show enormous differences. 
+The graph of Ramped half-and-half strategy evidences
+occurence of well known problem of GP called \textit{bloat}.
+Bloat is defined as \textit{program growth without (significant) return in
+terms of fitness} \cite{fg}.
+
+As is further stated in \cite{fg}, bloat has significant practical effects.
+
+~\begin{enum}
+ & Large programs are computationally expensive to evolve and run.
+ & Large programs can be hard to interpret.
+ & Large programs may exhibit poor generalisation. 
+\end{enum}~
+
+For these reasons bloat has been a subject of intense study in GP.
+
+On the contrary, our strategy seems to
+be successfully fighting with bloat  in this particular problem.
+
+It is also noticeable from the correct solutions that have been found.
+
+Here follow all 7 correct solutions found using ramped half-and-half strategy.\\
+
+\begin{easylist}[enumerate]
+& \texttt{p2 (ifa (p2 l r) (ifa (ifa (p2 m m) (p2 m l)) (p2 (ifa l r) (p2 m r)))) (ifa (i
+fa (p3 m (ifa m m) l) (p2 r r)) (ifa (ifa (p3 (p2 m l) (ifa (p3 (p2 m (ifa l m))
+ (p3 (ifa r (p2 (ifa (p2 m r) (p3 l r m)) (p2 m r))) m l) (p2 m l)) (ifa (p3 r l
+ l) (p3 m m r))) (ifa (p3 l l r) (p2 (ifa l r) (p2 m r)))) (p2 l r)) (ifa (p3 m
+l l) (p2 r r))))}
+& \texttt{p3 (ifa (p3 (p3 m m r) (p3 (ifa (p3 (p3 m m r) (p3 r l l) (ifa m l)) (ifa (p3 l
+ r (ifa (ifa m m) (p3 r l (p2 (ifa m l) (ifa (p3 (p2 l (p3 (ifa m m) (p2 l m) (p2 l r)))
+  (p2 (ifa (ifa m r) (p3 r m l)) (p3 (ifa l l) (p3 l r l) (p2 l l))) (p2
+(ifa (ifa m l) (p2 l r)) (p2 (p2 l m) (p3 l r m)))) (ifa (ifa (p3 (p3 r m m) (p3
+ (p2 l m) (p2 l m) (ifa m m)) (ifa l m)) (ifa (p3 l r m) (ifa (ifa r r) (p2 m l)
+))) l)))))) (ifa l l))) m (p2 (p2 (ifa l l) (ifa m r)) r)) (ifa m l)) (ifa (p3 l
+ r (ifa (ifa m m) (p3 r l (p2 (ifa m l) (ifa (p3 (p2 l (p3 (ifa m m) (p2 l m) (p2 l r)))
+  (p2 (ifa (ifa m r) (p3 r m l)) (p3 (ifa l l) (p3 l r l) (p2 l l))) (p2
+(ifa r (p2 l r)) (p2 (p2 l m) (p3 l r m)))) (ifa (ifa (p3 (p3 r m m) (p3 (p2 l m
+) (p2 l m) (p2 l r)) (ifa l m)) (ifa (p3 l r m) (ifa (ifa r r) (p2 m l)))) l))))
+)) (ifa l l))) m (p2 (p2 (ifa l l) (ifa m r)) r)}
+& \texttt{p2 (ifa (p3 m m m) (p2 r m)) (p3 (ifa l r) (ifa (ifa l (p3 (p3 r r l) (ifa l m)
+ (ifa m (ifa (p2 (p3 (p3 (p3 (ifa m l) (p2 r r) (ifa m m)) (p3 (p3 r r r) (ifa r
+ r) (p2 m l)) (p3 (p2 m m) (p3 r m r) (ifa m r))) (p2 (p2 (p3 r l m) (p2 m m)) (
+p3 (ifa r l) (p2 m l) (p2 r m))) (p3 (ifa (ifa r l) (p3 m m r)) (p3 (ifa r m) (p3 l l r)
+ (p2 r r)) l)) (p2 (p2 (p2 (p2 r r) (p2 r l)) (p3 (ifa r r) (p2 l l) (p2
+ m l))) (p3 m r l))) (p2 (ifa (ifa (p3 (ifa m r) (p2 (p3 (p3 r r l) (ifa l m) (p3
+  m l r)) (p2 m r)) (p3 l l l)) (ifa (p3 l l m) (ifa l r))) (ifa (ifa (ifa m r)
+(ifa l l)) (p2 (ifa l m) (p2 l m)))) (p3 (ifa (p2 (p2 r m) (p3 m l m)) (ifa (p2
+r l) (p3 r m r))) (p2 (p2 (p2 r r) (ifa l r)) (p3 (p3 m m m) (ifa m r) (ifa l m)
+)) (p2 m l))))))) r) (ifa l r))}
+& \texttt{p2 (p3 (p2 m l) (ifa r l) (ifa m l)) (ifa (p2 m m) (ifa r l))}
+& \texttt{p2 (p2 m r) (ifa m (p2 (p2 r r) (ifa (ifa (ifa m r) (p3 m m l)) r)))}
+& \texttt{ifa (p2 (p2 m m) (p2 m l)) (p3 (p2 r r) (ifa m l) (p2 m l))}
+& \texttt{p2 (ifa (p2 (ifa r m) (p2 (p2 (p2 l m) (p3 m m r)) (ifa l l))) (p3 (p2 l l) 
+(ifa m l) m)) l}
+\end{easylist}
+
+\newpage	
+Compare it with 18 solutions obtained by using geometric strategy 
+(I accidentally lost one solution).\\
+
+\begin{easylist}[enumerate]
+& \texttt{ifa m (p3 r (ifa (ifa m (ifa m (p3 r (ifa m r) m))) (ifa m (p2 (p2 m (ifa m (ifa
+m (ifa m (p3 r (ifa m r) m))))) (ifa m r)))) m)}
+& \texttt{p3 (ifa m l) (p3 m l (ifa (p2 m (ifa l (p2 (p2 (p2 m (ifa m m)) (ifa (ifa m l)
+r)) (ifa m l)))) r)) r}
+& \texttt{ifa l (ifa (p3 (p2 (p2 r (ifa r m)) (p3 l (ifa m (ifa m l)) r)) l m) (p2 r (ifa
+ m (p2 l (p2 (p2 l (ifa m r)) m)))))}
+& \texttt{p3 m (p3 r (ifa m l) l) (ifa (p2 (p3 m m r) (ifa m l)) r)}
+& \texttt{p3 (ifa m (ifa (ifa m m) (ifa m r))) m (p3 r (ifa (p3 (ifa m r) m (p3 r (ifa m
+l) m)) l) l)}
+& \texttt{p3 (ifa m l) (p2 m l) (ifa (p3 (ifa m (ifa m l)) m r) (p2 (ifa m (ifa r l)) l))}
+& \texttt{p3 (p3 r (ifa (p2 (ifa m (ifa (ifa m r) m)) m) l) m) l (ifa m r)}
+& \texttt{ifa (p3 m (ifa (ifa r m) r) l) (p3 (p3 l (ifa m r) r) (ifa m (ifa m l)) m)}
+& \texttt{p3 m (p3 l (ifa (p3 m (ifa m m) m) r) r) (ifa m (ifa m l))}
+& \texttt{p2 (ifa (ifa m (p3 l (p2 (p2 (ifa (ifa m l) l) m) (ifa m r)) (p2 (ifa (ifa (ifa
+ (ifa (p3 r l m) r) l) l) (p3 l (ifa m r) (ifa m r))) m))) r) (p2 (p3 r (ifa (if
+a (p2 (ifa m (ifa (ifa (ifa (p3 m r (ifa l l)) l) l) l)) m) (p3 l (p2 (ifa m (p2
+ l m)) (ifa m r)) m)) l) m) l)}
+& \texttt{ifa (ifa m (ifa m r)) (ifa (ifa (ifa m l) r) (p3 (p3 l (ifa m r) (p3 r l r)) (
+ifa m l) m))}
+& \texttt{ifa (p2 m r) (ifa (p2 (ifa (p2 l m) r) m) (p3 l m (ifa r (p3 l (ifa m r) r))))}
+& \texttt{p3 (p3 l (ifa m r) r) (ifa (p3 m m (p2 r (ifa (p3 r (ifa m l) (ifa m l)) l))) l
+) m}
+& \texttt{p3 r (ifa m (p3 l l (ifa (ifa m m) r))) m,89.0,Just 4)
+(ifa (p3 (ifa m l) r (ifa m l)) (ifa m (p3 (p3 l (ifa m r) m) r (ifa m l)))}
+& \texttt{ifa m (p3 (ifa l (p3 r (ifa (ifa l l) (ifa (ifa r r) r)) (ifa r r))) (ifa (ifa
+m r) r) m)}
+& \texttt{p3 (ifa (ifa (ifa m l) l) l) (p3 m (ifa (ifa (ifa (p2 m l) l) (ifa r r)) l) (
+ifa m (ifa m r))) r}
+& \texttt{ifa (ifa (ifa m r) r) (p3 r (ifa (ifa m (p3 (ifa m (ifa m r)) (ifa (ifa r l) r)
+ (ifa m l))) (ifa l (p3 (ifa m l) (ifa m l) (ifa m (ifa m r))))) m)}
+& \texttt{p3 (p3 m r (ifa (p2 m m) l)) l (ifa m r)}
+\end{easylist}~
+
+\newpage
+
+\Lets also examine average term sizes of the Simple Simbolic Regression problem,
+again based on 10 runs.\\
+
+Ramped half-and-half strategy:
+
+\includegraphics[scale=0.65]{reports/SSR/term-sizes-SSR-10behu-ramped.png}
+
+Our geometric strategy:
+
+\includegraphics[scale=0.65]{reports/SSR/term-sizes-SSR-10behu-geom.png}
+
+In this case the difference is not that radical, but still we can see
+that runs of SSR using ramped half-and-half tend to have terms approximately twice
+as big as runs using geometric strategy.
+
+\newpage	
+\section{Even-parity problem}
+The even-parity function is a boolean function taking as inputs $N$
+boolean values and returning \textit{True} if an even number of inputs 
+are \textit{True}. For odd number it returns \textit{False}.
+
+The reason why examine this problem is that it has been used by many researchers
+as a benchmark for GP \cite{todo,yu01}.
+
+We compare our results with that obtained by Yu in \cite{yu01},
+where is presented approach to evolve recursive and modular programs
+by use of higher-order functions and $\lambda$-abstractions.
+Since the results are highly sensitive to set of \textit{building blocks}, 
+we use very similar set of \textit{building blocks} in order to make 
+the results comparable.
+
+\begin{align*}
+\sigma = [Boo&l] \ar Bool\\
+\Gamma = \{
+  and   &: Bool \ar Bool \ar Bool                              ,\\
+  or    &: Bool \ar Bool \ar Bool                              ,\\
+  nand  &: Bool \ar Bool \ar Bool                              ,\\
+  nor   &: Bool \ar Bool \ar Bool                              ,\\
+  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
+  head' &: [Bool] \ar Bool                                   ,\\
+  tail' &: [Bool] \ar [Bool]                              \}\\
+\end{align*}
+
+The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
+this problem is considered atomic.
+
+Unlike in \cite{yu01}, we use specific instance of polymorphic 
+function \texttt{foldr}. 
+And modifications of functions \texttt{head} 
+(returning the first element of the list) 
+and \texttt{tail} (returning the list without the first element) are used; 
+making them total by returning default value \textit{False}
+and \texttt{[]}, respectively.
+
+The \texttt{fold} function is higher order function widely used in \textit{Haskell}
+programming language, recursively defined as follows.
+
+\texttt{~\\
+foldr :: (a -> b -> b) -> b -> [a] -> b \\
+foldr f z []     = z \\
+foldr f z (x:xs) = f x (foldr f z xs)  
+}\\
+
+\Lets demonstrate \texttt{foldr}'s behavior on the following simple example.
+
+\texttt{~\\
+foldr xor False [True,False,True] \\
+= xor True (foldr False [False,True]) \\
+= xor True ( xor False ( foldr False [True] ) )\\
+= xor True ( xor False ( xor True (foldr False [] ) )\\
+= xor True ( xor False ( xor True False) )\\
+= False
+}\\
+
+The reasons why is this difficult problem 
+for GP to solve are following \cite{todo}:
+
+\begin{itemize}
+ \item Since a single change of one input generates a different output,
+       the output is highly sensitive to all inputs.
+ \item The building blocks set does not contain functions \textit{xor}
+       or \textit{eq}, which are keys to solving the problem.
+       But those two functions may be discover by GP.  
+\end{itemize}
+
+We use the same fitness function as in \cite{yu01}. 
+The fitness function examines the individual by giving
+it all possible boolean lists of length 2 and 3.
+Thus resulting in $2^2 + 2^3 = 12$ test cases.
+For each correct response it receives $1/12$, 
+therefore correct solution receives fitness value $1$.\\
+
 \begin{todo}
-   & rozdíli při křížení ruznejch reprezentací 
-      && \atTree umožňuje víc křížení než \sexprTree
-          &&& každý co de v \sexprTree umí i \atTree ale obráceně to neplatí
-      && \atTree je zase ekvivalentní tomu kozovskýmu  
+ & zmínit tu, že to pri týyle building block stačí na general solution...
 \end{todo}
 
-\subsection{ Repairing method }
-\label{repairing-method}
+\subsection{Experiments}
+
+Series of five experiments have been performed to test properties of our system.
+Each experiment consisted of 50 independent runs of GP algorithm.
+Each run had maximally 50 generations
+and 500 individuals as population size.\\
+
+In those experiments we examined influences of the following factors:
+
+\begin{itemize}
+ \item Option of preserving the best individual into the next population.
+ \item Option of performing $\eta$-normalization of generated individuals 
+      (conversion from \lnf to \benf, see \ref{benefits}).
+ \item Representation of individuals: \sexprTree vs \atTree.
+ \item Option of performing optimized abstraction elimination
+       (see \ref{toSKI}) instead of the simple one.
+\end{itemize}
+
+The setup of the first experiment was following:\\
+
+\begin{enum}
+ & \textit{Geometric} search strategy for generating individuals with default $q=0.75$.
+   This option is unchanged for all five experiments.
+ & No best individual preservation.
+ & $\eta$-normalization was not performed.
+ & Individuals represented as \sexprTree{}s. 
+ & Simple abstraction elimination.
+\end{enum}~
+
+Thus the simplest one setup; all extra features are off. 
+In each following experiment one extra feature will be turned on,
+in hope that it will enhance the performance.
+
+The correct individual was yield 
+in 17 out of 50 runs, i.e., 34\% success rate.
+The performance curve for $P(M,i)$ is following.
+
+\includegraphics[scale=0.65]{reports/ep/report1/probabs.png}
+
+Interestingly, in 2 runs out of 50, a correct solution was present in the initial population.
+
+The performance curve for $I(M,i,z)$ is following.
+
+\includegraphics[scale=0.65]{reports/ep/report1/indivs.png}
+
+The smallest value in this graph is used to indicate the minimum effort
+(i.e. number of individuals to be processed) for GP to yield a correct solution.
+
+Thus for this experiment this analysis suggest
+to perform only term generating and no evolution at all, 
+because the smallest value 56500 is in generation 0.
+
+This experiment took 59 minutes (on average desktop computer).\\
+
+Second experiment was different from the first one in that it 
+preserved the best individual to the next generation.
+
+It scored 22/50, i.e., 40\% success rate. A correct solution was
+created once in the initial generation and again it was
+sufficient to make the value in generation 0
+be the smallest value in the $I(M,i,z)$ curve; now 114000
+ individuals to be processed.
+This experiment also took 59 minutes.\\
+
+Third experiment was different from the second one in that 
+$\eta$-normalization of generated individuals was performed.
+
+It scored 27/50, i.e., 54\% success rate. 
+Two runs contained a correct solution in the initial population.
+But now the $I(M,i,z)$ was best for generation 1 with 56000
+ individuals to be processed.
+This experiment took 53 minutes.\\
+
+Fourth experiment was different from the third one in that
+\atTree representation was used instead of \sexprTree.
+
+It scored 34/50, i.e., 68\% success rate. Again two runs
+contained a correct solution in the initial population.
+The $I(M,i,z)$ was best for generation 0 with 56500 
+individuals to be processed.
+This experiment took 64 minutes.
+
+As we will see, this setup was the most successful one,
+thus we again show the performance curves.
+
+\includegraphics[scale=0.65]{reports/ep/report4/probabs.png}
+
+\includegraphics[scale=0.65]{reports/ep/report4/indivs.png}
+
+Fifth, last experiment was different from the fourth one in that
+optimized abstraction elimination was used instead of the simple one.
+
+It scored 33/50, i.e., 66\% success rate. One run
+contained a correct solution in the initial population.
+The $I(M,i,z)$ was best for generation 6 with 59500 
+individuals to be processed.
+This experiment took only 30 minutes. \\
+
+\subsection{Comparison with other results}
+
+Our results are less successful then those 
+presented by Yu \cite{yu01};
+they scored 40/50, i.e., 80\% success rate.  
+Her $I(M,i,z)$ was best for generation 4 with 17500 
+individuals to be processed.
+(We scored 34/50 (68\%); min $I(M,i,z) = 56,500$.)
+
+Yu also uses crossover as the only genetic operator, thus we
+can narrow the discussion to discussion of generating method and 
+crossover method.
+
+When we observe results of our experiments, it seems
+that our system is pretty good in term generation, but not so good 
+in crossover. Since the main focus point of this thesis is
+the term generation, it is positive rather then negative observation.
+
+Design of our crossover comes from that we wanted it to be
+very simple generalization of standard tree-swapping crossover.
+Whereas the crossover used by Yu is more structure oriented
+--- as is stated in \cite{yu01} the crossover \textit{can only operate 
+between nodes in two main programs or between nodes in the same
+kind of $\lambda$ abstraction (i.e., $\lambda$ abstractions that 
+represent the same function arguments to the same higher order function).}
+%This seems to be great advantage for optimizing the
+
+Since our method has better success rate for generation 0, it would 
+be interesting to try combination of our generating 
+method with crossover used by Yu.
+
+From an optimist's point of view 34/50 versus 40/50 
+is not such a huge difference when we take to account that
+our general tree-swapping crossover is competing with specialized one.
+
+Positive observation is that our result at least outperforms 
+other results mentioned in \cite{yu01} with which Yu's method is compared.
+
+\textit{Generic genetic programming} scored 17/60 (28\%); 
+min $I(M,i,z) = 220,000$.
+
+\textit{GP with ADFs} scored 10/29 (34\%); 
+min $I(M,i,z) = 1,440,000$.
 
 \begin{todo}
-& !!!
+& napsat tam jak vypadaj nějaký nalezeneý řešení
 \end{todo}
 
-\subsection{ Comparison of the two methods }
-\label{comarison-ski-repairing}
+\section{Flies and Apples}
+
+\textit{Flies and Apples} is problem consisting in breeding 
+control programs for a fly agent in a simple world.\\
+
+Whereas previous problem is a kind of problem where we know precisely
+what problem we are solving and where we are more interested in 
+performance statistics than in the solutions since we know the correct solution in 
+advance, problem we are going to present now is of more playful nature.
+In this problem we are not interested in numerical results, instead 
+we are hoping for interesting \textit{behavior} to emerge.
+
+We start with defining a simple simulation and goal for agents.
+Then we prepare first testing simulation world. 
+After that we write some agent programs by hand. After that we 
+identify some useful recurring 
+themes occurring in those programs and pack them into functions
+which will become building blocks. During this process are also identified
+useful sensory data to be prepared for the agent program.
+After that we run the evolution and
+iteratively continue in this process  --- we may add more testing worlds,
+add more functions or change them, etc.
+
+Since we do not need to satisfy the \textit{closure} constraint, 
+it is much easier to perform such a "human-computer jam" than it
+would be in standard GP. And it is further simplified by use of
+functional language (such as \textit{Haskell}) where almost
+every construct can be expressed as function (or value).
+
+Another advantage of typed GP over standard GP is that 
+it is more capable of handling huge set of building blocks,
+since the search space can be significantly smaller due to 
+fact that types must match.
+\\
+
+Fly world is a square grid. 
+Each cell is ether empty or contains precisely one object.
+There are three kinds of objects:
+
+~\begin{enum}
+ & Wall
+ & Apple
+ & Fly
+\end{enum}~
+
+Apples and flies have \textit{energy}.
+
+Moreover, each fly has its control program and inner state, 
+so called \textit{registers}.
+
+There is a queue of all flies in the fly world determining which 
+fly is next to make a move.
+After the fly in the front of the queue performs its move, 
+it is put to the back. In other words, flies are cyclically
+taking turns similarly like in common card games.\\
+
+A control program of a fly has as input a collection of various sensory information
+and returns as output agents move and agents registers for the next turn.
+
+Fly can perform two kinds of moves:
+
+~\begin{enum}
+ & \textit{Trave.} Moves the fly to one of four adjacent cells determined by 
+   direction specified in the move.  
+   Result of the travel move depends on the content of the target cell.    
+   && If it is empty, then the fly simply moves to that cell.
+   && If it contains wall, then no movement is performed.
+   && If it contains an apple, then the fly moves to that cell, the apple 
+      is eaten and its energy is added to energy of the fly.  
+   && If it contains another fly, then the fly moves to that cell, the fly with bigger 
+      energy stays alive and the weaker one is eaten
+      and its energy is added to energy of the stronger fly.  
+ & \textit{Split.} Puts a daughter fly to one of four adjacent cells.
+    Together with the direction, amount of energy and new registers 
+    for the fly are specified by this move. The mother gives this energy to 
+    its child. In order to be successful, the target cell must be empty
+    and the mother fly must have energy greater then 1. 
+\end{enum}~
+
+Agent's fitness is computed as sum of fitness cases; each 
+fitness case consists of one fly world and fitness in a
+fly world is computed as sum of agent's energy and of energy of all his descendants
+(daughters, granddaughter, ...) after predefined number of rounds.
+
+\Lets look in more detail on input for fly. It is collection of 
+various useful information. Its components are:
+
+~\begin{enum}
+ & Current value of fly's Registers. 
+ & Energy of the fly.
+ & Direction of the last successful travel move.
+ & Boolean value indicating whether the last move was successful.
+ & Direction of the nearest apple.
+ & Distance to the nearest apple.
+ & Energy of the nearest apple.
+ & Direction of the nearest fly.
+ & Distance to the nearest fly.
+ & Energy of the nearest fly.
+ & Direction of the "center of gravity" of all apples.
+ & Distance to the "center of gravity" of all apples.
+\end{enum} ~
+
+Set of building blocks is following.
+
+\begin{align*}
+\sigma = Input &\ar Output\\
+\Gamma = \{
+dUp              &: Direction                                       ,\\              
+dDown            &: Direction                                       ,\\              
+dLeft            &: Direction                                       ,\\              
+dRight           &: Direction                                       ,\\            
+output           &: Move \ar Registers \ar Output                   ,\\            
+travel           &: Direction \ar Move                              ,\\          
+split            &: Direction \ar Int \ar Registers \ar Move        ,\\       
+easySplit        &: Input \ar Move                                  ,\\     
+myEnergy         &: Input \ar Int                                   ,\\     
+myLastTravel     &: Input \ar Direction                             ,\\                
+myWasSuccess     &: Input \ar Bool                                  ,\\      
+nAppleDir        &: Input \ar Direction                             ,\\          
+nAppleDir        &: Input \ar Direction                             ,\\                    
+nAppleDist       &: Input \ar Distance                              ,\\                   
+nAppleEnergy     &: Input \ar Int                                   ,\\              
+nFlyDir          &: Input \ar Direction                             ,\\                      
+nFlyDist         &: Input \ar Distance                              ,\\                    
+nFlyEnergy       &: Input \ar Int                                   ,\\              
+\end{align*}
+\begin{align*}
+cAppleDir        &: Input \ar Direction                             ,\\          
+cAppleDist       &: Input \ar Distance                              ,\\        
+myRegs           &: Input \ar Registers                             ,\\         
+xGet             &: Input \ar Int                                   ,\\     
+yGet             &: Input \ar Int                                   ,\\     
+zGet             &: Input \ar Int                                   ,\\     
+dGet             &: Input \ar Direction                             ,\\          
+xSet             &: Int \ar Registers \ar Registers                 ,\\          
+ySet             &: Int \ar Registers \ar Registers                 ,\\          
+zSet             &: Int \ar Registers \ar Registers                 ,\\          
+dSet             &: Direction \ar Registers \ar Registers           ,\\                
+xInc             &: Registers \ar Registers                         ,\\              
+yInc             &: Registers \ar Registers                         ,\\              
+zInc             &: Registers \ar Registers                         ,\\              
+rotCW            &: Direction \ar Direction                         ,\\                                                             
+(==)             &: Int \ar Int \ar Bool                            ,\\                                      
+(<=)             &: Int \ar Int \ar Bool                            ,\\               
+if'              &: Bool \ar Output \ar Output \ar Output           ,\\                                  
+if'              &: Bool \ar Move \ar Move \ar Move                 ,\\                                  
+if'              &: Bool \ar Direction \ar Direction \ar Direction  ,\\              
+if'              &: Bool \ar Int \ar Int \ar Int                    ,\\     
+if'              &: Bool \ar Distance \ar Distance \ar Distance     ,\\        
+if'              &: Bool \ar Registers \ar Registers \ar Registers  ,\\                           
+0                &: Int                                             ,\\                                     
+1                &: Int                                             ,\\                                    
+2                &: Int                                             \}\\
+\end{align*}
+
+We will not discuss it in great detail, only briefly: First four 
+elements are four direction values. Next four functions are
+constructors for $Output$ and $Move$ type. Next thirteen functions
+are for accessing components of $Input$. Next eleven functions are 
+related to $Registers$. The $rotCW$ performs clock-wise rotation of a direction.
+Next two functions are comparison functions. Next six are various \textit{if}s. And last 
+three are integer constants.
+
+~\\\\\textbf{todo DODĚLAT ASI AŽ POTOM, AŤ SE S TIM MOC NEZDRŽUJU}\\\\
+
+Tedy celkem 44 prvků množiny stavebních bloků, tedy poměrně obrovská množina
+na GP poměry.
+
+Takže množina atomických typů je následující:
+Move Bool Registers Input Direction Int Output Distance
+
+Používáme tři levely. První level vypadá takle
+
+[obrázek]
+
+Další dva levely jsou reakce na typické chování mouchy 
+která používala pro předchozí level triviální strategii
+jdi furt za nejbližšim jablkemn. roto jsou následující dva levely 
+léčka na tuto strategii.
+
+V druhém levelu se zmíněná moucha nachytá a vleze do 
+chodby kde je mnohem mín jablek než v v tom mega jablku z jakblek.
+
+[obrázek]
+
+Třetí dovádí předchozí myšlenku do extrému. 
+Moucha se zasekne a nesebere žádné jablko.
+
+V počátečních pokusech byl problém s tím, že mochy s počáteční energii 1
+neměli moc tendenci se množit. Proto byla simulace předělána tak, aby
+že nyní je počáteční energie 100. To mělo na Mouchy blahodárný vliv.
+Ale i pro mouchy sw počátečni energii 1 vznikli některé zajímavé strategie
+
+[povídání o zajímavých strategiích pro poč en 1]
+
+Po navíšení poč energie vznikli tyto zajímavá chování
+
+[povídání o zajímavých strat por poč en 100]\\
+
+Aplikace: například v oboru počítačových her, kde nám často jde o
+zajímavé nelineární chování by se tento přístup k využití typovaného funkcionálního
+programování dal využít.
+
+Předvedený postup by taky ohlo být možné použít na
+
+\section{Results summary}
+
+\chapter{Previous related work}
+
+\section{Evolution of Constrained Syntactic Structures }
+\textit{Evolution of Constrained Syntactic Structures}
+is mechanism proposed by Koza in \cite{koza92} for relaxing 
+the \textit{closure} requirement.
+
+Problem specific \textit{rules of construction}
+are used to constrain which function node
+can have which child node.
+
+Those rules are used in generating method similar to the standard one.
+
+For purposes of crossover types of symbols are listed - 
+e.g. in the problem of \textit{symbolic multiple regression} 
+(and various other problems)
+those types are
+
+~\begin{easylist}[itemize]
+& the root, and
+& non root points.
+\end{easylist}~
+
+Crossover is performed by first selecting any node in the first
+parent and after that by selecting node of the same type as the first node.\\
+
+System based on types do this task more indirectly and in a
+simple systematic manner.
+
+\section{Strongly Typed Genetic Programming}
+One of first such type based systems was that described by Montana
+in \cite{montana95}. This system is called 
+Strongly Typed Genetic Programming (STGP).
+Beside simple types he also uses kind of polymorphism 
+called \textit{generic functions} based on table lookup.
+
+Used generating method is generalization of the standard one
+dealing with type restrictions and . Moreover, checking preventing
+selection of node which makes it impossible to complete the subtree 
+is performed.
+
+The crossover works in similar manner as that Koza used
+in his Constrained Syntactic Structures. Such a crossover
+is less cautious then that used in our system
+since it may fail to find a node with desired type in the second parent.
+If the crossover fails, then the crossover returns either parents or nothing.
+
+We were briefly measuring fail rates for such a crossover and
+it was somewhere around 15\% for even-parity problem. Thus we prefer
+to perform the more cautious version described in \ref{typed-swapping}.  \\
+
+This system does not involve use of lambda abstractions or
+higher-order functions.
+
+\section{Higher-Order Functions and Lambda Abstractions} 
+In \cite{yu01} Yu presents a GP system utilizing
+polymorphic higher-order functions 
+\footnote{Higher-order function is a function taking another function as 
+input parameter.} and lambda abstractions.
+
+Important point of interest in this work is use of
+\texttt{foldr} function as a tool for \textit{implicit recursion},
+i.e. recursion without explicit recursive calls.
+
+Approach described in this work is more specialized then our approach
+in the following ways.
+
+\begin{itemize}
+\item The terminal set for constructing lambda abstraction subtrees 
+      is limited to use only constants and variables of that particular
+      lambda abstraction, i.e., outer variables are not allowed to be used
+      as terminals in this work.
+\item  Whereas we see lambda head of lambda function as 
+another kind of node (as parent node of the body of the function), 
+in this work whole lambda function is considered more as single
+point with internal structure. Higher-order functions
+and its lambda abstractions are said to form \textit{two-layer-hierarchy}
+called in this work \textit{structure abstraction}.
+\item The crossover used in this work \textit{can only operate 
+between nodes in the two main programs or 
+between nodes in two lambda abstractions that represent 
+the same function arguments to the same higher order function. }
+\end{itemize}
+
+The difference between approach presented in this work and our approach 
+may be stated in the following way. Approach presented in \cite{yo01}
+tries to adopt some useful constructs from domain
+of functional programming and use them for a benefit in problem solving by GP.
+Whereas our approach is to perform genetic programming directly in the
+simply typed lambda calculus.
+
+% v even parity scoruje o něco lépe než my.
+
+\section{Functional Genetic Programming with Combinators}
+In \cite{kes} by Briggs and O’Neill is presented technique 
+utilizing GP with combinators in similar way as we do.
+
+The difference between approach presented in this work
+and our approach is that in this work terms are generated
+straight from \textit{library} (their term used for context) 
+of combinators, whereas we first generate $\lambda$-terms which are
+subsequently translated into combinator terms.
+
+Thus we can in our system see the combinator representation
+as option which can be turned off, whereas their system has this
+feature "hard-wired". On the other hand, their advantage is
+that direct generation of terms generates 
+shorter combinator terms than indirect generation.
+
+They are using more general type system then us
+--- the Hindley–Milner type system. So the system
+is polymorphic.
+
+They also discuss the properties of exhaustive enumeration
+of terms and compare it with GP search.  
+Since our successful term generating \textit{geometric} strategy
+is based on undisciplined exhaustive enumeration, this
+is another interesting thing that connects our two systems.
+
+Another really interesting concept is their \textit{Generalized
+genetic operator} based on term generation. Suppose we have
+some terms we want to operate on 
+(for one term it corresponds to mutation, for two (parent) 
+terms it is crossover, etc.). In order to do that we construct 
+\textit{library} of all subterms of all terms we want to operate
+on and from this library we generate new individuals by supplying
+it with desired type.
+
+It would be nice to use some ideas inspired from this work in
+the future improvements of our system.
+
+\section{Genetic Programming with Polymorphic Types and Higher-Order Functions}
+by Franck Binard and Amy Felty
 
 \begin{todo}
-& !!!
-& říct že nakonec sme přilnuli k zbavit se proměnejch
+ & dopsat nebo vyhodit
 \end{todo}
 
-\newpage
+%\chapter{Ideas yet not implemented}
+\chapter{Future work}
+
+\section{Polymorphic variant}
+
+\begin{todo}
+ & poznamky o čem se mimojiné zmínit:
+   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
+      jako alternativa k ADF ! Myslim že hodně zajimavý
+\end{todo}
+
 \section{Mutation}
@@
 
-%\subsection{Using term generation}
-
-\section{ Complexity discussion }
-
-\begin{todo}
-& !!!
-& zmínit že mi moc neřešíme formální/asymptotický složitosti
-  && de nám hlavně o realný časy
-& vyjít z komplexnosti A* asi
-  && složitost výrazně ovlivněná kvalitou heuristický funkce
-\end{todo}
-
-\chapter{ Implementation }
-
-Předběžný náhodně uspořádaný body osnovy ..\\
-
-\begin{easylist}[itemize]
-& že je to v haskellu
-&& Jak moc hrotit vysvětlení haskellu?
-& že je to servrově řešený
-& architektura evolučního algoritmu 
-&& ty tři typový třídy a ta čtvrtá typová třída
-& evaluace
-&& vychytavka s vyhodnocenim celýho seznamu aby s eto nemuselo furt inicializovat - inicializace vyhodnocovače jednou za generaci 
-& hlavní soubory
-&& Eva.hs     - Eva monada
-&& GP-Core.hs - Evoluční algoritmus
-&& GP-Data.hs - Instance tříd z GP-Core
-&& IMx.hs - Term generating
-&& TTerm.hs   - typed lambda term
-&& TTree.hs   - CTT
-&& Server.hs  - main serveru
-&& Job.hs     - Registrace problemu
-&& Heval.hs   - Evaluator 
-&& složka Problems
-&&& pro každej problém složka 
-&&& ta typicky strukturu
-&&&& Problem.hs
-&&&& Funs.hs
-&& Slozka server
-&&& index.html
-&&& slozka js/Problems
-&&& .js pro kazdej problem
-\end{easylist}
-
-\section{Top level view}
-\section{Comments about main source files}
- \subsubsection{ Eva.hs }
- \subsubsection{ GP\_{}Core.hs }
-\section{Term generating}
-\section{Crossover}
-\section{Mutation}
-
-\section{Brief history of the project}
-\begin{todo}
- & zmínit třeba tyhle fáze:
-  && že sem si delal evaluator sám co sem pak s performenc duvodu opustil
-  && že sem hledal intuicionistickej dokazovač ale že nejsem nenarazil na vhodnej
-     pač nejsou delany na to plivat hodně dukazu  
-  && že sem nejdriv pouzival ten vypsychlej dokazovač co tam měl nějaký ty 
-     spešl rozhodovací stromy
-  && pak sem to přepsal tak aby to nebylo jen na GP šitý ale obecně na jakoukoliv 
-     evoluci
-  && ...
-\end{todo}
-
-\chapter{Problems}
-	In this section will be presented usage of the system in order to solve specific problems. Several problems will be described and some of them compared with results
-of others.
-
-\section{Even-parity problem}
-The even-parity function is a boolean function taking as inputs $N$
-boolean values and returning \textit{True} if an even number of inputs 
-are \textit{True}. For odd number it returns \textit{False}.
-
-The reason why examine this problem is that it has been used by many researchers
-as a benchmark for GP \cite{todo,yu01}.
-
-We compare our results with that obtained by Yu in \cite{yu01},
-where is presented approach to evolve recursive and modular programs
-by use of higher-order functions and $\lambda$-abstractions.
-Since the results are highly sensitive to set of \textit{building blocks}, 
-we use very similar set of \textit{building blocks} in order to make 
-the results comparable.
-
-\begin{align*}
-\sigma = [Boo&l] \ar Bool\\
-\Gamma = \{
-  and   &: Bool \ar Bool \ar Bool                              ,\\
-  or    &: Bool \ar Bool \ar Bool                              ,\\
-  nand  &: Bool \ar Bool \ar Bool                              ,\\
-  nor   &: Bool \ar Bool \ar Bool                              ,\\
-  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
-  head' &: [Bool] \ar Bool                                   ,\\
-  tail' &: [Bool] \ar [Bool]                              \}\\
-\end{align*}
-
-The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
-this problem is considered atomic.
-
-Unlike in \cite{yu01}, we use specific instance of polymorphic 
-function \texttt{foldr}. 
-And modifications of functions \texttt{head} 
-(returning the first element of the list) 
-and \texttt{tail} (returning the list without the first element) are used; 
-making them total by returning default value \textit{False}
-and \texttt{[]}, respectively.
-
-The \texttt{fold} function is higher order function widely used in \textit{Haskell}
-programming language, recursively defined as follows.
-
-\texttt{~\\
-foldr :: (a -> b -> b) -> b -> [a] -> b \\
-foldr f z []     = z \\
-foldr f z (x:xs) = f x (foldr f z xs)  
-}\\
-
-\Lets demonstrate \texttt{foldr}'s behavior on the following simple example.
-
-\texttt{~\\
-foldr xor False [True,False,True] \\
-= xor True (foldr False [False,True]) \\
-= xor True ( xor False ( foldr False [True] ) )\\
-= xor True ( xor False ( xor True (foldr False [] ) )\\
-= xor True ( xor False ( xor True False) )\\
-= False
-}\\
-
-The reasons why is this difficult problem 
-for GP to solve are following \cite{todo}:
-
-\begin{itemize}
- \item Since a single change of one input generates a different output,
-       the output is highly sensitive to all inputs.
- \item The building blocks set does not contain functions \textit{xor}
-       or \textit{eq}, which are keys to solving the problem.
-       But those two functions may be discover by GP.  
-\end{itemize}
-
-We use the same fitness function as in \cite{yu01}. 
-The fitness function examines the individual by giving
-it all possible boolean lists of length 2 and 3.
-Thus resulting in $2^2 + 2^3 = 12$ test cases.
-For each correct response it receives $1/12$, 
-therefore correct solution receives fitness value $1$.\\
-
-\begin{todo}
- & zmínit tu, že to při týhle building block stačí na general solution...
-\end{todo}
-
-\subsection{Experiments}
-
-Series of five experiments have been performed to test properties of our system.
-Each experiment consisted of 50 independent runs of GP algorithm.
-Each run had maximally 50 generations
-(50 + 1 for generation 0) and 500 individuals as population size.\\
-
-In those experiments we examined influences of the following factors:
-
-\begin{itemize}
- \item Option of preserving the best individual into the next population.
- \item Option of performing $\eta$-normalization of generated individuals 
-      (conversion from \lnf to \benf, see \ref{benefits}).
- \item Representation of individuals: \sexprTree vs \atTree.
- \item Option of performing optimized abstraction elimination
-       (see \ref{toSKI}) instead of the simple one.
-\end{itemize}
-
-We analyze the ability of the system to produce a correct solution.
-We are interested in percentage of runs that led to discovery of
-correct solution. And we also use the popular measurement  
-method within GP field --- the \textit{performance curves}
-described in \cite{koza92}.
-
-First performance curve is the cumulative \textit{probability of success}
-$P(M,i)$, where $i$ is generation number ($i = 0$ for initial population) 
-and $M$ is the size of the population. 
-$P(M,i)$ is probability that a correct solution will
-be find in one run of GP with maximal number of generations $i$ and 
-population size $M$. $P(M,i)$ is computed as percentage of 
-runs yielding a correct solution in generation $i$ or earlier.
-
-From values of $P(M,i)$ are computed values of the second performance curve
-$I(M,i,z)$
---- the total number of individuals that must be processed to yield a correct
-solution with probability $z$ by multiple runs of GP with maximal 
-number of generations $i$ and population size $M$ (usually $z = 0.99$).
-
-As is described in \cite{koza92},
-The number $R(M,i,z)$ of independent runs required to yield a correct solution 
-is computed as
-
-$$ R(M,i,z) = \lceil\frac{\log(1-z)}{\log(1-P(M,i))}\rceil ,$$
-
-thus $I(M,i,z)$ is computed as
-
-$$I(M,i,z) = (i+1) \cdot M \cdot R(M,i,z) .$$\\
-
-The setup of the first experiment was following:\\
-
-\begin{enum}
- & \textit{Geometric} search strategy for generating individuals with default $q=0.75$.
-   This option is unchanged for all five experiments.
- & No best individual preservation.
- & $\eta$-normalization was not performed.
- & Individuals represented as \sexprTree{}s. 
- & Simple abstraction elimination.
-\end{enum}~
-
-Thus the simplest one setup; all extra features are off. 
-In each following experiment one extra feature will be turned on,
-in hope that it will enhance the performance.
-
-The correct individual was yield 
-in 17 out of 50 runs, i.e., 34\% success rate.
-The performance curve for $P(M,i)$ is following.
-
-\includegraphics[scale=0.65]{reports/ep/report1/probabs.png}
-
-Interestingly, in 2 runs out of 50, a correct solution was present in the initial population.
-
-The performance curve for $I(M,i,z)$ is following.
-
-\includegraphics[scale=0.65]{reports/ep/report1/indivs.png}
-
-The smallest value in this graph is used to indicate the minimum effort
-(i.e. number of individuals to be processed) for GP to yield a correct solution.
-
-Thus for this experiment this analysis suggest
-to perform only term generating and no evolution at all, 
-because the smallest value 56500 is in generation 0.
-
-This experiment took 59 minutes (on average desktop computer).\\
-
-Second experiment was different from the first one in that it 
-preserved the best individual to the next generation.
-
-It scored 22/50, i.e., 40\% success rate. A correct solution was
-created once in the initial generation and again it was
-sufficient to make the value in generation 0
-be the smallest value in the $I(M,i,z)$ curve; now 114000
- individuals to be processed.
-This experiment also took 59 minutes.\\
-
-Third experiment was different from the second one in that 
-$\eta$-normalization of generated individuals was performed.
-
-It scored 27/50, i.e., 54\% success rate. 
-Two runs contained a correct solution in the initial population.
-But now the $I(M,i,z)$ was best for generation 1 with 56000
- individuals to be processed.
-This experiment took 53 minutes.\\
-
-Fourth experiment was different from the third one in that
-\atTree representation was used instead of \sexprTree.
-
-It scored 34/50, i.e., 68\% success rate. Again two runs
-contained a correct solution in the initial population.
-The $I(M,i,z)$ was best for generation 0 with 56500 
-individuals to be processed.
-This experiment took 64 minutes.
-
-As we will see, this setup was the most successful one,
-thus we again show the performance curves.
-
-\includegraphics[scale=0.65]{reports/ep/report4/probabs.png}
-
-\includegraphics[scale=0.65]{reports/ep/report4/indivs.png}
-
-Fifth, last experiment was different from the fourth one in that
-optimized abstraction elimination was used instead of the simple one.
-
-It scored 33/50, i.e., 66\% success rate. One run
-contained a correct solution in the initial population.
-The $I(M,i,z)$ was best for generation 6 with 59500 
-individuals to be processed.
-This experiment took only 30 minutes. \\
-
-\subsection{Comparison with other results}
-
-Our results are less successful then those 
-presented by Yu \cite{yu01};
-they scored 40/50, i.e., 80\% success rate.  
-Her $I(M,i,z)$ was best for generation 4 with 17500 
-individuals to be processed.
-(We scored 34/50 (68\%); min $I(M,i,z) = 56,500$.)
-
-Yu also uses crossover as the only genetic operator, thus we
-can narrow the discussion to discussion of generating method and 
-crossover method.
-
-When we observe results of our experiments, it seems
-that our system is pretty good in term generation, but not so good 
-in crossover. Since the main focus point of this thesis is
-the term generation, it is positive rather then negative observation.
-
-Design of our crossover comes from that we wanted it to be
-very simple generalization of standard tree-swapping crossover.
-Whereas the crossover used by Yu is more structure oriented
---- as is stated in \cite{yu01} the crossover \textit{can only operate 
-between nodes in two main programs or between nodes in the same
-kind of $\lambda$ abstraction (i.e., $\lambda$ abstractions that 
-represent the same function arguments to the same higher order function).}
-%This seems to be great advantage for optimizing the
-
-Since our method has better success rate for generation 0, it would 
-be interesting to try combination of our generating 
-method with crossover used by Yu.
-
-From an optimist's point of view 34/50 versus 40/50 
-is not such a huge difference when we take to account that
-our general tree-swapping crossover is competing with specialized one.
-
-Positive observation is that our result at least outperforms 
-other results mentioned in \cite{yu01} with which Yu's method is compared.
-
-\textit{Generic genetic programming} scored 17/60 (28\%); 
-min $I(M,i,z) = 220,000$.
-
-\textit{GP with ADFs} scored 10/29 (34\%); 
-min $I(M,i,z) = 1,440,000$.
-
-\begin{todo}
-& napsat tam jak vypadaj nějaký nalezeneý řešení
-\end{todo}
-
-\section{Flies and Apples}
-
-\textit{Flies and Apples} is problem consisting in breeding 
-control programs for a fly agent in a simple world.\\
-
-Whereas previous problem is a kind of problem where we know precisely
-what problem we are solving and where we are more interested in 
-performance statistics than in the solutions since we know the correct solution in 
-advance, problem we are going to present now is of more playful nature.
-In this problem we are not interested in numerical results, instead 
-we are hoping for interesting \textit{behavior} to emerge.
-
-We start with defining a simple simulation and goal for agents.
-Then we prepare first testing simulation world. 
-After that we write some agent programs by hand. After that we 
-identify some useful recurring 
-themes occurring in those programs and pack them into functions
-which will become building blocks. During this process are also identified
-useful sensory data to be prepared for the agent program.
-After that we run the evolution and
-iteratively continue in this process  --- we may add more testing worlds,
-add more functions or change them, etc.
-
-Since we do not need to satisfy the \textit{closure} constraint, 
-it is much easier to perform such a "human-computer jam" than it
-would be in standard GP. And it is further simplified by use of
-functional language (such as \textit{Haskell}) where almost
-every construct can be expressed as function (or value).
-
-Another advantage of typed GP over standard GP is that 
-it is more capable of handling huge set of building blocks,
-since the search space can be significantly smaller due to 
-fact that types must match.
-\\
-
-Fly world is a square grid. 
-Each cell is ether empty or contains precisely one object.
-There are three kinds of objects:
-
-~\begin{enum}
- & Wall
- & Apple
- & Fly
-\end{enum}~
-
-Apples and flies have \textit{energy}.
-
-Moreover, each fly has its control program and inner state, 
-so called \textit{registers}.
-
-There is a queue of all flies in the fly world determining which 
-fly is next to make a move.
-After the fly in the front of the queue performs its move, 
-it is put to the back. In other words, flies are cyclically
-taking turns similarly like in common card games.\\
-
-A control program of a fly has as input a collection of various sensory information
-and returns as output agents move and agents registers for the next turn.
-
-Fly can perform two kinds of moves:
-
-~\begin{enum}
- & \textit{Trave.} Moves the fly to one of four adjacent cells determined by 
-   direction specified in the move.  
-   Result of the travel move depends on the content of the target cell.    
-   && If it is empty, then the fly simply moves to that cell.
-   && If it contains wall, then no movement is performed.
-   && If it contains an apple, then the fly moves to that cell, the apple 
-      is eaten and its energy is added to energy of the fly.  
-   && If it contains another fly, then the fly moves to that cell, the fly with bigger 
-      energy stays alive and the weaker one is eaten
-      and its energy is added to energy of the stronger fly.  
- & \textit{Split.} Puts a daughter fly to one of four adjacent cells.
-    Together with the direction, amount of energy and new registers 
-    for the fly are specified by this move. The mother gives this energy to 
-    its child. In order to be successful, the target cell must be empty
-    and the mother fly must have energy greater then 1. 
-\end{enum}~
-
-Agent's fitness is computed as sum of fitness cases; each 
-fitness case consists of one fly world and fitness in a
-fly world is computed as sum of agent's energy and of energy of all his descendants
-(daughters, granddaughter, ...) after predefined number of rounds.
-
-\Lets look in more detail on input for fly. It is collection of 
-various useful information. Its components are:
-
-~\begin{enum}
- & Current value of fly's Registers. 
- & Energy of the fly.
- & Direction of the last successful travel move.
- & Boolean value indicating whether the last move was successful.
- & Direction of the nearest apple.
- & Distance to the nearest apple.
- & Energy of the nearest apple.
- & Direction of the nearest fly.
- & Distance to the nearest fly.
- & Energy of the nearest fly.
- & Direction of the "center of gravity" of all apples.
- & Distance to the "center of gravity" of all apples.
-\end{enum} ~
-
-Set of building blocks is following.
-
-\begin{align*}
-\sigma = Input &\ar Output\\
-\Gamma = \{
-dUp              &: Direction                                       ,\\              
-dDown            &: Direction                                       ,\\              
-dLeft            &: Direction                                       ,\\              
-dRight           &: Direction                                       ,\\            
-output           &: Move \ar Registers \ar Output                   ,\\            
-travel           &: Direction \ar Move                              ,\\          
-split            &: Direction \ar Int \ar Registers \ar Move        ,\\       
-easySplit        &: Input \ar Move                                  ,\\     
-myEnergy         &: Input \ar Int                                   ,\\     
-myLastTravel     &: Input \ar Direction                             ,\\                
-myWasSuccess     &: Input \ar Bool                                  ,\\      
-nAppleDir        &: Input \ar Direction                             ,\\          
-nAppleDir        &: Input \ar Direction                             ,\\                    
-nAppleDist       &: Input \ar Distance                              ,\\                   
-nAppleEnergy     &: Input \ar Int                                   ,\\              
-nFlyDir          &: Input \ar Direction                             ,\\                      
-nFlyDist         &: Input \ar Distance                              ,\\                    
-nFlyEnergy       &: Input \ar Int                                   ,\\              
-cAppleDir        &: Input \ar Direction                             ,\\          
-cAppleDist       &: Input \ar Distance                              ,\\        
-myRegs           &: Input \ar Registers                             ,\\         
-\end{align*}
-\begin{align*}
-xGet             &: Input \ar Int                                   ,\\     
-yGet             &: Input \ar Int                                   ,\\     
-zGet             &: Input \ar Int                                   ,\\     
-dGet             &: Input \ar Direction                             ,\\          
-xSet             &: Int \ar Registers \ar Registers                 ,\\          
-ySet             &: Int \ar Registers \ar Registers                 ,\\          
-zSet             &: Int \ar Registers \ar Registers                 ,\\          
-dSet             &: Direction \ar Registers \ar Registers           ,\\                
-xInc             &: Registers \ar Registers                         ,\\              
-yInc             &: Registers \ar Registers                         ,\\              
-zInc             &: Registers \ar Registers                         ,\\              
-rotCW            &: Direction \ar Direction                         ,\\                                                             
-(==)             &: Int \ar Int \ar Bool                            ,\\                                      
-(<=)             &: Int \ar Int \ar Bool                            ,\\               
-if'              &: Bool \ar Output \ar Output \ar Output           ,\\                                  
-if'              &: Bool \ar Move \ar Move \ar Move                 ,\\                                  
-if'              &: Bool \ar Direction \ar Direction \ar Direction  ,\\              
-if'              &: Bool \ar Int \ar Int \ar Int                    ,\\     
-if'              &: Bool \ar Distance \ar Distance \ar Distance     ,\\        
-if'              &: Bool \ar Registers \ar Registers \ar Registers  ,\\                           
-0                &: Int                                             ,\\                                     
-1                &: Int                                             ,\\                                    
-2                &: Int                                             \}\\
-\end{align*}
-
-We will not discuss it in great detail, only briefly: First four 
-elements are four direction values. Next four functions are
-constructors for $Output$ and $Move$ type. Next thirteen functions
-are for accessing components of $Input$. Next eleven functions are 
-related to $Registers$. The $rotCW$ performs clock-wise rotation of a direction.
-Next two functions are comparison functions. Next six are various \textit{if}s. And last 
-three are integer constants.
-
-Tedy celkem 44 prvků množiny stavebních bloků, tedy poměrně obrovská množina
-na GP poměry.
-
-Takže množina atomických typů je následující:
-Move Bool Registers Input Direction Int Output Distance
-
-Používáme tři levely. První level vypadá takle
-
-[obrázek]
-
-Další dva levely jsou reakce na typické chování mouchy 
-která používala pro předchozí level triviální strategii
-jdi furt za nejbližšim jablkemn. roto jsou následující dva levely 
-léčka na tuto strategii.
-
-V druhém levelu se zmíněná moucha nachytá a vleze do 
-chodby kde je mnohem mín jablek než v v tom mega jablku z jakblek.
-
-[obrázek]
-
-Třetí dovádí předchozí myšlenku do extrému. 
-Moucha se zasekne a nesebere žádné jablko.
-
-V počátečních pokusech byl problém s tím, že mochy s počáteční energii 1
-neměli moc tendenci se množit. Proto byla simulace předělána tak, aby
-že nyní je počáteční energie 100. To mělo na Mouchy blahodárný vliv.
-Ale i pro mouchy sw počátečni energii 1 vznikli některé zajímavé strategie
-
-[povídání o zajímavých strategiích pro poč en 1]
-
-Po navíšení poč energie vznikli tyto zajímavá chování
-
-[povídání o zajímavých strat por poč en 100]\\
-
-Aplikace: například v oboru počítačových her, kde nám často jde o
-zajímavé nelineární chování by se tento přístup k využití typovaného funkcionálního
-programování dal využít.
-
-Předvedený postup by taky ohlo být možné použít na
-
-\section{Simple Symbolic Regression}
-
-\textit{Simple Symbolic Regression} is a problem described
-in \cite{koza92}. Objective of this problem is to 
-find a function $f(x)$ that fits a sample
-of twenty given points. The target function is 
-function $f_{t}(x) = x^4 + x^3 + x^2 + x$.
-
-Desired type of generated programs $\sigma$ and 
-building blocks context $\Gamma$ are following.
-\newcommand{\Real}{\mathbb{R}}
-\begin{align*}
-\sigma = \Real \ar &\Real\\
-\Gamma = \{
-  (+)  &: \Real \ar \Real \ar \Real    ,\\
-  (-)  &: \Real \ar \Real \ar \Real    ,\\
-  (*)  &: \Real \ar \Real \ar \Real    ,\\
-  rdiv &: \Real \ar \Real \ar \Real    ,\\
-  sin  &: \Real \ar \Real              ,\\
-  cos  &: \Real \ar \Real              ,\\
-  exp  &: \Real \ar \Real              ,\\ 
-  rlog &: \Real \ar \Real              \}\\
-\end{align*}
-where
-\begin{align*}
-rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
-p/q & \mbox{otherwise } \end{cases}  \\
-rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
-log(\vert x\vert) & \mbox{otherwise}. \end{cases}
-\end{align*}
-
-Fitness function is computed as follows
-
-$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
-
-where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.\\
-
-An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
-considered as a correct individual.\\
-
-On this problem we demonstrate that if control parameters are set to be
-compatible with standard GP, then the results obtained fit the results
-presented in \cite{koza92} relatively well.
-
-Those control parameters are following two:
-
-~\begin{enum}
- & \textit{Ramped half-and-half strategy} is used as search strategy.
- & Option of preserving the best individual into the next population is set \textbf{off}.
-\end{enum}~
-
-The experiment consisted of 50 independent runs of GP algorithm.
-Each run had maximally 50 generations and 500 individuals as population size.\\
-
-It scored 17/50 (34\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 23 
-with 192,000 individuals to be processed.
-The experiment took 46 minutes.\\
-
-Performance curves are following.
-
-\includegraphics[scale=0.65]{reports/SSR/1/probabs.png}
-
-\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}\\
-
-\Lets compare it with results from \cite{koza92} (based on 113 runs):
-
-Success rate 35\%; minimal $I(M,i,z)$ in generation 24 
-with 162,500 individuals to be processed.\\
-
-Second experiment was performed to compare 
-\textit{ramped half-and-half strategy} with our
-\textit{geometric strategy} (with default parameter $q=0.75$). 
-Thus all control parameters stay same except for the 
-search strategy for generating algorithm.
-
-It scored 21/50 (42\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 29 
-with 150,000 individuals to be processed.
-The experiment took 26 minutes.\\
-
-\includegraphics[scale=0.65]{reports/SSR/2/probabs.png}
-
-\includegraphics[scale=0.65]{reports/SSR/2/indivs.png}\\
-
-So our geometric strategy is slightly more successful then 
-ramped half-and-half strategy. But more interesting is that
-it made the whole experiment almost twice as fast.
-
-~\\TODO dát sem ty grafy průměrný term size
-
-\newpage
-\section{Artificial Ant}
-
-\textit{Artificial Ant} is another problem described
-in \cite{koza92}. Objective of this problem is to 
-find a control program for an artificial ant so
-that it can find all food located on "Santa Fe" trail.
-
-The Santa Fe trail lies on toroidal square grid.
-The ant is in the upper left corner, facing right.
-It has following layout.
-
-\includegraphics[scale=1]{santafe.png}
-
-The ant is able to move forward, turn left, and sense if a food 
-piece is ahead of him.\\
-
-Desired type of generated programs $\sigma$ and 
-building blocks context $\Gamma$ are following.
-\begin{align*}
-\sigma = An&tAction\\
-\Gamma = \{~~
-  l    &: AntAction                              ,\\
-  r    &: AntAction                              ,\\
-  m    &: AntAction                              ,\\
-  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
-  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
-  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
-\end{align*}
-
-Action $l$ turns the ant left.
-
-Action $r$ turns the ant right.
-
-Action $m$ moves the ant forward.
-
-Action $ifa x y$ (if-food-ahead) performs action $x$ if a food piece is ahead of the ant,
-otherwise it performs action $y$.
-
-Action $p2 x y$ first performs action $x$ and after it action $y$.
-
-Action $p3 x y z$ first performs action $x$, after that action $y$ and finally $z$.\\
-
-Actions $l, r$ and $m$ each take one time step to execute.
-
-Ants action is performed over and over again until it reaches predefined
-maximal number of steps. In \cite{koza92} this limit is set to be 400 time steps.
-
-Fitness value is equal to number of eaten food pieces.
-
-An individual such that eats all 89 pieces of food is 
-considered as a correct solution.\\
-
-There is also mentioned following ant program which is described as correct solution.
-
-\texttt{~\\
-(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))\\
-}
-
-But here arises a problem: In our implementation this program needs 545
-time steps; if it is given only 400 time steps, then it eats only 79 pieces
-of food. It is uncertain whether it is caused by my mistake in implementation
-or by inaccuracy in \cite{koza92}. Since everything else seems working just 
-fine, I have decided to set the limit to 600 in my experiments.
-
-Due to those uncertainties, we will not perform comparing experiments 
-as we did in the problem of simple symbolic regression.\\
-
-In the following experiments, the option of preserving the best individual 
-into the next generation is active (since we are not trying to compare 
-with results in \cite{koza92}).
-
-As usually, first experiment consisted of 50 independent runs of GP algorithm.
-Each run had maximally 50 generations and 500 individuals as population size.
-\textit{Ramped half-and-half strategy} is used as search strategy for the first 
-experiment.\\
-
-It scored 7/50 (14\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 28 
-with 449,500 individuals to be processed.
-The experiment took 265 minutes.
-
-Performance curves are following.
-
-Similarly as in previous problem, second experiment was performed to compare 
-\textit{ramped half-and-half strategy} with our
-\textit{geometric strategy} (with default parameter $q=0.75$). 
-Thus all control parameters stay same except for the 
-search strategy for generating algorithm.
-
-It scored 19/50 (38\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 10 
-with 115,500 individuals to be processed.
-The experiment took 107 minutes.\\
-
-\includegraphics[scale=0.65]{reports/Ant/1/probabs.png}
-
-\includegraphics[scale=0.65]{reports/Ant/1/indivs.png}\\
-
-This is big improvement in all three watched factors:
-
-\begin{enumerate}
- \item 38\% vs 14\% success rate.
- \item 115,500 vs 449,500 individuals needed to be processed.
- \item 107 vs 265 minutes.  
-\end{enumerate}
-
-Third factor, the time difference might give us clue about what is going on.
-
-Those two experiments differs only in used term generating method.
-During one run the vast majority of the time is spend by evolution,
-the generation process only slightly prolongs the time spend in generation 0.
-The run time of the evolution is influenced by two important factors:
-Number of generations and time that crossover takes.
-
-The number of generations of more successful method is surly smaller,
-but it seems improbable that it would have such a big impact 
-(this is idea is supported by results of experiments with
-even-parity problem, where times are almost the same but success 
-rates differs significantly).
-
-Time of crossover is proportionate to number of nodes that 
-an individual has. Thus it seems that observed slowdown is
-due to terms being too big. \\
-
-In order to test this hypothesis, we run additional experiments
-to get statistics about dynamics of individuals' sizes.
-
-~\\TODO dát sem ty grafy průměrný term size
-
-\section{Results summary}
-
-\chapter{Previous related work}
-
-\section{Evolution of Constrained Syntactic Structures }
-by Koza
-
-\section{STGP}
-by Montana
-
-\section{Higher-Order Functions and Lambda Abstractions} 
-by Yu
-Požívá dvou levelovou hierarchii (najít tu pasaž)
-
-oproti nám umí i polymorfní typy
-
-K  lambda abstrakcím přistupuje jako k jedinému bodu který se dělí až 
-vevnitř.
-
-Používá na to speciálně přichystaný křížení
-
-Generování počáteční generace je v zásadě to samí co koza, když příde na vygenerování 
-abstrakce tak se to zase generuje klasicky s tim že proměný jsou \#1 \#2 atd
-
-klade velkej důraz na foldr jakožto na implicitní rekurzi
-
-v even parity scoruje o něco lépe než my.
-
-\section{Genetic Programming with Polymorphic Types and Higher-Order Functions}
-by Franck Binard and Amy Felty
-
-%\chapter{Ideas yet not implemented}
-\chapter{Future work}
-
-\section{Polymorphic variant}
-
-\begin{todo}
- & poznamky o čem se mimojiné zmínit:
-   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
-      jako alternativa k ADF ! Myslim že hodně zajimavý
-\end{todo}
-
 \section{Big Context}
@@
 
+The goal of this thesis was to design and implement a system
+performing Genetic Programming over some typed functional programming language.
+As this typed functional programming language we have chosen the 
+\textit{simply typed lambda calculus}.
+We have also obligated ourselves to do it in a such way that 
+generalizes the standard Genetic Programming, rather than
+crates completely new system.
+
+We may say that we have chosen the simplest possible
+options in both areas --- the simply typed calculus and the standard GP.
+We see this decision as fortunate one --- since it enables us add more
+complex features after sufficient exploration of those simplest cases.
+
+We have started the journey toward this goal by study 
+of subject that intersects both those areas:
+Method which for desired type and context generates lambda terms.
+
+The way to obtain this method started with inference rules 
+(the defining concepts of the typed lambda calculus),
+it proceeded through term generating grammars and finished 
+with inhabitation trees.
+
+A* algorithm has been utilized to drive systematic enumeration
+of terms in \lnf.
+
+\newpage
 \begin{todo}
- & !!!!!!!!!!!!!!!!!
+ & my tu řikáme že to závadíme kuli typum, ale my děláme v víc! -dáváme tam
+   i abstrakce atd - ale mužeme říct, že tim že se to pak převede zase na 
+   to bez abstrakcí tak setrváváme dal v krasnym bez-lambda abstrakčí ale
+   s tou výhodou že naše metoda gerneruje všechny struktury pro ten danej typ ...
+   nebo tak něco - asi to dát spíš do design of our system nebo někam
+   ale někde to asi zmínit
 \end{todo}
@@
 
+\textbf{Jen když zbude čas:}
+
+\begin{todo}
+
+& Complexity discussion 
+&& !!!
+&& zmínit že mi moc neřešíme formální/asymptotický složitosti
+  &&& de nám hlavně o realný časy
+&& vyjít z komplexnosti A* asi
+  &&& složitost výrazně ovlivněná kvalitou heuristický funkce
+
+ & Brief history of the project\\
+  zmínit třeba tyhle fáze:
+  && že sem si delal evaluator sám co sem pak s performenc duvodu opustil
+  && že sem hledal intuicionistickej dokazovač ale že nejsem nenarazil na vhodnej
+     pač nejsou delany na to plivat hodně dukazu  
+  && že sem nejdriv pouzival ten vypsychlej dokazovač co tam měl nějaký ty 
+     spešl rozhodovací stromy
+  && pak sem to přepsal tak aby to nebylo jen na GP šitý ale obecně na jakoukoliv 
+     evoluci
+  && ...
+
+& pro \ref{benefits}
+&& možná eště zmínit něco jako : \\
+Jedna drobná nevýhoda je, že pokud chceme generovat termy od nejmenšího do největšího
+a pokud lnf generujeme od nejmenšího po největší tak pak generovaní \benf
+timhle trikem nemusí bejt od nejmenšího po nějvětší. V naší aplikaci (tzn GP)
+ale nijak nebazírtujeme na týhle vlastnosti a nakonec budeme spíš stejně používat
+nějaký nedeterminizmy aby to fungovalo líp, takže to vubec nevadí, že budou trochu přeházený.
+
+ & pro \ref{barlike}
+ && možná zmínit že podobně jako mužeme transformovat inference rules
+   na grammar rules, mužeme i obráceně grammar rules
+   převést na inference rules - to bylo ostatně uděláno v tom dukazu ale
+   je to podle mě hezký říct ty pravidla explicitně.
+ & pro \ref{lnf}
+ && potřebuju eště nějaký \textbf{propositions} z \cite{barendregt10} tu uvíst ?
+ && ta definice je tam daná pro $\tur{}{M}{\sigma}$,
+   ukazat ze ta naše je ok, že tam má navíc to $\Gamma$.
+   Stačí udělat tríček že beru funkci co má lambda hlavu odpovídající 
+   tomu kontextu.
+
+   & k \ref{Term-generating-grammar}
+    &&  buď dokázat ty převody na grammar rules explicitně
+     nebo říct že se to dokáže podobnym stylem jako v dukazu 
+     v sekci \ref{barlike}
+    && možná ocitovat toho van Wijngaarden [1981] s těma 2-level grammar
+     jak je to v Barendregtovi \cite{barendregt10} 
+ & možná konec (tree reprez of typed terms) subsekce trochu přeformulovat, je to takový   
+   kostrbatý...
+ & o odvoz pravidlech říct
+   && možná zmínit že se jedná o \textit{gentzen style proof}
+   && říct že jsou to modus ponens a jeden směr věty o dedukci  
+ & nějak přepsat \ref{GP-prog-trees} aby to bylo ve stejnym stylu jako v \ref{tree-reps}
+ & napsat transformace mezi \atTree a \sexprTree
+%& ujistit se že je v části o tree reprezentacích 
+%  je napsaný co to znamená že podstrom má nějakej typ,
+%  že podstromy odpovídaj podvýrazum atd
+% & EXAMPLES of tree representations of \lterms  
+% & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
+%   přímo \sexprTree a říct proč se tomu tak řiká 
+%   (na tom místě co je v textu dřív) 
+% & zkontrolovat že reference tree-reps se odkazuje na dobrý místo,
+%   mohlo by se to rozbít tim že tree reprezentations - typovany/netypovany 
+%   jsou od sebe oddelený
+\end{todo}
+
 \addcontentsline{toc}{chapter}{Conclusion}
@@
   Springer, 2005. ISBN 978-0-387-26417-2
+
+\bibitem{fg}
+ Riccardo Poli, William B. Langdon, Nicholas F. McPhee
+ \emph{A Field Guide to Genetic Programming}.
+ Lulu Enterprises, UK Ltd, 2008.
 
@@
 Dr scient thesis, University of Oslo, Norway, 1994.
+
+\bibitem{kes}
+Forrest Briggs, Melissa O’Neill.
+\emph{Functional Genetic Programming and Exhaustive
+Program Search with Combinator Expressions.}
+International Journal of Knowledge-based and Intelligent Engineering Systems,
+Volume 12 Issue 1, Pages 47-68, January 2008.
 

### 2013-07-31 v4/GP_Data.hs

@@ -15,7 +15,7 @@ import TTerm (Typ,Context,CTTerm(..),
 import TTree (CTT(..),ttreeDepth,ttreeChangeSubtree,ttreeSubtree,TTree,ttreePoses2, 
               TTPos,ttreePoses2WithTyps,ttreePoses2ByTyp, ttreePoses2WithTyps_onlyCompatible)
 
-import IM3 ( SearchOptions , prove ,proveCTTerm , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
+import InhabTree ( SearchOptions , prove ,proveCTTerm , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
 --import IM2 ( SearchOptions , prove , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )
 
 --import IM ( SearchOptions , prove , kozaSearchOptions , allEdgesSearchOptions , geomSearchOptions )

### 2013-07-31 v4/InhabTree.hs

@@ -1,36 +1,985 @@
+--  Nový pokus o Inhabitation Machines bez potřeby generovat graf dopředu a vůbec to udělat pořádně
+
+--  Asi by to chtělo celý předelat tak, že ZTree bude stavová monada, ale radši 
+-- to nejdřív už dodělat když je to takle rozdělaný a až pak dyštak překopat.
+
+
+-- IM2 = přidání polymorfizmu
+
 module InhabTree 
-( prove 
-, SearchOptions(..)  
-, kozaSearchOptions
-, allEdgesSearchOptions
-, geomSearchOptions  
-) where
+ ( SearchOptions(..) 
+ , defaultSearchOptions 
+ , kozaSearchOptions
+ , allEdgesSearchOptions
+ , geomSearchOptions 
+ , prove 
+ , proveCTTerm
+ ) where
 
-import TTerm (Typ,Context)  
---(Symbol,Typ(..),TTerm(..),Context,ttermTyp,toSki,toSki',checkTyp)
+import Debug.Trace
+
+import TTerm (Symbol,Typ(..),TTerm(..),Context,ttermTyp,toSki,toSki',checkTyp,fullEtaReduce,CTTerm(..),
+  optSki)
+
+import TTree (CTT,mkCTT2)
+
+import Data.List
+import Data.Maybe
+import qualified Data.Set as Set
+import qualified Data.Map as Map
+import Data.Set (Set)
+import Data.Map (Map)
+
+import System.Random
+
+import qualified Data.PSQueue as Q
+
+import PolyUtils ( Substi , Table, Entry(..), TypHead , Edge, MGU 
+                 , NextVar, CurrentTyp , GlobalsTable , LocalsTable
+                 , emptyTable, ctxToTable , addToTable , subFromTable_, showTable, showSubsti
+                 , typeArgz  
+                 , match , applySubsti , composeSubsti, emptySubsti , mgu
+                 , getEdges_all , updateByMGU )
+
+
+
+ 
+
+
+--proveAll :: Int -> Typ -> Context -> [CTT]
+
+proveCTTerm :: SearchOptions -> [CTTerm]
+proveCTTerm so = 
+  let ( so' , problemHead ) = problemHeadPreproccess so
+      trees = proveWith2 so'
+   in map ( CTTerm problemHead . optSki . fullEtaReduce . tree2tterm ) trees   -- <============== toSki s ' je s typeCheckem ...........
+
+
+prove :: SearchOptions -> [CTT]
+prove so = 
[…]

### 2013-07-31 v4/Problems/SSR/Problem.hs

@@ -10,7 +10,7 @@ import Problems.SSR.Funs (ff)
 
 import ProblemUtils 
 import GP_Core (FitFun(FF6))
-import GP_Data (CTTGen(..))
+import GP_Data (CTTGen(..),CTTermGen(..))
 
 
 reg = PO_CTTP_ PO_CTTP {
@@ -29,6 +29,25 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_ff          = FF6 (asType::Double->Double) ff "Problems.SSR.Funs", 
   
   cttp_saveBest    = False -- True 
+  
+}
+
+rege = PO_CTTeP_ PO_CTTeP {
+  cttep_code        = "ssr2"                                        ,
+  cttep_info        = "Simple Symbolic Regression. With @-tree."   ,
+  cttep_data        = jsData                                       ,
+  cttep_numRuns     = IntSlider "Runs"            1 100   1    1   ,
+  cttep_numGene     = IntSlider "Generations"     0 200   50   10  ,
+  cttep_popSize     = IntSlider "Population size" 0 5000  500  100 ,
+  
+  cttep_typ         = dou1                                         ,
+  cttep_ctx         = ctx                                          ,
+
+  cttep_gOpt        = CTTermG_Geom  dou1 ctx 0.75 , --CTTG_Geom  dou1 ctx 0.75  , --CTTG_Koza2 dou1 ctx , 
+  
+  cttep_ff          = FF6 (asType::Double->Double) ff "Problems.SSR.Funs", 
+  
+  cttep_saveBest    = False -- True 
   
 }
 

### 2013-07-31 v4/Register.hs

@@ -1,4 +1,4 @@
-module Job ( job, problemList ) where
+module Register ( job, problemList ) where
 
 import Text.JSON (decode, Result(..), JSValue )
 import JSONUtils (fromJsStr,jsProp,jsArr)
@@ -23,6 +23,7 @@ import qualified Problems.EvenParity.Problem as EvenParity
 
 regs = 
  [ reg SSR.reg 
+ , reg SSR.rege 
  , reg Ant.reg
  , reg EvenParity.reg
  , reg EvenParity.rege

### 2013-07-31 v4/Server.hs

@@ -17,7 +17,7 @@ import Data.ByteString.Lazy.UTF8 (fromString)
 import Text.JSON ( encode )
 
 import Utils ( unescape )
-import Job (problemList,job)
+import Register (problemList,job)
 import ServerInterface( OutputBuffer , OutRecord(..) )
 
 

### 2013-07-31 v4/data/js/Problems/Fly02.js

@@ -40,7 +40,8 @@ var onGo = function(){
   container
    .html('')
    .append( infoText )
-   .append( linksDiv ) 
+   .append( linksDiv )
+   .append( '<hr>' ) 
    .append( flySimContainer ); 
 
   flySim = FlySim.init( '_fly-sim' , function(sim){
@@ -61,15 +62,20 @@ var onGo = function(){
 
 };
 
-var fenotyp = function( solution , ffVal ){ 
+var fenotyp = function( solution , ffVal , actGen ){ 
 
   var currSolutionID = nextSolutionID ;
   nextSolutionID ++ ;
 
   solutions[ currSolutionID ] = solution ;
 
+  if( actGen == 0 ){
+    linksDiv.append('<hr>');  
+  }
+
   var newLink = $('<a>')
-   .attr('href','#')
+   //.attr('href','#')
+   .css('cursor','pointer')
    .attr('title', solution.toString() )
    .html( ffVal )
    .click( function(){ onLinkClick(currSolutionID); } );

### 2013-08-01 tex/thesis.tex
čtvrtek 6:36

@@
 %$<$dedication$>$
-I would like to express my gratitude to my supervisor RNDr. Petr Pudlák, Ph.D.
-for the useful comments, remarks and help with this master thesis.
+I would like to express my sincere gratitude to my supervisor RNDr. Petr Pudlák, Ph.D.
+for the useful comments, remarks and willingness to help me with 
+this thesis.
 I would also like to thank my parents, who have supported me throughout
-my studies.
+my studies and whole life.
 
@@
 % abstrakt v rozsahu 80-200 slov; nejedná se však o opis zadání diplomové práce
+{\fontsize{0.35cm}{1em}\selectfont 
+V této práci je představen design a implementace
+systému řešícího úlohu genetického programování v
+simply typed lambda kalkulu. Je zde představena
+metoda inicializace počáteční populace založená 
+na technice produkující typované lambda termy 
+v dlouhé normální formě. Tato metoda je parametrizována
+jednoduchou prohledávací strategií. Několik takových
+prohledávacích strategií je představeno, jako například
+strategie pro systematické generování nebo strategie
+odpovídající standardní ramped half-and-half metodě.
+Další z představených strategií, strategie jménem 
+geometrická strategie je blíže podrobena experimentům,
+které ukáží že má několik žádoucích efektů na průběh evoluce, 
+jakými jsou zlepšení míry úspěšnosti, nižší časové nároky 
+a menší průměrnou velikost termů v porovnání se standardní
+ramped half-and-half metodou generování jedinců. Další výkonnostní
+zlepšení jsou navržena a podpořena experimenty, jedná se o
+$\eta$-normalizaci vygenerovaných jedinců a @-tree
+reprezentaci jedinců. Použitý proces eliminace abstrakcí
+umožňuje použití jednoduchého podstromy měnícího 
+křížení.}
 
 Klíčová slova:
-geneticke programovani, funkcionalni programovani
+genetické programovaní, funkcionalní programovaní
 % 3 až 5 klíčových slov
@@
 Abstract:
+{\fontsize{0.35cm}{1em}\selectfont 
+In this thesis is presented design and implementation
+of a system performing genetic programming in simply
+typed lambda calculus. Population initialization
+method based on term generating technique producing
+typed lambda terms in long normal form is introduced. 
+This method is parameterized by simple search strategy.
+Several search strategies are presented, such as
+strategy for systematic generation or strategy
+corresponding to standard ramped half-and-half method.
+Another such a strategies called \textit{geometric} strategy
+is further examined in experiments and shown
+to have various desirable effects such as
+improved success rate, lesser time consumption and 
+smaller average term size in comparison with standard 
+ramped half-and-half generating method. Other performance
+enhancements are proposed and supported by experiments
+such as $\eta$-normalization of generated individuals and 
+@-tree representation of individuals. Abstraction elimination
+is utilized to enable use of simple tree-swapping crossover.
+}
+
 % abstrakt v rozsahu 80-200 slov v angličtině; nejedná se však o překlad
@@
 
-The contribution of this thesis can be seen 
-in that it deals with this task using concepts from
+%The contribution of this thesis can be seen 
+This thesis deals with this task using concepts from
 theory of typed $\lambda$-calculus and combinatory logic
@@
 \begin{itemize}
- \item \textit{Inhabitation tree} (a slight modification 
-   of \textit{inhabitation machine}\cite{barendregt10}
-   suited better for reasoning about implementation details) 
+ \item \textit{Inhabitation tree} 
+ %(a slight modification 
+ %  of \textit{inhabitation machine}\cite{barendregt10}
+ %  suited better for reasoning about implementation details) 
    is presented and used for \textit{generation of individuals}.
@@
    such as \textit{systematic} strategy performing exhaustive enumeration
-   of individuals and strategy performing same individual generation 
-   as standard GP method \textit{ramped half-and-half}.  
- \item One of those search strategies, the \textit{geometric} strategy, is
+   of individuals and strategy performing generalization of 
+   standard GP term generating method \textit{ramped half-and-half}.  
+ \item One of proposed search strategies, the \textit{geometric} strategy, is
    examined in a couple of experiments and shown to have 
@@
        two enhancements of the system inspired from the theoretical
-       background of the system are examined and shown to have 
+       background are examined and shown to have 
        improving effect on performance of the system.     
@@
    to enable use of simple \textit{mutation operator}.
+   Both simple and optimize versions of this transformation are examined.
 % \item \textit{Polymorphic} generalization of inhabitation tree,
@@
 
+\section{Evaluation}
+
+Evaluation is process of assigning fitness value to all 
+individuals in the population. It is also usual to check
+individuals for being correct solution.
+If such a correct solution is find, the algorithm terminates. 
+In our implementation fitness function returns
+pair of real number (fitness value) and boolean value (indicating
+whether tested individual is correct solution).
+
+\newpage
+
 \section{Selection}
@@
 It uses fitness value of each individual in straightforward way to determine probability of selecting this individual.\\
-\newpage
 Let there be $popSize$ individuals in the population.\\
@@
 
-Similarly as for crossover, a maximum permissible depth for mutant individual 
+Analogically as for crossover, a maximum permissible depth for mutant individual 
 is defined.
+
+Similarly as Koza in \cite{koza92}, we do not use the mutation operator
+in this thesis. This does not mean that I do not like it. 
+Conversely, I am looking forward too its enhancing effect in the future work.
 
@@
 
-Koza in \cite{koza92} uses as default probabilities those values:
+Koza\cite{koza92} uses as default probabilities those values:
 
@@
 \end{itemize}
-
-\section{Evaluation}
-
-Evaluation is process of assigning fitness value to all 
-individuals in the population. It is also usual to check
-individuals for being correct solution.
-If such a correct solution is find, the algorithm terminates. 
-In our implementation fitness function returns
-pair of real number (fitness value) and boolean value (indicating
-whether tested individual is correct solution).
 
@@
 \textit{lnf} term generating grammar.
-
-\begin{todo}
- & všude nahradit ty množinový za $\Gamma , x:\sigma$ a taky $(x,\Sigma)$ kde se to nehodí.
- & Až to bude dostatečně sepsáno, tak do této kapitoly taky přidat
-   část o abstraction elimination a o polymorfním zobecnění inhabitation tree.
-   Nezapomenout to pak zmínit i v tomhle úvodu.
-\end{todo}
 
@@
       that term generating grammars are not grammars in
-      the standard sense. 
+      the standard sense (but instead \textit{2-level-grammars}). 
       \item
@@
       We may think about the whole structure of generating 
-      at once since the whole tree, although infinite, is
+      at once since the whole tree, although usually infinite, is
       fairly simple structure of alternating layers of \textit{and} 
@@
       It enables us to ask such questions
-      as: What would be usage of inhabitation trees with 
-      number labeled edges coming from \textit{or}-nodes? Et cetera.
+      as: \textit{What would be usage of inhabitation trees with 
+      number labeled edges coming from \textit{or}-nodes?.}
       \end{enumerate}   
@@
       This is even more visible when we introduce
-      concept of an \textit{incomplete inhabitation tree}
+      concept of an \textit{incomplete inhabitation tree};
+      such a tree is created by omission of some subtrees
+      with and-node roots (i.e., those accessible by \textit{or} edge).
+      \textit{Finite incomplete inhabitation tree} corresponds
+      to finite collection of typed \lterms. In very natural 
+      sense we can call it a \textit{typed multiterm} --- a collection
+      of typed terms forming one tree where those terms share
+      the common parts.
+
+      Thus the process of term generation is bidirectional ---
+      from (incomplete) inhabitation tree (seen as collection) 
+      we may obtain its element by solving it. But we may also 
+      take collection of typed trees and compress them
+      into one tree by sharing their common parts, which
+      happens to be the (incomplete) inhabitation tree 
+      that can be solved precisely by those terms which 
+      constructed it.
 
@@
 inhabitation tree ideas to perform systematic generation and other
-kinds of generation of lambda terms;
+kinds of generation of \lterms;
 but maybe even more interesting application of inhabitation trees
@@
 
+\newpage
 \section{ Inhabitation machine }
@@
 continuously updated graphs and if 
-the \textit{fenotype} JavaScript file is present 
+the \textit{phenotype} JavaScript file is present 
 for the problem, then is also available immediate 
-presentation of the best individual from each generation.\\
+presentation of the best individual from each generation.
+JavaScript functions standing for best individuals are 
+also immediately accessible through JavaScript console,
+e.g. for debugging purposes.\\
 
@@
 \end{easylist}
+
+~\\[6em]
+
+\includegraphics[scale=0.4]{skryn.png}
 
@@
 
+\newpage
 \section{Flies and Apples}
@@
 output           &: Move \ar Registers \ar Output                   ,\\            
+\end{align*}
+\begin{align*}
 travel           &: Direction \ar Move                              ,\\          
@@
 nAppleDir        &: Input \ar Direction                             ,\\                    
-nAppleDir        &: Input \ar Direction                             ,\\                    
 nAppleDist       &: Input \ar Distance                              ,\\                   
@@
 nFlyEnergy       &: Input \ar Int                                   ,\\              
-\end{align*}
-\begin{align*}
 cAppleDir        &: Input \ar Direction                             ,\\          
@@
 
+\newpage
 We will not discuss it in great detail, only briefly: First four 
 elements are four direction values. Next four functions are
-constructors for $Output$ and $Move$ type. Next thirteen functions
+constructors for $Output$ and $Move$ type. Next twelve functions
 are for accessing components of $Input$. Next eleven functions are 
@@
 
-~\\\\\textbf{todo DODĚLAT ASI AŽ POTOM, AŤ SE S TIM MOC NEZDRŽUJU}\\\\
-
-Tedy celkem 44 prvků množiny stavebních bloků, tedy poměrně obrovská množina
-na GP poměry.
-
-Takže množina atomických typů je následující:
-Move Bool Registers Input Direction Int Output Distance
-
-Používáme tři levely. První level vypadá takle
-
-[obrázek]
-
-Další dva levely jsou reakce na typické chování mouchy 
-která používala pro předchozí level triviální strategii
-jdi furt za nejbližšim jablkemn. roto jsou následující dva levely 
-léčka na tuto strategii.
-
-V druhém levelu se zmíněná moucha nachytá a vleze do 
-chodby kde je mnohem mín jablek než v v tom mega jablku z jakblek.
-
-[obrázek]
-
-Třetí dovádí předchozí myšlenku do extrému. 
-Moucha se zasekne a nesebere žádné jablko.
-
-V počátečních pokusech byl problém s tím, že mochy s počáteční energii 1
-neměli moc tendenci se množit. Proto byla simulace předělána tak, aby
-že nyní je počáteční energie 100. To mělo na Mouchy blahodárný vliv.
-Ale i pro mouchy sw počátečni energii 1 vznikli některé zajímavé strategie
-
-[povídání o zajímavých strategiích pro poč en 1]
-
-Po navíšení poč energie vznikli tyto zajímavá chování
-
-[povídání o zajímavých strat por poč en 100]\\
+This gives 44 elements of building block context. It is quiet huge 
+context, especially if it was a standard GP \TuF.\\
+
+So the set of atomic types is following =\\
+$\{Move, Bool, Registers, Input,
+Direction, Int, Output, Distance\}$.\\
+
+We currently use three worlds.
+First one looks like this:\\
+
+~\\[2em]
+
+\includegraphics[scale=0.8]{fly/1.png}
+
+Other two worlds are reaction on typical behavior 
+of a trivial and boring fly which was successful
+enough to prevent further evolution.
+This boring strategy was just following the path 
+of the nearest apple.
+
+\newpage
+
+Following two worlds are therefore "Devil's gift" for this fly.\\
+
+\includegraphics[scale=0.4]{fly/2.png}
+\includegraphics[scale=0.4]{fly/3.png}\\
+
+In the second world the boring fly enters the corridor
+and misses the mega apple.
+The third world escalates previous idea into absurdity.
+The boring fly eats nothing since it dumbly tries to eat the
+apple behind the wall.
+
+By introduction of those two new worlds new kinds of behavior emerged,
+but not interesting enough to be worth explaining here. One can,
+however, check those behaviors out in the gallery of interesting fly 
+behaviors, included on the enclosed CD.\\
+
+In the initial experiments there was problem with that flies did not
+want to split. This was due to that they were given default energy 1.
+Since this problem is not tied by some unchangeable definition,
+we changed this default value to 100 and aded a siplified version of split constructor. That had aftermath in new kinds of behavior.
 
@@
 
-\section{Results summary}
+%\section{Results summary}
 
@@
 The goal of this thesis was to design and implement a system
-performing Genetic Programming over some typed functional programming language.
+performing GP over some typed functional programming language.
 As this typed functional programming language we have chosen the 
@@
 We have also obligated ourselves to do it in a such way that 
-generalizes the standard Genetic Programming, rather than
+generalizes the standard GP, rather than
 crates completely new system.
@@
 We may say that we have chosen the simplest possible
-options in both areas --- the simply typed calculus and the standard GP.
+option in both areas --- the simply typed calculus and the standard GP.
 We see this decision as fortunate one --- since it enables us add more
@@
 
-A* algorithm has been utilized to drive systematic enumeration
-of terms in \lnf.
+A* algorithm in combination with inhabitation trees
+has been utilized to drive systematic enumeration
+of typed \lterms in their \lnf. This enumeration was further 
+parameterized by simple search strategy in order to enable
+such different approaches as \textit{systematic} generation 
+and \textit{ramped half-and-half} generation to be definable
+in the means of simple search strategy.\\
+
+With this apparatus in hands we introduced novel approach
+to term generation by defining the \textit{geometric} search 
+strategy. This strategy is further parameterized by parameter
+$q$,  but since we wanted to avoid suspicion that success
+of this generating method depends on fine-tuning of this
+parameter, we used its default value $q = 0.75$ in every experiment.
+
+After being generated all terms undergo process of \textit{abstraction
+elimination}, which enables our simple tree swapping crossover operation
+which is only genetic operator that we have used in all experiments.
+
+We examined it in three different experiments, which all 
+supported the idea that it has very desirable qualities.
+
+First two experiments, \textit{Simple Symbolic Regression} and 
+\textit{Artificial Ant}, were performed in order to compare
+the \textit{geometric} strategy with the standard GP term 
+generating method \textit{ramped half-and-half}.
+
+In both experiments were observed improvements in the following aspects.
+
+\begin{enumerate}
+ \item Success rate was improved.
+ \item Minimal number of individuals needed to be processed in order to yield 
+       correct solution with probability 99\% was lowered.
+ \item Run time was significantly reduced. 
+ \item Average size of a term was decreased.
+\end{enumerate}
+
+In the case of Artificial Ant problem, all improvements were significant.
+
+These results make me believe that \textit{geometric} strategy might
+be welcomed reinforcements in the fight against the bogey of 
+the GP community --- the \textit{bloat}.
+Or at least, it seems to work for those two experiments in a such way.\\
+
+In the third experiment, the \textit{even-parity} problem, geometric strategy 
+showed ability to yield correct solution in the initial generation
+8 times out of 200 runs (throughout 5 experiments), which is interesting
+result since it is uninformed search.
+
+Five consecutive experiments with even-parity problem supported 
+hypothesis that $\eta$-normalization of generated terms enhances performance.
+We have also seen that use \atTree{}s instead of more traditional \sexprTree{}s
+as tree representation of individuals is able to enhanced performance.
+
+Use of optimized abstraction elimination instead of its basic variant 
+showed significant improvements in time consumption with little or 
+non effects on performance rates.
+
+Implemented system was designed with importance of interactivity in mind,
+resulting in server/client architecture for core/GUI components.
 

### 2013-08-01 tex/thesis.tex
čtvrtek 7:21

@@
 
-Aplikace: například v oboru počítačových her, kde nám často jde o
-zajímavé nelineární chování by se tento přístup k využití typovaného funkcionálního
-programování dal využít.
-
-Předvedený postup by taky ohlo být možné použít na
-
-%\section{Results summary}
+~\\
+\includegraphics[scale=0.33]{fly/4.png}
+
+\newpage
+
+Application of this kind of approach is for example 
+computer games design where non-linearity and adaptability
+to players actions are highly appreciated. 
+Also games are not tight by some fixed specification
+so this approach might be very appropriate in this industry.
+
+~\\[1em]
+
+\includegraphics[scale=0.35]{fly/5.png}
 
@@
 
-\section{Genetic Programming with Polymorphic Types and Higher-Order Functions}
-by Franck Binard and Amy Felty
-
-\begin{todo}
- & dopsat nebo vyhodit
-\end{todo}
+%\section{Genetic Programming with Polymorphic Types and Higher-Order Functions}
+%by Franck Binard and Amy Felty
+%
+%\begin{todo}
+% & dopsat nebo vyhodit
+%\end{todo}
 
@@
 \end{thebibliography}
+
+\appendix
+\chapter{First Appendix}
+
+abc
 

### 2013-08-01 tex/thesis.tex
čtvrtek 9:34

@@
 
-\begin{todo}
-& definovat \textbf{CR} a říct o tom lemátku, že má CR redukce má unikátní nf.
-&& V \cite{barendregt84} (ta nová kniha s žlutejma deskama) je ukázano že je eta CR (Church Rosserovská)
-& zmínit Theorem 2B.4 z \cite{barendregt10}
-\end{todo}
-
 \subsection{$\eta$-reduction}
@@
 
-\begin{todo}
-   & říct že se vždycky zastavěj (strong normalization)
-     && a jaký to má výhody pro nás v GP (duležitý),
-     && potřebný definice pro to uvést v \lterms sekci
-     && musí se říct co je strongly normalizing, 
-        tzn simpli typed $lambda$-kalkulus termy a $\beta$-redukce /
-        $\beta\eta$-redukce
-%     && řekl bych že eta redukce je SN vždycky (tzn i untyped lk)
-%        pač plati (CR+terminující pak SN) a je CR vzhledem k barendrecht-stará-žlutá
-%        a terminující je rpotože zmenšuje term s nemuže zmenšovat donekonečna    
-\end{todo}
+Nice thing about \textit{simply typed lambda calculus} is that
+program $M$ such that \GMS always terminates 
+(provided that implementations of function symbols in $\Gamma$ 
+also always terminate). This fact is particularly handy for
+GP --- we can evaluate bred programs without worrying about hanging
+in the infinite loop forever.
+
+%\begin{todo}
+%   & říct že se vždycky zastavěj (strong normalization)
+%     && a jaký to má výhody pro nás v GP (duležitý),
+%     && potřebný definice pro to uvést v \lterms sekci
+%     && musí se říct co je strongly normalizing, 
+%        tzn simpli typed $lambda$-kalkulus termy a $\beta$-redukce /
+%        $\beta\eta$-redukce
+%%     && řekl bych že eta redukce je SN vždycky (tzn i untyped lk)
+%%        pač plati (CR+terminující pak SN) a je CR vzhledem k barendrecht-stará-žlutá
+%%        a terminující je rpotože zmenšuje term s nemuže zmenšovat donekonečna    
+%\end{todo}
 
@@
        end at some point. Thus every \lterm has \enf.
- \item Since \ered is \textbf{CR}, \enf is unique (see \ref{todo}). 
+ \item Since \ered is \textit{Church–Rosser}, \enf is unique (see \cite{barendregt84}). 
 \end{enumerate}
@@
 is that they are the same representation as S-expressions
-used by standard GP.
+used by standard GP.\\
+
+Again, similarly as for standard version, 
+a maximum permissible depth $D_{created}$ 
+for offspring individuals is defined (e.g. $D_{created} = 17$).
+If one of the offspring has greater depth than this limit, then 
+this offspring is replaced by the first parent in the result of 
+the crossover operator. If both offspring exceeds this limit, than 
+both are replaced by both parents.
+
+For \atTree the $D_{created}$ must be larger since
+\atTree (without lambdas) is a binary tree. This 
+enlargement is approximately proportionate to average number of 
+function arguments. We use generous $D_{created} = 17\times3$.
 
@@
 
-Performance curves are following.
+%Performance curves are following.
 
@@
 The experiment took 107 minutes.\\
+
+\includegraphics[scale=0.5]{reports/Ant/p.png}
 
@@
 \appendix
-\chapter{First Appendix}
-
-abc
+\chapter{Enclosed CD}
+
+The enclosed CD contains implementation source codes,
+installation instructions and some other material, 
+all described in a greater detail in the \texttt{readme.txt}
+file.
 

### 2013-08-01 tex/thesis.tex
čt 1348

@@
 There is one important constrain on function implementations for functions from F:
-There should be type $A$ that for every function symbol $f \in F$ the corresponding function implementation standing behind this symbol should be of a type 
-$A \times ... \times A \rightarrow A$ and should be total (defined on all
+There must be type $A$ that for every function symbol $f \in F$ the corresponding function implementation standing behind this symbol must be of a type 
+$A \times ... \times A \rightarrow A$ and must be total (defined on all
 combinations of inputs). And analogically for $t \in T$ being of a type A.
@@
 
-In nodes with \textit{depth} $\leq d$ the \textit{full}
+In nodes with \textit{depth} $< d$ the \textit{full}
 method selects from the set $F$, whereas the \textit{grow} method 
@@
 
-It is usual to not distinguish between $V$ and $U$ and just use
+It is usual to not distinguish between $V$ and $C$ and just use
 one set of variable names. But because this distinction is quite convenient
@@
 
+This technique (or trick) is called \textit{Currying}.\footnote{More formally,
+currying is process of transforming a
+function which takes multiple inputs
+into function that takes one argument and
+returns function with one less argument.}
+
 \subsection{Statements of the form $M : \sigma$}
@@
    somehow analogous to \textit{Java}
-   expression \texttt{$M$ instanceof $\sigma$}.
+   expression \texttt{$M$ instanceof $\sigma$}.\footnote{
+   This analogy is inadequate in that instanceof is
+   construct of the language, whereas : is symbol
+   out of the language, it enables us to talk about
+   terms from outside perspective.  }
 
@@
 It is not a grammar in classical sense, because we will be operating with infinite sets of nonterminal symbols and rules. Technically it it
-2-level grammer \cite{todo}.\\
+2-level grammer \cite{barendregt10}.\\
 
@@
 
-\begin{todo}
-& skontrolovat ty dukazy pořádně sepsal sem 
-  je dost rychle a mužou bejt divný. 
-\end{todo}
+%\begin{todo}
+%& skontrolovat ty dukazy pořádně sepsal sem 
+%  je dost rychle a mužou bejt divný. 
+%\end{todo}
 
@@
 inhabitation tree ideas to perform systematic generation and other
-kinds of generation of \lterms;
-but maybe even more interesting application of inhabitation trees
-involving some above mentioned ideas will be revealed in the 
-chapter about future work, since those ideas are not yet implemented.
-
-\newpage
-\section{ Inhabitation machine }
-
-\begin{todo}
-  & Když na to zbude čas, tak popsat Inhabitation Machine...
-   && něco jako na závěr ještě ve stručnosti popišme 
-      IM, strukturu ze který je odvozenej IT. 
-  & říct, že je to hezkej konstrukt vzhledem k tomu že je konečnej
-  & ale je popsanej velmi zrychla (v tý knize) a implementační
-     detaily jsou celkem vynechaný
-  & na stromový verzi jsou implementační detaily hezky vidět
-  & a navíc pro polymorfní verzi jsou už i IM nekonečný takže
-     se strácí i ta konečnostní výhoda
-  & dát tam obrázek 
-\end{todo}
+kinds of generation of \lterms.
+%;
+%but maybe even more interesting application of inhabitation trees
+%involving some above mentioned ideas will be revealed in the 
+%chapter about future work, since those ideas are not yet implemented.
+
+%\newpage
+%\section{ Inhabitation machine }
+%
+%\begin{todo}
+%  & Když na to zbude čas, tak popsat Inhabitation Machine...
+%   && něco jako na závěr ještě ve stručnosti popišme 
+%      IM, strukturu ze který je odvozenej IT. 
+%  & říct, že je to hezkej konstrukt vzhledem k tomu že je konečnej
+%  & ale je popsanej velmi zrychla (v tý knize) a implementační
+%     detaily jsou celkem vynechaný
+%  & na stromový verzi jsou implementační detaily hezky vidět
+%  & a navíc pro polymorfní verzi jsou už i IM nekonečný takže
+%     se strácí i ta konečnostní výhoda
+%  & dát tam obrázek 
+%\end{todo}
 
@@
 
-Each generated \lterms $M$ should  satisfy \GMS. \\
+Each generated \lterms $M$ must  satisfy \GMS. \\
 
@@
 standing for unfinished tree of a type $\sigma$ with context $\Gamma$ to use
-as set of building blocks
-\footnote{\textbf{TODO}: Tady to napsat nějak líp s tim gamma }.\\
+as set of building blocks.\\
+%\footnote{\textbf{TODO}: Tady to napsat nějak líp s tim gamma }
 
@@
 
-\subsection{Future improvements}
-
-\begin{todo}
-& !!!
-\end{todo}
+\subsection{Further improvements}
+
+We may easily put all deterministic expansion steps into one step 
+--- by which we mean the lambda head expansion. This could be
+done when expanding atomic type by also expanding its 
+non-atomic arguments. This would result in that only atomic types
+wold by in unfinished leafs. It is important to accordingly change
+the \textit{dist} in A* which is now 1 generally.
+
+It is also worth mentioning that \textit{dist} 1 is also for
+lambda heads with long list of variable names in the current specification --- we do not bother ourselves with this detail.
+But when it is not the desirable behavior it is very easy to change
+by using mor complicated \textit{dist} than 1.
 
@@
 their $filter(nexts,depth)$ 
-and $init()$ methods. Again, in order to avoid confusion, 
+and $init()$ methods\footnote{It is also needed to 
+add forceUniqueness() boolean predicate to the strategy
+since Koza's ramped half and half requires uniqueness.
+Let us skip this step here in this explanation,since it can be easily implemented but adds lot of unnecessary distraction. May the reader forgive.}
+. Again, in order to avoid confusion, 
 \lets stress that $depth$ is really $\lfloor depth/2 \rfloor$ since
 our formal states contain above each symbol node a type node.
-
-\begin{todo}
- & Přidat že pak mužeme čekovat unikátnost při generování, protože už neni zajištěna 
-   jako u systematickýho. Nebo říct že musíme ale pak asi změnit implementaci tý geometrický což znevalidní experimenty
-\end{todo}
 
@@
 
-\begin{todo}
-& eště se musí nějak prohrotit to, že tam je navíc ta lambda hlava
-kerá ještě o něco kazí tu $depth$  - to už uplně neplatí tim že je v
-uvodu popsaný to že se ta hlava ukousne ale stejně to ještě pořádně zkontrolovat
-& ukázat že je to fakt to samý co kozovský klasický generování
-& zavést tam pojem run pro ten jeden vnější while cyklus 
- (i když to je asi jasný z kontextu)
-\end{todo}
+%\begin{todo}
+%& eště se musí nějak prohrotit to, že tam je navíc ta lambda hlava
+%kerá ještě o něco kazí tu $depth$  - to už uplně neplatí tim že je v
+%uvodu popsaný to že se ta hlava ukousne ale stejně to ještě pořádně zkontrolovat
+%& ukázat že je to fakt to samý co kozovský klasický generování
+%& zavést tam pojem run pro ten jeden vnější while cyklus 
+% (i když to je asi jasný z kontextu)
+%\end{todo}
 
@@
 
-\begin{todo}
-
-& sjednotit terminologii nexts successors atd
-
-& Odiskutovat nějak tohle:
-přirozeně chceme při produkci velkýho množství termu obhospodařovávat nějaký
-velký množství rozdělanejch termu v nějaký struktuře a postuipně je dodělávat.
-Proto přichází jako tahle struktura na mysl nějaká prioritní fronta,
-otázka je kde vzít ty priority a A* je jednoduchá ale přitom optimum hledající 
-odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.
-
-& určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
-a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
-udělali jeden dukaz typicky složitýho tvrzení, my sme v situaci kdy
-chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému.
-
-& nekde zmínit že s těma nexts nezacházíme takle dementně, že samozřejmě vybíráme 
-jen z těch přidávanejch podstromu ale že tim nechcem komplikovat popis
-
-\end{todo}
-
-\subsection{Other tried approach to generating individuals}
-
-\begin{todo}
- & popsat tu ten puvodní přístup, kde se bralo přímo 
-   zavedení a eliminace pro každej typovej konstrukt 
-   a podle toho se pak udělali příslušný term generating grammar pravidla
-   kerý odpovídaj manipulacímse strommem
- & ale má nevýhody
-   && hlavně to neni v normální formě, takže generujeme spoustu 
-      ruzně dopočítanejch beta-rovnáse-ekvivalentních termu
-   && u pravidla zavedení implikace neni uplně zřejmý
-      co tam použít za typ
-   && takže sice to na to ze začátku běhalo ale nakonec to bylo vyměněno
-      za IM metodu since ta je jednoduší, líp se přemejšlí o ní,
-      konstrukty jako and a or nemusíme zavádět na urovni pravidel
-      ale stačí tam dat vhodný polymorfní konstruktory/destruktory 
- & nějak tu metodu pomenovat a přemenovat podle toho nadpis
-\end{todo}
+\subsection{Final remarks}
+We should mention that since simply typed lambda calculus
+is equivalent to implicational fragment of intuitionistic logic
+it wold be possible to do term generation by prover for this logic. But logic prover is concerned with finding one proof of a difficult
+proposition, whereas types usually tend to be easy propositions
+but we need enormous amount of them. So this two fields, although 
+equivalent in theory do not have equivalent goals in practice.\\
+
+We have also tried generation directly from grammar rules
+corresponding to basic inference rules -- we did it before
+encountering the work of \cite{barendregt10}, but since this has two
+drawbacks we have not supported the implementation of the former
+term generator to the actual state of source code.
+
+Those drawbacks are that it is not in normal form and that
+it is not obvious which type shoud by added in the grammar 
+rule corresponding to the \textit{$\ar$-elimination}.
+
+%\begin{todo}
+% 
+%& sjednotit terminologii nexts successors atd
+% 
+%& Odiskutovat nějak tohle:
+%přirozeně chceme při produkci velkýho množství termu obhospodařovávat nějaký
+%velký množství rozdělanejch termu v nějaký struktuře a postuipně je dodělávat.
+%Proto přichází jako tahle struktura na mysl nějaká prioritní fronta,
+%otázka je kde vzít ty priority a A* je jednoduchá ale přitom optimum hledající 
+%odpověď na tuhle otázku, dostatečně volná skrz tu heuristiku.  
+%
+%& určitě někde zmínit ospravedlnění proč si napsat dokazovač sám
+%a né požít nějakej už hotovej - de o to, že dokazovače jsou delaný na to aby
+%udělali jeden dukaz typicky složitýho tvrzení, my sme v situaci kdy
+%chceme co nejvíc (a co možná co nejvíc rozmanitejch) dukazu jednoduchýho problému. 
+%
+%& nekde zmínit že s těma nexts nezacházíme takle dementně, že samozřejmě vybíráme 
+%jen z těch přidávanejch podstromu ale že tim nechcem komplikovat popis 
+% 
+%\end{todo}
+
+%\subsection{Other tried approach to generating individuals}
+%
+%\begin{todo}
+% & popsat tu ten puvodní přístup, kde se bralo přímo 
+%   zavedení a eliminace pro každej typovej konstrukt 
+%   a podle toho se pak udělali příslušný term generating grammar pravidla
+%   kerý odpovídaj manipulacímse strommem
+% & ale má nevýhody
+%   && hlavně to neni v normální formě, takže generujeme spoustu 
+%      ruzně dopočítanejch beta-rovnáse-ekvivalentních termu
+%   && u pravidla zavedení implikace neni uplně zřejmý
+%      co tam použít za typ
+%   && takže sice to na to ze začátku běhalo ale nakonec to bylo vyměněno
+%      za IM metodu since ta je jednoduší, líp se přemejšlí o ní,
+%      konstrukty jako and a or nemusíme zavádět na urovni pravidel
+%      ale stačí tam dat vhodný polymorfní konstruktory/destruktory 
+% & nějak tu metodu pomenovat a přemenovat podle toho nadpis
+%\end{todo}
 
@@
 
-Performance curves are following.
-
-\includegraphics[scale=0.65]{reports/SSR/1/probabs.png}
-
-\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}\\
+%Performance curves are following.
+%\includegraphics[scale=0.65]{reports/SSR/1/probabs.png}
+%\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}\\
 
@@
 with 162,500 individuals to be processed.
-
-We have 34\%; minimal $I(M,i,z)$ in generation 23 
-with 192,000. This seems like a pretty good match.
+ This seems like a pretty good match.
 \\
@@
 
-\includegraphics[scale=0.65]{reports/SSR/2/probabs.png}
-
-\includegraphics[scale=0.65]{reports/SSR/2/indivs.png}\\
-
 So our geometric strategy is slightly more successful then 
@@
 
+\includegraphics[scale=0.45]{reports/SSR/p.png}
+
+\includegraphics[scale=0.45]{reports/SSR/i.png}
+
+%\includegraphics[scale=0.65]{reports/SSR/2/probabs.png}
+%\includegraphics[scale=0.65]{reports/SSR/2/indivs.png}\\
+
 \newpage
@@
 The ant is able to move forward, turn left, and sense if a food 
-piece is ahead of him.\\
+piece is ahead of him.
+
+\newpage
 
@@
 %Performance curves are following.
-
-\includegraphics[scale=0.65]{reports/Ant/2/probabs.png}
-
-\includegraphics[scale=0.65]{reports/Ant/2/indivs.png}\\
+%\includegraphics[scale=0.65]{reports/Ant/2/probabs.png}
+%\includegraphics[scale=0.65]{reports/Ant/2/indivs.png}\\
 
@@
 
-\includegraphics[scale=0.5]{reports/Ant/p.png}
-
-\includegraphics[scale=0.65]{reports/Ant/1/probabs.png}
-
-\includegraphics[scale=0.65]{reports/Ant/1/indivs.png}\\
+\includegraphics[scale=0.45]{reports/Ant/p.png}
+
+\includegraphics[scale=0.45]{reports/Ant/i.png}
+
+\newpage
+
+%\includegraphics[scale=0.65]{reports/Ant/1/probabs.png}
+%\includegraphics[scale=0.65]{reports/Ant/1/indivs.png}\\
 
@@
 
-Ramped half-and-half strategy:
-
-\includegraphics[scale=0.65]{reports/Ant/term-sizes-ANT-10behu-ramped.png}
-
-Our geometric strategy:
-
-\includegraphics[scale=0.65]{reports/Ant/term-sizes-ANT-10behu-geom.png}
+\includegraphics[scale=0.45]{reports/Ant/ts.png}
+
+%Ramped half-and-half strategy:
+
+%\includegraphics[scale=0.65]{reports/Ant/term-sizes-ANT-10behu-ramped.png}
+
+%Our geometric strategy:
+
+%\includegraphics[scale=0.65]{reports/Ant/term-sizes-ANT-10behu-geom.png}
 
@@
 
-Ramped half-and-half strategy:
-
-\includegraphics[scale=0.65]{reports/SSR/term-sizes-SSR-10behu-ramped.png}
-
-Our geometric strategy:
-
-\includegraphics[scale=0.65]{reports/SSR/term-sizes-SSR-10behu-geom.png}
+\includegraphics[scale=0.5]{reports/SSR/ts.png}
 
@@
 The reason why examine this problem is that it has been used by many researchers
-as a benchmark for GP \cite{todo,yu01}.
+as a benchmark for GP \cite{yu01}.
 
@@
 The reasons why is this difficult problem 
-for GP to solve are following \cite{todo}:
+for GP to solve are following \cite{yu01}:
 
@@
 
-\begin{todo}
- & zmínit tu, že to pri týyle building block stačí na general solution...
-\end{todo}
+It is interesting that such a few fitness cases are
+sufficient to find general solution (it was so for every
+solution I tried). It comes from the $\Gamma$, the solution tends 
+always be \texttt{foldr} with \textit{xor} defined by the lambda
+abstraction.
+
+%\begin{todo}
+% & zmínit tu, že to pri týyle building block stačí na general solution...
+%\end{todo}
 
@@
 in 17 out of 50 runs, i.e., 34\% success rate.
-The performance curve for $P(M,i)$ is following.
-
-\includegraphics[scale=0.65]{reports/ep/report1/probabs.png}
+
+%The performance curve for $P(M,i)$ is following.
+%\includegraphics[scale=0.65]{reports/ep/report1/probabs.png}
 
@@
 
-The performance curve for $I(M,i,z)$ is following.
-
-\includegraphics[scale=0.65]{reports/ep/report1/indivs.png}
+%The performance curve for $I(M,i,z)$ is following.
+%\includegraphics[scale=0.65]{reports/ep/report1/indivs.png}
 
@@
 
-As we will see, this setup was the most successful one,
-thus we again show the performance curves.
-
-\includegraphics[scale=0.65]{reports/ep/report4/probabs.png}
-
-\includegraphics[scale=0.65]{reports/ep/report4/indivs.png}
+As we will see, this setup was the most successful one.
+%,thus we again show the performance curves.
+
+%\includegraphics[scale=0.65]{reports/ep/report4/probabs.png}
+%\includegraphics[scale=0.65]{reports/ep/report4/indivs.png}
 
@@
 This experiment took only 30 minutes. \\
+
+\includegraphics[scale=0.5]{reports/ep/p.png}
+
+\includegraphics[scale=0.5]{reports/ep/i.png}
 
@@
 
-When we observe results of our experiments, it seems
-that our system is pretty good in term generation, but not so good 
-in crossover. Since the main focus point of this thesis is
-the term generation, it is positive rather then negative observation.
-
 Design of our crossover comes from that we wanted it to be
@@
 kind of $\lambda$ abstraction (i.e., $\lambda$ abstractions that 
-represent the same function arguments to the same higher order function).}
+represent the same function arguments to the same higher order function).} And also variable terminals of Yu's lambda functions
+are constrained to be only the variables of that lambda function, not from some outer scope. This may be advantage for this particular problem because xor does not need any outer variable
+(recall that most common solution is combination of xor and foldr in some similar way as foldr exaple above).
+From an optimist's point of view 34/50 versus 40/50 
+is not such a huge difference when we take to account that
+our general tree-swapping crossover is competing with specialized one.   
 %This seems to be great advantage for optimizing the
@@
 
-From an optimist's point of view 34/50 versus 40/50 
-is not such a huge difference when we take to account that
-our general tree-swapping crossover is competing with specialized one.
-
-Positive observation is that our result at least outperforms 
+Another optimistic observation is that our result at least outperforms 
 other results mentioned in \cite{yu01} with which Yu's method is compared.
@@
 
-\begin{todo}
-& napsat tam jak vypadaj nějaký nalezeneý řešení
-\end{todo}
+%
+%
+%
+%\begin{todo}
+%& napsat tam jak vypadaj nějaký nalezeneý řešení
+%\end{todo}
 
@@
 
-\begin{todo}
- & poznamky o čem se mimojiné zmínit:
-   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
-      jako alternativa k ADF ! Myslim že hodně zajimavý
-\end{todo}
+At this point I have almost complete implementation
+of polymorphic generalization of the system.
+
+This involves extending the type system so that we
+also have type variables.
+
+Lets very briefly sketch the key phases.
+
+Then instead of simple equality checking of the
+current type with return type of the function, we 
+try to fined \textit{most general unification}
+of those two type terms. If it succeeds, than 
+we may use this function symbol (or constant).
+Important part is to deal in the right way wit obtained 
+substitution.
+
+%\begin{todo}
+% & poznamky o čem se mimojiné zmínit:
+%   && $(\$) : (\alpha \ar \beta) \ar \alpha \ar \beta$ 
+%      jako alternativa k ADF ! Myslim že hodně zajimavý
+%\end{todo}
 
@@
 
-\begin{todo}
-& říct že na tu sme se moc nesoustředili ze dvou duvodu
-  ktery to ospravedlňujou 
-	&& koza taky
-	&& ta s vygereovanim podstromu je v zásadě použití
-	   algoritmu na generovaní jedince a tak nám přišlo
-	   lepší pořádně se soustředit na generování jedince a
-	   začít do toho mutaci tahat až bude generování pořádně
-	   prozkoumáno.
-& To ale neznamená že jí nemáme radi, naopak se těšíme
-  že její vhodný zavedení celou metodu ještě zlepší.
-
-& Nápad na to jak to dělat:
-  && nutno ošetřit - co všechno do gammy alemyslim že v phodě když je uplne 
-     stejná jako ta generujicí, pač ve vy SKI-ovanym termu stejne nejsou zadny 
-     proměný 
-  && mít nějakej předem danej počet pohybu na frontě, abych mohl efektivně
-     kontrolovat že to přidání mutací moc nespomalý
-  && Po tom počtu pohybu končim generování
-  && mam nějakej počet vygenerovanejch termu
-  && pro chytrou mutaci mužu chtít třeba 10 termu
-     &&& pokud se mi jich vygenerovalo míň, tak beru míň
-     &&& pokud se mi jich vygenerovalo víc, tak vemu 10 náhodně
-     &&& těch 10 vyhodnotim (jakože dosazenejch na jejich místo)
-     &&& nejlepší si vezmu
-     &&& dobrý udělat to zase najednou pro všechny v tý generaci mutovaný
-         pánčto je nákladný spouštět Hint
-& Další mutace - beta eta redukce (?? asi moc naroční viz dale) 
-  && zmenšující mutace
-  && udělaná že převedem SKIBC do lambda termu a to redukujem? to asi ne pač
-     pak převod to spíš prodlouží, spíš asi mít redukční pravidla pro SKIBC
-     a těma redukovat, ale je otázka jak moc tam budou vznikat SKIBC-redexi
-     - ty mohou vznikat jen při @-tree reprezentaci
-  && to otevírá otázku toho, že by elementy gammy mohla mít k sobě redukční pravidla
-& Další mutace - jednoduše náhrada podstromu/uzlu za nějakej z gammy
-  && nezvětšující mutace
-& Další mutace - mutace číselnejch-like konstant
-& To mě přivádí k myšlence sice pro generování ale myslim podstatný:
-  jakmile jsou v problému číselný-like konstanty tak mužu udělat pro 
-  jeden tvar hodně jedincu - což šetří jak čas na generování, tak to prozkomává 
-  víc danej tvar
-
-\end{todo}
+Since we have term generating method, it is easy to 
+define subtree changing mutation.
+
+Also node changing mutations are easy.
+
+It would be also nice to perform some kind of informed
+(or clever) mutation  - that is mutation that tries more possibilities 
+and selects the best one (i.e. performs a local search).
+
+In the lambda calculus context, it would be interesting to try 
+some reduction based mutations --- simplifying the term
+by bringing it to normal form form which it was dragged 
+away by crossovers and other mutations.
+
+Also mutations of number-like constants would be useful.
+
+%\begin{todo}
+%& říct že na tu sme se moc nesoustředili ze dvou duvodu
+%  ktery to ospravedlňujou 
+%	&& koza taky
+%	&& ta s vygereovanim podstromu je v zásadě použití
+%	   algoritmu na generovaní jedince a tak nám přišlo
+%	   lepší pořádně se soustředit na generování jedince a
+%	   začít do toho mutaci tahat až bude generování pořádně
+%	   prozkoumáno.
+%& To ale neznamená že jí nemáme radi, naopak se těšíme
+%  že její vhodný zavedení celou metodu ještě zlepší.
+%  
+%& Nápad na to jak to dělat:
+%  && nutno ošetřit - co všechno do gammy alemyslim že v phodě když je uplne 
+%     stejná jako ta generujicí, pač ve vy SKI-ovanym termu stejne nejsou zadny 
+%     proměný 
+%  && mít nějakej předem danej počet pohybu na frontě, abych mohl efektivně
+%     kontrolovat že to přidání mutací moc nespomalý
+%  && Po tom počtu pohybu končim generování
+%  && mam nějakej počet vygenerovanejch termu
+%  && pro chytrou mutaci mužu chtít třeba 10 termu
+%     &&& pokud se mi jich vygenerovalo míň, tak beru míň
+%     &&& pokud se mi jich vygenerovalo víc, tak vemu 10 náhodně
+%     &&& těch 10 vyhodnotim (jakože dosazenejch na jejich místo)
+%     &&& nejlepší si vezmu
+%     &&& dobrý udělat to zase najednou pro všechny v tý generaci mutovaný
+%         pánčto je nákladný spouštět Hint
+%& Další mutace - beta eta redukce (?? asi moc naroční viz dale) 
+%  && zmenšující mutace
+%  && udělaná že převedem SKIBC do lambda termu a to redukujem? to asi ne pač
+%     pak převod to spíš prodlouží, spíš asi mít redukční pravidla pro SKIBC
+%     a těma redukovat, ale je otázka jak moc tam budou vznikat SKIBC-redexi
+%     - ty mohou vznikat jen při @-tree reprezentaci
+%  && to otevírá otázku toho, že by elementy gammy mohla mít k sobě redukční pravidla
+%& Další mutace - jednoduše náhrada podstromu/uzlu za nějakej z gammy
+%  && nezvětšující mutace
+%& Další mutace - mutace číselnejch-like konstant
+%& To mě přivádí k myšlence sice pro generování ale myslim podstatný:
+%  jakmile jsou v problému číselný-like konstanty tak mužu udělat pro 
+%  jeden tvar hodně jedincu - což šetří jak čas na generování, tak to prozkomává 
+%  víc danej tvar
+
+%\end{todo}
 
@@
 
-Tento problém je napůl rozpracovaný, hlavně protože by se na něj o dost více
-hodilo použít polymorfní typy, které jsou také ne uplně dodělané a véto thesis
-se o nich zminujeme jen velmi okrajově. Přestose myslím jedná o zajímavou myšlanku
-a proto ho popíšeme.
-
-\section{Roadmap}
-\section{Tree ants}
-\section{[Propojení těch dvou barandrechtskejch odvozovacích pravidel do jedinýho]}
-\section{[Chytřejší heuristika pro A* která si to předpočítá na $\overline{\Gamma}$ ]}
-\section{[šlechtění těch strategií na generování termů]}
-\section{[šlechtění fitness funkcí a nápad s "turnajem olympioniků"]}
-\section{[šlechtění search strategií pro prohledávání]}
-\section{[šlechtění alternative family trees]}
-
-\section{vypisování jedinců tak že nahradíme kombinátory \lterm{}ama a redukujem }
+\textit{Big Context} is an unfinished problem 
+involving breeding of clasical functional functions
+as \texttt{head,tail,map,elem,filter}. We first take
+small sufficient $\Gamma$ for each function to be evolved.
+After that we  fuse those contexts into one big context and check whether the functions are still evolvable. Or we may try to 
+evolve several functions at once by evolving tuples of them.
+
+We have postponed this problem since it is much more suitable
+for polymorphic system.
+
+\section{Parallelisation}
+Since Haskell is very good at Parallelisation 
+and GP is easily parallelised, it comes naturally to 
+connect the two.
+
+%Tento problém je napůl rozpracovaný, hlavně protože by se na něj o dost více
+%hodilo použít polymorfní typy, které jsou také ne uplně dodělané a véto thesis
+%se o nich zminujeme jen velmi okrajově. Přestose myslím jedná o zajímavou myšlanku
+%a proto ho popíšeme.
+
+\section{Future ideas with Inhabitation trees}
+
+I have lot of ideas involving inhabitation trees and 
+informed term generation. Number labeled 
+inhabitation tree may be used as more informed
+version of geometric strategy. 
+Variation of ant colony optimization
+could be used as vehicle for this informed
+term generation. This ACO would differ
+from classical one in that
+it would use trees instead of paths.
+Ant may select one edge (or-node)
+or clone himself (and-node). The global criterion 
+determining the pheromone trail intensity 
+are based upon fitness and local heuristic
+would be some term shortening factor.  \footnote{
+Unfortunately I have run out of time for writing this
+thesis so I must let this favorite theme of mine mostly unexplained.}
+
+%\section{Roadmap}
+%This is idea based on
+%\section{Tree ants}
+
+%\section{[Propojení těch dvou barandrechtskejch odvozovacích pravidel do jedinýho]}
+%\section{[Chytřejší heuristika pro A* která si to předpočítá na $\overline{\Gamma}$ ]}
+%\section{[šlechtění těch strategií na generování termů]}
+%\section{[šlechtění fitness funkcí a nápad s "turnajem olympioniků"]}
+%\section{[šlechtění search strategií pro prohledávání]}
+%\section{[šlechtění alternative family trees]}
+%
+%\section{vypisování jedinců tak že nahradíme kombinátory \lterm{}ama a redukujem }
 
@@
 
-\newpage
-\begin{todo}
- & my tu řikáme že to závadíme kuli typum, ale my děláme v víc! -dáváme tam
-   i abstrakce atd - ale mužeme říct, že tim že se to pak převede zase na 
-   to bez abstrakcí tak setrváváme dal v krasnym bez-lambda abstrakčí ale
-   s tou výhodou že naše metoda gerneruje všechny struktury pro ten danej typ ...
-   nebo tak něco - asi to dát spíš do design of our system nebo někam
-   ale někde to asi zmínit
-\end{todo}
-
-Další témata co zmínit v úvodu nebo v závěru :\\
-
-\begin{easylist}[itemize]
-
-& zmínit že existuje i jinej pohled na to proč tam zavádět typy:
-  && mužem koukat na to tak že konstrejnt je naše TuF tim, že 
-     v ní nemužou bejt libovolný funkce
-  && ale tuhle constrain mužeme omezit tim že konstrjnujeme 
-     právě použití tech bilding bloku, kde constrains jsou daný těma typama
-     - jakoby výměna global constrain za localní konstrejny na tech bilding  
-       blockách
-  && druhá možná ješte silnější motivace je že pomocí typu pak mohou vznikat 
-     smysluplnější programy, tím že dostavají vstupy které přesně čekají
-     a dávají výstupy které mají nějaké očekávané vlastnosti
-
-& Možná se zmínit o tý představě dvou pólu
-  induktivním a deduktivním -
-  klasickym GP a Dokazovačem, čim složitější
-  typovej sytem tim víc se jedná o deduktivní   
-  systém a blíží se to dokazovači.
-   && Ve světle toho, je pak hezky vidět, že
-      je chytrý exhaustivně systematicky prohledávat od nejmenšího
-      po největší- jak budeme zesložiťovat typovej systém, tak
-      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
-      jich generovat kvanta
-
-& že imlementace je delana těma vtipnejamtajpklasama umožnujicima libovolný
-  evoluční struktury na evolvovani
-
-& říct že vnekterejch aspektech je videt, že to je work in progress
-  ale že si myaslime že i tak to ukazuje pěkný výsledky 
-  a že doufame že ty nedodělky budou další prací odstraněny
-
-& Říct jak jde text:
-  && Začnem popsanim klasickýho GP podle Kozy
-  && Vyslovíme definice okolo lambda termu a typu
-  && Vyslovíme definici inhabitačního stromu
-     kterej je stromovou verzí IM od Barendrechta.
-  && Matematicky popíšem jak na to dem my.
-  && Popíšem rozšíření zakladního systému na 
-     polymorfní.
-  && Popíšem to z programátorskýho hlediska
-  && Popíšem příklady na kterejch to demonstrujem
-  && Srovnáme to s nějakejma přístupama co už 
-     existujou
-
-& Nějak naťuknout už tady v úvodu jak se to má při
-  srovnání s tim jak na to dou jiný (Yu atd)
-
-& Výkřiky co zmínit:
-  && Na co zaměřený:
-     &&& Proof of a concept
-     &&& Spíš jednoduchost
-         &&&& Ve většině aspektú to sleduje
-              Kozu - tzn i v málo řídících
-              parametru
-     &&& Duraz na generování termu (!)
-
-  && Na co nezaměřený (ale snad taky ok v tom):
-     &&& Performance   
-     &&& Mature knihovna co by šla použít robustně
-         v cizim kódu - spíš to počítá že to 
-         bude dál rozvíjeno a tohle je mezizastávka
-
- & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
-   defining EA and then specifying the differences or something 
-   like that...
-
-& bylo by asi mimojiný dobrý 
-ospravedlnit/okomentovat tak fancy (=javascript frontend) 
-implementaci při 
-dost nedodělanejch jinejch věcech (heuristika by 
-mohla být lepší, mutace by mohli bejt, atd)
-ale věřil sem, že je to proces iterativní
-a že je po celou dobu dobrý mít na čem to testovat.
-Javascript byl vybranej pač HTML je nejpromáklejší
-UI jaký vubec je na zemi teď (state of dzí art)
-a tak to snad nebyla stráta času.
-navíc se naskitla díky tomu další pěkná příležitost
-k ukázání toho, že to neni omezený na haskell
-ale že se kombinatorový konstrukty daj 
-v zásadě přeložit do libovolnýho jazyka celkem bez problému
-nebo minimalně do JS. Ale todle možná nepsat tolik do závěru
-spíš nějaky podrobnosti přehodit do implementačních detailu
-a tady to jen tak líznout.
-
-\end{easylist}
-
-\textbf{Jen když zbude čas:}
-
-\begin{todo}
-
-& Complexity discussion 
-&& !!!
-&& zmínit že mi moc neřešíme formální/asymptotický složitosti
-  &&& de nám hlavně o realný časy
-&& vyjít z komplexnosti A* asi
-  &&& složitost výrazně ovlivněná kvalitou heuristický funkce
-
- & Brief history of the project\\
-  zmínit třeba tyhle fáze:
-  && že sem si delal evaluator sám co sem pak s performenc duvodu opustil
-  && že sem hledal intuicionistickej dokazovač ale že nejsem nenarazil na vhodnej
-     pač nejsou delany na to plivat hodně dukazu  
-  && že sem nejdriv pouzival ten vypsychlej dokazovač co tam měl nějaký ty 
-     spešl rozhodovací stromy
-  && pak sem to přepsal tak aby to nebylo jen na GP šitý ale obecně na jakoukoliv 
-     evoluci
-  && ...
-
-& pro \ref{benefits}
-&& možná eště zmínit něco jako : \\
-Jedna drobná nevýhoda je, že pokud chceme generovat termy od nejmenšího do největšího
-a pokud lnf generujeme od nejmenšího po největší tak pak generovaní \benf
-timhle trikem nemusí bejt od nejmenšího po nějvětší. V naší aplikaci (tzn GP)
-ale nijak nebazírtujeme na týhle vlastnosti a nakonec budeme spíš stejně používat
-nějaký nedeterminizmy aby to fungovalo líp, takže to vubec nevadí, že budou trochu přeházený.
-
- & pro \ref{barlike}
- && možná zmínit že podobně jako mužeme transformovat inference rules
-   na grammar rules, mužeme i obráceně grammar rules
-   převést na inference rules - to bylo ostatně uděláno v tom dukazu ale
-   je to podle mě hezký říct ty pravidla explicitně.
- & pro \ref{lnf}
- && potřebuju eště nějaký \textbf{propositions} z \cite{barendregt10} tu uvíst ?
- && ta definice je tam daná pro $\tur{}{M}{\sigma}$,
-   ukazat ze ta naše je ok, že tam má navíc to $\Gamma$.
-   Stačí udělat tríček že beru funkci co má lambda hlavu odpovídající 
-   tomu kontextu.
-
-   & k \ref{Term-generating-grammar}
-    &&  buď dokázat ty převody na grammar rules explicitně
-     nebo říct že se to dokáže podobnym stylem jako v dukazu 
-     v sekci \ref{barlike}
-    && možná ocitovat toho van Wijngaarden [1981] s těma 2-level grammar
-     jak je to v Barendregtovi \cite{barendregt10} 
- & možná konec (tree reprez of typed terms) subsekce trochu přeformulovat, je to takový   
-   kostrbatý...
- & o odvoz pravidlech říct
-   && možná zmínit že se jedná o \textit{gentzen style proof}
-   && říct že jsou to modus ponens a jeden směr věty o dedukci  
- & nějak přepsat \ref{GP-prog-trees} aby to bylo ve stejnym stylu jako v \ref{tree-reps}
- & napsat transformace mezi \atTree a \sexprTree
-%& ujistit se že je v části o tree reprezentacích 
-%  je napsaný co to znamená že podstrom má nějakej typ,
-%  že podstromy odpovídaj podvýrazum atd
-% & EXAMPLES of tree representations of \lterms  
-% & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
-%   přímo \sexprTree a říct proč se tomu tak řiká 
-%   (na tom místě co je v textu dřív) 
-% & zkontrolovat že reference tree-reps se odkazuje na dobrý místo,
-%   mohlo by se to rozbít tim že tree reprezentations - typovany/netypovany 
-%   jsou od sebe oddelený
-\end{todo}
+%\newpage
+%\begin{todo}
+% & my tu řikáme že to závadíme kuli typum, ale my děláme v víc! -dáváme tam
+%   i abstrakce atd - ale mužeme říct, že tim že se to pak převede zase na 
+%   to bez abstrakcí tak setrváváme dal v krasnym bez-lambda abstrakčí ale
+%   s tou výhodou že naše metoda gerneruje všechny struktury pro ten danej typ ...
+%   nebo tak něco - asi to dát spíš do design of our system nebo někam
+%   ale někde to asi zmínit
+%\end{todo}
+%
+%
+%Další témata co zmínit v úvodu nebo v závěru :\\
+%
+%
+%
+%\begin{easylist}[itemize]
+%
+%
+% 
+%& zmínit že existuje i jinej pohled na to proč tam zavádět typy:
+%  && mužem koukat na to tak že konstrejnt je naše TuF tim, že 
+%     v ní nemužou bejt libovolný funkce
+%  && ale tuhle constrain mužeme omezit tim že konstrjnujeme 
+%     právě použití tech bilding bloku, kde constrains jsou daný těma typama
+%     - jakoby výměna global constrain za localní konstrejny na tech bilding  
+%       blockách
+%  && druhá možná ješte silnější motivace je že pomocí typu pak mohou vznikat 
+%     smysluplnější programy, tím že dostavají vstupy které přesně čekají
+%     a dávají výstupy které mají nějaké očekávané vlastnosti    
+%
+%& Možná se zmínit o tý představě dvou pólu
+%  induktivním a deduktivním -
+%  klasickym GP a Dokazovačem, čim složitější
+%  typovej sytem tim víc se jedná o deduktivní   
+%  systém a blíží se to dokazovači.
+%   && Ve světle toho, je pak hezky vidět, že
+%      je chytrý exhaustivně systematicky prohledávat od nejmenšího
+%      po největší- jak budeme zesložiťovat typovej systém, tak
+%      čim dál víc bude už stačit najít vubec nějaký řešení a bude čim dál těžší
+%      jich generovat kvanta
+%
+%
+%& že imlementace je delana těma vtipnejamtajpklasama umožnujicima libovolný
+%  evoluční struktury na evolvovani
+%
+%& říct že vnekterejch aspektech je videt, že to je work in progress
+%  ale že si myaslime že i tak to ukazuje pěkný výsledky 
+%  a že doufame že ty nedodělky budou další prací odstraněny
+%
+%
+%& Říct jak jde text:
+%  && Začnem popsanim klasickýho GP podle Kozy
+%  && Vyslovíme definice okolo lambda termu a typu
+%  && Vyslovíme definici inhabitačního stromu
+%     kterej je stromovou verzí IM od Barendrechta.
+%  && Matematicky popíšem jak na to dem my.
+%  && Popíšem rozšíření zakladního systému na 
+%     polymorfní.
+%  && Popíšem to z programátorskýho hlediska
+%  && Popíšem příklady na kterejch to demonstrujem
+%  && Srovnáme to s nějakejma přístupama co už 
+%     existujou 
+%
+%& Nějak naťuknout už tady v úvodu jak se to má při
+%  srovnání s tim jak na to dou jiný (Yu atd)
+%  
+%& Výkřiky co zmínit:
+%  && Na co zaměřený:
+%     &&& Proof of a concept
+%     &&& Spíš jednoduchost
+%         &&&& Ve většině aspektú to sleduje
+%              Kozu - tzn i v málo řídících
+%              parametru
+%     &&& Duraz na generování termu (!)              
+%
+%  && Na co nezaměřený (ale snad taky ok v tom):
+%     &&& Performance   
+%     &&& Mature knihovna co by šla použít robustně
+%         v cizim kódu - spíš to počítá že to 
+%         bude dál rozvíjeno a tohle je mezizastávka
+%
+% & TALK ABOUT GP is part of EA etc. and maybe define the GP by 
+%   defining EA and then specifying the differences or something 
+%   like that...   
+%
+%& bylo by asi mimojiný dobrý 
+%ospravedlnit/okomentovat tak fancy (=javascript frontend) 
+%implementaci při 
+%dost nedodělanejch jinejch věcech (heuristika by 
+%mohla být lepší, mutace by mohli bejt, atd)
+%ale věřil sem, že je to proces iterativní
+%a že je po celou dobu dobrý mít na čem to testovat.
+%Javascript byl vybranej pač HTML je nejpromáklejší
+%UI jaký vubec je na zemi teď (state of dzí art)
+%a tak to snad nebyla stráta času.
+%navíc se naskitla díky tomu další pěkná příležitost
+%k ukázání toho, že to neni omezený na haskell
+%ale že se kombinatorový konstrukty daj 
+%v zásadě přeložit do libovolnýho jazyka celkem bez problému
+%nebo minimalně do JS. Ale todle možná nepsat tolik do závěru
+%spíš nějaky podrobnosti přehodit do implementačních detailu
+%a tady to jen tak líznout.
+%   
+%\end{easylist}
+%
+%
+%\textbf{Jen když zbude čas:}
+%
+%\begin{todo}
+%
+%& Complexity discussion 
+%&& !!!
+%&& zmínit že mi moc neřešíme formální/asymptotický složitosti
+%  &&& de nám hlavně o realný časy
+%&& vyjít z komplexnosti A* asi
+%  &&& složitost výrazně ovlivněná kvalitou heuristický funkce
+%
+%
+% & Brief history of the project\\
+%  zmínit třeba tyhle fáze:
+%  && že sem si delal evaluator sám co sem pak s performenc duvodu opustil
+%  && že sem hledal intuicionistickej dokazovač ale že nejsem nenarazil na vhodnej
+%     pač nejsou delany na to plivat hodně dukazu  
+%  && že sem nejdriv pouzival ten vypsychlej dokazovač co tam měl nějaký ty 
+%     spešl rozhodovací stromy
+%  && pak sem to přepsal tak aby to nebylo jen na GP šitý ale obecně na jakoukoliv 
+%     evoluci
+%  && ...
+%
+%& pro \ref{benefits}
+%&& možná eště zmínit něco jako : \\
+%Jedna drobná nevýhoda je, že pokud chceme generovat termy od nejmenšího do největšího
+%a pokud lnf generujeme od nejmenšího po největší tak pak generovaní \benf
+%timhle trikem nemusí bejt od nejmenšího po nějvětší. V naší aplikaci (tzn GP)
+%ale nijak nebazírtujeme na týhle vlastnosti a nakonec budeme spíš stejně používat
+%nějaký nedeterminizmy aby to fungovalo líp, takže to vubec nevadí, že budou trochu přeházený. 
+%
+% & pro \ref{barlike}
+% && možná zmínit že podobně jako mužeme transformovat inference rules
+%   na grammar rules, mužeme i obráceně grammar rules
+%   převést na inference rules - to bylo ostatně uděláno v tom dukazu ale
+%   je to podle mě hezký říct ty pravidla explicitně.
+% & pro \ref{lnf}
+% && potřebuju eště nějaký \textbf{propositions} z \cite{barendregt10} tu uvíst ?
+% && ta definice je tam daná pro $\tur{}{M}{\sigma}$,
+%   ukazat ze ta naše je ok, že tam má navíc to $\Gamma$.
+%   Stačí udělat tríček že beru funkci co má lambda hlavu odpovídající 
+%   tomu kontextu.
+%   
+%   & k \ref{Term-generating-grammar}
+%    &&  buď dokázat ty převody na grammar rules explicitně
+%     nebo říct že se to dokáže podobnym stylem jako v dukazu 
+%     v sekci \ref{barlike}
+%    && možná ocitovat toho van Wijngaarden [1981] s těma 2-level grammar
+%     jak je to v Barendregtovi \cite{barendregt10} 
+% & možná konec (tree reprez of typed terms) subsekce trochu přeformulovat, je to takový   
+%   kostrbatý...
+% & o odvoz pravidlech říct
+%   && možná zmínit že se jedná o \textit{gentzen style proof}
+%   && říct že jsou to modus ponens a jeden směr věty o dedukci  
+% & nějak přepsat \ref{GP-prog-trees} aby to bylo ve stejnym stylu jako v \ref{tree-reps}
+% & napsat transformace mezi \atTree a \sexprTree
+%%& ujistit se že je v části o tree reprezentacích 
+%%  je napsaný co to znamená že podstrom má nějakej typ,
+%%  že podstromy odpovídaj podvýrazum atd
+%% & EXAMPLES of tree representations of \lterms  
+%% & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
+%%   přímo \sexprTree a říct proč se tomu tak řiká 
+%%   (na tom místě co je v textu dřív) 
+%% & zkontrolovat že reference tree-reps se odkazuje na dobrý místo,
+%%   mohlo by se to rozbít tim že tree reprezentations - typovany/netypovany 
+%%   jsou od sebe oddelený
+%\end{todo}
 

### 2013-08-01 v4/Problems/Ant/Problem.hs

@@ -21,8 +21,8 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_typ         = ant                                          ,
   cttp_ctx         = ctx                                          ,
   
---cttp_gOpt        = CTTG_Geom     ant ctx 0.75 ,  
-  cttp_gOpt        = CTTG_Koza2 ant ctx ,    
+  cttp_gOpt        = CTTG_Geom     ant ctx 0.75 ,  
+--  cttp_gOpt        = CTTG_Koza2 ant ctx ,    
 
   cttp_ff          = FF5 "ff" "Problems.Ant.Funs" (), 
   

### 2013-08-01 v4/Problems/SSR/Problem.hs

@@ -24,7 +24,7 @@ reg = PO_CTTP_ PO_CTTP {
   cttp_typ         = dou1                                         ,
   cttp_ctx         = ctx                                          ,
 
-  cttp_gOpt        = CTTG_Geom  dou1 ctx 0.75 , --CTTG_Geom  dou1 ctx 0.75  , --CTTG_Koza2 dou1 ctx , 
+  cttp_gOpt        = CTTG_Koza2 dou1 ctx , --CTTG_Geom  dou1 ctx 0.75  , --CTTG_Koza2 dou1 ctx , 
   
   cttp_ff          = FF6 (asType::Double->Double) ff "Problems.SSR.Funs", 
   

### 2013-10-13 tex/article.tex
[tex] article

\documentclass[conference]{IEEEtran}
%\usepackage{stmaryrd}
\usepackage{amsfonts}

%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
%\usepackage{hyperref}
%\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage[utf8]{inputenc}

\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}

\IEEEoverridecommandlockouts    % to create the author's affliation portion

\textwidth 178mm    % <------ These are the adjustments we made 10/18/2005
\textheight 239mm   % You may or may not need to adjust these numbers again
\oddsidemargin -7mm
\evensidemargin -7mm
\topmargin -6mm
\columnsep 5mm

\begin{document}

% paper title: Must keep \ \\ \LARGE\bf in it to leave enough margin.
\title{\ \\ \LARGE\bf Typed Functional Genetic Programming}

\author{Tom\'{a}\v{s} K\v{r}en}

\maketitle

\begin{abstract}
In this paper, generalization of the standard genetic programming (GP)
for simply typed lambda calculus is presented. We use population 
initialization method parameterized by simple search strategy. 
First described strategy corresponds to standard ramped half-and-half method, 
second one corresponds to exhaustive systematic search and third one is a 
novel geometric strategy, which outperforms standard method in success 
rate, time consumption and average individual size in two experiments. 
Other performance enhancements based on theory of lambda calculus are 
proposed and supported by experiment. Abstraction elimination is 
utilized to enable use of simple tree-swapping crossover.
\end{abstract}
% no key words

\section{Introduction}

\PARstart{O}{ur} approach aims to play with the full arsenal given by simply typed lambda calculus, thus we begin our reasoning with an exhaustive systematic search
in mind. Our second goal is to construct a system generalizing Standard 
GP \cite{koza92}.
In order to satisfy both these goals the designed system should be 
parameterized by some simple piece of code that makes the difference
between exhaustive systematic search and standard but 
quiet arbitrary ramped half-and-half generating method.

Those two design goals also differentiate our system from 
the three state of the art systems for typed GP known to us.

~\\
\textit{[Kurzívou jsou neformální poznámky]\\\\
\textbf{Yu} - (články: evenParity, polyGP co sem ted našel, ten o burzách co mám v kindlu)
Odlišnosti:
(1) - generování se nedělá systematicky: pokud strom dojde do místa
kde funkci nemá dát jaký parametr, tak místo tý funkce dá nějaký 
terminál.(uvádí 85\% uspěšnost)  
(2) - Ty křížení ma trochu jinak než v tom článku o even parity,
musim zjistit jak má udělaný že se jí nedostane např prom \#3 někam,
kde neni definovaná když dělá přesun podstromu, v tom starym to ale 
bylo myslim založený na tom, že nedovolovala vnější proměnný uvnitř
lambda termu - což platí i nadále. čili vtom to určitě nehraje full deck.
Naopak se víc zaměřuje na polymorfizmus a další věci.\\
\textbf{kanadani}\\ 
- o dost silnější typovej systém\\
- System F, i jim dynamicky vznikaj typy\\
- moc silný typoví systém, takže generování už je dost
  složitý, to se otiskuje v silně nestandardním algoritmu\\
\textbf{kombinátoři} \\
- vůbec nepoužívaj Lambda Abstrakce\\
- univerzální genetickej operator\\
- taky řešej systematický prohledávání\\
\textbf{A eště tak nak všeobecně}\\
- geometrická strategie generování termů jakožto hybrid systematického a náhodného sytylu --- to že jsme si vytičili ty dva cíle který plníme tou 
jednoduchou strategií, tak máme možnost nalízt jednoduchou strategii která 
je na pomezí obou cílů - taková strategie je právě ta naše geometrická 
a ukazuje se že se bohulibě chová právě k strašáku GP -- \textbf{bloatu}.\\
- teoretické LC konstrukty s výhodou použity - @-stromy a eta-redukce\\
- křížení by eliminace 
}

\section{Generating method}

\textit{
- Unifished term je term který může obsahovat jako list uzel tvaru (tau,Gamma), takzvaný unfinished leaf (UL).\\
- Pracujeme s prioritní frontou rozdělaných lambda termů.\\ 
- Priorita je dána počtem UL v termu.\\
- Na počátku je do fronty vložen jediný UT = (tau,gama)\\
- Z fronty vyndám term s nejnižším \#UL, pokud má \#UL = 0, pak je to vygenerovaný jedinec, pokud ne tak vezmu jeho DFS-první UL a expanduju ho. Výsledkem expandování je několik (possibly 0) unfin. termů, které jsou až na expandovaný uzel totožné s původním UT, daný UT je nahrazen novým pod-UT. \\
- Expandování probíhá následovně: \\
(a) je to (a-b,Gama) pak tento list nahradim novým podstromem x.(b;Gama,x:a)\\
(b) je to (alfa,Gama) pak pro každý f:(T1-T2-...-Tn-alpha) el Gama 
nahradíme tento list novým podstromem ( f (T1,Gama) (T2,Gama) .. (Tn,Gama) )
\\\\
- Prohledávací strategie má pak funkci filtru následníků. \\
- Pokud se fronta vyprázdní ještě před vygenerováním požadovaného 
  počtu termů tak zbytek dogenerujeme stejným zpusobem, tzn do 
  fronty přihodíme (tau,gama)  \\
- Systematická strategie nezahodí nic (Pro ní generování 
  kolapsuje do A* algoritmu).\\
- Ramped-half-and-half zahodí vždy vše krom jediného.\\
- Geometrická strategie nezahodí s pravděpodobností q na hloubka UL \\
}

\textbf{$\eta$-normalization ...}

\textit{- pač je to generovany v lnf, neboli beta-eta na -1 nf kde 
eta na -1 je eta expanze tak je chytrý transformovat 
to do beta-eta nf kde  
}

\section{Crossover}
\textit{- design goal: proper generalization of standard tree 
swapping crossover\\ 
- … výceméně zkopčit z diplomky\\
\textbf{@-tree representation ...}\\
...
}
\section{Experiments}

\subsection{Simple Symbolic Regression}
\subsection{Artificial Ant}
\subsection{Even-parity problem}

\section{Conclusions}

\begin{thebibliography}{22}

\bibitem{koza92}
  John R. Koza,
  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
  MIT Press, Cambridge, MA,
  1992.

\bibitem{koza05}
  Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
  \emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
  Springer, 2005. ISBN 978-0-387-26417-2

\bibitem{fg}
 Riccardo Poli, William B. Langdon, Nicholas F. McPhee
 \emph{A Field Guide to Genetic Programming}.
 Lulu Enterprises, UK Ltd, 2008.

\bibitem{yu01}
  T. Yu. 
  \emph{Hierachical processing for evolving recursive and modular 
        programs using higher order functions and lambda abstractions}. 
  Genetic Programming and Evolvable Machines,
  2(4):345–380, December 2001. ISSN 1389-2576.

\bibitem{montana95}
D. J. Montana. 
\emph{Strongly typed genetic programming.} 
Evolutionary Computation, 3(2): 199–230, 1995.
%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha

\bibitem{haynes96}
T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
\emph{Type inheritance in strongly typed genetic programming.} 
In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
in Genetic Programming 2, chapter 18, pages 359–376.
MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1. 
%URL \url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.

\bibitem{olsson94}
J. R. Olsson. 
\emph{Inductive functional programming using incremental program 
transformation and Execution of logic programs by 
iterative-deepening A* SLD-tree search.} 
Dr scient thesis, University of Oslo, Norway, 1994.

\bibitem{kes}
Forrest Briggs, Melissa O’Neill.
\emph{Functional Genetic Programming and Exhaustive
Program Search with Combinator Expressions.}
International Journal of Knowledge-based and Intelligent Engineering Systems,
Volume 12 Issue 1, Pages 47-68, January 2008.

\bibitem{barendregt84}
H. P. Barendregt,
\emph{The Lambda Calculus: its Syntax and Semantics}, 
revised ed., North-Holland, 1984.

\bibitem{barendregt92}
H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
\emph{Lambda Calculi with Types.} 
Handbook of Logic in Computer Science, 1992.

\bibitem{barendregt10}

  Henk Barendregt, Wil Dekkers, Richard Statman,
  \emph{Lambda Calculus With Types}.
  Cambridge University Press,
  2010. 
  %URL \url{http://www.cs.ru.nl/~henk/book.pdf}.

\bibitem{jones87}
Simon Peyton Jones. 
\emph{The Implementation of Functional Programming Languages}. 
Prentice Hall, 1987.

\bibitem{AIAMA}
	Stuart J. Russell, Peter Norvig,
	\emph{Artificial Intelligence: A Modern Approach}.
	Pearson Education,
	2003.

\end{thebibliography}

% that's all folks
\end{document}

### 2013-10-13 tex/article.tex
[tex] article

@@
 %\usepackage{hyperref}
-%\usepackage{amsmath}
+\usepackage{amsmath}
 %\usepackage{amssymb}
 \usepackage[utf8]{inputenc}
+\usepackage{qtree}
+\usepackage{xspace}
+
+\newcommand{\Lets}{Let us\xspace}
+\newcommand{\lets}{let us\xspace}
+\newcommand{\lterm}{$\lambda$-term\xspace}
+\newcommand{\lterms}{$\lambda$-terms\xspace}
+\newcommand{\lhead}{$\lambda$-head\xspace}
+\newcommand{\lheads}{$\lambda$-heads\xspace}
+\newcommand{\la}{\leftarrow\xspace}
+\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
+\newcommand{\tur}[3]{#1\vdash{}#2:#3}
+\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
+\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
+\newcommand{\atTree}{@-tree\xspace}
+\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
+\newcommand{\lh}[1]{\lambda #1}
+\newcommand{\sexprTree}{sexpr-tree\xspace}
+\newcommand{\SexprTree}{Sexpr-tree\xspace}
+\newcommand{\then}{\Rightarrow\xspace}
+\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
+\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
+\newcommand{\ST}{\mathop{\mathrm{ST}}}
+\newcommand{\FV}{\mathop{\mathrm{FV}}}
+\newcommand{\Scomb }{\mathbf{S}}
+\newcommand{\Kcomb }{\mathbf{K}}
+\newcommand{\Icomb }{\mathbf{I}}
+\newcommand{\bbarr}{\twoheadrightarrow_\beta}
+\newcommand{\barr}{\rightarrow_\beta}
+\newcommand{\beq}{=_\beta}
+\newcommand{\eearr}{\twoheadrightarrow_\eta}
+\newcommand{\earr}{\rightarrow_\eta}
+\newcommand{\eeq}{=_\eta}
+\newcommand{\bearr}{\rightarrow_{\beta\eta}}
+\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
+\newcommand{\beeq}{=_{\beta\eta}}
+\newcommand{\etar}{\twoheadrightarrow_\eta}
+\newcommand{\ered}{$\eta$-reduction\xspace}
+\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
+\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
+\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
+\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
+\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
+\newcommand{\bredex}{$\beta$-redex\xspace} 
+\newcommand{\lnf}{\textit{lnf}\xspace}
+\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
+\newcommand{\Bcomb }{\mathbf{B}}   
+\newcommand{\BBcomb }{\mathbf{B*}}
+\newcommand{\Ccomb }{\mathbf{C}}   
+\newcommand{\CCcomb }{\mathbf{C'}}
+\newcommand{\SScomb }{\mathbf{S'}}
+\newcommand{\ar}{\rightarrow\xspace}
+\newcommand{\T}{\mathbb{T}\xspace}
+\newcommand{\Real}{\mathbb{R}}
 
@@
 \section{Crossover}
-\textit{- design goal: proper generalization of standard tree 
-swapping crossover\\ 
-- … výceméně zkopčit z diplomky\\
-\textbf{@-tree representation ...}\\
-...
-}
+
+Design goal behind our approach to crossover operation is
+to try to generalize standard tree swapping crossover.
+The crossover operation in standard GP is performed 
+by swapping randomly selected subtrees in each parent 
+S-expression.
+
+For typed lambda terms two difficulties arise: Types and variables.
+We will show how to crossover typed \lterm trees in both 
+\atTree and \sexprTree notation.
+As in standard GP our crossover will be performed by swapping
+two subtrees. But now with constraint that both subtrees have
+the same type.
+
+Variables bring more difficulties then types do.
+This problem arises from variables that are free in subterms corresponding 
+to swapped subtrees.
+
+Following example illustrates the problem. 
+\Lets have these two parent trees with selected nodes in bold.\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]
+
+~\\The swap of subtrees results in following trees:\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]
+
+~\\The problem is that variable $x_2$ in the second tree
+is not bound by any $\lambda$-head and since
+it is not element of $\Gamma$, the second tree is not well-typed \lterm.
+
+We can avoid dealing with this problem by avoiding use of variables.
+This can be achieved by process called abstraction elimination.
+
+\subsection{Abstraction elimination}
+
+\textit{Abstraction elimination} is a process of transforming 
+an arbitrary \lterm into \lterm that contains no lambda abstractions
+and no bound variables.
+The newly produced \lterm may contain function applications, 
+free symbols from former \lterm and some new symbols standing for 
+combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
+
+Those combinators are defined as:
+\begin{align*}
+\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
+\Kcomb &= \lam{x\,y}{x} \\
+\Icomb &= \lam{x}{x} 
+\end{align*}
+
+\Lets describe transformation $\Ae$ performing this process.
+\begin{align*}
+\Ae[x]           &= x &\\[0.4em]
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
+\Ae[\lam{x}{x}]  &= \Icomb &\\
+\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
+\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
+&\textbf{if } x \in \FV(M)\\
+\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
+&\textbf{if } x \in \FV(M) \\
+&&\vee x \in \FV(N)
+\end{align*}
+
+This is simple version of this process. More optimized version,
+in the means of the size of resulting term and its performance
+is following one, presented in \cite{jones87}.
+
+This version operates with more combinators:
+\begin{align*}
+\Bcomb  &= \lam{f\,g\,x}{f\,(g\,x)} \\
+\Ccomb  &= \lam{f\,g\,x}{f\,x\,g} \\
+\SScomb &= \lam{c\,f\,g\,x}{c\,(f\,x)\,(g\,x)} \\
+\BBcomb &= \lam{c\,f\,g\,x}{c\,(f\,(g\,x))} \\
+\CCcomb &= \lam{c\,f\,g\,x}{c\,(f\,x)\,g} 
+\end{align*}
+
+And the transformation can be written as follows.
+\begin{align*}
+\Ae[x]           &= x &\\
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\
+\Ae[\lam{x}{M}]  &= A[ x ; \Ae[M] ] &
+\\[1em]
+A[x;x]           &= \Icomb &\\
+A[x;y]           &= (\Kcomb~y)\\
+A[x;\,(M\,N)\,]  &= Opt[~\Scomb~(A[x;M])~(A[x;N])~]
+\\[1em]
+Opt[~\Scomb~(\Kcomb~M)~(\Kcomb~N)~]   &= \Kcomb~(M~N)\\
+Opt[~\Scomb~(\Kcomb~M)~\Icomb~]       &= M\\
+Opt[~\Scomb~(\Kcomb~M)~(\Bcomb~N~L)~] &= \BBcomb~M~N~L\\
+Opt[~\Scomb~(\Kcomb~M)~N~]            &= \Bcomb~M~N\\
+Opt[~\Scomb~(\Bcomb~M~N)~(\Kcomb~L)~] &= \CCcomb~M~N~L\\
+Opt[~\Scomb~M~(\Kcomb~N)~]            &= \Ccomb~M~N\\
+Opt[~\Scomb~(\Bcomb~M~N)~L~]          &= \SScomb~M~N~L
+\end{align*}
+
+As is stated in \cite{jones87},
+the biggest disadvantage of this technique is that the translated
+term is often much larger than in its lambda form --- the size of
+the translated term can be proportional to the
+square of the size of the original term.
+
+But the advantage is also tempting --- no need to deal with variables
+and lambda heads.
+
+\subsection{Typed subtree swapping}
+\label{typed-swapping}
+
+First thing to do in standard subtree swapping is to select random node
+in the first parent.
+
+We modify this procedure so that we allow
+selection only of those nodes with such a type that there exists  
+a node in the second parent with the same type.
+
+Standard subtree swapping crossover as a first thing selects 
+whether the selected node will be inner node (usually with probability 
+$p_{ip} = 90\%$) or leaf node (with probability 10\%).
+
+We are in a more complicated situation, because one of those 
+sets may be empty, because of allowing only nodes with possible "partner"
+in the second parent. Thus we do this step only if both sets are
+nonempty.
+
+After selecting a node in the first parent we select node in the
+second parent such that type of that node must by the same as the type 
+of the first node. Again, this may eliminate the "90-10" step of
+first deciding whether the selected node will be internal node 
+or leaf node.
+
+When both nodes are selected we may swap the trees. \\
+
+If the abstraction elimination was performed, then 
+since the trees are of the same type and there are no variables to be 
+moved from their scope, the offspring trees are well typed.\\
+
+Both \sexprTree and \atTree are able to be crossed by this 
+mechanism. But \atTree has more possibilities then \atTree.
+This comes from the fact that every subtree of the \sexprTree
+corresponds to a subtree of \atTree, but there are subtrees
+of \atTree that do not correspond to a subtree of a \sexprTree.\\
+
+Following example should clarify this.
+
+\Tree[.@	
+   [.@ \textbf{f} x ]
+   [.y ]  		 			
+]
+\Tree[.\textbf{f} x y ]~\\
+
+In \atTree, \textbf{f} is leaf thus subtree, 
+whereas in \sexprTree it is internal node thus not a subtree.\\
+
+Another nice property of \sexprTree{}s with no lambdas 
+is that they are the same representation as S-expressions
+used by standard GP.\\
+
+Again, similarly as for standard version, 
+a maximum permissible depth $D_{created}$ 
+for offspring individuals is defined (e.g. $D_{created} = 17$).
+If one of the offspring has greater depth than this limit, then 
+this offspring is replaced by the first parent in the result of 
+the crossover operator. If both offspring exceeds this limit, than 
+both are replaced by both parents.
+
+For \atTree the $D_{created}$ must be larger since
+\atTree (without lambdas) is a binary tree. This 
+enlargement is approximately proportionate to average number of 
+function arguments. We use generous $D_{created} = 17\times3$.
+
 \section{Experiments}
@@
 \subsection{Simple Symbolic Regression}
+
+\textit{Simple Symbolic Regression} is a problem described
+in \cite{koza92}. Objective of this problem is to 
+find a function $f(x)$ that fits a sample
+of twenty given points. The target function is 
+function $f_{t}(x) = x^4 + x^3 + x^2 + x$.
+
+Desired type of generated programs $\sigma$ and 
+building blocks context $\Gamma$ are following.
+\begin{align*}
+\sigma = \Real \ar &\Real\\
+\Gamma = \{
+  (+)  &: \Real \ar \Real \ar \Real    ,\\
+  (-)  &: \Real \ar \Real \ar \Real    ,\\
+  (*)  &: \Real \ar \Real \ar \Real    ,\\
+  rdiv &: \Real \ar \Real \ar \Real    ,\\
+  sin  &: \Real \ar \Real              ,\\
+  cos  &: \Real \ar \Real              ,\\
+  exp  &: \Real \ar \Real              ,\\ 
+  rlog &: \Real \ar \Real              \}
+\end{align*}
+where
+\begin{align*}
+rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
+p/q & \mbox{otherwise } \end{cases}  \\
+rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
+log(\vert x\vert) & \mbox{otherwise}. \end{cases}
+\end{align*}
+
+Fitness function is computed as follows
+
+$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
+
+where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.\\
+
+An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
+considered as a correct individual.
+
 \subsection{Artificial Ant}
@@
 \section{Conclusions}
+
+Our goal was to design and implement a system
+performing GP over some typed functional programming language.
+As this typed functional programming language we have chosen the 
+\textit{simply typed lambda calculus}.
+We have also obligated ourselves to do it in a such way that 
+generalizes the standard GP, rather than
+crates completely new system.
+
+We may say that we have chosen the simplest possible
+option in both areas --- the simply typed calculus and the standard GP.
+We see this decision as fortunate one --- since it enables us add more
+complex features after sufficient exploration of those simplest cases.
+
+We have started the journey toward this goal by study 
+of subject that intersects both those areas:
+Method which for desired type and context generates lambda terms.
+
+The way to obtain this method started with inference rules 
+(the defining concepts of the typed lambda calculus),
+it proceeded through term generating grammars and finished 
+with inhabitation trees.
+
+A* algorithm in combination with inhabitation trees
+has been utilized to drive systematic enumeration
+of typed \lterms in their \lnf. This enumeration was further 
+parameterized by simple search strategy in order to enable
+such different approaches as \textit{systematic} generation 
+and \textit{ramped half-and-half} generation to be definable
+in the means of simple search strategy.\\
+
+With this apparatus in hands we introduced novel approach
+to term generation by defining the \textit{geometric} search 
+strategy. This strategy is further parameterized by parameter
+$q$,  but since we wanted to avoid suspicion that success
+of this generating method depends on fine-tuning of this
+parameter, we used its default value $q = 0.75$ in every experiment.
+
+After being generated all terms undergo process of \textit{abstraction
+elimination}, which enables our simple tree swapping crossover operation
+which is only genetic operator that we have used in all experiments.
+
+We examined it in three different experiments, which all 
+supported the idea that it has very desirable qualities.
+
+First two experiments, \textit{Simple Symbolic Regression} and 
+\textit{Artificial Ant}, were performed in order to compare
+the \textit{geometric} strategy with the standard GP term 
+generating method \textit{ramped half-and-half}.
+
+In both experiments were observed improvements in the following aspects.
+
+\begin{enumerate}
+ \item Success rate was improved.
+ \item Minimal number of individuals needed to be processed in order to yield 
+       correct solution with probability 99\% was lowered.
+ \item Run time was significantly reduced. 
+ \item Average size of a term was decreased.
+\end{enumerate}
+
+In the case of Artificial Ant problem, all improvements were significant.
+
+These results make me believe that \textit{geometric} strategy might
+be welcomed reinforcements in the fight against the bogey of 
+the GP community --- the \textit{bloat}.
+Or at least, it seems to work for those two experiments in a such way.\\
+
+In the third experiment, the \textit{even-parity} problem, geometric strategy 
+showed ability to yield correct solution in the initial generation
+8 times out of 200 runs (throughout 5 experiments), which is interesting
+result since it is uninformed search.
+
+Five consecutive experiments with even-parity problem supported 
+hypothesis that $\eta$-normalization of generated terms enhances performance.
+We have also seen that use \atTree{}s instead of more traditional \sexprTree{}s
+as tree representation of individuals is able to enhanced performance.
+
+Use of optimized abstraction elimination instead of its basic variant 
+showed significant improvements in time consumption with little or 
+non effects on performance rates.
+
+Implemented system was designed with importance of interactivity in mind,
+resulting in server/client architecture for core/GUI components.
 

### 2013-10-14 tex/article.tex
tex

@@
 \textit{[Kurzívou jsou neformální poznámky]\\\\
-\textbf{Yu} - (články: evenParity, polyGP co sem ted našel, ten o burzách co mám v kindlu)
+\textbf{Yu} - (články: evenParity, polyGP
+\textbf{[doplnit do citací]}
+ co sem ted našel, ten o burzách co mám v kindlu)
 Odlišnosti:
@@
 \textit{
-- Unifished term je term který může obsahovat jako list uzel tvaru (tau,Gamma), takzvaný unfinished leaf (UL).\\
+- Unifished term je term který může obsahovat jako list uzel tvaru 
+$(\tau,\Gamma)$, takzvaný unfinished leaf (UL).\\
 - Pracujeme s prioritní frontou rozdělaných lambda termů.\\ 
 - Priorita je dána počtem UL v termu.\\
-- Na počátku je do fronty vložen jediný UT = (tau,gama)\\
+- Na počátku je do fronty vložen jediný UT = $(\tau,\Gamma)$\\
 - Z fronty vyndám term s nejnižším \#UL, pokud má \#UL = 0, pak je to vygenerovaný jedinec, pokud ne tak vezmu jeho DFS-první UL a expanduju ho.
@@
 - Expandování probíhá následovně: \\
-(a) je to (a-b,Gama) pak tento list nahradim novým podstromem x.(b;Gama,x:a)\\
-(b) je to (alfa,Gama) pak pro každý f:(T1-T2-...-Tn-alpha) el Gama 
-nahradíme tento list novým podstromem ( f (T1,Gama) (T2,Gama) .. (Tn,Gama) )
+(a) je to  $(\sigma \ar \tau,\Gamma)$ pak tento list nahradim novým podstromem $\lam{x}{(\tau;\Gamma,x:a)}$\\
+(b) je to $(\alpha,\Gamma)$ pak pro každý\\ 
+$f : (\tau_1 \ar \dots \ar \tau_n \ar \alpha) \in \Gamma$ \\
+nahradíme tento list novým podstromem \\
+$(~f~(\tau_1,\Gamma)~\dots~(\tau_n,\Gamma)~)$
 \\\\
@@
   počtu termů tak zbytek dogenerujeme stejným zpusobem, tzn do 
-  fronty přihodíme (tau,gama)  \\
+  fronty přihodíme $(\tau,\Gamma)$ a jedem dál \\
 - Systematická strategie nezahodí nic (Pro ní generování 
@@
 eta na -1 je eta expanze tak je chytrý transformovat 
-to do beta-eta nf kde  
+to do beta-eta nf. To stačí opakovanou eta redukcí, protože beta normálnost
+se neporuší (asi ve zkratce uvést proč, pač je to celkem přímočarý) 
 }
@@
 \section{Crossover}
+
+\textbf{[Odtud dál jsou zkrácené pasáže z diplomky,co mi 
+přišli celkem ok, ale ještě budou určitě editováný 
+(zkracování, doplňování, vhodnější formulace,...)]\\}
 

### 2013-10-14 fishtron.js/js/ants.js
js ants

@@ -1,0 +1,12 @@
+
+var defaultOpts = {
+  rho : 0.5 ,
+
+};
+
+
+function updateFeromon( oldTau , deltaTaus , opts ){
+  return (1-opts.rho)*oldTau + sum(deltaTaus); // newTau
+}
+
+

### 2013-10-14 fishtron.js/js/tests.js

@@ -1,5 +1,13 @@
 var fishtronTests = function(){
 
+
+  test( 'utils', function(){
+
+    strictEqual( sum([1,2,3,1000]), 1006 ,'sum (1)' );
+    strictEqual( sum([])          , 0    ,'sum (2)' );
+
+
+  });
     
     test( 'PriorityQueue', function(){
 
@@ -219,6 +227,14 @@ var fishtronTests = function(){
 
       console.timeEnd('tests');
     });
+
+  test( 'ants', function(){
+
+    ok(true,'fake test...' );
+
+  });
+
+  
 
 };
 

### 2013-10-14 fishtron.js/js/utils.js

@@ -30,6 +30,10 @@ var cons = function(x,xs){
 };
 
 
+function sum(xs){
+  return _.reduce(xs,function(x,y){return x+y;},0);  
+}
+
 
 var partition = function(mustBeTrue,array){
   assert(_.isArray(array),'partition : the array argument mus be an array.');

### 2013-10-18 fishtron.js/js/ants.js
mkDist (ntbk v buse)

@@ -152,7 +152,8 @@ function selectSucc( from , path , tau , succsFun , heur , opts ){
 
   for( i = 0 ; i < ss.length ; i++ ){
     var to = ss[i];
-    var probab = Math.pow( tau[from][to], opts.alpha) * Math.pow( heur(from,to), opts.beta);
+    var probab = Math.pow( tau[from][to], opts.alpha) 
+               * Math.pow( heur(from,to), opts.beta);
     ps.push( probab );
     pSum += probab;
   }
@@ -175,15 +176,11 @@ function mkPath( from , tau , succsFun , heur , isGoal , opts ){
   var path = [];
 
   while( !isGoal(path) && next !== null ){
-
     path.push(next);
     var next = selectSucc( next , path , tau , succsFun , heur , opts );  
-    
   }
 
   return isGoal(path) ? path : null ;
-
-  
 }
 
 

### 2013-10-18 fishtron.js/js/eva.js

@@ -44,9 +44,30 @@ function gp(opts){
 function mkDist( distArr_ ){
   
   var distArr = distArr_;
+  var sum = 0;
+
+  for(var i = 0; i < distArr.length; i++){
+    sum += distArr[i][1];
+  }
+
+  var get = function(){
+
+    var ball = Math.random() * sum;
+    var sumNow = 0;
+    var i;
+    for(i = 0; i < distArr.length; i++){
+      sumNow += distArr[i][1];
+      if( ball < sumNow ){
+        break;
+      }
+    }
+
+    return distArr[i][0];
+
+  };
 
   return {
-
+    get : get
   };
 }
 

### 2013-10-21 tex/evogp.tex
evogp stub

--- 2013-10-13 tex/article.tex
@@
-\documentclass[conference]{IEEEtran}
-%\usepackage{stmaryrd}
+\documentclass{llncs}
+
+\usepackage[utf8]{inputenc}
 \usepackage{amsfonts}
+\usepackage{qtree}
+\usepackage{xspace}
 
+%\usepackage{makeidx}  % allows for indexgeneration
 %\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
 %\usepackage{hyperref}
-%\usepackage{amsmath}
+\usepackage{amsmath}
 %\usepackage{amssymb}
-\usepackage[utf8]{inputenc}
+
+\newcommand{\Lets}{Let us\xspace}
+\newcommand{\lets}{let us\xspace}
+\newcommand{\lterm}{$\lambda$-term\xspace}
+\newcommand{\lterms}{$\lambda$-terms\xspace}
+\newcommand{\lhead}{$\lambda$-head\xspace}
+\newcommand{\lheads}{$\lambda$-heads\xspace}
+\newcommand{\la}{\leftarrow\xspace}
+\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
+\newcommand{\tur}[3]{#1\vdash{}#2:#3}
+\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
+\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
+\newcommand{\atTree}{@-tree\xspace}
+\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
+\newcommand{\lh}[1]{\lambda #1}
+\newcommand{\sexprTree}{sexpr-tree\xspace}
+\newcommand{\SexprTree}{Sexpr-tree\xspace}
+\newcommand{\then}{\Rightarrow\xspace}
+\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
+\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
+\newcommand{\ST}{\mathop{\mathrm{ST}}}
+\newcommand{\FV}{\mathop{\mathrm{FV}}}
+\newcommand{\Scomb }{\mathbf{S}}
+\newcommand{\Kcomb }{\mathbf{K}}
+\newcommand{\Icomb }{\mathbf{I}}
+\newcommand{\bbarr}{\twoheadrightarrow_\beta}
+\newcommand{\barr}{\rightarrow_\beta}
+\newcommand{\beq}{=_\beta}
+\newcommand{\eearr}{\twoheadrightarrow_\eta}
+\newcommand{\earr}{\rightarrow_\eta}
+\newcommand{\eeq}{=_\eta}
+\newcommand{\bearr}{\rightarrow_{\beta\eta}}
+\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
+\newcommand{\beeq}{=_{\beta\eta}}
+\newcommand{\etar}{\twoheadrightarrow_\eta}
+\newcommand{\ered}{$\eta$-reduction\xspace}
+\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
+\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
+\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
+\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
+\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
+\newcommand{\bredex}{$\beta$-redex\xspace} 
+\newcommand{\lnf}{\textit{lnf}\xspace}
+\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
+\newcommand{\Bcomb }{\mathbf{B}}   
+\newcommand{\BBcomb }{\mathbf{B*}}
+\newcommand{\Ccomb }{\mathbf{C}}   
+\newcommand{\CCcomb }{\mathbf{C'}}
+\newcommand{\SScomb }{\mathbf{S'}}
+\newcommand{\ar}{\rightarrow\xspace}
+\newcommand{\T}{\mathbb{T}\xspace}
+\newcommand{\Real}{\mathbb{R}}
 
@@
 
-\IEEEoverridecommandlockouts    % to create the author's affliation portion
-
-\textwidth 178mm    % <------ These are the adjustments we made 10/18/2005
-\textheight 239mm   % You may or may not need to adjust these numbers again
-\oddsidemargin -7mm
-\evensidemargin -7mm
-\topmargin -6mm
-\columnsep 5mm
-
 \begin{document}
 
-% paper title: Must keep \ \\ \LARGE\bf in it to leave enough margin.
-\title{\ \\ \LARGE\bf Typed Functional Genetic Programming}
-
-\author{Tom\'{a}\v{s} K\v{r}en}
-
+\title{\textit{[něco jako?]} 
+Generating lambda term individuals using undisciplined A*}
+\author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}
+\institute{Matfyz}
 \maketitle
@@
 \begin{abstract}
+\textit{(Aby tu něco bylo, pak se udělá lepší (trochu upravenej abstract z diplomky))}
 In this paper, generalization of the standard genetic programming (GP)
@@
 \end{abstract}
-% no key words
 
@@
 
-\PARstart{O}{ur} approach aims to play with the full arsenal given by simply typed lambda calculus, thus we begin our reasoning with an exhaustive systematic search
+\subsection{gp a typy jsou dobrý / motivace}
+...
+\subsection{příběh}
+
+Our approach aims to play with the full arsenal given by simply typed lambda calculus, thus we begin our reasoning with an exhaustive systematic search
 in mind. Our second goal is to construct a system generalizing Standard 
@@
 ~\\
-\textit{[Kurzívou jsou neformální poznámky]\\\\
-\textbf{Yu} - (články: evenParity, polyGP co sem ted našel, ten o burzách co mám v kindlu)
+\textbf{poznamky naky}\textit{\\
+- geometrická strategie generování termů jakožto hybrid systematického a náhodného sytylu --- to že jsme si vytičili ty dva cíle který plníme tou 
+jednoduchou strategií, tak máme možnost nalízt jednoduchou strategii která 
+je na pomezí obou cílů - taková strategie je právě ta naše geometrická 
+a ukazuje se že se bohulibě chová právě k strašáku GP -- \textbf{bloatu}.\\
+- teoretické LC konstrukty s výhodou použity - @-stromy a eta-redukce\\
+- křížení by eliminace 
+}
+
+\subsection{obsah kapitol článku v 1 větě}
+...
+
+\section{Related work}
+\subsection{Yu}
+\textbf{Yu} - (články: evenParity, polyGP
+\textbf{[doplnit do citací]}
+ co sem ted našel, ten o burzách co mám v kindlu)
 Odlišnosti:
@@
 lambda termu - což platí i nadále. čili vtom to určitě nehraje full deck.
-Naopak se víc zaměřuje na polymorfizmus a další věci.\\
+Naopak se víc zaměřuje na polymorfizmus a další věci.
+
+\subsection{kombinatori}
+
+\textbf{kombinátoři} \\
+- vůbec nepoužívaj Lambda Abstrakce\\
+- univerzální genetickej operator\\
+- taky řešej systematický prohledávání
+
+\subsection{kanadani}
+
 \textbf{kanadani}\\ 
@@
 - moc silný typoví systém, takže generování už je dost
-  složitý, to se otiskuje v silně nestandardním algoritmu\\
-\textbf{kombinátoři} \\
-- vůbec nepoužívaj Lambda Abstrakce\\
-- univerzální genetickej operator\\
-- taky řešej systematický prohledávání\\
-\textbf{A eště tak nak všeobecně}\\
-- geometrická strategie generování termů jakožto hybrid systematického a náhodného sytylu --- to že jsme si vytičili ty dva cíle který plníme tou 
-jednoduchou strategií, tak máme možnost nalízt jednoduchou strategii která 
-je na pomezí obou cílů - taková strategie je právě ta naše geometrická 
-a ukazuje se že se bohulibě chová právě k strašáku GP -- \textbf{bloatu}.\\
-- teoretické LC konstrukty s výhodou použity - @-stromy a eta-redukce\\
-- křížení by eliminace 
-}
+  složitý, to se otiskuje v silně nestandardním algoritmu
 
-\section{Generating method}
+\subsection{barendregt}
+
+\section{Preliminaries}
+%definice, pojmy
+
+\subsection{Term}
+
+\subsection{Type}
+
+\subsection{Context}
+
+\subsection{Unfinished term}
+
+\textit{- Unifished term je term který může obsahovat jako list uzel tvaru 
+$(\tau,\Gamma)$, takzvaný unfinished leaf (UL).\\}
+
+\section{Our approach}
+\subsection{Introduction}
+úvod ke kapitole - systematicky, A*, filtrace, znova dám do fronty začátek
+
+\subsection{Generating algorithm}
+
+The inputs for the term generating algorithm are following.
+\begin{enumerate}
+ \item Desired type $\tau$ of generated terms.
+ \item Context $\Gamma$ representing set of building symbols.
+ \item Number $n$ of terms to be generated.
+\end{enumerate}
 
 \textit{
-- Unifished term je term který může obsahovat jako list uzel tvaru (tau,Gamma), takzvaný unfinished leaf (UL).\\
 - Pracujeme s prioritní frontou rozdělaných lambda termů.\\ 
 - Priorita je dána počtem UL v termu.\\
-- Na počátku je do fronty vložen jediný UT = (tau,gama)\\
+- Na počátku je do fronty vložen jediný UT = $(\tau,\Gamma)$\\
 - Z fronty vyndám term s nejnižším \#UL, pokud má \#UL = 0, pak je to vygenerovaný jedinec, pokud ne tak vezmu jeho DFS-první UL a expanduju ho.
@@
 eta na -1 je eta expanze tak je chytrý transformovat 
-to do beta-eta nf kde  
+to do beta-eta nf. To stačí opakovanou eta redukcí, protože beta normálnost
+se neporuší (asi ve zkratce uvést proč, pač je to celkem přímočarý) 
 }
 
-\section{Crossover}
-\textit{- design goal: proper generalization of standard tree 
-swapping crossover\\ 
-- … výceméně zkopčit z diplomky\\
-\textbf{@-tree representation ...}\\
+\subsection{Crossover???}
+
+\section{Experiments}
+\subsection{Simple symbolic regression}
+
+\textit{Simple Symbolic Regression} is a problem described
+in \cite{koza92}. Objective of this problem is to 
+find a function $f(x)$ that fits a sample
+of twenty given points. The target function is 
+function $f_{t}(x) = x^4 + x^3 + x^2 + x$.
+
+Desired type of generated programs $\sigma$ and 
+building blocks context $\Gamma$ are following.
+\begin{align*}
+\sigma = \Real \ar &\Real\\
+\Gamma = \{
+  (+)  &: \Real \ar \Real \ar \Real    ,
+  (-)   : \Real \ar \Real \ar \Real    ,
+  (*)   : \Real \ar \Real \ar \Real    ,
+  rdiv  : \Real \ar \Real \ar \Real    ,\\
+  sin  &: \Real \ar \Real              ,
+  cos   : \Real \ar \Real              ,
+  exp   : \Real \ar \Real              , 
+  rlog  : \Real \ar \Real              \}
+\end{align*}
+where
+\begin{align*}
+rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
+p/q & \mbox{otherwise } \end{cases}  \\
+rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
+log(\vert x\vert) & \mbox{otherwise}. \end{cases}
+\end{align*}
+
+Fitness function is computed as follows
+
+$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
+
+where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.\\
+
+An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
+considered as a correct individual.
+
+\subsection{Artificial ant}
 ...
-}
-\section{Experiments}
+\subsection{Even parity problem}
+...
+\section{Conclusions}
+...
 
-\subsection{Simple Symbolic Regression}
-\subsection{Artificial Ant}
-\subsection{Even-parity problem}
-
-\section{Conclusions}
-
-\begin{thebibliography}{22}
+\begin{thebibliography}{1}
 
@@
 
-% that's all folks
 \end{document}

### 2013-10-22 tex/evogp.tex
evogp tex

@@
 \newcommand{\Lp}  {\Lambda^{\prime}\xspace}
-\newcommand{\tur}[3]{#1\vdash{}#2:#3}
+\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
 \newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
@@
 
+\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
+
 \hyphenation{op-tical net-works semi-conduc-tor IEEEtran}
@@
 \author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}
-\institute{Matfyz}
+\institute{Matfyz ...}
 \maketitle
@@
 \begin{abstract}
-\textit{(Aby tu něco bylo, pak se udělá lepší (trochu upravenej abstract z diplomky))}
+\textbf{\textit{
+(Aby tu něco bylo, pak se udělá lepší 
+(tohle je trochu upravenej abstract z diplomky))}}
 In this paper, generalization of the standard genetic programming (GP)
@@
 
+(...)\\
+
 Our approach aims to play with the full arsenal given by simply typed lambda calculus, thus we begin our reasoning with an exhaustive systematic search
@@
 the three state of the art systems for typed GP known to us.
+
+~\\(...)
 
@@
 
+\subsection{Ramped half-and-half}
+
 \subsection{Term}
+
+\subsubsection{Reductions}
+\subsubsection{Tree representations}
+%\subsubsection{}
 
@@
 
+\textit{--- nezapomenout popsat $\Gamma,x:\tau$ notaci
+ a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
+ tzn je nová...}
+
+\subsection{\GMS}
+
 \subsection{Unfinished term}
@@
 \textit{- Unifished term je term který může obsahovat jako list uzel tvaru 
-$(\tau,\Gamma)$, takzvaný unfinished leaf (UL).\\}
+$(\tau,\Gamma)$, takzvaný unfinished leaf (UL).\\
+-Vysvětlit unfinished leaf.\\}
 
@@
  \item Number $n$ of terms to be generated.
+ \item Search strategy $S$. 
 \end{enumerate}
 
+Essential data structure of our term generating algorithm 
+is priority queue of unfinished terms. 
+Priority of an unfinished term is given by 
+the number of unfinished leafs in that term.
+Initially, the queue contains only one unfinished term; 
+$\ul{\tau}{\Gamma}$.\\
+
+At each step, the term $M$ with the lowest number of 
+unfinished leafs is pulled from the queue.
+According to the actual number of those leafs one of
+the following actions is performed.
+\begin{enumerate}
+ \item If the term has no unfinished leaf (i.e., it is a finished
+ term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
+ result collection of generated terms.   
+ \item Otherwise, \textit{successors} of the unfinished term $M$ are
+       filtered out by \textit{search strategy} $S$ and
+       those successors that outlast the filtration 
+       are inserted into the queue.
+\end{enumerate}
+
+\textit{Successors} of an unfinished term $M$ are obtained by 
+\textit{expansion} of the \mbox{\textit{DFS-first}} unfinished leaf $L$
+(i.e., the leftmost unfinished leaf of $M$).\\
+
+Expansion of the selected unfinished leaf $L$ leads to creation of 
+one or many (possibly zero) successors.
+In this process, $L$ is replaced
+by a new subterm defined by the following rules.
+\begin{enumerate}
+ \item If $L = \ul{\rho \ar \sigma}{\Gamma}$, 
+       then $L$ is replaced by $\lamb{x}{\ul{\sigma}{\Gamma,x \colon \rho}}$.
+       Thus this expansion results in exactly one successor.  
+ \item If $L = \ul{\alpha}{\Gamma}$ where $\alpha$ is \textit{atomic} type,
+       then for each 
+       \mbox{$f : (\tau_1 \ar \dots \ar \tau_n \ar \alpha) \in \Gamma$}
+       the unfinished leaf $L$ is replaced by 
+       $(~f~(\tau_1,\Gamma)~\dots~(\tau_n,\Gamma)~)$.
+       Thus this expansion results in many (possibly zero or one) successors.
+\end{enumerate}
+
+Now that we have all possible successors of $M$, we are about to apply
+the \textit{search strategy} $S$. A search strategy is a procedure
+which takes as input a set of unfinished terms and returns a subset
+of the input set. Therefore, search strategy acts as a filter reducing 
+the search space.
+
+If the queue becomes empty before the desired number $n$ of terms
+is generated, then the initial unfinished term $\ul{\tau}{\Gamma}$ 
+is inserted to the queue and the process continues.
+
+\Lets now discuss three such search strategies.
+
+\subsubsection{Systematic strategy}
+\subsubsection{Ramped half-and-half strategy}
+\subsubsection{Geometric strategy}
+
+~\\\\
 \textit{
-- Pracujeme s prioritní frontou rozdělaných lambda termů.\\ 
-- Priorita je dána počtem UL v termu.\\
-- Na počátku je do fronty vložen jediný UT = $(\tau,\Gamma)$\\
-- Z fronty vyndám term s nejnižším \#UL, pokud má \#UL = 0, pak je to vygenerovaný jedinec, pokud ne tak vezmu jeho DFS-první UL a expanduju ho.
-Výsledkem expandování je několik (possibly 0) unfin.
-termů, které jsou až na expandovaný uzel totožné s původním UT, daný UT je nahrazen novým pod-UT.
-\\
-- Expandování probíhá následovně: \\
-(a) je to  $(\sigma \ar \tau,\Gamma)$ pak tento list nahradim novým podstromem $\lam{x}{(\tau;\Gamma,x:a)}$\\
-(b) je to $(\alpha,\Gamma)$ pak pro každý\\ 
-$f : (\tau_1 \ar \dots \ar \tau_n \ar \alpha) \in \Gamma$ \\
-nahradíme tento list novým podstromem \\
-$(~f~(\tau_1,\Gamma)~\dots~(\tau_n,\Gamma)~)$
-\\\\
-- Prohledávací strategie má pak funkci filtru následníků. \\
-- Pokud se fronta vyprázdní ještě před vygenerováním požadovaného 
-  počtu termů tak zbytek dogenerujeme stejným zpusobem, tzn do 
-  fronty přihodíme $(\tau,\Gamma)$ a jedem dál \\
 - Systematická strategie nezahodí nic (Pro ní generování 
@@
 
-\textbf{$\eta$-normalization ...}
+\subsection{Discussion and further improvements}
+
+\subsection{$\eta$-normalization}
 
@@
 
-\subsection{Crossover???}
+\subsection{Crossover}
+
+\textit{Zmínil bych jí tu ale jen hodně rychle...}
 
@@
 where
+
+\noindent\begin{minipage}{.5\linewidth}
 \begin{align*}
 rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
-p/q & \mbox{otherwise } \end{cases}  \\
+p/q & \mbox{otherwise } \end{cases}  
+\end{align*}
+\end{minipage}%
+\begin{minipage}{.5\linewidth}
+\begin{align*}
 rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
@@
 \end{align*}
+\end{minipage}
 
+~\\\\
 Fitness function is computed as follows

### 2013-10-22 tex/evogp.tex
evogp tex

@@
 %\usepackage{amssymb}
+
+\usepackage[ampersand]{easylist}
+\usepackage{color}
 
@@
 \newcommand{\T}{\mathbb{T}\xspace}
+\newcommand{\C}{\mathbb{C}\xspace}
 \newcommand{\Real}{\mathbb{R}}
 
+\newenvironment{todo}
+{ ~\\[0.5em]
+  {\color{red}\textbf{TODO}}
+  \begin{easylist}[itemize]}
+{ \end{easylist}
+  ~}
+
+\newcommand{\Lpr}{\Lambda^\prime}
 \newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
@@
 \title{\textit{[něco jako?]} 
-Generating lambda term individuals using undisciplined A*}
+Generating lambda term individuals using 
+%undisciplined 
+forgetful
+A*}
 \author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}
@@
 
+Here we  %formally 
+describe programming language, 
+in which we generate individual programs --- so called \lterms.
+
+\begin{definition}
+Let $V$ be infinite countable set of {\it 
+variable names}. Let $C$ be set of {\it constant names}, 
+$V \cap C = \emptyset$.	 	
+Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
+\begin{align*}
+x   \in V \cup C  &\then x     \in \Lambda \\
+M,N \in \Lambda   &\then (M~N) \in \Lambda 
+\textit{~~~~~~(Function application)} \\
+x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
+\textit{~~~~($\lambda$-abstraction)} 
+\end{align*}
+\end{definition}
+
+\textit{Function application} and 
+\textit{$\lambda$-abstraction} are concepts
+well known from common programming languages. 
+For example in JavaScript 
+$(M~N)$ translates to expression \texttt{$M$($N$)} and
+$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
+In other words, the function application 
+corresponds to the act of supplying a function 
+with an argument and
+the $\lambda$-abstraction is equivalent to 
+\textit{anonymous function}.
+
+\begin{todo}
+ & asi říct o zkratkách notací
+\end{todo}
+
 \subsubsection{Reductions}
+(... zmiňovat ? a pokud ano tak tady nebo radši v sekci která diskutuje 
+normálnost vygenerovanejch termů a jejich zkrácení pomocí $eta$-redukce)
 \subsubsection{Tree representations}
-%\subsubsection{}
+(... zmiňovat ?)
 
@@
 
+A \lterm as described above
+corresponds to a program expression with no type information
+included. Now we will describe \textit{types} (or \textit{type terms}).
+After putting those two pieces 
+(\textit{\lterms} and \textit{types}) together 
+we will get system called \textit{simply typed $\lambda$-calculus}.
+
+\begin{definition}
+Let $A$ be set of {\it atomic type names}. 
+Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
+\begin{align*}
+\alpha      \in A  &\then   \alpha \in \T \\
+\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
+\end{align*}~
+
+\end{definition}
+
+Type $\sigma \ar \tau$ is type for functions taking as input
+something of a type $\sigma$ and returning 
+as output something of a type $\tau$.
+
+\begin{todo}
+ & asi říct o zkratce notace
+\end{todo}
+
 \subsection{Context}
 
-\textit{--- nezapomenout popsat $\Gamma,x:\tau$ notaci
+\begin{definition} 
+~
+\begin{enumerate}
+ \item 	Let $\Lambda$ be set of {\it \lterms}. 
+	Let $\mathbb{T}$ be set of {\it types}.       
+	A {\it statement} $M : \sigma$ is a pair 
+	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
+	Statement $M : \sigma$ is vocalized as 
+	{\it "$M$ has type $\sigma$"}.
+	The term $M$ is called the {\it subject} of the 
+	statement $M : \sigma$.
+ \item A \textit{declaration} is a statement 
+ $x : \sigma$ where $x \in V \cup C$.
+
+ \item A \textit{context} 
+ %(or \textit{basis}) 
+ is set of declarations with distinct variables as subjects.
+\end{enumerate}~
+\end{definition}
+
+\begin{todo}
+   & říct že $T \cup F$ v Std GP je u nás $\Gamma$
+   & popsat $\Gamma,x:\tau$ notaci
  a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
- tzn je nová...}
+ tzn je nová...
+  & že píšem $x:\sigma \in \Gamma$ místo $(x,\sigma) \in \Gamma$
+\end{todo}
 
-\subsection{\GMS}
+\subsection{Simply typed $\lambda$-calculus}
+
+\begin{definition}
+A statement $M\colon\sigma$ is \textit{derivable from}
+a context $\Gamma$ (notation 
+\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
+if it can be produced by the following rules.
+\begin{align*}
+x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
+\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
+&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
+\tur{\Gamma,x:\sigma}{M}{\tau}
+&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
+\end{align*}~
+\end{definition}
+
+...Our goal is to produce terms $M$
+for a given pair $\ul{\tau}{\Gamma}$
+such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.
+
+\begin{todo}
+    & From this definition we can derive following rules...
+
+	& citovat něco, asi barendrechta 
+	(on to tam ale nějak moc nedokazuje) 
+	případně mojí diplomku (tam to dokazuju)
+\end{todo}
 
@@
 
-\textit{- Unifished term je term který může obsahovat jako list uzel tvaru 
-$(\tau,\Gamma)$, takzvaný unfinished leaf (UL).\\
--Vysvětlit unfinished leaf.\\}
+\begin{definition}
+Let $V$ be infinite countable set of {\it 
+variable names}. Let $C$ be set of {\it constant names}, 
+$V \cap C = \emptyset$.	
+Let $\T$ be set of types.
+Let $\C$ be set of all contexts on ($V \cup C$, $\T$).
+Then $\Lpr$ is set of 
+\textit{unfinished  \lterms} defined inductively as follows.	
+\begin{align*}
+\tau \in \T , \Gamma \in \C &\then \ul{\tau}{\Gamma} \in \Lpr
+\textit{~~~~~~~~(Unfinished leaf)}\\
+x   \in V \cup C  &\then x     \in \Lpr \\
+M,N \in \Lpr   &\then (M~N) \in \Lpr 
+\textit{~~~~~~(Function application)} \\
+x   \in V , M \in \Lpr &\then \lamb{x}{M} \in \Lpr
+\textit{~~~~($\lambda$-abstraction)} 
+\end{align*}
+\end{definition}
+
+\begin{todo}
+   & Vysvětlit k čemu je to dobrý.
+\end{todo}
 
@@
 \subsection{Introduction}
-úvod ke kapitole - systematicky, A*, filtrace, znova dám do fronty začátek
+
+\begin{todo}
+ & úvod ke kapitole - systematicky, A*, filtrace, znova dám do fronty začátek
+\end{todo}
 
@@
 is priority queue of unfinished terms. 
-Priority of an unfinished term is given by 
-the number of unfinished leafs in that term.
+Priority of an unfinished term is given by its size.
 Initially, the queue contains only one unfinished term; 
@@
 
-At each step, the term $M$ with the lowest number of 
-unfinished leafs is pulled from the queue.
+At each step, the term $M$ with the smallest size
+is pulled from the queue.
 According to the actual number of those leafs one of
@@
 \begin{enumerate}
- \item If the term has no unfinished leaf (i.e., it is a finished
+ \item If the term $M$ has no unfinished leaf (i.e., it is a finished
  term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
@@
 
-\textit{- pač je to generovany v lnf, neboli beta-eta na -1 nf kde 
-eta na -1 je eta expanze tak je chytrý transformovat 
-to do beta-eta nf. To stačí opakovanou eta redukcí, protože beta normálnost
+\textbf{(Otázka zda vůbec zminovat?)}
+
+\textit{- pač je to generovany v lnf, neboli \beenf kde 
+$\eta^{-1}$ je $\eta$-expanze tak je chytrý transformovat 
+to do \benf. To stačí opakovanou $\eta$-redukcí, protože beta normálnost
 se neporuší (asi ve zkratce uvést proč, pač je to celkem přímočarý) 
@@
 
-\textit{Zmínil bych jí tu ale jen hodně rychle...}
+\textit{Zmínil bych jí tu ale jen hodně rychle..., nutno zmínit
+pokud budem uvádět even parity problém }
 

### 2013-10-23 tex/evogp.tex
evogp tex

@@
 \newenvironment{todo}
-{ ~\\[0.5em]
-  {\color{red}\textbf{TODO}}
+{~\\ {\color{red}\textbf{TODO}}
   \begin{easylist}[itemize]}
-{ \end{easylist}
-  ~}
+{ \end{easylist}}
 
@@
 
-\subsection{Ramped half-and-half}
+%\subsection{Term}
 
-\subsection{Term}
-
-Here we  %formally 
+\Lets 
+%Here we  %formally 
 describe programming language, 
@@
 \textit{anonymous function}.
-
-\begin{todo}
- & asi říct o zkratkách notací
-\end{todo}
-
-\subsubsection{Reductions}
-(... zmiňovat ? a pokud ano tak tady nebo radši v sekci která diskutuje 
-normálnost vygenerovanejch termů a jejich zkrácení pomocí $eta$-redukce)
-\subsubsection{Tree representations}
-(... zmiňovat ?)
-
-\subsection{Type}
-
+%\begin{todo}
+% & asi říct o zkratkách notací
+%\end{todo}
+%
+%
+%\subsubsection{Reductions}
+%(... zmiňovat ? a pokud ano tak tady nebo radši v sekci která diskutuje 
+%normálnost vygenerovanejch termů a jejich zkrácení pomocí $eta$-redukce)
+%\subsubsection{Tree representations}
+%(... zmiňovat ?)
+%\subsection{Type}
 A \lterm as described above
@@
 \sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
-\end{align*}~
-
+\end{align*}
 \end{definition}
@@
 
-\begin{todo}
- & asi říct o zkratce notace
-\end{todo}
+%\begin{todo}
+% & asi říct o zkratce notace
+%\end{todo}
 
-\subsection{Context}
+%\subsection{Context}
 
-\begin{definition} 
-~
-\begin{enumerate}
+\begin{definition}\begin{enumerate}
  \item 	Let $\Lambda$ be set of {\it \lterms}. 
@@
  is set of declarations with distinct variables as subjects.
-\end{enumerate}~
+\end{enumerate}
 \end{definition}
 
-\begin{todo}
-   & říct že $T \cup F$ v Std GP je u nás $\Gamma$
-   & popsat $\Gamma,x:\tau$ notaci
- a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
- tzn je nová...
-  & že píšem $x:\sigma \in \Gamma$ místo $(x,\sigma) \in \Gamma$
-\end{todo}
+%\begin{todo}
+%   & říct že $T \cup F$ v Std GP je u nás $\Gamma$
+%   & popsat $\Gamma,x:\tau$ notaci
+% a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
+% tzn je nová...
+%  & že píšem $x:\sigma \in \Gamma$ místo $(x,\sigma) \in \Gamma$
+%\end{todo}
 
-\subsection{Simply typed $\lambda$-calculus}
+%\subsection{Simply typed $\lambda$-calculus}
 
@@
 &~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
-\end{align*}~
+\end{align*}
 \end{definition}
@@
 
-\begin{todo}
-    & From this definition we can derive following rules...
+%\begin{todo}
+%    & From this definition we can derive following rules...
+%
+%	& citovat něco, asi barendrechta 
+%	(on to tam ale nějak moc nedokazuje) 
+%	případně mojí diplomku (tam to dokazuju)
+%\end{todo}
 
-	& citovat něco, asi barendrechta 
-	(on to tam ale nějak moc nedokazuje) 
-	případně mojí diplomku (tam to dokazuju)
-\end{todo}
-
-\subsection{Unfinished term}
+%\subsection{Unfinished term}
 
@@
 
-\begin{todo}
-   & Vysvětlit k čemu je to dobrý.
-\end{todo}
+%\begin{todo}
+%   & Vysvětlit k čemu je to dobrý.
+%\end{todo}
 
@@
 
-Essential data structure of our term generating algorithm 
+Essential data structure of our algorithm 
 is priority queue of unfinished terms. 
 Priority of an unfinished term is given by its size.
-Initially, the queue contains only one unfinished term; 
-$\ul{\tau}{\Gamma}$.\\
+At the beginning, the queue contains only one unfinished term; 
+$\ul{\tau}{\Gamma}$. The search strategy $S$ also 
+initializes its internal state (if it has one).\\
 
@@
 \begin{enumerate}
- \item If $L = \ul{\rho \ar \sigma}{\Gamma}$, 
-       then $L$ is replaced by $\lamb{x}{\ul{\sigma}{\Gamma,x \colon \rho}}$.
+ \item If $L = \ul{\rho_1 \ar \dots \ar \rho_n \ar \alpha}{\Gamma}$,
+ 	   where $\alpha$ is atomic type and $n \geq 1$, 
+       then $L$ is replaced by 
+       $\lamb{x_1 \dots x_n}{\ul{\sigma}
+       {\Gamma,x_1 \colon \rho_1,\dots,x_n \colon \rho_n}}$.
        Thus this expansion results in exactly one successor.  
@@
        then for each 
-       \mbox{$f : (\tau_1 \ar \dots \ar \tau_n \ar \alpha) \in \Gamma$}
+       \mbox{$f : (\tau_1 \ar \dots \ar \tau_m \ar \alpha) \in \Gamma$}
        the unfinished leaf $L$ is replaced by 
-       $(~f~(\tau_1,\Gamma)~\dots~(\tau_n,\Gamma)~)$.
+       $(~f~(\tau_1,\Gamma)~\dots~(\tau_m,\Gamma)~)$.
        Thus this expansion results in many (possibly zero or one) successors.
@@
 is generated, then the initial unfinished term $\ul{\tau}{\Gamma}$ 
-is inserted to the queue and the process continues.
+is inserted to the queue, search strategy $S$
+again initializes its internal state and the process continues.
 
@@
 
+\begin{todo}
+ & Okomentovat jak se počítá velikost a při tý příležitosti říct
+   že A* heuristika se skrejvá v tom jak velký určíme unfinished leafs,
+   že my používáme 1, ale že si de představit o dost sofistikovanější 
+   verze.   
+\end{todo}
+
 \subsubsection{Systematic strategy}
+
+If we use trivial strategy that returns all the inputs, 
+then the algorithm systematically generates 
+first $n$ smallest lambda terms in their
+\textit{long normal form}.
+
 \subsubsection{Ramped half-and-half strategy}
+
+The internal state of this strategy consists of two variables.
+It is the only one strategy described here that uses an internal state.
+
+\begin{enumerate}
+ \item \textit{isFull} - A boolean value, determining whether \textit{full}
+                     or \textit{grow} method will be performed.
+ \item \textit{d} - A integer value from $\setDots{2}{D_{init}}$, where 
+                $D_{init}$ is predefined maximal depth (e.g. 6).                    
+\end{enumerate}
+
+This strategy returns precisely one randomly
+(uniformly) selected  element from 
+the \textit{selection subset} of input set
+(or zero elements if the input set is empty). 
+The \textit{selection subset} 
+to select from is determined by $depth, d$ and $isFull$.
+The $depth$ parameter is the depth (in the term tree) 
+of the unfinished leaf that was expanded.
+Those elements of input set whose newly added subtree contains one ore more 
+unfinished leafs are regarded as \textit{non-terminals}, whereas 
+those whose newly added subtree contains no unfinished leaf are regarded as 
+\textit{terminals}.
+If $depth = 0$, then the subset to select from is  
+set of all \textit{non-terminals} of the input set.
+If $depth = d$, then the subset to select from is
+set of all \textit{terminals} of the input set.
+In other cases of $depth$ it depends on value of $isFull$.
+If $isFull = true$, then the subset to select from is 
+set of all \textit{non-terminals} of the input set.
+If $isFull = false$, then the subset to select from is 
+the whole input set.
+
+\begin{todo}
+ & napsat o tom že ji používa koza [citace] a že je to standard,
+ pokud je použita na gamu co splnuje closure podmínku, tak 
+ generuje přesně stejně jako
+ & ...This means that the queue always contains only one (or zero) state.
+\end{todo}
+
 \subsubsection{Geometric strategy}
 
-~\\\\
-\textit{
-- Systematická strategie nezahodí nic (Pro ní generování 
-  kolapsuje do A* algoritmu).\\
-- Ramped-half-and-half zahodí vždy vše krom jediného.\\
-- Geometrická strategie nezahodí s pravděpodobností q na hloubka UL \\
-}
+We can see those two previous strategies as two extremes on the spectrum of 
+possible strategies. 
+\textit{Systematic strategy} filters no successor state thus performing
+exhaustive search resulting in discovery of $n$ smallest terms in one run.
+On the other hand, \textit{ramped half-and-half strategy} filters 
+all but one successor states resulting in degradation of 
+the priority queue into "fancy variable".
+\textit{Geometric strategy} is simple yet fairly effective term generating 
+strategy somewhere in the middle of this spectrum.
+It is parameterized by parameter $q \in (0,1)$, its default well-performing 
+value is $q = 0.75$.
+For each element of the input set 
+it is probabilistically decided whether
+it will be returned or omitted. A probability $p$ of returning is
+same for all elements, but depends on the $depth$, 
+which is defined in the same way as in previous strategy. 
+It is computed as follows.
+$$ p = q^{depth} $$
+This formula is motivated by idea that it is important to
+explore all possible root symbols, but as the $depth$ 
+increases it becomes less "dangerous" to omit 
+an exploration branch. 
+We can see this by considering that this strategy results in
+somehow forgetful A* search.
+With each omission we make the search space smaller. But with
+increasing depth these omissions have smaller impact on the search space,
+i.e., they cut out lesser portion of the search space.
+Another slightly esoteric argument supporting this formula is that "root 
+parts" of a program usually stand for more crucial parts
+with radical impact on global behavior of a program, 
+whereas "leaf parts" of a program usually
+stand for less important local parts (e.g. constants).  
+This strategy also plays nicely with the idea that 
+"too big trees should be killed".
 
@@
 \begin{align*}
-\sigma = \Real \ar &\Real\\
+\tau = \Real \ar &\Real\\
 \Gamma = \{
@@
 where
-
-\noindent\begin{minipage}{.5\linewidth}
+\noindent
+\begin{minipage}{.5\linewidth}
 \begin{align*}
@@
 \end{minipage}
-
-~\\\\
-Fitness function is computed as follows
-
+~\\
+Fitness function is computed as follows.
 $$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
-
-where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.\\
-
+where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
 An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
@@
 \subsection{Artificial ant}
-...
+
+\textit{Artificial Ant} is another problem described
+in \cite{koza92}. Objective of this problem is to 
+find a control program for an artificial ant so
+that it can find all food located on "Santa Fe" trail.
+The Santa Fe trail lies on toroidal square grid.
+The ant is in the upper left corner, facing right.
+The ant is able to move forward, turn left, and sense if a food 
+piece is ahead of him.
+
+\begin{align*}
+\tau = An&tAction\\
+\Gamma = \{~~
+  l    &: AntAction                              ,
+  r    : AntAction                               ,
+  m    : AntAction                               ,\\
+  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
+  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
+  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
+\end{align*}
+
+Action $l$ turns the ant left. 
+Action $r$ turns the ant right.
+Action $m$ moves the ant forward.
+Action $ifa~x~y$ (if-food-ahead) performs action $x$ 
+if a food piece is ahead of the ant,
+otherwise it performs action $y$.
+Action $p2~x~y$ first performs action $x$ and after it action $y$.
+Action $p3~x~y~z$ first performs action $x$, 
+after that action $y$ and finally $z$.
+Actions $l, r$ and $m$ each take one time step to execute.
+Ants action is performed over and over again until it reaches predefined
+maximal number of steps. 
+Fitness value is equal to number of eaten food pieces.
+An individual such that eats all 89 pieces of food is 
+considered as a correct solution.
+This limit is set to be 600\footnote{
+In \cite{koza92} this limit is said to be 400 time steps.
+But there is also mentioned following solution, 
+which is described as correct solution:
+\texttt{(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))}.
+This program needs 545
+time steps; if it is given only 400 time steps, then it eats only 79 pieces
+of food. Thus we use 600 time steps. 
+} time steps.
+
 \subsection{Even parity problem}
@@
 \section{Conclusions}
-...
+
+\begin{todo}
+ & do future work bych napsal, že díky jednoduchosti 
+   toho co strategie dělá je to idealní kandidát na 
+   optimalizaci pomocí samotnýho GP
+\end{todo}
 

### 2013-11-07 tex/evogp13/evogp.tex
evogp13

--- 2013-10-23 tex/evogp.tex
@@
 
+%\usepackage{times}
 \usepackage[utf8]{inputenc}
 \usepackage{amsfonts}
-\usepackage{qtree}
+%\usepackage{qtree}
 \usepackage{xspace}
@@
 \usepackage[ampersand]{easylist}
-\usepackage{color}
+\usepackage[usenames,dvipsnames]{color}
 
@@
 \newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
+\newcommand{\ro}[1]{{\color{blue} #1}}
+\newcommand{\tom}[1]{{\color{ForestGreen} #1}}
 
-\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}
+%\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}
 
@@
 
-\title{\textit{[něco jako?]} 
-Generating lambda term individuals using 
+\title{
+Generating Lambda Term Individuals in Typed Genetic Programming Using 
 %undisciplined 
-forgetful
-A*}
-\author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}
-\institute{Matfyz ...}
+Forgetful A*}
+\author{Tom\'{a}\v{s} K\v{r}en\inst{1} \and Roman Neruda\inst{2}}
+\institute{%
+Faculty of Mathematics and Physics\\
+Charles University in Prague\\
+Malostransk\'e n\'am\v{e}st\'\i~25, 11000 Prague, Czech Republic\\
+\email{tomas.kren@mff.cuni.cz}
+\and
+Institute of Computer Science\\
+Academy of Sciences of the Czech Republic\\
+Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207, Prague, Czech Republic\\
+\email{roman@cs.cas.cz}
+}
 \maketitle
@@
 \begin{abstract}
-\textbf{\textit{
-(Aby tu něco bylo, pak se udělá lepší 
-(tohle je trochu upravenej abstract z diplomky))}}
-In this paper, generalization of the standard genetic programming (GP)
-for simply typed lambda calculus is presented. We use population 
-initialization method parameterized by simple search strategy. 
-First described strategy corresponds to standard ramped half-and-half method, 
-second one corresponds to exhaustive systematic search and third one is a 
-novel geometric strategy, which outperforms standard method in success 
-rate, time consumption and average individual size in two experiments. 
+In this paper, a generalization of genetic programming 
+for simply typed lambda calculus is presented. We propose three population 
+initialization methods depending on different search strategies. 
+The first strategy corresponds to standard ramped half-and-half initialization, 
+the second one corresponds to exhaustive systematic search, and the third one 
+represents a novel geometric strategy which outperforms standard genetic 
+programming in success rate, time consumption and average individual size 
+on two experiments. 
 Other performance enhancements based on theory of lambda calculus are 
-proposed and supported by experiment. Abstraction elimination is 
-utilized to enable use of simple tree-swapping crossover.
+proposed and supported by experiments, including abstraction elimination 
+that enables us to utilize a simple tree-swapping crossover.
 \end{abstract}
@@
 
-\subsection{gp a typy jsou dobrý / motivace}
-...
-\subsection{příběh}
+%\subsection{gp a typy jsou dobrý / motivace}
+%...
+%\subsection{příběh}
 
-(...)\\
+%(...)\\
 
-Our approach aims to play with the full arsenal given by simply typed lambda calculus, thus we begin our reasoning with an exhaustive systematic search
-in mind. Our second goal is to construct a system generalizing Standard 
-GP \cite{koza92}.
-In order to satisfy both these goals the designed system should be 
-parameterized by some simple piece of code that makes the difference
-between exhaustive systematic search and standard but 
-quiet arbitrary ramped half-and-half generating method.
+Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}.
+Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
 
-Those two design goals also differentiate our system from 
-the three state of the art systems for typed GP known to us.
+\begin{todo}
+& \ro{... něco k tomu, že se to teda dělá i skrz lambda kalkul}
+\end{todo}
 
-~\\(...)
+The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
+During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies.
+Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus.
+Thus, the natural idea is to employ an exhaustive systematic search.
+On the other hand, if we were to mimic the standard GP approach, a quite arbitrary yet common and successful ramped half-and-half generating heuristic~\cite{fg} should probably be used.
+These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies.
+This design goal also differentiate our approach from
+the three state of the art proposals for typed GP known to us that are discussed in the following section. 
+Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
+Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.
 
-~\\
-\textbf{poznamky naky}\textit{\\
-- geometrická strategie generování termů jakožto hybrid systematického a náhodného sytylu --- to že jsme si vytičili ty dva cíle který plníme tou 
-jednoduchou strategií, tak máme možnost nalízt jednoduchou strategii která 
-je na pomezí obou cílů - taková strategie je právě ta naše geometrická 
-a ukazuje se že se bohulibě chová právě k strašáku GP -- \textbf{bloatu}.\\
-- teoretické LC konstrukty s výhodou použity - @-stromy a eta-redukce\\
-- křížení by eliminace 
-}
+%~\\(...)
+%
+%~\\
+%\textbf{poznamky naky}\textit{\\
+%- geometrická strategie generování termů jakožto hybrid systematického a náhodného sytylu --- to že jsme si vytičili ty dva cíle který plníme tou 
+%jednoduchou strategií, tak máme možnost nalízt jednoduchou strategii která 
+%je na pomezí obou cílů - taková strategie je právě ta naše geometrická 
+%a ukazuje se že se bohulibě chová právě k strašáku GP -- \textbf{bloatu}.\\
+%- teoretické LC konstrukty s výhodou použity - @-stromy a eta-redukce\\
+%- křížení by eliminace 
+%}
 
-\subsection{obsah kapitol článku v 1 větě}
-...
+%\subsection{obsah kapitol článku v 1 větě}
+
+The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions.
+Main original results about search strategies in individual generating are described in section~\ref{approach}.
+Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.
 
 \section{Related work}
+\label{related}
 \subsection{Yu}
@@
 \section{Preliminaries}
+\label{preliminaries}
 %definice, pojmy
-
 %\subsection{Term}
 
-\Lets 
+In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
+First, \lets 
 %Here we  %formally 
-describe programming language, 
-in which we generate individual programs --- so called \lterms.
+describe a programming language, 
+in which the GP algorithm generates individual programs --- the so called \lterms.
 
@@
 corresponds to the act of supplying a function 
-with an argument and
+with an argument, and
 the $\lambda$-abstraction is equivalent to 
-\textit{anonymous function}.
-%\begin{todo}
-% & asi říct o zkratkách notací
-%\end{todo}
-%
-%
+\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
+they were recently introduced to C++, and they are expected to be supported in Java 8.}.
+\begin{todo}
+ & asi říct o zkratkách notací
+\end{todo}
+
 %\subsubsection{Reductions}
@@
 %\subsection{Type}
+
 A \lterm as described above
@@
 included. Now we will describe \textit{types} (or \textit{type terms}).
-After putting those two pieces 
-(\textit{\lterms} and \textit{types}) together 
-we will get system called \textit{simply typed $\lambda$-calculus}.
 
@@
 
-%\begin{todo}
-% & asi říct o zkratce notace
-%\end{todo}
+\begin{todo}
+ & asi říct o zkratce notace
+\end{todo}
+
+The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
+combining the previously defined \textit{\lterms} and \textit{types} together.
 
@@
 
-%\begin{todo}
-%   & říct že $T \cup F$ v Std GP je u nás $\Gamma$
-%   & popsat $\Gamma,x:\tau$ notaci
-% a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
-% tzn je nová...
-%  & že píšem $x:\sigma \in \Gamma$ místo $(x,\sigma) \in \Gamma$
-%\end{todo}
+\begin{todo}
+   & říct že $T \cup F$ v Std GP je u nás $\Gamma$
+   & popsat $\Gamma,x:\tau$ notaci
+ a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
+ tzn je nová...
+  & že píšem $x:\sigma \in \Gamma$ místo $(x,\sigma) \in \Gamma$
+\end{todo}
 
@@
 
-%\begin{todo}
-%   & Vysvětlit k čemu je to dobrý.
-%\end{todo}
+\begin{todo}
+   & Vysvětlit k čemu je to dobrý. Prostě říct že 
+     to $\ul{\tau}{\Gamma}$ představuje zatim nespecifikovaný term
+     který je ale typu $\tau$ a mohu se v něm objevit stavební symboly
+     z $\Gamma$.
+\end{todo}
 
 \section{Our approach}
+\label{approach}
 \subsection{Introduction}
+\tom{
+Our approach to \lterm gerating is based on technique 
+briefly described in \cite{barendregt10}, which generates
+well-typed \lterms in their long normal form. 
+We use this technique to perform systematic exhaustive enumeration
+of \lterms in their long normal form in order from smallest to largest.
+We use well known \textit{A* algorithm} \cite{AIMA} for this task.
+A* is used to search in a given state space for a goal state. 
+It maintains a priority queue to organize states yet to be explored.
+Initially this queue contains only the initial state.
+
+Our state space to search in is the space of unfinished \lterms. 
+The initial state is the unfinished term $\ul{\tau}{\Gamma}$, 
+where $\tau$ is the desired type of
+terms to be generated and $\Gamma$ is the context
+representing the set of building symbols to be used in construction of
+terms (it corresponds to the set $T \cup F$ in
+standard GP enriched with types). The process of determining 
+successors of a state described below is designed so it constructs well-typed 
+\lterms and omits no \lterm in its long normal form. 
+A state is considered a goal state if it contains no unfinished
+leaf, i.e., it is a finished \lterm.
+
+Our generating method is based on simple modification of the
+standard A*, which we call \textit{forgetful A*}. This modification consist in 
+additional parameter for the A* algorithm -- the \textit{search strategy}. 
+It is a simple filtration function (along with initialization procedure)
+that is given the set of all successors of the state that is being examined
+and returns a subset of this input. This subset is added to the priority queue 
+to be further explored. In this way the search space may be reduced as 
+the filtration function may \textit{forget} some successors.
+If the queue becomes empty before the desired number of \lterms
+is generated, then the initial statete is inserted to the queue
+and the process continues. For the standard A* this would be meaningless,
+but since our A* is forgetful this kind of restart makes sense.
+
+A* keeps a priority queue of states during the generation process,
+on the other hand the \textit{ramped half-and-half method}, 
+the standard GP algorithm for generating individuals, 
+keeps only one individual which is gradually constructed. This 
+behavior is easily achieved by use of suitable search strategy 
+that returns subset consisting of only one successor.
+The systematic search is obtained by search strategy that 
+returns whole input set.      
+Our novel \textit{geometric strategy} can be understood as
+point somewhere between those two extremes.
+}
 
 \begin{todo}
- & úvod ke kapitole - systematicky, A*, filtrace, znova dám do fronty začátek
+& \tom{Kde nejlíp zmínit citaci diplomky? Něco jako, popisujem to tu velmi ve ztručnosti,
+  more elaborate description may be found in [diplomka]?}
+& \tom{definovat long normal form? (asi nejde uplně krátce bohužel)}
 \end{todo}
@@
 is pulled from the queue.
-According to the actual number of those leafs one of
+According to the number of unfinished leafs in $M$ one of
 the following actions is performed.
@@
  term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
- result collection of generated terms.   
+ result set\footnote{
+ TODO: říct že tim že je to set se řeší problém toho, aby se tam nevyskytovali
+ některý vygenerovaný víckrát, pokud však toto nechcem můžem použít jinou collection
+ , případně to vůbec nezminovbat když by to kradlo moc místa
+ } of generated terms.   
  \item Otherwise, \textit{successors} of the unfinished term $M$ are
@@
        then $L$ is replaced by 
-       $\lamb{x_1 \dots x_n}{\ul{\sigma}
+       $\lamb{x_1 \dots x_n}{\ul{\alpha}
        {\Gamma,x_1 \colon \rho_1,\dots,x_n \colon \rho_n}}$.
@@
 \section{Experiments}
+\label{experiments}
+
 \subsection{Simple symbolic regression}
@@
 ...
+
 \section{Conclusions}
+\label{conclusions}
 
@@
 
-\begin{thebibliography}{1}
+\nocite{*}
 
-\bibitem{koza92}
-  John R. Koza,
-  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
-  MIT Press, Cambridge, MA,
-  1992.
-
-\bibitem{koza05}
-  Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
-  \emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
-  Springer, 2005. ISBN 978-0-387-26417-2
-
-\bibitem{fg}
- Riccardo Poli, William B. Langdon, Nicholas F. McPhee
- \emph{A Field Guide to Genetic Programming}.
- Lulu Enterprises, UK Ltd, 2008.
-
-\bibitem{yu01}
-  T. Yu. 
-  \emph{Hierachical processing for evolving recursive and modular 
-        programs using higher order functions and lambda abstractions}. 
-  Genetic Programming and Evolvable Machines,
-  2(4):345–380, December 2001. ISSN 1389-2576.
-
-\bibitem{montana95}
-D. J. Montana. 
-\emph{Strongly typed genetic programming.} 
-Evolutionary Computation, 3(2): 199–230, 1995.
-%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha
-
-\bibitem{haynes96}
-T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
-\emph{Type inheritance in strongly typed genetic programming.} 
-In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
-in Genetic Programming 2, chapter 18, pages 359–376.
-MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1. 
-%URL \url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.
-
-\bibitem{olsson94}
-J. R. Olsson. 
-\emph{Inductive functional programming using incremental program 
-transformation and Execution of logic programs by 
-iterative-deepening A* SLD-tree search.} 
-Dr scient thesis, University of Oslo, Norway, 1994.
-
-\bibitem{kes}
-Forrest Briggs, Melissa O’Neill.
-\emph{Functional Genetic Programming and Exhaustive
-Program Search with Combinator Expressions.}
-International Journal of Knowledge-based and Intelligent Engineering Systems,
-Volume 12 Issue 1, Pages 47-68, January 2008.
-
-\bibitem{barendregt84}
-H. P. Barendregt,
-\emph{The Lambda Calculus: its Syntax and Semantics}, 
-revised ed., North-Holland, 1984.
-
-\bibitem{barendregt92}
-H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
-\emph{Lambda Calculi with Types.} 
-Handbook of Logic in Computer Science, 1992.
-
-\bibitem{barendregt10}
-
-  Henk Barendregt, Wil Dekkers, Richard Statman,
-  \emph{Lambda Calculus With Types}.
-  Cambridge University Press,
-  2010. 
-  %URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
-
-\bibitem{jones87}
-Simon Peyton Jones. 
-\emph{The Implementation of Functional Programming Languages}. 
-Prentice Hall, 1987.
-
-\bibitem{AIAMA}
-	Stuart J. Russell, Peter Norvig,
-	\emph{Artificial Intelligence: A Modern Approach}.
-	Pearson Education,
-	2003.
-
-\end{thebibliography}
+\bibliographystyle{splncs}
+\bibliography{gp}
 
 \end{document}
+
+%
+%\begin{thebibliography}{1}
+%
+%
+%\bibitem{koza92}
+  %John R. Koza,
+  %\emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
+  %MIT Press, Cambridge, MA,
+  %1992. 
+%
+%\bibitem{koza05}
+  %Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
+  %\emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
+  %Springer, 2005. ISBN 978-0-387-26417-2 
+%
+%\bibitem{fg}
+ %Riccardo Poli, William B. Langdon, Nicholas F. McPhee
+ %\emph{A Field Guide to Genetic Programming}.
+ %Lulu Enterprises, UK Ltd, 2008.
+%
+%\bibitem{yu01}
+  %T. Yu. 
+  %\emph{Hierachical processing for evolving recursive and modular 
+        %programs using higher order functions and lambda abstractions}. 
+  %Genetic Programming and Evolvable Machines,
+  %2(4):345–380, December 2001. ISSN 1389-2576.
+%
+%
+%\bibitem{montana95}
+%D. J. Montana. 
+%\emph{Strongly typed genetic programming.} 
+%Evolutionary Computation, 3(2): 199–230, 1995.
+%%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha
+%
+%\bibitem{haynes96}
+%T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
+%\emph{Type inheritance in strongly typed genetic programming.} 
+%In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
+%in Genetic Programming 2, chapter 18, pages 359–376.
+%MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1. 
+%%URL \url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.
+%
+%\bibitem{olsson94}
+%J. R. Olsson. 
+%\emph{Inductive functional programming using incremental program 
+%transformation and Execution of logic programs by 
+%iterative-deepening A* SLD-tree search.} 
+%Dr scient thesis, University of Oslo, Norway, 1994.
+%
+%\bibitem{kes}
+%Forrest Briggs, Melissa O’Neill.
+%\emph{Functional Genetic Programming and Exhaustive
+%Program Search with Combinator Expressions.}
+%International Journal of Knowledge-based and Intelligent Engineering Systems,
+%Volume 12 Issue 1, Pages 47-68, January 2008. 
+%
+%
+%\bibitem{barendregt84}
+%H. P. Barendregt,
+%\emph{The Lambda Calculus: its Syntax and Semantics}, 
+%revised ed., North-Holland, 1984.
+%
+%\bibitem{barendregt92}
+%H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
+%\emph{Lambda Calculi with Types.} 
+%Handbook of Logic in Computer Science, 1992. 
+%
+%\bibitem{barendregt10}
+%
+  %Henk Barendregt, Wil Dekkers, Richard Statman,
+  %\emph{Lambda Calculus With Types}.
+  %Cambridge University Press,
+  %2010. 
+  %%URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
+%
+%\bibitem{jones87}
+%Simon Peyton Jones. 
+%\emph{The Implementation of Functional Programming Languages}. 
+%Prentice Hall, 1987.
+%
+%
+%\bibitem{AIAMA}
+	%Stuart J. Russell, Peter Norvig,
+	%\emph{Artificial Intelligence: A Modern Approach}.
+	%Pearson Education,
+	%2003. 
+%
+%
+%\end{thebibliography}

### 2013-11-11 tex/evogp13/evogp.tex
článek

@@
 %\usepackage{amssymb}
+\usepackage{graphicx}
 
@@
 \newcommand{\tom}[1]{{\color{ForestGreen} #1}}
+\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 \title{
-Generating Lambda Term Individuals in Typed Genetic Programming Using 
-%undisciplined 
-Forgetful A*}
-\author{Tom\'{a}\v{s} K\v{r}en\inst{1} \and Roman Neruda\inst{2}}
-\institute{%
-Faculty of Mathematics and Physics\\
-Charles University in Prague\\
-Malostransk\'e n\'am\v{e}st\'\i~25, 11000 Prague, Czech Republic\\
-\email{tomas.kren@mff.cuni.cz}
-\and
-Institute of Computer Science\\
-Academy of Sciences of the Czech Republic\\
-Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207, Prague, Czech Republic\\
-\email{roman@cs.cas.cz}
-}
+Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*}
+%\author{Tom\'{a}\v{s} K\v{r}en\inst{1} \and Roman Neruda\inst{2}}
+%\institute{%
+%Faculty of Mathematics and Physics\\
+%Charles University in Prague\\
+%Malostransk\'e n\'am\v{e}st\'\i~25, 11000 Prague, Czech Republic\\
+%\email{tomas.kren@mff.cuni.cz}
+%\and
+%Institute of Computer Science\\
+%Academy of Sciences of the Czech Republic\\
+%Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207, Prague, Czech Republic\\
+%\email{roman@cs.cas.cz}
+%}
 \maketitle
@@
 the second one corresponds to exhaustive systematic search, and the third one 
-represents a novel geometric strategy which outperforms standard genetic 
-programming in success rate, time consumption and average individual size 
-on two experiments. 
-Other performance enhancements based on theory of lambda calculus are 
-proposed and supported by experiments, including abstraction elimination 
-that enables us to utilize a simple tree-swapping crossover.
+represents a novel geometric strategy. Three well-known benchmark experiments
+support that the geometric strategy outperforms
+the standard generating method in success rate, time consumption and average individual size. 
+%Other performance enhancements based on theory of lambda calculus are 
+%proposed and supported by experiments, including abstraction elimination 
+%that enables us to utilize a simple tree-swapping crossover.
 \end{abstract}
@@
 Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
-
-\begin{todo}
-& \ro{... něco k tomu, že se to teda dělá i skrz lambda kalkul}
-\end{todo}
+Use of types naturally opens door to enriching S-expressions,
+the traditional GP representation of individuals, with concepts from
+lambda calculus, which is simple yet powerful functional mathematical and programming 
+language extensively used in type theory. Such attempts has shown to be 
+successful \cite{yu01}.
 
@@
 \label{related}
-\subsection{Yu}
-\textbf{Yu} - (články: evenParity, polyGP
-\textbf{[doplnit do citací]}
- co sem ted našel, ten o burzách co mám v kindlu)
-Odlišnosti:
-(1) - generování se nedělá systematicky: pokud strom dojde do místa
-kde funkci nemá dát jaký parametr, tak místo tý funkce dá nějaký 
-terminál.(uvádí 85\% uspěšnost)  
-(2) - Ty křížení ma trochu jinak než v tom článku o even parity,
-musim zjistit jak má udělaný že se jí nedostane např prom \#3 někam,
-kde neni definovaná když dělá přesun podstromu, v tom starym to ale 
-bylo myslim založený na tom, že nedovolovala vnější proměnný uvnitř
-lambda termu - což platí i nadále. čili vtom to určitě nehraje full deck.
-Naopak se víc zaměřuje na polymorfizmus a další věci.
 
-\subsection{kombinatori}
+In \cite{yu01} Yu presents a GP system utilizing
+polymorphic higher-order functions\footnote{Higher-order 
+function is a function taking another function as 
+input parameter.} and lambda abstractions.
+Important point of interest in this work is use of
+\texttt{foldr} function as a tool for \textit{implicit recursion},
+i.e. recursion without explicit recursive calls. 
+The terminal set for constructing lambda abstraction subtrees 
+is limited to use only constants and variables of that particular
+lambda abstraction, i.e., outer variables are not allowed to be used
+as terminals in this work. This is significant difference from our approach 
+since we permit all well-typed normalized \lterms. From this difference also
+comes different crossover operation. We focus more on term generating process; 
+their term generation is performed in a similar way as the standard one, 
+whereas our term generation also tries to utilize techniques of systematic enumeration.
 
-\textbf{kombinátoři} \\
-- vůbec nepoužívaj Lambda Abstrakce\\
-- univerzální genetickej operator\\
-- taky řešej systematický prohledávání
+In \cite{kes} Briggs and O’Neill present technique 
+utilizing typed GP with combinators.
+The difference between approach presented in this work
+and our approach is that in this work terms are generated
+straight from \textit{library} of combinators and no lambda abstractions
+are used. They are using more general polymorphic type system then us
+-- the Hindley–Milner type system. They also discuss the 
+properties of exhaustive enumeration of terms and compare it with GP search.  
+They also present interesting concept of \textit{Generalized
+genetic operator} based on term generation.
 
-\subsection{kanadani}
+In \cite{binard2008genetic} by Binard and Felty even 
+stronger type system (\textit{System F}) is used.  
+But with increasing power of the type system comes increasing difficulty of term generation.
+For this reason evolution in this work takes interesting and nonstandard shape 
+(fitness is associated with \textit{genes} which are evolved together with \textit{species}
+which together participate in creation of individuals).
+This differs from our approach, which tries to be generalization of
+the standard GP\cite{koza92}.
 
-\textbf{kanadani}\\ 
-- o dost silnější typovej systém\\
-- System F, i jim dynamicky vznikaj typy\\
-- moc silný typoví systém, takže generování už je dost
-  složitý, to se otiskuje v silně nestandardním algoritmu
-
-\subsection{barendregt}
+In contrast with above mentioned works our approach uses very simple type system 
+(simply typed lambda calculus) and concentrates on process of generation  
+able to generate all possible well-typed normalized lambda terms. In order to do
+so we use technique based on \textit{inhabitation machines} 
+described by Barendregt in~\cite{barendregt10}.
 
@@
 they were recently introduced to C++, and they are expected to be supported in Java 8.}.
-\begin{todo}
- & asi říct o zkratkách notací
-\end{todo}
+$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for $(\dots((M_1~M_2)~M_3)~\dots~M_n)$
+and $\lam{x_1 x_2 \dots x_n }{M}$ for $\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.
 
@@
 as output something of a type $\tau$. 
-
-\begin{todo}
- & asi říct o zkratce notace
-\end{todo}
-
+$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
+$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
 The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
@@
 
-\begin{todo}
-   & říct že $T \cup F$ v Std GP je u nás $\Gamma$
-   & popsat $\Gamma,x:\tau$ notaci
- a to že to automaticky implikuje že x neni obsažena v $\Gamma$,
- tzn je nová...
-  & že píšem $x:\sigma \in \Gamma$ místo $(x,\sigma) \in \Gamma$
-\end{todo}
+Context is a basic type theoretic concept suitable as a typed alternative
+for terminal and function set in standard GP. 
+Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
+such that $\Gamma$ does not contain any declaration with $x$ as subject.
+We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.
 
@@
 
-...Our goal is to produce terms $M$
+Our goal in term generation is to produce terms $M$
 for a given pair $\ul{\tau}{\Gamma}$
@@
 
-\begin{todo}
-   & Vysvětlit k čemu je to dobrý. Prostě říct že 
-     to $\ul{\tau}{\Gamma}$ představuje zatim nespecifikovaný term
-     který je ale typu $\tau$ a mohu se v něm objevit stavební symboly
-     z $\Gamma$.
-\end{todo}
+\textit{Unfinished leaf} $\ul{\tau}{\Gamma}$ 
+stands for yet not specified \lterm of the type $\tau$ 
+build from symbols of $\Gamma$.
 
@@
 \subsection{Introduction}
-\tom{
+
 Our approach to \lterm gerating is based on technique 
@@
 A* is used to search in a given state space for a goal state. 
+It finds the optimal solution (in our case the smallest term)
+and uses "advising" heuristic function.
 It maintains a priority queue to organize states yet to be explored.
@@
 point somewhere between those two extremes.
-}
 
-\begin{todo}
-& \tom{Kde nejlíp zmínit citaci diplomky? Něco jako, popisujem to tu velmi ve ztručnosti,
-  more elaborate description may be found in [diplomka]?}
-& \tom{definovat long normal form? (asi nejde uplně krátce bohužel)}
-\end{todo}
-
-\subsection{Generating algorithm}
+\subsection{Algorithm}
 
@@
 is priority queue of unfinished terms. 
-Priority of an unfinished term is given by its size.
+Priority of an unfinished term is given by its size\footnote{
+A* heuristic function is hidden in method of computing
+size of unfinished leafs $\ul{\tau}{\Gamma}$. Our algorithm uses
+trivial estimate $\vert\ul{\tau}{\Gamma}\vert = 1$ which is trivially admissible.
+This heuristic is not as silly as it might seem since it is
+quite usual to have $x$ such that $x : \tau \in \Gamma$.
+Since the true value of $\vert\ul{\tau}{\Gamma}\vert$ depends only on
+$\tau$ and \textit{types} in $\Gamma$ (the \textit{signature}), 
+no matter how many variables/constants of
+each type there are, it is should by pretty effective to compute this
+value precisely and store them for later. 
+%This topic has other interesting 
+%subtopics (such as surprisingly small 
+%number of such "signatures" for a given problem), 
+%but it is unfortunately beyond the scope of this paper.
+}.
 At the beginning, the queue contains only one unfinished term; 
 $\ul{\tau}{\Gamma}$. The search strategy $S$ also 
-initializes its internal state (if it has one).\\
+initializes its internal state (if it has one).
 
@@
  term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
- result set\footnote{
- TODO: říct že tim že je to set se řeší problém toho, aby se tam nevyskytovali
- některý vygenerovaný víckrát, pokud však toto nechcem můžem použít jinou collection
- , případně to vůbec nezminovbat když by to kradlo moc místa
- } of generated terms.   
+ result set
+% \footnote{
+% \red{TODO: říct že tim že je to set se řeší problém toho, aby se tam nevyskytovali
+% některý vygenerovaný víckrát, pokud však toto nechcem můžem použít jinou collection
+% , případně to vůbec nezminovbat když by to kradlo moc místa
+% }} 
+ of generated terms.   
  \item Otherwise, \textit{successors} of the unfinished term $M$ are
@@
 \textit{expansion} of the \mbox{\textit{DFS-first}} unfinished leaf $L$
-(i.e., the leftmost unfinished leaf of $M$).\\
+(i.e., the leftmost unfinished leaf of $M$).
 
@@
 In this process, $L$ is replaced
-by a new subterm defined by the following rules.
+by a new subterm defined by the following rules\footnote{
+For the sake of simplicity it is presented as two separate rules. 
+Since the first rule results in exactly one successor it is smarter
+to combine those two rules into one resulting in that unfinished leafs
+have only atomic types. At the beginning we transform the initial type into atomic by first rule (if necessary). After that
+only second rule is applied, but if it results in creation of some unatomic 
+$\ul{\tau_i}{\Gamma}$, then first rules are applied, but during the same successor creation This step eliminates all unatomic unfinished leafs. }.
 \begin{enumerate}
@@
        the unfinished leaf $L$ is replaced by 
-       $(~f~(\tau_1,\Gamma)~\dots~(\tau_m,\Gamma)~)$.
+       $(~f~\ul{\tau_1}{\Gamma}~\dots~\ul{\tau_m}{\Gamma}~)$.
        Thus this expansion results in many (possibly zero or one) successors.
@@
 
-\begin{todo}
- & Okomentovat jak se počítá velikost a při tý příležitosti říct
-   že A* heuristika se skrejvá v tom jak velký určíme unfinished leafs,
-   že my používáme 1, ale že si de představit o dost sofistikovanější 
-   verze.   
-\end{todo}
-
 \subsubsection{Systematic strategy}
@@
 \subsubsection{Ramped half-and-half strategy}
-
+is generalization of the standard ramped half-and-half method described 
+in \cite{koza92}. If applied to context satisfying closure requirement
+(all constants/variables are of the same type, all functions are operations
+on this type), it will behave in the same way as the standard method.
 The internal state of this strategy consists of two variables.
@@
                      or \textit{grow} method will be performed.
- \item \textit{d} - A integer value from $\setDots{2}{D_{init}}$, where 
+ \item \textit{d} - An integer value from $\setDots{2}{D_{init}}$, where 
                 $D_{init}$ is predefined maximal depth (e.g. 6).                    
@@
 the whole input set.
-
-\begin{todo}
- & napsat o tom že ji používa koza [citace] a že je to standard,
- pokud je použita na gamu co splnuje closure podmínku, tak 
- generuje přesně stejně jako
- & ...This means that the queue always contains only one (or zero) state.
-\end{todo}
 
@@
 
-\subsection{Discussion and further improvements}
-
-\subsection{$\eta$-normalization}
-
-\textbf{(Otázka zda vůbec zminovat?)}
-
-\textit{- pač je to generovany v lnf, neboli \beenf kde 
-$\eta^{-1}$ je $\eta$-expanze tak je chytrý transformovat 
-to do \benf. To stačí opakovanou $\eta$-redukcí, protože beta normálnost
-se neporuší (asi ve zkratce uvést proč, pač je to celkem přímočarý) 
-}
-
-\subsection{Crossover}
-
-\textit{Zmínil bych jí tu ale jen hodně rychle..., nutno zmínit
-pokud budem uvádět even parity problém }
+~\\
+Furthermore, our system utilizes generalization of the standard 
+tree-swapping crossover operator. Since it is beyond the scope of this 
+paper we mention it only briefly. Two main concerns with swapping typed subtrees 
+are types and free variables. Well-typed offspring is obtained 
+by swapping only subtrees of the same type. Only subtrees with 
+corresponding counterpart in the second parent are randomly chosen from.
+More interesting problem lies in free variables, which may cause trouble
+if swapped somewhere where it is suddenly not bounded. In order to circumvent this
+difficulty we utilize technique called \textit{abstraction elimination}\cite{jones87}
+that transforms an arbitrary \lterm into \lterm that contains no lambda abstractions
+and no bound variables. After the initial population is generated,  it is transformed
+by abstraction elimination. Another possible transformation taking place
+after initialization is $\eta$-normalization shortening 
+rather long long normal form into $\beta\eta$-normal form.  
+Another performence enhancing transformation is use of "applicative" tree 
+representation (coming directly from inductive definition of \lterms).
 
@@
 
+We made tree experiments comparing performance of standard 
+\textit{ramped half-and-half strategy} with our
+\textit{geometric strategy} (with default parameter $q=0.75$). 
+Each experiment consisted of 50 independent runs 
+of GP algorithm. Each run had maximally 51 generations and population size
+was 500 individuals.
+In those experiments we analyze the ability of the system to produce 
+a correct solution, computational cost estimates and average size 
+of an individual throughout generations. 
+For the first two metrics we use the popular measurement  
+methods within GP field --- the \textit{performance curves}
+described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
+and $I(M,i,z)$ (the total number of individuals that must be processed 
+to yield a correct solution with probability $z =99\%$).
+
 \subsection{Simple symbolic regression}
-
 \textit{Simple Symbolic Regression} is a problem described
@@
 function $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
-
-Desired type of generated programs $\sigma$ and 
+Desired type of generated programs $\tau$ and 
 building blocks context $\Gamma$ are following.
@@
 \end{minipage}
-~\\
+
+\begin{figure}[h!]
+  \caption{Performance curves for Simple symbolic regression.}
+  \centering
+    \includegraphics[scale=0.135]{imgs/SSR_FINAL.png}
+\end{figure}
 Fitness function is computed as follows.
@@
 
+Standard \textit{ramped half-and-half strategy} scored 17/50 (34\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 23 with 192,000 individuals to be processed.
+The average individual size for generation 50 was 76.
+The experiment took 46 minutes.
+Our \textit{geometric strategy} scored 21/50 (42\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 29 with 150,000 individuals to be processed.
+The average individual size for generation 50 was 47.
+The experiment took 26 minutes.
+Figure~1 shows the progress of those metrics throughout generations.
+So for the \textit{Simple symbolic regression}
+our geometric strategy is slightly more successful then 
+ramped half-and-half strategy in all observed metrics.
+
 \subsection{Artificial ant}
-
 \textit{Artificial Ant} is another problem described
@@
 maximal number of steps.
+
+\begin{figure}[h!]
+  \caption{Performance curves for Artificial ant problem.}
+  \centering
+    \includegraphics[scale=0.135]{imgs/ANT_FINAL.png}
+\end{figure}
+
 Fitness value is equal to number of eaten food pieces.
@@
 considered as a correct solution.
-This limit is set to be 600\footnote{
+This limit is set to be 600 time steps\footnote{
 In \cite{koza92} this limit is said to be 400 time steps.
@@
 of food. Thus we use 600 time steps. 
-} time steps.
+}.
+
+Standard \textit{ramped half-and-half strategy} scored 7/50 (14\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 28 with 449,500 individuals to be processed.
+The average individual size for generation 50 was circa 400.
+The experiment took 265 minutes.
+Our \textit{geometric strategy} scored 19/50 (38\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 10 with 115,500 individuals to be processed.
+The average individual size for generation 50 was circa 90.
+The experiment took 107 minutes.
+This is a big improvement in all watched factors.
+Figure~2 shows it in greater detail.
 
 \subsection{Even parity problem}
-...
+
+Previous two problems were instances of standard GP, i.e. their building symbols
+obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
+The even-parity function is a boolean function taking as inputs $N$
+boolean values and returning \textit{True} if an even number of inputs 
+are \textit{True}. For odd number it returns \textit{False}.
+This problem has been used by many researchers
+as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
+where is presented approach to evolve recursive and modular programs
+by use of higher-order functions and $\lambda$-abstractions.
+We use very similar set of \textit{building symbols} as in \cite{yu01}. 
+\begin{align*}
+\tau = [Boo&l] \ar Bool\\
+\Gamma = \{
+  and   &: Bool \ar Bool \ar Bool                              ,
+  or    : Bool \ar Bool \ar Bool                              ,\\
+  nand  &: Bool \ar Bool \ar Bool                              ,
+  nor    : Bool \ar Bool \ar Bool                              ,\\
+  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
+  head' &: [Bool] \ar Bool                                   ,
+  tail' : [Bool] \ar [Bool]                              \}
+\end{align*}
+\begin{figure}[h!]
+  \caption{Performance curves for Even parity problem.}
+  \centering
+    \includegraphics[scale=0.135]{imgs/EP_FINAL.png}
+\end{figure}
+
+The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
+this problem is considered atomic.
+Unlike in \cite{yu01}, we use specific instance of polymorphic 
+function \texttt{foldr}. 
+And modifications of functions \texttt{head} 
+(returning the first element of the list) 
+and \texttt{tail} (returning the list without the first element) are used; 
+making them total by returning default value \textit{False}
+and \texttt{[]}, respectively.
+We use the same fitness function as in \cite{yu01}. 
+The fitness function examines the individual by giving
+it all possible boolean lists of length 2 and 3.
+
+Standard \textit{ramped half-and-half strategy} scored 9/50 (18\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 17 with 333,000 individuals to be processed.
+The average individual size for generation 50 was 39.
+The experiment took 28 minutes.
+Our \textit{geometric strategy} scored 32/50 (64\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 0 with 28,000 individuals to be processed.
+The average individual size for generation 50 was circa 100.
+The experiment took 33 minutes.
+Figure~3 shows it in greater detail.
+Once again this is a big improvement against standard method. 
+One may get the impression that geometric strategy suffers from bloat here,
+but considering that it performs significantly better and
+that term size around 100 is similar as for the previous two problems, 
+it seems to us that ramped half-and-half maybe has difficulties with 
+constructing more complex terms, and therefore is "staying small".
+Another remarkable observation is that 4 times out of 50 (8\% of all runs) 
+the 100\% correct solution was present in the generation 0. This is pretty good
+result for uninformed search. However, our results are less successful 
+then those presented in \cite{yu01} (with slightly different set of 
+building symbols); they scored 40/50 (80\%) success rate and $I(M,i,z)$ of 
+17,500 in generation 4. But by observation of typical solution which is often
+some modification of \texttt{foldr xor False inputList} one sees that important task
+for GP here is to create \texttt{xor} function. This is advantage for the more specialized term representation used in \cite{yu01} 
+(restricting use of outer variables). 
+The difference in crossover operators may also be involved in the difference, 
+but such analysis is beyond the scope of this paper. Finally, our result at least outperforms other results mentioned in \cite{yu01}: \textit{Generic genetic programming} 
+scored 17/60 (28\%); min $I(M,i,z) = 220,000$. \textit{GP with ADFs} 
+scored 10/29 (34\%); min $I(M,i,z) = 1,440,000$.
 
@@
 
-\begin{todo}
- & do future work bych napsal, že díky jednoduchosti 
-   toho co strategie dělá je to idealní kandidát na 
-   optimalizaci pomocí samotnýho GP
-\end{todo}
+In this work, we have proposed a generalization of genetic programming 
+for simply typed lambda calculus. The main focus of this paper is on individual generation algorithms. 
+The main idea is to explore the spectrum of available approaches bounded on one end by the traditional
+ramped half-and-half strategy, while the second bound represents an exhaustive systematic search. 
+Thus, three population initialization methods have been designed, 
+depending on different search strategies. 
+The first strategy corresponds to the above mentioned traditional genetic programming initialization,  
+the second one corresponds to exhaustive search, and the third one 
+represents a novel geometric strategy which outperforms standard genetic 
+programming in success rate, resource consumption and average individual size 
+on the presented experiments.
+
+%\red{ [větu dvě o experimentech]}
+
+The relevance of these algorithms does not effect only initialization, but it can be used during mutation
+operator in a straightforward way. 
+%Other performance enhancements based on theory of lambda calculus are 
+%proposed and supported by experiments, including abstraction 
+%elimination that enables us to utilize
+%a simple tree-swapping crossover.
+
+Our future research will focus on exploring other possible search strategies positioned in the middle ground 
+identified in this paper. Since the search strategy can be easily described by a rather simple filtration algorithm, it might be
+interesting to allow the typed GP system to evolve the search strategy it 
+utilizes in a kind of meta-evolution way.
+
+%\red{ [ještě něco navíc nebo podrobněji, říct poly že neni ale že se na něm pracuje a de hezky zandat do již systemu] }
 
@@
 \bibliographystyle{splncs}
-\bibliography{gp}
+\bibliography{evogp}
 

### 2013-11-11 tex/thesis.tex
[tex] article
[nic]
článek

@@
 Design goal behind our approach to crossover operation is
-to try to generalize standard tree swapping mutation.
+to try to generalize standard tree swapping crossover.
 This is not because of dislike of more specialized crossovers;
@@
 The difference between approach presented in this work and our approach 
-may be stated in the following way. Approach presented in \cite{yo01}
+may be stated in the following way. Approach presented in \cite{yu01}
 tries to adopt some useful constructs from domain

### 2013-11-11 v4/Problems/EvenParity/Problem.hs

@@ -47,8 +47,8 @@ rege = PO_CTTeP_ PO_CTTeP {
   cttep_ctx         = prog_ctx                                     ,
   
   cttep_gOpt        = CTTermG_Geom     prog_typ prog_ctx 0.75 , 
---cttp_gOpt        = CTTG_Koza2    prog_typ prog_ctx  , 
---cttp_gOpt        = CTTG_AllEdges prog_typ prog_ctx  , 
+--cttep_gOpt        = CTTermG_Koza     prog_typ prog_ctx  , 
+--cttep_gOpt        = CTTG_AllEdges prog_typ prog_ctx  , 
 
   cttep_ff          = FF6 prog_type ff "Problems.EvenParity.Funs", 
   

### 2013-11-11 v4/TTerm.hs

@@ -74,7 +74,16 @@ instance Show CTTerm where
          in "\\ "++ vars ++ " -> " ++ show tt 
 
 instance JShow CTTerm where
+  jss_size (CTTerm _ tt) = Just $ ttermSize tt
   jshow_js ctt = Just . jsStr . jsShow $ ctt
+
+ttermSize :: TTerm -> Int
+ttermSize tterm = case tterm of
+ TVar _   _ -> 1
+ TVal _   _ -> 1
+ TLam _ m _ -> 1 + ttermSize m
+ TApp m n _ -> 0 + ttermSize m + ttermSize n
+
 
 
 jsShow :: CTTerm -> String
